### 2022-07-07 GroupPolarizationStatmod/papers/ms.tex
3/10 analyzed; saving all data, means displayed correctly

% Author: Matthew Turner

% \documentclass[11pt,letterpaper]{scrartcl}
\documentclass[letterpaper,man,natbib]{apa6}
% \documentclass[11pt]{report}
% \documentclass{report}
% \documentclass{book}
\usepackage[bookmarks, hidelinks]{hyperref}
\usepackage{amssymb,amsmath}
% \usepackage{fullpage}
\usepackage{tabulary}
\usepackage{tabularx}
\usepackage{float}
% \usepackage[margin=1.00in]{geometry}
% \usepackage[margin=0.90in]{geometry}

\usepackage{caption}
\usepackage{booktabs}
\usepackage{pslatex}
\usepackage{apacite}
\usepackage{authblk}
\usepackage{subcaption}
\usepackage{pgfplots}
\usepackage{wrapfig}
\usepackage[english]{babel}
\usepackage{lmodern}
% \usepackage{setspace}
% \doublespace
% \usepackage{url}
\usepackage{bigfoot}
\usepackage[export]{adjustbox}
% \setlength\intextsep{0pt}

\usepackage{graphicx}

\title{How simple consensus to the mean can be mistaken for ``group polarization'' extremity-shifted consensus}
\shorttitle{Group polarization or simple consensus?}
\author[1]{{Matthew A.~Turner}}
\affil[1]{Department of Earth System Science, Stanford University}

\author[2,3]{{Paul E.~Smaldino}}
\affil[2]{Cognitive and Information Sciences, University of California, Merced}
\affil[3]{Center for Advanced Study in the Behavioral Sciences, Stanford University}

\abstract{
  Group polarization is the name given by social scientists for the observed tendency of
  like-minded groups to become more extreme as they discuss some topic of interest.
  Clearly it is important to understand such a phenomenon to understand broader
  problematic trends towards increased polariztion across society as a whole.
  This group-level shift in opinion extremity is often measured 
  Unfortunately, recent analyses have demonstrated that this measurement and
  analysis approach can result in false positive detections of changes in mean
  that are actually due to floor/ceiling effects from using ordinal scales to
  measuring continuously-distributed latent psychological variables, and not
  accounting for this mismatch. In this paper we analyze a selection of influential
  group polarization studies to demonstrate that many group polarization 
  detections reported throughout decades of literature are plausibly false
  due to these floor/ceiling effects. We develop a generative model of group
  polarization data, use it to represent each group polarization study, and
  investigate whether we can generate observed opinion shifts to extremity by
  setting the latent opinion mean to be constant, but allow latent opinion variance
  to decrease from pre- to post-discussion, which theoritcally 
  occurs in \emph{simple consensus} where the mean stays constant. After we identify which
  positive detections of group polarization are plausibly false in this way, we further
  investigate whether the latent constant mean opinion and pre- and post-discussion
  opinion variances can fool the statistical tests used in each study to 
  generate false positives. Finally, we demonstrate how a more sophisticated
  thresholded-normal statistical model can account for the metric-ordinal 
  mismatch to estimate latent distribution parameters and accurately identify
  simple consensus.
}

\begin{document}
\maketitle
\setlength{\parindent}{15pt}
% \maketitle

\section{Introduction}

\subsection{Group polarization}

\subsection{Statistical inference in opinion change}

2-3 par: the right way; what GP studies do; why it's a problem (N-1 sents) and how
to fix it (1 sent)

When

\begin{itemize}
  \item 
    Group polarization studies use null-hypothesis testing that implicitly
    assume (1) \emph{opinions} (or other equivalent latent psychological variables
    such as \emph{attitudes} or \emph{beliefs}) are continuously-valued, and
    (2) 
  \item
    The exact null-hypothesis testing procedure varies between studies $t$-tests, ANOVA, etc. 
\end{itemize}

Recently, problems with using have been made clear~\citep{Liddell2018}.
The way to fix this problem going forward is to use a thresholded-normal 
statistical model for group polarization data, or other related statistical 
model that accounts for opinion binning~\citep[Ch. 23]{Liddell2018,KruschkeDBDA}.

\subsection{Study selection and overview}

Studies were selected to be either influential papers, based on citations or
author prominence, that test either examine
group polarization in political opinions, and to be representative of investigations
of four primary group polarization explanations extant in the literature.
At the time of paper selection we did not examine the statistical methods in detail beyond
checking that they indeed used

Because of the difficulties casting individual research articles into our modeling
framework, and the central task of creating the model and analyzing this first
batch of studies with it, we limit our analysis to 10 studies. It is problem
enough that these 10 prominent studies have the problems we demonstrate they
plausibly do. Problems among these 10 suggest the problem may be more wide spread. 
When we do more analyses
they will be aided by improved data management systems for applying this model to 
additional studies (eventually outside of group polarization).

Statistical problems among these studies are not limited to a mismatch between
measurement and statistical procedures, which requires the detailed treatment
we provide in this paper. Other problems are more pedestrian.
One major issue shared by several studies
is the use of unpaired $t$-tests to detect group polarization. Since the same
group has their opinions measured twice over time, the opinions are correlated
and so a paired $t$-test would be required for test assumptions to be met 
(assuming all other paired $t$-test assumptions were met, 
which of course they are not, which is why our paper here is necessary).
In some cases the reported shifts were ``significant'', 
but not truly ``group polarization'' shifts, i.e., 
shifts from initially biased group opinions
to more extreme group opinions biased in the same direction; instead shifts 
were from a neutral initial group opinion to a biased group opinion 
(as in, e.g.,~\citet{Abrams1990} Uncategorized2), or from
an opinion biased in one direction (as in, e.g.,~\citet{Abrams1990} Categorized1,
Categorized2, AND OTHERS???). In some results of \citet{Freidkin1999a}, 
shift values are reported with their
standard deviations---but the shift standard deviations are greater than the
shifts themselves, indiciating that zero shift is included in the observed
shift confidence interval. This is \emph{prima facie} a negative result that
should not have been counted as evidence of group polarization.

Some group polarization studies do not measure continuous latent opinions using
a metric scale, and so are not affected by the floor and ceiling effects that
undermine a good number of experimental results, as we examine here. There may,
however, be other problems, such as non-group polarization shifts and the use
of unpaired $t$-tests when paired tests are required.

\subsection{Research overview}

In this paper we analyze whether published statistical detections of group 
polarization obtained through null-hypothesis statistical tests
are plausibly false because they can be explained equally well by simple consensus
where the mean group opinion does not change over time, but opinion variance does.
This analysis requires several preliminary steps. 
First we formally develop our generative model of simple consensus and group 
polarization in terms of changes in group opinion mean and variance from 
pre- to post-deliberation. Then we developed a search algorithm for identifying 
which, if any, simple consensus scenarios with 
constant latent mean and pre- and post-deliberation opinion variances 
could generate published observations of group polarization. To apply this model
and algorithm to published results, we developed an interactive web application
for data input and to test which different potential simple consensus scenarios 
could generate group polarization opinion shifts. After identifying plausibly false
detections of group polarization we examine whether the identified simple
consensus scenario (parameters?) can indeed generate data to fool the 
paper's null-hypothesis tests into thinking the data actually support the inference
that simple consensus did not occur. Finally, we examine whether, in these same
plausibly false detections, a Bayesian thresholded-normal statistical model
will accurately identify simple consensus instead of falsely detecting group
polarization.

\section{Model}

Our model represnts the generative process of finding different forms of group consensus
in group polarization experiments, and the process of measuring that consensus using
an ordinal measurement scale, e.g.\ a Likert-type scale. This model considers two
relevant forms of consensus: \emph{simple consensus} and \emph{group polarization}.
Either process is \emph{consensus} since opinion variance among the group 
decreases as individuals come to agree more than they did previously. 
Group polarization experiments are designed to foster group consensus
it is not necessary to consider clustering, negative influence, or other
processes known to generally occur when opinions within a group significantly
diverge~\citep{Turner2018}.

\section{Analysis}

\bibliographystyle{apacite}

\setlength{\bibleftmargin}{.125in}
\setlength{\bibindent}{-\bibleftmargin}

\bibliography{/Users/mt/workspace/Writing/library.bib}

\appendix

\section{Notes on studies}

\paragraph{\citet{Abrams1990}}

\paragraph{\citet{Burnstein1973}}

\paragraph{\citet{Hogg1990}} It is not clear if there are indeed significant pre-post
deliberation differences reported in this paper. The Results are a tangle of
estimates of effect sizes and diagnostic statistical measures. It is also
not clear that they actually report the pre- and post-deliberation opinions;
it appears they only report the shifts. ``Opinion Difference Measures'' 
does not seem to include a test that the shifts themselves are reliable.

\end{document}

### 2022-07-07 GroupPolarizationStatmod/app.R

@@ -9,7 +9,7 @@ source("numerical.R")
 
 # STUDIES_DB = "data/CaseStudies_ProtoDB-TEST.xlsx"
 # STUDIES <- read_excel(STUDIES_DB, "ArticleStudiesMinimal")
-STUDIES_DB = "data/StudiesAnalysis-TEST.csv"
+STUDIES_DB = "data/StudiesAnalysis.csv"
 STUDIES <- read.csv(STUDIES_DB)
 # STUDIES <- read_excel("data/CaseStudies_ProtoDB.xlsx", "ArticleStudiesMinimal")
 FULL_TAGS <- sort(paste(STUDIES$ArticleTag, STUDIES$TreatmentTag, sep=" - "))
@@ -116,7 +116,7 @@ server <- function(input, output, session) {
 
     # STUDIES <- read_excel("data/CaseStudies_ProtoDB-TEST.xlsx", "ArticleStudiesMinimal")
     # FULL_TAGS <- sort(paste(STUDIES$ArticleTag, STUDIES$TreatmentTag, sep=" - "))
-    STUDIES_DB = "data/StudiesAnalysis-TEST.csv"
+    STUDIES_DB = "data/StudiesAnalysis.csv"
     STUDIES <- read.csv(STUDIES_DB)
 
     STUDIES$LatentMean <- as.numeric(STUDIES$LatentMean)
@@ -184,8 +184,10 @@ server <- function(input, output, session) {
             # ", latPostSD=", fmtVal(latentPostSDResult()[1]^0.5), "\n",
             "simObsPreSD=", fmtVal(sdObsPre()), 
             ", simObsPostSD=", fmtVal(sdObsPost()), 
-            ".\nAbs. Error: pre = ", fmtVal(sqrt(latentPreSDResult()[2])),
-            ", post = ", fmtVal(sqrt(latentPostSDResult()[2])),
+            ".\nMeans: mu_pre = ", fmtVal(meanObs(kVec(), probVecPre())),
+            ", mu_post = ", fmtVal(meanObs(kVec(), probVecPost())),
+            ".\nError in means: mu_pre = ", fmtVal(latentPreSDResult()[2]),
+            ", mu_post = ", fmtVal(latentPostSDResult()[2]),
             sep = ""
         )
 
@@ -264,15 +266,17 @@ server <- function(input, output, session) {
         treatmentRow$LatentSDPre <- latentPreSDResult()[1]
         treatmentRow$LatentSDPost <- latentPostSDResult()[1]
 
[…]

### 2022-07-07 GroupPolarizationStatmod/numerical.R

@@ -16,17 +16,18 @@ solveForLatentSD <- function(kVec, latentMean, observedMean, guess,
 {
     # Keep the squared error function to be minimized through hillclimbing
     # separate from hillclimbing itself.
-    sqErr <- function(latentSD)
+    errfunc <- function(latentSD)
     {
         simulatedObservedMean <- meanObs(
             kVec,
             makeProbVec(kVec, latentMean, latentSD)
         )
-
-        return ((observedMean - simulatedObservedMean)^2);
+        
+        # return ((observedMean - simulatedObservedMean)^2);
+        return (simulatedObservedMean - observedMean);
     }
 
-    return (hillclimbing(sqErr, guess, step, maxIts, tol))
+    return (hillclimbing(errfunc, guess, step, maxIts, tol))
 }
 
 
@@ -36,10 +37,13 @@ solveForLatentSD <- function(kVec, latentMean, observedMean, guess,
 hillclimbing <- function(errfunc, sd_guess, stepSize = 0.1, maxIts = 1e5, tol = 1e-4)
 {
     sd_curr <- sd_guess
+    # sqerr_curr <- errfunc(sd_curr)[2]
+    # err_curr <-  errfunc(sd_curr)[2]
+    # sqerr_curr <- err_curr^2
     err_curr <- errfunc(sd_curr)
+    sqerr_curr <- err_curr^2
     
     its <- 0
-    # totalIts <- 0
     change <- 0.0
 
     while(its < maxIts)
@@ -50,20 +54,22 @@ hillclimbing <- function(errfunc, sd_guess, stepSize = 0.1, maxIts = 1e5, tol = 
         # generate pre or post observed mean.
         sd_next <- rnorm(1, sd_curr, stepSize)
         err_next <- errfunc(sd_next)
+        sqerr_next <- err_next^2
 
-        change <- err_next - err_curr
+        change <- sqerr_next - sqerr_curr
 
         if (abs(change) < tol)
         {
             return (c(sd_curr, err_curr, its));
         }
         
-        # A negative change means err_next is less that err_curr, and closer/better
+        # A negative change means err_next is less that sqerr_curr, and closer/better
         # to minimum solution.
         if (change < 0)
         {
             sd_curr <- sd_next
             err_curr <- err_next
+            sqerr_curr <- sqerr_next
         }
     }
     return (c(sd_curr, err_curr, its));

### 2022-07-07 GroupPolarizationStatmod/papers/ms.tex
updated Hogg1990 supplement

@@
 
+Different group polarization studies have different goals, and measure shifts
+in a variety of ways, using different (or no) statistical tests to confirm
+the presence of group polarization opinion shifts. This supplemental section
+reviews the ten papers one-by-one and explains how group 
+polarization was measured in each paper. I also explain how I decided which
+group polarization detections I included from each paper, and how we decided
+if they were plausibly false positives, including whether we needed to
+
 \paragraph{\citet{Abrams1990}}
@@
 
-\paragraph{\citet{Hogg1990}} It is not clear if there are indeed significant pre-post
-deliberation differences reported in this paper. The Results are a tangle of
-estimates of effect sizes and diagnostic statistical measures. It is also
-not clear that they actually report the pre- and post-deliberation opinions;
-it appears they only report the shifts. ``Opinion Difference Measures'' 
-does not seem to include a test that the shifts themselves are reliable.
+\paragraph{\citet{Hogg1990}} It is not clear if there are indeed significant
+pre-post deliberation differences reported in this paper. The Results are a
+tangle of estimates of effect sizes and diagnostic statistical measures. It
+is also not clear that they actually report the pre- and post-deliberation
+opinions; it appears they only report the shifts. ``Opinion Difference
+Measures'' does not seem to include a test that the shifts themselves are
+reliable. I decided to count a finding from this study as a false positive if
+the model generated pre- and post-deliberation mean opinions within an
+absolute error of 0.01 of the reported value. The argument seems to be that
+since they are testing social ``frame[s] of reference'' (Table 1) that it is
+not necessary to show that group polarization itself occurs, only that the
+various experimental treatments yield different shifts---but if the shifts
+themselves are insignificant, then what is being measured, exactly?
 

### 2022-07-07 GroupPolarizationStatmod/app.R
updates to database

@@ -252,6 +252,7 @@ server <- function(input, output, session) {
     
     observeEvent(input$saveBtn, {
         treatmentSplit <- strsplit(input$treatmentTag, " - ")
+        print(treatmentSplit)
         article <- treatmentSplit[[1]][1]
         treatment <- treatmentSplit[[1]][2]
         

### 2022-11-04 GroupPolarizationStatmod/papers/ms.tex
paper and data updates

@@
 
-\title{How simple consensus to the mean can be mistaken for ``group polarization'' extremity-shifted consensus}
+\title{Group polarization is often a mirage caused by inappropriate statistical analyses}
 \shorttitle{Group polarization or simple consensus?}
-\author[1]{{Matthew A.~Turner}}
+\author[1,2]{{Matthew A.~Turner}}
 \affil[1]{Department of Earth System Science, Stanford University}
+\affil[2]{Division of Social Sciences, Stanford Doerr School of Sustainability, Stanford University}
 
-\author[2,3]{{Paul E.~Smaldino}}
-\affil[2]{Cognitive and Information Sciences, University of California, Merced}
-\affil[3]{Center for Advanced Study in the Behavioral Sciences, Stanford University}
+\author[3,4,5]{{Paul E.~Smaldino}}
+\affil[3]{Cognitive and Information Sciences, University of California, Merced}
+\affil[4]{Center for Advanced Study in the Behavioral Sciences, Stanford University}
+\affil[5]{Santa Fe Institute}
 
 \abstract{
-  Group polarization is the name given by social scientists for the observed tendency of
-  like-minded groups to become more extreme as they discuss some topic of interest.
+  Group polarization is the observed tendency of
+  like-minded groups to arrive at a more extreme consensus opinion than
+  the average opinion after discussing some topic of interest.
   Clearly it is important to understand such a phenomenon to understand broader
   problematic trends towards increased polariztion across society as a whole.
-  This group-level shift in opinion extremity is often measured 
   Unfortunately, recent analyses have demonstrated that this measurement and
-  analysis approach can result in false positive detections of changes in mean
-  that are actually due to floor/ceiling effects from using ordinal scales to
-  measuring continuously-distributed latent psychological variables, and not
-  accounting for this mismatch. In this paper we analyze a selection of influential
-  group polarization studies to demonstrate that many group polarization 
-  detections reported throughout decades of literature are plausibly false
-  due to these floor/ceiling effects. We develop a generative model of group
-  polarization data, use it to represent each group polarization study, and
-  investigate whether we can generate observed opinion shifts to extremity by
-  setting the latent opinion mean to be constant, but allow latent opinion variance
-  to decrease from pre- to post-discussion, which theoritcally 
-  occurs in \emph{simple consensus} where the mean stays constant. After we identify which
-  positive detections of group polarization are plausibly false in this way, we further
+  analysis approach can result in false positive detections group polarization
+  from failing to account for floor/ceiling effects induced when using ordinal scales to
+  measure continuously-distributed latent psychological variables. 
+  Here we demonstrate that many high profile detections
+  of group polarization are plausibly false
+  due to these floor/ceiling effects using generative modeling seeded by 
+  reported group polarization values and parameters. After we identify which
+  reported detections of group polarization are plausibly false detections, we further
   investigate whether the latent constant mean opinion and pre- and post-discussion
   opinion variances can fool the statistical tests used in each study to 
-  generate false positives. Finally, we demonstrate how a more sophisticated
-  thresholded-normal statistical model can account for the metric-ordinal 
-  mismatch to estimate latent distribution parameters and accurately identify
-  simple consensus.
+  generate false positives. Finally, we evaluate whether a more appropriate
+  Bayesian cumulative normal statistical model can more accurately distinguish
+  group polarization from simple consensus. Group polarization continues to 
+  pose a problem for psychological theory with an ever-expanding set of potential theoretical
+  explanations. A search for explanations is pointless, however, without sound
+  evidence. We believe this work will provide a sturdier analytical foundation for 
+  future group-polarization research.
 }
@@
 
-\section{Model}
+\section{Model and Data}
 
@@
 
+We collected group polarization data from ten empirical studies with sixty constituent
+experimental treatments across all studies, specifically the opinion shift and pre- and
+post-deliberation mean opinions, and the number of participants
+in each treatment (\citet{Burnstein1975} did not report pre- and post-deliberation
+means, only the shift in means). We did not include experimental treatments that did not 
+yield a positive detection of group polarization since we are only estimating
+the rate of plausible false positive detections of group polarization. 
+We then created new data that indicated whether a reported.
+
+In two cases, reported opinion and shift values were converted from an ordinal
+scale to a percentage scale, specifically \citet{Friedkin1999a} and \citet{Krizan2007}.
+The authors in these papers did report the original ordinal measurement scale,
+so we converted the means back to the orignial measurement scale.
+In a few cases from \citet{Friedkin1999a} we did not need to apply our 
+model to the reported data because we could determine a reported opinion shift was 
+\emph{prima facie} false since reported opinion shift magnitudes
+were smaller than the standard deviation in shifts (see Table 1, p. 868, \emph{ibid}),
+despite Friedkin's claim that regressions (not clearly defined) on all shifts 
+yielded $p < 0.05$~\cite{Kruschke2018c}.
+
+\subsection{Latent opinions, ordinal measurements}
+
+Our analysis relies on two important formal concepts: latent opinions and their 
+measurement on an ordinal scale. \emph{Latent opinions} are unobservable psychological,
+continuous-valued variables that are somehow represented and stored in the
+brain-body. 
+When researchers use continuous statistical models to infer 
+that group polarization occurred (which all ten considered here do), 
+they are implicitly assuming that latent opinions are themselves continuous.
+We can only measure behaviors in group polarization experiments,
+specifically the behavior of participants marking one ordinal value or another
+to indicate their opinion---a process that is not well understood in itself,
+but which is unimportant for our analyses. \emph{Ordinal measurements} (e.g.\
+a Likert or other discrete-valued opinion scale) of latent opinions force
+participants to appropriately bin their opinion in one of several categories
+representing the valence and extremity of their opinion, e.g., on a
+7-point Likert scale -3 could indcate
+``strongly disagree'', +3 could be ``strongly agree'', 0 would be neutral, and 
+intermediate values are less extreme opinions of either valence.
+
+All ten studies analyzed here used an ordinal scale, but group polarization experiments seem to
+lack any systematicity in experimental design beyond that. (EXAMPLES)
+
+\subsection{Simple consensus and group polarization are statistical patterns}
+
+\subsection{Analytical approach}
+
+The goal of this study is to understand if published detections of group polarization are
+plausibly false.
+
+Therefore we designed a series of tests to understand if, and to what degree,
+published results are false positive detections of group polarization that
+are at least as well described by a simple-consensus model.
+
+\subsubsection{Identifying parameters that plausibly generate false detections of group polarization}
+
+\subsubsection{$t$-tests of generated simple-consensus data}
+
+\subsubsection{Bayesian ordered-probit models of generated simple-consensus data}
+
 \section{Analysis}
+
+\subsection{Plausible parameters for false positives}
+
+We found that X\% of group polarization detections across all studies were 
+plausibly false positives (Table~\ref{tab:XXXXX}).
+
+\subsection{Do $t$-tests accurately detect simple consensus?}
+
+We found that in X\% of generated simple consensus data led to false positive
+detections
+
+When simulated data reliably result in false positive $t$-test results (where
+the $t$-test $p \le 0.1$) this strengthens the case that
+
+\subsection{Do Bayesian ordered-probit models accurately detect simple consensus?}
+
+We found that in many cases Bayesian ordered-probit models avoid falsely identifying simple
+consensus as group polarization. This supports the claim that it is the choice of statistical
+procedure, including the failure to account for the ordinal nature of, and variance within,
+opinion measuremenets in group polarization experiments that leads to false
+positives, not anything intrinsic to the phenomenon of group polarization.
+
+\section{Discussion}
+
+We showed that many prominent detections of group polarization are plausibly
+false, and that future studies \emph{must} use ordered-probit statistical
+models and estimate highest-density or confidence intervals, not point-estimates 
+and significance values~\cite{Meehl1997}.
+
+Due to a lack of data sharing in general, and often from decades-old papers, and a persistent 
+failure of group polarization work to provide opinion variance information, 
+it is impossible to know whether an historical detection of group polarization
+is a true or false positive.
+
+This problem is made all the worse by the lack of a consistent experimental design for group
+polarization experiments. This makes it impossible to systematically account for
+additional sources of vairance known to lead to (orthogonal) over-estimates
+of effect sizes~\cite{Yarkoni2021}. This is due to the rush to explain why 
+group polarization occurs using weak theory~\cite{Cartwright1971,Cartwright1973} and makes
+group polarization meta-analyses (e.g. Brown, 1986; and Isenberg, 1986)\nocite{Brown1986,Isenberg1986}
+unreliable~\cite{Meehl1990,Meehl1997}.
+
+Group polariztion is an essential social phenomenon that could be a major driver of
+progress-halting political polarization more broadly. If echo chambers of like-minded
+individuals tend to become radicalized it is essential to measure by how much
+in different contexts, and to develop rigorous theoretical explanations of
+this phenomenon. It appears, unfortunately, that decades of group polarization
+research may have published many plausibly false inferences of the
+occurrence of group polarization, as demonstrated by our simple consensus
+model of false group polarization detection. As mentioned, a lack of
+consistency and failure to account for additional sources of variance could
+be two additional ways published research is underpowered, and associated
+theoretical insights undermined. It may not be easy to get the group polarization field to 
+consolidate experimental designs and use appropriate statistical procedures,
+but it is a straightforward solution. Indeed, combining inter-study consistency and statistical
+rigor is the only way to a reliable understanding of group polarization.
 

### 2022-11-16 GroupPolarizationStatmod/papers/ms.tex
t-test table being made on small examples

@@
 When simulated data reliably result in false positive $t$-test results (where
-the $t$-test $p \le 0.1$) this strengthens the case that
+the $t$-test $p \le 0.1$) this strengthens the case that.
+
+One complicating factor to this approach is that statistical power, $W$, varies
+between treatments. Therefore, simulated data should yield statistically-significant
+group polarization with frequency close to $W$ for a given treatment at a given
+significance value, $\alpha$.
 
@@
 
-\paragraph{\citet{Abrams1990}}
+\paragraph{\citet{Abrams1990}} This study is meant to support the presence of the "self-categorization" effect in
+social psychology. The paper presents three experiments and their results--Experiment 3 is the group polarization
+experiment, designed to show that self-categorization is at work in the emergence of group polarization. Uses high
+schoolers aged 16-17 in Hawkes Bay, New Zealand. There are five deliberation topics (p. 112) across two conditions
+(p. 113): one condition where participants are made aware of the relationship between their opinions and other group
+members' opinions ("Categorized") and another condition where participants were given no information about their
+opinion relative to other group members' opinions ("Uncategorized"). It was confusing to me what scale is used for
+Table 2 (p. 114), but parsing the "Dependent measures" subsection several times led me to understand the last
+paragraph, that "[a]ttitude shift overall was...assessed...on the unconverted +4 to -4 scale."
 
-\paragraph{\citet{Burnstein1973}}
+\paragraph{\citet{Burnstein1973}} First I want to note this study was funded in part by a Guggenheim fellowship (see
+note on first page) and the authors acknolwedge the assistance of "vinous" discussions with colleagues at the
+University of Provence and the "spirited and intelligent efforts" of two female research assistants. The study
+presents the results of two experiments (really one experiment with two pools of participants) that test the
+persuasive arguments hypothesis by having students perform the standard group polarization experiment with five CDQ
+discussion items, items A,B,E,F, and H (A,B risk-inducing; E, F caution-inducing; H neutral). Experiment 1 used 149
+male intro psych students and Experiment 2 used 76 male intro psych students from the University of Michigan. There
+are three conditions across the two experiments--one hypothesized to show an emergent shift and the other two
+hypothesized to not show a shift. None of the conditions allowed participants to know each others' ratings exactly,
+the participants only deliberated the various topics. In the one hypothesized to show a shift, participants were
+instructed to argue for their position. In the other two, participants argued against their position. The "against"
+condition was further divided into a case where everyone knew everyone was prompted to argue against their position,
+and another sub-condition where participants did not know if others were arguing for or against their true position.
+We include only the first condition where participants argued for their position, which is the only condition
+expected to show shifts (though others do, but the authors let themselves off the hook saying, "They are curious
+findings and defy straightforward explanation...we will merely note their occurrence and not venture to speculate as
+to their cause"). Thus, in our table we have conditions A, B, E, F, and H. Even though H is supposed to not show a
+shift, it does in this study, which is possibly further evidence that this study merely shows participant consensus
+formation, not opinion shifts.
+
+\paragraph{\citet{Burnstein1975}} In this study, Burnstein and Vinokur ran an experiment to evaluate their theory
+that generating persuasive arguments are necessary for group polarization to occur. The authors design a study with
+three conditions, one Experimental and two Control, where participants (60 male students from intro psych at the
+University of Michigan) first answered four CDQ items that typically generate shifts towards more risk, then, in the
+Experimental condition, saw group members' responses and produced arguments that might support the positions of
+other participants. In the first control condition, Condition II in the paper and Control1 in our table,
+participants were exposed to others' opinions, but does not get to reflect on their opinions or give arguments. In
+condition III, the second control condtition (we call Control2) participants do not get to know other participants'
+opinions at all. Thus, Burnstein and Vinokur hypothesize that significant shifts will only occur in Condition I, the
+Experimental condition. We thus only include Condition I in our analysis. The authors only report the shift, meaning
+we are free to choose any pre- and post-deliberation value to see if the null hypothesis is actually plausible.
+
+\paragraph{\citet{Schkade2010}} Friedkin uses three CDQ topics, "Sports", "School", and "Surgery" that use a
+20-point CDQ risk scale in fractional units, 0.05, 0.10, ..., 1.00 across three group sizes (2, 3, and 4, which he
+calls "dyads", "triads", and "tetrads", which we use in our CaseStudyTag). Friedkin only gives opinion shifts, and
+apparently transforms the fractional CDQ responses into percentages, since the shifts range between 5 and 9.
+However, closer inspection of Friedkin's results shows that the reported "Choice Shift" is not equal to the
+difference between mean final and initial opinions. Friedkin also uses "different" means for pre- and
+post-discussion opinions, which seems easy enough to show are really the same by re-organizing terms in the sum.
+Friedkin uses two topics in the tetrad condition where responses are in units of money where participants can freely
+give any number they choose ("Asbestos" and "Disaster"), and another topic called "Institutions" where participants
+rate 13 institutions, which participants rated on a scale from 0 to 100 percent to indicate the degree of trust the
+participant has in each institution. We only consider the CDQ topics since these are the ones rated on a discrete
+scale.
 
@@
 
+\paragraph{\citet{Krizan2007}} Uses five CDQ topics ("scenarios") from Kogan and
+Wallach (1967) that typically provide shifts toward more risky decisions.  These are
+given in Footnote 2 on p. 194. Krizan and Baron report shifts and mean pre-test
+opinions on the risk scale, meaning that shifts are expected to be negative towards
+lower odds of success required to take the risk, i.e., the acceptance of more risk.
+Krizan and Baron report mean pre-test opinions, but average over all three conditions
+for each scenario, then report shifts for each scenario and condition, so we
+effectively have no way for an exact match of initial mean to condition x scenario,
+where conditions are no-discussion, deliberation without knowledge of an out-group's
+opinion, and deliberation with knowledge of an out-group's opinion. We are only
+interested in the non-control cases--K\&B found no difference in knowing out-group
+opinion, which they say is evidence against the self-categorization hypothesis of group
+polarization.  Therefore, we have ten cases we test across five scenarios and two
+conditions with our model, and we use the reported initial mean for each scenario
+twice--once for each scenario across the two conditions. Post-deliberation opinions are
+calculated in the spreadsheet using the reported shifts. K\&B also report the "Overall"
+shift, but I find this meaningless since it is across different items with no
+statistical control for inter-item variability. I noted that false detections are
+plausible in all cases due to the fact that the standard deviations of shifts overwhelm
+the shifts themselves, but I will still run the model on these cases to confirm.
+
+\paragraph{\citet{Moscovici1969}} Moscovici and Zavalloni studied whether group
+polarization is more general than CDQ by using opinions about political topics. There
+were two general topics--attitudes about "Americans" and attitudes about then-president
+of France Charles "DeGaulle"--with 12 and 11 specific items, respectively, on which
+participants gave opinions and deliberated. Participants deliberated on one of the two
+general topics, giving opinions on each item on a seven-point Likert scale, -3 to +3.
+The authors reported pre-consensus and post-consensus mean opinions taken over all
+participants and items for each of the two general topics I have called the "Americans"
+and "DeGaulle" conditions.
+
+\paragraph{\citet{Myers1970}} Myers and Bishop test whether group polarization occurs
+in groups with different levels of racial bias determined by a survey that is not also
+the deliberation topic. The participants were high school students from western
+Michigan. In the experimental condition, social influence occurs through deliberation;
+in the control condition there is no deliberation or social influence at all, just
+re-tested opinions after a distraction task. Obviously we only report on whether there
+was a significant shift in the experimental condition. In both conditions, three groups
+were created according to their racial prejudice as measured by a 100-item racial
+attitude measurement instrument, called "low", "medium", and "high" prejudice groups.
+The participants then gave their opinions, deliberated, and gave opinions again on
+eight deliberation topics. Means were taken across items and prejudice group.
+
+\paragraph{\citet{Myers1975}} Myers performs two experiments each with experimental and control conditions for two group types with initial biases.
+In Experiment 1, Myers creates biased groups by having participants (intro psychology students) read a description of six different teachers, three "bad" professors and three "good" professors, which the participants are assumed to infer through vignettes about the professors.
+The participants in the experimental condition of Experiment 1 deliberated in small groups after giving their initial opinions on six items a 0-to-10 bad-to-good scale.
+In Experiment 2, participants were put into conservatively or liberally biased small groups based on their pre-deliberation opinions on six deliberation items regarding womens' rights and the role of women in the family and society.
+Again, in the experimental group, participants deliberated each item before giving their final opinions on the items.
+In the control group in each Experiment, there was no deliberation, just some filler task.
+Curiously, significant shifts were still found in the control groups--it is unclear to us why.
+Experiment 1 also had participants adjust the bad and good professors' pay to obtain non-ordinal measures of opinion regarding these professors, which we do not consider here.
+However, as we will discuss in the paper, this does not make it free of problems--like so many other group polarization studies, Myers statistical model does not account for inter-item, inter-group, or inter-participant variability.
+
+\paragraph{\citet{Schkade2010}} Participants in Schkade, Sunstein, and Hastie's study were put into two politically ideologically biased groups based on their geographical location in the US state of Colorado: participants were drawn from Boulder, Colorado (abbreviated CO), to be liberally biased and from Colorado Springs, CO, to be conservatively biased.
+Participants gave their opinions and discussed three topics related to Global Warming, Civil Unions, and Affirmative Action.
+Pre- and post-deliberation mean are given in Table 1 (p.
+232) for all six (topics x geographies).
+
 \end{document}

### 2022-11-16 GroupPolarizationStatmod/analysis.R

@@ -1,0 +1,106 @@
+library(dplyr)
+
+##
+# Arguments:
+#  studiesAnalysisDfPath (string): path where exact analysis was done using web app
+#    to determine whether a result was a false positive or not.
+#  
+makePlausibleFPTable <- function(studiesAnalysisDfPath = "data/StudiesAnalysis.csv", useSynced = FALSE) 
+{
+  plausibleFPTable <-
+    read.csv(studiesAnalysisDfPath) %>%
+    filter(IncludeInt == 1) %>% 
+    group_by(ArticleTag) %>%
+    summarize(PlausibleFPRate = mean(PlausibleInt))
+  
+  write.csv(plausibleFPTable, file = "data/output/PlausibleFPTable.csv", row.names = FALSE)
+  
+  return (plausibleFPTable)
+}
+
+
+##
+# Generate synthetic data based using parameters from experimental conditions
+# that yielded plausibly false positives to test how frequently simulated simple
+# consensus data yields false positives.
+#
+# Arguments:
+#  studiesAnalysisDfPath (string): path where exact analysis was done using web app
+#    to determine whether a result was a false positive or not.
+#  ntrials (int): Number of synthetic datasets to generate and t-test for each of the
+#    treatments yielding plausibly false positive results.
+#  limit (int): Limit the number of treatments to test for initial development.
+#
+makeTtestReliabilityTable <- function(studiesAnalysisDfPath = "data/StudiesAnalysis.csv",
+                                      outputPath = "data/output/TtestFitTable.csv",
+                                      diagnosticSavePath = "data/diagnostic/tTestFits.RDS",
+                                      ntrials = 2, limit = 2)
+{
+  # Load analysis of studies from web interface stored as CSV.
+  studiesDf <- read.csv(studiesAnalysisDfPath)
+
+  # Filter out implausible false positives.
+  studiesDf <- filter(studiesDf, (IncludeInt == 1) && (PlausibleInt == 1))
+  
+  # Limit number of plausible false positives for preliminary development.
+  if (limit > 0)
+  {
+    studiesDf <- head(studiesDf, limit)
+  }
+  
+  ## Apply t-tests to generated data over n trials.
+  # Initialize results dataframe with same length of studiesDf and appropriate
+  # additional columns.
+  resultsDf <- data.frame(ArticleTag = c(), 
+                          TreatmentTag = c(),
+                          TrialIndex = c(),
[…]

### 2022-11-16 GroupPolarizationStatmod/experiments.R

@@ -182,20 +182,15 @@ calculateBayesian <- function(N, firstBinValue, nBins, latentMean, latentSd)
 
 
 tTestExperiment <- function(N, firstBinValue, nBins, latentMean, latentSdPre, 
-                            latentSdPost, paired = TRUE, var.equal = FALSE)
+                            latentSdPost, paired = FALSE, var.equal = FALSE)
 {
-    simObsPre <- simulatedObservation(N, firstBinValue, nBins, latentMean, latentSdPre)
-    # print(simObsPre)
-    print(mean(simObsPre))
-    # print("")
-    simObsPost <- simulatedObservation(N, firstBinValue, nBins, latentMean, latentSdPost)
-    # print(simObsPost)
-    print(mean(simObsPost))
+  print(latentMean)
+  print(latentSdPre)
+  simObsPre <- simulatedObservation(N, firstBinValue, nBins, latentMean, latentSdPre)
 
-    # df <- data.frame("pre" = simObsPre, "post" = simObsPost)
+  simObsPost <- simulatedObservation(N, firstBinValue, nBins, latentMean, latentSdPost)
 
-    return (t.test(simObsPre, simObsPost, paired = paired, var.equal = var.equal))
-    # return (t.test(pre, post, data=df, paired = paired, var.equal = var.equal))
+  return (t.test(simObsPre, simObsPost, paired = paired, var.equal = var.equal))
 }
 
 
@@ -212,37 +207,40 @@ tTestExperiment <- function(N, firstBinValue, nBins, latentMean, latentSdPre,
 
 simulatedObservation <- function(N, firstBinValue, nBins, latentMean, latentSd)
 {
-    # Draw latent opinion data for each participant; parameters are set
-    # to empirical, as opposed to population, values.
-    
-    # mvrnorm uses variance, not standard deviation, for its spread parameter.
-    # latentVariance <- latentSd^2
-    # latentData <- mvrnorm(N, latentMean, latentVariance, empirical=TRUE)
-    latentData <- rnorm(N, latentMean, latentSd)
+  # Draw latent opinion data for each participant; parameters are set
+  # to empirical, as opposed to population, values.
+  
+  # mvrnorm uses variance, not standard deviation, for its spread parameter.
+  # latentVariance <- latentSd^2
+  # latentData <- mvrnorm(N, latentMean, latentVariance, empirical=TRUE)
+  print(N)
+  print(latentMean)
+  print(latentSd)
[…]

### 2022-11-18 GroupPolarizationStatmod/papers/ms.tex
t-test analysis done for first pass

@@
-% Author: Matthew Turner
+% % Author: Matthew Turner
 
-% \documentclass[11pt,letterpaper]{scrartcl}
-\documentclass[letterpaper,man,natbib]{apa6}
-% \documentclass[11pt]{report}
-% \documentclass{report}
-% \documentclass{book}
-\usepackage[bookmarks, hidelinks]{hyperref}
-\usepackage{amssymb,amsmath}
-% \usepackage{fullpage}
-\usepackage{tabulary}
-\usepackage{tabularx}
-\usepackage{float}
-% \usepackage[margin=1.00in]{geometry}
+% % \documentclass[11pt,letterpaper]{scrartcl}
+% % \documentclass[letterpaper,man,natbib]{apa6}
+% \documentclass[11pt,letterpaper]{amsart}
+% % \documentclass[11pt, letterpaper]{article}
+% % \usepackage{natbib}
+% % \documentclass[11pt]{report}
+% % \documentclass{report}
+% % \documentclass{book}
+% \usepackage[bookmarks, hidelinks]{hyperref}
+% \usepackage{amssymb,amsmath}
+% % \usepackage{fullpage}
+% \usepackage{tabulary}
+% \usepackage{tabularx}
+% \usepackage{float}
+% % \usepackage[margin=1.00in]{geometry}
 % \usepackage[margin=0.90in]{geometry}
 
-\usepackage{caption}
+% % \usepackage{fontspec}
+
+% % \usepackage{xunicode}
+% % \usepackage{xltxtra}
+% % \defaultfontfeatures{Mapping=tex-text}
+% % \setmainfont[Ligatures={Common}, Numbers={OldStyle}, Variant=01]{Hoefler Text}
+% % \setromanfont[Ligatures={Common}, Numbers={OldStyle}, Variant=01]{Hoefler Text}
+% % \setmonofont[Scale=0.95]{PT Sans}
+% \usepackage{caption}
+% \usepackage{apacite}
+% \usepackage{booktabs}
+% \usepackage{pslatex}
+% \usepackage{subcaption}
+% \usepackage{pgfplots}
+% \usepackage{wrapfig}
+% \usepackage[english]{babel}
+% \usepackage{lmodern}
+% % \doublespace
+% % \usepackage{url}
+% \usepackage{bigfoot}
+% \usepackage[export]{adjustbox}
+% % \setlength\intextsep{0pt}
+
+% \usepackage{graphicx}
+
+\documentclass[12pt, letterpaper]{article}
+\usepackage{fontspec} 
+\usepackage{marvosym}
+\usepackage{setspace}
+\usepackage{longtable}
+% DOCUMENT LAYOUT
+% \usepackage{geometry} 
+\usepackage[margin=0.90in]{geometry}
+\usepackage{lineno}
+\usepackage[usenames,dvipsnames]{xcolor}
+% \geometry{letterpaper, textwidth=5.5in, textheight=8.5in, marginparsep=7pt, marginparwidth=.6in}
+% \setlength\parindent{0in}
+
+\definecolor{linenocolor}{gray}{0.6}
+\renewcommand\thelinenumber{\color{linenocolor}\arabic{linenumber}}
+% \calclayout %centers text on page.
+
+% \renewcommand{\thesubsection}{\arabic{subsection}.}
+% FONTS
+% \usepackage[usenames,dvipsnames]{color}
+% \usepackage{xunicode}
+% \usepackage{xltxtra}
+\defaultfontfeatures{Mapping=tex-text}
+% \setromanfont[Ligatures={Common}, Variant=01]{Hoefler Text}
+% \setromanfont[Ligatures={Common}, Variant=01]{Palatino}
+\setromanfont[Ligatures={Common}, Variant=01]{Baskerville}
+% \setromanfont[Ligatures={Common}, Variant=01]{Plantagenet Cherokee}
+% \setmainfont [BoldFont={STIX Two Text SemiBold}, Ligatures={Common},
+% Numbers={OldStyle}, Variant=01]{STIX Two Text}
+% \setmainfont [BoldFont={STIX Two Text SemiBold}, Ligatures={Common},
+% Numbers={OldStyle}, Variant=01]{STIX Two Text}
+% \setromanfont [Ligatures={Common},
+% Numbers={OldStyle}, Variant=01]{STIX Two Text}
+\setmonofont[Scale=0.95]{PT Sans}
+
+% ---- CUSTOM COMMANDS
+% \chardef\&="E050
+\newcommand{\html}[1]{\href{#1}{\scriptsize\textsc{[html]}}}
+\newcommand{\pdf}[1]{\href{#1}{\scriptsize\textsc{[pdf]}}}
+\newcommand{\doi}[1]{\href{#1}{\scriptsize\textsc{[doi]}}}
+\newcommand{\lurl}[1]{\href{#1}{\scriptsize\textsc{[link]}}}
+
+\newcommand{\githubsymbol}{\raisebox{-2pt}{\includegraphics[height=10.5pt]{github}}}
+\newcommand{\twittersymbol}{\raisebox{-2pt}{\includegraphics[height=10.5pt]{twitter}}}
+\newcommand{\monitorsymbol}{\raisebox{-2pt}{\includegraphics[height=10.5pt]{monitor}}}
+
+% ---- MARGIN YEARS
+\usepackage{marginnote}
+\newcommand{\amper{}}{\chardef\amper="E0BD }
+\newcommand{\years}[1]{\marginnote{\small #1}}
+\renewcommand*{\raggedleftmarginnote}{}
+\setlength{\marginparsep}{7pt}
+\reversemarginpar
+
+% HEADINGS
+\usepackage{sectsty} 
+\usepackage[normalem]{ulem} 
+\sectionfont{\upshape\Large}
+\subsectionfont{\scshape\normalsize} 
+\subsubsectionfont{\upshape\large}
+
+% PDF SETUP
+% ---- FILL IN HERE THE DOC TITLE AND AUTHOR
+\usepackage[%dvipdfm, 
+bookmarks, colorlinks, breaklinks, 
+% ---- FILL IN HERE THE TITLE AND AUTHOR
+	pdftitle={Matthew A. Turner - Vita},
+	pdfauthor={Matthew A. Turner, PhD},
+	pdfproducer={https://mt.digital}
+]{hyperref}  
+\hypersetup{linkcolor=blue,citecolor=blue,filecolor=black,urlcolor=MidnightBlue}
+
+\usepackage{apacite}
 \usepackage{booktabs}
-\usepackage{pslatex}
-\usepackage{apacite}
 \usepackage{authblk}
-\usepackage{subcaption}
-\usepackage{pgfplots}
-\usepackage{wrapfig}
-\usepackage[english]{babel}
-\usepackage{lmodern}
-% \usepackage{setspace}
-% \doublespace
-% \usepackage{url}
-\usepackage{bigfoot}
-\usepackage[export]{adjustbox}
-% \setlength\intextsep{0pt}
 
-\usepackage{graphicx}
+\begin{document}
 
 \title{Group polarization is often a mirage caused by inappropriate statistical analyses}
-\shorttitle{Group polarization or simple consensus?}
+% \shorttitle{Group polarization or simple consensus?}
 \author[1,2]{{Matthew A.~Turner}}
@@
 
-\abstract{
+\maketitle
+% \setlength{\parindent}{15pt}
+% \maketitle
+
+\spacing{1.5}
+
+\linenumbers
+\modulolinenumbers[5]
+
+\begin{abstract}
+  \noindent
   Group polarization is the observed tendency of
@@
   future group-polarization research.
-}
-
-\begin{document}
-\maketitle
-\setlength{\parindent}{15pt}
-% \maketitle
+\end{abstract}
 
@@
 
-Recently, problems with using have been made clear~\citep{Liddell2018}.
+Recently, problems with using have been made clear~\cite{Liddell2018}.
 The way to fix this problem going forward is to use a thresholded-normal 
 statistical model for group polarization data, or other related statistical 
-model that accounts for opinion binning~\citep[Ch. 23]{Liddell2018,KruschkeDBDA}.
+model that accounts for opinion binning~\cite[Ch. 23]{Liddell2018,KruschkeDBDA}.
 
@@
 were from a neutral initial group opinion to a biased group opinion 
-(as in, e.g.,~\citet{Abrams1990} Uncategorized2), or from
-an opinion biased in one direction (as in, e.g.,~\citet{Abrams1990} Categorized1,
-Categorized2, AND OTHERS???). In some results of \citet{Freidkin1999a}, 
+(as in, e.g.,~\citeA{Abrams1990} Uncategorized2), or from
+an opinion biased in one direction (as in, e.g.,~\citeA{Abrams1990} Categorized1,
+Categorized2, AND OTHERS???). In some results of \citeA{Freidkin1999a}, 
 shift values are reported with their
@@
 processes known to generally occur when opinions within a group significantly
-diverge~\citep{Turner2018}.
+diverge~\cite{Turner2018}.
 
@@
 post-deliberation mean opinions, and the number of participants
-in each treatment (\citet{Burnstein1975} did not report pre- and post-deliberation
+in each treatment (\citeA{Burnstein1975} did not report pre- and post-deliberation
 means, only the shift in means). We did not include experimental treatments that did not 
@@
 In two cases, reported opinion and shift values were converted from an ordinal
-scale to a percentage scale, specifically \citet{Friedkin1999a} and \citet{Krizan2007}.
+scale to a percentage scale, specifically \citeA{Friedkin1999a} and \citeA{Krizan2007}.
 The authors in these papers did report the original ordinal measurement scale,
 so we converted the means back to the orignial measurement scale.
-In a few cases from \citet{Friedkin1999a} we did not need to apply our 
+In a few cases from \citeA{Friedkin1999a} we did not need to apply our 
 model to the reported data because we could determine a reported opinion shift was 
@@
 
+\spacing{1.0}
+\input{plausibleFP-table.tex}
+\spacing{1.5}
+
 \subsection{Do $t$-tests accurately detect simple consensus?}
@@
 When simulated data reliably result in false positive $t$-test results (where
-the $t$-test $p \le 0.1$) this strengthens the case that.
+the $t$-test $p \le 0.1$) this strengthens the case that reported detections
+of a false group polarization detection.
 
-One complicating factor to this approach is that statistical power, $W$, varies
-between treatments. Therefore, simulated data should yield statistically-significant
-group polarization with frequency close to $W$ for a given treatment at a given
-significance value, $\alpha$.
+\spacing{1.0}
+% \input{t-test-table.tex}
+\input{t-test-table-custom.tex}
+\spacing{1.5}
 
@@
 
-\paragraph{\citet{Abrams1990}} This study is meant to support the presence of the "self-categorization" effect in
+\paragraph{{Abrams1990}} This study is meant to support the presence of the "self-categorization" effect in
 social psychology. The paper presents three experiments and their results--Experiment 3 is the group polarization
@@
 
-\paragraph{\citet{Burnstein1973}} First I want to note this study was funded in part by a Guggenheim fellowship (see
+\paragraph{{Burnstein1973}} First I want to note this study was funded in part by a Guggenheim fellowship (see
 note on first page) and the authors acknolwedge the assistance of "vinous" discussions with colleagues at the
@@
 
-\paragraph{\citet{Burnstein1975}} In this study, Burnstein and Vinokur ran an experiment to evaluate their theory
+\paragraph{{Burnstein1975}} In this study, Burnstein and Vinokur ran an experiment to evaluate their theory
 that generating persuasive arguments are necessary for group polarization to occur. The authors design a study with
@@
 
-\paragraph{\citet{Schkade2010}} Friedkin uses three CDQ topics, "Sports", "School", and "Surgery" that use a
+\paragraph{{Schkade2010}} Friedkin uses three CDQ topics, "Sports", "School", and "Surgery" that use a
 20-point CDQ risk scale in fractional units, 0.05, 0.10, ..., 1.00 across three group sizes (2, 3, and 4, which he
@@
 
-\paragraph{\citet{Hogg1990}} It is not clear if there are indeed significant
+\paragraph{{Hogg1990}} It is not clear if there are indeed significant
 pre-post deliberation differences reported in this paper. The Results are a
@@
 
-\paragraph{\citet{Krizan2007}} Uses five CDQ topics ("scenarios") from Kogan and
+\paragraph{{Krizan2007}} Uses five CDQ topics ("scenarios") from Kogan and
 Wallach (1967) that typically provide shifts toward more risky decisions.  These are
@@
 
-\paragraph{\citet{Moscovici1969}} Moscovici and Zavalloni studied whether group
+\paragraph{{Moscovici1969}} Moscovici and Zavalloni studied whether group
 polarization is more general than CDQ by using opinions about political topics. There
@@
 
-\paragraph{\citet{Myers1970}} Myers and Bishop test whether group polarization occurs
+\paragraph{{Myers1970}} Myers and Bishop test whether group polarization occurs
 in groups with different levels of racial bias determined by a survey that is not also
@@
 
-\paragraph{\citet{Myers1975}} Myers performs two experiments each with experimental and control conditions for two group types with initial biases.
+\paragraph{{Myers1975}} Myers performs two experiments each with experimental and control conditions for two group types with initial biases.
 In Experiment 1, Myers creates biased groups by having participants (intro psychology students) read a description of six different teachers, three "bad" professors and three "good" professors, which the participants are assumed to infer through vignettes about the professors.
@@
 
-\paragraph{\citet{Schkade2010}} Participants in Schkade, Sunstein, and Hastie's study were put into two politically ideologically biased groups based on their geographical location in the US state of Colorado: participants were drawn from Boulder, Colorado (abbreviated CO), to be liberally biased and from Colorado Springs, CO, to be conservatively biased.
+\paragraph{{Schkade2010}} Participants in Schkade, Sunstein, and Hastie's study were put into two politically ideologically biased groups based on their geographical location in the US state of Colorado: participants were drawn from Boulder, Colorado (abbreviated CO), to be liberally biased and from Colorado Springs, CO, to be conservatively biased.
 Participants gave their opinions and discussed three topics related to Global Warming, Civil Unions, and Affirmative Action.

### 2022-11-18 GroupPolarizationStatmod/analysis.R

@@ -1,5 +1,7 @@
 library(dplyr)
+library(tidystats)
 
+source("experiments.R")
 ##
 # Arguments:
 #  studiesAnalysisDfPath (string): path where exact analysis was done using web app
@@ -31,7 +33,7 @@ makePlausibleFPTable <- function(studiesAnalysisDfPath = "data/StudiesAnalysis.c
 #    treatments yielding plausibly false positive results.
 #  limit (int): Limit the number of treatments to test for initial development.
 #
-makeTtestReliabilityTable <- function(studiesAnalysisDfPath = "data/StudiesAnalysis.csv",
+makeTtestSimulationsTable <- function(studiesAnalysisDfPath = "data/StudiesAnalysis.csv",
                                       outputPath = "data/output/TtestFitTable.csv",
                                       diagnosticSavePath = "data/diagnostic/tTestFits.RDS",
                                       ntrials = 2, limit = 2)
@@ -40,7 +42,7 @@ makeTtestReliabilityTable <- function(studiesAnalysisDfPath = "data/StudiesAnaly
   studiesDf <- read.csv(studiesAnalysisDfPath)
 
   # Filter out implausible false positives.
-  studiesDf <- filter(studiesDf, (IncludeInt == 1) && (PlausibleInt == 1))
+  studiesDf <- filter(studiesDf, (IncludeInt == 1) & (PlausibleInt == 1))
   
   # Limit number of plausible false positives for preliminary development.
   if (limit > 0)
@@ -97,7 +99,21 @@ makeTtestReliabilityTable <- function(studiesAnalysisDfPath = "data/StudiesAnaly
   write.csv(resultsDf, outputPath)
   saveRDS(allTestsDiagnostic, file = diagnosticSavePath)
   
-  return (resultsDf)
+  # return (resultsDf)
+}
+
+summarizeTTestFitTable <- function(fitTablePath = "data/output/TtestFitTable.csv",
+                                   summaryTTablePath = "data/output/TtestSummaryTable.csv",
+                                   significanceVal = 0.1)
+{
+  fitTableDf <- 
+    read.csv(fitTablePath) %>%
+    group_by(ArticleTag, TreatmentTag, ExpectedPower) %>%
+    summarize(FractionSignificant = mean(tTestPvalue < significanceVal), .groups='keep')
+  
+  write.csv(fitTableDf, summaryTTablePath, row.names = FALSE)
+  
[…]

### 2022-11-18 GroupPolarizationStatmod/experiments.R

@@ -133,11 +133,6 @@ makeBayesianFitTable <- function(studies.data.csv = "data/StudiesAnalysis.csv",
                 LatentMeanPostPosteriorSD = muPostPost["SD"]
             )
 
-            # print("")
-            # print("Tibble row:")
-            # print(tibbleRow)
-            # print("")
-
             outputDf <- add_row(outputDf, tibbleRow)
         }
     }
@@ -184,8 +179,6 @@ calculateBayesian <- function(N, firstBinValue, nBins, latentMean, latentSd)
 tTestExperiment <- function(N, firstBinValue, nBins, latentMean, latentSdPre, 
                             latentSdPost, paired = FALSE, var.equal = FALSE)
 {
-  print(latentMean)
-  print(latentSdPre)
   simObsPre <- simulatedObservation(N, firstBinValue, nBins, latentMean, latentSdPre)
 
   simObsPost <- simulatedObservation(N, firstBinValue, nBins, latentMean, latentSdPost)
@@ -211,11 +204,6 @@ simulatedObservation <- function(N, firstBinValue, nBins, latentMean, latentSd)
   # to empirical, as opposed to population, values.
   
   # mvrnorm uses variance, not standard deviation, for its spread parameter.
-  # latentVariance <- latentSd^2
-  # latentData <- mvrnorm(N, latentMean, latentVariance, empirical=TRUE)
-  print(N)
-  print(latentMean)
-  print(latentSd)
   latentData <- rnorm(N, latentMean, latentSd)
 
   # Create the bins and thresholds to be used to bin latentData.

### 2022-11-19 GroupPolarizationStatmod/paper/ms.tex
finished refactor FP and t-test tables

--- 2022-11-18 GroupPolarizationStatmod/papers/ms.tex
@@
-% % Author: Matthew Turner
-
-% % \documentclass[11pt,letterpaper]{scrartcl}
-% % \documentclass[letterpaper,man,natbib]{apa6}
-% \documentclass[11pt,letterpaper]{amsart}
-% % \documentclass[11pt, letterpaper]{article}
-% % \usepackage{natbib}
-% % \documentclass[11pt]{report}
-% % \documentclass{report}
-% % \documentclass{book}
-% \usepackage[bookmarks, hidelinks]{hyperref}
-% \usepackage{amssymb,amsmath}
-% % \usepackage{fullpage}
-% \usepackage{tabulary}
-% \usepackage{tabularx}
-% \usepackage{float}
-% % \usepackage[margin=1.00in]{geometry}
-% \usepackage[margin=0.90in]{geometry}
-
-% % \usepackage{fontspec}
-
-% % \usepackage{xunicode}
-% % \usepackage{xltxtra}
-% % \defaultfontfeatures{Mapping=tex-text}
-% % \setmainfont[Ligatures={Common}, Numbers={OldStyle}, Variant=01]{Hoefler Text}
-% % \setromanfont[Ligatures={Common}, Numbers={OldStyle}, Variant=01]{Hoefler Text}
-% % \setmonofont[Scale=0.95]{PT Sans}
-% \usepackage{caption}
-% \usepackage{apacite}
-% \usepackage{booktabs}
-% \usepackage{pslatex}
-% \usepackage{subcaption}
-% \usepackage{pgfplots}
-% \usepackage{wrapfig}
-% \usepackage[english]{babel}
-% \usepackage{lmodern}
-% % \doublespace
-% % \usepackage{url}
-% \usepackage{bigfoot}
-% \usepackage[export]{adjustbox}
-% % \setlength\intextsep{0pt}
-
-% \usepackage{graphicx}
-
-\documentclass[12pt, letterpaper]{article}
-\usepackage{fontspec} 
+\documentclass[11pt, letterpaper]{article}
 \usepackage{marvosym}
@@
 % \usepackage{geometry} 
-\usepackage[margin=0.90in]{geometry}
+\usepackage[margin=1.0in]{geometry}
 \usepackage{lineno}
@@
 
+% \usepackage{fontspec} 
 % \renewcommand{\thesubsection}{\arabic{subsection}.}
@@
 % \usepackage{xltxtra}
-\defaultfontfeatures{Mapping=tex-text}
-% \setromanfont[Ligatures={Common}, Variant=01]{Hoefler Text}
-% \setromanfont[Ligatures={Common}, Variant=01]{Palatino}
-\setromanfont[Ligatures={Common}, Variant=01]{Baskerville}
-% \setromanfont[Ligatures={Common}, Variant=01]{Plantagenet Cherokee}
-% \setmainfont [BoldFont={STIX Two Text SemiBold}, Ligatures={Common},
-% Numbers={OldStyle}, Variant=01]{STIX Two Text}
-% \setmainfont [BoldFont={STIX Two Text SemiBold}, Ligatures={Common},
-% Numbers={OldStyle}, Variant=01]{STIX Two Text}
-% \setromanfont [Ligatures={Common},
-% Numbers={OldStyle}, Variant=01]{STIX Two Text}
-\setmonofont[Scale=0.95]{PT Sans}
+% \defaultfontfeatures{Mapping=tex-text}
+% % \setromanfont[Ligatures={Common}, Variant=01]{Hoefler Text}
+% % \setromanfont[Ligatures={Common}, Variant=01]{Palatino}
+% \setromanfont[Ligatures={Common}, Variant=01]{Baskerville}
+% % \setromanfont[Ligatures={Common}, Variant=01]{STIX Two Math}
+% % \setmathrm{Baskerville}
+% % \setromanfont[Ligatures={Common}, Variant=01]{Plantagenet Cherokee}
+% % \setmainfont [BoldFont={STIX Two Text SemiBold}, Ligatures={Common},
+% % Numbers={OldStyle}, Variant=01]{STIX Two Math}
+% % \setmathrm{STIX Two Math}
+% % \setmainfont [BoldFont={STIX Two Text SemiBold}, Ligatures={Common},
+% % Numbers={OldStyle}, Variant=01]{STIX Two Text}
+% % \setromanfont [Ligatures={Common},
+% % Numbers={OldStyle}, Variant=01]{STIX Two Text}
+% \setmonofont[Scale=0.95]{PT Sans}
 
-% ---- CUSTOM COMMANDS
-% \chardef\&="E050
-\newcommand{\html}[1]{\href{#1}{\scriptsize\textsc{[html]}}}
-\newcommand{\pdf}[1]{\href{#1}{\scriptsize\textsc{[pdf]}}}
-\newcommand{\doi}[1]{\href{#1}{\scriptsize\textsc{[doi]}}}
-\newcommand{\lurl}[1]{\href{#1}{\scriptsize\textsc{[link]}}}
-
-\newcommand{\githubsymbol}{\raisebox{-2pt}{\includegraphics[height=10.5pt]{github}}}
-\newcommand{\twittersymbol}{\raisebox{-2pt}{\includegraphics[height=10.5pt]{twitter}}}
-\newcommand{\monitorsymbol}{\raisebox{-2pt}{\includegraphics[height=10.5pt]{monitor}}}
-
-% ---- MARGIN YEARS
-\usepackage{marginnote}
-\newcommand{\amper{}}{\chardef\amper="E0BD }
-\newcommand{\years}[1]{\marginnote{\small #1}}
-\renewcommand*{\raggedleftmarginnote}{}
-\setlength{\marginparsep}{7pt}
-\reversemarginpar
+\usepackage{mathspec}
+% \setallmainfonts(Digits,Latin){STIX Two Text}
+% \setallmainfonts(Digits,Latin){Baskerville}
+% \setallmainfonts(Digits,Latin){Plantagenet Cherokee}
+\setallmainfonts(Digits,Latin){Palatino}
+% \setallmainfonts(Digits,Latin){Hoefler Text}
 
@@
 % ---- FILL IN HERE THE TITLE AND AUTHOR
-	pdftitle={Matthew A. Turner - Vita},
-	pdfauthor={Matthew A. Turner, PhD},
+	pdftitle={Many claims of group polarization are plausibly false due to inappropriate
+    statistics},
+	pdfauthor={Matthew A. Turner},
 	pdfproducer={https://mt.digital}
@@
 
-\title{Group polarization is often a mirage caused by inappropriate statistical analyses}
+\title{Many claims of group polarization are plausibly false due to inappropriate
+    statistics}
 % \shorttitle{Group polarization or simple consensus?}
@@
 
+\begin{equation}
+  \int_{-\infty}^\infty x^2 dx \neq \pi
+\end{equation}
+\noindent
+
 \subsubsection{Identifying parameters that plausibly generate false detections of group polarization}
@@
 
+\vspace{.5em}
 \spacing{1.0}
-\input{plausibleFP-table.tex}
+% \input{plausibleFP-table.tex}
+  \begin{table}
+    \begin{center}
+      \caption{Number and fraction of experimental conditions where group polarization was
+               reported to have occurred, but plausibly did not.}
+      \vspace{.5em}
+      \input{plausibleFP-table-custom.tex}
+    \end{center}
+  \end{table}
 \spacing{1.5}

### 2022-11-19 GroupPolarizationStatmod/analysis.R

@@ -1,5 +1,7 @@
 library(dplyr)
 library(tidystats)
+library(knitr)
+library(kableExtra)
 
 source("experiments.R")
 
@@ -16,7 +18,7 @@ makePlausibleFPTable <- function(studiesAnalysisDfPath = "data/StudiesAnalysis.c
     read.csv(studiesAnalysisDfPath) %>%
     filter(IncludeInt == 1) %>% 
     group_by(ArticleTag) %>%
-    summarize(PlausibleFPRate = mean(PlausibleInt))
+    summarize(PlausibleFPs = sum(PlausibleInt), ConditionCount = n(), PlausibleFPRate = mean(PlausibleInt), )
   
   write.csv(plausibleFPTable, file = "data/output/PlausibleFPTable.csv", row.names = FALSE)
   
@@ -118,6 +120,31 @@ summarizeTTestFitTable <- function(fitTablePath = "data/output/TtestFitTable.csv
   
   return (fitTableDf)
 }
+
+
+latexifyPlausibleFPTable <- function(plausibleFPTablePath = "data/output/PlausibleFPTable.csv",
+                                     latexifiedPath = "papers/plausibleFP-table.tex")
+{
+    plausibleFP_table <- read.csv(plausibleFPTablePath)
+    plausibleFP_latex <- 
+        kable(plausibleFP_table, "latex", longtable = T, booktabs = T, digits=2) %>%
+        kable_styling(latex_options = c("repeat_header"))
+
+    writeLines(plausibleFP_latex, latexifiedPath) 
+}
+
+
+latexifyTTestExperiment <- function(tTestTablePath = "data/output/TtestSummaryTable.csv",
+                                    latexifiedPath = "papers/t-test-table.tex")
+{
+    tTestTable <- read.csv(tTestTablePath)
+    ttestTable_latex <- 
+        kable(tTestTable, "latex", longtable = T, booktabs = T, digits=2) %>%
+        kable_styling(latex_options = c("repeat_header"))
+
+    writeLines(ttestTable_latex, latexifiedPath)
+}
+
 
 doAnalyses <- function(studiesAnalysisDfPath = "data/StudiesAnalysis.csv")
 {

### 2022-11-20 GroupPolarizationStatmod/paper/ms.tex
expanded t-test table
added quotes from dissertation

@@
 % \calclayout %centers text on page.
-
-% \usepackage{fontspec} 
-% \renewcommand{\thesubsection}{\arabic{subsection}.}
-% FONTS
-% \usepackage[usenames,dvipsnames]{color}
-% \usepackage{xunicode}
-% \usepackage{xltxtra}
-% \defaultfontfeatures{Mapping=tex-text}
-% % \setromanfont[Ligatures={Common}, Variant=01]{Hoefler Text}
-% % \setromanfont[Ligatures={Common}, Variant=01]{Palatino}
-% \setromanfont[Ligatures={Common}, Variant=01]{Baskerville}
-% % \setromanfont[Ligatures={Common}, Variant=01]{STIX Two Math}
-% % \setmathrm{Baskerville}
-% % \setromanfont[Ligatures={Common}, Variant=01]{Plantagenet Cherokee}
-% % \setmainfont [BoldFont={STIX Two Text SemiBold}, Ligatures={Common},
-% % Numbers={OldStyle}, Variant=01]{STIX Two Math}
-% % \setmathrm{STIX Two Math}
-% % \setmainfont [BoldFont={STIX Two Text SemiBold}, Ligatures={Common},
-% % Numbers={OldStyle}, Variant=01]{STIX Two Text}
-% % \setromanfont [Ligatures={Common},
-% % Numbers={OldStyle}, Variant=01]{STIX Two Text}
-% \setmonofont[Scale=0.95]{PT Sans}
 
@@
 \end{abstract}
+
+\begin{quote}
+In our introductory social psychology course, 
+we have for many years used the [group polarization experimental paradigm] as
+a laboratory exercise. The exercise works beautifully, but one must be
+careful to forewarn a class that [group polarization] does not occur with every 
+group\ldots and that the effect is not large. 
+\par\raggedleft\cite[p. 205]{Brown1986}
+\end{quote}
+
+\begin{quote}
+One of the most robust findings in social psychology is that of attitude polarization 
+following discussion with like-minded others.
+\par\raggedleft\cite[p. 267]{Cooper2001}
+\end{quote}
 

### 2022-11-20 GroupPolarizationStatmod/paper/ms.tex
some progress writing up model and outlining results/discussion

@@
 \usepackage[usenames,dvipsnames]{xcolor}
+\definecolor{myorange}{RGB}{240, 96, 0}
+\newcommand{\mt}[1]{{\textcolor{myorange} {({\tiny MT:} #1)}}}
 % \geometry{letterpaper, textwidth=5.5in, textheight=8.5in, marginparsep=7pt, marginparwidth=.6in}
@@
 \usepackage{authblk}
+
+\newcommand{\sigmapre}{\sigma_\mathrm{pre}}
+\newcommand{\sigmapost}{\sigma_\mathrm{post}}
+\newcommand{\normalpre}{\mathcal{N}(\mu, \sigmapre)}
+\newcommand{\normalpost}{\mathcal{N}(\mu, \sigmapost)}
 
@@
 relevant forms of consensus: \emph{simple consensus} and \emph{group polarization}.
-Either process is \emph{consensus} since opinion variance among the group 
-decreases as individuals come to agree more than they did previously. 
-Group polarization experiments are designed to foster group consensus
-it is not necessary to consider clustering, negative influence, or other
-processes known to generally occur when opinions within a group significantly
-diverge~\cite{Turner2018}.
+Either process is \emph{consensus}, if we define consensus as a reduction
+in opinion variance over time.  
+Group polarization experiments are designed to result group consensus
+since they group together like-minded individuals to see if they become more
+extreme after discussion~\cite{Turner2018}.
 
 We collected group polarization data from ten empirical studies with sixty constituent
-experimental treatments across all studies, specifically the opinion shift and pre- and
+experimental conditions across all studies. We recorded the reported opinion shift and pre- and
 post-deliberation mean opinions, and the number of participants
-in each treatment (\citeA{Burnstein1975} did not report pre- and post-deliberation
-means, only the shift in means). We did not include experimental treatments that did not 
+in each condition, as long as this data was available---\citeA{Burnstein1975} 
+did not report pre- and post-deliberation
+means, only the shift in means. We did not include experimental treatments that did not 
 yield a positive detection of group polarization since we are only estimating
 the rate of plausible false positive detections of group polarization.
-We then created new data that indicated whether a reported.
 
-In two cases, reported opinion and shift values were converted from an ordinal
-scale to a percentage scale, specifically \citeA{Friedkin1999a} and \citeA{Krizan2007}.
-The authors in these papers did report the original ordinal measurement scale,
-so we converted the means back to the orignial measurement scale.
-In a few cases from \citeA{Friedkin1999a} we did not need to apply our 
-model to the reported data because we could determine a reported opinion shift was 
-\emph{prima facie} false since reported opinion shift magnitudes
-were smaller than the standard deviation in shifts (see Table 1, p. 868, \emph{ibid}),
-despite Friedkin's claim that regressions (not clearly defined) on all shifts 
-yielded $p < 0.05$~\cite{Kruschke2018c}.
+Based on this data
 
-\subsection{Latent opinions, ordinal measurements}
+% In two cases, reported opinion and shift values were converted from an ordinal
+% scale to a percentage scale, specifically \citeA{Friedkin1999a} and \citeA{Krizan2007}.
+% The authors in these papers did report the original ordinal measurement scale,
+% so we converted the means back to the orignial measurement scale.
+% In a few cases from \citeA{Friedkin1999a} we did not need to apply our 
+% model to the reported data because we could determine a reported opinion shift was 
+% \emph{prima facie} false since reported opinion shift magnitudes
+% were smaller than the standard deviation in shifts (see Table 1, p. 868, \emph{ibid}),
+% despite Friedkin's claim that regressions (not clearly defined) on all shifts 
+% yielded $p < 0.05$~\cite{Kruschke2018c}.
+
+\subsection{Latent opinions, ordinal measurements, and how simple consensus can look like
+group polarization}
 
 Our analysis relies on two important formal concepts: latent opinions and their 
-measurement on an ordinal scale. \emph{Latent opinions} are unobservable psychological,
-continuous-valued variables that are somehow represented and stored in the
-brain-body. 
-When researchers use continuous statistical models to infer 
-that group polarization occurred (which all ten considered here do), 
-they are implicitly assuming that latent opinions are themselves continuous.
-We can only measure behaviors in group polarization experiments,
-specifically the behavior of participants marking one ordinal value or another
-to indicate their opinion---a process that is not well understood in itself,
-but which is unimportant for our analyses. \emph{Ordinal measurements} (e.g.\
-a Likert or other discrete-valued opinion scale) of latent opinions force
-participants to appropriately bin their opinion in one of several categories
-representing the valence and extremity of their opinion, e.g., on a
-7-point Likert scale -3 could indcate
-``strongly disagree'', +3 could be ``strongly agree'', 0 would be neutral, and 
-intermediate values are less extreme opinions of either valence.
+measurement on an ordinal scale. \emph{Latent opinions} are theoretical,
+unobservable, psychological, continuous-valued variables that are 
+somehow represented and stored in the brain-body, and can be reproduced with
+some accuracy when a person is prompted. \emph{Ordinal measurements} of 
+of latent opinions are ordinal-valued representations of latent opinions 
+in some physical form, such as a mark on a sheet of paper or radio button on a web form, that
+indicate the polarity and strength of a person's opinion opinion in terms of
+one of a number of opinion bin values.
+
+For example, on a 7-point ordinal scale -3 could indcate 
+``strongly disagree'', +3 could be ``strongly agree'', 0 would be psychologically 
+neutral, and intermediate values are less extreme opinions of either valence, 
+such as was used by \citeA{Moscovici1969}. Or, one might
 
@@
 
-\subsection{Simple consensus and group polarization are statistical patterns}
+\subsection{Evaluating whether a group polarization finding is plausibly false}
 
-\subsection{Analytical approach}
+We used large-$N$ and generative modeling supported by $t$-tests on simulated
+data to demonstrate that many published findings are plausibly false.
 
-The goal of this study is to understand if published detections of group polarization are
-plausibly false.
-
-Therefore we designed a series of tests to understand if, and to what degree,
-published results are false positive detections of group polarization that
-are at least as well described by a simple-consensus model.
-
-\begin{equation}
-  \int_{-\infty}^\infty x^2 dx \neq \pi
-\end{equation}
-\noindent
-
-\subsubsection{Identifying parameters that plausibly generate false detections of group polarization}
+\subsubsection{Identify null-model parameters that yield plausibly-false positives}
 
@@
 
-\subsubsection{Bayesian ordered-probit models of generated simple-consensus data}
+After identifying null-model parameters that yield plausibly-false positives 
+in the $N \to \infty$ case, we generated binned samples from normal opinion 
+distributions $\normalpre$ and $\normalpost$.
+
+\subsection{Bayesian ordered-probit models reliably identify simple consensus}
 
@@
 
-We found that in X\% of generated simple consensus data led to false positive
-detections
-
-When simulated data reliably result in false positive $t$-test results (where
-the $t$-test $p \le 0.1$) this strengthens the case that reported detections
-of a false group polarization detection.
-
 \spacing{1.0}
-% \input{t-test-table.tex}
-\input{t-test-table-custom.tex}
+\input{t-test-table.tex}
+% \input{t-test-table-custom.tex}
 \spacing{1.5}
@@
 polarization experiments. This makes it impossible to systematically account for
-additional sources of vairance known to lead to (orthogonal) over-estimates
-of effect sizes~\cite{Yarkoni2021}. This is due to the rush to explain why 
-group polarization occurs using weak theory~\cite{Cartwright1971,Cartwright1973} and makes
-group polarization meta-analyses (e.g. Brown, 1986; and Isenberg, 1986)\nocite{Brown1986,Isenberg1986}
-unreliable~\cite{Meehl1990,Meehl1997}.
+additional sources of vairance known to lead to over-estimates
+of effect sizes~\cite{Yarkoni2021}, independent of the false positives problem
+we identified in the current paper.
+
+% This is due to the rush to explain why 
+% group polarization occurs using weak theory~\cite{Cartwright1971,Cartwright1973} and makes
+% group polarization meta-analyses (e.g. Brown, 1986; and Isenberg, 1986)\nocite{Brown1986,Isenberg1986}
+% unreliable~\cite{Meehl1990,Meehl1997}.
 

### 2022-11-20 GroupPolarizationStatmod/paper/ms.tex
ready to start on model figures for n->infty and t-test analyses

@@
 
+\usepackage{amsmath}
 \usepackage{mathspec}
@@
 
-All ten studies analyzed here used an ordinal scale, but group polarization experiments seem to
-lack any systematicity in experimental design beyond that. (EXAMPLES)
+We assume that participant $i$'s internal, latent psychological opinion at 
+time of reporting ($t \in \{pre, post\}$) is drawn from a normal distribution
+with mean $\mu_t$ and standard deviation $\sigma_t$,
+\begin{equation}
+  o_{i,t} \sim \mathcal{N}(\mu_t, \sigma_t).
+  \label{eq:opinionDistribution}
+\end{equation}
+\noindent
+This is the condition-level distribution of opinions at time, $t$, chosen to
+represent the common experimental design used by group polarization studies,
+not because it is statistically appropriate---in fact it is not, which is what
+we aim to show in this paper.
 
@@
 
+We model measurement of $N \to \infty$ participant opinions,
+which results in a histogram of frequency of opinions
+in each bin, i.e., $o_{i,t} = b_k$. The frequency of responses in each bin is 
+the integral over the continuous normal probability density function from 
+one bin threshold to another. This transforms the probability density 
+function from $p(o_{i,t};~\mu_t, \sigma_t)$ to the probability
+of observing a reported opinion of each bin value. 
+The probability of observing an opinion in bin
+$b_k$ is calculated by integrating the normal probability density 
+function over the range of $\theta_{k-1}$ to $\theta_{k}$. 
+Formally, we write the 
+probability of observing an opinion in bin $k$ at time $t$ as
+\begin{equation}
+\begin{aligned}
+  p(o = b_k;\mu_t, \sigma_t, \theta) 
+    &= \int_{\theta_{k-1}}^{\theta_k} p(o;\mu_t, \sigma_t) do \\
+    &= \Phi \left( \frac{o - \mu_t}{\sigma_t} \right)\Big|_{o=\theta_{k-1}}^{\theta_k} \\
+    &= \Phi \left( \frac{\theta_k - \mu_t}{\sigma_t} \right) - 
+       \Phi \left( \frac{\theta_{k-1} - \mu_t}{\sigma_t} \right)
+\end{aligned}
+\label{eq:binFrequency}
+\end{equation}
+\noindent
+where $\Phi(\frac{o - \mu}{\sigma})$ is the normalized 
+normal cumulative distribution function over opinions $o$, shifted by an amount $\mu$ with
+standard deviation $\sigma$.
 \subsubsection{Identify null-model parameters that yield plausibly-false positives}
@@
 \spacing{1.0}
+\hspace{-1em}
+\begin{center}
 \input{t-test-table.tex}
+\end{center}
 % \input{t-test-table-custom.tex}

### 2022-11-21 GroupPolarizationStatmod/paper/ms.tex
added table with experiment conditions info and GP schematic figure

@@
 \usepackage{longtable}
+\usepackage{pdflscape}
+\usepackage{graphicx}
 % DOCUMENT LAYOUT
@@
 \end{abstract}
+\clearpage
 
@@
 
+\begin{itemize}
+  \item 
+    \emph{Group polarization} is a process where the mean opinion of a 
+    like-minded group becomes more extreme over time~\cite{Brown1986,Brown2000}. 
+    It is therefore an important component process in
+    the broader process of increasing political polarization that can hamstring
+    governments~\cite{Mason2018UncivilAgreementBook,Klein2020}.
+  \item
+    Individuals' opinions in group polarization experiments are typically 
+    measured using an ordinal scale, such
+    as a Likert scale, where behavioral experiment participants' internal, theoretically
+    continuous, opinions are converted to one of several opinion bins indiciating
+    their degree of agreement or disagreement with some prompt statement.
+  \item
+    Recently, however, it has been shown that it is critical to use 
+    appropriate statistical methods when using experimental designs that 
+    measure continuous latent psychological variables with ordinal
+    measurements---failure to do so can result in false positive support for
+    the presence of experimental effects~\cite{Liddell2018}.
+  \item
+    In this paper, we develop a generative modeling approach for evaluating whether or
+    not published detections of group polarization are plausibly false, i.e.,
+    whether they are \emph{plausibly-false positives} (PFPs). 
+  \item
+    We found significant evidence that many published group polarization findings
+    are indeed plausibly false since we found distribution parameters that
+    generated simulated pre- and post-discussion data that a $t$-test found
+    to be from distributions with different means, although there was no difference
+    in mean, only a difference in variance. 
+  \item
+    Although a $t$-test may be fooled by simple consensus under ordinal measurements,
+    we show that a more appropriate ordered probit model of the data yield no
+    such plausibly-false positives.
+  \item
+    Our work shows the need for using appropriate statistical tests, as well as
+    the need for a consistent experimental paradigm so that meaningful meta-analyses
+    can be performed for theoretical integration and rigorous understanding of
+    the group polarization effect, should it prove to occur in future group
+    polarization studies that use proper statistics. In addition to avoiding
+    $t$-tests, future group polarization studies can be more reliable by
+    accounting for random effects arising from, e.g., socioeconomic status,
+    gender, and other individual-level factors that may bias how much and in what
+    direction a participant changes their opinions. Failure to do so can result 
+    in artificially inflated effect sizes~\cite{Yarkoni2021}.
+\end{itemize}
+
 \subsection{Group polarization}
+
+Group polarization is often induced in behavioral experiments by first answering
+an opinion prompt, grouping
+relatively like-minded individuals together to discuss the prompt, then 
+again polling opinions; the change in mean opinion towards a more extreme
+consensus than the initial mean is the amount of group polarization that has occurred
+(Figure~\ref{fig:groupPolarizationSchematic}).
+
+While all ten studies scrutinized here use the general experimental procedure
+outlined above (Figure~\ref{fig:groupPolarizationSchematic}), each study and constituent 
+experimental condition have idosyncracies that makes the results themselves
+more opaque, and that makes comparison of different studies' findings more
+difficult. (REVIEW A FEW SIMPLE EXAMPLES). A detailed examination of each study's
+idiosyncracies is given in Table~\ref{tab:conditionsComparison}.
+
+\begin{figure}
+  \caption{Group polarization is experimentally induced by first gather
+  participant responses to an opinion prompt and grouping together like-minded
+  participants (A). Group polarization is theorized to occur when individuals
+  are put together with like-minded individuals discuss the prompt (B). Group 
+  polarization is detected by again polling all participants' opinions (C) and comparing
+  these to the initial opinions.}
+  \label{fig:groupPolarizationSchematic}
+  \centering
+  \includegraphics[width=\textwidth]{Figures/GPExperimentSketch.pdf}
+\end{figure}
+
+\spacing{1.0}
+\input{tables/conditions_table.tex}
+\spacing{1.5}
 

### 2022-11-24 GroupPolarizationStatmod/paper/ms.tex
some writing up model; figures.R script for making explanatory stat/model figs

@@
 
-Our analysis relies on two important formal concepts: latent opinions and their 
-measurement on an ordinal scale. \emph{Latent opinions} are theoretical,
-unobservable, psychological, continuous-valued variables that are 
+Our model represents continuous latent opinions and their 
+measurement on an ordinal scale. \emph{Latent opinions} are 
+unobservable, psychological variables that are 
 somehow represented and stored in the brain-body, and can be reproduced with
-some accuracy when a person is prompted. \emph{Ordinal measurements} of 
-of latent opinions are ordinal-valued representations of latent opinions 
+some accuracy when a person is prompted to give their opinion. 
+\emph{Ordinal measurements} of 
+of latent opinions are discrete-valued representations of latent opinions 
 in some physical form, such as a mark on a sheet of paper or radio button on a web form, that
 indicate the polarity and strength of a person's opinion opinion in terms of
-one of a number of opinion bin values.
+one of a number of opinion bin values. All group polarization experiments must somehow
+measure opinions, and all but a few use ordinal measurements (those few exceptions include
+using bets made during blackjack as a proxy for opinions~\cite{Blascovich1973,Blascovich1975,Blascovich1976}
+and using mock jury punitive rewards as a proxy for opinions~\cite{Schkade2000}).
 
-For example, on a 7-point ordinal scale -3 could indcate 
-``strongly disagree'', +3 could be ``strongly agree'', 0 would be psychologically 
-neutral, and intermediate values are less extreme opinions of either valence, 
-such as was used by \citeA{Moscovici1969}. Or, one might
+% For example, on a 7-point ordinal scale -3 could indcate 
+% ``strongly disagree'', +3 could be ``strongly agree'', 0 would be psychologically 
+% neutral, and intermediate values are less extreme opinions of either valence, 
+% such as was used by \citeA{Moscovici1969}. Or, one might
 
-We assume that participant $i$'s internal, latent psychological opinion at 
-time of reporting ($t \in \{pre, post\}$) is drawn from a normal distribution
-with mean $\mu_t$ and standard deviation $\sigma_t$,
+We assume that participant $i$ has latent opinion at time $t$ that is drawn from
+a group-level normal distribution,
 \begin{equation}
@@
 \noindent 
-This is the condition-level distribution of opinions at time, $t$, chosen to
-represent the common experimental design used by group polarization studies,
+This distribution has mean $\mu_t$ and standard deviation $\sigma_t$.
+This form is chosen to
+reflect the common statistical analysis used by group polarization studies,
 not because it is statistically appropriate---in fact it is not, which is what
@@
 
+It is appropriate to assume opinions are normally distributed because published
+group polarization studies use some sort of $t$-tests, or similar tests, which are designed
+to detect whether the mean of two distributions are the same or not. This imposes the
+implicit assumption that opinions themselves are normal. But as we will show, 
+metric statistical tests such as $t$-tests reliably generate false rejections
+of the null hypothesis that there is no difference between two distributions' means.
+
+(HOW SIMPLE CONSENSUS LOOKS LIKE GROUP POLARIZATION)
+
 \subsection{Evaluating whether a group polarization finding is plausibly false}
+
+To evaluate whether a positive group polarization observation is plausibly false,
+we develop two layers of evidence for plausible falsehood. The first layer
+tests whether we can identify $\sigmapre$ and
 

### 2022-11-24 GroupPolarizationStatmod/figures.R

@@ -1,0 +1,39 @@
+library(ggplot2)
+
+plot_latent_pdf_integration = function() {
+
+  mytheme = theme(
+    panel.border = element_blank(), axis.line = element_line(), 
+    text = element_text(size=16, family = 'PT Sans'), 
+    panel.background = element_rect(fill = "white")
+  )
+
+  p = ggplot(data.frame(x = c(-3, 3)), aes(x)) + 
+
+    stat_function(fun = dnorm,
+                  args = c(0, 2.5),
+                  geom = "line",
+                  xlim = c(-5, 5)) +
+    stat_function(fun = dnorm,
+                  geom = "area",
+                  args = c(0, 2.5),
+                  fill = "steelblue",
+                  xlim = c(0.5, 1.5)) +
+    stat_function(fun = dnorm,
+                  geom = "area",
+                  args = c(0, 2.5),
+                  fill = "steelblue",
+                  xlim = c(2.5, 5)) +
+
+    xlab("Latent opnion") + ylab("Probability density") +
+
+    mytheme
+  
+  return (p)
+}
+
+
+plot_ordinal_distribution = function() {
+  # Use same parameters as latent distribution to calculate integration over
+  # bin threshold limits \theta_{k-1}..\theta_k
+}

### 2022-11-27 GroupPolarizationStatmod/paper/ms.tex
added GP from simple consensus fig, model text

@@
 \usepackage{marvosym}
+\usepackage{float}
 \usepackage{setspace}
 \usepackage{longtable}
+\usepackage{caption}
 \usepackage{pdflscape}
@@
 \linenumbers
-\modulolinenumbers[5]
+\modulolinenumbers[3]
 
@@
     direction a participant changes their opinions. Failure to do so can result 
-    in artificially inflated effect sizes~\cite{Yarkoni2021}.
+    in artificially-inflated effect sizes~\cite{Yarkoni2021}.
 \end{itemize}
@@
 \begin{equation}
-  o_{i,t} \sim \mathcal{N}(\mu_t, \sigma_t).
+  o_{i,t} \sim p(o; \mu_t, \sigma_t) = \mathcal{N}(\mu_t, \sigma_t).
   \label{eq:opinionDistribution}
@@
 
-(HOW SIMPLE CONSENSUS LOOKS LIKE GROUP POLARIZATION)
+\subsubsection{How simple consensus can look like group polarization}
+
+\begin{table}[h]
+  \begin{tabular}{cp{5.0in}} \toprule
+   Symbol & Description  \\ \midrule  
+   $t \in \{\mathrm{pre}, \mathrm{post}\}$ & Time, either \emph{pre}- or
+                                           \emph{post}-discussion \\
+   $k \in \{0,\ldots,K\} $ & Bin index ($k$) between 0 and total number of
+         ordinal bins, $K$ \\
+   $\mu$  &   Latent mean, assumed constant in simple consensus \\ 
+   $p(o;\mu,\sigma_t)$ & Normal latent distribution of opinions with mean
+                         $\mu$ and variance $\sigma^2$ \\
+   $p(o \in b_k; \mu, \sigma_t)$ & Ordinal density of binned opinions in bin $b_k$ 
+            (Equation~\ref{eq:binFrequency}) \\
+   $\theta_k$  & The upper threshold value for binning latent opinions into
+   bin $b_k$ \\
+   $\langle o_t \rangle$  &   Observed mean opinion at time $t$ \\
+   \bottomrule
+  \end{tabular} 
+\end{table}
+
+\begin{figure}
+  \caption{\emph{Simple consensus} occurs in a group when its latent opinion
+  distribution variance decreases, but the mean does not change (A). When 
+latent opinions are measured on an ordinal scale, the mean of the distribution
+can be distorted by ceiling (or floor) effects (B). Group polarization 
+can appear to have occurred where the mean of the observed opinions appears
+to increase despite the latent mean being constant (C).}
+  \label{fig:consensusDistroIllustration}
+  \vspace{1em}
+  \centering
+    \includegraphics[width=1.0\textwidth]{ConsensusDistroIllustration/figure.pdf}
+\end{figure}
 
@@
 \begin{aligned}
-  p(o = b_k;\mu_t, \sigma_t, \theta) 
+  p(o \in b_k; \mu_t, \sigma_t, \theta) 
     &= \int_{\theta_{k-1}}^{\theta_k} p(o;\mu_t, \sigma_t) do \\
@@
 % \input{plausibleFP-table.tex}
-  \begin{table}
+\begin{table}[H]
     \begin{center}
@@
 \hspace{-1em}
-\begin{center}
 \input{t-test-table.tex}
-\end{center}
 % \input{t-test-table-custom.tex}

### 2022-11-27 GroupPolarizationStatmod/figures.R

@@ -1,6 +1,8 @@
 library(ggplot2)
 
-plot_latent_pdf_integration = function() {
+plot_latent_pdf_integration = function(mu = 0, sd = 2.5, xlim = c(-6.0, 6.0), 
+                                       bin1 = c(-0.5, 0.5), bin2 = c(1.5, 5.0),
+                                       save_path = "test_latent_pdf.pdf") {
 
   mytheme = theme(
     panel.border = element_blank(), axis.line = element_line(), 
@@ -8,32 +10,117 @@ plot_latent_pdf_integration = function() {
     panel.background = element_rect(fill = "white")
   )
 
-  p = ggplot(data.frame(x = c(-3, 3)), aes(x)) + 
+  p = ggplot(data.frame(x = xlim), aes(x)) + 
 
     stat_function(fun = dnorm,
-                  args = c(0, 2.5),
+                  args = c(mu, sd),
                   geom = "line",
-                  xlim = c(-5, 5)) +
+                  xlim = xlim + 1) +
+    
     stat_function(fun = dnorm,
                   geom = "area",
-                  args = c(0, 2.5),
-                  fill = "steelblue",
-                  xlim = c(0.5, 1.5)) +
+                  args = c(mu, sd),
+                  fill = "lightpink",
+                  xlim = c(-5.0, -1.5)) +
+    
     stat_function(fun = dnorm,
                   geom = "area",
-                  args = c(0, 2.5),
-                  fill = "steelblue",
-                  xlim = c(2.5, 5)) +
+                  args = c(mu, sd),
+                  fill = "lightblue",
+                  xlim = c(-1.5, -0.5)) +
 
-    xlab("Latent opnion") + ylab("Probability density") +
+    stat_function(fun = dnorm,
+                  geom = "area",
+                  args = c(mu, sd),
+                  fill = "#ffcb58",
+                  xlim = c(-0.5, 0.5)) +
+
+    stat_function(fun = dnorm,
+                  geom = "area",
+                  args = c(mu, sd),
+                  fill = "lightgreen",
+                  xlim = c(0.5, 1.5)) +
+    
+    stat_function(fun = dnorm,
+                  geom = "area",
+                  args = c(mu, sd),
+                  fill = "#F6CFFF",
+                  xlim = c(1.5, 5)) +
+
[…]

### 2022-11-27 GroupPolarizationStatmod/model.R

@@ -3,7 +3,7 @@
 ##
 
 # Probability that a participant reports opinion in bin k, i.e., between
-# theta_{k-1} and theta_k.
+# theta_{k-1} and theta_k. 
 binProb <- function(thisBin, bin1, K, latentMean, latentSD)
 {
     # Thresholds in theta vector start at bin1 + 0.5 (see Eq. 6 in L&K supplement).
@@ -11,6 +11,7 @@ binProb <- function(thisBin, bin1, K, latentMean, latentSD)
 
     # Get this by solving thisbin = (bin1 - 1) + binIdx, where binIdx starts from 1.
     binIdx <- thisBin - bin1 + 1
+    
     # But it only works if k is not the first...
     if (thisBin == bin1)
     {
@@ -21,7 +22,10 @@ binProb <- function(thisBin, bin1, K, latentMean, latentSD)
         theta_km1 <- interiorThresholds[binIdx - 1]
     }
     # ... or the last values.
-    if (thisBin == bin1 + K - 1)
+    # print(thisBin)
+    # print(bin1)
+    # print(K)
+    if (thisBin == (bin1 + K - 1))
     {
         theta_k = Inf
     }
@@ -29,6 +33,8 @@ binProb <- function(thisBin, bin1, K, latentMean, latentSD)
     {
         theta_k <- interiorThresholds[binIdx]
     }
+    # print(theta_k)
+    # print(theta_km1)
     # From Equation 2 in Liddell & Kruschke (2018), p. 340.
     return (pnorm((theta_k - latentMean) / latentSD) - 
             pnorm((theta_km1 - latentMean) / latentSD))

### 2022-11-30 GroupPolarizationStatmod/paper/ms.tex
paper updates to model section especially

@@
 % HEADINGS
-\usepackage{sectsty} 
-\usepackage[normalem]{ulem} 
-\sectionfont{\upshape\Large}
-\subsectionfont{\scshape\normalsize} 
-\subsubsectionfont{\upshape\large}
+% \usepackage{sectsty} 
+% \usepackage[normalem]{ulem} 
+% \sectionfont{\upshape\Large}
+% \subsectionfont{\scshape\normalsize} 
+% \subsubsectionfont{\upshape\large}
 
@@
 \newcommand{\sigmapost}{\sigma_\mathrm{post}}
+
 \newcommand{\normalpre}{\mathcal{N}(\mu, \sigmapre)}
 \newcommand{\normalpost}{\mathcal{N}(\mu, \sigmapost)}
+
+\newcommand{\ordmeanpre}{\langle o_{\mathrm{pre}}}
+\newcommand{\ordmeanpost}{\langle o_{\mathrm{post}}}
 
@@
 
+Simple consensus can look like group polarization when (1) the mean latent opinion
+does not change; (2) both more and less extreme opinions move toward this 
+unchanged latent mean (i.e., opinion variance shrinks due to consensus); 
+and (3) an ordinal scale is used that bins all extreme
+latent opinions into the same extreme bin even though those means have changed,
+but opinion bins that are less extreme or of opposite polarity more accurately
+reflect the underlying change in latent opinion distribution. This process is
+illustrated in Figure~\ref{fig:consensusDistroIllustration}, where we show
+a toy example of what can occur when the pre-consensus
+opinions (upper left plot)
+
 \begin{table}[h]
@@
                                            \emph{post}-discussion \\
-   $k \in \{0,\ldots,K\} $ & Bin index ($k$) between 0 and total number of
+   $\mu$  &   Latent mean, assumed constant for simple consensus \\ 
+   $\sigma_t$ & Latent standard deviation at time $t$ \\
+   $p(o;\mu,\sigma_t)$ & Normal latent distribution of opinions with mean
+                         $\mu$ and variance $\sigma_t^2$ \\
+   $k \in \{0,\ldots,K\} $ & Threshold index ($k$) between 0 and total number of
          ordinal bins, $K$ \\
-   $\mu$  &   Latent mean, assumed constant in simple consensus \\ 
-   $p(o;\mu,\sigma_t)$ & Normal latent distribution of opinions with mean
-                         $\mu$ and variance $\sigma^2$ \\
+   $b_k$ & Opinion bin $k$; note there is no $k=0$ bin. \\
+   $\theta_{k-1},~\theta_k$  & Lower, upper threshold value for binning latent opinions into
+     bin $b_k$ \\
    $p(o \in b_k; \mu, \sigma_t)$ & Ordinal density of binned opinions in bin $b_k$ 
             (Equation~\ref{eq:binFrequency}) \\
-   $\theta_k$  & The upper threshold value for binning latent opinions into
-   bin $b_k$ \\
-   $\langle o_t \rangle$  &   Observed mean opinion at time $t$ \\
+   $\langle o_t \rangle$  &   Observed mean opinion calculated from ordinal data at time $t$ \\
    \bottomrule
@@
 
-\subsection{Evaluating whether a group polarization finding is plausibly false}
-
-To evaluate whether a positive group polarization observation is plausibly false,
-we develop two layers of evidence for plausible falsehood. The first layer
-tests whether we can identify $\sigmapre$ and
-
-We used large-$N$ and generative modeling supported by $t$-tests on simulated
-data to demonstrate that many published findings are plausibly false.
-
-We model measurement of $N \to \infty$ participant opinions,
+This distributional view models the measurement of $N \to \infty$ participant opinions,
 which results in a histogram of frequency of opinions
@@
 standard deviation $\sigma$.
+
+\subsection{Evaluating whether a group polarization finding is plausibly false}
+
+We have shown in one toy example that group polarization can appear to occur
+when ordinal measures are used, without accounting for the binning process; now
+we show how to .
+
+We develop two measures for how plausible it is that a reported positive detection
+of group polarization is indeed a false positive, building our confidence 
+in the plausibility at each step. In this way, we take a modern Bayesian approach
+to scientific evidence that typically does not dichotomize ``true'' and ``false''
+into absolutes, but evaluates degrees of belief instead.
+
+The first measure we use for whether a reported detection of group polarization
+is plausibly false is whether we can find a combination of latent simple
+consensus parameters $\mu$, $\sigmapre$, and $\sigmapost$ representing latent
+distributions that, when binned (Equation~\ref{eq:binFrequency} and
+Figure~\ref{fig:consensusDistroIllustration}), result in ordinal distributions
+that have the same difference in means reported in the literature that the authors
+took to mean group polarization had occurred. Note that these calculations are
+done for the case where effectively we have $N \to \infty$; next we turn to
+the question of whether data generated from these simple consensus distributions
+generate a ``significant'' group polarization result.
+
+The second measure of how plausible it is that a group polarization is false
+is whether ordinal data generated by binning draws from the latent distributions
+is found to be significantly different according to Welch's $t$-test.
+
 \subsubsection{Identify null-model parameters that yield plausibly-false positives}
+
+To find simple consensus parameters $\mu$, $\sigmapre$, and $\sigmapost$ that
+generate ordinal distributions showing group polarization, we used the above
+framework in an optimization routine that takes an initial guess for these latent
+parameters and hillclimbs to find $\sigmapre$ and $\sigmapost$ that most accurately
+generate ordinal data that appears to have had group polarization occur.
 
@@
 in the $N \to \infty$ case, we generated binned samples from normal opinion 
-distributions $\normalpre$ and $\normalpost$.
+distributions $\normalpre$ and $\normalpost$. Specifically, for each experimental
+condition that passed the initial $N \to \infty$ test, we generated 1000 
+pre- and post-deliberation ordinal datasets.
+
+We generated these datasets by...
 
 \subsection{Bayesian ordered-probit models reliably identify simple consensus}
+
+\mt{Is this necessary or can we just reference Liddell \& Kruschke 2018 who show
+that ordered-probit models accurately identify latent parameters when normal-distribution
+models do not?}
 

### 2022-12-01 GroupPolarizationStatmod/paper/ms.tex
updated paper model section

@@
 
-We assume that participant $i$ has latent opinion at time $t$ that is drawn from
+We begin by developing a formal model of ordinal measurement of latent continuous
+opinions (model variables and parameters are given in Table~\ref{tab:modelVariables}).
+First, we assume that participant $i$ has latent opinion at time $t$ that is drawn from
 a group-level normal distribution,
@@
 \begin{table}[h]
+  \caption{Variables used in the formal model of group polarization from simple consensus.}
+  \label{tab:modelVariables}
   \begin{tabular}{cp{5.0in}} \toprule

### 2022-12-07 GroupPolarizationStatmod/paper/ms.tex
updated abstract and started updating figures to add sigma-finding diagram

@@
   \noindent
-  Group polarization is the observed tendency of
-  like-minded groups to arrive at a more extreme consensus opinion than
-  the average opinion after discussing some topic of interest.
-  Clearly it is important to understand such a phenomenon to understand broader
-  problematic trends towards increased polariztion across society as a whole.
-  Unfortunately, recent analyses have demonstrated that this measurement and
-  analysis approach can result in false positive detections group polarization
-  from failing to account for floor/ceiling effects induced when using ordinal scales to
-  measure continuously-distributed latent psychological variables. 
-  Here we demonstrate that many high profile detections
-  of group polarization are plausibly false
-  due to these floor/ceiling effects using generative modeling seeded by 
-  reported group polarization values and parameters. After we identify which
-  reported detections of group polarization are plausibly false detections, we further
-  investigate whether the latent constant mean opinion and pre- and post-discussion
-  opinion variances can fool the statistical tests used in each study to 
-  generate false positives. Finally, we evaluate whether a more appropriate
-  Bayesian cumulative normal statistical model can more accurately distinguish
-  group polarization from simple consensus. Group polarization continues to 
-  pose a problem for psychological theory with an ever-expanding set of potential theoretical
-  explanations. A search for explanations is pointless, however, without sound
-  evidence. We believe this work will provide a sturdier analytical foundation for 
-  future group-polarization research.
+  Psychology is beset by three crises: the reproducibility crisis, the
+  generalizability crisis, and the theory crisis. New verbal theories for various
+  phenomena are constantly being developed orthogonal to other theories, all based on
+  mostly unreproducible analyses; even if a positive result is in some sense ``true''
+  it may not account for all the variance present in a set of observations---studies
+  thus fail to generalize because their results do not account for this variance.
+  There is another related problem: a phenomenon may be ``reproducible'' only
+  because studies of the phenomenon systematically misuse statistics in such a way
+  that induces false positives for each succeeding ``replication.'' Here we show that
+  replicated detections of group polarization are often plausibly false due to
+  treating ordinal measurements of opinions, e.g., on a Likert-type scale, as drawn
+  directly from an underlying continuous latent opinion distribution, which masks
+  extreme latent opinions. Our intensive review of sixty experimental conditions
+  across ten foundational group polarization papers revealed that 95\% of the
+  conditions generated plausibly false positive results. To show this, we developed a
+  novel generative model that provides two levels of evidence that an historical
+  published empirical finding of group polarization is plausibly false.  Our model
+  can be used to evaluate additional studies in group polarization and beyond, even
+  if the reported data and significant metrics are incompletely reported, and the
+  original raw data has been lost to time.  This may be just the tip of the
+  plausibly-false-positives-in-psychology iceberg, both for group polarization
+  studies and for the legion psychological results that model ordinal observations of
+  opinions, beliefs, emotions, or any other latent psychological variable that is
+  theoretically continuous, but measured using an ordinal scale.  
 \end{abstract}
@@
 \begin{figure}
-  \caption{Group polarization is experimentally induced by first gather
-  participant responses to an opinion prompt and grouping together like-minded
-  participants (A). Group polarization is theorized to occur when individuals
-  are put together with like-minded individuals discuss the prompt (B). Group 
-  polarization is detected by again polling all participants' opinions (C) and comparing
-  these to the initial opinions.}
+  \caption{Group polarization is experimentally induced by first gathering
+  participant responses to an opinion prompt (A). Then groups of like-minded 
+  individuals discuss the prompt (B). Group 
+  polarization occurs if the mean post-discussion opinion is more extreme 
+  than the mean pre-discussion opinion (C).}
   \label{fig:groupPolarizationSchematic}
@@
 
-\section{Model and Data}
+\section{Model}
 
@@
 in group polarization experiments, and the process of measuring that consensus using
-an ordinal measurement scale, e.g.\ a Likert-type scale. This model considers two
+an ordinal measurement scale, e.g.\ a Likert-type scale. This model operationalizes two
 relevant forms of consensus: \emph{simple consensus} and \emph{group polarization}.
-Either process is \emph{consensus}, if we define consensus as a reduction
-in opinion variance over time.  
-Group polarization experiments are designed to result group consensus
-since they group together like-minded individuals to see if they become more
-extreme after discussion~\cite{Turner2018}.
+\mt{Need to explain in Intro why consensus should occur along with group
+polarization} 
+% Group polarization experiments are designed to result group consensus
+% since they group together like-minded individuals to see if they become more
+% extreme after discussion~\cite{Turner2018}.
 
@@
 group polarization}
-
-Our model represents continuous latent opinions and their 
-measurement on an ordinal scale. \emph{Latent opinions} are 
-unobservable, psychological variables that are 
-somehow represented and stored in the brain-body, and can be reproduced with
-some accuracy when a person is prompted to give their opinion. 
-\emph{Ordinal measurements} of 
-of latent opinions are discrete-valued representations of latent opinions 
-in some physical form, such as a mark on a sheet of paper or radio button on a web form, that
-indicate the polarity and strength of a person's opinion opinion in terms of
-one of a number of opinion bin values. All group polarization experiments must somehow
-measure opinions, and all but a few use ordinal measurements (those few exceptions include
-using bets made during blackjack as a proxy for opinions~\cite{Blascovich1973,Blascovich1975,Blascovich1976}
-and using mock jury punitive rewards as a proxy for opinions~\cite{Schkade2000}).
-
-% For example, on a 7-point ordinal scale -3 could indcate 
-% ``strongly disagree'', +3 could be ``strongly agree'', 0 would be psychologically 
-% neutral, and intermediate values are less extreme opinions of either valence, 
-% such as was used by \citeA{Moscovici1969}. Or, one might
 

### 2022-12-07 GroupPolarizationStatmod/figures.R

@@ -63,7 +63,7 @@ plot_latent_pdf_integration = function(mu = 0, sd = 2.5, xlim = c(-6.0, 6.0),
 
 plot_ordinal_distribution = function(mu = 0, sd = 2.5, min_bin = -2, 
                                      max_bin = 2, bins = -2:2,
-                                     these_colors = c("lightpink", "lightblue", 
+                                     bin_colors = c("lightpink", "lightblue", 
                                                       "#f7ae3d", "lightgreen", 
                                                       "#F6CFFF"),
                                      save_path = "test_ordinal_pdf.pdf") {
@@ -90,7 +90,7 @@ plot_ordinal_distribution = function(mu = 0, sd = 2.5, min_bin = -2,
   print(paste("Mean observed for sd = ", sd, ": ", mean_observed, sep = ""))
   
   p <- ggplot(data=ord_dist_df, aes(x=OpinionBin, y=Density)) +
-         geom_bar(stat = "identity", fill = these_colors) + 
+         geom_bar(stat = "identity", fill = bin_colors) + 
          geom_vline(xintercept = mean_observed, linetype="dashed") +
          xlab("Binned ordinal opinion") + ylab("Probability density\n\n") +
          mytheme

### 2023-10-18 GroupPolarizationStatmod/paper/ms.tex
update paper abstract

@@
 % \shorttitle{Group polarization or simple consensus?}
-\author[1,2]{{Matthew A.~Turner}}
-\affil[1]{Department of Earth System Science, Stanford University}
-\affil[2]{Division of Social Sciences, Stanford Doerr School of Sustainability, Stanford University}
+% \author[1,2]{{Matthew A.~Turner}}
+\author[1]{{Matthew A.~Turner}}
+% \affil[1]{Department of Earth System Science, Stanford University}
+\affil[1]{Division of Social Sciences, Stanford Doerr School of Sustainability, Stanford University}
 
-\author[3,4,5]{{Paul E.~Smaldino}}
-\affil[3]{Cognitive and Information Sciences, University of California, Merced}
-\affil[4]{Center for Advanced Study in the Behavioral Sciences, Stanford University}
-\affil[5]{Santa Fe Institute}
+% \author[3,4,5]{{Paul E.~Smaldino}}
+\author[2,3]{{Paul E.~Smaldino}}
+\affil[2]{Cognitive and Information Sciences, University of California, Merced}
+% \affil[4]{Center for Advanced Study in the Behavioral Sciences, Stanford University}
+\affil[3]{Santa Fe Institute}
 
@@
 \linenumbers
-\modulolinenumbers[3]
+% \modulolinenumbers[3]
 
@@
   \noindent
-  Psychology is beset by three crises: the reproducibility crisis, the
-  generalizability crisis, and the theory crisis. New verbal theories for various
-  phenomena are constantly being developed orthogonal to other theories, all based on
-  mostly unreproducible analyses; even if a positive result is in some sense ``true''
-  it may not account for all the variance present in a set of observations---studies
-  thus fail to generalize because their results do not account for this variance.
-  There is another related problem: a phenomenon may be ``reproducible'' only
-  because studies of the phenomenon systematically misuse statistics in such a way
-  that induces false positives for each succeeding ``replication.'' Here we show that
-  replicated detections of group polarization are often plausibly false due to
-  treating ordinal measurements of opinions, e.g., on a Likert-type scale, as drawn
-  directly from an underlying continuous latent opinion distribution, which masks
-  extreme latent opinions. Our intensive review of sixty experimental conditions
-  across ten foundational group polarization papers revealed that 95\% of the
-  conditions generated plausibly false positive results. To show this, we developed a
-  novel generative model that provides two levels of evidence that an historical
-  published empirical finding of group polarization is plausibly false.  Our model
-  can be used to evaluate additional studies in group polarization and beyond, even
-  if the reported data and significant metrics are incompletely reported, and the
-  original raw data has been lost to time.  This may be just the tip of the
-  plausibly-false-positives-in-psychology iceberg, both for group polarization
-  studies and for the legion psychological results that model ordinal observations of
-  opinions, beliefs, emotions, or any other latent psychological variable that is
-  theoretically continuous, but measured using an ordinal scale.  
+  % Psychology is beset by three crises: the reproducibility crisis, the
+  % generalizability crisis, and the theory crisis. New verbal theories for various
+  % phenomena are constantly being developed orthogonal to other theories, all based on
+  % mostly unreproducible analyses; even if a positive result is in some sense ``true''
+  % it may not account for all the variance present in a set of observations---studies
+  % thus fail to generalize because their results do not account for this variance.
+  % There is another related problem: a phenomenon may be ``reproducible'' only
+  % because studies of the phenomenon systematically misuse statistics in such a way
+  % that induces false positives for each succeeding ``replication.'' 
+  \emph{Group polarization} refers to the observation that like-minded novel groups tend to
+  become more extreme in their opinions after discussion as group members found consensus.
+  \emph{Consensus} here refers to decreasing opinion variance; we further define
+  \emph{simple consensus} as a process where group opinion variance decreases but
+  the mean opinion does not change.
+  Here we show that many published experimental detections of group polarization are explained
+  equally rigorously as \emph{simple consensus}. 
+  The reports we analyze assumed in their statistical analyses 
+  that both latent psychological opinions \emph{and} the ordinal 
+  reported opinions (e.g., on a Likert-type scale) are continuous. We use
+  a generative model to demonstrate that 95\% of sixty experimental conditions
+  across ten foundational group polarization papers report plausibly false detections of
+  group polarization. We show that simulated data from these plausibly false
+  cases often fools a $t$-test into reporting significant group polarizaiton
+  effects, but a Bayesian implementation of the more appropriate ordered-probit
+  model reliably identifies these data as
+  representing simple consensus, not group polarization.
+  Our results demonstrate the need for a new wave of group polarization research
+  that uses rigorous statistical methods appropriate for the study design. 
+  Our approach can be used beyond group polarization
+  research to analyze results from any experimental paradigm that may use statistical models that
+  are poorly matched to the study design. Replications are worthless if the statistics are wrong.
+
+  % To show this, we developed a
+  % novel generative model that provides two levels of evidence that an historical
+  % published empirical finding of group polarization is plausibly false.  Our model
+  % can be used to evaluate additional studies in group polarization and beyond, even
+  % if the reported data and significant metrics are incompletely reported, and the
+  % original raw data has been lost to time.  This may be just the tip of the
+  % plausibly-false-positives-in-psychology iceberg, both for group polarization
+  % studies and for the legion psychological results that model ordinal observations of
+  % opinions, beliefs, emotions, or any other latent psychological variable that is
+  % theoretically continuous, but measured using an ordinal scale.  
 \end{abstract}

### 2023-10-18 GroupPolarizationStatmod/analysis.R

@@ -155,7 +155,7 @@ latexifyTTestExperiment <- function(tTestTablePath = "data/output/TtestSummaryTa
     
     tTestTable <- tTestTable[, col_order]
     
-    tTestTable <- tTestTable %>% rename(FPRate = FractionSignificant, 
+    tTestTable <- tTestTable %>% rename(FPRate = FractionSignificant,
                                         Est.FPRate = ExpectedPower)
     
     ttestTable_latex <- 

### 2023-12-15 GroupPolarizationStatmod/paper/ms.tex
calculating cohen's d for, parallelized metric model

@@
 
-\subsubsection{$t$-tests of generated simple-consensus data}
+\subsubsection{Metric model inference of Cohen's $d$ for simulated simple-consensus data}
 
@@
 
-We generated these datasets by...
+We generated these simulated data using the latent mean, pre-deliberation 
+standard deviation, and post-deliberation standard deviations as parameters of
+normal distributions of pre- and post-deliberation opinions.
 
-\subsection{Bayesian ordered-probit models reliably identify simple consensus}
+\subsection{Bayesian ordered-probit models inference of Cohen's $d$ for simulated
+simple-consensus data}
 
-\mt{Is this necessary or can we just reference Liddell \& Kruschke 2018 who show
-that ordered-probit models accurately identify latent parameters when normal-distribution
-models do not?}
+Non-metric ordered probit models, which account for the data collection 
+process, theoretically should be better able to correctly
+distinguish simple consensus compared to metric models. Therefore, we also fit
+the following ordered probit model to simulated group polarization data.
 
@@
 
-\subsection{Do $t$-tests accurately detect simple consensus?}
+\subsection{Metric linear models fail to identify simple consensus}
 
@@
 
-\subsection{Do Bayesian ordered-probit models accurately detect simple consensus?}
+\subsection{Ordered-probit models also fail?}
 
@@
 positives, not anything intrinsic to the phenomenon of group polarization.
+
+\subsection{More data enables ordered probit models, but not metric models, to reliably
+  identify simple consensus}
 

### 2023-12-15 GroupPolarizationStatmod/experiments.R

@@ -205,17 +205,23 @@ calculateBayesian <- function(N, firstBinValue, nBins, latentMean, latentSd)
 }
 
 
-tTestExperiment <- function(N, firstBinValue, nBins, latentMean, latentSdPre, 
+# Uses R's t.test function to fit a linear model to estimate pre- and post-
+# deliberation means in a simple consensus context where there is only one
+# latent mean.
+simulate_metric_cohens_d <- function(N, firstBinValue, nBins, latentMean, latentSdPre, 
                             latentSdPost, paired = FALSE, var.equal = FALSE)
 {
-  simObsPre <- simulatedObservation(N, firstBinValue, nBins, latentMean, latentSdPre)
+  sim_obs_pre <- simulatedObservation(N, firstBinValue, nBins, latentMean, latentSdPre)
 
-  simObsPost <- simulatedObservation(N, firstBinValue, nBins, latentMean, latentSdPost)
+  sim_obs_post <- simulatedObservation(N, firstBinValue, nBins, latentMean, latentSdPost)
   
-  t_result <- t.test(simObsPre, simObsPost, paired = paired, var.equal = var.equal)
+  t_result <- t.test(sim_obs_pre, sim_obs_post, paired = paired, var.equal = var.equal)
   
   pre_mean_estimate <- t_result$estimate[[1]] 
   post_mean_estimate <- t_result$estimate[[2]] 
+  
+  pre_sd <- sd(sim_obs_pre)
+  post_sd <- sd(sim_obs_post)
   
   return (cohens_d(pre_mean_estimate, post_mean_estimate, pre_sd, post_sd))
 }
@@ -223,7 +229,7 @@ tTestExperiment <- function(N, firstBinValue, nBins, latentMean, latentSdPre,
 
 cohens_d <- function(pre_mean_estimate, post_mean_estimate, pre_sd, post_sd) {
   
-  numerator <- pre_mean_estimate - post_mean_estimate
+  numerator <- post_mean_estimate - pre_mean_estimate
   
   denominator = sqrt((pre_sd**2 + post_sd**2) / 2.0)
   
@@ -429,68 +435,6 @@ latentMeanRanges <- function(caseStudyDataFile = "caseStudies.csv",
     # Save processed dataframe to disk.
     write.csv(csData, "caseStudiesProcessed.csv")
 }
-
-
-##
-# Arguments:
-#     case (data.frame row): Single experimental condition case from case
-#                            studies.
-#
[…]

### 2024-01-18 GroupPolarizationStatmod/paper/ms.tex
updated abstract

--- 2023-10-18 GroupPolarizationStatmod/paper/ms.tex
@@
   % that induces false positives for each succeeding ``replication.'' 
-  \emph{Group polarization} refers to the observation that like-minded novel groups tend to
-  become more extreme in their opinions after discussion as group members found consensus.
-  \emph{Consensus} here refers to decreasing opinion variance; we further define
-  \emph{simple consensus} as a process where group opinion variance decreases but
-  the mean opinion does not change.
-  Here we show that many published experimental detections of group polarization are explained
-  equally rigorously as \emph{simple consensus}. 
-  The reports we analyze assumed in their statistical analyses 
-  that both latent psychological opinions \emph{and} the ordinal 
-  reported opinions (e.g., on a Likert-type scale) are continuous. We use
-  a generative model to demonstrate that 95\% of sixty experimental conditions
-  across ten foundational group polarization papers report plausibly false detections of
-  group polarization. We show that simulated data from these plausibly false
-  cases often fools a $t$-test into reporting significant group polarizaiton
-  effects, but a Bayesian implementation of the more appropriate ordered-probit
-  model reliably identifies these data as
-  representing simple consensus, not group polarization.
-  Our results demonstrate the need for a new wave of group polarization research
-  that uses rigorous statistical methods appropriate for the study design. 
-  Our approach can be used beyond group polarization
-  research to analyze results from any experimental paradigm that may use statistical models that
-  are poorly matched to the study design. Replications are worthless if the statistics are wrong.
+  \emph{Group polarization} is a theoretical process where like-minded novel group members 
+  become more extreme in their opinions after group discussion, but also find \emph{consensus}, reducing 
+  group-level opinion variance.
+  \emph{Simple consensus} may occur instead, where opinion variance decreases but
+  the mean group opinion does not change.
+  Here we show that many published detections of group polarization are explained
+  equally rigorously as simple consensus.
+  This is the case because experimenters failed
+  to account for the measurement of continous-valued psychological opinions 
+  as ordinal-valued reported opinions, e.g., on a Likert-type scale. 
+  This creates ceiling effects that mask whether extreme opinions become less extreme.
+  We developed a generative model to demonstrate that 95\% of sixty experimental conditions
+  across ten group polarization papers report plausibly false detections of
+  group polarization. Our approach can be used to evaluate whether other detections of
+  group polarization may be plausibly false---our approach could be applied beyond
+  group polarization as well. This work also provides a
+  framework for new group polarization studies that could
+  share a common, rigorous, and interoperable study design to multiply our knowledge
+  of this potentially important group process, avoiding the historically fragmented
+  approach using disparate, incommensurate study designs.
 

### 2024-01-18 GroupPolarizationStatmod/paper/ms.tex
starting analyses writing and cleaning up model setup

--- 2023-12-15 GroupPolarizationStatmod/paper/ms.tex
@@
 \begin{figure}
-  \caption{Group polarization is experimentally induced by first gathering
-  participant responses to an opinion prompt (A). Then groups of like-minded 
-  individuals discuss the prompt (B). Group 
-  polarization occurs if the mean post-discussion opinion is more extreme 
-  than the mean pre-discussion opinion (C).}
+  \caption{Group polarization experiments trials in the ten studies
+    analyzed here proceed in three steps: 
+    (A) \textbf{Pre-deliberation.} Participant opinions are measured using an 
+    ordinal scale. 
+    (B) \textbf{Deliberation.} Groups discuss their opinions on the prompt. 
+    (C) \textbf{Post-deliberation.} Participant opinions are measured again.
+    Group polarization in one trial is said to occur if 
+    the mean post-deliberation opinion is greater than the pre-deliberation 
+    opinion.
+}
   \label{fig:groupPolarizationSchematic}
@@
 
-\spacing{1.0}
-\input{tables/conditions_table.tex}
-\spacing{1.5}
+% \spacing{1.0}
+% \input{tables/conditions_table.tex}
+% \spacing{1.5}
 
@@
 
-\subsection{Metric linear models fail to identify simple consensus}
+\subsection{Metric linear models often fail to identify simple consensus}
 
-\spacing{1.0}
-\hspace{-1em}
-\input{t-test-table.tex}
-% \input{t-test-table-custom.tex}
-\spacing{1.5}
+Here we evaluate if metric linear models inferred group polarization from
+simulated simple consensus data using parameters identified above. Those
+parameters specified continous pre- and post-deliberation
+
+% \spacing{1.0}
+% \hspace{-1em}
+% \input{t-test-table.tex}
+% % \input{t-test-table-custom.tex}
+% \spacing{1.5}
 
@@
   identify simple consensus}
+
+  \mt{Idea: increase $N$ for simulations and see effect on distribution of Cohen's $d$
+  for both metric and ordinal models}
 

### 2024-01-18 GroupPolarizationStatmod/analysis.R

@@ -44,7 +44,7 @@ makePlausibleFPTable <- function(studiesAnalysisDfPath = "data/StudiesAnalysis.c
 make_metric_cohens_d_table <- function(studiesAnalysisDfPath = "data/StudiesAnalysis.csv",
                                       outputPath = "data/output/cohens_sample_test.csv",
                                       diagnosticSavePath = "data/diagnostic/tTestFits.RDS",
-                                      ntrials = 2, limit = 2)
+                                      ntrials = 2, limit = 2, N_multiplier = 1)
 {
   # Load analysis of studies from web interface stored as CSV.
   studiesDf <- read.csv(studiesAnalysisDfPath)
@@ -82,7 +82,7 @@ make_metric_cohens_d_table <- function(studiesAnalysisDfPath = "data/StudiesAnal
   my_cluster <- makeCluster(detectCores() - 1, type = "FORK")
   registerDoParallel(my_cluster)
   
-result <- foreach (studiesRowIdx = 1:nrow(studiesDf)) %dopar% 
+  result <- foreach (studiesRowIdx = 1:nrow(studiesDf)) %dopar% 
   {
     # Get treatment row of interest.
     row <- studiesDf[studiesRowIdx, ]
@@ -95,8 +95,8 @@ result <- foreach (studiesRowIdx = 1:nrow(studiesDf)) %dopar%
                            TrialIndex = c(), LatentMean = c(), LatentSDPre = c(),
                            LatentSDPost = c(), Cohens_d = c(), ObservedShift = c(),
                            ExpectedPower = c())
-    
-    expectedPower = power.t.test(delta = row$ObservedShift, n = row$N, 
+    N = as.integer(row$N * N_multiplier)
+    expectedPower = power.t.test(delta = row$ObservedShift, n = N, 
                                  sd = (row$LatentSDPost + row$LatentSDPre) / 2.0,
                                  sig.level = 0.1)$power
     print("HERE")
@@ -104,9 +104,9 @@ result <- foreach (studiesRowIdx = 1:nrow(studiesDf)) %dopar%
     for (trial_idx in 1:ntrials)
     {
       # Run t-test ntrials experiments for treatment row and trial index.
-      this_cohens_d <- simulate_metric_cohens_d(row$N, row$MinBinValue, nBins, 
-                                                row$LatentMean, row$LatentSDPre, 
-                                                row$LatentSDPost)
[…]

### 2024-01-18 GroupPolarizationStatmod/plot.R

@@ -62,3 +62,37 @@ plotFreq <- function(inputDf, fittedFreq, nBins)
              adj=c(0.5,0.5), padj=c(-0.5,-0.5))
     }
 }
+
+
+plot_metric_cohens <- function(metric_cohens_csv = "data/output/metric_cohens_d.csv",
+                               output_file = "figures/metric_cohens.pdf") {
+  
+  read_csv(metric_cohens_csv) %>% 
+    unite(ExperimentID, ArticleTag, TreatmentTag) %>% 
+    ggplot(mapping = aes(x = Cohens_d, y = ExperimentID)) + 
+      geom_vline(xintercept = 0, color = "red") + 
+      geom_vline(xintercept = c(-1, 1), color="red", linetype="dotted") + 
+      geom_boxplot() +
+      xlim(-2.0, 2.0)
+  
+  ggsave(output_file)
+}
+
+
+plot_ordinal_cohens <- function(ordinal_data_dir = "data/probit_fits",
+                                output_file = "figures/ordinal_cohens.pdf") {
+  
+  df <- dir_ls(ordinal_data_dir, glob = "*.csv") %>%
+    read_csv() %>% unite(ExperimentID, ArticleTag, TreatmentTag)
+  
+  df$Cohens_d <- cohens_d(df$LatentMeanPrePosteriorMean, df$LatentMeanPostPosteriorMean,
+                          df$LatentMeanPrePosteriorSD, df$LatentMeanPostPosteriorSD)
+  
+  ggplot(df, mapping = aes(x = Cohens_d, y = ExperimentID)) + 
+    geom_vline(xintercept = 0, color = "red") + 
+    geom_vline(xintercept = c(-1, 1), color="red", linetype="dotted") + 
+    geom_boxplot() #+
+    # xlim(-2.0, 2.0)
+  
+  ggsave(output_file)
+}

### 2024-01-19 GroupPolarizationStatmod/paper/ms.tex
updated figures and started metric model analysis

@@
 \usepackage{caption}
+\usepackage{subcaption}
 \usepackage{pdflscape}
@@
   \noindent
-  % Psychology is beset by three crises: the reproducibility crisis, the
-  % generalizability crisis, and the theory crisis. New verbal theories for various
-  % phenomena are constantly being developed orthogonal to other theories, all based on
-  % mostly unreproducible analyses; even if a positive result is in some sense ``true''
-  % it may not account for all the variance present in a set of observations---studies
-  % thus fail to generalize because their results do not account for this variance.
-  % There is another related problem: a phenomenon may be ``reproducible'' only
-  % because studies of the phenomenon systematically misuse statistics in such a way
-  % that induces false positives for each succeeding ``replication.'' 
   \emph{Group polarization} is a theoretical process where like-minded novel group members 
@@
   as ordinal-valued reported opinions, e.g., on a Likert-type scale. 
-  This creates ceiling effects that mask whether extreme opinions become less extreme.
-  We developed a generative model to demonstrate that 95\% of sixty experimental conditions
+  This creates a ceiling effect that masks extreme opinions becoming less extreme.
+  We developed a generative model to 
+  demonstrate that 95\% of sixty experimental conditions
   across ten group polarization papers report plausibly false detections of
@@
   approach using disparate, incommensurate study designs.
+\end{abstract}
 
-  % To show this, we developed a
-  % novel generative model that provides two levels of evidence that an historical
-  % published empirical finding of group polarization is plausibly false.  Our model
-  % can be used to evaluate additional studies in group polarization and beyond, even
-  % if the reported data and significant metrics are incompletely reported, and the
-  % original raw data has been lost to time.  This may be just the tip of the
-  % plausibly-false-positives-in-psychology iceberg, both for group polarization
-  % studies and for the legion psychological results that model ordinal observations of
-  % opinions, beliefs, emotions, or any other latent psychological variable that is
-  % theoretically continuous, but measured using an ordinal scale.  
-\end{abstract}
 \clearpage
@@
                          $\mu$ and variance $\sigma_t^2$ \\
-   $k \in \{0,\ldots,K\} $ & Threshold index ($k$) between 0 and total number of
-         ordinal bins, $K$ \\
+   $k \in \{0,\ldots,B\} $ & Threshold index ($k$) between 0 and total number of
+         ordinal bins, $B$ \\
    $b_k$ & Opinion bin $k$; note there is no $k=0$ bin. \\
@@
 
-Here we evaluate if metric linear models inferred group polarization from
-simulated simple consensus data using parameters identified above. Those
-parameters specified continous pre- and post-deliberation
+Here we evaluate if metric linear models inferred group polarization from simulated
+simple consensus data using parameters identified above. Those parameters specified
+continous pre- and post-deliberation means and variance of normal opinion
+distributions that represent simple consensus but yields
+pre- and post-deliberation ordinal distributions with different, group-polarized
+means. This parameter-finding process implicitly assumes $N\to\infty$ participants.
 
-% \spacing{1.0}
-% \hspace{-1em}
-% \input{t-test-table.tex}
-% % \input{t-test-table-custom.tex}
-% \spacing{1.5}
+It is therefore necessary to evaluate whether the simple consensus parameters so 
+identified in the previous step yield simulated data that also appear to be
+group-polarized when measured on an ordinal scale. We draw $N$ samples from the
+pre- and post-deliberation normal distributions and bin them into one of 
+$B$ ordinal values. $N$ and $B$ are set specifically for each of the 
+60 experimental conditions analyzed here. We then fit a linear model to the 
+simulated observed pre- and post-deliberation data to infer whether the mean
+shifted, indicating group polarization. We measure the pre- and
+post-deliberation standard deviations as well. Then we calculate Cohen's $d$
+to quantify the effect size for each simulation. A suite of simulations are
+necessary since there might be significant variance in $d$ for different
+generations of simulated pre- and post-deliberation observations. Therefore we
+perform 1000 trials for each experimental condition and inspect the resulting
+distributions of $d$ across trials in a boxplot.
+
+\vspace{0.5em}
+\begin{center}
+{[ FIGURE \ref{fig:MetricBoxplot_N} ABOUT HERE ]} \\
+\end{center}
 
@@
 is a true or false positive.
+
+We did make some simplifying assumptions that could affect model outcomes.
 
@@
 
-\section{Notes on studies}
+\section*{Figures}
+
+\begin{figure}
+  \caption{
+    Boxplots visualizing the distribution of Cohen's $d$ values (x-axis in
+    A and B) inferred through a metric linear model over 1000 simulation 
+    trials for each experimental condition (y-axis in A and B). All conditions
+    have many trials that result in non-zero Cohen's $d$ values with number of
+    simulated participants, $N$, set to their empirical values reported in the
+    associated journal article (A). However, when the number of simulated
+    participants is increased to $10N$, there are several conditions that 
+    indeed yield distributions closer to $d = 0$ as expected, however some
+    conditions have $d$ distributions centered on a nonzero mean with little to
+    no overlap with $d=0$; variance in $d$ decreases as expected
+    across experimental conditions (B).
+  }  
+  \label{fig:MetricBoxplot}
+  \centering
+  \begin{subfigure}[]{0.9\textwidth}
+    % \hspace{-7.5em} 
+    \centering
+    \textsf{\large \textbf{A}, empirical $N$} \\
+    \includegraphics[width=\textwidth]{Figures/boxplots/metric_cohens.pdf}
+    \label{fig:MetricBoxplot_N}
+  \end{subfigure} \\
+\end{figure}
+\begin{figure}\ContinuedFloat
+  \begin{subfigure}[]{0.9\textwidth}
+    % \hspace{-5.5em} 
+    \centering
+    \textsf{\large \textbf{B}, 10 $\times$ empirical $N$} \\
+    \includegraphics[width=\textwidth]{Figures/boxplots/10N_metric_cohens.pdf}
+    \label{fig:MetricBoxplot_10N}
+  \end{subfigure}
+\end{figure}
+
+\clearpage
+
+\section*{Notes on studies}
 

### 2024-04-09 GroupPolarizationStatmod/paper.bk/tables/conditions_table.tex
paper rewrite

% \newgeometry{margin=1cm} % modify this if you need even more space
\begin{landscape}

%put your table here

% \begin{table}
%   \small
%   \begin{tabular}{p{1.5in}lp{2.5in}p{3.5in}}
  % \begin{longtable}{p
  \small
  \begin{longtable}{p{1.5in}lp{2.0in}p{4.0in}}
    \caption{Summary of idiosyncracies across experimental conditions in the
    studies we scrutinized for plausibly-false positives.}
    \label{tab:conditionsComparison}\\
  % \begin{longtable}{llll}
  % \begin{tabular}{llll}
  \toprule
  \textbf{Article (Article Tag)} & \textbf{Condition Tag} & \textbf{Prompt} & \textbf{Description} \\
  \midrule
  \citeA{Moscovici1969} (Moscovici1969) & Americans-Opinion & American economic aid is always used for
  political pressure.  & 11 total prompts (1 reported), all 7-point scales. Unclear how data was aggregated. \\
  \midrule
  ~(Moscovici1969) & DeGaulle-Opinion & Charles De Gaulle is too old to (be president) & 12 total prompts (1 reported), all 7-point scales. Unclear how data was aggregated. \\
  \midrule
  ~(Moscovici1969) & DeGaulle-Judgment & Charles De Gaulle is too old to (be president) & Same as
  above, except ``subjects were asked first to judge individually and then to reach consensus as a
  group in evaluating the varoableness or unfavorableness of each item toward de Gaulle,
  regardless of whether they agreed...~with the content of the item.'' \\
  \bottomrule
  \end{longtable}
  % \end{tabular}
% \end{table}

\end{landscape}
\restoregeometry

### 2024-04-09 GroupPolarizationStatmod/paper/ms.qmd
paper rewrite

---
title: "Prominent claims of group polarization are plausibly false with lessons for studying psychological dynamics"
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
    cite-method: biblatex
    biblatexoptions: sorting=none, style=numeric-comp
    pdf-engine: latexmk
    pdf-engine-opt: -xelatex
    # template-partials:
    #   - _extensions/partials/title.tex
authors:
  - name: Matthew Adam Turner
    institute: Stanford University Doerr School of Sustainability, Stanford University
    roles: writing, model development, project development
    corresponding: true
    email: maturner@stanford.edu
  - name: Paul E. Smaldino
    institute: 
      - Cognitive and Information Sciences, Univeristy of California, Merced
      - Santa Fe Institute
    roles: writing
    corresponding: true
bibliography: ../library.bib
classoption: abstract
keep-tex: true
abstract: |
  This is the abstract
geometry:
  - margin=1.1in
  - letterpaper
biblio-title: References
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).

### 2024-04-09 GroupPolarizationStatmod/paper/ms.tex
paper rewrite

--- 2024-01-18 GroupPolarizationStatmod/paper/ms.tex
@@
-\documentclass[11pt, letterpaper]{article}
-\usepackage{marvosym}
-\usepackage{float}
-\usepackage{setspace}
-\usepackage{longtable}
-\usepackage{caption}
-\usepackage{pdflscape}
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
+%
+\documentclass[
+  letterpaper,
+  abstract]{scrartcl}
+
+\usepackage{amsmath,amssymb}
+\usepackage{iftex}
+\ifPDFTeX
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else  
+    % xetex/luatex font selection
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\usepackage[margin=1.1in,letterpaper]{geometry}
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\setcounter{secnumdepth}{5}
+% Make \paragraph and \subparagraph free-standing
+\ifx\paragraph\undefined\else
+  \let\oldparagraph\paragraph
+  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+  \let\oldsubparagraph\subparagraph
+  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
+
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{241,243,245}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
+\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
+\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
+\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
+\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
+\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
+
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
 \usepackage{graphicx}
-% DOCUMENT LAYOUT
-% \usepackage{geometry} 
-\usepackage[margin=1.0in]{geometry}
-\usepackage{lineno}
-\usepackage[usenames,dvipsnames]{xcolor}
-\definecolor{myorange}{RGB}{240, 96, 0}
-\newcommand{\mt}[1]{{\textcolor{myorange} {({\tiny MT:} #1)}}}
-% \geometry{letterpaper, textwidth=5.5in, textheight=8.5in, marginparsep=7pt, marginparwidth=.6in}
-% \setlength\parindent{0in}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
 
-\definecolor{linenocolor}{gray}{0.6}
-\renewcommand\thelinenumber{\color{linenocolor}\arabic{linenumber}}
-% \calclayout %centers text on page.
+\KOMAoption{captions}{tableheading}
+\makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\AtBeginDocument{%
+\ifdefined\contentsname
+  \renewcommand*\contentsname{Table of contents}
+\else
+  \newcommand\contentsname{Table of contents}
+\fi
+\ifdefined\listfigurename
+  \renewcommand*\listfigurename{List of Figures}
+\else
+  \newcommand\listfigurename{List of Figures}
+\fi
+\ifdefined\listtablename
+  \renewcommand*\listtablename{List of Tables}
+\else
+  \newcommand\listtablename{List of Tables}
+\fi
+\ifdefined\figurename
+  \renewcommand*\figurename{Figure}
+\else
+  \newcommand\figurename{Figure}
+\fi
+\ifdefined\tablename
+  \renewcommand*\tablename{Table}
+\else
+  \newcommand\tablename{Table}
+\fi
+}
+\@ifpackageloaded{float}{}{\usepackage{float}}
+\floatstyle{ruled}
+\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
+\floatname{codelisting}{Listing}
+\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
+\makeatother
+\makeatletter
+\makeatother
+\makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
+\makeatother
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\usepackage[sorting=none, style=numeric-comp]{biblatex}
+\addbibresource{../library.bib}
+\usepackage{bookmark}
 
-\usepackage{amsmath}
-\usepackage{mathspec}
-% \setallmainfonts(Digits,Latin){STIX Two Text}
-% \setallmainfonts(Digits,Latin){Baskerville}
-% \setallmainfonts(Digits,Latin){Plantagenet Cherokee}
-\setallmainfonts(Digits,Latin){Palatino}
-% \setallmainfonts(Digits,Latin){Hoefler Text}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  pdftitle={Prominent claims of group polarization are plausibly false with lessons for studying psychological dynamics},
+  colorlinks=true,
+  linkcolor={blue},
+  filecolor={Maroon},
+  citecolor={Blue},
+  urlcolor={Blue},
+  pdfcreator={LaTeX via pandoc}}
 
-% HEADINGS
-% \usepackage{sectsty} 
-% \usepackage[normalem]{ulem} 
-% \sectionfont{\upshape\Large}
-% \subsectionfont{\scshape\normalsize} 
-% \subsubsectionfont{\upshape\large}
-
-% PDF SETUP
-% ---- FILL IN HERE THE DOC TITLE AND AUTHOR
-\usepackage[%dvipdfm, 
-bookmarks, colorlinks, breaklinks, 
-% ---- FILL IN HERE THE TITLE AND AUTHOR
-	pdftitle={Many claims of group polarization are plausibly false due to inappropriate
-    statistics},
-	pdfauthor={Matthew A. Turner},
-	pdfproducer={https://mt.digital}
-]{hyperref}  
-\hypersetup{linkcolor=blue,citecolor=blue,filecolor=black,urlcolor=MidnightBlue}
-
-\usepackage{apacite}
-\usepackage{booktabs}
-\usepackage{authblk}
-
-\newcommand{\sigmapre}{\sigma_\mathrm{pre}}
-\newcommand{\sigmapost}{\sigma_\mathrm{post}}
-
-\newcommand{\normalpre}{\mathcal{N}(\mu, \sigmapre)}
-\newcommand{\normalpost}{\mathcal{N}(\mu, \sigmapost)}
-
-\newcommand{\ordmeanpre}{\langle o_{\mathrm{pre}}}
-\newcommand{\ordmeanpost}{\langle o_{\mathrm{post}}}
+\title{Prominent claims of group polarization are plausibly false with
+lessons for studying psychological dynamics}
+\author{Matthew Adam Turner \and Paul E. Smaldino}
+\date{}
 
 \begin{document}
+\maketitle
+\begin{abstract}
+This is the abstract
+\end{abstract}
 
-\title{Many claims of group polarization are plausibly false due to inappropriate
-    statistics}
-% \shorttitle{Group polarization or simple consensus?}
-% \author[1,2]{{Matthew A.~Turner}}
-\author[1]{{Matthew A.~Turner}}
-% \affil[1]{Department of Earth System Science, Stanford University}
-\affil[1]{Division of Social Sciences, Stanford Doerr School of Sustainability, Stanford University}
+\renewcommand*\contentsname{Table of contents}
+{
+\hypersetup{linkcolor=}
+\setcounter{tocdepth}{3}
+\tableofcontents
+}
+\section{Quarto}\label{quarto}
 
-% \author[3,4,5]{{Paul E.~Smaldino}}
-\author[2,3]{{Paul E.~Smaldino}}
-\affil[2]{Cognitive and Information Sciences, University of California, Merced}
-% \affil[4]{Center for Advanced Study in the Behavioral Sciences, Stanford University}
-\affil[3]{Santa Fe Institute}
+Quarto enables you to weave together content and executable code into a
+finished document. To learn more about Quarto see
+\url{https://quarto.org}.
 
-\maketitle
-% \setlength{\parindent}{15pt}
-% \maketitle
+\section{Running Code}\label{running-code}
 
-\spacing{1.5}
+When you click the \textbf{Render} button a document will be generated
+that includes both content and the output of embedded code. You can
+embed code like this:
 
-\linenumbers
-% \modulolinenumbers[3]
+\begin{Shaded}
+\begin{Highlighting}[]
+\DecValTok{1} \SpecialCharTok{+} \DecValTok{1}
+\end{Highlighting}
+\end{Shaded}
 
-\begin{abstract}
-  \noindent
-  % Psychology is beset by three crises: the reproducibility crisis, the
-  % generalizability crisis, and the theory crisis. New verbal theories for various
-  % phenomena are constantly being developed orthogonal to other theories, all based on
-  % mostly unreproducible analyses; even if a positive result is in some sense ``true''
-  % it may not account for all the variance present in a set of observations---studies
-  % thus fail to generalize because their results do not account for this variance.
-  % There is another related problem: a phenomenon may be ``reproducible'' only
-  % because studies of the phenomenon systematically misuse statistics in such a way
-  % that induces false positives for each succeeding ``replication.'' 
-  \emph{Group polarization} is a theoretical process where like-minded novel group members 
-  become more extreme in their opinions after group discussion, but also find \emph{consensus}, reducing 
-  group-level opinion variance.
-  \emph{Simple consensus} may occur instead, where opinion variance decreases but
-  the mean group opinion does not change.
-  Here we show that many published detections of group polarization are explained
-  equally rigorously as simple consensus.
-  This is the case because experimenters failed
-  to account for the measurement of continous-valued psychological opinions 
-  as ordinal-valued reported opinions, e.g., on a Likert-type scale. 
-  This creates ceiling effects that mask whether extreme opinions become less extreme.
-  We developed a generative model to demonstrate that 95\% of sixty experimental conditions
-  across ten group polarization papers report plausibly false detections of
-  group polarization. Our approach can be used to evaluate whether other detections of
-  group polarization may be plausibly false---our approach could be applied beyond
-  group polarization as well. This work also provides a
-  framework for new group polarization studies that could
-  share a common, rigorous, and interoperable study design to multiply our knowledge
-  of this potentially important group process, avoiding the historically fragmented
-  approach using disparate, incommensurate study designs.
+\begin{verbatim}
+[1] 2
+\end{verbatim}
 
-  % To show this, we developed a
-  % novel generative model that provides two levels of evidence that an historical
-  % published empirical finding of group polarization is plausibly false.  Our model
-  % can be used to evaluate additional studies in group polarization and beyond, even
-  % if the reported data and significant metrics are incompletely reported, and the
-  % original raw data has been lost to time.  This may be just the tip of the
-  % plausibly-false-positives-in-psychology iceberg, both for group polarization
-  % studies and for the legion psychological results that model ordinal observations of
-  % opinions, beliefs, emotions, or any other latent psychological variable that is
-  % theoretically continuous, but measured using an ordinal scale.  
-\end{abstract}
-\clearpage
+You can add options to executable code like this
 
-\begin{quote}
-In our introductory social psychology course, 
-we have for many years used the [group polarization experimental paradigm] as
-a laboratory exercise. The exercise works beautifully, but one must be
-careful to forewarn a class that [group polarization] does not occur with every 
-group\ldots and that the effect is not large. 
-\par\raggedleft\cite[p. 205]{Brown1986}
-\end{quote}
+\begin{verbatim}
+[1] 4
+\end{verbatim}
 
-\begin{quote}
-One of the most robust findings in social psychology is that of attitude polarization 
-following discussion with like-minded others.
-\par\raggedleft\cite[p. 267]{Cooper2001}
-\end{quote}
+The \texttt{echo:\ false} option disables the printing of code (only
+output is displayed).
 
-\section{Introduction}
-
-\begin{itemize}
-  \item 
-    \emph{Group polarization} is a process where the mean opinion of a 
-    like-minded group becomes more extreme over time~\cite{Brown1986,Brown2000}. 
-    It is therefore an important component process in
-    the broader process of increasing political polarization that can hamstring
-    governments~\cite{Mason2018UncivilAgreementBook,Klein2020}.
-  \item
-    Individuals' opinions in group polarization experiments are typically 
-    measured using an ordinal scale, such
-    as a Likert scale, where behavioral experiment participants' internal, theoretically
-    continuous, opinions are converted to one of several opinion bins indiciating
-    their degree of agreement or disagreement with some prompt statement.
-  \item
-    Recently, however, it has been shown that it is critical to use 
-    appropriate statistical methods when using experimental designs that 
-    measure continuous latent psychological variables with ordinal
-    measurements---failure to do so can result in false positive support for
-    the presence of experimental effects~\cite{Liddell2018}.
-  \item
-    In this paper, we develop a generative modeling approach for evaluating whether or
-    not published detections of group polarization are plausibly false, i.e.,
-    whether they are \emph{plausibly-false positives} (PFPs). 
-  \item
-    We found significant evidence that many published group polarization findings
-    are indeed plausibly false since we found distribution parameters that
-    generated simulated pre- and post-discussion data that a $t$-test found
-    to be from distributions with different means, although there was no difference
-    in mean, only a difference in variance. 
-  \item
-    Although a $t$-test may be fooled by simple consensus under ordinal measurements,
-    we show that a more appropriate ordered probit model of the data yield no
-    such plausibly-false positives.
-  \item
-    Our work shows the need for using appropriate statistical tests, as well as
-    the need for a consistent experimental paradigm so that meaningful meta-analyses
-    can be performed for theoretical integration and rigorous understanding of
-    the group polarization effect, should it prove to occur in future group
-    polarization studies that use proper statistics. In addition to avoiding
-    $t$-tests, future group polarization studies can be more reliable by
-    accounting for random effects arising from, e.g., socioeconomic status,
-    gender, and other individual-level factors that may bias how much and in what
-    direction a participant changes their opinions. Failure to do so can result 
-    in artificially-inflated effect sizes~\cite{Yarkoni2021}.
-\end{itemize}
-
-\subsection{Group polarization}
-
-Group polarization is often induced in behavioral experiments by first answering
-an opinion prompt, grouping
-relatively like-minded individuals together to discuss the prompt, then 
-again polling opinions; the change in mean opinion towards a more extreme
-consensus than the initial mean is the amount of group polarization that has occurred
-(Figure~\ref{fig:groupPolarizationSchematic}).
-
-While all ten studies scrutinized here use the general experimental procedure
-outlined above (Figure~\ref{fig:groupPolarizationSchematic}), each study and constituent 
-experimental condition have idosyncracies that makes the results themselves
-more opaque, and that makes comparison of different studies' findings more
-difficult. (REVIEW A FEW SIMPLE EXAMPLES). A detailed examination of each study's
-idiosyncracies is given in Table~\ref{tab:conditionsComparison}.
-
-\begin{figure}
-  \caption{Group polarization experiments trials in the ten studies
-    analyzed here proceed in three steps: 
-    (A) \textbf{Pre-deliberation.} Participant opinions are measured using an 
-    ordinal scale. 
-    (B) \textbf{Deliberation.} Groups discuss their opinions on the prompt. 
-    (C) \textbf{Post-deliberation.} Participant opinions are measured again.
-    Group polarization in one trial is said to occur if 
-    the mean post-deliberation opinion is greater than the pre-deliberation 
-    opinion.
-}
-  \label{fig:groupPolarizationSchematic}
-  \centering
-  \includegraphics[width=\textwidth]{Figures/GPExperimentSketch.pdf}
-\end{figure}
-
-% \spacing{1.0}
-% \input{tables/conditions_table.tex}
-% \spacing{1.5}
-
-\subsection{Statistical inference in opinion change}
-
-2-3 par: the right way; what GP studies do; why it's a problem (N-1 sents) and how
-to fix it (1 sent)
-
-When
-
-\begin{itemize}
-  \item 
-    Group polarization studies use null-hypothesis testing that implicitly
-    assume (1) \emph{opinions} (or other equivalent latent psychological variables
-    such as \emph{attitudes} or \emph{beliefs}) are continuously-valued, and
-    (2) 
-  \item
-    The exact null-hypothesis testing procedure varies between studies $t$-tests, ANOVA, etc. 
-\end{itemize}
-
-Recently, problems with using have been made clear~\cite{Liddell2018}.
-The way to fix this problem going forward is to use a thresholded-normal 
-statistical model for group polarization data, or other related statistical 
-model that accounts for opinion binning~\cite[Ch. 23]{Liddell2018,KruschkeDBDA}.
-
-\subsection{Study selection and overview}
-
-Studies were selected to be either influential papers, based on citations or
-author prominence, that test either examine
-group polarization in political opinions, and to be representative of investigations
-of four primary group polarization explanations extant in the literature.
-At the time of paper selection we did not examine the statistical methods in detail beyond
-checking that they indeed used
-
-Because of the difficulties casting individual research articles into our modeling
-framework, and the central task of creating the model and analyzing this first
-batch of studies with it, we limit our analysis to 10 studies. It is problem
-enough that these 10 prominent studies have the problems we demonstrate they
-plausibly do. Problems among these 10 suggest the problem may be more wide spread. 
-When we do more analyses
-they will be aided by improved data management systems for applying this model to 
-additional studies (eventually outside of group polarization).
-
-Statistical problems among these studies are not limited to a mismatch between
-measurement and statistical procedures, which requires the detailed treatment
-we provide in this paper. Other problems are more pedestrian.
-One major issue shared by several studies
-is the use of unpaired $t$-tests to detect group polarization. Since the same
-group has their opinions measured twice over time, the opinions are correlated
-and so a paired $t$-test would be required for test assumptions to be met 
-(assuming all other paired $t$-test assumptions were met, 
-which of course they are not, which is why our paper here is necessary).
-In some cases the reported shifts were ``significant'', 
-but not truly ``group polarization'' shifts, i.e., 
-shifts from initially biased group opinions
-to more extreme group opinions biased in the same direction; instead shifts 
-were from a neutral initial group opinion to a biased group opinion 
-(as in, e.g.,~\citeA{Abrams1990} Uncategorized2), or from
-an opinion biased in one direction (as in, e.g.,~\citeA{Abrams1990} Categorized1,
-Categorized2, AND OTHERS???). In some results of \citeA{Freidkin1999a}, 
-shift values are reported with their
-standard deviations---but the shift standard deviations are greater than the
-shifts themselves, indiciating that zero shift is included in the observed
-shift confidence interval. This is \emph{prima facie} a negative result that
-should not have been counted as evidence of group polarization.
-
-Some group polarization studies do not measure continuous latent opinions using
-a metric scale, and so are not affected by the floor and ceiling effects that
-undermine a good number of experimental results, as we examine here. There may,
-however, be other problems, such as non-group polarization shifts and the use
-of unpaired $t$-tests when paired tests are required.
-
-\subsection{Research overview}
-
-In this paper we analyze whether published statistical detections of group 
-polarization obtained through null-hypothesis statistical tests
-are plausibly false because they can be explained equally well by simple consensus
-where the mean group opinion does not change over time, but opinion variance does.
-This analysis requires several preliminary steps. 
-First we formally develop our generative model of simple consensus and group 
-polarization in terms of changes in group opinion mean and variance from 
-pre- to post-deliberation. Then we developed a search algorithm for identifying 
-which, if any, simple consensus scenarios with 
-constant latent mean and pre- and post-deliberation opinion variances 
-could generate published observations of group polarization. To apply this model
-and algorithm to published results, we developed an interactive web application
-for data input and to test which different potential simple consensus scenarios 
-could generate group polarization opinion shifts. After identifying plausibly false
-detections of group polarization we examine whether the identified simple
-consensus scenario (parameters?) can indeed generate data to fool the 
-paper's null-hypothesis tests into thinking the data actually support the inference
-that simple consensus did not occur. Finally, we examine whether, in these same
-plausibly false detections, a Bayesian thresholded-normal statistical model
-will accurately identify simple consensus instead of falsely detecting group
-polarization.
-
-\section{Model}
-
-Our model represnts the generative process of finding different forms of group consensus
-in group polarization experiments, and the process of measuring that consensus using
-an ordinal measurement scale, e.g.\ a Likert-type scale. This model operationalizes two
-relevant forms of consensus: \emph{simple consensus} and \emph{group polarization}.
-\mt{Need to explain in Intro why consensus should occur along with group
-polarization} 
-% Group polarization experiments are designed to result group consensus
-% since they group together like-minded individuals to see if they become more
-% extreme after discussion~\cite{Turner2018}.
-
-We collected group polarization data from ten empirical studies with sixty constituent
-experimental conditions across all studies. We recorded the reported opinion shift and pre- and
-post-deliberation mean opinions, and the number of participants
-in each condition, as long as this data was available---\citeA{Burnstein1975} 
-did not report pre- and post-deliberation
-means, only the shift in means. We did not include experimental treatments that did not 
-yield a positive detection of group polarization since we are only estimating
-the rate of plausible false positive detections of group polarization.
-
-Based on this data
-
-% In two cases, reported opinion and shift values were converted from an ordinal
-% scale to a percentage scale, specifically \citeA{Friedkin1999a} and \citeA{Krizan2007}.
-% The authors in these papers did report the original ordinal measurement scale,
-% so we converted the means back to the orignial measurement scale.
-% In a few cases from \citeA{Friedkin1999a} we did not need to apply our 
-% model to the reported data because we could determine a reported opinion shift was 
-% \emph{prima facie} false since reported opinion shift magnitudes
-% were smaller than the standard deviation in shifts (see Table 1, p. 868, \emph{ibid}),
-% despite Friedkin's claim that regressions (not clearly defined) on all shifts 
-% yielded $p < 0.05$~\cite{Kruschke2018c}.
-
-\subsection{Latent opinions, ordinal measurements, and how simple consensus can look like
-group polarization}
-
-We begin by developing a formal model of ordinal measurement of latent continuous
-opinions (model variables and parameters are given in Table~\ref{tab:modelVariables}).
-First, we assume that participant $i$ has latent opinion at time $t$ that is drawn from
-a group-level normal distribution,
-\begin{equation}
-  o_{i,t} \sim p(o; \mu_t, \sigma_t) = \mathcal{N}(\mu_t, \sigma_t).
-  \label{eq:opinionDistribution}
-\end{equation}
-\noindent 
-This distribution has mean $\mu_t$ and standard deviation $\sigma_t$.
-This form is chosen to
-reflect the common statistical analysis used by group polarization studies,
-not because it is statistically appropriate---in fact it is not, which is what
-we aim to show in this paper.
-
-It is appropriate to assume opinions are normally distributed because published
-group polarization studies use some sort of $t$-tests, or similar tests, which are designed
-to detect whether the mean of two distributions are the same or not. This imposes the
-implicit assumption that opinions themselves are normal. But as we will show, 
-metric statistical tests such as $t$-tests reliably generate false rejections
-of the null hypothesis that there is no difference between two distributions' means.
-
-\subsubsection{How simple consensus can look like group polarization}
-
-Simple consensus can look like group polarization when (1) the mean latent opinion
-does not change; (2) both more and less extreme opinions move toward this 
-unchanged latent mean (i.e., opinion variance shrinks due to consensus); 
-and (3) an ordinal scale is used that bins all extreme
-latent opinions into the same extreme bin even though those means have changed,
-but opinion bins that are less extreme or of opposite polarity more accurately
-reflect the underlying change in latent opinion distribution. This process is
-illustrated in Figure~\ref{fig:consensusDistroIllustration}, where we show
-a toy example of what can occur when the pre-consensus
-opinions (upper left plot)
-
-\begin{table}[h]
-  \caption{Variables used in the formal model of group polarization from simple consensus.}
-  \label{tab:modelVariables}
-  \begin{tabular}{cp{5.0in}} \toprule
-   Symbol & Description  \\ \midrule  
-   $t \in \{\mathrm{pre}, \mathrm{post}\}$ & Time, either \emph{pre}- or
-                                           \emph{post}-discussion \\
-   $\mu$  &   Latent mean, assumed constant for simple consensus \\ 
-   $\sigma_t$ & Latent standard deviation at time $t$ \\
-   $p(o;\mu,\sigma_t)$ & Normal latent distribution of opinions with mean
-                         $\mu$ and variance $\sigma_t^2$ \\
-   $k \in \{0,\ldots,K\} $ & Threshold index ($k$) between 0 and total number of
-         ordinal bins, $K$ \\
-   $b_k$ & Opinion bin $k$; note there is no $k=0$ bin. \\
-   $\theta_{k-1},~\theta_k$  & Lower, upper threshold value for binning latent opinions into
-     bin $b_k$ \\
-   $p(o \in b_k; \mu, \sigma_t)$ & Ordinal density of binned opinions in bin $b_k$ 
-            (Equation~\ref{eq:binFrequency}) \\
-   $\langle o_t \rangle$  &   Observed mean opinion calculated from ordinal data at time $t$ \\
-   \bottomrule
-  \end{tabular} 
-\end{table}
-
-\begin{figure}
-  \caption{\emph{Simple consensus} occurs in a group when its latent opinion
-  distribution variance decreases, but the mean does not change (A). When 
-latent opinions are measured on an ordinal scale, the mean of the distribution
-can be distorted by ceiling (or floor) effects (B). Group polarization 
-can appear to have occurred where the mean of the observed opinions appears
-to increase despite the latent mean being constant (C).}
-  \label{fig:consensusDistroIllustration}
-  \vspace{1em}
-  \centering
-    \includegraphics[width=1.0\textwidth]{ConsensusDistroIllustration/figure.pdf}
-\end{figure}
-
-This distributional view models the measurement of $N \to \infty$ participant opinions,
-which results in a histogram of frequency of opinions
-in each bin, i.e., $o_{i,t} = b_k$. The frequency of responses in each bin is 
-the integral over the continuous normal probability density function from 
-one bin threshold to another. This transforms the probability density 
-function from $p(o_{i,t};~\mu_t, \sigma_t)$ to the probability
-of observing a reported opinion of each bin value. 
-The probability of observing an opinion in bin
-$b_k$ is calculated by integrating the normal probability density 
-function over the range of $\theta_{k-1}$ to $\theta_{k}$. 
-Formally, we write the 
-probability of observing an opinion in bin $k$ at time $t$ as
-\begin{equation}
-\begin{aligned}
-  p(o \in b_k; \mu_t, \sigma_t, \theta) 
-    &= \int_{\theta_{k-1}}^{\theta_k} p(o;\mu_t, \sigma_t) do \\
-    &= \Phi \left( \frac{o - \mu_t}{\sigma_t} \right)\Big|_{o=\theta_{k-1}}^{\theta_k} \\
-    &= \Phi \left( \frac{\theta_k - \mu_t}{\sigma_t} \right) - 
-       \Phi \left( \frac{\theta_{k-1} - \mu_t}{\sigma_t} \right)
-\end{aligned}
-\label{eq:binFrequency}
-\end{equation}
-\noindent
-where $\Phi(\frac{o - \mu}{\sigma})$ is the normalized 
-normal cumulative distribution function over opinions $o$, shifted by an amount $\mu$ with
-standard deviation $\sigma$.
-
-\subsection{Evaluating whether a group polarization finding is plausibly false}
-
-We have shown in one toy example that group polarization can appear to occur
-when ordinal measures are used, without accounting for the binning process; now
-we show how to .
-
-We develop two measures for how plausible it is that a reported positive detection
-of group polarization is indeed a false positive, building our confidence 
-in the plausibility at each step. In this way, we take a modern Bayesian approach
-to scientific evidence that typically does not dichotomize ``true'' and ``false''
-into absolutes, but evaluates degrees of belief instead.
-
-The first measure we use for whether a reported detection of group polarization
-is plausibly false is whether we can find a combination of latent simple
-consensus parameters $\mu$, $\sigmapre$, and $\sigmapost$ representing latent
-distributions that, when binned (Equation~\ref{eq:binFrequency} and
-Figure~\ref{fig:consensusDistroIllustration}), result in ordinal distributions
-that have the same difference in means reported in the literature that the authors
-took to mean group polarization had occurred. Note that these calculations are
-done for the case where effectively we have $N \to \infty$; next we turn to
-the question of whether data generated from these simple consensus distributions
-generate a ``significant'' group polarization result.
-
-The second measure of how plausible it is that a group polarization is false
-is whether ordinal data generated by binning draws from the latent distributions
-is found to be significantly different according to Welch's $t$-test.
-
-\subsubsection{Identify null-model parameters that yield plausibly-false positives}
-
-To find simple consensus parameters $\mu$, $\sigmapre$, and $\sigmapost$ that
-generate ordinal distributions showing group polarization, we used the above
-framework in an optimization routine that takes an initial guess for these latent
-parameters and hillclimbs to find $\sigmapre$ and $\sigmapost$ that most accurately
-generate ordinal data that appears to have had group polarization occur.
-
-\subsubsection{Metric model inference of Cohen's $d$ for simulated simple-consensus data}
-
-After identifying null-model parameters that yield plausibly-false positives 
-in the $N \to \infty$ case, we generated binned samples from normal opinion 
-distributions $\normalpre$ and $\normalpost$. Specifically, for each experimental
-condition that passed the initial $N \to \infty$ test, we generated 1000 
-pre- and post-deliberation ordinal datasets.
-
-We generated these simulated data using the latent mean, pre-deliberation 
-standard deviation, and post-deliberation standard deviations as parameters of
-normal distributions of pre- and post-deliberation opinions.
-
-\subsection{Bayesian ordered-probit models inference of Cohen's $d$ for simulated
-simple-consensus data}
-
-Non-metric ordered probit models, which account for the data collection 
-process, theoretically should be better able to correctly
-distinguish simple consensus compared to metric models. Therefore, we also fit
-the following ordered probit model to simulated group polarization data.
-
-\section{Analysis}
-
-\subsection{Plausible parameters for false positives}
-
-We found that X\% of group polarization detections across all studies were 
-plausibly false positives (Table~\ref{tab:XXXXX}).
-
-\vspace{.5em}
-\spacing{1.0}
-% \input{plausibleFP-table.tex}
-\begin{table}[H]
-    \begin{center}
-      \caption{Number and fraction of experimental conditions where group polarization was
-               reported to have occurred, but plausibly did not.}
-      \vspace{.5em}
-      \input{plausibleFP-table-custom.tex}
-    \end{center}
-  \end{table}
-\spacing{1.5}
-
-\subsection{Metric linear models often fail to identify simple consensus}
-
-Here we evaluate if metric linear models inferred group polarization from
-simulated simple consensus data using parameters identified above. Those
-parameters specified continous pre- and post-deliberation
-
-% \spacing{1.0}
-% \hspace{-1em}
-% \input{t-test-table.tex}
-% % \input{t-test-table-custom.tex}
-% \spacing{1.5}
-
-\subsection{Ordered-probit models also fail?}
-
-We found that in many cases Bayesian ordered-probit models avoid falsely identifying simple
-consensus as group polarization. This supports the claim that it is the choice of statistical
-procedure, including the failure to account for the ordinal nature of, and variance within,
-opinion measuremenets in group polarization experiments that leads to false
-positives, not anything intrinsic to the phenomenon of group polarization.
-
-\subsection{More data enables ordered probit models, but not metric models, to reliably
-  identify simple consensus}
-
-  \mt{Idea: increase $N$ for simulations and see effect on distribution of Cohen's $d$
-  for both metric and ordinal models}
-
-\section{Discussion}
-
-We showed that many prominent detections of group polarization are plausibly
-false, and that future studies \emph{must} use ordered-probit statistical
-models and estimate highest-density or confidence intervals, not point-estimates 
-and significance values~\cite{Meehl1997}.
-
-Due to a lack of data sharing in general, and often from decades-old papers, and a persistent 
-failure of group polarization work to provide opinion variance information, 
-it is impossible to know whether an historical detection of group polarization
-is a true or false positive.
-
-This problem is made all the worse by the lack of a consistent experimental design for group
-polarization experiments. This makes it impossible to systematically account for
-additional sources of vairance known to lead to over-estimates
-of effect sizes~\cite{Yarkoni2021}, independent of the false positives problem
-we identified in the current paper.
-
-% This is due to the rush to explain why 
-% group polarization occurs using weak theory~\cite{Cartwright1971,Cartwright1973} and makes
-% group polarization meta-analyses (e.g. Brown, 1986; and Isenberg, 1986)\nocite{Brown1986,Isenberg1986}
-% unreliable~\cite{Meehl1990,Meehl1997}.
-
-Group polariztion is an essential social phenomenon that could be a major driver of
-progress-halting political polarization more broadly. If echo chambers of like-minded
-individuals tend to become radicalized it is essential to measure by how much
-in different contexts, and to develop rigorous theoretical explanations of
-this phenomenon. It appears, unfortunately, that decades of group polarization
-research may have published many plausibly false inferences of the
-occurrence of group polarization, as demonstrated by our simple consensus
-model of false group polarization detection. As mentioned, a lack of
-consistency and failure to account for additional sources of variance could
-be two additional ways published research is underpowered, and associated
-theoretical insights undermined. It may not be easy to get the group polarization field to 
-consolidate experimental designs and use appropriate statistical procedures,
-but it is a straightforward solution. Indeed, combining inter-study consistency and statistical
-rigor is the only way to a reliable understanding of group polarization.
-
-\bibliographystyle{apacite}
-
-\setlength{\bibleftmargin}{.125in}
-\setlength{\bibindent}{-\bibleftmargin}
-
-\bibliography{/Users/mt/workspace/Writing/library.bib}
-
-\appendix
-
-\section{Notes on studies}
-
-Different group polarization studies have different goals, and measure shifts
-in a variety of ways, using different (or no) statistical tests to confirm
-the presence of group polarization opinion shifts. This supplemental section
-reviews the ten papers one-by-one and explains how group 
-polarization was measured in each paper. I also explain how I decided which
-group polarization detections I included from each paper, and how we decided
-if they were plausibly false positives, including whether we needed to
-
-\paragraph{{Abrams1990}} This study is meant to support the presence of the "self-categorization" effect in
-social psychology. The paper presents three experiments and their results--Experiment 3 is the group polarization
-experiment, designed to show that self-categorization is at work in the emergence of group polarization. Uses high
-schoolers aged 16-17 in Hawkes Bay, New Zealand. There are five deliberation topics (p. 112) across two conditions
-(p. 113): one condition where participants are made aware of the relationship between their opinions and other group
-members' opinions ("Categorized") and another condition where participants were given no information about their
-opinion relative to other group members' opinions ("Uncategorized"). It was confusing to me what scale is used for
-Table 2 (p. 114), but parsing the "Dependent measures" subsection several times led me to understand the last
-paragraph, that "[a]ttitude shift overall was...assessed...on the unconverted +4 to -4 scale."
-
-\paragraph{{Burnstein1973}} First I want to note this study was funded in part by a Guggenheim fellowship (see
-note on first page) and the authors acknolwedge the assistance of "vinous" discussions with colleagues at the
-University of Provence and the "spirited and intelligent efforts" of two female research assistants. The study
-presents the results of two experiments (really one experiment with two pools of participants) that test the
-persuasive arguments hypothesis by having students perform the standard group polarization experiment with five CDQ
-discussion items, items A,B,E,F, and H (A,B risk-inducing; E, F caution-inducing; H neutral). Experiment 1 used 149
-male intro psych students and Experiment 2 used 76 male intro psych students from the University of Michigan. There
-are three conditions across the two experiments--one hypothesized to show an emergent shift and the other two
-hypothesized to not show a shift. None of the conditions allowed participants to know each others' ratings exactly,
-the participants only deliberated the various topics. In the one hypothesized to show a shift, participants were
-instructed to argue for their position. In the other two, participants argued against their position. The "against"
-condition was further divided into a case where everyone knew everyone was prompted to argue against their position,
-and another sub-condition where participants did not know if others were arguing for or against their true position.
-We include only the first condition where participants argued for their position, which is the only condition
-expected to show shifts (though others do, but the authors let themselves off the hook saying, "They are curious
-findings and defy straightforward explanation...we will merely note their occurrence and not venture to speculate as
-to their cause"). Thus, in our table we have conditions A, B, E, F, and H. Even though H is supposed to not show a
-shift, it does in this study, which is possibly further evidence that this study merely shows participant consensus
-formation, not opinion shifts.
-
-\paragraph{{Burnstein1975}} In this study, Burnstein and Vinokur ran an experiment to evaluate their theory
-that generating persuasive arguments are necessary for group polarization to occur. The authors design a study with
-three conditions, one Experimental and two Control, where participants (60 male students from intro psych at the
-University of Michigan) first answered four CDQ items that typically generate shifts towards more risk, then, in the
-Experimental condition, saw group members' responses and produced arguments that might support the positions of
-other participants. In the first control condition, Condition II in the paper and Control1 in our table,
-participants were exposed to others' opinions, but does not get to reflect on their opinions or give arguments. In
-condition III, the second control condtition (we call Control2) participants do not get to know other participants'
-opinions at all. Thus, Burnstein and Vinokur hypothesize that significant shifts will only occur in Condition I, the
-Experimental condition. We thus only include Condition I in our analysis. The authors only report the shift, meaning
-we are free to choose any pre- and post-deliberation value to see if the null hypothesis is actually plausible.
-
-\paragraph{{Schkade2010}} Friedkin uses three CDQ topics, "Sports", "School", and "Surgery" that use a
-20-point CDQ risk scale in fractional units, 0.05, 0.10, ..., 1.00 across three group sizes (2, 3, and 4, which he
-calls "dyads", "triads", and "tetrads", which we use in our CaseStudyTag). Friedkin only gives opinion shifts, and
-apparently transforms the fractional CDQ responses into percentages, since the shifts range between 5 and 9.
-However, closer inspection of Friedkin's results shows that the reported "Choice Shift" is not equal to the
-difference between mean final and initial opinions. Friedkin also uses "different" means for pre- and
-post-discussion opinions, which seems easy enough to show are really the same by re-organizing terms in the sum.
-Friedkin uses two topics in the tetrad condition where responses are in units of money where participants can freely
-give any number they choose ("Asbestos" and "Disaster"), and another topic called "Institutions" where participants
-rate 13 institutions, which participants rated on a scale from 0 to 100 percent to indicate the degree of trust the
-participant has in each institution. We only consider the CDQ topics since these are the ones rated on a discrete
-scale.
-
-\paragraph{{Hogg1990}} It is not clear if there are indeed significant
-pre-post deliberation differences reported in this paper. The Results are a
-tangle of estimates of effect sizes and diagnostic statistical measures. It
-is also not clear that they actually report the pre- and post-deliberation
-opinions; it appears they only report the shifts. ``Opinion Difference
-Measures'' does not seem to include a test that the shifts themselves are
-reliable. I decided to count a finding from this study as a false positive if
-the model generated pre- and post-deliberation mean opinions within an
-absolute error of 0.01 of the reported value. The argument seems to be that
-since they are testing social ``frame[s] of reference'' (Table 1) that it is
-not necessary to show that group polarization itself occurs, only that the
-various experimental treatments yield different shifts---but if the shifts
-themselves are insignificant, then what is being measured, exactly?
-
-\paragraph{{Krizan2007}} Uses five CDQ topics ("scenarios") from Kogan and
-Wallach (1967) that typically provide shifts toward more risky decisions.  These are
-given in Footnote 2 on p. 194. Krizan and Baron report shifts and mean pre-test
-opinions on the risk scale, meaning that shifts are expected to be negative towards
-lower odds of success required to take the risk, i.e., the acceptance of more risk.
-Krizan and Baron report mean pre-test opinions, but average over all three conditions
-for each scenario, then report shifts for each scenario and condition, so we
-effectively have no way for an exact match of initial mean to condition x scenario,
-where conditions are no-discussion, deliberation without knowledge of an out-group's
-opinion, and deliberation with knowledge of an out-group's opinion. We are only
-interested in the non-control cases--K\&B found no difference in knowing out-group
-opinion, which they say is evidence against the self-categorization hypothesis of group
-polarization.  Therefore, we have ten cases we test across five scenarios and two
-conditions with our model, and we use the reported initial mean for each scenario
-twice--once for each scenario across the two conditions. Post-deliberation opinions are
-calculated in the spreadsheet using the reported shifts. K\&B also report the "Overall"
-shift, but I find this meaningless since it is across different items with no
-statistical control for inter-item variability. I noted that false detections are
-plausible in all cases due to the fact that the standard deviations of shifts overwhelm
-the shifts themselves, but I will still run the model on these cases to confirm.
-
-\paragraph{{Moscovici1969}} Moscovici and Zavalloni studied whether group
-polarization is more general than CDQ by using opinions about political topics. There
-were two general topics--attitudes about "Americans" and attitudes about then-president
-of France Charles "DeGaulle"--with 12 and 11 specific items, respectively, on which
-participants gave opinions and deliberated. Participants deliberated on one of the two
-general topics, giving opinions on each item on a seven-point Likert scale, -3 to +3.
-The authors reported pre-consensus and post-consensus mean opinions taken over all
-participants and items for each of the two general topics I have called the "Americans"
-and "DeGaulle" conditions.
-
-\paragraph{{Myers1970}} Myers and Bishop test whether group polarization occurs
-in groups with different levels of racial bias determined by a survey that is not also
-the deliberation topic. The participants were high school students from western
-Michigan. In the experimental condition, social influence occurs through deliberation;
-in the control condition there is no deliberation or social influence at all, just
-re-tested opinions after a distraction task. Obviously we only report on whether there
-was a significant shift in the experimental condition. In both conditions, three groups
-were created according to their racial prejudice as measured by a 100-item racial
-attitude measurement instrument, called "low", "medium", and "high" prejudice groups.
-The participants then gave their opinions, deliberated, and gave opinions again on
-eight deliberation topics. Means were taken across items and prejudice group.
-
-\paragraph{{Myers1975}} Myers performs two experiments each with experimental and control conditions for two group types with initial biases.
-In Experiment 1, Myers creates biased groups by having participants (intro psychology students) read a description of six different teachers, three "bad" professors and three "good" professors, which the participants are assumed to infer through vignettes about the professors.
-The participants in the experimental condition of Experiment 1 deliberated in small groups after giving their initial opinions on six items a 0-to-10 bad-to-good scale.
-In Experiment 2, participants were put into conservatively or liberally biased small groups based on their pre-deliberation opinions on six deliberation items regarding womens' rights and the role of women in the family and society.
-Again, in the experimental group, participants deliberated each item before giving their final opinions on the items.
-In the control group in each Experiment, there was no deliberation, just some filler task.
-Curiously, significant shifts were still found in the control groups--it is unclear to us why.
-Experiment 1 also had participants adjust the bad and good professors' pay to obtain non-ordinal measures of opinion regarding these professors, which we do not consider here.
-However, as we will discuss in the paper, this does not make it free of problems--like so many other group polarization studies, Myers statistical model does not account for inter-item, inter-group, or inter-participant variability.
-
-\paragraph{{Schkade2010}} Participants in Schkade, Sunstein, and Hastie's study were put into two politically ideologically biased groups based on their geographical location in the US state of Colorado: participants were drawn from Boulder, Colorado (abbreviated CO), to be liberally biased and from Colorado Springs, CO, to be conservatively biased.
-Participants gave their opinions and discussed three topics related to Global Warming, Civil Unions, and Affirmative Action.
-Pre- and post-deliberation mean are given in Table 1 (p.
-232) for all six (topics x geographies).
+\printbibliography[title=References]
 

### 2024-04-09 GroupPolarizationStatmod/plot.R

@@ -1,6 +1,13 @@
 ##
 # Some methods for plotting results from statmods.
 #
+
+require(fs)
+require(tidyverse)
+
+source("experiments.R")
+
+
 LIKERT_SCALE_DENS_MAX <- 0.7
 
 ##

### 2024-04-09 GroupPolarizationStatmod/paper.bk/ms.tex
init quarto paper rewrite

--- 2022-11-18 GroupPolarizationStatmod/papers/ms.tex
@@
-% % Author: Matthew Turner
-
-% % \documentclass[11pt,letterpaper]{scrartcl}
-% % \documentclass[letterpaper,man,natbib]{apa6}
-% \documentclass[11pt,letterpaper]{amsart}
-% % \documentclass[11pt, letterpaper]{article}
-% % \usepackage{natbib}
-% % \documentclass[11pt]{report}
-% % \documentclass{report}
-% % \documentclass{book}
-% \usepackage[bookmarks, hidelinks]{hyperref}
-% \usepackage{amssymb,amsmath}
-% % \usepackage{fullpage}
-% \usepackage{tabulary}
-% \usepackage{tabularx}
-% \usepackage{float}
-% % \usepackage[margin=1.00in]{geometry}
-% \usepackage[margin=0.90in]{geometry}
-
-% % \usepackage{fontspec}
-
-% % \usepackage{xunicode}
-% % \usepackage{xltxtra}
-% % \defaultfontfeatures{Mapping=tex-text}
-% % \setmainfont[Ligatures={Common}, Numbers={OldStyle}, Variant=01]{Hoefler Text}
-% % \setromanfont[Ligatures={Common}, Numbers={OldStyle}, Variant=01]{Hoefler Text}
-% % \setmonofont[Scale=0.95]{PT Sans}
-% \usepackage{caption}
-% \usepackage{apacite}
-% \usepackage{booktabs}
-% \usepackage{pslatex}
-% \usepackage{subcaption}
-% \usepackage{pgfplots}
-% \usepackage{wrapfig}
-% \usepackage[english]{babel}
-% \usepackage{lmodern}
-% % \doublespace
-% % \usepackage{url}
-% \usepackage{bigfoot}
-% \usepackage[export]{adjustbox}
-% % \setlength\intextsep{0pt}
-
-% \usepackage{graphicx}
-
-\documentclass[12pt, letterpaper]{article}
-\usepackage{fontspec} 
+\documentclass[11pt, letterpaper]{article}
 \usepackage{marvosym}
+\usepackage{float}
 \usepackage{setspace}
 \usepackage{longtable}
+\usepackage{caption}
+\usepackage{subcaption}
+\usepackage{pdflscape}
+\usepackage{graphicx}
 % DOCUMENT LAYOUT
 % \usepackage{geometry} 
-\usepackage[margin=0.90in]{geometry}
+\usepackage[margin=1.0in]{geometry}
 \usepackage{lineno}
 \usepackage[usenames,dvipsnames]{xcolor}
+\definecolor{myorange}{RGB}{240, 96, 0}
+\newcommand{\mt}[1]{{\textcolor{myorange} {({\tiny MT:} #1)}}}
 % \geometry{letterpaper, textwidth=5.5in, textheight=8.5in, marginparsep=7pt, marginparwidth=.6in}
@@
 
-% \renewcommand{\thesubsection}{\arabic{subsection}.}
-% FONTS
-% \usepackage[usenames,dvipsnames]{color}
-% \usepackage{xunicode}
-% \usepackage{xltxtra}
-\defaultfontfeatures{Mapping=tex-text}
-% \setromanfont[Ligatures={Common}, Variant=01]{Hoefler Text}
-% \setromanfont[Ligatures={Common}, Variant=01]{Palatino}
-\setromanfont[Ligatures={Common}, Variant=01]{Baskerville}
-% \setromanfont[Ligatures={Common}, Variant=01]{Plantagenet Cherokee}
-% \setmainfont [BoldFont={STIX Two Text SemiBold}, Ligatures={Common},
-% Numbers={OldStyle}, Variant=01]{STIX Two Text}
-% \setmainfont [BoldFont={STIX Two Text SemiBold}, Ligatures={Common},
-% Numbers={OldStyle}, Variant=01]{STIX Two Text}
-% \setromanfont [Ligatures={Common},
-% Numbers={OldStyle}, Variant=01]{STIX Two Text}
-\setmonofont[Scale=0.95]{PT Sans}
-
-% ---- CUSTOM COMMANDS
-% \chardef\&="E050
-\newcommand{\html}[1]{\href{#1}{\scriptsize\textsc{[html]}}}
-\newcommand{\pdf}[1]{\href{#1}{\scriptsize\textsc{[pdf]}}}
-\newcommand{\doi}[1]{\href{#1}{\scriptsize\textsc{[doi]}}}
-\newcommand{\lurl}[1]{\href{#1}{\scriptsize\textsc{[link]}}}
-
-\newcommand{\githubsymbol}{\raisebox{-2pt}{\includegraphics[height=10.5pt]{github}}}
-\newcommand{\twittersymbol}{\raisebox{-2pt}{\includegraphics[height=10.5pt]{twitter}}}
-\newcommand{\monitorsymbol}{\raisebox{-2pt}{\includegraphics[height=10.5pt]{monitor}}}
-
-% ---- MARGIN YEARS
-\usepackage{marginnote}
-\newcommand{\amper{}}{\chardef\amper="E0BD }
-\newcommand{\years}[1]{\marginnote{\small #1}}
-\renewcommand*{\raggedleftmarginnote}{}
-\setlength{\marginparsep}{7pt}
-\reversemarginpar
+\usepackage{amsmath}
+\usepackage{mathspec}
+% \setallmainfonts(Digits,Latin){STIX Two Text}
+% \setallmainfonts(Digits,Latin){Baskerville}
+% \setallmainfonts(Digits,Latin){Plantagenet Cherokee}
+\setallmainfonts(Digits,Latin){Palatino}
+% \setallmainfonts(Digits,Latin){Hoefler Text}
 
 % HEADINGS
-\usepackage{sectsty} 
-\usepackage[normalem]{ulem} 
-\sectionfont{\upshape\Large}
-\subsectionfont{\scshape\normalsize} 
-\subsubsectionfont{\upshape\large}
+% \usepackage{sectsty} 
+% \usepackage[normalem]{ulem} 
+% \sectionfont{\upshape\Large}
+% \subsectionfont{\scshape\normalsize} 
+% \subsubsectionfont{\upshape\large}
 
@@
 
+\newcommand{\sigmapre}{\sigma_\mathrm{pre}}
+\newcommand{\sigmapost}{\sigma_\mathrm{post}}
+
+\newcommand{\normalpre}{\mathcal{N}(\mu, \sigmapre)}
+\newcommand{\normalpost}{\mathcal{N}(\mu, \sigmapost)}
+
+\newcommand{\ordmeanpre}{\langle o_{\mathrm{pre}}}
+\newcommand{\ordmeanpost}{\langle o_{\mathrm{post}}}
+
 \begin{document}
 
-\title{Group polarization is often a mirage caused by inappropriate statistical analyses}
-% \shorttitle{Group polarization or simple consensus?}
-\author[1,2]{{Matthew A.~Turner}}
-\affil[1]{Department of Earth System Science, Stanford University}
-\affil[2]{Division of Social Sciences, Stanford Doerr School of Sustainability, Stanford University}
+\title{Many claims of group polarization are plausibly false due to inappropriate
+    statistics}
+\author[1]{{Matthew A.~Turner}}
+\affil[1]{Division of Social Sciences, Stanford Doerr School of Sustainability, Stanford University}
 
-\author[3,4,5]{{Paul E.~Smaldino}}
-\affil[3]{Cognitive and Information Sciences, University of California, Merced}
-\affil[4]{Center for Advanced Study in the Behavioral Sciences, Stanford University}
-\affil[5]{Santa Fe Institute}
+\author[2,3]{{Paul E.~Smaldino}}
+\affil[2]{Cognitive and Information Sciences, University of California, Merced}
+\affil[3]{Santa Fe Institute}
+
+\date{}
 
@@
 \linenumbers
-\modulolinenumbers[5]
+% \modulolinenumbers[3]
 
@@
   \noindent
-  Group polarization is the observed tendency of
-  like-minded groups to arrive at a more extreme consensus opinion than
-  the average opinion after discussing some topic of interest.
-  Clearly it is important to understand such a phenomenon to understand broader
-  problematic trends towards increased polariztion across society as a whole.
-  Unfortunately, recent analyses have demonstrated that this measurement and
-  analysis approach can result in false positive detections group polarization
-  from failing to account for floor/ceiling effects induced when using ordinal scales to
-  measure continuously-distributed latent psychological variables. 
-  Here we demonstrate that many high profile detections
-  of group polarization are plausibly false
-  due to these floor/ceiling effects using generative modeling seeded by 
-  reported group polarization values and parameters. After we identify which
-  reported detections of group polarization are plausibly false detections, we further
-  investigate whether the latent constant mean opinion and pre- and post-discussion
-  opinion variances can fool the statistical tests used in each study to 
-  generate false positives. Finally, we evaluate whether a more appropriate
-  Bayesian cumulative normal statistical model can more accurately distinguish
-  group polarization from simple consensus. Group polarization continues to 
-  pose a problem for psychological theory with an ever-expanding set of potential theoretical
-  explanations. A search for explanations is pointless, however, without sound
-  evidence. We believe this work will provide a sturdier analytical foundation for 
-  future group-polarization research.
+  \emph{Group polarization} is a theoretical process where like-minded novel group members 
+  become more extreme in their opinions after group discussion, but also find \emph{consensus}, reducing 
+  group-level opinion variance.
+  \emph{Simple consensus} may occur instead, where opinion variance decreases but
+  the mean group opinion does not change.
+  Here we show that many published detections of group polarization are explained
+  equally rigorously as simple consensus.
+  This is the case because experimenters failed
+  to account for the measurement of continous-valued psychological opinions 
+  as ordinal-valued reported opinions, e.g., on a Likert-type scale. 
+  This creates a ceiling effect that masks extreme opinions becoming less extreme.
+  We developed a generative model to 
+  demonstrate that 95\% of sixty experimental conditions
+  across ten group polarization papers report plausibly false detections of
+  group polarization. Our approach can be used to evaluate whether other detections of
+  group polarization may be plausibly false---our approach could be applied beyond
+  group polarization as well. This work also provides a
+  framework for new group polarization studies that could
+  share a common, rigorous, and interoperable study design to multiply our knowledge
+  of this potentially important group process, avoiding the historically fragmented
+  approach using disparate, incommensurate study designs.
 \end{abstract}
+
+\clearpage
+
+\begin{quote}
+In our introductory social psychology course, 
+we have for many years used the [group polarization experimental paradigm] as
+a laboratory exercise. The exercise works beautifully, but one must be
+careful to forewarn a class that [group polarization] does not occur with every 
+group\ldots and that the effect is not large. 
+\par\raggedleft\cite[p. 205]{Brown1986}
+\end{quote}
+
+\begin{quote}
+One of the most robust findings in social psychology is that of attitude polarization 
+following discussion with like-minded others.
+\par\raggedleft\cite[p. 267]{Cooper2001}
+\end{quote}
 
@@
 
+\begin{itemize}
+  \item 
+    \emph{Group polarization} is a process where the mean opinion of a 
+    like-minded group becomes more extreme over time~\cite{Brown1986,Brown2000}. 
+    It is therefore an important component process in
+    the broader process of increasing political polarization that can hamstring
+    governments~\cite{Mason2018UncivilAgreementBook,Klein2020}.
+  \item
+    Individuals' opinions in group polarization experiments are typically 
+    measured using an ordinal scale, such
+    as a Likert scale, where behavioral experiment participants' internal, theoretically
+    continuous, opinions are converted to one of several opinion bins indiciating
+    their degree of agreement or disagreement with some prompt statement.
+  \item
+    Recently, however, it has been shown that it is critical to use 
+    appropriate statistical methods when using experimental designs that 
+    measure continuous latent psychological variables with ordinal
+    measurements---failure to do so can result in false positive support for
+    the presence of experimental effects~\cite{Liddell2018}.
+  \item
+    In this paper, we develop a generative modeling approach for evaluating whether or
+    not published detections of group polarization are plausibly false, i.e.,
+    whether they are \emph{plausibly-false positives} (PFPs). 
+  \item
+    We found significant evidence that many published group polarization findings
+    are indeed plausibly false since we found distribution parameters that
+    generated simulated pre- and post-discussion data that a $t$-test found
+    to be from distributions with different means, although there was no difference
+    in mean, only a difference in variance. 
+  \item
+    Although a $t$-test may be fooled by simple consensus under ordinal measurements,
+    we show that a more appropriate ordered probit model of the data yield no
+    such plausibly-false positives.
+  \item
+    Our work shows the need for using appropriate statistical tests, as well as
+    the need for a consistent experimental paradigm so that meaningful meta-analyses
+    can be performed for theoretical integration and rigorous understanding of
+    the group polarization effect, should it prove to occur in future group
+    polarization studies that use proper statistics. In addition to avoiding
+    $t$-tests, future group polarization studies can be more reliable by
+    accounting for random effects arising from, e.g., socioeconomic status,
+    gender, and other individual-level factors that may bias how much and in what
+    direction a participant changes their opinions. Failure to do so can result 
+    in artificially-inflated effect sizes~\cite{Yarkoni2021}.
+\end{itemize}
+
 \subsection{Group polarization}
+
+Group polarization is often induced in behavioral experiments by first answering
+an opinion prompt, grouping
+relatively like-minded individuals together to discuss the prompt, then 
+again polling opinions; the change in mean opinion towards a more extreme
+consensus than the initial mean is the amount of group polarization that has occurred
+(Figure~\ref{fig:groupPolarizationSchematic}).
+
+While all ten studies scrutinized here use the general experimental procedure
+outlined above (Figure~\ref{fig:groupPolarizationSchematic}), each study and constituent 
+experimental condition have idosyncracies that makes the results themselves
+more opaque, and that makes comparison of different studies' findings more
+difficult. (REVIEW A FEW SIMPLE EXAMPLES). A detailed examination of each study's
+idiosyncracies is given in Table~\ref{tab:conditionsComparison}.
+
+\begin{figure}
+  \caption{Group polarization experiments trials in the ten studies
+    analyzed here proceed in three steps: 
+    (A) \textbf{Pre-deliberation.} Participant opinions are measured using an 
+    ordinal scale. 
+    (B) \textbf{Deliberation.} Groups discuss their opinions on the prompt. 
+    (C) \textbf{Post-deliberation.} Participant opinions are measured again.
+    Group polarization in one trial is said to occur if 
+    the mean post-deliberation opinion is greater than the pre-deliberation 
+    opinion.
+}
+  \label{fig:groupPolarizationSchematic}
+  \centering
+  \includegraphics[width=\textwidth]{Figures/GPExperimentSketch.pdf}
+\end{figure}
+
+% \spacing{1.0}
+% \input{tables/conditions_table.tex}
+% \spacing{1.5}
 
@@
 in group polarization experiments, and the process of measuring that consensus using
-an ordinal measurement scale, e.g.\ a Likert-type scale. This model considers two
+an ordinal measurement scale, e.g.\ a Likert-type scale. This model operationalizes two
 relevant forms of consensus: \emph{simple consensus} and \emph{group polarization}.
-Either process is \emph{consensus} since opinion variance among the group 
-decreases as individuals come to agree more than they did previously. 
-Group polarization experiments are designed to foster group consensus
-it is not necessary to consider clustering, negative influence, or other
-processes known to generally occur when opinions within a group significantly
-diverge~\cite{Turner2018}.
+\mt{Need to explain in Intro why consensus should occur along with group
+polarization} 
+% Group polarization experiments are designed to result group consensus
+% since they group together like-minded individuals to see if they become more
+% extreme after discussion~\cite{Turner2018}.
 
 We collected group polarization data from ten empirical studies with sixty constituent
-experimental treatments across all studies, specifically the opinion shift and pre- and
+experimental conditions across all studies. We recorded the reported opinion shift and pre- and
 post-deliberation mean opinions, and the number of participants
-in each treatment (\citeA{Burnstein1975} did not report pre- and post-deliberation
-means, only the shift in means). We did not include experimental treatments that did not 
+in each condition, as long as this data was available---\citeA{Burnstein1975} 
+did not report pre- and post-deliberation
+means, only the shift in means. We did not include experimental treatments that did not 
 yield a positive detection of group polarization since we are only estimating
 the rate of plausible false positive detections of group polarization.
-We then created new data that indicated whether a reported.
 
-In two cases, reported opinion and shift values were converted from an ordinal
-scale to a percentage scale, specifically \citeA{Friedkin1999a} and \citeA{Krizan2007}.
-The authors in these papers did report the original ordinal measurement scale,
-so we converted the means back to the orignial measurement scale.
-In a few cases from \citeA{Friedkin1999a} we did not need to apply our 
-model to the reported data because we could determine a reported opinion shift was 
-\emph{prima facie} false since reported opinion shift magnitudes
-were smaller than the standard deviation in shifts (see Table 1, p. 868, \emph{ibid}),
-despite Friedkin's claim that regressions (not clearly defined) on all shifts 
-yielded $p < 0.05$~\cite{Kruschke2018c}.
+Based on this data
 
-\subsection{Latent opinions, ordinal measurements}
+% In two cases, reported opinion and shift values were converted from an ordinal
+% scale to a percentage scale, specifically \citeA{Friedkin1999a} and \citeA{Krizan2007}.
+% The authors in these papers did report the original ordinal measurement scale,
+% so we converted the means back to the orignial measurement scale.
+% In a few cases from \citeA{Friedkin1999a} we did not need to apply our 
+% model to the reported data because we could determine a reported opinion shift was 
+% \emph{prima facie} false since reported opinion shift magnitudes
+% were smaller than the standard deviation in shifts (see Table 1, p. 868, \emph{ibid}),
+% despite Friedkin's claim that regressions (not clearly defined) on all shifts 
+% yielded $p < 0.05$~\cite{Kruschke2018c}.
 
-Our analysis relies on two important formal concepts: latent opinions and their 
-measurement on an ordinal scale. \emph{Latent opinions} are unobservable psychological,
-continuous-valued variables that are somehow represented and stored in the
-brain-body. 
-When researchers use continuous statistical models to infer 
-that group polarization occurred (which all ten considered here do), 
-they are implicitly assuming that latent opinions are themselves continuous.
-We can only measure behaviors in group polarization experiments,
-specifically the behavior of participants marking one ordinal value or another
-to indicate their opinion---a process that is not well understood in itself,
-but which is unimportant for our analyses. \emph{Ordinal measurements} (e.g.\
-a Likert or other discrete-valued opinion scale) of latent opinions force
-participants to appropriately bin their opinion in one of several categories
-representing the valence and extremity of their opinion, e.g., on a
-7-point Likert scale -3 could indcate
-``strongly disagree'', +3 could be ``strongly agree'', 0 would be neutral, and 
-intermediate values are less extreme opinions of either valence.
+\subsection{Latent opinions, ordinal measurements, and how simple consensus can look like
+group polarization}
 
-All ten studies analyzed here used an ordinal scale, but group polarization experiments seem to
-lack any systematicity in experimental design beyond that. (EXAMPLES)
+We begin by developing a formal model of ordinal measurement of latent continuous
+opinions (model variables and parameters are given in Table~\ref{tab:modelVariables}).
+First, we assume that participant $i$ has latent opinion at time $t$ that is drawn from
+a group-level normal distribution,
+\begin{equation}
+  o_{i,t} \sim p(o; \mu_t, \sigma_t) = \mathcal{N}(\mu_t, \sigma_t).
+  \label{eq:opinionDistribution}
+\end{equation}
+\noindent 
+This distribution has mean $\mu_t$ and standard deviation $\sigma_t$.
+This form is chosen to
+reflect the common statistical analysis used by group polarization studies,
+not because it is statistically appropriate---in fact it is not, which is what
+we aim to show in this paper.
 
-\subsection{Simple consensus and group polarization are statistical patterns}
+It is appropriate to assume opinions are normally distributed because published
+group polarization studies use some sort of $t$-tests, or similar tests, which are designed
+to detect whether the mean of two distributions are the same or not. This imposes the
+implicit assumption that opinions themselves are normal. But as we will show, 
+metric statistical tests such as $t$-tests reliably generate false rejections
+of the null hypothesis that there is no difference between two distributions' means.
 
-\subsection{Analytical approach}
+\subsubsection{How simple consensus can look like group polarization}
 
-The goal of this study is to understand if published detections of group polarization are
-plausibly false.
+Simple consensus can look like group polarization when (1) the mean latent opinion
+does not change; (2) both more and less extreme opinions move toward this 
+unchanged latent mean (i.e., opinion variance shrinks due to consensus); 
+and (3) an ordinal scale is used that bins all extreme
+latent opinions into the same extreme bin even though those means have changed,
+but opinion bins that are less extreme or of opposite polarity more accurately
+reflect the underlying change in latent opinion distribution. This process is
+illustrated in Figure~\ref{fig:consensusDistroIllustration}, where we show
+a toy example of what can occur when the pre-consensus
+opinions (upper left plot)
 
-Therefore we designed a series of tests to understand if, and to what degree,
-published results are false positive detections of group polarization that
-are at least as well described by a simple-consensus model.
+\begin{table}[h]
+  \caption{Variables used in the formal model of group polarization from simple consensus.}
+  \label{tab:modelVariables}
+  \begin{tabular}{cp{5.0in}} \toprule
+   Symbol & Description  \\ \midrule  
+   $t \in \{\mathrm{pre}, \mathrm{post}\}$ & Time, either \emph{pre}- or
+                                           \emph{post}-discussion \\
+   $\mu$  &   Latent mean, assumed constant for simple consensus \\ 
+   $\sigma_t$ & Latent standard deviation at time $t$ \\
+   $p(o;\mu,\sigma_t)$ & Normal latent distribution of opinions with mean
+                         $\mu$ and variance $\sigma_t^2$ \\
+   $k \in \{0,\ldots,B\} $ & Threshold index ($k$) between 0 and total number of
+         ordinal bins, $B$ \\
+   $b_k$ & Opinion bin $k$; note there is no $k=0$ bin. \\
+   $\theta_{k-1},~\theta_k$  & Lower, upper threshold value for binning latent opinions into
+     bin $b_k$ \\
+   $p(o \in b_k; \mu, \sigma_t)$ & Ordinal density of binned opinions in bin $b_k$ 
+            (Equation~\ref{eq:binFrequency}) \\
+   $\langle o_t \rangle$  &   Observed mean opinion calculated from ordinal data at time $t$ \\
+   \bottomrule
+  \end{tabular} 
+\end{table}
 
-\subsubsection{Identifying parameters that plausibly generate false detections of group polarization}
+\begin{figure}
+  \caption{\emph{Simple consensus} occurs in a group when its latent opinion
+  distribution variance decreases, but the mean does not change (A). When 
+latent opinions are measured on an ordinal scale, the mean of the distribution
+can be distorted by ceiling (or floor) effects (B). Group polarization 
+can appear to have occurred where the mean of the observed opinions appears
+to increase despite the latent mean being constant (C).}
+  \label{fig:consensusDistroIllustration}
+  \vspace{1em}
+  \centering
+    \includegraphics[width=1.0\textwidth]{ConsensusDistroIllustration/figure.pdf}
+\end{figure}
 
-\subsubsection{$t$-tests of generated simple-consensus data}
+This distributional view models the measurement of $N \to \infty$ participant opinions,
+which results in a histogram of frequency of opinions
+in each bin, i.e., $o_{i,t} = b_k$. The frequency of responses in each bin is 
+the integral over the continuous normal probability density function from 
+one bin threshold to another. This transforms the probability density 
+function from $p(o_{i,t};~\mu_t, \sigma_t)$ to the probability
+of observing a reported opinion of each bin value. 
+The probability of observing an opinion in bin
+$b_k$ is calculated by integrating the normal probability density 
+function over the range of $\theta_{k-1}$ to $\theta_{k}$. 
+Formally, we write the 
+probability of observing an opinion in bin $k$ at time $t$ as
+\begin{equation}
+\begin{aligned}
+  p(o \in b_k; \mu_t, \sigma_t, \theta) 
+    &= \int_{\theta_{k-1}}^{\theta_k} p(o;\mu_t, \sigma_t) do \\
+    &= \Phi \left( \frac{o - \mu_t}{\sigma_t} \right)\Big|_{o=\theta_{k-1}}^{\theta_k} \\
+    &= \Phi \left( \frac{\theta_k - \mu_t}{\sigma_t} \right) - 
+       \Phi \left( \frac{\theta_{k-1} - \mu_t}{\sigma_t} \right)
+\end{aligned}
+\label{eq:binFrequency}
+\end{equation}
+\noindent
+where $\Phi(\frac{o - \mu}{\sigma})$ is the normalized 
+normal cumulative distribution function over opinions $o$, shifted by an amount $\mu$ with
+standard deviation $\sigma$.
 
-\subsubsection{Bayesian ordered-probit models of generated simple-consensus data}
+\subsection{Evaluating whether a group polarization finding is plausibly false}
+
+We have shown in one toy example that group polarization can appear to occur
+when ordinal measures are used, without accounting for the binning process; now
+we show how to .
+
+We develop two measures for how plausible it is that a reported positive detection
+of group polarization is indeed a false positive, building our confidence 
+in the plausibility at each step. In this way, we take a modern Bayesian approach
+to scientific evidence that typically does not dichotomize ``true'' and ``false''
+into absolutes, but evaluates degrees of belief instead.
+
+The first measure we use for whether a reported detection of group polarization
+is plausibly false is whether we can find a combination of latent simple
+consensus parameters $\mu$, $\sigmapre$, and $\sigmapost$ representing latent
+distributions that, when binned (Equation~\ref{eq:binFrequency} and
+Figure~\ref{fig:consensusDistroIllustration}), result in ordinal distributions
+that have the same difference in means reported in the literature that the authors
+took to mean group polarization had occurred. Note that these calculations are
+done for the case where effectively we have $N \to \infty$; next we turn to
+the question of whether data generated from these simple consensus distributions
+generate a ``significant'' group polarization result.
+
+The second measure of how plausible it is that a group polarization is false
+is whether ordinal data generated by binning draws from the latent distributions
+is found to be significantly different according to Welch's $t$-test.
+
+\subsubsection{Identify null-model parameters that yield plausibly-false positives}
+
+To find simple consensus parameters $\mu$, $\sigmapre$, and $\sigmapost$ that
+generate ordinal distributions showing group polarization, we used the above
+framework in an optimization routine that takes an initial guess for these latent
+parameters and hillclimbs to find $\sigmapre$ and $\sigmapost$ that most accurately
+generate ordinal data that appears to have had group polarization occur.
+
+\subsubsection{Metric model inference of Cohen's $d$ for simulated simple-consensus data}
+
+After identifying null-model parameters that yield plausibly-false positives 
+in the $N \to \infty$ case, we generated binned samples from normal opinion 
+distributions $\normalpre$ and $\normalpost$. Specifically, for each experimental
+condition that passed the initial $N \to \infty$ test, we generated 1000 
+pre- and post-deliberation ordinal datasets.
+
+We generated these simulated data using the latent mean, pre-deliberation 
+standard deviation, and post-deliberation standard deviations as parameters of
+normal distributions of pre- and post-deliberation opinions.
+
+\subsection{Bayesian ordered-probit models inference of Cohen's $d$ for simulated
+simple-consensus data}
+
+Non-metric ordered probit models, which account for the data collection 
+process, theoretically should be better able to correctly
+distinguish simple consensus compared to metric models. Therefore, we also fit
+the following ordered probit model to simulated group polarization data.
 
@@
 
+\vspace{.5em}
 \spacing{1.0}
-\input{plausibleFP-table.tex}
+% \input{plausibleFP-table.tex}
+\begin{table}[H]
+    \begin{center}
+      \caption{Number and fraction of experimental conditions where group polarization was
+               reported to have occurred, but plausibly did not.}
+      \vspace{.5em}
+      \input{plausibleFP-table-custom.tex}
+    \end{center}
+  \end{table}
 \spacing{1.5}
 
-\subsection{Do $t$-tests accurately detect simple consensus?}
+\subsection{Metric linear models often fail to identify simple consensus}
 
-We found that in X\% of generated simple consensus data led to false positive
-detections
+Here we evaluate if metric linear models inferred group polarization from simulated
+simple consensus data using parameters identified above. Those parameters specified
+continous pre- and post-deliberation means and variance of normal opinion
+distributions that represent simple consensus but yields
+pre- and post-deliberation ordinal distributions with different, group-polarized
+means. This parameter-finding process implicitly assumes $N\to\infty$ participants.
 
-When simulated data reliably result in false positive $t$-test results (where
-the $t$-test $p \le 0.1$) this strengthens the case that reported detections
-of a false group polarization detection.
+It is therefore necessary to evaluate whether the simple consensus parameters so 
+identified in the previous step yield simulated data that also appear to be
+group-polarized when measured on an ordinal scale. We draw $N$ samples from the
+pre- and post-deliberation normal distributions and bin them into one of 
+$B$ ordinal values. $N$ and $B$ are set specifically for each of the 
+60 experimental conditions analyzed here. We then fit a linear model to the 
+simulated observed pre- and post-deliberation data to infer whether the mean
+shifted, indicating group polarization. We measure the pre- and
+post-deliberation standard deviations as well. Then we calculate Cohen's $d$
+to quantify the effect size for each simulation. A suite of simulations are
+necessary since there might be significant variance in $d$ for different
+generations of simulated pre- and post-deliberation observations. Therefore we
+perform 1000 trials for each experimental condition and inspect the resulting
+distributions of $d$ across trials in a boxplot.
 
-\spacing{1.0}
-% \input{t-test-table.tex}
-\input{t-test-table-custom.tex}
-\spacing{1.5}
+\vspace{0.5em}
+\begin{center}
+{[ FIGURE \ref{fig:MetricBoxplot_N} ABOUT HERE ]} \\
+\end{center}
 
-\subsection{Do Bayesian ordered-probit models accurately detect simple consensus?}
+\subsection{Ordered-probit models also fail?}
 
@@
 positives, not anything intrinsic to the phenomenon of group polarization.
+
+\subsection{More data enables ordered probit models, but not metric models, to reliably
+  identify simple consensus}
+
+  \mt{Idea: increase $N$ for simulations and see effect on distribution of Cohen's $d$
+  for both metric and ordinal models}
 
@@
 
+We did make some simplifying assumptions that could affect model outcomes.
+
 This problem is made all the worse by the lack of a consistent experimental design for group
 polarization experiments. This makes it impossible to systematically account for
-additional sources of vairance known to lead to (orthogonal) over-estimates
-of effect sizes~\cite{Yarkoni2021}. This is due to the rush to explain why 
-group polarization occurs using weak theory~\cite{Cartwright1971,Cartwright1973} and makes
-group polarization meta-analyses (e.g. Brown, 1986; and Isenberg, 1986)\nocite{Brown1986,Isenberg1986}
-unreliable~\cite{Meehl1990,Meehl1997}.
+additional sources of vairance known to lead to over-estimates
+of effect sizes~\cite{Yarkoni2021}, independent of the false positives problem
+we identified in the current paper.
+
+% This is due to the rush to explain why 
+% group polarization occurs using weak theory~\cite{Cartwright1971,Cartwright1973} and makes
+% group polarization meta-analyses (e.g. Brown, 1986; and Isenberg, 1986)\nocite{Brown1986,Isenberg1986}
+% unreliable~\cite{Meehl1990,Meehl1997}.
 

### 2024-04-09 GroupPolarizationStatmod/paper/_extensions/partials/title.tex
init quarto paper rewrite

\usepackage{authblk}
\usepackage{etoolbox}
\makeatletter
\title{$title$}
\makeatother
\author[1,*]{{Matthew A.~Turner}}
\affil[1]{\small Division of Social Sciences, Stanford Doerr School of Sustainability, Stanford University}

\author[2,3]{{Paul E.~Smaldino}}
\affil[2]{\small Cognitive and Information Sciences, University of California, Merced}
\affil[3]{\small Santa Fe Institute} 
% \vspace{2em}
\affil[*]{\small Correspondence: \href{mailto:maturner@stanford.edu}{maturner@stanford.edu}}

\date{}

### 2024-04-09 GroupPolarizationStatmod/paper/ms.qmd
init quarto paper rewrite

@@
     pdf-engine-opt: -xelatex
-    # template-partials:
-    #   - _extensions/partials/title.tex
+    keep-tex: true
+    template-partials:
+      - _extensions/partials/title.tex
 authors:
   - name: Matthew Adam Turner
-    institute: Stanford University Doerr School of Sustainability, Stanford University
+    affiliation: Stanford University Doerr School of Sustainability, Stanford University
     roles: writing, model development, project development
@@
   - name: Paul E. Smaldino
-    institute: 
+    affiliation: 
       - Cognitive and Information Sciences, Univeristy of California, Merced

### 2024-04-11 GroupPolarizationStatmod/paper/ms.qmd
Updated abstract and title, stub intro

--- 2024-04-09 GroupPolarizationStatmod/paper/ms.qmd
@@
 ---
-title: "Prominent claims of group polarization are plausibly false with lessons for studying psychological dynamics"
+title: "Foundational claims of group polarization are plausibly false, with lessons for studying psychological dynamics"
 format:
@@
 abstract: |
-  This is the abstract
+  \noindent _Group polarization_ is the accepted name for a form of consensus where members of a like-minded group become, on average, more extreme in their opinions after discussing a topic.
+Group polarization is important because it may drive broader polarization if relatively moderate groups become more extreme.
+Group polarization is celebrated as a regular, reliable outcome of a certain experimental design based on dozens of published studies that purportedly demonstrate the phenomena.
+However, in this paper we show that 55 of 60 published observations of purported group polarization, from ten foundational studies, are equally well-explained by a simple consensus model where group opinion variance decreases, but extremity does not increase.
+We show this through a novel computational meta-analysis framework designed to identify plausible pre- and post-discussion opinion distributions that are consistent with simple consensus, but appear to be group polarization when measured on ordinal, Likert-style scales.
+This occurs due to ceiling effects induced by ordinal measurements.
+At best this shows that a selection of foundational findings in social psychology must be discarded.
+At worst this is the tip of the iceberg, and many similar social psychological findings within and without the group polarization tradition will similarly be shown to be unreliable.
 geometry:
@@
 
-## Quarto
-
-Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.
-
-## Running Code
-
-When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:
-
-```{r}
-1 + 1
-```
-
-You can add options to executable code like this
-
-```{r}
-#| echo: false
-2 * 2
-```
-
-The `echo: false` option disables the printing of code (only output is displayed).
+## Introduction

### 2024-04-11 GroupPolarizationStatmod/paper/ms.tex
init quarto paper rewrite
Updated abstract and title, stub intro

@@
 \documentclass[
-  letterpaper,
-  abstract]{scrartcl}
+  abstract]{article}
 
@@
 
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{241,243,245}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
-\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
-
 \providecommand{\tightlist}{%
@@
 
-\KOMAoption{captions}{tableheading}
 \makeatletter
@@
 \hypersetup{
-  pdftitle={Prominent claims of group polarization are plausibly false with lessons for studying psychological dynamics},
+  pdftitle={Foundational claims of group polarization are plausibly false with lessons for studying psychological dynamics},
   colorlinks=true,
@@
 
-\title{Prominent claims of group polarization are plausibly false with
-lessons for studying psychological dynamics}
-\author{Matthew Adam Turner \and Paul E. Smaldino}
+\usepackage{authblk}
+\usepackage{etoolbox}
+\makeatletter
+\title{Foundational claims of group polarization are plausibly false
+with lessons for studying psychological dynamics}
+\makeatother
+\author[1,*]{{Matthew A.~Turner}}
+\affil[1]{\small Division of Social Sciences, Stanford Doerr School of Sustainability, Stanford University}
+
+\author[2,3]{{Paul E.~Smaldino}}
+\affil[2]{\small Cognitive and Information Sciences, University of California, Merced}
+\affil[3]{\small Santa Fe Institute} 
+% \vspace{2em}
+\affil[*]{\small Correspondence: \href{mailto:maturner@stanford.edu}{maturner@stanford.edu}}
+
 \date{}
-
 \begin{document}
@@
 \begin{abstract}
-This is the abstract
+\noindent Group polarization is the accepted name for a form of
+consensus where members of a like-minded group become, on average, more
+extreme in their opinions after discussing a topic. Group polarization
+is important because it may drive broader polarization if relatively
+moderate groups become more extreme. Group polarization is celebrated as
+a regular, reliable outcome of a certain experimental design based on
+dozens of published studies that purportedly demonstrate the phenomena.
+However, in this paper we show that 55 of 60 published observations of
+purported group polarization, from ten foundational studies, are equally
+well-explained by a simple consensus model where group opinion variance
+decreases, but extremity does not increase. We show this through a novel
+meta-analytical framework designed to identify plausible pre- and
+post-discussion opinion distributions that are consistent with simple
+consensus, but appear to be group polarization when measured on ordinal,
+Likert-style scales. This occurs due to ceiling effects induced by
+ordinal measurements. At best this shows that a selection of
+foundational findings in social psychology must be discarded. At worst
+this is the tip of the iceberg, and many similar social psychological
+findings within and without the group polarization tradition will
+similarly be shown to be unreliable.
 \end{abstract}
@@
 }
-\section{Quarto}\label{quarto}
-
-Quarto enables you to weave together content and executable code into a
-finished document. To learn more about Quarto see
-\url{https://quarto.org}.
-
-\section{Running Code}\label{running-code}
-
-When you click the \textbf{Render} button a document will be generated
-that includes both content and the output of embedded code. You can
-embed code like this:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\DecValTok{1} \SpecialCharTok{+} \DecValTok{1}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-[1] 2
-\end{verbatim}
-
-You can add options to executable code like this
-
-\begin{verbatim}
-[1] 4
-\end{verbatim}
-
-The \texttt{echo:\ false} option disables the printing of code (only
-output is displayed).
 

### 2024-04-11 GroupPolarizationStatmod/paper/ms.qmd
Full intro outline and two intro paragraphs; paper and Disc outline

@@
 ---
-title: "Foundational claims of group polarization are plausibly false, with lessons for studying psychological dynamics"
+title: "Foundational claims of group polarization are plausibly false, with lessons for the empirical study of psychological dynamics"
 format:
@@
     corresponding: true
-bibliography: ../library.bib
+bibliography: "/Users/mt/workspace/Writing/library.bib"
 classoption: abstract
@@
 abstract: |
-  \noindent _Group polarization_ is the accepted name for a form of consensus where members of a like-minded group become, on average, more extreme in their opinions after discussing a topic.
-Group polarization is important because it may drive broader polarization if relatively moderate groups become more extreme.
-Group polarization is celebrated as a regular, reliable outcome of a certain experimental design based on dozens of published studies that purportedly demonstrate the phenomena.
+  \noindent _Group polarization_ is the name for a form of consensus where members of a like-minded group become, on average, more extreme in their opinions after discussing a topic.
+Group polarization is important because it may drive political polarization if moderate partisans become more extreme.
+Theoretically, group polarization is celebrated as a regular, reliable outcome of a certain experimental design based on dozens of published studies that purportedly demonstrate the phenomena.
 However, in this paper we show that 55 of 60 published observations of purported group polarization, from ten foundational studies, are equally well-explained by a simple consensus model where group opinion variance decreases, but extremity does not increase.
-We show this through a novel computational meta-analysis framework designed to identify plausible pre- and post-discussion opinion distributions that are consistent with simple consensus, but appear to be group polarization when measured on ordinal, Likert-style scales.
+We use a novel computational meta-analysis framework designed to identify plausible pre- and post-discussion opinion distributions that are consistent with simple consensus, but appear to be group polarization when measured on ordinal, Likert-style scales.
 This occurs due to ceiling effects induced by ordinal measurements.
-At best this shows that a selection of foundational findings in social psychology must be discarded.
+At best this shows that a selection of foundational research in social psychology must be discarded.
 At worst this is the tip of the iceberg, and many similar social psychological findings within and without the group polarization tradition will similarly be shown to be unreliable.
@@
 
+```{=tex}
+\begin{quote}
+In our introductory social psychology course, 
+we have for many years used the [group polarization experimental paradigm] as
+a laboratory exercise. The exercise works beautifully, but one must be
+careful to forewarn a class that [group polarization] does not occur with every 
+group\ldots and that the effect is not large. 
+\par\raggedleft(Brown, 1986, p.\cite[p. 205]{Brown1986})
+\end{quote}
+```
+```{=tex}
+\begin{quote}
+One of the most robust findings in social psychology is that of attitude polarization 
+following discussion with like-minded others.
+\par\raggedleft(Cooper et al, 2001, p. 267 \cite[p. 267]{Cooper2001})
+\end{quote}
+```
 ## Introduction
+
+*Group polarization* is the name for a group-level process that leads to greater opinion extremity in a group as the group deliberates some topic (e.g., "Do you approve, disapprove, or neither approve nor disapprove of the current administration's policies?").
+Extremity only increases, though, assuming that the mean opinion of the group is initially biased towards one extreme or the other [@Brown1986; @Brown2020].
+Theoretically, group polarization occurs as groups find consensus, meaning that the variance in group opinion decreases [@DeGroot1974].
+Group polarization is important to understand rigorously, therefore, since if groups become increasingly extreme over time, while group opinions are consolidated, this would exacerbate political polarization between political opposition groups, which is already undermines the ability of partisans in democratic governments to work together [@Mason2018UncivilAgreementBook; @Klein2020].
+Dozens of studies and meta-analyses over decades have apparently established the empirical reality of group polarization \[@Moscovici1969; @Myers1970; @Isenberg1986; @Sunstein2009; @Sunstein2019\] [@Brown2000, Ch.
+5].
+However, many or most of the evidence for group polarization relies on experimental designs that use ordinal measurements that are then analyzed as if they were continuous.
+Unfortunately, this procedure is unfortunately known to induce false positives if it is not properly accounted for [@Liddell2018].
+
+In this paper, we perform a meta-analysis of sixty detections of group polarization from ten published research articles and show that fifty-five of the sixty cannot be reliably identified as group polarization.
+We show that these 55 detections are equally well-explained by a *simple consensus* model, where the mean group opinion is unchanged by group discussion but opinion variance decreases.
+We use a novel approach, described herein, that enabled us to identify continuous pre- and post-discussion simple consensus opinion distributions that, when measured with an ordinal scale, result in data that appear to be group-polarized.
+(How this procedure works in detail: continuous case to identify means, then More broadly, this approach could be used in similar studies that used ordinal measures to determine whether change occurred in distributions of continuous psychological variables.
+For social psychology and group polarization, this means that an empirical fact must be re-evaluated and new experimental procedures established to rigorously understand whether and when group polarization occurs.
+
+More about group polarization history and evidence; why these ten studies and how they represent prominent theories (see Rupert Brown book, Isenberg review; Sunstein highly prominent and elevated GP in trade books)
+
+NEXT =\> More about the logic of this study, why showing that these results are plausibly false strictly means they are not rigorously true.
+
+RESEARCH OVERVIEW
+
+## Methods
+
+## Results
+
+## Discussion
+
+Ordinal scales are a useful experimental
+
+Replication crisis, generalizability crisis [@Yarkoni2022], and general lack of coordination in asking the right questions to efficiently understand the mechanisms behind .
+Understanding mechanisms is the goal of science [@Machamer2000; @Craver2006].
+However, in social and psychological sciences, finding mechanisms is more difficult due to the complexity of social psychological systems [@Dubova2023], which require a high density of conceptual definitions that may be difficult to harmonize.

### 2024-04-13 GroupPolarizationStatmod/paper/ms.qmd
expanded plan for next two intro paragraphs to do

@@
 Group polarization is important to understand rigorously, therefore, since if groups become increasingly extreme over time, while group opinions are consolidated, this would exacerbate political polarization between political opposition groups, which is already undermines the ability of partisans in democratic governments to work together [@Mason2018UncivilAgreementBook; @Klein2020].
-Dozens of studies and meta-analyses over decades have apparently established the empirical reality of group polarization \[@Moscovici1969; @Myers1970; @Isenberg1986; @Sunstein2009; @Sunstein2019\] [@Brown2000, Ch.
+Dozens of studies and meta-analyses over decades have apparently established the empirical reality of group polarization [@Moscovici1969; @Myers1970; @Isenberg1986; @Sunstein2009; @Sunstein2019] [@Brown2000, Ch.
 5].
@@
 In this paper, we perform a meta-analysis of sixty detections of group polarization from ten published research articles and show that fifty-five of the sixty cannot be reliably identified as group polarization.
-We show that these 55 detections are equally well-explained by a *simple consensus* model, where the mean group opinion is unchanged by group discussion but opinion variance decreases.
+We show that these 55 are plausibly false detections because they are equally well-explained by a *simple consensus* model, where the mean group opinion is unchanged by group discussion but opinion variance decreases.
 We use a novel approach, described herein, that enabled us to identify continuous pre- and post-discussion simple consensus opinion distributions that, when measured with an ordinal scale, result in data that appear to be group-polarized.
@@
 
-NEXT =\> More about the logic of this study, why showing that these results are plausibly false strictly means they are not rigorously true.
+NEXT =\> More about the logic of this study, why showing that these results are plausibly false strictly means they are not rigorously true, overview of methods in somewhat more detail.
+Maybe need two paragraphs.
 

### 2024-04-13 GroupPolarizationStatmod/paper/ms.qmd
wrote paragraph explaining selection criteria

@@
 
-More about group polarization history and evidence; why these ten studies and how they represent prominent theories (see Rupert Brown book, Isenberg review; Sunstein highly prominent and elevated GP in trade books)
+In this study, we take the first step towards a new kind of meta-analysis needed to demonstrate that published detections of group polarization are plausibly false.
+We selected sixty foundational group polarization detections from ten studies.
+We chose the ten studies guided by a desire to identify foundational studies that contribute to the general consensus in social psychology that group polarization reliably occurs across experimental settings—one author of one of the ten journal articles is a law professor who has written two trade books on group polarization to argue that group polarization occurs in real-world settings in addition to the laboratory, with political and other practical importance [@Sunstein2009; @Sunstein2019].
+Our criteria were that the studies were either (1) written by influential authors; (2) had an outsize influence on the study of group polarization in terms of citations and journal prestige; or (3) are well-cited studies that were designed to support one of four recognized theoretical explanations of group polarization.
+The studies must also have used an ordinal scale for measuring participant opinions (all but a handful of exceptions use ordinal measurements to measure group polarization).
+Follow-on studies could establish more objective standards for inclusion, but the above criteria have provided a useful selection of ten diverse studies as we explain in more detail now.
+
+The first prominent group polarization studies of broad scientific and social interest examined .
+In tracing the historical development of group polarization, four theoretical explanations have been identified that could possibly be compatible, with some practitioners considering them incompatible.
+Some influential studies we identify We have suggested elsewhere that a simpler explanation could go farther [@Turner2020], but here we are not interested in evaluating or comparing the these theories, but instead selecting influential studies from different recognized theoretical traditions for our selective meta-analysis here.
 

### 2024-04-16 GroupPolarizationStatmod/paper/ms.qmd
Intro outlined; nearly complete

@@
 keep-tex: true
+geometry:
+  - margin=1.1in
+  - letterpaper
+biblio-title: References
+indent: true
 abstract: |
@@
 At worst this is the tip of the iceberg, and many similar social psychological findings within and without the group polarization tradition will similarly be shown to be unreliable.
-geometry:
-  - margin=1.1in
-  - letterpaper
-biblio-title: References
 ---
@@
 
+Unchecked social and political polarization undermines cooperation and political efficacy [@Mason2018UncivilAgreementBook; @Klein2020].
 *Group polarization* is the name for a group-level process that leads to greater opinion extremity in a group as the group deliberates some topic (e.g., "Do you approve, disapprove, or neither approve nor disapprove of the current administration's policies?").
 Extremity only increases, though, assuming that the mean opinion of the group is initially biased towards one extreme or the other [@Brown1986; @Brown2020].
-Theoretically, group polarization occurs as groups find consensus, meaning that the variance in group opinion decreases [@DeGroot1974].
-Group polarization is important to understand rigorously, therefore, since if groups become increasingly extreme over time, while group opinions are consolidated, this would exacerbate political polarization between political opposition groups, which is already undermines the ability of partisans in democratic governments to work together [@Mason2018UncivilAgreementBook; @Klein2020].
+Group polarization is important to understand rigorously, therefore, since if groups become increasingly extreme over time, while group opinions are consolidated, this would exacerbate political polarization between political opposition groups, which is already undermines the ability of partisans in democratic governments to work together.
 Dozens of studies and meta-analyses over decades have apparently established the empirical reality of group polarization [@Moscovici1969; @Myers1970; @Isenberg1986; @Sunstein2009; @Sunstein2019] [@Brown2000, Ch.
 5].
+We believe group polarization should occur in certain contexts since more extreme group members are likely also to be more stubborn [@Guazzini2015; @Lewandowsky2019], which should pull consensus to extremes [@Turner2020].
 However, many or most of the evidence for group polarization relies on experimental designs that use ordinal measurements that are then analyzed as if they were continuous.
@@
 We show that these 55 are plausibly false detections because they are equally well-explained by a *simple consensus* model, where the mean group opinion is unchanged by group discussion but opinion variance decreases.
-We use a novel approach, described herein, that enabled us to identify continuous pre- and post-discussion simple consensus opinion distributions that, when measured with an ordinal scale, result in data that appear to be group-polarized.
-(How this procedure works in detail: continuous case to identify means, then More broadly, this approach could be used in similar studies that used ordinal measures to determine whether change occurred in distributions of continuous psychological variables.
-For social psychology and group polarization, this means that an empirical fact must be re-evaluated and new experimental procedures established to rigorously understand whether and when group polarization occurs.
+Theoretically, consensus occurs when sufficiently like-minded individuals influence, which decreases group opinion variance [@DeGroot1974; @Turner2018].
+Group polarization, then, is consensus plus an overall increase in group opinion extremity, i.e., an increase in mean group opinion after discussion.
+We use a novel approach, described herein, that enabled us to identify continuous pre- and post-discussion simple consensus opinion distributions that result in data that appear to be group-polarized when opinions are drawn from these distributions and measured with an ordinal scale.
+This approach could be adapted for similar social psychological studies that used ordinal measures to determine whether change occurred in distributions of continuous psychological variables such as opinions and beliefs.
+For social psychology and group polarization, this means that an accepted, heretofore unquestioned, empirical fact must be re-evaluated.
+Experimental designs in the field must be updated if we are to rigorously understand whether and when group polarization occurs.
 
@@
 Our criteria were that the studies were either (1) written by influential authors; (2) had an outsize influence on the study of group polarization in terms of citations and journal prestige; or (3) are well-cited studies that were designed to support one of four recognized theoretical explanations of group polarization.
-The studies must also have used an ordinal scale for measuring participant opinions (all but a handful of exceptions use ordinal measurements to measure group polarization).
-Follow-on studies could establish more objective standards for inclusion, but the above criteria have provided a useful selection of ten diverse studies as we explain in more detail now.
+The studies must also have used an ordinal scale for measuring participant opinions (all but a handful of exceptions use ordinal measurements to measure group polarization; e.g.
+blackjack bets [@Blascovich1973; @Blascovich1976] and mock jury monetary damages [@Schkade2000]).
+Follow-on studies could establish more objective standards for inclusion, but the above criteria have provided a useful selection of ten diverse studies.
 
-The first prominent group polarization studies of broad scientific and social interest examined .
-In tracing the historical development of group polarization, four theoretical explanations have been identified that could possibly be compatible, with some practitioners considering them incompatible.
-Some influential studies we identify We have suggested elsewhere that a simpler explanation could go farther [@Turner2020], but here we are not interested in evaluating or comparing the these theories, but instead selecting influential studies from different recognized theoretical traditions for our selective meta-analysis here.
+The first prominent group polarization studies of broad scientific interest examined social and political issues without much theoretical development, i.e., lacking specific predictions or explanations.
+Of these, we include two journal articles, Moscovici and Zavalloni (1969)[@Moscovici1969] who analyzed responses from high school students about domestic and foreign politics and Myers (1970) [@Myers1970] who .
+The first article is a study of French high school students responding to questions about .
+The second is \
+since group polarization was In tracing the historical development of group polarization, four theoretical explanations have been identified that could possibly be compatible, with some practitioners considering them incompatible.
+Some influential studies we identify Here we are not interested in evaluating or comparing the these theories, but instead we select influential studies from different theoretical traditions for our selective meta-analysis here.
 
-NEXT =\> More about the logic of this study, why showing that these results are plausibly false strictly means they are not rigorously true, overview of methods in somewhat more detail.
-Maybe need two paragraphs.
+Each of the ten group polarization studies/journal articles we analyze varied significantly in how groups were formed, how opinions were gathered, and other details. Some
 
-RESEARCH OVERVIEW
+To evaluate whether published detections of group polarization are plausibly false we developed and combined formal and generative models of group polarization to identify plausible simple consensus pre- and post-discussion opinion distributions that seem like group polarization when binned or measured on ordinal scales.
+Opinion distributions are assumed to be normal.
+We define a *simple consensus pair* of *latent opinion* distributions to be a pair of pre- and post-discussion distributions where the mean opinion is constant, but the pre-discussion variance is greater than the post-discussion variance.
+A *latent distribution* of opinions represents the internal, unobservable distribution of opinions for a group or collection of groups.
+The *observed distribution* is the distribution of reported ordinal opinion measurements.
+We then define a *plausibly false detection* of group polarization to be one where the mean post-discussion observed distribution is greater in magnitude (i.e., more extreme) than the post-discussion observed distribution's, but both were measurements of a pair of latent simple consensus distributions.
+
+In the formal model, a simulated observed distribution is generated by binning a latent distribution into a finite number of ordinal scale bins, then integrating over the probability density in each bin.
+In the generative model, an observed distribution is simulated by first drawing participant opinions from a latent distribution, then counting the number of opinions falling within each bin.
+The formal model is used to identify which, if any, simple consensus pairs exist for a given reported detection of group polarization through an optimization routine that finds pre- and post-discussion variances that generate simulated observations that appear to be group polarization (or reports a failure to such variances).
+If the optimization routine finds a constant latent mean that generates simulated observed group polarization, then we say the published group polarization detection is plausibly false.
+If the routine fails to find a constant latent mean that generates simulated group polarization, then we cannot and do not claim the detection is plausibly false.
+We use the generative model to evaluate how frequently we would expect plausibly false detections when there are finite-size groups, since the formal model implicitly applies infinite-size groups.
+Therefore in our analyses, we also generate one thousand simulated group polarization datasets for each of the 55 plausibly false detections identified with the formal model and examine the distribution of group polarization effect sizes in terms of Cohen's $d$.
+In the first round of simulations we set the number of observations to match those in the ten published studies we analyzed.
+Many or most simulated effect sizes are non-zero with these numbers of observations.
+To see if perhaps these non-zero effects were due to low statistical power from too few observations, we also tested a hypothetical study with ten times more participants.
+Finally, we evaluate whether an appropriate statistical model that accounts for the ordinal measurement of latent opinions can reliably identify simple consensus in latent opinions when observed opinions appear as group polarization.
 
@@
 
-Ordinal scales are a useful experimental
+Ordinal scales are an indispensable tool for psychological research on how social influence changes people's beliefs, opinions, knowledge, or similar psychological content.
+We show here that we must be careful not to mistake the measurement procedure as psychological reality when designing experiments.
 

### 2024-04-18 GroupPolarizationStatmod/paper/ms.qmd
More intro content and structure

@@
 indent: true
+suppress-bibliography: true
 abstract: |
@@
 
+Experimetnal designs in group polarization studies share a certain experimental boilerplate (Figure 1 — NEED TO INCLUDE) with the goal of testing the group polarization hypothesis that a group of like-minded people will develop more extreme opinions on some salient topic over the course of discussing some topic.
+Group polarization experiments are thus initialized by forming like-minded groups of people and surveying their initial opinions on some topic.
+In some cases these initial opinions are used for assigning participants to like-minded groups, but other methods have been used, such as geographic location for producing, on average, like-minded groups [@Schkade2010].
+Participants then discuss the topic in their groups.
+After discussion, participant opinions are again measured.
+Several studies we analyze here determined group polarization occurred (i.e., defined a positive detection) only as a non-zero difference in mean opinion, with no other statistical testing.
+Some others performed null-hypothesis $t$-tests to test whether there was enough evidence to reject the null-hypothesis that *simple consensus* occurredand not group polarization.
+In our meta-analysis simulations, we measure simulated group polarization effect size in terms of Cohen's $d$, which is an abstraction beyond these two practices from the group polarization literature that can also be used to compare model fits between metric models that treat group polarization measurements as continuous although they are ordinal, and more appropriate statistical models that account for data ordinality.
+
 In this paper, we perform a meta-analysis of sixty detections of group polarization from ten published research articles and show that fifty-five of the sixty cannot be reliably identified as group polarization.
@@
 
-In this study, we take the first step towards a new kind of meta-analysis needed to demonstrate that published detections of group polarization are plausibly false.
+<!-- In this study, we take the first step towards a new kind of meta-analysis needed to demonstrate that published detections of group polarization are plausibly false.
 We selected sixty foundational group polarization detections from ten studies.
+-->
+
+### Studies included in this meta-analysis
+
 We chose the ten studies guided by a desire to identify foundational studies that contribute to the general consensus in social psychology that group polarization reliably occurs across experimental settings—one author of one of the ten journal articles is a law professor who has written two trade books on group polarization to argue that group polarization occurs in real-world settings in addition to the laboratory, with political and other practical importance [@Sunstein2009; @Sunstein2019].
-Our criteria were that the studies were either (1) written by influential authors; (2) had an outsize influence on the study of group polarization in terms of citations and journal prestige; or (3) are well-cited studies that were designed to support one of four recognized theoretical explanations of group polarization.
+We chose studies that either had a significant influence on the study of group polarization and social psychology, in terms of a high citation count or adaptation in popularized trade books, or were designed to support one of four prominent theoretical explanations of group polarization.
 The studies must also have used an ordinal scale for measuring participant opinions (all but a handful of exceptions use ordinal measurements to measure group polarization; e.g.
 blackjack bets [@Blascovich1973; @Blascovich1976] and mock jury monetary damages [@Schkade2000]).
-Follow-on studies could establish more objective standards for inclusion, but the above criteria have provided a useful selection of ten diverse studies.
+We included at least one study that supported or was motivated by one of four established theoretical explanations (introduced below) of group polarization to show all have at least some supporting evidence that must be discarded.
+We also included two studies that lack a clear theoretical motivation.
+For the sake of orienting the reader to the group polarization literature and as a useful organizing feature for our analysis, we now introduce the theories that motivate eight of the ten studies.
+To further orient the reader, we then provide details of the experimental designs used to produce the 60 published positive detections of group polarization across 10 research articles that we scrutinize here.
+Group polarization experimental designs are diverse and inconsistent with one another in many ways, which complicates this or other potential meta-analyses and may imperil our ability to rigorously understand how regular the group polarization effect really is [@Yarkoni2022; @Almaatouq2022].
 
-The first prominent group polarization studies of broad scientific interest examined social and political issues without much theoretical development, i.e., lacking specific predictions or explanations.
-Of these, we include two journal articles, Moscovici and Zavalloni (1969)[@Moscovici1969] who analyzed responses from high school students about domestic and foreign politics and Myers (1970) [@Myers1970] who .
-The first article is a study of French high school students responding to questions about .
-The second is \
-since group polarization was In tracing the historical development of group polarization, four theoretical explanations have been identified that could possibly be compatible, with some practitioners considering them incompatible.
-Some influential studies we identify Here we are not interested in evaluating or comparing the these theories, but instead we select influential studies from different theoretical traditions for our selective meta-analysis here.
+The first theoretical explanation whose supporting findings we include here is (1) social comparisons theory, which posits that an individual's opinions are updated so they stand out in a group enough to be "desirably distinctive" (Myers, 1978, p.
+562) but not too distinct to be viewed as anti-conformist [@Brown1974; @Sanders1977; @Myers1978].
+Second, *persuasive arguments theory* posits that more extreme individuals have a greater number of more convincing arguments for their view, which draw more moderate group members to extreme opinions [@Burnstein1973; @Burnstein1975; @Burnstein1977].
+Third, the *self-categorization theory* predicts that group polarization will occur as more moderate individuals judge the mean group opinion to be more extreme [@Turner1989; @Abrams1990; @Krizan2007].
+Fourth and finally, the social decision schemes explanation posits that the structure of [@Davis1973; @Zuber1992; @Friedkin1999a].
+We do not evaluate or compare these theories here, though a simpler explanation that extremists are more stubborn may be a more parsimonious explanation consistent with these four alternatives [@Acemoglu2013; @Guazzini2015; @Turner2020].
 
-Each of the ten group polarization studies/journal articles we analyze varied significantly in how groups were formed, how opinions were gathered, and other details. Some
+Each of the ten group polarization studies we analyze here varied significantly in how groups were formed, how opinions were gathered, and other details.
+There was significant variation across experimental designs, participant population sizes, group sizes, and how like-minded groups were formed in the first place.
+We review some of these details below to provide the reader with some insight into the experimental designs that produced what we show are plausibly false positive detections of group polarization.
+We focus here on the questions that were asked, how like-minded novel experimental groups were formed, organizing our presentation by theoretical tradition.
+
+#### Details of experimental conditions included in this meta-analysis
+
+\paragraph{Atheoretical} The first of ten studies we included in our meta-analysis was by Moscovici and Zavalloni (1969) [@Moscovici1969] who sought to measure group polarization among groups of French high school students who discussed domestic and foreign politics.
+Moscovici and Zavalloni tested four conditions whether group polarization occurred for two different questions and two different rating systems: participants were asked their opinions on then-president Charles De Gaulle and their opinion about foreign aid from the USA either in the "Opinion" setting or the "Judgement" condition (see Table 4 on p.
+132).
+
+The second, relatively more recent article is Schkade, Sunstein, and Hastie (2010) [@Schkade2010] who used geographic location to construct like-minded groups by running a group polarization experiment in two cities in the state of Colorado in the United States: Colorado Springs, whose residents historically voted for republican candidates, and Boulder, whose residents historically voted for democratic candidates.
+Discussion groups were formed in which they measured group polarization when group members discussed global warming, affirmative action, and same-sex civil unions.
+
+\paragraph{Social comparisons} In developing the social comparisons theory of group polarization (described in Myers (1978) [@Myers1978]), Myers and Bishop (1970), writing in *Science,* [@Myers1970] sought to measure group polarization among high school students in the United States who discussed racial attitudes and Myers (1975) [@Myers1975] measured group polarization in groups of college students discussing the quality of a faculty member in one condition, and women's rights and issues in another condition.
+
+\paragraph{Persuasive arguments}
+
+\paragraph{Self-categorization}
+
+\paragraph{Social decisions scheme}
+
+The ten studies we included each tested several experimental conditions meant to elicit group polarization in novel experimental groups, 60 in total, which we briefly introduce here. \
+
+### Research overview
 
@@
 To see if perhaps these non-zero effects were due to low statistical power from too few observations, we also tested a hypothetical study with ten times more participants.
-Finally, we evaluate whether an appropriate statistical model that accounts for the ordinal measurement of latent opinions can reliably identify simple consensus in latent opinions when observed opinions appear as group polarization.
+Finally, we evaluate whether an appropriate statistical model that accounts for the ordinal measurement of latent opinions can reliably identify simulated simple consensus in latent opinions when simulated observations appear as group polarization based on mean group opinion alone.
 
@@
 However, in social and psychological sciences, finding mechanisms is more difficult due to the complexity of social psychological systems [@Dubova2023], which require a high density of conceptual definitions that may be difficult to harmonize.
+
+# Tables {.unnumbered}
+
+asdf

### 2024-05-02 GroupPolarizationStatmod/paper/ms.qmd
paper updates

@@
 
-\paragraph{Atheoretical} The first of ten studies we included in our meta-analysis was by Moscovici and Zavalloni (1969) [@Moscovici1969] who sought to measure group polarization among groups of French high school students who discussed domestic and foreign politics.
+\paragraph{Atheoretical}
+
+The first of ten studies we included in our meta-analysis was by Moscovici and Zavalloni (1969) [@Moscovici1969] who sought to measure group polarization among groups of French high school students who discussed domestic and foreign politics.
 Moscovici and Zavalloni tested four conditions whether group polarization occurred for two different questions and two different rating systems: participants were asked their opinions on then-president Charles De Gaulle and their opinion about foreign aid from the USA either in the "Opinion" setting or the "Judgement" condition (see Table 4 on p.
@@
 
-\paragraph{Social comparisons} In developing the social comparisons theory of group polarization (described in Myers (1978) [@Myers1978]), Myers and Bishop (1970), writing in *Science,* [@Myers1970] sought to measure group polarization among high school students in the United States who discussed racial attitudes and Myers (1975) [@Myers1975] measured group polarization in groups of college students discussing the quality of a faculty member in one condition, and women's rights and issues in another condition.
+\paragraph{Social comparisons}
+
+In developing the social comparisons theory of group polarization (described in Myers (1978) [@Myers1978]), Myers and Bishop (1970), writing in *Science,* [@Myers1970] sought to measure group polarization among high school students in the United States who discussed racial attitudes and Myers (1975) [@Myers1975] measured group polarization in groups of college students discussing the quality of a faculty member in one condition, and women's rights and issues in another condition.
 
@@
 
-The ten studies we included each tested several experimental conditions meant to elicit group polarization in novel experimental groups, 60 in total, which we briefly introduce here. \
+The ten studies we included each tested several experimental conditions meant to elicit group polarization in novel experimental groups, 60 in total, which we briefly introduce here.\
 

### 2024-10-29 GroupPolarizationStatmod/paper.bk/ms.tex
updates and started FDR module

@@
 This distribution has mean $\mu_t$ and standard deviation $\sigma_t$.
-This form is chosen to
+\mt{The following here goes in intro to justify model choices:} This form is chosen to
 reflect the common statistical analysis used by group polarization studies,

### 2024-10-29 GroupPolarizationStatmod/false_detection_rate.R

@@ -1,0 +1,55 @@
+##
+# 
+#
+
+load_fdr_data = function(probit_fits_dir = "data/probit_fits", 
+                         sync_file = "data/probit_fits/all.csv",
+                         overwrite = FALSE) {
+   
+  if (file_exists(sync_file) && !overwrite) {
+
+    ret <- dir_ls(ordinal_data_dir, glob = "*.csv") %>%
+      read_csv() %>% unite(ExperimentID, ArticleTag, TreatmentTag)
+
+    write_csv(df, sync_file)
+
+  } else {
+
+    ret <- read_csv(sync_file)
+  }
+
+  return (ret)
+}
+
+
+false_detection_rate_by_condition = function(probit_fits_dir = "data/probit_fits",
+                                             sigvals = seq(0.2, 2.0, 0.15)) {
+  
+  fdr_data = read_csv(load_fdr_data(probit_fits_dir)) 
+
+  unique_treatments = unique(fdr_data$TreatmentTag)
+
+  n_plausible = length(unique_treatments)
+
+  Treatment = unlist(
+    map(unique_treatments, \(treatment) rep(treatment, sigvals))
+  )
+
+  alpha_v_sigval = tibble(Alpha = rep(NA, n_plausible*length(sigvals)),
+                          Sigval = rep(sigvals, n_plausible),
+                          Treatment = Treatment)
+
+  for (this_sigval in sigvals) {
+    gb = load_fdr_data(fdr_data_file) %>%
+           select(Sigval == this_sigval) %>%
+           group_by(TreatmentTag)
+
+    # Calculate alpha from this gb and put in alpha_v_sigval.
+    # Or somehow need to save one for each treatment or sig val then
+    # rbind all the results
+    
+  }
+         
+
+}
+

### 2024-10-29 GroupPolarizationStatmod/figures.R
figures and paper to tex

@@ -1,5 +1,7 @@
 library(ggplot2)
 
+source("model.R")  # for binProb function used in ordinal plot function below.
+
 plot_latent_pdf_integration = function(mu = 0, sd = 2.5, xlim = c(-6.0, 6.0), 
                                        bin1 = c(-0.5, 0.5), bin2 = c(1.5, 5.0),
                                        save_path = "test_latent_pdf.pdf") {

### 2024-10-29 GroupPolarizationStatmod/paper/ms.tex

@@
 }{}
-\makeatletter
-\@ifundefined{KOMAClassName}{% if non-KOMA class
-  \IfFileExists{parskip.sty}{%
-    \usepackage{parskip}
-  }{% else
-    \setlength{\parindent}{0pt}
-    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
-}{% if KOMA class
-  \KOMAoptions{parskip=half}}
-\makeatother
 \usepackage{xcolor}
@@
 \usepackage[sorting=none, style=numeric-comp]{biblatex}
-\addbibresource{../library.bib}
+\addbibresource{/Users/mt/workspace/Writing/library.bib}
 \usepackage{bookmark}
@@
 \hypersetup{
-  pdftitle={Foundational claims of group polarization are plausibly false with lessons for studying psychological dynamics},
+  pdftitle={Foundational claims of group polarization are plausibly false, with lessons for the empirical study of psychological dynamics},
   colorlinks=true,
@@
 \makeatletter
-\title{Foundational claims of group polarization are plausibly false
-with lessons for studying psychological dynamics}
+\title{Foundational claims of group polarization are plausibly false,
+with lessons for the empirical study of psychological dynamics}
 \makeatother
@@
 \begin{abstract}
-\noindent Group polarization is the accepted name for a form of
-consensus where members of a like-minded group become, on average, more
-extreme in their opinions after discussing a topic. Group polarization
-is important because it may drive broader polarization if relatively
-moderate groups become more extreme. Group polarization is celebrated as
-a regular, reliable outcome of a certain experimental design based on
+\noindent \emph{Group polarization} is the name for a form of consensus
+where members of a like-minded group become, on average, more extreme in
+their opinions after discussing a topic. Group polarization is important
+because it may drive political polarization if moderate partisans become
+more extreme. Theoretically, group polarization is celebrated as a
+regular, reliable outcome of a certain experimental design based on
 dozens of published studies that purportedly demonstrate the phenomena.
@@
 well-explained by a simple consensus model where group opinion variance
-decreases, but extremity does not increase. We show this through a novel
-meta-analytical framework designed to identify plausible pre- and
+decreases, but extremity does not increase. We use a novel computational
+meta-analysis framework designed to identify plausible pre- and
 post-discussion opinion distributions that are consistent with simple
@@
 ordinal measurements. At best this shows that a selection of
-foundational findings in social psychology must be discarded. At worst
+foundational research in social psychology must be discarded. At worst
 this is the tip of the iceberg, and many similar social psychological
@@
 }
+\begin{quote}
+In our introductory social psychology course, 
+we have for many years used the [group polarization experimental paradigm] as
+a laboratory exercise. The exercise works beautifully, but one must be
+careful to forewarn a class that [group polarization] does not occur with every 
+group\ldots and that the effect is not large. 
+\par\raggedleft(Brown, 1986, p.\cite[p. 205]{Brown1986})
+\end{quote}
+
+\begin{quote}
+One of the most robust findings in social psychology is that of attitude polarization 
+following discussion with like-minded others.
+\par\raggedleft(Cooper et al, 2001, p. 267 \cite[p. 267]{Cooper2001})
+\end{quote}
+
+\section{Introduction}\label{introduction}
+
+Unchecked social and political polarization undermines cooperation and
+political efficacy \autocite{Mason2018UncivilAgreementBook,Klein2020}.
+\emph{Group polarization} is the name for a group-level process that
+leads to greater opinion extremity in a group as the group deliberates
+some topic (e.g., ``Do you approve, disapprove, or neither approve nor
+disapprove of the current administration's policies?''). Extremity only
+increases, though, assuming that the mean opinion of the group is
+initially biased towards one extreme or the other
+\autocite{Brown1986,Brown2020}. Group polarization is important to
+understand rigorously, therefore, since if groups become increasingly
+extreme over time, while group opinions are consolidated, this would
+exacerbate political polarization between political opposition groups,
+which is already undermines the ability of partisans in democratic
+governments to work together. Dozens of studies and meta-analyses over
+decades have apparently established the empirical reality of group
+polarization
+\autocite{Moscovici1969,Myers1970,Isenberg1986,Sunstein2009,Sunstein2019}
+\autocite[Ch. 5]{Brown2000}. We believe group polarization should occur
+in certain contexts since more extreme group members are likely also to
+be more stubborn \autocite{Guazzini2015,Lewandowsky2019}, which should
+pull consensus to extremes \autocite{Turner2020}. However, many or most
+of the evidence for group polarization relies on experimental designs
+that use ordinal measurements that are then analyzed as if they were
+continuous. Unfortunately, this procedure is unfortunately known to
+induce false positives if it is not properly accounted for
+\autocite{Liddell2018}.
+
+Experimetnal designs in group polarization studies share a certain
+experimental boilerplate (Figure 1 --- NEED TO INCLUDE) with the goal of
+testing the group polarization hypothesis that a group of like-minded
+people will develop more extreme opinions on some salient topic over the
+course of discussing some topic. Group polarization experiments are thus
+initialized by forming like-minded groups of people and surveying their
+initial opinions on some topic. In some cases these initial opinions are
+used for assigning participants to like-minded groups, but other methods
+have been used, such as geographic location for producing, on average,
+like-minded groups \autocite{Schkade2010}. Participants then discuss the
+topic in their groups. After discussion, participant opinions are again
+measured. Several studies we analyze here determined group polarization
+occurred (i.e., defined a positive detection) only as a non-zero
+difference in mean opinion, with no other statistical testing. Some
+others performed null-hypothesis \(t\)-tests to test whether there was
+enough evidence to reject the null-hypothesis that \emph{simple
+consensus} occurredand not group polarization. In our meta-analysis
+simulations, we measure simulated group polarization effect size in
+terms of Cohen's \(d\), which is an abstraction beyond these two
+practices from the group polarization literature that can also be used
+to compare model fits between metric models that treat group
+polarization measurements as continuous although they are ordinal, and
+more appropriate statistical models that account for data ordinality.
+
+In this paper, we perform a meta-analysis of sixty detections of group
+polarization from ten published research articles and show that
+fifty-five of the sixty cannot be reliably identified as group
+polarization. We show that these 55 are plausibly false detections
+because they are equally well-explained by a \emph{simple consensus}
+model, where the mean group opinion is unchanged by group discussion but
+opinion variance decreases. Theoretically, consensus occurs when
+sufficiently like-minded individuals influence, which decreases group
+opinion variance \autocite{DeGroot1974,Turner2018}. Group polarization,
+then, is consensus plus an overall increase in group opinion extremity,
+i.e., an increase in mean group opinion after discussion. We use a novel
+approach, described herein, that enabled us to identify continuous pre-
+and post-discussion simple consensus opinion distributions that result
+in data that appear to be group-polarized when opinions are drawn from
+these distributions and measured with an ordinal scale. This approach
+could be adapted for similar social psychological studies that used
+ordinal measures to determine whether change occurred in distributions
+of continuous psychological variables such as opinions and beliefs. For
+social psychology and group polarization, this means that an accepted,
+heretofore unquestioned, empirical fact must be re-evaluated.
+Experimental designs in the field must be updated if we are to
+rigorously understand whether and when group polarization occurs.
+
+\subsection{Studies included in this
+meta-analysis}\label{studies-included-in-this-meta-analysis}
+
+We chose the ten studies guided by a desire to identify foundational
+studies that contribute to the general consensus in social psychology
+that group polarization reliably occurs across experimental
+settings---one author of one of the ten journal articles is a law
+professor who has written two trade books on group polarization to argue
+that group polarization occurs in real-world settings in addition to the
+laboratory, with political and other practical importance
+\autocite{Sunstein2009,Sunstein2019}. We chose studies that either had a
+significant influence on the study of group polarization and social
+psychology, in terms of a high citation count or adaptation in
+popularized trade books, or were designed to support one of four
+prominent theoretical explanations of group polarization. The studies
+must also have used an ordinal scale for measuring participant opinions
+(all but a handful of exceptions use ordinal measurements to measure
+group polarization; e.g.~blackjack bets
+\autocite{Blascovich1973,Blascovich1976} and mock jury monetary damages
+\autocite{Schkade2000}). We included at least one study that supported
+or was motivated by one of four established theoretical explanations
+(introduced below) of group polarization to show all have at least some
+supporting evidence that must be discarded. We also included two studies
+that lack a clear theoretical motivation. For the sake of orienting the
+reader to the group polarization literature and as a useful organizing
+feature for our analysis, we now introduce the theories that motivate
+eight of the ten studies. To further orient the reader, we then provide
+details of the experimental designs used to produce the 60 published
+positive detections of group polarization across 10 research articles
+that we scrutinize here. Group polarization experimental designs are
+diverse and inconsistent with one another in many ways, which
+complicates this or other potential meta-analyses and may imperil our
+ability to rigorously understand how regular the group polarization
+effect really is \autocite{Yarkoni2022,Almaatouq2022}.
+
+The first theoretical explanation whose supporting findings we include
+here is (1) social comparisons theory, which posits that an individual's
+opinions are updated so they stand out in a group enough to be
+``desirably distinctive'' (Myers, 1978, p.~562) but not too distinct to
+be viewed as anti-conformist \autocite{Brown1974,Sanders1977,Myers1978}.
+Second, \emph{persuasive arguments theory} posits that more extreme
+individuals have a greater number of more convincing arguments for their
+view, which draw more moderate group members to extreme opinions
+\autocite{Burnstein1973,Burnstein1975,Burnstein1977}. Third, the
+\emph{self-categorization theory} predicts that group polarization will
+occur as more moderate individuals judge the mean group opinion to be
+more extreme \autocite{Turner1989,Abrams1990,Krizan2007}. Fourth and
+finally, the social decision schemes explanation posits that the
+structure of \autocite{Davis1973,Zuber1992,Friedkin1999a}. We do not
+evaluate or compare these theories here, though a simpler explanation
+that extremists are more stubborn may be a more parsimonious explanation
+consistent with these four alternatives
+\autocite{Acemoglu2013,Guazzini2015,Turner2020}.
+
+Each of the ten group polarization studies we analyze here varied
+significantly in how groups were formed, how opinions were gathered, and
+other details. There was significant variation across experimental
+designs, participant population sizes, group sizes, and how like-minded
+groups were formed in the first place. We review some of these details
+below to provide the reader with some insight into the experimental
+designs that produced what we show are plausibly false positive
+detections of group polarization. We focus here on the questions that
+were asked, how like-minded novel experimental groups were formed,
+organizing our presentation by theoretical tradition.
+
+\subsubsection{Details of experimental conditions included in this
+meta-analysis}\label{details-of-experimental-conditions-included-in-this-meta-analysis}
+
+\paragraph*{Atheoretical}
+
+The first of ten studies we included in our meta-analysis was by
+Moscovici and Zavalloni (1969) \autocite{Moscovici1969} who sought to
+measure group polarization among groups of French high school students
+who discussed domestic and foreign politics. Moscovici and Zavalloni
+tested four conditions whether group polarization occurred for two
+different questions and two different rating systems: participants were
+asked their opinions on then-president Charles De Gaulle and their
+opinion about foreign aid from the USA either in the ``Opinion'' setting
+or the ``Judgement'' condition (see Table 4 on p.~132).
+
+The second, relatively more recent article is Schkade, Sunstein, and
+Hastie (2010) \autocite{Schkade2010} who used geographic location to
+construct like-minded groups by running a group polarization experiment
+in two cities in the state of Colorado in the United States: Colorado
+Springs, whose residents historically voted for republican candidates,
+and Boulder, whose residents historically voted for democratic
+candidates. Discussion groups were formed in which they measured group
+polarization when group members discussed global warming, affirmative
+action, and same-sex civil unions.
+
+\paragraph{Social comparisons}
+
+In developing the social comparisons theory of group polarization
+(described in Myers (1978) \autocite{Myers1978}), Myers and Bishop
+(1970), writing in \emph{Science,} \autocite{Myers1970} sought to
+measure group polarization among high school students in the United
+States who discussed racial attitudes and Myers (1975)
+\autocite{Myers1975} measured group polarization in groups of college
+students discussing the quality of a faculty member in one condition,
+and women's rights and issues in another condition.
+
+\paragraph{Persuasive arguments}
+
+\paragraph{Self-categorization}
+
+\paragraph{Social decisions scheme}
+
+The ten studies we included each tested several experimental conditions
+meant to elicit group polarization in novel experimental groups, 60 in
+total, which we briefly introduce here.\\
+
+\subsection{Research overview}\label{research-overview}
+
+To evaluate whether published detections of group polarization are
+plausibly false we developed and combined formal and generative models
+of group polarization to identify plausible simple consensus pre- and
+post-discussion opinion distributions that seem like group polarization
+when binned or measured on ordinal scales. Opinion distributions are
+assumed to be normal. We define a \emph{simple consensus pair} of
+\emph{latent opinion} distributions to be a pair of pre- and
+post-discussion distributions where the mean opinion is constant, but
+the pre-discussion variance is greater than the post-discussion
+variance. A \emph{latent distribution} of opinions represents the
+internal, unobservable distribution of opinions for a group or
+collection of groups. The \emph{observed distribution} is the
+distribution of reported ordinal opinion measurements. We then define a
+\emph{plausibly false detection} of group polarization to be one where
+the mean post-discussion observed distribution is greater in magnitude
+(i.e., more extreme) than the post-discussion observed distribution's,
+but both were measurements of a pair of latent simple consensus
+distributions.
+
+In the formal model, a simulated observed distribution is generated by
+binning a latent distribution into a finite number of ordinal scale
+bins, then integrating over the probability density in each bin. In the
+generative model, an observed distribution is simulated by first drawing
+participant opinions from a latent distribution, then counting the
+number of opinions falling within each bin. The formal model is used to
+identify which, if any, simple consensus pairs exist for a given
+reported detection of group polarization through an optimization routine
+that finds pre- and post-discussion variances that generate simulated
+observations that appear to be group polarization (or reports a failure
+to such variances). If the optimization routine finds a constant latent
+mean that generates simulated observed group polarization, then we say
+the published group polarization detection is plausibly false. If the
+routine fails to find a constant latent mean that generates simulated
+group polarization, then we cannot and do not claim the detection is
+plausibly false. We use the generative model to evaluate how frequently
+we would expect plausibly false detections when there are finite-size
+groups, since the formal model implicitly applies infinite-size groups.
+Therefore in our analyses, we also generate one thousand simulated group
+polarization datasets for each of the 55 plausibly false detections
+identified with the formal model and examine the distribution of group
+polarization effect sizes in terms of Cohen's \(d\). In the first round
+of simulations we set the number of observations to match those in the
+ten published studies we analyzed. Many or most simulated effect sizes
+are non-zero with these numbers of observations. To see if perhaps these
+non-zero effects were due to low statistical power from too few
+observations, we also tested a hypothetical study with ten times more
+participants. Finally, we evaluate whether an appropriate statistical
+model that accounts for the ordinal measurement of latent opinions can
+reliably identify simulated simple consensus in latent opinions when
+simulated observations appear as group polarization based on mean group
+opinion alone.
+
+\section{Methods}\label{methods}
+
+We begin by developing a formal model of ordinal measurement of latent
+continuous opinions (model variables and parameters are given in
+Table~\ref{tab:modelVariables}). First, we assume that participant $i$ has
+latent opinion at time $t$ that is drawn from a group-level normal distribution,
+\begin{equation}
+  o_{i,t} \sim p(o; \mu_t, \sigma_t) = \mathcal{N}(\mu_t, \sigma_t).
+  \label{eq:opinionDistribution}
+\end{equation} 
+\noindent 
+The initial opinion of agent $i$ is are denoted by $o_{i,t=0}$, and thus
+in the model are drawn from a normal distribution with mean $\mu_0$ and standard
+deviation $\sigma_0$. The final opinion of agent $i$ is written $o_{i,t=T}$,
+drawn from a normal distribution with mean $\mu_T$ and standard deviation
+$\sigma_T$.
+
+\section{Results}\label{results}
+
+\vspace{0.5em}
+\begin{center}
+{[ FIGURE \ref{fig:MetricBoxplot} ABOUT HERE ]} \\
+\end{center}
+
+\vspace{0.5em}
+\begin{center}
+{[ FIGURE \ref{fig:OrdinalBoxplot} ABOUT HERE ]} \\
+\end{center}
+
+\section{Discussion}\label{discussion}
+
+Ordinal scales are an indispensable tool for psychological research on
+how social influence changes people's beliefs, opinions, knowledge, or
+similar psychological content. We show here that we must be careful not
+to mistake the measurement procedure as psychological reality when
+designing experiments.
+
+Replication crisis, generalizability crisis \autocite{Yarkoni2022}, and
+general lack of coordination in asking the right questions to
+efficiently understand the mechanisms behind . Understanding mechanisms
+is the goal of science \autocite{Machamer2000,Craver2006}. However, in
+social and psychological sciences, finding mechanisms is more difficult
+due to the complexity of social psychological systems
+\autocite{Dubova2023}, which require a high density of conceptual
+definitions that may be difficult to harmonize.
 
@@
 
+\appendix
+
+\section{Figures}
+
+\begin{figure}
+  \caption{
+    Boxplots visualizing the distribution of Cohen's $d$ values (x-axis in
+    A and B) inferred through a metric linear model over 1000 simulation 
+    trials for each experimental condition (y-axis in A and B). All conditions
+    have many trials that result in non-zero Cohen's $d$ values with number of
+    simulated participants, $N$, set to their empirical values reported in the
+    associated journal article (A). However, when the number of simulated
+    participants is increased to $10N$, there are several conditions that 
+    indeed yield distributions closer to $d = 0$ as expected, however some
+    conditions have $d$ distributions centered on a nonzero mean with little to
+    no overlap with $d=0$; variance in $d$ decreases as expected
+    across experimental conditions (B).
+  }  
+  \label{fig:MetricBoxplots}
+  \centering
+  \begin{subfigure}[]{0.9\textwidth}
+    % \hspace{-7.5em} 
+    \centering
+    \textsf{\large \textbf{A}, empirical $N$} \\
+    \includegraphics[width=\textwidth]{Figures/boxplots/metric_cohens.pdf}
+    \label{fig:MetricBoxplot_N}
+  \end{subfigure} \\
+\end{figure}
+\begin{figure}\ContinuedFloat
+  \begin{subfigure}[]{0.9\textwidth}
+    % \hspace{-5.5em} 
+    \centering
+    \textsf{\large \textbf{B}, 10 $\times$ empirical $N$} \\
+    \includegraphics[width=\textwidth]{Figures/boxplots/10N_metric_cohens.pdf}
+    \label{fig:MetricBoxplot_10N}
+  \end{subfigure}
+\end{figure}
+
+\begin{figure}
+  \caption{Boxplot of Cohen's $d$ for the ordered probit model over 100
+    simulation trias for each experimental condition (y-axis in A and B). Each
+    trial represents a possible outcome of a group polarization experiment where
+    the true opinion shift is zero. The mean across trials is closer to 0 than
+    in the metric case, but many simulated zero-shift experiments still result in
+    an observed shift. This illustrates the weak statistical power of these group
+    polarization experiments.}
+  \label{fig:OrdinalBoxplot}
+  \centering
+    \includegraphics[width=0.9\textwidth]{Figures/boxplots/ordinal_cohens.pdf}
+\end{figure}
+
 \end{document}

### 2024-10-29 GroupPolarizationStatmod/false_detection_rate.R
finishing FDR calculation

@@ -40,9 +40,12 @@ false_detection_rate_by_condition = function(probit_fits_dir = "data/probit_fits
                           Treatment = Treatment)
 
   for (this_sigval in sigvals) {
-    gb = load_fdr_data(fdr_data_file) %>%
+    # Create a tibble with only this_sigval and for one treatment tag
+     alpha_v_this_sigval <- load_fdr_data(fdr_data_file) %>%
            select(Sigval == this_sigval) %>%
            group_by(TreatmentTag)
+
+    alpha_v_sigval %>% filter(mutate(alpha_v_sigval, 
 
     # Calculate alpha from this gb and put in alpha_v_sigval.
     # Or somehow need to save one for each treatment or sig val then

### 2024-11-01 GroupPolarizationStatmod/paper/ms.tex
cleaning up for submission-ready and starting FDR calcs

@@
 \usepackage{amsmath,amssymb}
+\usepackage{algorithm}
+\usepackage{algpseudocode}
 \usepackage{iftex}
@@
 
+\newcommand{\mupre}{\mu_\mathrm{pre}}
+\newcommand{\mupost}{\mu_\mathrm{post}}
+
+\newcommand{\sigmapre}{\sigma_\mathrm{pre}}
+\newcommand{\sigmapost}{\sigma_\mathrm{post}}
+
+\newcommand{\thetafit}{\theta_\mathrm{fit}}
+
+\newcommand{\fdr}{\mathrm{FDR}}
+
+\definecolor{myorange}{RGB}{240, 96, 0}
+\newcommand{\mt}[1]{{\textcolor{myorange} {({\tiny MT:} #1)}}}
+
 \usepackage{authblk}
@@
 {
-\hypersetup{linkcolor=}
+\hypersetup{linkcolor=gray}
 \setcounter{tocdepth}{3}
@@
 }
+
+\begin{center} \noindent\rule{4cm}{0.4pt} \end{center}
+\clearpage
+
 \begin{quote}
@@
 
+\begin{quote}
+A medium effect size is conceived as one large enough to be visible to the naked eye.
+\par\raggedleft(Cohen, 1988 \cite[p. 26]{CohenBook1988})
+\end{quote}
+
 \section{Introduction}\label{introduction}
 
-Unchecked social and political polarization undermines cooperation and
-political efficacy \autocite{Mason2018UncivilAgreementBook,Klein2020}.
-\emph{Group polarization} is the name for a group-level process that
-leads to greater opinion extremity in a group as the group deliberates
-some topic (e.g., ``Do you approve, disapprove, or neither approve nor
-disapprove of the current administration's policies?''). Extremity only
-increases, though, assuming that the mean opinion of the group is
-initially biased towards one extreme or the other
-\autocite{Brown1986,Brown2020}. Group polarization is important to
-understand rigorously, therefore, since if groups become increasingly
-extreme over time, while group opinions are consolidated, this would
-exacerbate political polarization between political opposition groups,
-which is already undermines the ability of partisans in democratic
-governments to work together. Dozens of studies and meta-analyses over
-decades have apparently established the empirical reality of group
-polarization
+the ten-article corpus, which is representative of the
+false discovery rate for group polarization studies.
+
+Unchecked social and political polarization undermines cooperation and political
+efficacy \autocite{Mason2018UncivilAgreementBook,Klein2020}.  \emph{Group
+polarization} is the name that has been given to a group-level process that leads
+to greater opinion extremity in a group as the group deliberates some topic (e.g.,
+``Do you approve, disapprove, or neither approve nor disapprove of the current
+administration's policies?''). Extremity only increases, though, assuming that the
+mean opinion of the group is initially biased towards one extreme or the other
+\autocite{Brown1986,Brown2020}. Group polarization is important to understand
+rigorously, therefore, since if groups become increasingly extreme over time,
+while group opinions are consolidated, this would exacerbate political
+polarization between political opposition groups, which is already undermines the
+ability of partisans in democratic governments to work together. Dozens of studies
+and meta-analyses over decades have apparently established the empirical reality
+of group polarization
 \autocite{Moscovici1969,Myers1970,Isenberg1986,Sunstein2009,Sunstein2019}
-\autocite[Ch. 5]{Brown2000}. We believe group polarization should occur
-in certain contexts since more extreme group members are likely also to
-be more stubborn \autocite{Guazzini2015,Lewandowsky2019}, which should
-pull consensus to extremes \autocite{Turner2020}. However, many or most
-of the evidence for group polarization relies on experimental designs
-that use ordinal measurements that are then analyzed as if they were
-continuous. Unfortunately, this procedure is unfortunately known to
-induce false positives if it is not properly accounted for
+\autocite[Ch. 5]{Brown2000}. We believe group polarization should occur in certain
+contexts since more extreme group members are likely also to be more stubborn
+\autocite{Guazzini2015,Lewandowsky2019}, which should pull consensus to extremes
+\autocite{Turner2020}. However, many or most of the evidence for group
+polarization relies on experimental designs that use ordinal measurements that are
+then analyzed as if they were continuous. Unfortunately, this procedure is
+unfortunately known to induce false positives if it is not properly accounted for
 \autocite{Liddell2018}.
@@
 
-We begin by developing a formal model of ordinal measurement of latent
-continuous opinions (model variables and parameters are given in
-Table~\ref{tab:modelVariables}). First, we assume that participant $i$ has
+% We begin by developing a formal model of ordinal measurement of latent
+% continuous opinions (model variables and parameters are given in
+% Table~\ref{tab:modelVariables}).
+
+Metascience frames our analysis since metascientific theory provides a framework for
+our goal of calculating the false discovery rate across experimental conditions,
+journal articles, and the ten-paper corpus. The false discovery rate is then
+the Type I error rate relative to the rate of positive discoveries, true or false.
+We calculate via simulation and data fitting
+the Type I error rate for the experimental designs used in experimental conditions
+that, under a preliminary large-$N$ model analysis, generated group
+polarization upon ordinal measurement of simple consensus latent opinion 
+distributions. These can then be aggregated to estimate the false discovery rate
+for each journal article and for the ten-article corpus, which is representative of the
+false discovery rate for group polarization studies.
+
+\subsection{Metascience of the false detection rate}
+
+The false detection rate is the scaled probability that a detection, $D$, of
+an effect, $E$, is in fact a false detection and there is \emph{not} an effect,
+denoted $\neg E$. The probability that a detection $D$ has occurred when there
+was no effect, i.e. when $\neg E$, is denoted $\alpha = \Pr(D | \neg E)$.
+This is known as the Type I or family-wise error rate. To calculate the 
+false detection rate ($\fdr$), one must compare $\alpha$ to the statistical power,
+the probability of a detection when there is an effect, denoted $W = \Pr(D | E)$.
+Both these values must be scaled by the rate at which effects occur, denoted $b$.
+Using the notation outlined above, the false detection rate is
+\begin{equation}
+  \fdr = \frac{(1 - b)\alpha}{(1 - b)\alpha + bW}.
+\end{equation}
+
+We calculate $\alpha = \Pr(D | \neg E)$ through simulation for experimental
+conditions identified as plausibly generating false detections, 
+and use generously optimistic values of
+$b=0.1$~\cite{Ioannidis2005,McElreath2015} and $W=0.8$~\cite{Smaldino2016NatSel}
+in the main text and in the compare with more pessimistic, yet realistic, estimates
+of these values, as well as even more optimistic estimates.
+
+In group polarization, $E$ represents the group polarization \emph{effect}, and
+$\neg E$ represents its absence. Simple consensus is one way for group polarization
+to be absent, i.e., simple consensus is one form of $\neg E$. The presence of
+an effect $E$ is independent of whether an effect was \emph{detected}, 
+represented by $D$ if it was detected and $\neg D$ if not.
+
+\subsection{How study design can induce false detections of group polarization}
+
+First, we assume that participant $i$ has
 latent opinion at time $t$ that is drawn from a group-level normal distribution,
@@
 
+\subsection{Root-finding algorithm to identify plausible false detections}
+
+Simple consensus can appear as group polarization since a change in
+
+\begin{algorithm}
+  \caption{User-guided routine to find pre- and post-deliberation variances that
+  yield group polarization in the absence of mean opinion change.
+\mt{NOT YET DONE...}}\label{alg:routine}
+  \begin{algorithmic}
+    \Require $\mu$, $\sigmapre$, $\sigmapost$ \Comment{Static guess for common mean,
+    initial guesses of variances.}
+    \Require $\epsilon$, $\delta$, $i_\mathrm{max}$ \Comment{Root-finding tolerance, step size, and maximum
+    iterations.}
+    % \Require $\delta$ \Comment{Root-finding step size.}
+    % \Require $i_\mathrm{max}$ \Comment{Maximum number of iterations.}
+    \State $\Delta \gets \mupre' - \mupost'$ 
+    \While {$\epsilon < |\Delta| $ \and $i < i_max$}
+      \State $\Delta \gets (\mupre' - \mupost')^2$
+    \EndWhile
+  \end{algorithmic}
+\end{algorithm}
+
+\subsection{Calculate false detection rate from simulations}
+
+False detection rates are calculated using Cohen's $d$ to measure effect
+size, defined as~\cite[p. 331]{Liddell2018}
+\begin{equation}
+  d = \frac{\mupost - \mupre}{\sqrt{\frac{1}{2}(\sigmapost^2 + \sigmapre^2)}}.
+\end{equation}. 
+\noindent
+Denote the collection of fitted parameters from the $i^\mathrm{th}$ simulation trial
+for experimental condition $c$ and journal article $a$ as
+% $\thetafit = \left{\mu_t = \left{\mupre,~\mupost\right},~\sigma_t = \left{\mupre,~\mupost\right}\right}$.
+$\thetafit_{a,c,i} = \{\mupre,~\mupost,~\mupre,~\mupost\}_{a,c,i}$.
+These parameters are the modal estimates for each parameter obtained from
+an ordered probit statistical model fit with
+Bayesian Markov-Chain Monte Carlo (MCMC)~\cite[p. 338]{Liddell2018}.
+
+For each $a,c$, we calculate $\alpha_{a,c}(d^*) =
+\Pr(D |\neg E;~a,~c,~d^*)$ for a given significance value, $d^*$, directly as
+the fraction of $d$ values that are greater than $d^*$.
+
+To calculate article-level false detection curves we calculate the average
+across , we set $\alpha_{a,c} = 0.0$ for those
+conditions for which we failed to generate spurious group polarization in the
+large-$N$ model.
+
 \section{Results}\label{results}
 
-\vspace{0.5em}
-\begin{center}
-{[ FIGURE \ref{fig:MetricBoxplot} ABOUT HERE ]} \\
-\end{center}
+\subsection{Ninety percent of detections are plausibly false}
 
-\vspace{0.5em}
-\begin{center}
-{[ FIGURE \ref{fig:OrdinalBoxplot} ABOUT HERE ]} \\
-\end{center}
+We identified 54 of 60 experimental 
+conditions where a $N \to \infty$ model falsely indicated group polarization 
+instead of simple consensus.
+
+\subsection{Eighty percent false detection rate across corpus for $d^*=0.8$?}
+
+\subsection{What $d^*$ is necessary to achieve magical $\alpha = 0.05$?}
 
@@
 
-\section{Figures}
+% \section{Figures}
 
-\begin{figure}
-  \caption{
-    Boxplots visualizing the distribution of Cohen's $d$ values (x-axis in
-    A and B) inferred through a metric linear model over 1000 simulation 
-    trials for each experimental condition (y-axis in A and B). All conditions
-    have many trials that result in non-zero Cohen's $d$ values with number of
-    simulated participants, $N$, set to their empirical values reported in the
-    associated journal article (A). However, when the number of simulated
-    participants is increased to $10N$, there are several conditions that 
-    indeed yield distributions closer to $d = 0$ as expected, however some
-    conditions have $d$ distributions centered on a nonzero mean with little to
-    no overlap with $d=0$; variance in $d$ decreases as expected
-    across experimental conditions (B).
-  }  
-  \label{fig:MetricBoxplots}
-  \centering
-  \begin{subfigure}[]{0.9\textwidth}
-    % \hspace{-7.5em} 
-    \centering
-    \textsf{\large \textbf{A}, empirical $N$} \\
-    \includegraphics[width=\textwidth]{Figures/boxplots/metric_cohens.pdf}
-    \label{fig:MetricBoxplot_N}
-  \end{subfigure} \\
-\end{figure}
-\begin{figure}\ContinuedFloat
-  \begin{subfigure}[]{0.9\textwidth}
-    % \hspace{-5.5em} 
-    \centering
-    \textsf{\large \textbf{B}, 10 $\times$ empirical $N$} \\
-    \includegraphics[width=\textwidth]{Figures/boxplots/10N_metric_cohens.pdf}
-    \label{fig:MetricBoxplot_10N}
-  \end{subfigure}
-\end{figure}
+% \begin{figure}
+%   \caption{
+%     Boxplots visualizing the distribution of Cohen's $d$ values (x-axis in
+%     A and B) inferred through a metric linear model over 1000 simulation 
+%     trials for each experimental condition (y-axis in A and B). All conditions
+%     have many trials that result in non-zero Cohen's $d$ values with number of
+%     simulated participants, $N$, set to their empirical values reported in the
+%     associated journal article (A). However, when the number of simulated
+%     participants is increased to $10N$, there are several conditions that 
+%     indeed yield distributions closer to $d = 0$ as expected, however some
+%     conditions have $d$ distributions centered on a nonzero mean with little to
+%     no overlap with $d=0$; variance in $d$ decreases as expected
+%     across experimental conditions (B).
+%   }  
+%   \label{fig:MetricBoxplots}
+%   \centering
+%   \begin{subfigure}[]{0.9\textwidth}
+%     % \hspace{-7.5em} 
+%     \centering
+%     \textsf{\large \textbf{A}, empirical $N$} \\
+%     \includegraphics[width=\textwidth]{Figures/boxplots/metric_cohens.pdf}
+%     \label{fig:MetricBoxplot_N}
+%   \end{subfigure} \\
+% \end{figure}
+% \begin{figure}\ContinuedFloat
+%   \begin{subfigure}[]{0.9\textwidth}
+%     % \hspace{-5.5em} 
+%     \centering
+%     \textsf{\large \textbf{B}, 10 $\times$ empirical $N$} \\
+%     \includegraphics[width=\textwidth]{Figures/boxplots/10N_metric_cohens.pdf}
+%     \label{fig:MetricBoxplot_10N}
+%   \end{subfigure}
+% \end{figure}
 
-\begin{figure}
-  \caption{Boxplot of Cohen's $d$ for the ordered probit model over 100
-    simulation trias for each experimental condition (y-axis in A and B). Each
-    trial represents a possible outcome of a group polarization experiment where
-    the true opinion shift is zero. The mean across trials is closer to 0 than
-    in the metric case, but many simulated zero-shift experiments still result in
-    an observed shift. This illustrates the weak statistical power of these group
-    polarization experiments.}
-  \label{fig:OrdinalBoxplot}
-  \centering
-    \includegraphics[width=0.9\textwidth]{Figures/boxplots/ordinal_cohens.pdf}
-\end{figure}
+% \begin{figure}
+%   \caption{Boxplot of Cohen's $d$ for the ordered probit model over 100
+%     simulation trias for each experimental condition (y-axis in A and B). Each
+%     trial represents a possible outcome of a group polarization experiment where
+%     the true opinion shift is zero. The mean across trials is closer to 0 than
+%     in the metric case, but many simulated zero-shift experiments still result in
+%     an observed shift. This illustrates the weak statistical power of these group
+%     polarization experiments.}
+%   \label{fig:OrdinalBoxplot}
+%   \centering
+%     \includegraphics[width=0.9\textwidth]{Figures/boxplots/ordinal_cohens.pdf}
+% \end{figure}
 

### 2024-11-01 GroupPolarizationStatmod/analysis.R

@@ -111,7 +111,7 @@ make_metric_cohens_d_table <- function(studiesAnalysisDfPath = "data/StudiesAnal
       threadlocal_result_rows[[ trial_idx ]] <- 
         data.frame(ArticleTag = row$ArticleTag, 
                    TreatmentTag = row$TreatmentTag,
-                   TrialIndex = trial_idx, 
+                   TrialIndex = trial_idx,
                    LatentMean = row$LatentMean,
                    ObservedShift = row$ObservedShift,
                    LatentSDPre = row$LatentSDPre, 

### 2024-11-01 GroupPolarizationStatmod/experiments.R

@@ -205,26 +205,31 @@ calculateBayesian <- function(N, firstBinValue, nBins, latentMean, latentSd)
 }
 
 
+
+### XXX Outdated as of 2024-10-30 due to Bayesian-only statistical fits XXX ###
+
+
+
 # Uses R's t.test function to fit a linear model to estimate pre- and post-
 # deliberation means in a simple consensus context where there is only one
 # latent mean.
-simulate_metric_cohens_d <- function(N, firstBinValue, nBins, latentMean, latentSdPre, 
-                            latentSdPost, paired = FALSE, var.equal = FALSE)
-{
-  sim_obs_pre <- simulatedObservation(N, firstBinValue, nBins, latentMean, latentSdPre)
+# simulate_metric_cohens_d <- function(N, firstBinValue, nBins, latentMean, latentSdPre, 
+#                             latentSdPost, paired = FALSE, var.equal = FALSE)
+# {
+#   sim_obs_pre <- simulatedObservation(N, firstBinValue, nBins, latentMean, latentSdPre)
 
-  sim_obs_post <- simulatedObservation(N, firstBinValue, nBins, latentMean, latentSdPost)
+#   sim_obs_post <- simulatedObservation(N, firstBinValue, nBins, latentMean, latentSdPost)
   
-  t_result <- t.test(sim_obs_pre, sim_obs_post, paired = paired, var.equal = var.equal)
+#   t_result <- t.test(sim_obs_pre, sim_obs_post, paired = paired, var.equal = var.equal)
   
-  pre_mean_estimate <- t_result$estimate[[1]] 
-  post_mean_estimate <- t_result$estimate[[2]] 
+#   pre_mean_estimate <- t_result$estimate[[1]] 
+#   post_mean_estimate <- t_result$estimate[[2]] 
   
-  pre_sd <- sd(sim_obs_pre)
-  post_sd <- sd(sim_obs_post)
+#   pre_sd <- sd(sim_obs_pre)
+#   post_sd <- sd(sim_obs_post)
   
-  return (cohens_d(pre_mean_estimate, post_mean_estimate, pre_sd, post_sd))
-}
+#   return (cohens_d(pre_mean_estimate, post_mean_estimate, pre_sd, post_sd))
+# }
 
 
 cohens_d <- function(pre_mean_estimate, post_mean_estimate, pre_sd, post_sd) {

### 2024-11-01 GroupPolarizationStatmod/false_detection_rate.R

@@ -1,14 +1,20 @@
 ##
-# 
+# Calculate the false detection rate from bayesian fits data.
 #
+# Author: Matthew A. Turner
+# Date: 2024-10-30
+#
+
+library(tidyverse)
+library(checkmate)
 
 load_fdr_data = function(probit_fits_dir = "data/probit_fits", 
                          sync_file = "data/probit_fits/all.csv",
                          overwrite = FALSE) {
    
-  if (file_exists(sync_file) && !overwrite) {
+  if (!file.exists(sync_file) || overwrite) {
 
-    ret <- dir_ls(ordinal_data_dir, glob = "*.csv") %>%
+    ret <- dir.ls(ordinal_data_dir, glob = "*.csv") %>%
       read_csv() %>% unite(ExperimentID, ArticleTag, TreatmentTag)
 
     write_csv(df, sync_file)

### 2024-11-01 GroupPolarizationStatmod/numerical.R

@@ -14,8 +14,8 @@ source('model.R')
 solveForLatentSD <- function(kVec, latentMean, observedMean, guess, 
                              step = 0.01, maxIts = 10000, tol = 1e-6)
 {
-    # Keep the squared error function to be minimized through hillclimbing
-    # separate from hillclimbing itself.
+    # Keep the error function to be minimized through hillclimbing
+    # separate from hillclimbing itself. ### XXXX BUT WHYYYY seems confusing...
     errfunc <- function(latentSD)
     {
         simulatedObservedMean <- meanObs(
@@ -27,6 +27,9 @@ solveForLatentSD <- function(kVec, latentMean, observedMean, guess,
         return (simulatedObservedMean - observedMean);
     }
 
+    # This structure is kind of weird with this implicitly defined "errfunc" that
+    # only takes one variable because the other one has been set "privately"
+    # and implicitly within this function definition.
     return (hillclimbing(errfunc, guess, step, maxIts, tol))
 }
 
@@ -37,9 +40,7 @@ solveForLatentSD <- function(kVec, latentMean, observedMean, guess,
 hillclimbing <- function(errfunc, sd_guess, stepSize = 0.1, maxIts = 1e5, tol = 1e-4)
 {
     sd_curr <- sd_guess
-    # sqerr_curr <- errfunc(sd_curr)[2]
-    # err_curr <-  errfunc(sd_curr)[2]
-    # sqerr_curr <- err_curr^2
+    # "Error" between the published shift and the one simulated here via simple consensus.
     err_curr <- errfunc(sd_curr)
     sqerr_curr <- err_curr^2
     

### 2024-11-01 GroupPolarizationStatmod/util.R

@@ -117,6 +117,7 @@ convertFriedkinTo20Bin <- function(friedkinVal)
     return ( 20 * (friedkinVal / 100.0) )
 }
 
+
 pooled.sd <- function(x1, x2)
 {
   #calculate sample size of each group

### 2024-11-04 GroupPolarizationStatmod/paper/ms.tex
outlined results subsections

@@
 
-\newcommand{\thetafit}{\theta_\mathrm{fit}}
+\newcommand{\thetafit}{\theta^\mathrm{fit}}
 
@@
 
-\subsection{Eighty percent false detection rate across corpus for $d^*=0.8$?}
+\subsection{False detection rates exceed ??\% even at ``high significance''}
 
-\subsection{What $d^*$ is necessary to achieve magical $\alpha = 0.05$?}
+\mt{Graph with $\alpha(d^*)$ for $d^* = 0.2,~0.5,~0.8$ for each condition,
+each article aggregated, and across all conditions/articles. Use color-coding for
+each article, stems with dots at end, red to yellow gg2 theme, in order of
+appearance.}
+
+\subsection{Minimum $d^*$ for FWER of 0.05}
+
+Solving the inverse problem of which $d^*$ achieves a low family-wise error rate,
+which drives the false detection rate. Here we present our calculations to find
+$d^*$ that achieve $\alpha(d^*) = 0.05$, a commonly-assumed parameter value of
+metascience.
 

### 2024-11-08 GroupPolarizationStatmod/paper/ms.tex
updates to Discussion

@@
 
-Ordinal scales are an indispensable tool for psychological research on
-how social influence changes people's beliefs, opinions, knowledge, or
-similar psychological content. We show here that we must be careful not
-to mistake the measurement procedure as psychological reality when
-designing experiments.
+\mt{Restate results and importance and outline this section.}
 
-Replication crisis, generalizability crisis \autocite{Yarkoni2022}, and
-general lack of coordination in asking the right questions to
-efficiently understand the mechanisms behind . Understanding mechanisms
-is the goal of science \autocite{Machamer2000,Craver2006}. However, in
-social and psychological sciences, finding mechanisms is more difficult
-due to the complexity of social psychological systems
-\autocite{Dubova2023}, which require a high density of conceptual
-definitions that may be difficult to harmonize.
+\mt{One more brick to fall in the ongoing replication and generalizability crisis \autocite{Yarkoni2022}.}
+
+Since there seems to be no selection pressure on false discovery rates, 
+experiment design for studying group polarization was free to evolve randomly.
+Various choices for the number of bins and other instrumental details may have
+then become entrenched in different communities defined by 
+common theoretical perspectives taken by researchers in those communities. 
+Therefore, when 
+Isenberg (1986)~\cite{Isenberg1986} reports that his meta-analysis found more
+support for the XX theory than the YY theory in the form of greater effect sizes,
+it may just be random cultural drift that the XX community instruments were
+more prone to greater .
+
+This leads to one of many follow-up research questions generated by this work.
+First, are there general principles of experimental design that could reduce
+the false detection rate for group polarization or similar experimental protocols?
+Second, what is the best way to calculate false detection rate? 
+Likely, even if we cannot identify parameters consistent with simple consensus
+that generate spurious group polarization shifts in the large $N$ case, one
+may still guess at some simple consensus parameters and caluclate the false
+detection rate as we did. On the flip side, we likely found only one of infinitely
+many simple consensus parameters that could result in a false detection. Does this
+have an effect on the calculated false positive rate? More generally, our
+work motivates similar analyses of sub-fields of social 
+psychology and that do not account for Likert-based experimental designs in
+their statistical analyses. We provided a template for this future work.
+
+\mt{concluding paragraph: first sentences summarize the Discussion, then
+in 1-2 sentences summarize the paper.}
 

### 2024-11-09 GroupPolarizationStatmod/paper/ms.tex
added model sketches and docs to scripts

@@
 people will develop more extreme opinions on some salient topic over the
-course of discussing some topic. Group polarization experiments are thus
+course of discussing some topic (Figure~\ref{fig:ModelSketch}). 
+Group polarization experiments are thus
 initialized by forming like-minded groups of people and surveying their
@@
 more appropriate statistical models that account for data ordinality.
+
+\begin{figure}
+  \caption{\textbf{Common components of a group polarization experiment.}
+  Participants first give their opinion in response to some prompt (A). Afterwards,
+  participants deliberate the topic (B). Participants then give their opinion again
+  (C). Opinions are implicitly behavioral choices that are most consistent with some
+  complex aggregation of brain states and dynamics, and so are theoretically
+  continuous (D).}
+  \centering
+    \includegraphics[width=0.7\textwidth]{Figures/ModelSketchesA-C.pdf}
+    \includegraphics[width=0.65\textwidth]{Figures/ModelSketchesD.pdf}
+  \label{fig:ModelSketch}
+\end{figure}
 

### 2024-11-09 GroupPolarizationStatmod/scripts/plot.R

@@ -1,3 +1,14 @@
+##
+#
+# Plot model illustrations and analysis of family-wise error rate simulations.
+# 
+# Currently model illustrations are in the root project directory and will
+# soon be ported here.
+# 
+# Author: Matthew A. Turner <maturner01@gmail.com>
+# Date: 2024-11-05
+#
+
 library(ggplot2)
 
 source("scripts/analysis.R")

### 2024-11-09 GroupPolarizationStatmod/scripts/analysis.R
added model sketches and docs to scripts

@@ -78,20 +78,6 @@ calc_fdr_vs_significance = function(
 }
 
 
-sigval_for_low_fwer = function(fdr_vs_sig_tbl, target_fwer = 0.05) {
-
-  return(
-
-    fdr_vs_sig_tbl %>%
-      filter(FWER <= target_fwer) %>% 
-      group_by(StudyID, ExperimentID) %>% 
-      filter(SigVal == min(SigVal))  %>% 
-      arrange(desc(SigVal)) 
-
-  )
-}
-
-
 add_default_rows = function(fdr_vs_sig_tbl, default_fwer = 0.05) {
 
   # Figure out the number of sig values used.
@@ -153,3 +139,21 @@ NON_IDENTIFIED =
     DefaultFWER = c(0.5, 0.5, 0.2)
   )
 
+
+##
+# Given fdr_vs_sig_tbl (which includes an FWER column), 
+# calculate the smallest significance value that achieves an FWER of 
+# target_fwer, set to a default main text value of 0.05.
+#
+sigval_for_low_fwer = function(fdr_vs_sig_tbl, target_fwer = 0.05) {
+
+  return(
+
+    fdr_vs_sig_tbl %>%
+      filter(FWER <= target_fwer) %>% 
+      group_by(StudyID, ExperimentID) %>% 
+      filter(SigVal == min(SigVal))  %>% 
+      arrange(desc(SigVal)) 
+
+  )
+}

### 2024-11-18 GroupPolarizationStatmod/paper/ms.tex
good first draft of bar with markers for FWER, FDR

@@
 \makeatletter
-\title{Foundational claims of group polarization are plausibly false,
-with lessons for the empirical study of psychological dynamics}
+\title{Decades of ``group polarization'' replications may be undermined by high false discovery rates}
+% \title{Foundational claims of group polarization are plausibly false,
+% with lessons for the empirical study of psychological dynamics}
 \makeatother
@@
 \begin{abstract}
-\noindent \emph{Group polarization} is the name for a form of consensus
-where members of a like-minded group become, on average, more extreme in
-their opinions after discussing a topic. Group polarization is important
-because it may drive political polarization if moderate partisans become
-more extreme. Theoretically, group polarization is celebrated as a
-regular, reliable outcome of a certain experimental design based on
-dozens of published studies that purportedly demonstrate the phenomena.
-However, in this paper we show that 55 of 60 published observations of
-purported group polarization, from ten foundational studies, are equally
-well-explained by a simple consensus model where group opinion variance
-decreases, but extremity does not increase. We use a novel computational
-meta-analysis framework designed to identify plausible pre- and
-post-discussion opinion distributions that are consistent with simple
-consensus, but appear to be group polarization when measured on ordinal,
-Likert-style scales. This occurs due to ceiling effects induced by
-ordinal measurements. At best this shows that a selection of
-foundational research in social psychology must be discarded. At worst
-this is the tip of the iceberg, and many similar social psychological
-findings within and without the group polarization tradition will
-similarly be shown to be unreliable.
+\noindent 
+\emph{Group polarization} is the name for a form of consensus where members of
+a like-minded group become, on average, more extreme in their opinions after
+discussing a topic. Group polarization is important because it may increase
+social and political tensions if moderates become more extreme. Decades of
+studies have replicated detections of group polarization, typically using an
+ordinal scale to measure opinions.  However, these studies did not account for
+ordinal measurements, which can result in spurious detections of opinion
+change.  Lacking original data we can still calculate the probability that a
+published detection of group polarization was spurious, i.e., to calculate the
+\emph{false detection rate}.  Across 54 group polarization experimental
+conditions from ten representative journal articles, we calculated false
+detection rates to be between 0.52 and 0.88, with a median of 0.75, using a
+generative model of group polarization experiments seeded with data from the
+studies themselves.  We also use our model to develop experimental designs that
+achieve an acceptable false discovery rate.  Much of group polarization
+research may be unreliable, but this analysis can help change that. More
+broadly, this work demonstrates one way that replication success alone does not
+imply epistemic reliability.
 \end{abstract}
@@
 support for the XX theory than the YY theory in the form of greater effect sizes,
-it may just be random cultural drift that the XX community instruments were
-more prone to greater .
+it may just be random cultural drift that the XX community instruments had higher
+intrinsic false detection rates. In some cases it may be debatable whether to
+control for the Type I error rate, or may depend on context~\cite{Rubin2021}. 
+In this context, however, it is clear that we must apply such selection pressure 
+to group polarization experiments or it may be nearly impossible to trust 
+statistical analyses of these studies.
 

### 2024-11-18 GroupPolarizationStatmod/scripts/plot.R

@@ -9,10 +9,10 @@
 # Date: 2024-11-05
 #
 
+library(forcats)
 library(ggplot2)
 
 source("scripts/analysis.R")
-
 
 
 plot_sigval_for_low_fwer = function(target_fwer = 0.05, 
@@ -32,15 +32,37 @@ plot_sigval_for_low_fwer = function(target_fwer = 0.05,
 
 plot_fwer_fdr = function(probit_fits_dir = "data/probit_fits",
                          base_rate = 0.1, power = 0.8, 
-                         significance_vals = c(0.2, 0.5, 0.8)) {
+                         significance_vals = c(0.2, 0.5, 0.8),
+                         medium_sig_val = 0.5) {
 
   # Load data.
-  p = calc_fdr_vs_significance(significance_vals = significance_vals) %>%
-    # Concatenate ExperimentID name.
-    mutate(ExperimentID = paste(StudyID, ExperimentID, sep = "_"), 
-           SigVal = as_factor(SigVal)) %>%
-    arrange(desc(FWER)) %>%
-    ggplot(aes(x = FWER, y = ExperimentID, shape = SigVal)) + geom_point()
+  dat = calc_fdr_vs_significance(significance_vals = significance_vals) %>%
+    mutate(ExperimentID = paste(StudyID, ExperimentID, sep = "_")) %>%
+    select(!c(StudyID)) %>% 
+    mutate(SigVal = as_factor(SigVal))
+
+  # Filter to keep just "medium" effect sizes and sort factors.
+  dat_med = 
+    filter(dat, SigVal == medium_sig_val) %>% 
+    mutate(ExperimentID = fct_reorder(ExperimentID, FDR))
+    
+  p = 
+    ggplot(dat_med, aes(x = FDR, y = ExperimentID)) + 
+      geom_bar(stat = "identity", fill = "red", color = "red", width = 0.15)  +
+      geom_point(aes(x = FWER, y = ExperimentID), 
+                 dat %>% filter(SigVal == medium_sig_val), size = 2) + 
+      geom_point(aes(x = FDR, y = ExperimentID, shape = SigVal), dat) +
+      xlim(c(0, 1))
+
+      
+
+
+  # %>%
+  #   # Concatenate ExperimentID name.
+  #   mutate(ExperimentID = paste(StudyID, ExperimentID, sep = "_"), 
+  #          SigVal = as_factor(SigVal)) %>%
+  #   # arrange(desc(FWER)) %>%
[…]

### 2024-11-18 GroupPolarizationStatmod/src/false_detection_rate.R

@@ -8,6 +8,8 @@
 # library(tidyverse)
 library(dplyr)
 library(fs)
+library(readr)
+library(tidyr)
 
 
 cohens_d <- function(pre_mean_estimate, post_mean_estimate, pre_sd, post_sd) {

### 2024-11-22 GroupPolarizationStatmod/paper/ms.tex
plot good enough for now

@@
 \noindent 
-\emph{Group polarization} is the name for a form of consensus where members of
-a like-minded group become, on average, more extreme in their opinions after
-discussing a topic. Group polarization is important because it may increase
-social and political tensions if moderates become more extreme. Decades of
-studies have replicated detections of group polarization, typically using an
-ordinal scale to measure opinions.  However, these studies did not account for
-ordinal measurements, which can result in spurious detections of opinion
-change.  Lacking original data we can still calculate the probability that a
-published detection of group polarization was spurious, i.e., to calculate the
-\emph{false detection rate}.  Across 54 group polarization experimental
-conditions from ten representative journal articles, we calculated false
-detection rates to be between 0.52 and 0.88, with a median of 0.75, using a
-generative model of group polarization experiments seeded with data from the
-studies themselves.  We also use our model to develop experimental designs that
+\emph{Group polarization} is the name for a form of consensus where
+members of a like-minded group become, on average, more extreme in their
+opinions after discussing a topic. Group polarization is important because
+it may increase social and political tensions if moderates become more
+extreme. Decades of studies have replicated detections of group
+polarization, typically using an ordinal scale to measure opinions.
+However, these studies did not account for ordinal measurements, which can
+result in spurious detections of opinion change.  Lacking original data we
+can still calculate the probability that a published detection of group
+polarization was spurious, i.e., to calculate the \emph{false detection
+rate}.  Across 54 group polarization experimental conditions from ten
+representative journal articles, we calculated false detection rates to be
+between 0.52 and 0.88, with a median of 0.75, using a generative model of
+group polarization experiments seeded with data from the studies
+themselves.  We also use our model to develop experimental designs that
 achieve an acceptable false discovery rate.  Much of group polarization
 research may be unreliable, but this analysis can help change that. More
-broadly, this work demonstrates one way that replication success alone does not
-imply epistemic reliability.
+broadly, this work demonstrates one way that replication success alone
+does not imply epistemic reliability.
 \end{abstract}

### 2024-11-22 GroupPolarizationStatmod/scripts/plot.R

@@ -11,6 +11,7 @@
 
 library(forcats)
 library(ggplot2)
+library(latex2exp)
 
 source("scripts/analysis.R")
 
@@ -30,7 +31,9 @@ plot_sigval_for_low_fwer = function(target_fwer = 0.05,
 }
 
 
-plot_fwer_fdr = function(probit_fits_dir = "data/probit_fits",
+plot_fwer_fdr = function(save_path = "paper/Figures/fwer_fdr_experiment.pdf", 
+                         aggregate = FALSE,
+                         probit_fits_dir = "data/probit_fits",
                          base_rate = 0.1, power = 0.8, 
                          significance_vals = c(0.2, 0.5, 0.8),
                          medium_sig_val = 0.5) {
@@ -38,7 +41,7 @@ plot_fwer_fdr = function(probit_fits_dir = "data/probit_fits",
   # Load data.
   dat = calc_fdr_vs_significance(significance_vals = significance_vals) %>%
     mutate(ExperimentID = paste(StudyID, ExperimentID, sep = "_")) %>%
-    select(!c(StudyID)) %>% 
+    # select(!c(StudyID)) %>% 
     mutate(SigVal = as_factor(SigVal))
 
   # Filter to keep just "medium" effect sizes and sort factors.
@@ -46,31 +49,25 @@ plot_fwer_fdr = function(probit_fits_dir = "data/probit_fits",
     filter(dat, SigVal == medium_sig_val) %>% 
     mutate(ExperimentID = fct_reorder(ExperimentID, FDR))
     
+  # Redefine default shape palette
+  scale_shape_discrete = function(...) {
+    scale_shape_manual(values = c(6, 1, 0))
+  }
+
   p = 
     ggplot(dat_med, aes(x = FDR, y = ExperimentID)) + 
-      geom_bar(stat = "identity", fill = "red", color = "red", width = 0.15)  +
-      geom_point(aes(x = FWER, y = ExperimentID), 
-                 dat %>% filter(SigVal == medium_sig_val), size = 2) + 
-      geom_point(aes(x = FDR, y = ExperimentID, shape = SigVal), dat) +
-      xlim(c(0, 1))
+      geom_bar(aes(fill = StudyID, color = StudyID), stat = "identity", width = 0.15)  +
+      labs(fill = "Study ID", color = "Study ID") +
[…]

### 2024-11-24 GroupPolarizationStatmod/paper/ms.tex
made figures and started writing analysis text

@@
 
-\mt{Graph with $\alpha(d^*)$ for $d^* = 0.2,~0.5,~0.8$ for each condition,
-each article aggregated, and across all conditions/articles. Use color-coding for
-each article, stems with dots at end, red to yellow gg2 theme, in order of
-appearance.}
+We calculated the false detection rate across experimental conditions where
+we identified parameters that generate spurious group polarization in the 
+large-$N$ model (Figure~\ref{fig:fdr_experiments}). We then aggregated over
+experimental conditions within each study to understand the false detection rate
+across the ten studies (Figure~\ref{fig:fdr_studies}). Across the identified
+spurious conditions, we calculated a mean false detection rate of 0.74 with
+a significance level of $d^* = 0.5$. We also
+examined the mean false detection rates across studies if we set a highly
+optimistic estimate of the family-wise error rate, setting $\alpha = 0.05$. 
+This only reduces the false discovery rate to 0.71, but does slightly change the
+rankings of which study had the highest mean false discovery rates, with 
+Schkade, et al., (2010) going from highest mean false discovery rate to 
+second-lowest false discovery rate. The minimum false discovery rate in a study
+was Myers (1975), with a false discovery rate of 0.58, meaning that probability
+a group polarization detection in the best-designed study is greater than one in
+two.
 

### 2024-11-24 GroupPolarizationStatmod/scripts/analysis.R

@@ -12,6 +12,8 @@
 # Date: 2024-11-05
 #
 
+library(purrr)
+
 
 source("src/false_detection_rate.R")
 
@@ -81,12 +83,13 @@ calc_fdr_vs_significance = function(
 add_default_rows = function(fdr_vs_sig_tbl, default_fwer = 0.05) {
 
   # Figure out the number of sig values used.
-  sigvals = fdr_vs_sig_tbl$SigVal
-  n_sigvals = length(unique(sigvals))
+  sigvals = unique(fdr_vs_sig_tbl$SigVal)
+  n_sigvals = length(sigvals)
+  # print(sigvals)
 
   # Extract power and base rate.
-  W = fdr_vs_sig_tbl[1, Power]
-  b = fdr_vs_sig_tbl[1, BaseRate]
+  W = pull(fdr_vs_sig_tbl[1, "Power"])
+  b = pull(fdr_vs_sig_tbl[1, "BaseRate"])
   
   # Load and use non-identified plausible.
   non_id = NON_IDENTIFIED
@@ -94,41 +97,53 @@ add_default_rows = function(fdr_vs_sig_tbl, default_fwer = 0.05) {
   n_non_id = n_distinct(non_id)
   n_default_rows = n_non_id * n_sigvals
 
-  # Set default_fwer to be half of the least fwer calculated for that article
-  # by first finding the minimum fwer for the study.
-  study_ids = non_id$StudyID
+  # # Set default_fwer to be half of the least fwer calculated for that article
+  # # by first finding the minimum fwer for the study.
+  # study_ids = non_id$StudyID
 
-  default_fwers_tbl = 
-    fdr_vs_sig_tbl %>% 
-    filter(StudyID %in% study_ids) %>%
-    group_by(StudyID) %>%
-    summarize(default_fwer = min(FWER))
+  # default_fwers_tbl = 
+  #   fdr_vs_sig_tbl %>% 
+  #   filter(StudyID %in% study_ids) %>%
+  #   group_by(StudyID) %>%
+  #   summarize(default_fwer = min(FWER))
 
-  # Then create a vector that fits with default_rows.
-  default_fwers_vec = map(default_frews$DEFAULT, 
-                          \(frew) rep(frew / 2.0, n_sigvals))
+  # # Then create a vector that fits with default_rows.
+  # default_fwers_vec = map(default_fwers_tbl$default_fwer, 
+  #                         \(fwer) rep(fwer / 2.0, n_sigvals))
+  # print(default_fwers_vec)
 
[…]

### 2024-11-24 GroupPolarizationStatmod/scripts/plot.R

@@ -22,66 +22,124 @@ plot_sigval_for_low_fwer = function(target_fwer = 0.05,
                                     base_rate = 0.1, power = 0.8, 
                                     significance_vals = seq(0.2, 3.0, 0.01)) {
   
-  tbl = 
+  # Going to create ggplot, p.
+  p = 
+    # First load and pre-process calculation of FWER ≤ target_fwer by experiment.
     sigval_for_low_fwer(calc_fdr_vs_significance(significance_vals = 
                                                  significance_vals)) %>%
-      mutate(ExperimentID = paste(StudyID, ExperimentID, "_")) # %>%
+      mutate(ExperimentID = paste(StudyID, ExperimentID, "_"),
+             SigVal = as.numeric(SigVal)) %>%
 
-  return (tbl)
+    # Then plot the result.
+    ggplot(aes(x = SigVal, y = ExperimentID, fill = StudyID, color = StudyID)) +
+      geom_bar(stat = "identity", width = 0.15)  +
+      xlab(TeX("Significance $d^*$ for $\\alpha \\leq 0.05")) + 
+      ylab("Experiment ID")
+        
+     ggsave(save_path, p) 
 
-  # ggplot(tbl, aes(x = SigVal, y = ExperimentID)) +
-  #   geom_bar(aes(fill = StudyID, color = StudyID), 
-  #            stat = "identity", width = 0.15)  +
-  #   xlab(TeX("Significance $d^*$ for $\\alpha \\leq 0.05")) + 
-  #   ylab("Experiment ID")
-      
-  #  ggsave(save_path, p) 
-
-  #  return (p)
+   return (p)
 }
 
 
-plot_fdr = function(save_path = "paper/Figures/fwer_fdr_experiment.pdf", 
-                         aggregate = FALSE,
-                         probit_fits_dir = "data/probit_fits",
-                         base_rate = 0.1, power = 0.8, 
-                         significance_vals = c(0.2, 0.5, 0.8),
-                         medium_sig_val = 0.5) {
+plot_fdr = function(save_path = "paper/Figures/fdr_vs_experiment.pdf", 
+                    aggregate = FALSE, use_non_identified = FALSE,
+                    probit_fits_dir = "data/probit_fits",
+                    default_fwer = 0.05,
[…]

### 2024-12-06 GroupPolarizationStatmod/paper/ms.tex
final draft of plots before sending to p

@@
 
-\subsection{Ninety percent of detections are plausibly false}
+\subsection{Ninety percent of detections identified as plausibly false}
 
@@
 metascience.
+
+\begin{figure}
+  \caption{\textbf{Significance value, $d^*$, to limit family-wise error 
+  rate of 0.05 for each identified study.} \mt{Does this make sense? Order by the
+rankings in fdr\_vs\_experiment plots to better understand/compare.}}
+  \centering
+    \includegraphics[width=0.55\textwidth]{Figures/sigval_for_low_fwer.pdf}
+  \label{fig:sigval_for_low_fwer}
+\end{figure}
 

### 2024-12-06 GroupPolarizationStatmod/scripts/analysis.R

@@ -142,7 +142,7 @@ aggregate_fdr_vs_sig = function(fdr_vs_sig_tbl, use_non_identified = FALSE,
   }  
 
   return (
-    group_by(tbl, StudyID, SigVal) %>% summarise(FDR = mean(FDR))
+    group_by(tbl, StudyID, SigVal) %>% summarise(FDR = mean(FDR), FWER = mean(FWER))
   )
 }
 
@@ -162,14 +162,20 @@ NON_IDENTIFIED =
 #
 sigval_for_low_fwer = function(fdr_vs_sig_tbl, target_fwer = 0.05) {
 
+  # return (fdr_vs_sig_tbl)
+
   return(
 
     fdr_vs_sig_tbl %>%
+      # Reorder ExperimentID factors to be by max FWER, i.e., FWER for lowest sig. val.
+      mutate(StudyID = fct_reorder(StudyID, FWER, mean), 
+             ExperimentID = fct_reorder(ExperimentID, FWER, mean)) %>%
       filter(FWER <= target_fwer) %>% 
-      group_by(StudyID, ExperimentID) %>% 
+      # group_by(StudyID, ExperimentID) %>% 
+      group_by(ExperimentID) %>% 
       filter(SigVal == min(SigVal))  %>% 
-      mutate(SigVal = as_factor(SigVal)) %>%
-      arrange(desc(SigVal)) 
+      mutate(SigVal = as_factor(SigVal), ExperimentID = fct_reorder(ExperimentID, FDR)) # %>%
+      # arrange(desc(SigVal)) 
 
   )
 }

### 2024-12-06 GroupPolarizationStatmod/scripts/plot.R

@@ -12,34 +12,47 @@
 library(forcats)
 library(ggplot2)
 library(latex2exp)
+library(emojifont)
 
+source("src/false_detection_rate.R")
 source("scripts/analysis.R")
 
+mytheme = theme(
+  axis.line = element_line(), legend.key=element_rect(fill = NA),
+  text = element_text(size=16),# family = 'PT Sans'),
+  # legend.key.width = unit(2, 'cm'),
+  # legend.key.size = unit(1.5, 'lines'),
+  panel.background = element_rect(fill = "white"),
+  panel.grid.major.x = element_line(color = "lightgrey", linewidth = 0.1, linetype = 2)
+)
 
-plot_sigval_for_low_fwer = function(target_fwer = 0.05, 
-                                    save_path = "paper/Figures/sigval_for_low_fwer.pdf",
-                                    probit_fits_dir = "data/probit_fits",
-                                    base_rate = 0.1, power = 0.8, 
-                                    significance_vals = seq(0.2, 3.0, 0.01)) {
+rhg_cols <- rev(c("#771C19", "#AA3929", "#E25033", "#F27314", "#F8A31B", 
+              "#E2C59F", "#B6C5CC", "#8E9CA3", "#556670", "#556AAA", "#000000"))
+rhg_cols_10 <- rev(c("#771C19", "#AA3929", "#E25033", "#F27314", "#F8A31B", 
+              "#E2C59F", "#B6C5CC", "#8E9CA3", "#556670", "#556AAA"))
+
+plot_ordinal_cohens <- function(ordinal_data_dir = "data/probit_fits",
+                                sync_file = "data/probit_fits/all.csv",
+                                output_file = "paper/Figures/ordinal_cohens.pdf",
+                                overwrite = FALSE) {
   
-  # Going to create ggplot, p.
-  p = 
-    # First load and pre-process calculation of FWER ≤ target_fwer by experiment.
-    sigval_for_low_fwer(calc_fdr_vs_significance(significance_vals = 
-                                                 significance_vals)) %>%
-      mutate(ExperimentID = paste(StudyID, ExperimentID, "_"),
-             SigVal = as.numeric(SigVal)) %>%
-
-    # Then plot the result.
[…]

### 2024-12-08 GroupPolarizationStatmod/paper/ms.tex
starting to finalzie draft

@@
 \begin{figure}
-  \caption{\textbf{Common components of a group polarization experiment.}
-  Participants first give their opinion in response to some prompt (A). Afterwards,
-  participants deliberate the topic (B). Participants then give their opinion again
-  (C). Opinions are implicitly behavioral choices that are most consistent with some
+  \caption{\textbf{Opinions are implicitly behavioral choices that are most consistent with some
   complex aggregation of brain states and dynamics, and so are theoretically
-  continuous (D).}
+  continuous (A). Common components of a group polarization experiment.}
+  Participants first give their opinion in response to some prompt (B). Afterwards,
+  participants deliberate the topic (C). Participants then give their opinion again
+  (D). }
   \centering
-    \includegraphics[width=0.7\textwidth]{Figures/ModelSketchesA-C.pdf}
-    \includegraphics[width=0.65\textwidth]{Figures/ModelSketchesD.pdf}
+    \includegraphics[width=0.65\textwidth]{Figures/Model/ModelSketchesA.pdf}
+    \includegraphics[width=0.7\textwidth]{Figures/Model/ModelSketchesB-D.pdf}
   \label{fig:ModelSketch}
@@
 
+An opinion is never measured directly, only behaviors meant to indicate one's opinion.
+For example, only the location of a mark that a participant makes is measured, not
+the complex neural activity that somehow aggregates and expresses some latent
+representation. Because so much neural activity must be integrated to yield an opinion,
+it is most reasonable to assume opinions are continuous. We call these 
+\textit{latent opinions} because they are unobservable. In contrast, ordinal-valued
+\text{observed opinions} are the value participants report before and after group
+polarization deliberation.
+
 \subsection{Root-finding algorithm to identify plausible false detections}
@@
 
+\begin{figure}
+
+  \caption{\textbf{Family-wise error rates and false discovery rates across studies and experiments for
+    low, medium, and high Cohen's $d$ significance.}}
+  \label{fig:fwer_fdr_synthesis}
+
+  \centering
+  \includegraphics[width=1.1\textwidth]{Figures/Results/fwer_fdr_synthesis.pdf}
+
+\end{figure}
+
 \subsection{Minimum $d^*$ for FWER of 0.05}
@@
   \centering
-    \includegraphics[width=0.55\textwidth]{Figures/sigval_for_low_fwer.pdf}
+    \includegraphics[width=0.75\textwidth]{Figures/Results/sigval_for_low_fwer.pdf}
   \label{fig:sigval_for_low_fwer}

### 2024-12-08 GroupPolarizationStatmod/scripts/plot.R

@@ -1,233 +1,0 @@
-##
-#
-# Plot model illustrations and analysis of family-wise error rate simulations.
-# 
-# Currently model illustrations are in the root project directory and will
-# soon be ported here.
-# 
-# Author: Matthew A. Turner <maturner01@gmail.com>
-# Date: 2024-11-05
-#
-
-library(forcats)
-library(ggplot2)
-library(latex2exp)
-library(emojifont)
-
-source("src/false_detection_rate.R")
-source("scripts/analysis.R")
-
-mytheme = theme(
-  axis.line = element_line(), legend.key=element_rect(fill = NA),
-  text = element_text(size=16),# family = 'PT Sans'),
-  # legend.key.width = unit(2, 'cm'),
-  # legend.key.size = unit(1.5, 'lines'),
-  panel.background = element_rect(fill = "white"),
-  panel.grid.major.x = element_line(color = "lightgrey", linewidth = 0.1, linetype = 2)
-)
-
-rhg_cols <- rev(c("#771C19", "#AA3929", "#E25033", "#F27314", "#F8A31B", 
-              "#E2C59F", "#B6C5CC", "#8E9CA3", "#556670", "#556AAA", "#000000"))
-rhg_cols_10 <- rev(c("#771C19", "#AA3929", "#E25033", "#F27314", "#F8A31B", 
-              "#E2C59F", "#B6C5CC", "#8E9CA3", "#556670", "#556AAA"))
-
-plot_ordinal_cohens <- function(ordinal_data_dir = "data/probit_fits",
-                                sync_file = "data/probit_fits/all.csv",
-                                output_file = "paper/Figures/ordinal_cohens.pdf",
-                                overwrite = FALSE) {
-  
-  # df <- dir_ls(ordinal_data_dir, glob = "*.csv") %>%
-  #   read_csv() %>% unite(ExperimentID, ArticleTag, TreatmentTag)
-  df <- load_probit_data(probit_fits_dir, sync_file, overwrite)
-  
-  df$Cohens_d <- cohens_d(df$LatentMeanPrePosteriorMean, df$LatentMeanPostPosteriorMean,
-                          df$LatentMeanPrePosteriorSD, df$LatentMeanPostPosteriorSD)
-  
-  ggplot(df, mapping = aes(x = Cohens_d, y = ExperimentID)) + 
-    geom_vline(xintercept = 0, color = "red") + 
[…]

### 2024-12-08 GroupPolarizationStatmod/scripts/plot/results.R

@@ -1,0 +1,232 @@
+##
+#
+# Plot analysis of family-wise error rate simulations.
+# 
+# Currently model illustrations are in the root project directory and will
+# soon be ported here.
+# 
+# Author: Matthew A. Turner <maturner01@gmail.com>
+# Date: 2024-12-07
+#
+
+library(forcats)
+library(ggplot2)
+library(latex2exp)
+library(emojifont)
+
+source("src/false_detection_rate.R")
+source("scripts/analysis.R")
+
+mytheme = theme(
+  axis.line = element_line(), legend.key=element_rect(fill = NA),
+  text = element_text(size=16),# family = 'PT Sans'),
+  # legend.key.width = unit(2, 'cm'),
+  # legend.key.size = unit(1.5, 'lines'),
+  panel.background = element_rect(fill = "white"),
+  panel.grid.major.x = element_line(color = "lightgrey", linewidth = 0.1, linetype = 2)
+)
+
+rhg_cols <- rev(c("#771C19", "#AA3929", "#E25033", "#F27314", "#F8A31B", 
+              "#E2C59F", "#B6C5CC", "#8E9CA3", "#556670", "#556AAA", "#000000"))
+
+rhg_cols_10 <- rev(c("#771C19", "#AA3929", "#E25033", "#F27314", "#F8A31B", 
+              "#E2C59F", "#B6C5CC", "#8E9CA3", "#556670", "#556AAA"))
+
+plot_ordinal_cohens <- function(ordinal_data_dir = "data/probit_fits",
+                                sync_file = "data/probit_fits/all.csv",
+                                output_file = "paper/Figures/ordinal_cohens.pdf",
+                                overwrite = FALSE) {
+  
+  # df <- dir_ls(ordinal_data_dir, glob = "*.csv") %>%
+  #   read_csv() %>% unite(ExperimentID, ArticleTag, TreatmentTag)
+  df <- load_probit_data(probit_fits_dir, sync_file, overwrite)
+  
+  df$Cohens_d <- cohens_d(df$LatentMeanPrePosteriorMean, df$LatentMeanPostPosteriorMean,
+                          df$LatentMeanPrePosteriorSD, df$LatentMeanPostPosteriorSD)
+  
+  ggplot(df, mapping = aes(x = Cohens_d, y = ExperimentID)) + 
+    geom_vline(xintercept = 0, color = "red") + 
[…]

### 2024-12-19 GroupPolarizationStatmod/paper/ms.tex
all materials ready for final draft

--- 2024-12-06 GroupPolarizationStatmod/paper/ms.tex
@@
 \usepackage{iftex}
-\ifPDFTeX
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provide euro and other symbols
-\else % if luatex or xetex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-\fi
-\usepackage{lmodern}
-\ifPDFTeX\else  
-    % xetex/luatex font selection
-\fi
-% Use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{% use microtype if available
-  \usepackage[]{microtype}
-  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\usepackage{xcolor}
-\usepackage[margin=1.1in,letterpaper]{geometry}
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\setcounter{secnumdepth}{5}
-% Make \paragraph and \subparagraph free-standing
-\ifx\paragraph\undefined\else
-  \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-  \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
-\usepackage{calc} % for calculating minipage widths
-% Correct order of tables after \paragraph or \subparagraph
-\usepackage{etoolbox}
-\makeatletter
-\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
-\makeatother
-% Allow footnotes in longtable head/foot
-\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
-\makesavenoteenv{longtable}
-\usepackage{graphicx}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-
-\makeatletter
-\@ifpackageloaded{caption}{}{\usepackage{caption}}
-\AtBeginDocument{%
-\ifdefined\contentsname
-  \renewcommand*\contentsname{Table of contents}
-\else
-  \newcommand\contentsname{Table of contents}
-\fi
-\ifdefined\listfigurename
-  \renewcommand*\listfigurename{List of Figures}
-\else
-  \newcommand\listfigurename{List of Figures}
-\fi
-\ifdefined\listtablename
-  \renewcommand*\listtablename{List of Tables}
-\else
-  \newcommand\listtablename{List of Tables}
-\fi
-\ifdefined\figurename
-  \renewcommand*\figurename{Figure}
-\else
-  \newcommand\figurename{Figure}
-\fi
-\ifdefined\tablename
-  \renewcommand*\tablename{Table}
-\else
-  \newcommand\tablename{Table}
-\fi
-}
-\@ifpackageloaded{float}{}{\usepackage{float}}
-\floatstyle{ruled}
-\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
-\floatname{codelisting}{Listing}
-\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
-\makeatother
-\makeatletter
-\makeatother
-\makeatletter
-\@ifpackageloaded{caption}{}{\usepackage{caption}}
-\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
-\makeatother
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
-\usepackage[sorting=none, style=numeric-comp]{biblatex}
-\addbibresource{/Users/mt/workspace/Writing/library.bib}
-\usepackage{bookmark}
-
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same} % disable monospaced font for URLs
+\input{preamble.tex}
 \hypersetup{
-  pdftitle={Foundational claims of group polarization are plausibly false, with lessons for the empirical study of psychological dynamics},
+  pdftitle={Group polarization replications may be marred by high false discovery rates},
   colorlinks=true,
@@
   citecolor={Blue},
-  urlcolor={Blue},
-  pdfcreator={LaTeX via pandoc}}
+  urlcolor={Blue}}
 
@@
 
+\newcommand{\meanobs}{\bar{o}}
+\newcommand{\meanobsemp}{\bar{\mathbf{o}}}
+\newcommand{\meanobst}{\meanobs_t}
+\newcommand{\meanobsjt}{\meanobs_{j,t}}
+
+\newcommand{\meanobspre}{\meanobs_\mathrm{pre}}
+\newcommand{\meanobspost}{\meanobs_\mathrm{post}}
+
 \definecolor{myorange}{RGB}{240, 96, 0}
@@
 \usepackage{etoolbox}
+\usepackage{wrapfig}
 \makeatletter
-\title{Decades of ``group polarization'' replications may be undermined by high false discovery rates}
-% \title{Foundational claims of group polarization are plausibly false,
-% with lessons for the empirical study of psychological dynamics}
+\title{Group polarization replications may be marred by high false discovery rates}
 \makeatother
@@
 between 0.52 and 0.88, with a median of 0.75, using a generative model of
-group polarization experiments seeded with data from the studies
-themselves.  We also use our model to develop experimental designs that
+group polarization experiments seeded with empirical data.  
+We also use our model to develop experimental designs that
 achieve an acceptable false discovery rate.  Much of group polarization
-research may be unreliable, but this analysis can help change that. More
-broadly, this work demonstrates one way that replication success alone
+research may be unreliable. This analysis can help change that by enabling others to avoid this
+hidden pitfall. More
+broadly this work demonstrates one important way that replication success alone
 does not imply epistemic reliability.
@@
 
-\begin{figure}
-  \caption{\textbf{Common components of a group polarization experiment.}
-  Participants first give their opinion in response to some prompt (A). Afterwards,
-  participants deliberate the topic (B). Participants then give their opinion again
-  (C). Opinions are implicitly behavioral choices that are most consistent with some
-  complex aggregation of brain states and dynamics, and so are theoretically
-  continuous (D).}
-  \centering
-    \includegraphics[width=0.7\textwidth]{Figures/ModelSketchesA-C.pdf}
-    \includegraphics[width=0.65\textwidth]{Figures/ModelSketchesD.pdf}
-  \label{fig:ModelSketch}
-\end{figure}
+% \begin{figure}
+%   \caption{\textbf{Opinions are implicitly behavioral choices that are most consistent with some
+%   complex aggregation of brain states and dynamics, and so are theoretically
+%   continuous (A). Common components of a group polarization experiment.}
+%   Participants first give their opinion in response to some prompt (B). Afterwards,
+%   participants deliberate the topic (C). Participants then give their opinion again
+%   (D). }
+%   \centering
+%     \includegraphics[width=0.65\textwidth]{Figures/Model/ModelSketchesA.pdf}
+%     \includegraphics[width=0.7\textwidth]{Figures/Model/ModelSketchesB-D.pdf}
+%   \label{fig:ModelSketch}
+% \end{figure}
 
@@
 
+% \begin{wrapfigure}{l}{0.5\textwidth}
+\begin{figure}
+  \caption{\textbf{Group polarization experimental model.} Latent opinions are 
+    represented as numbers corresponding to the valence and extremism of one's opinion
+(A). Before deliberating a topic of interest, all participants report their 
+opinion (B). Participants then deliberate the topic with others (C). Following
+deliberation participants report their opinion once more (D).}
+  \centering
+  \includegraphics[width=0.65\textwidth]{Figures/Model/ModelSketches.pdf}
+  \label{fig:experiment_model}
+% \end{wrapfigure}
+\end{figure}
+
 \section{Methods}\label{methods}
@@
 
-\subsection{Metascience of the false detection rate}
+\subsection{Metascience of the false detection rate in group polarization}
 
@@
 
-We calculate $\alpha = \Pr(D | \neg E)$ through simulation for experimental
-conditions identified as plausibly generating false detections, 
-and use generously optimistic values of
-$b=0.1$~\cite{Ioannidis2005,McElreath2015} and $W=0.8$~\cite{Smaldino2016NatSel}
-in the main text and in the compare with more pessimistic, yet realistic, estimates
-of these values, as well as even more optimistic estimates.
-
 In group polarization, $E$ represents the group polarization \emph{effect}, and
@@
 
+The group polarization studies are journal articles that present
+the results of multiple experiments to support some hypothesis. To organize
+the presentation of false detection rates across the ten studies in our
+analysis, we first will organize false detection rates by experiment, where
+we created `ExperimentID` tags for each of the sixty experiments across ten
+studies. We also created `StudyID` tags for each of the ten journal articles
+with the key `{Author}{Year}`, e.g., \texttt{Schkade2010} for Schkade, et al.,
+(2010) \cite{Schkade2010}.
+
+We calculate $\alpha = \Pr(D | \neg E)$ through simulation for experimental
+conditions, 
+and use optimistic values of
+$b=0.1$~\cite{Ioannidis2005,McElreath2015} and $W=0.8$~\cite{Smaldino2016NatSel}
+in the main text and in the compare with more pessimistic, yet realistic, estimates
+of these values, as well as even more optimistic estimates.
+
 \subsection{How study design can induce false detections of group polarization}
 
-First, we assume that participant $i$ has
-latent opinion at time $t$ that is drawn from a group-level normal distribution,
+An opinion is never measured directly, only a behavior that indicates an opinion
+in a behavioral experiment or survey.  For example, the location of a mark
+that a participant makes is measured, not the complex neural activity that
+underlies forming opinions (Figure~\ref{fig:experiment_model}A) \mt{Some of this
+needs to be Intro'd}.
+Theoretically, an internal opinion in the form of neural activity is a
+\textit{latent opinion}, from the Latin \emph{latentem} meaning ``lie
+hidden''.  Participants report \text{observed opinions}, typically as one of
+a finite number of ranked, successive integer values that form the
+\emph{selection set}. The minimum possible value in the selection set might
+mean ``strongly disagree'' and the maximum value represents ``strongly
+agree''.
+
+% \begin{wrapfigure}{r}{0.5\textwidth}
+\begin{figure}
+  \caption{\textbf{Opinion measurement model and spurious group polarization.} 
+  Latent model participant opinions are drawn from a normal distribution with the
+  listed parameters (A,C). The pre-deliberation latent distribution has high
+  variance, but appears polarized when measured via integration (A).  The latent
+  distribution's variance is expected to decrease during deliberation (B) via
+  consensus, but in this example the latent mean is static, 
+$\mupre = \mupost$. However, when the post-deliberation distribution is measured,
+the simulated observed mean increases, i.e., spurious
+group polarization occurs (C).}
+  \centering
+  \includegraphics[width=0.65\textwidth]{Figures/Model/latent-ordinal-distros.pdf}
+  \label{fig:distros}
+\end{figure}
+% \end{wrapfigure}
+
+Group polarization experiments share a common experimental design where,
+first, participant opinions are gathered on a topic of interest. Next, groups
+are formed that have some initial opinion extremism. Participants deliberate
+or discuss the topic for some set time. After the set time, participants
+again give their opinion. Group polarization is said to occur if the
+mean group opinion became more extreme, i.e., if the magnitude of the mean
+group opinion is greater \emph{post-deliberation} compared to
+\emph{pre-deliberation}. For example,
+\cite{Moscovici1969} asked Parisian high school students to indicate their
+degree of agreement with the statement, ``American economic aid is always
+used for political pressure'' (Figure~\ref{fig:experiment_model}A). There
+were seven options in the selection set, -3 indicating ``strongly disagree'',
++3 indicating ``strongly agree'', and 0 was neutral. Moscovici and Zavalloni
+report that the mean pre-deliberation opinion was -0.61 and the mean
+post-deliberation opinion was -1.09 (Moscovici and Zavalloni, 1969; Table 4,
+p. 132). This is a -0.48 shift to more extreme disagreement with the 
+statement, reported by Moscovici and Zavalloni as statistically significant
+in Table 5 on p. 132.
+
+We formalize this experimental design in the $N\to\infty$ case where the
+distribution of ordinal measurements is created by integrating continuous latent
+opinion distributions.  Latent opinions in the model are normally distributed with
+mean $\mu_t$ and variance $\sigma_t$ for time $t$. We are concerned with only two
+times, \emph{pre-} and \emph{post-} deliberation ($t=0$ and $t=T$, respectively).
+For short we can write $\mu_\mathrm{pre}$ and $\sigma_\mathrm{pre}$ for
+pre-deliberation mean and variance (and similar for \emph{post}-deliberation).  
+The latent opinion of participant $i$ at time $t$ is drawn from a normal 
+distribution in the model, 
 \begin{equation} 
@@
 \noindent 
-The initial opinion of agent $i$ is are denoted by $o_{i,t=0}$, and thus
-in the model are drawn from a normal distribution with mean $\mu_0$ and standard
-deviation $\sigma_0$. The final opinion of agent $i$ is written $o_{i,t=T}$,
-drawn from a normal distribution with mean $\mu_T$ and standard deviation
-$\sigma_T$.
+Latent opinions are assumed to be normally distributed since they are unbounded 
+aggregations of neurobiological activity.
 
-\subsection{Root-finding algorithm to identify plausible false detections}
+The measurement process with $N\to\infty$ participants is simulated by 
+integrating the latent opinion distribution between ordinal bin thresholds, 
+of which there are $B+1$, where $B$ is the number of opinion bins, with sequential
+integer values $b$. In the Moscovici and Zavalloni (1969) example, $b \in
+\{-3,-2,\ldots,2,3\}$ and $B = 7$.
+Bin $b$ is defined by its lower and upper continuous-valued thresholds, 
+$\theta_{b-1}$ and
+$\theta_{b}$, respectively. Thresholds are given by the following equations
+(and visualized as dotted lines in Figure~\ref{fig:distros}).
+\begin{equation}
+\begin{cases}
+  \theta_0 &= -\infty \\
+  \theta_{B} &= \infty \\
+  \theta_b &= b + \frac{1}{2} \text{ if } 0 < b < B.
+\end{cases}
+\end{equation}
+\noindent
+With thresholds defined, the frequency of ordinal measurment $b$ at time $t$ 
+is (Figure~\ref{fig:distros})
+\begin{equation}
+  p(b;~\mu_t, \sigma_t, B) = \int_{\theta_{b-1}}^{\theta_b} p(o; \mu_t, \sigma_t)
+  do.
+  \label{eq:ordinal-frequency}
+\end{equation}
 
-Simple consensus can appear as group polarization since a change in
+We can now define spurious group polarization and give examples of when it occurs.
+First, the observed group polarization opinion shift is
+\begin{equation}
+  s = \meanobs_\mathrm{post} - \meanobs_\mathrm{pre}~,
+  \label{eq:group-polarization}
+\end{equation}
+\noindent
+where the expected value of observed opinions at time $t$ is
+\begin{equation}
+  \meanobst = \frac{1}{B} \sum_{b=1}^B b \cdot p(b; \mu_t, \sigma_t, B).
+  \label{eq:meanobs}
+\end{equation}
+\noindent
+\emph{Spurious} group polarization is when mean observed opinions seem to increase in 
+extremity when the latent mean held constant. We simulate spurious
+group polarization by setting $\mupre = \mupost = \mu$ and setting the variances
+to values known to induce spurious group polarization. 
+We explain how to find a constant latent mean, $\mu$, and pre- and post-deliberation
+variances, $\sigmapre$ and $\sigmapost$, respectively, that generate spurious 
+detections of group polarization in the next section. 
+% We describe how we find $\mu$,
+% $\sigmapre$ and $\sigmapost$ next in Section~\ref{sec:finding-variances}.
+
+\subsection{Web app controls algorithm to find parameters inducing spurious group polarization}
+
+To calculate the false discovery rate for each experimental condition, we first must
+identify a constant latent mean, $\mu$, and latent variances $\sigmapre$ and
+$\sigmapost$ that generate observation distributions\mt{define earlier?} with
+reported mean pre- and post-deliberation opinions and resulting shift $s$ (calculated
+using Equation~\ref{eq:group-polarization}). Define the collection of these
+parameters $\beta = \{\mu, \sigmapre, \sigmapost\}$. We identify $\beta$ through a
+hillclimbing optimization algorithm that finds $\beta$ that minimizes the difference
+between reported pre- and post-deliberation mean opinions and simulated ones
+calculated with Equation~\ref{eq:meanobs}.
+
+The algorithm to find $\beta$ holds $\mu$ constant while searching for $\sigmapre$
+and $\sigmapost$ that generate observed $\meanobspre$ and $\meanobspost$, within a
+tolerance of $\pm10^{-3}$ (Algorithm~\ref{alg:hillclimbing}). 
+The algorithm requires a researcher to specify an initial
+guess for $\beta$ and auxiliary parameter settings, which generally vary between
+experimental conditions.  Furthermore, the algorithm does not always produce
+reasonable distributions with a given $\beta$, and sometimes it does not produce
+spurious group polarization.  Therefore, it is necessary to sanity check the latent
+and ordinal opinion distributions for the $\beta$ found by the algorithm. If all is
+well, the experimental condition can be marked as plausibly spurious.  To tame the
+complexity involved in data entry and model fitting for each of these sixty
+experimental conditions, we developed a web application to input data from control
+the algorithm and inspect its results (Figure~\ref{fig:webapp}). Web application code is
+available online in our GitHub
+repository\footnote{\url{https://github.com/mt-digital/gp-statmod/blob/main/GroupPolarizationStatmod/app.R}}.
 
 \begin{algorithm}
-  \caption{User-guided routine to find pre- and post-deliberation variances that
-  yield group polarization in the absence of mean opinion change.
-\mt{NOT YET DONE...}}\label{alg:routine}
+  \caption{Routine to find latent mean and variance that generate spurious group polarization.}
+  \label{alg:hillclimbing}
   \begin{algorithmic}
-    \Require $\mu$, $\sigmapre$, $\sigmapost$ \Comment{Static guess for common mean,
-    initial guesses of variances.}
-    \Require $\epsilon$, $\delta$, $i_\mathrm{max}$ \Comment{Root-finding tolerance, step size, and maximum
-    iterations.}
-    % \Require $\delta$ \Comment{Root-finding step size.}
-    % \Require $i_\mathrm{max}$ \Comment{Maximum number of iterations.}
-    \State $\Delta \gets \mupre' - \mupost'$ 
-    \While {$\epsilon < |\Delta| $ \and $i < i_max$}
-      \State $\Delta \gets (\mupre' - \mupost')^2$
-    \EndWhile
+
+    \Require $\mu$, $\sigma$, $\bar{\mathbf{o}}$ 
+      \Comment{Static guess for common mean, initial guess of variance.}
+
+    \Require $\epsilon$, $\delta$, $i_\mathrm{max}$ 
+      \Comment{Hillclimbing tolerance, step size, and maximum iterations.}
+
+    \State $\Delta \gets \meanobs(\mu, \sigma) - \meanobsemp$ 
+      \Comment{Initialize simulation error.}
+    \While {$\Delta^2 > \epsilon$ and $i < i_\mathrm{max}$} \Comment{Search until desired
+    accuracy or max iterations reached.}
+      \State $\sigma' \sim \mathcal{N}(\sigma, \delta)$ \Comment{Draw new variance to test.}
+      \State $\Delta' \gets \bar{o}(\mu, \sigma') - \bar{\mathbf{o}}$
+      \If {$\Delta' - \Delta < 0$}
+        \State $\sigma \gets \sigma'$
+          \Comment{Update $\sigma$ if the error is less than before.}
+      \EndIf
+      \State $\Delta \gets \bar{o}(\mu, \sigma) - \bar{\mathbf{o}}$
+        \Comment{Update simulation error if $\sigma$ is updated.}
+    \EndWhile \\
+
+    \Return $\sigma$
   \end{algorithmic}
 \end{algorithm}
+
+\begin{figure}
+  \caption{\textbf{Shiny web application for user-guided hillclimbing to find
+  parameters that generate spurious group polarization.} This web application enables users to
+create new analyses of experimental conditions from published studies. Users 
+provide initial guesses for the algorithm to use as starting points in 
+finding variances that generate spurious group polarization. Users inspect 
+model fits provided by the algorithm, adjusting parameters if necessary, and
+mark an experimental condition as ``Plausibly spurious'' if a reasonable fit
+is indeed found.\mt{Maybe add some A, B, C guides to the figure and include
+in parentheses here to go with the text?}}
+  \centering
+    \includegraphics[width=0.75\textwidth]{Figures/Model/WebApp.png}
+  \label{fig:webapp}
+\end{figure}
+
+Note that so far, we have only identified cases where spurious group polarization plausibly
+occurs in the asymptotic case where $N\to\infty$ ($N$ being the number of participants).
+Experimental conditions yielding plausibly spurious group polarization warrant 
+some suspicion. If there exists set of latent simple consensus 
+parameters $\beta$ that create binned (Equation~\ref{eq:ordinal-frequency})
+opinion distributions that naively seem to be group polarized, then the means
+of at least some samples from these distibutions with empirical numbers of 
+participants would also appear to be shifted consistent with group
+polarization. This process itself does not tell us the false detection rate.
+We use the identified simple consensus parameters to define pre- and post-deliberation
+latent opinion distributions in simulated trials of group polarization 
+experiments.
 

### 2024-12-19 GroupPolarizationStatmod/app.R

@@ -36,7 +36,7 @@ ui <- function(request) { fluidPage(
     
     "))),
   
-    titlePanel("Group polarization counterexample generator."),
+    titlePanel("Spurious group polarization generator."),
 
     h3(textOutput("treatmentTag")),
 
@@ -89,7 +89,7 @@ ui <- function(request) { fluidPage(
                                          0.05, step = 0.005)
                         )
                     ),
-                    checkboxInput("Plausible", "Plausibly false positive?", FALSE)
+                    checkboxInput("Plausible", "Plausibly spurious?", FALSE)
                 ),
                 column(12,
                     textAreaInput("Notes", "Notes", 
@@ -136,7 +136,7 @@ server <- function(input, output, session) {
     output$MinBinValue <- reactive({treatmentRow()$MinBinValue})
     output$MaxBinValue <- reactive({treatmentRow()$MaxBinValue})
     output$treatmentTag <- renderText({
-        paste("Currently analyzing treatment:", input$treatmentTag)
+        paste("Currently analyzing ExperimentID:", input$treatmentTag)
     })
 
     # Use a guess of 1.5 for latentSD.

### 2024-12-19 GroupPolarizationStatmod/scripts/plot/model.R

@@ -21,12 +21,13 @@ source("model.R")  # for binProb function used in ordinal plot function below.
 
 plot_latent_pdf_integration = function(mu = 0, sd = 2.5, min_bin, max_bin, 
                                        bins, bin_colors, # xlim = c(-6.0, 6.0), 
-                                       save_path = "test_latent_pdf.pdf") {
+                                       save_path = "test_latent_pdf.pdf", 
+                                       max_prob = 1.0) {
   print(save_path)
 
 
-  xmin = min_bin - 1
-  xmax = max_bin + 1
+  xmin = min_bin - 0.5
+  xmax = max_bin + 0.5
 
   xlim = c(xmin, xmax)
 
@@ -63,13 +64,13 @@ plot_latent_pdf_integration = function(mu = 0, sd = 2.5, min_bin, max_bin,
                         geom = "area",
                         args = c(mu, sd),
                         fill = bin_colors[n_bins],
-                        xlim = c(xmax, max_bin + 0.5))
+                        xlim = c(xmax, max_bin - 0.5))
 
   p = p + xlab("Latent opinion") + ylab("Probability density\n\n") + 
   geom_vline(xintercept=c((bins[1:(length(bins) - 1)] + 0.5)), linetype="dotted") + 
   geom_vline(xintercept=mu, linetype="dashed") + 
   xlim(c(xmin, xmax)) + 
-  ylim(c(0, 1)) +
+  ylim(c(0, max_prob)) +
   scale_x_continuous(breaks = bins) +
   mytheme
   
@@ -84,7 +85,8 @@ plot_ordinal_distribution = function(mu = 0, sd = 2.5, min_bin = -2,
                                      bin_colors = c("lightpink", "lightblue", 
                                                       "#f7ae3d", "lightgreen", 
                                                       "#F6CFFF"),
-                                     save_path = "test_ordinal_pdf.pdf") {
+                                     save_path = "test_ordinal_pdf.pdf",
+                                     max_prob = 1.0) {
   
   # Use same parameters as latent distribution to calculate integration over
   # bin threshold limits \theta_{k-1}..\theta_k
@@ -108,8 +110,8 @@ plot_ordinal_distribution = function(mu = 0, sd = 2.5, min_bin = -2,
[…]

### 2024-12-21 GroupPolarizationStatmod/paper/ms.tex
model/methods good enough to move forward

@@
 
-the ten-article corpus, which is representative of the
-false discovery rate for group polarization studies.
+Unchecked social and political polarization undermines cooperation and
+political efficacy \autocite{Mason2018UncivilAgreementBook,Klein2020}.
+\emph{Group polarization} is the name given the social phenomenon of 
+when a group becomes more extreme in mean opinion over
+time, assuming the mean opinion is biased to one extreme to begin with.
+\mt{now hit them with the real shit to intro the problem, our analysis, and
+how this can be used constructively going forward.}
 
-Unchecked social and political polarization undermines cooperation and political
-efficacy \autocite{Mason2018UncivilAgreementBook,Klein2020}.  \emph{Group
-polarization} is the name that has been given to a group-level process that leads
-to greater opinion extremity in a group as the group deliberates some topic (e.g.,
-``Do you approve, disapprove, or neither approve nor disapprove of the current
-administration's policies?''). Extremity only increases, though, assuming that the
-mean opinion of the group is initially biased towards one extreme or the other
-\autocite{Brown1986,Brown2020}. Group polarization is important to understand
-rigorously, therefore, since if groups become increasingly extreme over time,
-while group opinions are consolidated, this would exacerbate political
-polarization between political opposition groups, which is already undermines the
-ability of partisans in democratic governments to work together. Dozens of studies
-and meta-analyses over decades have apparently established the empirical reality
-of group polarization
+\mt{Seems like a good second paragraph structurally. We should give more
+details about the scope of the study, group polarization and common
+elements of experimental designs.} 
+Group polarization experiments tend to employ a common set of design
+elements incorporated in our computational model analysis. Group
+polarization experiments mostly use ordinal (see exceptions
+\cite{Blascovich1974,Blascovich1975,Blascovich1976,Keating2016}). The opinions (e.g., ``Do you approve,
+disapprove, or neither approve nor disapprove of the current
+administration's policies?''). Extremity only increases, though, assuming
+that the mean opinion of the group is initially biased towards one extreme
+or the other \autocite{Brown1986,Brown2020}. Group polarization is important
+to understand rigorously, therefore, since if groups become increasingly
+extreme over time, while group opinions are consolidated, this would
+exacerbate political polarization between political opposition groups, which
+is already undermines the ability of partisans in democratic governments to
+work together. Dozens of studies and meta-analyses over decades have
+apparently established the empirical reality of group polarization
 \autocite{Moscovici1969,Myers1970,Isenberg1986,Sunstein2009,Sunstein2019}
-\autocite[Ch. 5]{Brown2000}. We believe group polarization should occur in certain
-contexts since more extreme group members are likely also to be more stubborn
-\autocite{Guazzini2015,Lewandowsky2019}, which should pull consensus to extremes
-\autocite{Turner2020}. However, many or most of the evidence for group
-polarization relies on experimental designs that use ordinal measurements that are
-then analyzed as if they were continuous. Unfortunately, this procedure is
-unfortunately known to induce false positives if it is not properly accounted for
-\autocite{Liddell2018}.
+\autocite[Ch. 5]{Brown2000}. We believe group polarization should occur in
+certain contexts since more extreme group members are likely also to be more
+stubborn \autocite{Guazzini2015,Lewandowsky2019}, which should pull
+consensus to extremes \autocite{Turner2020}. However, many or most of the
+evidence for group polarization relies on experimental designs that use
+ordinal measurements that are then analyzed as if they were continuous.
+Unfortunately, this procedure is unfortunately known to induce false
+positives if it is not properly accounted for \autocite{Liddell2018}.
 
@@
   \fdr = \frac{(1 - b)\alpha}{(1 - b)\alpha + bW}.
+  \label{eq:fdr}
 \end{equation}
@@
 \begin{equation}
-\begin{cases}
-  \theta_0 &= -\infty \\
-  \theta_{B} &= \infty \\
-  \theta_b &= b + \frac{1}{2} \text{ if } 0 < b < B.
+\theta_b = \begin{cases}
+  -\infty         & \text{ if } b=0 \\
+  \infty          & \text{ if } b=B \\
+  b + \frac{1}{2} & \text{ otherwise.}
 \end{cases}
@@
     \Require $\mu$, $\sigma$, $\bar{\mathbf{o}}$ 
-      \Comment{Static guess for common mean, initial guess of variance.}
+      \Comment{Set latent mean, guess variance that yields, input empirical mean.}
 
@@
       \State $\Delta' \gets \bar{o}(\mu, \sigma') - \bar{\mathbf{o}}$ 
+        \Comment{Calculate error using new variance.}
       \If {$\Delta' - \Delta < 0$}
@@
 
-\subsection{Calculate false detection rate from simulations}
+\subsection{Outcomes: Estimating and limiting false detection rates}
 
-False detection rates are calculated using Cohen's $d$ to measure effect
-size, defined as~\cite[p. 331]{Liddell2018}
+To calculate the false detection rate we begin by simulating 
+pre- and post-deliberation 
+opinions by drawing $N_e$ samples from normal latent opinion
+distributions that generated spurious group polarization, i.e., with
+parameters labelled $\beta_e$ found in the hillclimbing step. Binning these
+random values simulates observations. $N_e$ is the number
+of participants in experimental condition $e$. Each trial $i$ of $N_e$
+simulated pre- and post-deliberation observations is then inspected to see
+if spurious group polarization occurred in the simulation. Spurious group
+polarization is determined by first inferring the latent parameters of each
+simulated experimental trial by fitting an ordered probit model. 
+The inferred parameters are then used to calculate the effect size for that
+trial, Cohen's $d$, defined as~\cite[p. 331]{Liddell2018}
 \begin{equation}
   d = \frac{\mupost - \mupre}{\sqrt{\frac{1}{2}(\sigmapost^2 + \sigmapre^2)}}.
-\end{equation}. 
+\end{equation}
 \noindent
-Denote the collection of fitted parameters from the $i^\mathrm{th}$ simulation trial
-for experimental condition $c$ and journal article $a$ as
-% $\thetafit = \left{\mu_t = \left{\mupre,~\mupost\right},~\sigma_t = \left{\mupre,~\mupost\right}\right}$.
-$\thetafit_{a,c,i} = \{\mupre,~\mupost,~\mupre,~\mupost\}_{a,c,i}$.
-These parameters are the modal estimates for each parameter obtained from
-an ordered probit statistical model fit with
-Bayesian Markov-Chain Monte Carlo (MCMC)~\cite[p. 338]{Liddell2018}.
+The family-wise error rate is, $\alpha = \Pr(D|\neg E)$, is the 
+fraction of trials with $d > d^*$. To calculate the family-wise error rate
+for a study, $s$ which contains several experimental conditions, $e$,
+\begin{equation}
+  \alpha_s = \frac{1}{|s|} \sum_{e \in s} \alpha_e
+\end{equation}
+\noindent
+The grand mean across all studies for the corpus is 
+\begin{equation}
+  \alpha_\mathrm{all} = \frac{1}{|E|} \sum_{e \in E} \alpha_e.
+\end{equation}
+With $\alpha$ calculated, we calculate the false discovery rate using
+Equation~\ref{eq:fdr}.
 
-For each $a,c$, we calculate $\alpha_{a,c}(d^*) =
-\Pr(D |\neg E;~a,~c,~d^*)$ for a given significance value, $d^*$, directly as
-the fraction of $d$ values that are greater than $d^*$.
-
-To calculate article-level false detection curves we calculate the average
-across , we set $\alpha_{a,c} = 0.0$ for those
-conditions for which we failed to generate spurious group polarization in the
-large-$N$ model.
+We invert this procedure to find what significance value $d^*$ is necessary
+to achieve a low family-wise error rate $\alpha_e = 0.05$ for experimental
+condition, $e$. To do this we first calculate $\alpha_e(d^*)$ for a
+sufficiently wide range of $d^*$. We then find the lowest $d^*$ for which
+$\alpha_e(d^*) \leq 0.05$.
 
@@
 
+Our analysis brings some additional questions into focus. Most
+significantly, recall that even the commensurate ordered probit Bayesian
+model often failed to accurately detect negative results. A more rigorous
+analysis would include additional sources of variance that will likely
+further inflate the family-wise error rate, $\alpha$, and the false
+discovery rate~\cite{Yarkoni2022}. First and foremost, there is variance
+within each experimental deliberation group; no group polarization study
+accounts for this, instead treating participants in each experimental
+condition as if they were one large group. Furthermore, survey data are
+known to be noisy~\cite{Zaller1992}. People report opinions differently
+over time for no apparent reason (i.e., opinions are \emph{unstable}).
+Context matters: the order in which a survey question is asked, and which
+question framing is chosen among logically equivalent alternatives, can
+both significantly influence participant responses. The effect of 
+accounting for these sources of variance must be understood to estimate
+$\alpha$ (and power, $W$) for the design of the next generation of group
+polarization experiments.
+
 \mt{concluding paragraph: first sentences summarize the Discussion, then

### 2024-12-28 GroupPolarizationStatmod/paper/ms.tex
calculating quantiles and writing analysis

@@
 \section{Introduction}\label{introduction}
+
+If an extremist opinion falls and it is measured with a Likert scale, will it make a noise?
 
@@
 \subsection{How study design can induce false detections of group polarization}
+\label{subsec:false-detections}
 
@@
 
-\subsection{Outcomes: Estimating and limiting false detection rates}
+\subsection{Estimating and limiting false detection rates}
 
@@
 polarization is determined by first inferring the latent parameters of each
-simulated experimental trial by fitting an ordered probit model. 
+simulated experimental trial by fitting an ordered probit model~\cite{Liddell2018}. 
 The inferred parameters are then used to calculate the effect size for that
@@
   d = \frac{\mupost - \mupre}{\sqrt{\frac{1}{2}(\sigmapost^2 + \sigmapre^2)}}.
+  \label{eq:cohens}
 \end{equation}
 \noindent
-The family-wise error rate is, $\alpha = \Pr(D|\neg E)$, is the 
-fraction of trials with $d > d^*$. To calculate the family-wise error rate
+The family-wise error rate is, $\alpha_{e} = \Pr(D|\neg E)$, is the 
+fraction of trials, 
+with $d_{ei} > d^*$ for the $i^\text{th}$ trial for experimental
+condition $e$, written
+\begin{equation}
+  \alpha_e = \frac{1}{N_T} \sum_i^{N_T} \Theta_i,
+\end{equation}
+\noindent
+where
+\[
+  \Theta_i = 
+    \begin{cases}
+      1 & \text{ if } d_{ei} > d^* \\
+      0 & \text{ otherwise.}
+    \end{cases}
+\]
+and $N_T = 1000$ is the number of trials.
+
+To calculate the family-wise error rate
 for a study, $s$ which contains several experimental conditions, $e$,
@@
 
-\section{Results}\label{results}
+\section{Analysis}\label{analysis}
 
-\subsection{Ninety percent of detections identified as plausibly false}
+Our analysis estimates how the probability that a published detection of group
+polarization is actually a false detection. We first identified latent opinion
+parameters that yield spurious group polarization in an asymptotic $N\to\infty$
+model. This gives us our first suggestion that experimental conditions may be
+plausibly false. Briefly, we identified latent opinion distribution parameters
+that generated spurious group polarization in 54 out of 57 experimental
+conditions. In other words, 95\% of all group polarization detections are
+plausibly spurious due to ceiling artifacts from the continuous latent
+opinions-ordinal measurement experimental design. We then analyze family-wise
+error and false detection rates for each experimental condition and aggregated 
+up to the study level and across all studies and experimental conditions. 
+Family-wise error rates varied from a minimum of 0.097 for the
+\texttt{Feminists-Experimental} condition from Myers (1975)~\cite{Myers1975a}
+and a maximum of 0.67 for the \texttt{COSprings-CivilUnions} condition from
+Schkade, Sunstein, and Hastie (2010)~\cite{Schkade2010}. The median family-wise error
+rate is 0.28 for Krizan and Baron's experimental condition
+\texttt{NoOutgroupScenario1}. We close by inverting our procedure of calculating
+family-wise error rates to calculating what $d^*$ would be required in each of the 54
+experimental conditions to reduce the family-wise error rate to $\alpha = 0.05$.
 
-We identified 54 of 60 experimental 
-conditions where a $N \to \infty$ model falsely indicated group polarization 
-instead of simple consensus.
+\subsection{Nearly all group polarization detections are plausibly false}
 
-\subsection{False detection rates exceed ??\% even at ``high significance''}
+We used our formal distribution model of group polarization experiments
+to identify plausible latent simple consensus distributions that generate spurious
+group polarization in observed ordinal distributions 
+in the asymptotic case of taking $N \to \infty$ ordinal measurements.
+Across the ten studies we analyzed, we excluded 3 out of 60 experimental
+conditions that did not claim to be positive findings of group polarization.
+Of the remaining 57, we identified plausible $\beta$ that generate spurious
+group polarization in 54 experimental conditions. In the other three, we did
+identify $\beta$ that generate spurious group polarization, but the initial
+simulated distributions were highly polarized. Two of these
+simulations of experimental conditions from Schkade, et al., (2010) because
+the histogram plots of the observed opinion distributions are not
+polarized~\cite[Figure 1, p. 234]{Schkade2010}. The other condition where we
+rejected candidate $\beta$ was in a question about a ``bad teacher'' in Myers
+(1975)~\cite{Myers1975a}, where lower values signify increasingly 
+worse impressions of the teacher and higher values signify increasingly good
+impressions. It is unlikely that any participants would rate a ``bad teacher''
+as good, so we excluded this from our analysis as well. This exclusion may be
+overly generous since Myers (1975) did not provide details beyond mean values
+and opinion shifts, and the article was published before the importance
+of open data best practices was recognized.
+
+\clearpage
+
+\begin{table}[ht]
+\rotatebox{90} {
+  \input{tex/quantile_table.tex}
+}
+\end{table}
+
+\subsection{Half of experiments exceed a 75\% false discovery rate at ``medium'' significance}
+
+The $N\to\infty$ model enabled us to identify pre- and post-deliberation latent
+opinion distribution parameters that caused spurious group polarization, denoted
+$\beta_e$ for experimental condition $e$. Using these $\beta$ we drew 
+$N_\mathrm{emp}$ samples from the pre- and post-deliberation latent opinion 
+distributions, then binned them (Section~\ref{subsec:false-detections}) 
+to simulate participants reporting their pre- and post-deliberation opinions. 
+To estimate optimistic family-wise error and false detection rates, we fit an
+ordered probit model to the simulated observations and obtained estimates of
+the group polarization effect size, $d$ (Equation~\ref{eq:cohens}).
+If the effect size was greater than a significance value, $d > d^*$, then there
+was a detection $D$ of group polarization, even though by design the simulated latent mean
+group opinion was constant, i.e., there in fact was no effect, written $\neg E$.
+The family-wise error rate is the calculation of $\Pr(D|\neg E)$ as the frequency with which
+$d > d^*$ out of 1000 simulation trials. The false detection rate is the family-wise
+error rate scaled using the base rate of effects and the statistical power
+(Equation~\ref{eq:fdr}). We test three difference significance values,
+$d = 0.2,~0.5$ and $0.8$, corresponding to what Cohen prescribed to use for
+``low'', ``medium'', and ``high'' significance, respectively.
 
@@
 
-Our analysis brings some additional questions into focus. Most
-significantly, recall that even the commensurate ordered probit Bayesian
-model often failed to accurately detect negative results. A more rigorous
-analysis would include additional sources of variance that will likely
-further inflate the family-wise error rate, $\alpha$, and the false
-discovery rate~\cite{Yarkoni2022}. First and foremost, there is variance
-within each experimental deliberation group; no group polarization study
-accounts for this, instead treating participants in each experimental
-condition as if they were one large group. Furthermore, survey data are
-known to be noisy~\cite{Zaller1992}. People report opinions differently
-over time for no apparent reason (i.e., opinions are \emph{unstable}).
-Context matters: the order in which a survey question is asked, and which
-question framing is chosen among logically equivalent alternatives, can
-both significantly influence participant responses. The effect of 
-accounting for these sources of variance must be understood to estimate
-$\alpha$ (and power, $W$) for the design of the next generation of group
+Our analysis brings some additional questions into focus. Most significantly, recall that even
+the more appropriate ordered probit Bayesian model often failed to accurately detect negative
+results with the number of participants used in the original papers. Many of the studies
+analyzed here did not use any statistical test to establish significance; those that did used
+some form of a $t$-test, and it is not clear if the $t$-test that was used 
+accounted for a change in pre- and post-deliberation opinion variance, which we expect due to
+consensus.
+
+The ordered probit model we used still likely underestimates the variance in outcomes, which
+could require even larger sample sizes. A more rigorous analysis
+would include additional sources of variance that will likely further inflate the Type I
+error rate, $\alpha$, and the false discovery rate~\cite{Yarkoni2022}. First and foremost,
+there is variance within each experimental deliberation group; no group polarization study
+accounts for this, instead treating participants in each experimental condition as if they were
+one large group. Furthermore, survey data are known to be noisy~\cite{Zaller1992}. People
+report opinions differently over time for no apparent reason (i.e., opinions are
+\emph{unstable}).  Context matters: the order in which a survey question is asked, and which
+question framing is chosen among logically equivalent alternatives, can both significantly
+influence participant responses. The effect of accounting for these sources of variance must be
+understood to estimate $\alpha$ (and power, $W$) for the design of the next generation of group
 polarization experiments.

### 2024-12-28 GroupPolarizationStatmod/scripts/analysis.R

@@ -15,7 +15,8 @@
 library(purrr)
 
 
-source("src/false_detection_rate.R")
+source("src/fwer.R")
+source("src/util.R")
 
 
 # b for base rate, W for power. See 
@@ -27,57 +28,6 @@ fdr = function(fwer, b, W) {
   return (num / denom)
 }
 
-
-calc_fdr_vs_significance = function(
-  probit_fits_dir = "data/probit_fits", base_rate = 0.1, power = 0.8, 
-  significance_vals = c(0.2, 0.5, 0.8)
-) {
-
-  probit_fits_data = load_probit_data()
-  n_experiments = length(unique(probit_fits_data$ExperimentID))
-
-  # Initialize return tibble.
-  ret_tbl = tibble(StudyID = character(n_experiments),
-                   ExperimentID = character(n_experiments),
-                   FWER = numeric(n_experiments),
-                   FDR = numeric(n_experiments),
-                   SigVal = numeric(n_experiments),
-                   Power = numeric(n_experiments),
-                   BaseRate = numeric(n_experiments))
-  
-  # Calculate FWER for each significance value.
-  ii = 1
-  while (ii <= length(significance_vals)) {
-    
-    # Calculate row indices to be updated in return tibble.
-    low_rng = ((ii-1) * n_experiments + 1)
-    high_rng = (ii * n_experiments)
-    update_rng = low_rng:high_rng
-
-    # Extract this significance value for readability.
-    significance_val = significance_vals[ii]
-
-    # Get tibble with StudyID, ExperimentID, and FWER columns...
-    fwer_tbl = calculate_fwer(probit_fits_data, significance_val)
-    # ...to calculate FDR and create a combined table.
-    fdr_col = tibble(FDR = fdr(fwer_tbl$FWER, base_rate, power))
-    fwer_fdr = bind_cols(fwer_tbl, fdr_col)
-                
-    # Update the return tibble with this significance_val.
-    ret_tbl[update_rng, ] = 
-      bind_cols(
-        fwer_fdr,
-        # Three metadata columns.
-        tibble(SigVal = rep(significance_vals[ii], n_experiments),
-               Power = rep(power, n_experiments),
-               BaseRate = rep(base_rate, n_experiments))
-      )
-    
-    ii = ii + 1
-  }
-
-  return (ret_tbl)
-}
 
 
 add_default_rows = function(fdr_vs_sig_tbl, default_fwer = 0.05) {
[…]

### 2024-12-28 GroupPolarizationStatmod/scripts/calculations.R

@@ -1,0 +1,82 @@
+##
+# Calculations that are included in paper prose.
+#
+# Author: Matthew A. Turner
+# Date: 2024-12-27
+#
+
+
+source("src/util.R")
+
+
+fraction_at_or_above = function(measure = "FWER", significance_val = 0.5, 
+                                base_rate = 0.1, power = 0.8, value = 0.75) {
+
+  # Load data. Copied from scripts/plot/analysis.R
+  tbl <- calc_fdr_vs_significance(significance_vals = c(significance_val)) 
+
+  return (mean(tbl[, measure] >= value))
+}
+
+
+# Create a CSV list of quantile calculations that may be included in publications.
+quantiles_tbl = function(tex_output_file = "paper/tex/quantile_table.tex", 
+                         q_resolution = 0.05, 
+                         significance_vals = c(0.2, 0.5, 0.8),
+                         base_rate = 0.1, power = 0.8
+                         ) {
+  
+  probs <- seq(0.0, 1.0, q_resolution)
+
+  initialized = FALSE
+  for (sigval in significance_vals) {
+
+    tbl <- calc_fdr_vs_significance(significance_vals = sigval)
+
+    for (measure in c("FWER", "FDR")) {
+
+      # Get quantile for this measure, base rate, and power.
+      q <- quantile(as_vector(tbl[, measure]), probs = probs)
+
+      if (!initialized) {
+        # Load quantile named list into tbl with quantiles as colnames sans "%".
+        q_tbl <- tibble(!!!q, .name_repair = ~ str_replace(., "%", ""))
+        q_tbl$BaseRate <- base_rate
+        q_tbl$Power <- power
+        q_tbl$SigVal <- sigval
+        q_tbl$Measure <- measure
+
+        initialized = TRUE
+      }
+      else {
+        # Create new row.
+        new_row <- tibble(!!!q, .name_repair = ~ str_replace(., "%", ""))
+        new_row$BaseRate <- base_rate
+        new_row$Power <- power
+        new_row$SigVal <- sigval
+        new_row$Measure <- measure
+
+        # Bind new row to existing.
+        q_tbl <- rbind(q_tbl, new_row)
+      }
+    }
+  }
+
+  # Reorder columns to be easier to read.
+  q_tbl <- 
+    q_tbl %>% 
+      relocate(Measure, .before = `0`) %>% 
[…]

### 2024-12-28 GroupPolarizationStatmod/scripts/plot/analysis.R

@@ -14,7 +14,7 @@ library(ggplot2)
 library(latex2exp)
 library(emojifont)
 
-source("src/false_detection_rate.R")
+source("src/fwer.R")
 source("scripts/analysis.R")
 
 mytheme = theme(
@@ -58,7 +58,6 @@ plot_ordinal_cohens <- function(ordinal_data_dir = "data/probit_fits",
 
 plot_fdr = function(save_path = "paper/Figures/fdr_vs_experiment.pdf", 
                     aggregate = FALSE, use_non_identified = FALSE,
-                    probit_fits_dir = "data/probit_fits",
                     default_fwer = 0.05,
                     base_rate = 0.1, power = 0.8, 
                     significance_vals = c(0.2, 0.5, 0.8),
@@ -67,7 +66,6 @@ plot_fdr = function(save_path = "paper/Figures/fdr_vs_experiment.pdf",
   # Load data.
   tbl = calc_fdr_vs_significance(significance_vals = significance_vals) %>%
     mutate(ExperimentID = paste(StudyID, ExperimentID, sep = "_")) %>%
-    # select(!c(StudyID)) %>% 
     mutate(SigVal = as_factor(SigVal), Power = power, BaseRate = base_rate)
 
   if (aggregate)

### 2024-12-28 GroupPolarizationStatmod/src/fwer.R

@@ -5,11 +5,11 @@
 # Date: 2024-10-30
 #
 
-# library(tidyverse)
 library(dplyr)
-library(fs)
 library(readr)
 library(tidyr)
+
+source("src/util.R")
 
 
 cohens_d <- function(pre_mean_estimate, post_mean_estimate, pre_sd, post_sd) {
@@ -19,26 +19,6 @@ cohens_d <- function(pre_mean_estimate, post_mean_estimate, pre_sd, post_sd) {
   denominator = sqrt((pre_sd**2 + post_sd**2) / 2.0)
   
   return (numerator / denominator)
-}
-
-
-load_probit_data = function(probit_fits_dir = "data/probit_fits", 
-                            sync_file = "data/probit_fits/all.csv",
-                            overwrite = FALSE) {
-     
-  if (!file.exists(sync_file) || overwrite) {
-
-    ret <- dir_ls(probit_fits_dir, glob = "*.csv") %>%
-      read_csv() %>% unite(ExperimentID, ArticleTag, TreatmentTag)
-
-    write_csv(ret, sync_file)
-
-  } else {
-
-    ret <- read_csv(sync_file)
-  }
-
-  return (ret)
 }
 
 

### 2024-12-28 GroupPolarizationStatmod/src/util.R

@@ -1,0 +1,32 @@
+##
+# Utility functions used in other src and scripts files.
+#
+# Author: Matthew A. Turner
+# Date: 2024-12-27
+#
+
+library(dplyr)
+library(readr)
+library(tidyr)
+library(fs)
+
+load_probit_data = function(probit_fits_dir = "data/probit_fits", 
+                            sync_file = "data/probit_fits/all.csv",
+                            overwrite = FALSE) {
+     
+  if (!file.exists(sync_file) || overwrite) {
+
+    ret <- dir_ls(probit_fits_dir, glob = "*.csv") %>%
+      read_csv() %>% unite(ExperimentID, ArticleTag, TreatmentTag)
+
+    write_csv(ret, sync_file)
+
+  } else {
+
+    ret <- read_csv(sync_file)
+  }
+
+  return (ret)
+}
+
+

### 2024-12-30 GroupPolarizationStatmod/paper/ms.tex
eoy goalz

@@
 plausibly spurious due to ceiling artifacts from the continuous latent
-opinions-ordinal measurement experimental design. We then analyze family-wise
-error and false detection rates for each experimental condition and aggregated 
+opinions-ordinal measurement experimental design. We then analyze the family-wise
+error ($\alpha$) and false detection rate (FDR) 
+for each experimental condition and aggregated 
 up to the study level and across all studies and experimental conditions. 
-Family-wise error rates varied from a minimum of 0.097 for the
-\texttt{Feminists-Experimental} condition from Myers (1975)~\cite{Myers1975a}
-and a maximum of 0.67 for the \texttt{COSprings-CivilUnions} condition from
-Schkade, Sunstein, and Hastie (2010)~\cite{Schkade2010}. The median family-wise error
-rate is 0.28 for Krizan and Baron's experimental condition
-\texttt{NoOutgroupScenario1}. We close by inverting our procedure of calculating
-family-wise error rates to calculating what $d^*$ would be required in each of the 54
-experimental conditions to reduce the family-wise error rate to $\alpha = 0.05$.
+The lower twentieth percentile of $\alpha$
+We close by showing that $d^*$ must be greater than the ``high'' effect size
+$d^*=0.8$ to limit family-wise error rates to $\alpha \leq 0.05$ in all but six
+experimental conditions.
 
@@
 
-\clearpage
+\begin{table}[ht]
+  \caption{yo}
+  \label{tab:quantiles}
+\rotatebox{90} {\textbf{Quantiles for the two error-rate measures under different
+  significance values.}\mt{Not sure about the best way to format; probably don't
+  need this in the main text, but could be valuable in supplement?}}
+  \input{tex/quantile_table.tex}
 
-\begin{table}[ht]
-\rotatebox{90} {
-  \input{tex/quantile_table.tex}
-}
 \end{table}
@@
 
-The $N\to\infty$ model enabled us to identify pre- and post-deliberation latent
-opinion distribution parameters that caused spurious group polarization, denoted
-$\beta_e$ for experimental condition $e$. Using these $\beta$ we drew 
-$N_\mathrm{emp}$ samples from the pre- and post-deliberation latent opinion 
-distributions, then binned them (Section~\ref{subsec:false-detections}) 
-to simulate participants reporting their pre- and post-deliberation opinions. 
-To estimate optimistic family-wise error and false detection rates, we fit an
-ordered probit model to the simulated observations and obtained estimates of
-the group polarization effect size, $d$ (Equation~\ref{eq:cohens}).
-If the effect size was greater than a significance value, $d > d^*$, then there
-was a detection $D$ of group polarization, even though by design the simulated latent mean
-group opinion was constant, i.e., there in fact was no effect, written $\neg E$.
-The family-wise error rate is the calculation of $\Pr(D|\neg E)$ as the frequency with which
-$d > d^*$ out of 1000 simulation trials. The false detection rate is the family-wise
-error rate scaled using the base rate of effects and the statistical power
-(Equation~\ref{eq:fdr}). We test three difference significance values,
-$d = 0.2,~0.5$ and $0.8$, corresponding to what Cohen prescribed to use for
-``low'', ``medium'', and ``high'' significance, respectively.
+We simulated trials from each experimental condition, $e$, 
+by sampling and binning pre- and
+post-deliberation latent opinions from distributions with parameters identified
+with the $N\to\infty$ model as causing plausibly spurious group
+polarization. For each $e$,
+we drew a number of samples equal to the number of participants in 
+the original experiments, $N_\mathrm{emp}$. To estimate
+family-wise error and false detection rates, we fit an ordered probit
+model to the simulated observations and obtained estimates of the group polarization
+effect size, $d$ (Equation~\ref{eq:cohens}). If the effect size was greater than a
+significance value, $d > d^*$, then by definition there was a detection $D$ of group 
+polarization,
+even though there was no effect by design, written $\neg E$. 
+The family-wise error rate is $\Pr(D|\neg E)$, calculated as the frequency with which $d > d^*$ out of 1000
+simulation trials. The false detection rate is the family-wise error rate scaled
+using the base rate of effects and the statistical power (Equation~\ref{eq:fdr}). We
+test three difference significance values, $d = 0.2,~0.5$ and $0.8$, corresponding
+to what Cohen prescribed to use for ``low'', ``medium'', and ``high'' significance.
 
-We calculated the false detection rate across experimental conditions where
-we identified parameters that generate spurious group polarization in the 
-large-$N$ model (Figure~\ref{fig:fdr_experiments}). We then aggregated over
-experimental conditions within each study to understand the false detection rate
-across the ten studies (Figure~\ref{fig:fdr_studies}). Across the identified
-spurious conditions, we calculated a mean false detection rate of 0.74 with
-a significance level of $d^* = 0.5$. We also
-examined the mean false detection rates across studies if we set a highly
-optimistic estimate of the family-wise error rate, setting $\alpha = 0.05$. 
-This only reduces the false discovery rate to 0.71, but does slightly change the
-rankings of which study had the highest mean false discovery rates, with 
-Schkade, et al., (2010) going from highest mean false discovery rate to 
-second-lowest false discovery rate. The minimum false discovery rate in a study
-was Myers (1975), with a false discovery rate of 0.58, meaning that probability
-a group polarization detection in the best-designed study is greater than one in
-two.
+Family-wise error rates (Figure~\ref{fig:fwer_fdr_synthesis} and
+Table~\ref{tab:quantiles}) varied from a minimum of one in ten for the
+\texttt{Feminists-Experimental} condition from Myers (1975)~\cite{Myers1975a}
+to a maximum of two in three for the \texttt{COSprings-CivilUnions} condition from
+Schkade, Sunstein, and Hastie (2010)~\cite{Schkade2010}. The median family-wise error
+rate is 0.28 for Krizan and Baron's (2007) experimental condition
+\texttt{NoOutgroupScenario1}~\cite{Krizan2007}.
 

### 2024-12-30 GroupPolarizationStatmod/scripts/calculations.R

@@ -9,16 +9,6 @@
 source("src/util.R")
 
 
-fraction_at_or_above = function(measure = "FWER", significance_val = 0.5, 
-                                base_rate = 0.1, power = 0.8, value = 0.75) {
-
-  # Load data. Copied from scripts/plot/analysis.R
-  tbl <- calc_fdr_vs_significance(significance_vals = c(significance_val)) 
-
-  return (mean(tbl[, measure] >= value))
-}
-
-
 # Create a CSV list of quantile calculations that may be included in publications.
 quantiles_tbl = function(tex_output_file = "paper/tex/quantile_table.tex", 
                          q_resolution = 0.05, 
@@ -80,3 +70,8 @@ quantiles_tbl = function(tex_output_file = "paper/tex/quantile_table.tex",
   return (q_tbl)
 }
 
+
+all_calcuations <- function() {
+  fwer_fdr_synthesis_data()
+  quantiles_tbl()
+}

### 2025-01-01 GroupPolarizationStatmod/paper/ms.tex
nearly done with Analysis

@@
 \begin{equation}
-  \alpha_e = \frac{1}{N_T} \sum_i^{N_T} \Theta_i,
+  \alpha_e = \frac{1}{N_T} \sum_i^{N_T} \Theta(d_{ei}; d^*),
+  \label{eq:study_aggregate}
 \end{equation}
@@
 \[
-  \Theta_i = 
+  \Theta(d_{ei}; d^*) = 
     \begin{cases}
@@
 \]
-and $N_T = 1000$ is the number of trials.
+and $N_T = 1000$ is the number of trials. For simplicity we assume that the sign of
+$d_{ei}$ is positive, meaning a shift to greater extremisim. It is possible that
+$d_{ei}$ could be negative, indicating a shift to lesser extremity.
 
@@
 \noindent
-The grand mean across all studies for the corpus is 
+The mean across all results identified as plausibly spurious is 
 \begin{equation}
@@
 \end{equation}
-With $\alpha$ calculated, we calculate the false discovery rate using
-Equation~\ref{eq:fdr}.
+We finally may calculate the false discovery rate by plugging this $\alpha$
+into Equation~\ref{eq:fdr} and assuming empirically-motivated values for 
+base rate, $b$, and statistical power, $W$, as well as more pessimistic and
+optimistic settings for base rate and power.
 
-We invert this procedure to find what significance value $d^*$ is necessary
-to achieve a low family-wise error rate $\alpha_e = 0.05$ for experimental
-condition, $e$. To do this we first calculate $\alpha_e(d^*)$ for a
+We close our Analysis by inverting this procedure to find what significance value
+$d^*$ is necessary to achieve a low family-wise error rate $\alpha_e = 0.05$ for
+experimental condition, $e$. To do this we first calculate $\alpha_e(d^*)$ for a
 sufficiently wide range of $d^*$. We then find the lowest $d^*$ for which
@@
 up to the study level and across all studies and experimental conditions. 
-The lower twentieth percentile of $\alpha$
-We close by showing that $d^*$ must be greater than the ``high'' effect size
-$d^*=0.8$ to limit family-wise error rates to $\alpha \leq 0.05$ in all but six
-experimental conditions.
+We estimate the lower twentieth percentile of experimental conditions 
+have \emph{at least} $\alpha \leq 0.16$, 
+with median $\alpha \geq 0.27$ and upper eightieth percentile with $\alpha \geq
+0.40$ (Table~\ref{tab:quantiles}\mt{A.  –but don't have A and B; TODO: separate
+FWER and FDR tables, only use 0, 20, 50, 80, 100 for probs. Rename 0 -> min, 100 ->
+max.}). For the fiftieth and eightieth percentiles this
+corresponds to false discovery rates of 0.75 and 0.82, respectively. 
+We emphasize that these estimates and others detailed below are lower bounds on
+$\alpha$ and FDR since we are using what is likely an overgeneralized,
+underspecified statistical model that does not account for all the potential
+variance in experimental outcomes, which would increase error rates.
+We close our Analysis by calculating which $d^*$ limits family-wise error rates to
+$\alpha \leq 0.05$. This is greater than $d^* = 0.8$, classically taken to 
+represent ``high'' significance, in all but six experimental conditions.
 
-\subsection{Nearly all group polarization detections are plausibly false}
+\subsection{Nearly all group polarization detections are plausibly spurious}
 
@@
 conditions that did not claim to be positive findings of group polarization.
-Of the remaining 57, we identified plausible $\beta$ that generate spurious
+Of the remaining 57, we identified simple consensus $\beta$ that generate spurious
 group polarization in 54 experimental conditions. In the other three, we did
 identify $\beta$ that generate spurious group polarization, but the initial
-simulated distributions were highly polarized. Two of these
-simulations of experimental conditions from Schkade, et al., (2010) because
+simulated distributions were highly polarized. We rejected $\beta$ generated for
+two spurious group polarization from Schkade, et al., (2010) because
 the histogram plots of the observed opinion distributions are not
 polarized~\cite[Figure 1, p. 234]{Schkade2010}. The other condition where we
-rejected candidate $\beta$ was in a question about a ``bad teacher'' in Myers
-(1975)~\cite{Myers1975a}, where lower values signify increasingly 
-worse impressions of the teacher and higher values signify increasingly good
-impressions. It is unlikely that any participants would rate a ``bad teacher''
-as good, so we excluded this from our analysis as well. This exclusion may be
-overly generous since Myers (1975) did not provide details beyond mean values
-and opinion shifts, and the article was published before the importance
-of open data best practices was recognized.
+found but rejected spurious group polarization $\beta$ was in a question 
+about a ``bad teacher'' in Myers (1975)~\cite{Myers1975a}, where lower values
+signify increasingly worse impressions of the teacher and higher values signify
+increasingly good impressions. It is unlikely that a
+``bad teacher'' would be rated as good in this case, so we excluded this 
+from our analysis as well. This exclusion may be overly generous since Myers (1975)
+did not provide details beyond mean values and opinion shifts, and the article was
+published before open data practices.
 
 \begin{table}[ht]
-  \caption{yo}
+  \caption{{\textbf{Quantiles for the two error-rate measures under different
+  significance values.}\mt{Working on splitting in two and having more base rate
+and power conditions.}}}
   \label{tab:quantiles}
-\rotatebox{90} {\textbf{Quantiles for the two error-rate measures under different
-  significance values.}\mt{Not sure about the best way to format; probably don't
-  need this in the main text, but could be valuable in supplement?}}
+\rotatebox{90} {
   \input{tex/quantile_table.tex}
-
+}
 \end{table}
 
-\subsection{Half of experiments exceed a 75\% false discovery rate at ``medium'' significance}
+\subsection{Half of experiments exceed a 70\% false discovery rate at ``medium'' significance}
 
-We simulated trials from each experimental condition, $e$, 
-by sampling and binning pre- and
-post-deliberation latent opinions from distributions with parameters identified
-with the $N\to\infty$ model as causing plausibly spurious group
-polarization. For each $e$,
-we drew a number of samples equal to the number of participants in 
-the original experiments, $N_\mathrm{emp}$. To estimate
-family-wise error and false detection rates, we fit an ordered probit
-model to the simulated observations and obtained estimates of the group polarization
-effect size, $d$ (Equation~\ref{eq:cohens}). If the effect size was greater than a
-significance value, $d > d^*$, then by definition there was a detection $D$ of group 
-polarization,
-even though there was no effect by design, written $\neg E$. 
-The family-wise error rate is $\Pr(D|\neg E)$, calculated as the frequency with which $d > d^*$ out of 1000
-simulation trials. The false detection rate is the family-wise error rate scaled
-using the base rate of effects and the statistical power (Equation~\ref{eq:fdr}). We
-test three difference significance values, $d = 0.2,~0.5$ and $0.8$, corresponding
-to what Cohen prescribed to use for ``low'', ``medium'', and ``high'' significance.
+We simulated trials from each experimental condition, $e$, by sampling and binning
+pre- and post-deliberation latent opinions from distributions with parameters
+identified with the $N\to\infty$ model as causing plausibly spurious group
+polarization. For each $e$, we drew a number of samples equal to the number of
+participants in the original experiments, $N_\mathrm{emp}$. To estimate family-wise
+error and false detection rates, we fit an ordered probit model to the simulated
+observations and obtained estimates of the group polarization effect size, $d$
+(Equation~\ref{eq:cohens}). If the effect size was greater than a significance
+value, $d > d^*$, then by definition there was a detection $D$ of group
+polarization, even though there was no effect by design, written $\neg E$.  The
+family-wise error rate is $\Pr(D|\neg E)$, calculated as the frequency with which $d
+> d^*$ out of 1000 simulation trials. The false detection rate is the family-wise
+error rate scaled using the base rate of effects and the statistical power
+(Equation~\ref{eq:fdr}). We test three difference significance values, $d =
+0.2,~0.5$ and $0.8$, corresponding to what Cohen prescribed to use for ``low'',
+``medium'', and ``high'' significance. If not otherwise noted, we set the base
+rate to be $b=0.1$ and the power $W = 0.8$ as metascience parameters to calculate 
+the false discovery rate (Equation~\ref{eq:fdr}). Note that if the family-wise error
+rate is low at $\alpha = 0.05$, the false discovery rate is $\fdr = 0.36$.
 
-Family-wise error rates (Figure~\ref{fig:fwer_fdr_synthesis} and
-Table~\ref{tab:quantiles}) varied from a minimum of one in ten for the
+Different conditions yield different effect size distributions
+(Figure~\ref{fig:effect_size_distros}), which 
+determine the magnitude of error rates and difference between the error rates for low,
+medium, and high significance values $d^*=0.2,0.5,0.8$ magnitude of error rates 
+(Figure~\ref{fig:fwer_fdr_synthesis}B). 
+Family-wise error rates varied from a minimum of one in ten for the
 \texttt{Feminists-Experimental} condition from Myers (1975)~\cite{Myers1975a}
 to a maximum of two in three for the \texttt{COSprings-CivilUnions} condition from
-Schkade, Sunstein, and Hastie (2010)~\cite{Schkade2010}. The median family-wise error
+Schkade, Sunstein, and Hastie (2010)~\cite{Schkade2010} under medium
+significance effect size $d^* = 0.5$ (Figure~\ref{fig:fwer_fdr_synthesis} and
+Table~\ref{tab:quantiles}). The median family-wise error
 rate is 0.28 for Krizan and Baron's (2007) experimental condition
-\texttt{NoOutgroupScenario1}~\cite{Krizan2007}.
+\texttt{NoOutgroupScenario1}~\cite{Krizan2007}. This translated to a minimum
+false detection rate of 0.52 for the Myers (1975) condition, a maximum false
+detection rate of 0.88 for the Schkade, et al., (2010) condition, and a median false
+detection rate of 0.75 for the Krizan and Baron (2007) condition.
+
+In the conditions
+with lower family-wise error rates, there were several conditions where using the low 
+significance value $\alpha$ resulted in a much greater error rate than the medium or 
+high significance values. In the ten conditions with the lowest $\alpha$ for 
+$d^* = 0.5$, with $\alpha$ near 0.1, five of these conditions have $\alpha > 0.25$
+when $d^* = 0.2$ corresponding to a small effect size. For low $\alpha$, the
+difference in false detection rates for each $d^*$ value are more widely
+distributed, following the distribution of $\alpha$, meaning that 
+a modest increase in $d^*$ can have greater effects on the false discovery rate. In
+the conditions above the 50th percentile, $\alpha$ approaches and exceeds one in
+two, even when requiring ``high'' significance with $d^*=0.8$. With $d^* = 0.8$,
+$\alpha$ approaches and exceeds one in four, with four near one in two. When 
+requiring ``low'' signifance,
+
+Experimental designs within each study or article tend to share certain
+characteristics, such as the number of bins in the ordinal measurement design
+or the number of participants per condition. Therefore, the 
+average false detection rate across experimental trials within a study
+(Equation~\ref{eq:study_aggregate}) show that
 
@@
   \centering
-  \includegraphics[width=1.1\textwidth]{Figures/Results/fwer_fdr_synthesis.pdf}
+  \includegraphics[width=1.1\textwidth]{Figures/Analysis/fwer_fdr_synthesis.pdf}
+\end{figure}
 
-\end{figure}
+\begin{itemize}
+  \item 
+    Other $b$ and $W$ values in quantiles.
+  \item
+    Note that there were many ``significant'' results---in the wrong direction
+    (Figure~\ref{fig:OrdinalBoxplot}).
+\end{itemize}
 
@@
 which drives the false detection rate. Here we present our calculations to find
-$d^*$ that achieve $\alpha(d^*) = 0.05$, a commonly-assumed parameter value of
-metascience.
+$d^*$ that achieve $\alpha(d^*) = 0.05$.
 
 \begin{figure}
-  \caption{\textbf{Significance value, $d^*$, to limit family-wise error 
-  rate of 0.05 for each identified study.} \mt{Does this make sense? Order by the
-rankings in fdr\_vs\_experiment plots to better understand/compare.}}
+  \caption{
+    \textbf{Significance value, $d^*$, to limit family-wise error 
+    rate of 0.05 for each identified study.} The Moscovici and Zavalloni (1969)
+    ``Americans'' question would require the greatest $d^*$ to achieve $\alpha \leq
+    0.05$, even though it does not have the highest error rates. This is because of
+    larger outliers in simulated $d$ for this condition compared to others with
+    greater error rates (Figure~\ref{fig:OrdinalBoxplot}).
+  }
   \centering
-    \includegraphics[width=0.75\textwidth]{Figures/Results/sigval_for_low_fwer.pdf}
+    \includegraphics[width=0.75\textwidth]{Figures/Analysis/sigval_for_low_fwer.pdf}
   \label{fig:sigval_for_low_fwer}
@@
 
-% \section{Figures}
+\renewcommand{\thefigure}{A\arabic{figure}}
 
-% \begin{figure}
-%   \caption{
-%     Boxplots visualizing the distribution of Cohen's $d$ values (x-axis in
-%     A and B) inferred through a metric linear model over 1000 simulation 
-%     trials for each experimental condition (y-axis in A and B). All conditions
-%     have many trials that result in non-zero Cohen's $d$ values with number of
-%     simulated participants, $N$, set to their empirical values reported in the
-%     associated journal article (A). However, when the number of simulated
-%     participants is increased to $10N$, there are several conditions that 
-%     indeed yield distributions closer to $d = 0$ as expected, however some
-%     conditions have $d$ distributions centered on a nonzero mean with little to
-%     no overlap with $d=0$; variance in $d$ decreases as expected
-%     across experimental conditions (B).
-%   }  
-%   \label{fig:MetricBoxplots}
-%   \centering
-%   \begin{subfigure}[]{0.9\textwidth}
-%     % \hspace{-7.5em} 
-%     \centering
-%     \textsf{\large \textbf{A}, empirical $N$} \\
-%     \includegraphics[width=\textwidth]{Figures/boxplots/metric_cohens.pdf}
-%     \label{fig:MetricBoxplot_N}
-%   \end{subfigure} \\
-% \end{figure}
-% \begin{figure}\ContinuedFloat
-%   \begin{subfigure}[]{0.9\textwidth}
-%     % \hspace{-5.5em} 
-%     \centering
-%     \textsf{\large \textbf{B}, 10 $\times$ empirical $N$} \\
-%     \includegraphics[width=\textwidth]{Figures/boxplots/10N_metric_cohens.pdf}
-%     \label{fig:MetricBoxplot_10N}
-%   \end{subfigure}
-% \end{figure}
+\setcounter{figure}{0}
 
-% \begin{figure}
-%   \caption{Boxplot of Cohen's $d$ for the ordered probit model over 100
-%     simulation trias for each experimental condition (y-axis in A and B). Each
-%     trial represents a possible outcome of a group polarization experiment where
-%     the true opinion shift is zero. The mean across trials is closer to 0 than
-%     in the metric case, but many simulated zero-shift experiments still result in
-%     an observed shift. This illustrates the weak statistical power of these group
-%     polarization experiments.}
-%   \label{fig:OrdinalBoxplot}
-%   \centering
-%     \includegraphics[width=0.9\textwidth]{Figures/boxplots/ordinal_cohens.pdf}
-% \end{figure}
+\clearpage
+
+%%%%%%%%%% Supplement %%%%%%%%%%
+\pagebreak
+\begin{center}
+  \textbf{\Large \textsf{Minority-group incubators and majority-group
+reservoirs support the diffusion of climate change adaptations (Supplementary Material)}}
+\end{center}
+
+%%%%%%%%%% Prefix a "S" to all equations, figures, tables and reset the counter %%%%%%%%%%
+\setcounter{equation}{1}
+\setcounter{figure}{0}
+\setcounter{section}{0}
+\setcounter{table}{0}
+\setcounter{page}{1}
+\makeatletter
+\renewcommand{\theequation}{S\arabic{equation}}
+\renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\thetable}{S\arabic{table}}
+\renewcommand{\thesection}{S\arabic{section}}
+\renewcommand{\thepage}{S\arabic{page}}
+
+\section{Analysis}
+
+\begin{figure}
+  \caption{Boxplot of Cohen's $d$ for the ordered probit model over 1000
+    simulation trias for each experimental condition (y-axis in A and B). Each
+    trial represents a possible outcome of a group polarization experiment where
+    the true opinion shift is zero. The mean across trials is closer to 0 than
+    in the metric case, but many simulated zero-shift experiments still result in
+    an observed shift. This illustrates the weak statistical power of these group
+    polarization experiments.}
+  \label{fig:OrdinalBoxplot}
+  \centering
+    \includegraphics[width=1.0\textwidth]{Figures/Analysis/ordinal_cohens.pdf}
+\end{figure}
 

### 2025-01-01 GroupPolarizationStatmod/scripts/analysis.R

@@ -14,7 +14,6 @@
 
 library(purrr)
 
-
 source("src/fwer.R")
 source("src/util.R")
 

### 2025-01-01 GroupPolarizationStatmod/scripts/calculations.R

@@ -6,22 +6,28 @@
 #
 
 
-source("src/util.R")
+source("scripts/analysis.R")
 
 
 # Create a CSV list of quantile calculations that may be included in publications.
 quantiles_tbl = function(tex_output_file = "paper/tex/quantile_table.tex", 
-                         q_resolution = 0.05, 
+                         q_resolution_or_probs = 0.05, 
                          significance_vals = c(0.2, 0.5, 0.8),
                          base_rate = 0.1, power = 0.8
                          ) {
   
-  probs <- seq(0.0, 1.0, q_resolution)
+  # Hacky but just need this to proceed for now.
+  if (length(q_resolution_or_probs) > 1) {
+    probs <- q_resolution_or_probs
+  } else {
+    probs <- seq(0.0, 1.0, q_resolution_or_probs)
+  }
 
   initialized = FALSE
   for (sigval in significance_vals) {
 
-    tbl <- calc_fdr_vs_significance(significance_vals = sigval)
+    tbl <- calc_fdr_vs_significance(significance_vals = sigval, 
+                                    base_rate = base_rate, power = power)
 
     for (measure in c("FWER", "FDR")) {
 
@@ -64,14 +70,42 @@ quantiles_tbl = function(tex_output_file = "paper/tex/quantile_table.tex",
   # q_tbl_lim <- filter(SigVal == 0.5) %>% select(!c("SigVal", 
   q_tbl_lim <- q_tbl %>% select(!c("BaseRate", "Power"))
   
-  print(xtable(q_tbl_lim, type = "latex"), file = tex_output_file, 
-        include.rownames = FALSE, booktabs = TRUE, floating = FALSE)
+  if (!is_null(tex_output_file)) {
+    print(xtable(q_tbl_lim, type = "latex"), file = tex_output_file, 
+          include.rownames = FALSE, booktabs = TRUE, floating = FALSE)
+  }
 
   return (q_tbl)
 }
 
 
-all_calcuations <- function() {
-  fwer_fdr_synthesis_data()
-  quantiles_tbl()
+build_quantiles_across_contexts <- 
+  function(pessimistic_b_W = c(0.05, 0.5), 
+           middle_b_W = c(0.1, 0.8),
+           optimistic_b_W = c(0.5, 0.95),
+           quantile_probs = as.vector(c(0.0, 0.2, 0.5, 0.8, 1.0))) {
+
+  pbw <- pessimistic_b_W
+  mbw <- middle_b_W
+  obw <- optimistic_b_W
+
[…]

### 2025-01-01 GroupPolarizationStatmod/scripts/plot/analysis.R

@@ -19,7 +19,7 @@ source("scripts/analysis.R")
 
 mytheme = theme(
   axis.line = element_line(), legend.key=element_rect(fill = NA),
-  text = element_text(size=16),# family = 'PT Sans'),
+  text = element_text(size=18),# family = 'PT Sans'),
   # legend.key.width = unit(2, 'cm'),
   # legend.key.size = unit(1.5, 'lines'),
   panel.background = element_rect(fill = "white"),
@@ -34,7 +34,7 @@ rhg_cols_10 <- rev(c("#771C19", "#AA3929", "#E25033", "#F27314", "#F8A31B",
 
 plot_ordinal_cohens <- function(ordinal_data_dir = "data/probit_fits",
                                 sync_file = "data/probit_fits/all.csv",
-                                output_file = "paper/Figures/ordinal_cohens.pdf",
+                                output_file = "paper/Figures/Analysis/ordinal_cohens.pdf",
                                 overwrite = FALSE) {
   
   # df <- dir_ls(ordinal_data_dir, glob = "*.csv") %>%
@@ -56,7 +56,7 @@ plot_ordinal_cohens <- function(ordinal_data_dir = "data/probit_fits",
 
 
 
-plot_fdr = function(save_path = "paper/Figures/fdr_vs_experiment.pdf", 
+plot_fdr = function(save_path = "paper/Figures/Analysis/fdr_vs_experiment.pdf", 
                     aggregate = FALSE, use_non_identified = FALSE,
                     default_fwer = 0.05,
                     base_rate = 0.1, power = 0.8, 
@@ -80,12 +80,9 @@ plot_fdr = function(save_path = "paper/Figures/fdr_vs_experiment.pdf",
 plot_fdr_aggregated = function(tbl, focal_sig_val, 
                                save_path, default_fwer = 0.05,
                                use_non_identified = FALSE) {
-
   
   tbl = 
     as_tibble(aggregate_fdr_vs_sig(tbl, use_non_identified, default_fwer))
-      # mutate(StudyID = fct_reorder(StudyID, FWER
-    
 
   tbl_foc = 
     filter(tbl, SigVal == focal_sig_val) %>%
@@ -117,8 +114,10 @@ plot_fdr_aggregated = function(tbl, focal_sig_val,
     ggplot(tbl_foc, aes(x = FWER, y = StudyID, fill = StudyID)) +
[…]

### 2025-01-01 GroupPolarizationStatmod/src/fwer.R

@@ -26,23 +26,29 @@ cohens_d <- function(pre_mean_estimate, post_mean_estimate, pre_sd, post_sd) {
 # Family-wise error rate calculation using whatever ExperimentIDs are present.
 # Note probit_fits may be a tibble for avoiding reading data every calculation.
 #
-calculate_fwer = function (probit_fits = "data/probit_fits", sigval = 0.8) {
+calculate_fwer <- function (probit_fits = "data/probit_fits", sigval = 0.8) {
 
   # Load probit data if necessary.
   if (typeof(probit_fits) == "character") {
-    probit_data = load_probit_data(probit_fits)
+    probit_data <- load_probit_data(probit_fits)
   } else {
-    probit_data = probit_fits
+    probit_data <- probit_fits
   }
      
-  probit_data$Cohens_d = cohens_d(probit_data$LatentMeanPrePosteriorMean, 
+  probit_data$Cohens_d <- cohens_d(probit_data$LatentMeanPrePosteriorMean, 
                                   probit_data$LatentMeanPostPosteriorMean,
                                   probit_data$LatentMeanPrePosteriorSD, 
                                   probit_data$LatentMeanPostPosteriorSD)
 
-  probit_data$Significant = ifelse(probit_data$Cohens_d >= sigval, 1, 0)
+  # Get the sign of the hypothesized shift, where these columns come from 
+  # the original data.
+  shift_sign <- sign(probit_data$ObservedMeanPost - probit_data$ObservedMeanPre)
 
-  # return (probit_data)
+  # Only significant if it's positive in direction of shift. TODO: alternative
+  # formulation could check absolute value.
+  probit_data$Significant <- 
+    ifelse((shift_sign * probit_data$Cohens_d) >= sigval, 1, 0)
+
   ret = probit_data %>%
     group_by(ExperimentID) %>% 
     summarise(FWER = mean(Significant)) %>%

### 2025-01-01 GroupPolarizationStatmod/src/util.R

@@ -10,6 +10,7 @@ library(readr)
 library(tidyr)
 library(fs)
 
+
 load_probit_data = function(probit_fits_dir = "data/probit_fits", 
                             sync_file = "data/probit_fits/all.csv",
                             overwrite = FALSE) {
@@ -28,5 +29,4 @@ load_probit_data = function(probit_fits_dir = "data/probit_fits",
 
   return (ret)
 }
-
 

### 2025-01-25 GroupPolarizationStatmod/paper/ms.tex
nearly done with Analysis
fixed up table in final push for final draft

@@
 \hypersetup{linkcolor=gray}
-\setcounter{tocdepth}{3}
+\setcounter{tocdepth}{2}
 \tableofcontents
@@
 $d_{ei}$ is positive, meaning a shift to greater extremisim. It is possible that
-$d_{ei}$ could be negative, indicating a shift to lesser extremity.
+$d_{ei}$ could be negative, indicating a shift to lesser extremity. Indeed we
+observe many simulations with a shift to lesser extremity. For now we do not
+include these as indicating a ``significant'' effect, which we return to in the
+Analysis below.
 
@@
   \caption{{\textbf{Quantiles for the two error-rate measures under different
-  significance values.}\mt{Working on splitting in two and having more base rate
-and power conditions.}}}
+  significance values.}}}
   \label{tab:quantiles}
-\rotatebox{90} {
-  \input{tex/quantile_table.tex}
-}
+  \centering
+  \input{tex/quantile_table_min_20_50_80_max.tex}
 \end{table}
@@
 
-\subsection{Minimum $d^*$ for FWER of 0.05}
+\subsection{High significance values ($d^*$) often necessary for FWER of 0.05}
 
+Our simulations can be used a different way, in this case to calculate what
+significance value $d^*$ limits the family-wise error rate to the low value of
+0.05. 
 Solving the inverse problem of which $d^*$ achieves a low family-wise error rate,
@@
 \begin{center}
-  \textbf{\Large \textsf{Minority-group incubators and majority-group
-reservoirs support the diffusion of climate change adaptations (Supplementary Material)}}
+  \textbf{\Large \textsf{Group polarization replications may be marred by high false discovery rates (Supplementary Material)}}
 \end{center}

### 2025-01-25 GroupPolarizationStatmod/JAGSModels.R

@@ -1,106 +1,0 @@
-makeJAGSModelData <- function(N, firstBinValue, nBins, latentMean, latentSd)  # Pre, latentSdPost)
-{
-    thetaVec = as.vector(matrix(NA, ncol=nBins - 1, nrow=1))
-    thetaVec[1] = firstBinValue + 0.5
-    thetaVec[nBins - 1] = (firstBinValue + nBins - 1.5)  # e.g. fbv=1, nbins=10, this val is 9.5
-
-    opinions = simulatedObservation(N, firstBinValue, nBins, latentMean, latentSd)
-
-    modelDataList = list(
-        N = N,
-        o = opinions,
-        thetaVec = thetaVec,
-        nBins = nBins
-    )
-        
-    return (modelDataList)
-}
-
-
-# JAGS model definition 
-# ordSingleModelString = "
-
-#   model {
-
-#     for (i in 1:N) {
-
-#       o[i] ~ dcat(pr[i, 1:nBins])
-#       pr[i,1] <- pnorm(thetaVec[1], mu, 1/sigma^2)
-
-#       for (k in 2:(nBins - 1)) {
-#         pr[i, k] <- max(
-#           0, 
-#           pnorm(thetaVec[k], mu, 1 / (sigma^2)) - pnorm(thetaVec[k - 1], mu, 1 / (sigma^2)) 
-#         )
-#       }
-
-#       pr[i, nBins <- (1 - pnorm(thetaVec[nBins - 1], mu, 1/sigma^2))
-#     }
-
-#     mu ~ dnorm((1+max(nBins)) / 2, 1 / (max(nBins)^2))
-#     sigma ~ dunif(0.01 , max(nBins)*10)
-
-#     # Prior on thetaVec[q,k]. Stochastic for all except thetaVec[1,1] and thetaVec[1,last].
-#     for (kIdx in 2:(nBins - 2)) { # 1 and nBins-1 are fixed
-#       thetaVec[kIdx] ~ dnorm(kIdx + 0.5, 1/(2^2))
-#     }
-# }
-
-# " # close quote for ordModelString
-
-
-
-#   # ASSEMBLE THE DATA FOR JAGS. 
-#   #
-#   # N.B. THE JAGS MODEL ASSUMES THAT ALL ITEMS ARE SCORED WITH POSITIVE
[…]

### 2025-01-25 GroupPolarizationStatmod/experiments.R

@@ -6,7 +6,7 @@
 # require(glue)
 require(MASS)
 
-source("JAGSModels.R")
+
 
 ## INITIALIZE JAGS PARAMETERS (Following Kruschke (2015) DBDA2E-utilities.R)'
 # 
@@ -96,12 +96,12 @@ makeBayesianFitTable <- function(studies.data.csv = "data/StudiesAnalysis.csv",
     } else {
       lim <- 3
     }
-    # for (rowIdx in 1:nrow(studiesDf))
+    
     for (rowIdx in 1:lim)
     {
         row <- studiesDf[rowIdx, ]
-        #if (row$Include && row$Plausible && row$MinBinValue > 0)
-        if (row$Include && row$Plausible)  # && row$MinBinValue > 0)
+        
+        if (row$Include && row$Plausible)  
         {
             N <- row$N; firstBinValue <- row$MinBinValue; 
             nBins <- row$MaxBinValue - row$MinBinValue + 1; 
@@ -125,48 +125,81 @@ makeBayesianFitTable <- function(studies.data.csv = "data/StudiesAnalysis.csv",
                         "\n**************************************************\n\n",
                         sep=""))
             
-            # tryCatch({
+            # Get summary of ordered probit fit to simulated pre-deliberation data.
             suPre <- summary(
                 calculateBayesian(N, firstBinValue, nBins, latentMean, latentSdPre)
             )
 
+            # Get summary of ordered probit fit to simulated post-deliberation data.
             suPost <- summary(
                 calculateBayesian(N, firstBinValue, nBins, latentMean, latentSdPost)
             )
 
+            # Extract the posterior estimates for pre-...
             muPrePost = suPre["mu",]
+            # ...and post-deliberation latent opinion means.
             muPostPost = suPost["mu",]
 
+            # Add row for this experimental condition extracting several variables
+            # from the posterior estimates for pre- and post-deliberation latent
+            # opinion means and variances.
             tibbleRow <- tibble_row(
+                
+                # Metadata and parameters read from the `row` from `studiesDf`.
                 TreatmentTag = row$TreatmentTag, ArticleTag = row$ArticleTag,
                 N = N, 
                 MinBinValue = firstBinValue, 
                 MaxBinValue = firstBinValue + nBins - 1, 
[…]

### 2025-06-19 GroupPolarizationStatmod/paper/intro.qmd
init intro.qmd for only intro drafting

---
title: "Group Polarization Revisited"
author: "Your Name"
format:
  pdf:
    documentclass: article
    keep-tex: true
bibliography: 
  - this.bib
  - /Users/mt/workspace/Writing/library.bib
output: pdf
---

# Introduction

Group polarization is one of the most widely cited but unevenly evidenced phenomena in social psychology...

*Write like you mean it.*  
*You're not drafting — you're building.*

### 2025-06-19 GroupPolarizationStatmod/paper/ms.tex
finalized table loc for now
init intro.qmd for only intro drafting

@@
 
-If an extremist opinion falls and it is measured with a Likert scale, will it make a noise?
+If an extremist's opinion falls and it is measured with a Likert scale, will it
+make a noise? The answer had better be yes if we want to properly measure opinion
+change. Accurate measurements are critical
+for developing rigorous and reliable methods to promote more centrist,
+sustainable, and responsive governments and institutions
+\autocite{Mason2018UncivilAgreementBook,Klein2020}. Unfortunately, social and
+behavioral scientists have often used statistical methods 
+that distort
+opinion change when it's measured on an ordinal, Likert-style scale (e.g. -3
+indicates ``Strongly disagree'', +3 indicates ``Strongly agree'', and 0 is
+neutral)~\autocite{Liddell2018}. 
+This includes research that Obama official and law professor Cass Sunstein
+marshals to explain why ``people become extremists'' and why ``political and
+cultural polarization'' is ``so pervasive in America''\footnote{From the
+  Publisher's Summary. Sunstein does not make such general claims in the book,
+  but does suggest that the research he reviews collectively provides
+  a ``clue'' to explaining how
+  ``facism...student radicalism...Islamic terrorism...[t]he Rwandan
+genocide...[e]thnic Conflict.}
+administration official and law professor Cass Sunstein writing a book called
+*researchers construct
 
-Unchecked social and political polarization undermines cooperation and
-political efficacy \autocite{Mason2018UncivilAgreementBook,Klein2020}.
-\emph{Group polarization} is the name given the social phenomenon of 
+\emph{Group polarization} is that essentially
+measures how much being in an opinion echo-chamber the name given the social phenomenon of 
 when a group becomes more extreme in mean opinion over
@@
 
-Experimetnal designs in group polarization studies share a certain
+The null-consistency procedure functions as a test of model indistinguishability: it demonstrates that a reported group polarization effect could plausibly arise from a stable-latent process combined with changing opinion precision.
+This logic echoes foundational concerns in causal inference, where two competing models—one causal, one not—can imply the same observed data distribution and are therefore empirically indistinguishable without further assumptions \cite{pearl2000causality,spirtes2000causation}.
+In such cases, inference from observation to mechanism is invalid.
+Similarly, in psychometrics, it is well known that different configurations of latent traits and response thresholds can yield identical observed ordinal patterns, making the true structure unidentifiable without additional constraints \cite{borsboom2005measuring}.
+The present approach does not attempt to estimate the probability of false detection, but rather shows that under ordinal measurement, a null model can reproduce the observed pattern—rendering the result non-diagnostic of a true latent shift.
+This establishes a stringent standard: an effect must not only be detected, but shown to be incompatible with plausible null-generating mechanisms.
+
+Experimental designs in group polarization studies share a certain
 experimental boilerplate (Figure 1 --- NEED TO INCLUDE) with the goal of
@@
 
-\begin{table}[ht]
-  \caption{{\textbf{Quantiles for the two error-rate measures under different
-  significance values.}}}
-  \label{tab:quantiles}
-  \centering
-  \input{tex/quantile_table_min_20_50_80_max.tex}
-\end{table}
-
 \subsection{Half of experiments exceed a 70\% false discovery rate at ``medium'' significance}
@@
 detection rate of 0.75 for the Krizan and Baron (2007) condition.
+
+\begin{table}[ht]
+  \caption{{\textbf{Quantiles for the two error-rate measures under different
+  significance values.}}}
+  \label{tab:quantiles}
+  \centering
+  \input{tex/quantile_table_min_20_50_80_max.tex}
+\end{table}
 

### 2025-06-19 GroupPolarizationStatmod/paper/intro.qmd
updated title

@@
 ---
-title: "Group Polarization Revisited"
-author: "Your Name"
+title: "If the Null Fits, You Must Acquit: Ubiquitous False Detections of Group Polarization"
+author: "Matt Turner"
 format:

### 2025-06-20 GroupPolarizationStatmod/paper/intro.qmd
intro.qmd started

--- 2025-06-19 GroupPolarizationStatmod/paper/intro.qmd
@@
 ---
-title: "Group Polarization Revisited"
-author: "Your Name"
+title: "If the Null Fits, You Must Omit: Ubiquitous False Detections of Group Polarization"
+author: "Matt Turner"
 format:
+  html:
+    toc: true
+    include-in-header: header.html
+    theme: 
+      dark:
+        - darkly 
+        - theme-dark.scss
+      light:
+        - theme-light.scss
+    fontsize: 0.9em
+    linestretch: 1.2
   pdf:
@@
 
+## Abstract {.unnumbered}
+
+*Group polarization* is the name for a form of consensus where
+members of a like-minded group become, on average, more extreme in their
+opinions after discussing a topic. Group polarization is important because
+it may increase social and political tensions if moderates become more
+extreme. Decades of studies have replicated detections of group
+polarization, typically using an ordinal scale to measure opinions.
+However, these studies did not account for ordinal measurements, which can
+result in spurious detections of opinion change.  Lacking original data we
+can still calculate the probability that a published detection of group
+polarization was spurious, i.e., to calculate the \emph{false detection
+rate}.  Across 54 group polarization experimental conditions from ten
+representative journal articles, we calculated false detection rates to be
+between 0.52 and 0.88, with a median of 0.75, using a generative model of
+group polarization experiments seeded with empirical data.  
+We also use our model to develop experimental designs that
+achieve an acceptable false discovery rate.  Much of group polarization
+research may be unreliable. This analysis can help change that by enabling others to avoid this
+hidden pitfall. More
+broadly this work demonstrates one important way that replication success alone
+does not imply epistemic reliability.
+
 # Introduction
 
-Group polarization is one of the most widely cited but unevenly evidenced phenomena in social psychology...
+> In our introductory social psychology course, 
+we have for many years used the [group polarization experimental paradigm] as
+a laboratory exercise. The exercise works beautifully, but one must be
+careful to forewarn a class that [group polarization] does not occur with every 
+group...and that the effect is not large [@Brown1986].
 
-*Write like you mean it.*  
-*You're not drafting — you're building.*
+> One of the most robust findings in social psychology is that of attitude polarization  following discussion with like-minded others [@Cooper2001].
+
+> A medium effect size is conceived as one large enough to be visible to the naked eye. [@CohenBook1988]
+
+If an extremist's opinion falls and it is measured with a Likert scale, will it
+make a noise? The answer had better be yes if we want to properly measure opinion
+change. Accurate measurements are critical
+for developing rigorous and reliable methods to promote more stable,
+sustainable, responsive, and efficient governments and institutions
+[@Mason2018UncivilAgreementBook;@Klein2020]. Unfortunately, social and
+behavioral scientists have often used statistical methods 
+that distort
+opinion change when it's measured on an ordinal, Likert-style scale (e.g. -3
+indicates "Strongly disagree", +3 indicates "Strongly agree", and 0 is
+neutral) [@Liddell2018]. 
+This includes a phenomenon called *group polarization* that
+former Obama White House official and law professor Cass Sunstein relied on to 
+explain why "people become extremists" and why "political and
+cultural polarization" is "so pervasive in America", as the publisher summarizes [@Sunstein2009]. Sunstein suggests group polarization research provides a "clue" to explaining how
+"facism...student radicalism...Islamic terrorism...[t]he Rwandan
+genocide...[e]thnic Conflict...acts of torture and humiliation by American soldiers at Abu Ghraib", and more (p. 1).
+
+This work is motivated by a core conviction that better thinking leads to better
+science. Theoretical clarity, careful measurement, and principled statistical
+modeling are not academic luxuries, but essential tools for empirical science.
+Theoretical and statistical confusions in the group polarization literature have
+lain dormant, undermining measurements of opinion change in group polarization
+unbeknownst to generations of researchers [@Brown1986].
+
+**🧭 Next session, 6/20: Map paragraph set bridging theory to methods — experiments → ordinal pitfalls → null-consistency model → power analysis → transition to results**
+
+## Bridge: From Philosophy to Mechanics {.unnumbered}
+
+<!-- 🧭 Writing plan for next session -->
+
+### Experimental structure of group polarization studies
+
+- Brief overview of how group polarization experiments are typically designed  
+- Emphasize reliance on **pre/post ordinal opinion measures**  
+- Highlight the assumption that ordinal differences reflect latent opinion shifts
+
+### The ordinal-continuous mismatch problem
+
+- Explain why ordinal measurements do **not** uniquely reflect latent shifts  
+- Use an example or intuition (e.g., compressing/shrinking variance without changing means)  
+- Show how this opens the door to *null-consistent but polarization-looking* results
+
+### Description of the null-consistency modeling procedure
+
+- Introduce the generative model with fixed latent mean and changing variance  
+- Explain what counts as a plausible fit and how the model reverse-engineers the observation  
+- Clarify that this is a *constructive falsification tool*, not a probabilistic inference
+
+### Implications for statistical power and detection credibility
+
+- Discuss how, even if the detection is real, underpowered studies can't distinguish it from null  
+- Mention use of Bayesian ordered probit models to test for minimum detectable effect sizes  
+- Preview that even with ideal modeling, many studies likely lack sufficient power
+
+### Transition to empirical review and FDR estimation
+
+- Signal that the next section tests this logic across all experimental conditions  
+- Frame the upcoming analysis as assessing the **robustness** of published detections under this new lens
+
+The null-consistency procedure functions as a test of model indistinguishability: it demonstrates that a reported group polarization effect could plausibly arise from a stable-latent process combined with changing opinion precision.
+This logic echoes foundational concerns in causal inference, where two competing models—one causal, one not—can imply the same observed data distribution and are therefore empirically indistinguishable without further assumptions [@pearl2000causality;@spirtes2000causation].
+In such cases, inference from observation to mechanism is invalid.
+Similarly, in psychometrics, it is well known that different configurations of latent traits and response thresholds can yield identical observed ordinal patterns, making the true structure unidentifiable without additional constraints [@borsboom2005measuring].
+The present approach does not attempt to estimate the probability of false detection, but rather shows that under ordinal measurement, a null model can reproduce the observed pattern—rendering the result non-diagnostic of a true latent shift.
+This establishes a stringent standard: an effect must not only be detected, but shown to be incompatible with plausible null-generating mechanisms.

### 2025-06-20 GroupPolarizationStatmod/paper/ms.tex
intro.qmd started

@@
 *researchers construct
+
+This work is motivated by a core conviction that better thinking leads to better
+science.
+Theoretical clarity, careful measurement, and principled statistical modeling are not academic luxuries, but essential tools for empirical science.
+The group polarization literature has long suffered from ambiguities in theory, ordinal-to-continuous mismatches in measurement, and default inferential techniques that mask model non-identifiability.
+By reconstructing these issues from the ground up, this paper aims not only to assess the credibility of existing evidence, but to show that rigorous thinking — about what we measure, how we model, and what we can legitimately conclude — has direct, practical consequences for the claims we make about the social world.
 

### 2025-06-23 GroupPolarizationStatmod/paper/intro.qmd
worked on Intro

@@
         - theme-light.scss
-    fontsize: 0.9em
-    linestretch: 1.2
+    fontsize: 0.8em
+    linestretch: 1.1
   pdf:
@@
 can still calculate the probability that a published detection of group
-polarization was spurious, i.e., to calculate the \emph{false detection
-rate}.  Across 54 group polarization experimental conditions from ten
+polarization was spurious, i.e., to calculate the *false detection
+rate*.  Across 54 group polarization experimental conditions from ten
 representative journal articles, we calculated false detection rates to be
@@
 group polarization experiments seeded with empirical data.
-We also use our model to develop experimental designs that
-achieve an acceptable false discovery rate.  Much of group polarization
-research may be unreliable. This analysis can help change that by enabling others to avoid this
-hidden pitfall. More
-broadly this work demonstrates one important way that replication success alone
+We also use our model to develop experimental designs that achieve an acceptable false discovery rate.
+Much of group polarization research may be unreliable.
+This analysis can help change that by enabling others to avoid this hidden pitfall.
+More broadly this work demonstrates one important way that replication success alone
 does not imply epistemic reliability.
@@
 
-> A medium effect size is conceived as one large enough to be visible to the naked eye. [@CohenBook1988]
+<!-- > A medium effect size is conceived as one large enough to be visible to the naked eye. [@CohenBook1988] -->
 
@@
 make a noise? The answer had better be yes if we want to properly measure opinion
-change. Accurate measurements are critical
-for developing rigorous and reliable methods to promote more stable,
-sustainable, responsive, and efficient governments and institutions
+change. Accurate measurements are critical for developing rigorous and reliable methods to promote more stable, sustainable, responsive, and efficient governments and institutions
 [@Mason2018UncivilAgreementBook;@Klein2020]. Unfortunately, social and
@@
 
+- We demonstrate here that there is a pervasive problem observed over decades of group polarization research where faulty statistical methods and incomplete reporting that should force the retraction of these studies from the canon.
+Effectively, then, empirical support for the theoretical concept *group polarization* stops being so: the epistemic integrity of *group polarization* is undermined with its empirical foundation losing mass.
+
+We complement this critical analysis with a prescriptive analysis to demonstrate that more appropriate statistical models do indeed lower the false discovery rate by committing fewer false positive detections of group polarization under a null model with shrinking opinion variance.
+Recently, statistical best-practice is to use Bayesian models of statistics, where social dynamics and opinion-reporting behaviors are represented as causally-dependent probability distributions, where a sample taken from one distribution (e.g., the opinion distribution of a group in a group polarization experiment) acts as an input parameter to another distribution (e.g., that represents the individually-heterogeneous distortion of opinions when reported on an ordinal scale).
+Our models follow current best practices in psychometrics, social science, and Bayesian inference [@Liddell2018; @Kruschke2015; @Rethinking; @Deffner2024], but consolidate them in a new, useful way that enables the
+retroactive discovery of null model parameters that generate spurious group polarization, forcing published evidence for group polarization to be classified as non-identifiable, and removed from evidence that can be said to support group polarization.
+
+Our methods can be turned on any data-driven social science analysis of data where participants report a latent, subjective psychological variable (mood, physiological state, etc.), and where various theories supposedly battle it out to best represent the phenomenon of interest via NHST performed on ordinal observations with weak, informal theoretical models to motivate the studies.
+This may just be the first of many bricks to fall since *all 68* of the articles that mentioned the word "Likert" and analyzed ordinal data used metric models to detect latent psychological variable change across three influential psychology journals, *Psychological Science*, the *Journal of Experimental Psychology: General*, and the *Journal of Personality and Social Psychology* [@Liddell2018].
+
 This work is motivated by a core conviction that better thinking leads to better
@@
 
-**🧭 Next session, 6/20: Map paragraph set bridging theory to methods — experiments → ordinal pitfalls → null-consistency model → power analysis → transition to results**
+## Identifiability and the Illusion of Change
 
-## Bridge: From Philosophy to Mechanics {.unnumbered}
+<!-- XXX BULLET POINTS ALL GENERATED BY CHATGPT BASED ON CONVO TO HELP 
+     MAKE A SNAPSHOT OF ALL THE KEY POINTS THAT I HAVE DEVELOPED INDEPENDENTLY
+     BUT AM STRUGGLING TO HOLD TOGETHER IN MY HEAD AT ONCE. I WILL BE EDITING, 
+     CUSTOMIZING, ETC., AFTER WHICH THIS MESSAGE WILL BE REMOVED. 6/23/25 -->
 
-<!-- 🧭 Writing plan for next session -->
+- Many group polarization studies report opinion change based on pre/post differences, but assume a transparent mapping between observed scores and latent beliefs.
+- Yet ordinal opinion data can mask deeper structural ambiguities that invalidate these assumptions.
+- In what follows, we first describe the basic structure of group polarization experiments, then explain how ordinal measurement can lead to illusory shifts.
+- Finally, we introduce a null-consistency modeling procedure that operationalizes this insight, and show how it informs the credibility of reported effects.
 
@@
 
-- Brief overview of how group polarization experiments are typically designed  
-- Emphasize reliance on **pre/post ordinal opinion measures**  
-- Highlight the assumption that ordinal differences reflect latent opinion shifts
+- Most group polarization experiments follow a familiar structure: measure individual opinions, allow group discussion, then re-measure.
+- The central claim is that like-minded individuals become more extreme after deliberation.
+- Crucially, opinions are typically measured with Likert-style ordinal scales (e.g., from “Strongly disagree” to “Strongly agree”).
+- Analyses often treat these as interval-scaled, calculating mean differences and testing for significance.
 
@@
 
-- Explain why ordinal measurements do **not** uniquely reflect latent shifts  
-- Use an example or intuition (e.g., compressing/shrinking variance without changing means)  
-- Show how this opens the door to *null-consistent but polarization-looking* results
+- Ordinal scores do not have a fixed interval structure — they represent ranked categories, not true distances.
+- A given change in score may reflect a shift in opinion, increased certainty, or both.
+- For example, if opinions become more tightly clustered (lower variance) post-deliberation, the group mean can appear to shift even when the latent mean remains unchanged.
+- This opens the door to *null-consistent detections* — apparent effects that emerge from measurement artifacts rather than real opinion movement.
 
@@
 
-- Introduce the generative model with fixed latent mean and changing variance  
-- Explain what counts as a plausible fit and how the model reverse-engineers the observation  
-- Clarify that this is a *constructive falsification tool*, not a probabilistic inference
+- We formalize this concern using a generative model that holds the latent opinion mean constant while allowing variance to shrink.
+- The goal is to test whether an observed group polarization result could arise from such a stable-mean, shifting-variance process.
+- If this model can reproduce the pre/post measurement pattern, the observed result is non-diagnostic — it is consistent with a null hypothesis of no true change.
+- This is a form of constructive falsification: it doesn’t prove the effect is false, but shows it is not uniquely explained by change.
 
@@
 
-- Discuss how, even if the detection is real, underpowered studies can't distinguish it from null  
-- Mention use of Bayesian ordered probit models to test for minimum detectable effect sizes  
-- Preview that even with ideal modeling, many studies likely lack sufficient power
+- Even if an effect is real, detecting it reliably requires statistical power — the ability to distinguish true shifts from noise or artifacts.
+- Studies using ordinal measures and small samples may be especially vulnerable to false positives under the null.
+- To assess this, we fit Bayesian ordered probit models to simulated data and estimate the power to detect effects of various sizes.
 
@@
 
-- Signal that the next section tests this logic across all experimental conditions  
-- Frame the upcoming analysis as assessing the **robustness** of published detections under this new lens
+- Having established a model-based definition of null-consistent detection, we now apply it to published group polarization experiments.
+- By combining this logic with simulations seeded from published data, we estimate false discovery rates across conditions.
+- This isn’t a judgment about whether past detections were true.
+It’s a demonstration of how to estimate false detection rates properly—using models that actually fit the structure of the data.
+We calculate FWER and FDR for the original study designs at three effect size levels—small, medium, and large—based on Cohen’s guidelines [@CohenBook1988, Ch.
+2.2.3] and Bayesian methods from Liddell and Kruschke [@Liddell2018].
+The verdict is clear: even if these studies had used better models, most still wouldn’t have had the power to rule out false positives.
+That makes replication a weak signal of reliability.
+If future group polarization studies want to produce meaningful results, they need stronger designs—and that starts with a clean break from outdated statistical habits.
 

### 2025-06-23 GroupPolarizationStatmod/paper/ms.tex
worked on Intro

@@
 This work is motivated by a core conviction that better thinking leads to better
-science.
-Theoretical clarity, careful measurement, and principled statistical modeling are not academic luxuries, but essential tools for empirical science.
-The group polarization literature has long suffered from ambiguities in theory, ordinal-to-continuous mismatches in measurement, and default inferential techniques that mask model non-identifiability.
+science. Theoretical clarity, careful measurement, and principled statistical
+modeling are not academic luxuries, but essential tools for empirical science.
+Theoretical and statistical confusions in the group polarization literature have
+lain dormant, undermining measurements of opinion change in group polarization
+unbeknownst to generations of researchers~\autocite{Brown1986}.
+
+literature has long suffered from ambiguities in theory, ordinal-to-continuous mismatches in measurement, and default inferential techniques that mask model non-identifiability.
 By reconstructing these issues from the ground up, this paper aims not only to assess the credibility of existing evidence, but to show that rigorous thinking — about what we measure, how we model, and what we can legitimately conclude — has direct, practical consequences for the claims we make about the social world.

### 2025-08-06 GroupPolarizationStatmod/paper/intro.qmd
an update to Intro.qmd from June; merge with the rest by compiling intro.tex and including with the rest of the latex draft

@@
 That makes replication a weak signal of reliability.
-If future group polarization studies want to produce meaningful results, they need stronger designs—and that starts with a clean break from outdated statistical habits.
+If future group polarization studies want to produce meaningful result need stronger designs—and that starts with a clean break from outdated statistical habits.
 

### 2025-08-08 GroupPolarizationStatmod/paper/ms.tex
updates; continuing draft

--- 2025-06-19 GroupPolarizationStatmod/paper/ms.tex
@@
 \makeatletter
-\title{Group polarization replications may be marred by high false discovery rates}
+\title{If the Null Fits, You Must Omit: Ubiquitous False Detections of Group Polarization}
 \makeatother
 \author[1,*]{{Matthew A.~Turner}}
-\affil[1]{\small Division of Social Sciences, Stanford Doerr School of Sustainability, Stanford University}
+\affil[1]{\small Environmental Social Sciences, Stanford Doerr School of Sustainability, Stanford University}
 
@@
 \maketitle
-\begin{abstract}
-\noindent 
-\emph{Group polarization} is the name for a form of consensus where
-members of a like-minded group become, on average, more extreme in their
-opinions after discussing a topic. Group polarization is important because
-it may increase social and political tensions if moderates become more
-extreme. Decades of studies have replicated detections of group
-polarization, typically using an ordinal scale to measure opinions.
-However, these studies did not account for ordinal measurements, which can
-result in spurious detections of opinion change.  Lacking original data we
-can still calculate the probability that a published detection of group
-polarization was spurious, i.e., to calculate the \emph{false detection
-rate}.  Across 54 group polarization experimental conditions from ten
-representative journal articles, we calculated false detection rates to be
-between 0.52 and 0.88, with a median of 0.75, using a generative model of
-group polarization experiments seeded with empirical data.  
-We also use our model to develop experimental designs that
-achieve an acceptable false discovery rate.  Much of group polarization
-research may be unreliable. This analysis can help change that by enabling others to avoid this
-hidden pitfall. More
-broadly this work demonstrates one important way that replication success alone
-does not imply epistemic reliability.
-\end{abstract}
+% \begin{abstract}
+% \noindent 
+% \emph{Group polarization} is the name for a form of consensus where
+% members of a like-minded group become, on average, more extreme in their
+% opinions after discussing a topic. Group polarization is important because
+% it may increase social and political tensions if moderates become more
+% extreme. Decades of studies have replicated detections of group
+% polarization, typically using an ordinal scale to measure opinions.
+% However, these studies did not account for ordinal measurements, which can
+% result in spurious detections of opinion change.  Lacking original data we
+% can still calculate the probability that a published detection of group
+% polarization was spurious, i.e., to calculate the \emph{false detection
+% rate}.  Across 54 group polarization experimental conditions from ten
+% representative journal articles, we calculated false detection rates to be
+% between 0.52 and 0.88, with a median of 0.75, using a generative model of
+% group polarization experiments seeded with empirical data.  
+% We also use our model to develop experimental designs that
+% achieve an acceptable false discovery rate.  Much of group polarization
+% research may be unreliable. This analysis can help change that by enabling others to avoid this
+% hidden pitfall. More
+% broadly this work demonstrates one important way that replication success alone
+% does not imply epistemic reliability.
+% \end{abstract}
 
@@
 \begin{center} \noindent\rule{4cm}{0.4pt} \end{center}
-\clearpage
+% \clearpage
 
@@
 
-\section{Introduction}\label{introduction}
+\input{intro-latest.tex}
 
-If an extremist's opinion falls and it is measured with a Likert scale, will it
-make a noise? The answer had better be yes if we want to properly measure opinion
-change. Accurate measurements are critical
-for developing rigorous and reliable methods to promote more centrist,
-sustainable, and responsive governments and institutions
-\autocite{Mason2018UncivilAgreementBook,Klein2020}. Unfortunately, social and
-behavioral scientists have often used statistical methods 
-that distort
-opinion change when it's measured on an ordinal, Likert-style scale (e.g. -3
-indicates ``Strongly disagree'', +3 indicates ``Strongly agree'', and 0 is
-neutral)~\autocite{Liddell2018}. 
-This includes research that Obama official and law professor Cass Sunstein
-marshals to explain why ``people become extremists'' and why ``political and
-cultural polarization'' is ``so pervasive in America''\footnote{From the
-  Publisher's Summary. Sunstein does not make such general claims in the book,
-  but does suggest that the research he reviews collectively provides
-  a ``clue'' to explaining how
-  ``facism...student radicalism...Islamic terrorism...[t]he Rwandan
-genocide...[e]thnic Conflict.}
-administration official and law professor Cass Sunstein writing a book called
-*researchers construct
-
-\emph{Group polarization} is that essentially
-measures how much being in an opinion echo-chamber the name given the social phenomenon of 
-when a group becomes more extreme in mean opinion over
-time, assuming the mean opinion is biased to one extreme to begin with.
-\mt{now hit them with the real shit to intro the problem, our analysis, and
-how this can be used constructively going forward.}
-
-\mt{Seems like a good second paragraph structurally. We should give more
-details about the scope of the study, group polarization and common
-elements of experimental designs.} 
-Group polarization experiments tend to employ a common set of design
-elements incorporated in our computational model analysis. Group
-polarization experiments mostly use ordinal (see exceptions
-\cite{Blascovich1974,Blascovich1975,Blascovich1976,Keating2016}). The opinions (e.g., ``Do you approve,
-disapprove, or neither approve nor disapprove of the current
-administration's policies?''). Extremity only increases, though, assuming
-that the mean opinion of the group is initially biased towards one extreme
-or the other \autocite{Brown1986,Brown2020}. Group polarization is important
-to understand rigorously, therefore, since if groups become increasingly
-extreme over time, while group opinions are consolidated, this would
-exacerbate political polarization between political opposition groups, which
-is already undermines the ability of partisans in democratic governments to
-work together. Dozens of studies and meta-analyses over decades have
-apparently established the empirical reality of group polarization
-\autocite{Moscovici1969,Myers1970,Isenberg1986,Sunstein2009,Sunstein2019}
-\autocite[Ch. 5]{Brown2000}. We believe group polarization should occur in
-certain contexts since more extreme group members are likely also to be more
-stubborn \autocite{Guazzini2015,Lewandowsky2019}, which should pull
-consensus to extremes \autocite{Turner2020}. However, many or most of the
-evidence for group polarization relies on experimental designs that use
-ordinal measurements that are then analyzed as if they were continuous.
-Unfortunately, this procedure is unfortunately known to induce false
-positives if it is not properly accounted for \autocite{Liddell2018}.
-
-The null-consistency procedure functions as a test of model indistinguishability: it demonstrates that a reported group polarization effect could plausibly arise from a stable-latent process combined with changing opinion precision.
-This logic echoes foundational concerns in causal inference, where two competing models—one causal, one not—can imply the same observed data distribution and are therefore empirically indistinguishable without further assumptions \cite{pearl2000causality,spirtes2000causation}.
-In such cases, inference from observation to mechanism is invalid.
-Similarly, in psychometrics, it is well known that different configurations of latent traits and response thresholds can yield identical observed ordinal patterns, making the true structure unidentifiable without additional constraints \cite{borsboom2005measuring}.
-The present approach does not attempt to estimate the probability of false detection, but rather shows that under ordinal measurement, a null model can reproduce the observed pattern—rendering the result non-diagnostic of a true latent shift.
-This establishes a stringent standard: an effect must not only be detected, but shown to be incompatible with plausible null-generating mechanisms.
-
-Experimental designs in group polarization studies share a certain
-experimental boilerplate (Figure 1 --- NEED TO INCLUDE) with the goal of
-testing the group polarization hypothesis that a group of like-minded
-people will develop more extreme opinions on some salient topic over the
-course of discussing some topic (Figure~\ref{fig:ModelSketch}). 
-Group polarization experiments are thus
-initialized by forming like-minded groups of people and surveying their
-initial opinions on some topic. In some cases these initial opinions are
-used for assigning participants to like-minded groups, but other methods
-have been used, such as geographic location for producing, on average,
-like-minded groups \autocite{Schkade2010}. Participants then discuss the
-topic in their groups. After discussion, participant opinions are again
-measured. Several studies we analyze here determined group polarization
-occurred (i.e., defined a positive detection) only as a non-zero
-difference in mean opinion, with no other statistical testing. Some
-others performed null-hypothesis \(t\)-tests to test whether there was
-enough evidence to reject the null-hypothesis that \emph{simple
-consensus} occurredand not group polarization. In our meta-analysis
-simulations, we measure simulated group polarization effect size in
-terms of Cohen's \(d\), which is an abstraction beyond these two
-practices from the group polarization literature that can also be used
-to compare model fits between metric models that treat group
-polarization measurements as continuous although they are ordinal, and
-more appropriate statistical models that account for data ordinality.
-
-% \begin{figure}
-%   \caption{\textbf{Opinions are implicitly behavioral choices that are most consistent with some
-%   complex aggregation of brain states and dynamics, and so are theoretically
-%   continuous (A). Common components of a group polarization experiment.}
-%   Participants first give their opinion in response to some prompt (B). Afterwards,
-%   participants deliberate the topic (C). Participants then give their opinion again
-%   (D). }
-%   \centering
-%     \includegraphics[width=0.65\textwidth]{Figures/Model/ModelSketchesA.pdf}
-%     \includegraphics[width=0.7\textwidth]{Figures/Model/ModelSketchesB-D.pdf}
-%   \label{fig:ModelSketch}
-% \end{figure}
-
-In this paper, we perform a meta-analysis of sixty detections of group
-polarization from ten published research articles and show that
-fifty-five of the sixty cannot be reliably identified as group
-polarization. We show that these 55 are plausibly false detections
-because they are equally well-explained by a \emph{simple consensus}
-model, where the mean group opinion is unchanged by group discussion but
-opinion variance decreases. Theoretically, consensus occurs when
-sufficiently like-minded individuals influence, which decreases group
-opinion variance \autocite{DeGroot1974,Turner2018}. Group polarization,
-then, is consensus plus an overall increase in group opinion extremity,
-i.e., an increase in mean group opinion after discussion. We use a novel
-approach, described herein, that enabled us to identify continuous pre-
-and post-discussion simple consensus opinion distributions that result
-in data that appear to be group-polarized when opinions are drawn from
-these distributions and measured with an ordinal scale. This approach
-could be adapted for similar social psychological studies that used
-ordinal measures to determine whether change occurred in distributions
-of continuous psychological variables such as opinions and beliefs. For
-social psychology and group polarization, this means that an accepted,
-heretofore unquestioned, empirical fact must be re-evaluated.
-Experimental designs in the field must be updated if we are to
-rigorously understand whether and when group polarization occurs.
-
-\subsection{Studies included in this
-meta-analysis}\label{studies-included-in-this-meta-analysis}
-
-We chose the ten studies guided by a desire to identify foundational
-studies that contribute to the general consensus in social psychology
-that group polarization reliably occurs across experimental
-settings---one author of one of the ten journal articles is a law
-professor who has written two trade books on group polarization to argue
-that group polarization occurs in real-world settings in addition to the
-laboratory, with political and other practical importance
-\autocite{Sunstein2009,Sunstein2019}. We chose studies that either had a
-significant influence on the study of group polarization and social
-psychology, in terms of a high citation count or adaptation in
-popularized trade books, or were designed to support one of four
-prominent theoretical explanations of group polarization. The studies
-must also have used an ordinal scale for measuring participant opinions
-(all but a handful of exceptions use ordinal measurements to measure
-group polarization; e.g.~blackjack bets
-\autocite{Blascovich1973,Blascovich1976} and mock jury monetary damages
-\autocite{Schkade2000}). We included at least one study that supported
-or was motivated by one of four established theoretical explanations
-(introduced below) of group polarization to show all have at least some
-supporting evidence that must be discarded. We also included two studies
-that lack a clear theoretical motivation. For the sake of orienting the
-reader to the group polarization literature and as a useful organizing
-feature for our analysis, we now introduce the theories that motivate
-eight of the ten studies. To further orient the reader, we then provide
-details of the experimental designs used to produce the 60 published
-positive detections of group polarization across 10 research articles
-that we scrutinize here. Group polarization experimental designs are
-diverse and inconsistent with one another in many ways, which
-complicates this or other potential meta-analyses and may imperil our
-ability to rigorously understand how regular the group polarization
-effect really is \autocite{Yarkoni2022,Almaatouq2022}.
-
-The first theoretical explanation whose supporting findings we include
-here is (1) social comparisons theory, which posits that an individual's
-opinions are updated so they stand out in a group enough to be
-``desirably distinctive'' (Myers, 1978, p.~562) but not too distinct to
-be viewed as anti-conformist \autocite{Brown1974,Sanders1977,Myers1978}.
-Second, \emph{persuasive arguments theory} posits that more extreme
-individuals have a greater number of more convincing arguments for their
-view, which draw more moderate group members to extreme opinions
-\autocite{Burnstein1973,Burnstein1975,Burnstein1977}. Third, the
-\emph{self-categorization theory} predicts that group polarization will
-occur as more moderate individuals judge the mean group opinion to be
-more extreme \autocite{Turner1989,Abrams1990,Krizan2007}. Fourth and
-finally, the social decision schemes explanation posits that the
-structure of \autocite{Davis1973,Zuber1992,Friedkin1999a}. We do not
-evaluate or compare these theories here, though a simpler explanation
-that extremists are more stubborn may be a more parsimonious explanation
-consistent with these four alternatives
-\autocite{Acemoglu2013,Guazzini2015,Turner2020}.
-
-Each of the ten group polarization studies we analyze here varied
-significantly in how groups were formed, how opinions were gathered, and
-other details. There was significant variation across experimental
-designs, participant population sizes, group sizes, and how like-minded
-groups were formed in the first place. We review some of these details
-below to provide the reader with some insight into the experimental
-designs that produced what we show are plausibly false positive
-detections of group polarization. We focus here on the questions that
-were asked, how like-minded novel experimental groups were formed,
-organizing our presentation by theoretical tradition.
-
-\subsubsection{Details of experimental conditions included in this
-meta-analysis}\label{details-of-experimental-conditions-included-in-this-meta-analysis}
-
-\paragraph*{Atheoretical}
-
-The first of ten studies we included in our meta-analysis was by
-Moscovici and Zavalloni (1969) \autocite{Moscovici1969} who sought to
-measure group polarization among groups of French high school students
-who discussed domestic and foreign politics. Moscovici and Zavalloni
-tested four conditions whether group polarization occurred for two
-different questions and two different rating systems: participants were
-asked their opinions on then-president Charles De Gaulle and their
-opinion about foreign aid from the USA either in the ``Opinion'' setting
-or the ``Judgement'' condition (see Table 4 on p.~132).
-
-The second, relatively more recent article is Schkade, Sunstein, and
-Hastie (2010) \autocite{Schkade2010} who used geographic location to
-construct like-minded groups by running a group polarization experiment
-in two cities in the state of Colorado in the United States: Colorado
-Springs, whose residents historically voted for republican candidates,
-and Boulder, whose residents historically voted for democratic
-candidates. Discussion groups were formed in which they measured group
-polarization when group members discussed global warming, affirmative
-action, and same-sex civil unions.
-
-\paragraph{Social comparisons}
-
-In developing the social comparisons theory of group polarization
-(described in Myers (1978) \autocite{Myers1978}), Myers and Bishop
-(1970), writing in \emph{Science,} \autocite{Myers1970} sought to
-measure group polarization among high school students in the United
-States who discussed racial attitudes and Myers (1975)
-\autocite{Myers1975} measured group polarization in groups of college
-students discussing the quality of a faculty member in one condition,
-and women's rights and issues in another condition.
-
-\paragraph{Persuasive arguments}
-
-\paragraph{Self-categorization}
-
-\paragraph{Social decisions scheme}
-
-The ten studies we included each tested several experimental conditions
-meant to elicit group polarization in novel experimental groups, 60 in
-total, which we briefly introduce here.\\
-
-\subsection{Research overview}\label{research-overview}
-
-To evaluate whether published detections of group polarization are
-plausibly false we developed and combined formal and generative models
-of group polarization to identify plausible simple consensus pre- and
-post-discussion opinion distributions that seem like group polarization
-when binned or measured on ordinal scales. Opinion distributions are
-assumed to be normal. We define a \emph{simple consensus pair} of
-\emph{latent opinion} distributions to be a pair of pre- and
-post-discussion distributions where the mean opinion is constant, but
-the pre-discussion variance is greater than the post-discussion
-variance. A \emph{latent distribution} of opinions represents the
-internal, unobservable distribution of opinions for a group or
-collection of groups. The \emph{observed distribution} is the
-distribution of reported ordinal opinion measurements. We then define a
-\emph{plausibly false detection} of group polarization to be one where
-the mean post-discussion observed distribution is greater in magnitude
-(i.e., more extreme) than the post-discussion observed distribution's,
-but both were measurements of a pair of latent simple consensus
-distributions.
-
-In the formal model, a simulated observed distribution is generated by
-binning a latent distribution into a finite number of ordinal scale
-bins, then integrating over the probability density in each bin. In the
-generative model, an observed distribution is simulated by first drawing
-participant opinions from a latent distribution, then counting the
-number of opinions falling within each bin. The formal model is used to
-identify which, if any, simple consensus pairs exist for a given
-reported detection of group polarization through an optimization routine
-that finds pre- and post-discussion variances that generate simulated
-observations that appear to be group polarization (or reports a failure
-to such variances). If the optimization routine finds a constant latent
-mean that generates simulated observed group polarization, then we say
-the published group polarization detection is plausibly false. If the
-routine fails to find a constant latent mean that generates simulated
-group polarization, then we cannot and do not claim the detection is
-plausibly false. We use the generative model to evaluate how frequently
-we would expect plausibly false detections when there are finite-size
-groups, since the formal model implicitly applies infinite-size groups.
-Therefore in our analyses, we also generate one thousand simulated group
-polarization datasets for each of the 55 plausibly false detections
-identified with the formal model and examine the distribution of group
-polarization effect sizes in terms of Cohen's \(d\). In the first round
-of simulations we set the number of observations to match those in the
-ten published studies we analyzed. Many or most simulated effect sizes
-are non-zero with these numbers of observations. To see if perhaps these
-non-zero effects were due to low statistical power from too few
-observations, we also tested a hypothetical study with ten times more
-participants. Finally, we evaluate whether an appropriate statistical
-model that accounts for the ordinal measurement of latent opinions can
-reliably identify simulated simple consensus in latent opinions when
-simulated observations appear as group polarization based on mean group
-opinion alone.
-
-% \begin{wrapfigure}{l}{0.5\textwidth}
-\begin{figure}
-  \caption{\textbf{Group polarization experimental model.} Latent opinions are 
-    represented as numbers corresponding to the valence and extremism of one's opinion
-(A). Before deliberating a topic of interest, all participants report their 
-opinion (B). Participants then deliberate the topic with others (C). Following
-deliberation participants report their opinion once more (D).}
-  \centering
-  \includegraphics[width=0.65\textwidth]{Figures/Model/ModelSketches.pdf}
-  \label{fig:experiment_model}
-% \end{wrapfigure}
-\end{figure}
+% \input{old-intro.tex}
 

### 2025-08-12 GroupPolarizationStatmod/paper/ms.tex
fuller explanation of code in README, started in Methods with new data-model flow fig

@@
 % Table~\ref{tab:modelVariables}).
+
+\begin{figure}
+  \centering
+    \includegraphics[width=0.8\textwidth]{Figures/Model/data-model-flow.png}
+  \caption{\textbf{Workflow of the GP-stats simulation and analysis pipeline
+  for estimating false detection rates under null-consistent conditions.} Colored boxes indicate file types: teal = R scripts, blue = data files, orange = plotting scripts, cyan = Slurm batch script.
+Arrows show the flow from parameter estimation (hillclimbing or Shiny app) to Bayesian null model fitting on the cluster, aggregation of simulated results, and final visualization.}
+  \label{fig:}
+\end{figure}
 

### 2025-08-12 GroupPolarizationStatmod/paper/intro-latest.tex
update to make ms build work

\subsection*{Abstract}\label{abstract}
\addcontentsline{toc}{subsection}{Abstract}

\emph{Group polarization} is the name for a form of consensus where
members of a like-minded group become, on average, more extreme in their
opinions after discussing a topic. Group polarization is important
because it may increase social and political tensions if moderates
become more extreme. Decades of studies have replicated detections of
group polarization, typically using an ordinal scale to measure
opinions. However, these studies did not account for ordinal
measurements, which can result in spurious detections of opinion change.
Lacking original data we can still calculate the probability that a
published detection of group polarization was spurious, i.e., to
calculate the \emph{false detection rate}. Across 54 group polarization
experimental conditions from ten representative journal articles, we
calculated false detection rates to be between 0.52 and 0.88, with a
median of 0.75, using a generative model of group polarization
experiments seeded with empirical data. We also use our model to develop
experimental designs that achieve an acceptable false discovery rate.
Much of group polarization research may be unreliable. This analysis can
help change that by enabling others to avoid this hidden pitfall. More
broadly this work demonstrates one important way that replication
success alone does not imply epistemic reliability.

\section{Introduction}\label{introduction}

\begin{quote}
In our introductory social psychology course, we have for many years
used the {[}group polarization experimental paradigm{]} as a laboratory
exercise. The exercise works beautifully, but one must be careful to
forewarn a class that {[}group polarization{]} does not occur with every
group\ldots and that the effect is not large (Brown 1986).
\end{quote}

\begin{quote}
One of the most robust findings in social psychology is that of attitude
polarization following discussion with like-minded others (Cooper,
Kelly, and Weaver 2001).
\end{quote}

If an extremist's opinion falls and it is measured with a Likert scale,
will it make a noise? The answer had better be yes if we want to
properly measure opinion change. Accurate measurements are critical for
developing rigorous and reliable methods to promote more stable,
sustainable, responsive, and efficient governments and institutions
(Mason 2018; Klein 2020). Unfortunately, social and behavioral
scientists have often used statistical methods that distort opinion
change when it's measured on an ordinal, Likert-style scale (e.g.~-3
indicates ``Strongly disagree'', +3 indicates ``Strongly agree'', and 0
is neutral) (Liddell and Kruschke 2018). This includes a phenomenon
called \emph{group polarization} that former Obama White House official
and law professor Cass Sunstein relied on to explain why ``people become
extremists'' and why ``political and cultural polarization'' is ``so
pervasive in America'', as the publisher summarizes (Sunstein 2009).
Sunstein suggests group polarization research provides a ``clue'' to
explaining how ``facism\ldots student radicalism\ldots Islamic
terrorism\ldots{[}t{]}he Rwandan genocide\ldots{[}e{]}thnic
Conflict\ldots acts of torture and humiliation by American soldiers at
Abu Ghraib'', and more (p.~1).

\begin{itemize}
\tightlist
\item
  We demonstrate here that there is a pervasive problem observed over
  decades of group polarization research where faulty statistical
  methods and incomplete reporting that should force the retraction of
  these studies from the canon. Effectively, then, empirical support for
  the theoretical concept \emph{group polarization} stops being so: the
  epistemic integrity of \emph{group polarization} is undermined with
  its empirical foundation losing mass.
\end{itemize}

We complement this critical analysis with a prescriptive analysis to
demonstrate that more appropriate statistical models do indeed lower the
false discovery rate by committing fewer false positive detections of
group polarization under a null model with shrinking opinion variance.
Recently, statistical best-practice is to use Bayesian models of
statistics, where social dynamics and opinion-reporting behaviors are
represented as causally-dependent probability distributions, where a
sample taken from one distribution (e.g., the opinion distribution of a
group in a group polarization experiment) acts as an input parameter to
another distribution (e.g., that represents the
individually-heterogeneous distortion of opinions when reported on an
ordinal scale). Our models follow current best practices in
psychometrics, social science, and Bayesian inference (Liddell and
Kruschke 2018; Kruschke 2015; McElreath 2020; Deffner et al. 2024), but
consolidate them in a new, useful way that enables the retroactive
discovery of null model parameters that generate spurious group
polarization, forcing published evidence for group polarization to be
classified as non-identifiable, and removed from evidence that can be
said to support group polarization.

Our methods can be turned on any data-driven social science analysis of
data where participants report a latent, subjective psychological
variable (mood, physiological state, etc.), and where various theories
supposedly battle it out to best represent the phenomenon of interest
via NHST performed on ordinal observations with weak, informal
theoretical models to motivate the studies. This may just be the first
of many bricks to fall since \emph{all 68} of the articles that
mentioned the word ``Likert'' and analyzed ordinal data used metric
models to detect latent psychological variable change across three
influential psychology journals, \emph{Psychological Science}, the
\emph{Journal of Experimental Psychology: General}, and the
\emph{Journal of Personality and Social Psychology} (Liddell and
Kruschke 2018).

This work is motivated by a core conviction that better thinking leads
to better science. Theoretical clarity, careful measurement, and
principled statistical modeling are not academic luxuries, but essential
tools for empirical science. Theoretical and statistical confusions in
the group polarization literature have lain dormant, undermining
measurements of opinion change in group polarization unbeknownst to
generations of researchers (Brown 1986).

\subsection{Identifiability and the Illusion of
Change}\label{identifiability-and-the-illusion-of-change}

\begin{itemize}
\tightlist
\item
  Many group polarization studies report opinion change based on
  pre/post differences, but assume a transparent mapping between
  observed scores and latent beliefs.
\item
  Yet ordinal opinion data can mask deeper structural ambiguities that
  invalidate these assumptions.
\item
  In what follows, we first describe the basic structure of group
  polarization experiments, then explain how ordinal measurement can
  lead to illusory shifts.
\item
  Finally, we introduce a null-consistency modeling procedure that
  operationalizes this insight, and show how it informs the credibility
  of reported effects.
\end{itemize}

\subsubsection{Experimental structure of group polarization
studies}\label{experimental-structure-of-group-polarization-studies}

\begin{itemize}
\tightlist
\item
  Most group polarization experiments follow a familiar structure:
  measure individual opinions, allow group discussion, then re-measure.
\item
  The central claim is that like-minded individuals become more extreme
  after deliberation.
\item
  Crucially, opinions are typically measured with Likert-style ordinal
  scales (e.g., from ``Strongly disagree'' to ``Strongly agree'').
\item
  Analyses often treat these as interval-scaled, calculating mean
  differences and testing for significance.
\end{itemize}

\subsubsection{The ordinal-continuous mismatch
problem}\label{the-ordinal-continuous-mismatch-problem}

\begin{itemize}
\tightlist
\item
  Ordinal scores do not have a fixed interval structure --- they
  represent ranked categories, not true distances.
\item
  A given change in score may reflect a shift in opinion, increased
  certainty, or both.
\item
  For example, if opinions become more tightly clustered (lower
  variance) post-deliberation, the group mean can appear to shift even
  when the latent mean remains unchanged.
\item
  This opens the door to \emph{null-consistent detections} --- apparent
  effects that emerge from measurement artifacts rather than real
  opinion movement.
\end{itemize}

\subsubsection{Description of the null-consistency modeling
procedure}\label{description-of-the-null-consistency-modeling-procedure}

\begin{itemize}
\tightlist
\item
  We formalize this concern using a generative model that holds the
  latent opinion mean constant while allowing variance to shrink.
\item
  The goal is to test whether an observed group polarization result
  could arise from such a stable-mean, shifting-variance process.
\item
  If this model can reproduce the pre/post measurement pattern, the
  observed result is non-diagnostic --- it is consistent with a null
  hypothesis of no true change.
\item
  This is a form of constructive falsification: it doesn't prove the
  effect is false, but shows it is not uniquely explained by change.
\end{itemize}

\subsubsection{Implications for statistical power and detection
credibility}\label{implications-for-statistical-power-and-detection-credibility}

\begin{itemize}
\tightlist
\item
  Even if an effect is real, detecting it reliably requires statistical
  power --- the ability to distinguish true shifts from noise or
  artifacts.
\item
  Studies using ordinal measures and small samples may be especially
  vulnerable to false positives under the null.
\item
  To assess this, we fit Bayesian ordered probit models to simulated
  data and estimate the power to detect effects of various sizes.
\end{itemize}

\subsubsection{Transition to empirical review and FDR
estimation}\label{transition-to-empirical-review-and-fdr-estimation}

\begin{itemize}
\tightlist
\item
  Having established a model-based definition of null-consistent
  detection, we now apply it to published group polarization
  experiments.
\item
  By combining this logic with simulations seeded from published data,
  we estimate false discovery rates across conditions.
\item
  This isn't a judgment about whether past detections were true. It's a
  demonstration of how to estimate false detection rates
  properly---using models that actually fit the structure of the data.
  We calculate FWER and FDR for the original study designs at three
  effect size levels---small, medium, and large---based on Cohen's
  guidelines (Cohen 1988, Ch. 2.2.3) and Bayesian methods from Liddell
  and Kruschke (Liddell and Kruschke 2018). The verdict is clear: even
  if these studies had used better models, most still wouldn't have had
  the power to rule out false positives. That makes replication a weak
  signal of reliability. If future group polarization studies want to
  produce meaningful result need stronger designs---and that starts with
  a clean break from outdated statistical habits.
\end{itemize}

The null-consistency procedure functions as a test of model
indistinguishability: it demonstrates that a reported group polarization
effect could plausibly arise from a stable-latent process combined with
changing opinion precision. This logic echoes foundational concerns in
causal inference, where two competing models---one causal, one not---can
imply the same observed data distribution and are therefore empirically
indistinguishable without further assumptions (Pearl 2000; Spirtes,
Glymour, and Scheines 2000). In such cases, inference from observation
to mechanism is invalid. Similarly, in psychometrics, it is well known
that different configurations of latent traits and response thresholds
can yield identical observed ordinal patterns, making the true structure
unidentifiable without additional constraints (Borsboom 2005). The
present approach does not attempt to estimate the probability of false
detection, but rather shows that under ordinal measurement, a null model
can reproduce the observed pattern---rendering the result non-diagnostic
of a true latent shift. This establishes a stringent standard: an effect
must not only be detected, but shown to be incompatible with plausible
null-generating mechanisms.

### 2025-08-13 GroupPolarizationStatmod/paper/intro-latest.tex
revised streamlined intro

@@
 
-\begin{itemize}
-\tightlist
-\item
-  We demonstrate here that there is a pervasive problem observed over
-  decades of group polarization research where faulty statistical
-  methods and incomplete reporting that should force the retraction of
-  these studies from the canon. Effectively, then, empirical support for
-  the theoretical concept \emph{group polarization} stops being so: the
-  epistemic integrity of \emph{group polarization} is undermined with
-  its empirical foundation losing mass.
-\end{itemize}
+% \begin{itemize}
+% \tightlist
+% \item
+We developed a generative model of opinion change to identify, if possible,
+null opinion change models that could spuriously appear like group polarization
+when naively measured using a Likert scale without accounting for the measurement procedure. 
+We further applied our method to estimate the family-wise error rate of
+detections of group polarization in a corpus of ten influential papers from 
+across several decades and from various theoretical traditions. We turn this criticism of past
+blunders constructive by providing a procedure to calculate what Cohen's $d$ should be
+used to guarantee a family-wise error rate (i.e., Type I error rate) of 5\% or less.
 
-We complement this critical analysis with a prescriptive analysis to
-demonstrate that more appropriate statistical models do indeed lower the
-false discovery rate by committing fewer false positive detections of
-group polarization under a null model with shrinking opinion variance.
-Recently, statistical best-practice is to use Bayesian models of
-statistics, where social dynamics and opinion-reporting behaviors are
-represented as causally-dependent probability distributions, where a
-sample taken from one distribution (e.g., the opinion distribution of a
-group in a group polarization experiment) acts as an input parameter to
-another distribution (e.g., that represents the
-individually-heterogeneous distortion of opinions when reported on an
-ordinal scale). Our models follow current best practices in
-psychometrics, social science, and Bayesian inference (Liddell and
-Kruschke 2018; Kruschke 2015; McElreath 2020; Deffner et al. 2024), but
-consolidate them in a new, useful way that enables the retroactive
-discovery of null model parameters that generate spurious group
-polarization, forcing published evidence for group polarization to be
-classified as non-identifiable, and removed from evidence that can be
-said to support group polarization.
+We calculate the severity of the problem in two steps. In the first step we 
+provide proof by example that there exists a null opinion change model in terms of \emph{latent, continuous}
+psychological opinions that looks like group polarization when measured via participant behavior, i.e.,
+by their response on an \emph{ordinal survey scale}. If such a null model is found, this means that we cannot
+rule out a null effect, given the published data. This means, therefore, that the published detection is
+just as likely false as it is true, so we say it is a \emph{plausibly false detection of group polarization},
+or a \emph{plausibly spurious detection} for short.
+
+In the second step of the analysis of whether, we simulate observations
+of group polarization experiments where participant opinions are drawn from the 
+null model distributions identified in the first step,
+then these are measured on a Likert scale for simulated group polarization survey
+responses on an ordinal scale. We then calculate the
+family-wise error rate (i.e., Type I error rate) and false detection rate using
+various effect sizes to determine significance of the effect in terms of Cohen's $d$.
+The family-wise error rate is defined as the probability that a known null 
+effect is detected as significant. We use this to then estimate the *false detection rate*, which is the
+probability that a detection of an effect is a false one, which depends on the relative prevalence of
+true and false positive detections in the literature.
+Finally, we put our model to constructive use to show what Cohen's $d$ must be set to
+to limit the family-wise error rate to 5\%, denoted $d*$. This will not help rescue any of
+the published results here found to be plausibly spurious. But, it provides a measure of the weakness of
+the published experimental designs independent of their shared fatal failure to 
+account for the measurement procedure: greater $d*$ indicates weaker designs.
+
+We find that 95\% (54 of 57) of group polarization detections are plausibly spurious.
+In other words, empirical support for group polarization has been seriously undermined
+just by removing these results from the count of group polarization replications.
+There is no reason to expect other papers to be more reliable. (Summarize other
+two findings, 1. FWER/FDR and 2. $d*$ for FWER < 5\%)
 
@@
 supposedly battle it out to best represent the phenomenon of interest
-via NHST performed on ordinal observations with weak, informal
-theoretical models to motivate the studies. This may just be the first
+via null-hypothesis significance testing using ordinal data as continuous. 
+This may just be the first
 of many bricks to fall since \emph{all 68} of the articles that
@@
 \emph{Journal of Personality and Social Psychology} (Liddell and
-Kruschke 2018).
+Kruschke 2018). This further confirms that replication is far from the only
+foundational crisis that behavioral science faces\cite{Yarkoni2022,Turner2022}.
 
-This work is motivated by a core conviction that better thinking leads
-to better science. Theoretical clarity, careful measurement, and
-principled statistical modeling are not academic luxuries, but essential
-tools for empirical science. Theoretical and statistical confusions in
-the group polarization literature have lain dormant, undermining
-measurements of opinion change in group polarization unbeknownst to
-generations of researchers (Brown 1986).
+Replication is not enough:
+strong, mechanistic models must be developed to be able to predict Type I and II
+error rates *ahead of time*, as we demonstrate here. We only demonstrate how to
+calculate expected Type I error rates for a measurement design, i.e., the rate of false 
+positives. Low statistical power exacerbates false detection rates. So, to be confident in replication,
+a pre-registered measurement design should use this style of simulation to 
+estimate the statistical power. Tangled, weak, informal
+verbal models of complex psychological and social phenomena like those that motivated
+empirical studies of group polarization tend to be vulnerable
+to stastical weaknesses. It is difficult to incorporate prior knowledge such
+as experiment structure without a sufficiently rigorous and appropriate 
+formal system \cite{Smaldino2017,Smaldino2020,Flake2020,Deffner2024}.
+
+% This work is motivated by a core conviction that better thinking leads
+% to better science. Theoretical clarity, careful measurement, and
+% principled statistical modeling are not academic luxuries, but essential
+% tools for empirical science. Theoretical and statistical confusions in
+% the group polarization literature have lain dormant, undermining
+% measurements of opinion change in group polarization unbeknownst to
+% generations of researchers (Brown 1986).
 

### 2025-08-13 GroupPolarizationStatmod/paper/intro-latest.tex
revised, expanded, focused intro-latest

@@
 strong, mechanistic models must be developed to be able to predict Type I and II
-error rates *ahead of time*, as we demonstrate here. We only demonstrate how to
+error rates \emph{ahead of time}, as we demonstrate here. We only demonstrate how to
 calculate expected Type I error rates for a measurement design, i.e., the rate of false 
@@
 
-% This work is motivated by a core conviction that better thinking leads
-% to better science. Theoretical clarity, careful measurement, and
-% principled statistical modeling are not academic luxuries, but essential
-% tools for empirical science. Theoretical and statistical confusions in
-% the group polarization literature have lain dormant, undermining
-% measurements of opinion change in group polarization unbeknownst to
-% generations of researchers (Brown 1986).
+\subsection{Group polarization research}
 
-\subsection{Identifiability and the Illusion of
-Change}\label{identifiability-and-the-illusion-of-change}
+(DO THE MINIMUM AFTER EVERYTHING ELSE)
 
-\begin{itemize}
-\tightlist
-\item
-  Many group polarization studies report opinion change based on
-  pre/post differences, but assume a transparent mapping between
-  observed scores and latent beliefs.
-\item
-  Yet ordinal opinion data can mask deeper structural ambiguities that
-  invalidate these assumptions.
-\item
-  In what follows, we first describe the basic structure of group
-  polarization experiments, then explain how ordinal measurement can
-  lead to illusory shifts.
-\item
-  Finally, we introduce a null-consistency modeling procedure that
-  operationalizes this insight, and show how it informs the credibility
-  of reported effects.
-\end{itemize}
+To situate this current paper it will help to introduce some relevant research
+traditions and common experimental designs in group polarization, using the ten
+studies included in our corpus as touchstones. To begin, *group polarization*
+is defined as...(continue basic definition with some examples from our corpus).
 
-\subsubsection{Experimental structure of group polarization
-studies}\label{experimental-structure-of-group-polarization-studies}
+Paragraph dos: 
+Common experimental, measurement, and statistical procedures to induce and
+detect group polarization
 
-\begin{itemize}
-\tightlist
-\item
-  Most group polarization experiments follow a familiar structure:
-  measure individual opinions, allow group discussion, then re-measure.
-\item
-  The central claim is that like-minded individuals become more extreme
-  after deliberation.
-\item
-  Crucially, opinions are typically measured with Likert-style ordinal
-  scales (e.g., from ``Strongly disagree'' to ``Strongly agree'').
-\item
-  Analyses often treat these as interval-scaled, calculating mean
-  differences and testing for significance.
-\end{itemize}
+Paragraph tres: How this procedure has been used in the great pointless theoretical milieu
+that raged for decades with apparently no strong determination on the outcome
+or even what the fight was all about.
 
-\subsubsection{The ordinal-continuous mismatch
-problem}\label{the-ordinal-continuous-mismatch-problem}
+Paragraph quatro e final: our corpus.
 
-\begin{itemize}
-\tightlist
-\item
-  Ordinal scores do not have a fixed interval structure --- they
-  represent ranked categories, not true distances.
-\item
-  A given change in score may reflect a shift in opinion, increased
-  certainty, or both.
-\item
-  For example, if opinions become more tightly clustered (lower
-  variance) post-deliberation, the group mean can appear to shift even
-  when the latent mean remains unchanged.
-\item
-  This opens the door to \emph{null-consistent detections} --- apparent
-  effects that emerge from measurement artifacts rather than real
-  opinion movement.
-\end{itemize}
+\subsection{Overview}
 
-\subsubsection{Description of the null-consistency modeling
-procedure}\label{description-of-the-null-consistency-modeling-procedure}
-
-\begin{itemize}
-\tightlist
-\item
-  We formalize this concern using a generative model that holds the
-  latent opinion mean constant while allowing variance to shrink.
-\item
-  The goal is to test whether an observed group polarization result
-  could arise from such a stable-mean, shifting-variance process.
-\item
-  If this model can reproduce the pre/post measurement pattern, the
-  observed result is non-diagnostic --- it is consistent with a null
-  hypothesis of no true change.
-\item
-  This is a form of constructive falsification: it doesn't prove the
-  effect is false, but shows it is not uniquely explained by change.
-\end{itemize}
-
-\subsubsection{Implications for statistical power and detection
-credibility}\label{implications-for-statistical-power-and-detection-credibility}
-
-\begin{itemize}
-\tightlist
-\item
-  Even if an effect is real, detecting it reliably requires statistical
-  power --- the ability to distinguish true shifts from noise or
-  artifacts.
-\item
-  Studies using ordinal measures and small samples may be especially
-  vulnerable to false positives under the null.
-\item
-  To assess this, we fit Bayesian ordered probit models to simulated
-  data and estimate the power to detect effects of various sizes.
-\end{itemize}
-
-\subsubsection{Transition to empirical review and FDR
-estimation}\label{transition-to-empirical-review-and-fdr-estimation}
-
-\begin{itemize}
-\tightlist
-\item
-  Having established a model-based definition of null-consistent
-  detection, we now apply it to published group polarization
-  experiments.
-\item
-  By combining this logic with simulations seeded from published data,
-  we estimate false discovery rates across conditions.
-\item
-  This isn't a judgment about whether past detections were true. It's a
-  demonstration of how to estimate false detection rates
-  properly---using models that actually fit the structure of the data.
-  We calculate FWER and FDR for the original study designs at three
-  effect size levels---small, medium, and large---based on Cohen's
-  guidelines (Cohen 1988, Ch. 2.2.3) and Bayesian methods from Liddell
-  and Kruschke (Liddell and Kruschke 2018). The verdict is clear: even
-  if these studies had used better models, most still wouldn't have had
-  the power to rule out false positives. That makes replication a weak
-  signal of reliability. If future group polarization studies want to
-  produce meaningful result need stronger designs---and that starts with
-  a clean break from outdated statistical habits.
-\end{itemize}
-
-The null-consistency procedure functions as a test of model
-indistinguishability: it demonstrates that a reported group polarization
-effect could plausibly arise from a stable-latent process combined with
-changing opinion precision. This logic echoes foundational concerns in
-causal inference, where two competing models---one causal, one not---can
-imply the same observed data distribution and are therefore empirically
-indistinguishable without further assumptions (Pearl 2000; Spirtes,
-Glymour, and Scheines 2000). In such cases, inference from observation
-to mechanism is invalid. Similarly, in psychometrics, it is well known
-that different configurations of latent traits and response thresholds
-can yield identical observed ordinal patterns, making the true structure
-unidentifiable without additional constraints (Borsboom 2005). The
-present approach does not attempt to estimate the probability of false
-detection, but rather shows that under ordinal measurement, a null model
-can reproduce the observed pattern---rendering the result non-diagnostic
-of a true latent shift. This establishes a stringent standard: an effect
-must not only be detected, but shown to be incompatible with plausible
-null-generating mechanisms.
+In the next section, we introduce the various formal and computational Methods 
+used to obtain our findings outlined above. We then explain our findings in detail in the
+Analysis section. We close with a Discussion of how this is not a defeat, but
+a new beginning for group polariation, given that there are good reasons to expect
+group polarization to occur. We also explain how to expanding this analysis to
+examine other published studies on group polarization and across social and behavioral
+science disciplines.

### 2025-08-13 GroupPolarizationStatmod/scripts/plot/analysis.R
minor cleanups

@@ -37,8 +37,6 @@ plot_ordinal_cohens <- function(ordinal_data_dir = "data/probit_fits",
                                 output_file = "paper/Figures/Analysis/ordinal_cohens.pdf",
                                 overwrite = FALSE) {
   
-  # df <- dir_ls(ordinal_data_dir, glob = "*.csv") %>%
-  #   read_csv() %>% unite(ExperimentID, ArticleTag, TreatmentTag)
   df <- load_probit_data(probit_fits_dir, sync_file, overwrite)
   
   df$Cohens_d <- cohens_d(df$LatentMeanPrePosteriorMean, df$LatentMeanPostPosteriorMean,
@@ -190,29 +188,27 @@ plot_sigval_for_low_fwer = function(target_fwer = 0.05,
   study_id_fct = levels(pre_tbl_focal$StudyID)
 
   tbl = sigval_for_low_fwer(pre_tbl) %>%
-      # calc_fdr_vs_significance(significance_vals = significance_vals)
-    # ) %>%
     mutate(ExperimentID = as.factor(paste(StudyID, ExperimentID, sep = "_")),
-           SigVal = as.numeric(as.character(SigVal))) #%>%
-  #     mutate(StudyID = fct_reorder(StudyID, FWER, mean), 
-  #            ExperimentID = fct_reorder(ExperimentID, FWER, mean))
-
-  # tbl_med = filter(
+           SigVal = as.numeric(as.character(SigVal))) 
+  
   # Then plot the result.
   p = ggplot(tbl, aes(x = SigVal, y = ExperimentID, fill = StudyID, color = StudyID)) +
       geom_vline(xintercept = c(0.2, 0.5, 0.8), linetype = "dashed", 
                  linewidth = 0.25, color = "darkgrey") + 
-      geom_bar(aes(fill = StudyID, color = StudyID), stat = "identity", width = 0.15)  +
-      geom_point(aes(fill = StudyID, color = StudyID), shape = 18, size = 5)  +
-      xlab(TeX("Significance $d^*$ for $\\alpha \\leq 0.05")) + 
+      geom_bar(aes(fill = StudyID, color = StudyID), 
+               stat = "identity", width = 0.15)  +
+      geom_point(aes(fill = StudyID, color = StudyID), shape = 18, size = 5) +
+      xlab(TeX("Significance $d^*$ for $\\alpha \\leq 0.05")) +
       ylab("Experiment ID") +
[…]

### 2025-08-13 GroupPolarizationStatmod/scripts/plot/model.R

@@ -5,19 +5,11 @@ library(dplyr)
 source("model.R")  # for binProb function used in ordinal plot function below.
 
 
-# mytheme = theme(
-#   axis.line = element_line(), legend.key=element_rect(fill = NA),
-#   text = element_text(size=16),# family = 'PT Sans'),
-#   # legend.key.width = unit(2, 'cm'),
-#   # legend.key.size = unit(1.5, 'lines'),
-#   panel.background = element_rect(fill = "white"),
-#   panel.grid.major.x = element_line(color = "lightgrey", linewidth = 0.1, linetype = 2)
-# )
-  mytheme = theme(
-    panel.border = element_blank(), axis.line = element_line(), 
-    text = element_text(size=16, family = 'PT Sans'), 
-    panel.background = element_rect(fill = "white")
-  )
+mytheme <- theme(
+  panel.border = element_blank(), axis.line = element_line(), 
+  text = element_text(size=16, family = 'PT Sans'), 
+  panel.background = element_rect(fill = "white")
+)
 
 plot_latent_pdf_integration = function(mu = 0, sd = 2.5, min_bin, max_bin, 
                                        bins, bin_colors, # xlim = c(-6.0, 6.0), 
@@ -162,40 +154,27 @@ make_distros_consensus_figure =
 }
 
 
-# Run figure-making routines using data from CaseStudies.xlsx
-# Get schkade2010 data
+# Run figure-making routines using real seed data from a plausibly false detection.
 case_studies_tbl = read_csv("data/StudiesAnalysis.csv")
 
-# XXX note that the analyzed study CSV uses column name "ArticleTag" not StudyID
-# and "TreatmentTag" not ExperimentID. 
-# TODO Not sure right now where this change is done, need to find out
-# schkade2010 = filter(case_studies_tbl, ArticleTag == "Schkade2010" &
-#                      TreatmentTag == "COSprings-CivilUnions")
-                         
-# mu = schkade2010$LatentMean
-# sigma_pre = schkade2010$LatentSDPre
-# sigma_post = schkade2010$LatentSDPost
-# min_bin = schkade2010$MinBinValue
-# max_bin = schkade2010$MaxBinValue
-# bins = min_bin:max_bin
-
-
 # Trying Moscovici "Americans" as example to match other model figure.
[…]

### 2025-08-13 GroupPolarizationStatmod/paper/ms.tex
combined and cleaned up software sketch and shiny app screenshot

@@
 
-% We begin by developing a formal model of ordinal measurement of latent
-% continuous opinions (model variables and parameters are given in
-% Table~\ref{tab:modelVariables}).
-
 \begin{figure}
   \centering
-    \includegraphics[width=0.8\textwidth]{Figures/Model/data-model-flow.png}
+    % \includegraphics[width=0.8\textwidth]{Figures/Model/data-model-flow.png}
+    \includegraphics[width=\textwidth]{Figures/Model/data-model-flow-w-webapp.pdf}
   \caption{\textbf{Workflow of the GP-stats simulation and analysis pipeline
-  for estimating false detection rates under null-consistent conditions.} Colored boxes indicate file types: teal = R scripts, blue = data files, orange = plotting scripts, cyan = Slurm batch script.
-Arrows show the flow from parameter estimation (hillclimbing or Shiny app) to Bayesian null model fitting on the cluster, aggregation of simulated results, and final visualization.}
-  \label{fig:}
+  for estimating false detection rates under null-consistent conditions.} Colored boxes indicate file types: teal = R scripts, blue = data files, orange = plotting scripts, cyan
+= Slurm batch script. Arrows show the flow from parameter estimation (hillclimbing or Shiny app) to Bayesian null model fitting on the cluster, aggregation of simulated results,
+and final visualization. (D) shows the web app interface that takes researcher guesses at
+simple consensus models that generate observations, and the app passes the guess to initialize
+the hillclimbing search algorithm for finding spurious group polarization.}
+  \label{fig:data-model-flow.png}
 \end{figure}
@@
 
-\begin{figure}
-  \caption{\textbf{Shiny web application for user-guided hillclimbing to find
-  parameters that generate spurious group polarization.} This web application enables users to
-create new analyses of experimental conditions from published studies. Users 
-provide initial guesses for the algorithm to use as starting points in 
-finding variances that generate spurious group polarization. Users inspect 
-model fits provided by the algorithm, adjusting parameters if necessary, and
-mark an experimental condition as ``Plausibly spurious'' if a reasonable fit
-is indeed found.\mt{Maybe add some A, B, C guides to the figure and include
-in parentheses here to go with the text?}}
-  \centering
-    \includegraphics[width=0.75\textwidth]{Figures/Model/WebApp.png}
-  \label{fig:webapp}
-\end{figure}
+% \begin{figure}
+%   \caption{\textbf{Shiny web application for user-guided hillclimbing to find
+%   parameters that generate spurious group polarization.} This web application enables users to
+% create new analyses of experimental conditions from published studies. Users 
+% provide initial guesses for the algorithm to use as starting points in 
+% finding variances that generate spurious group polarization. Users inspect 
+% model fits provided by the algorithm, adjusting parameters if necessary, and
+% mark an experimental condition as ``Plausibly spurious'' if a reasonable fit
+% is indeed found.\mt{Maybe add some A, B, C guides to the figure and include
+% in parentheses here to go with the text?}}
+%   \centering
+%     \includegraphics[width=0.75\textwidth]{Figures/Model/WebApp.png}
+%   \label{fig:webapp}
+% \end{figure}
 

### 2025-08-14 GroupPolarizationStatmod/paper/ms.tex
restructured Methods; ready for final drafting

@@
 
-\begin{figure}
-  \centering
-    % \includegraphics[width=0.8\textwidth]{Figures/Model/data-model-flow.png}
-    \includegraphics[width=\textwidth]{Figures/Model/data-model-flow-w-webapp.pdf}
-  \caption{\textbf{Workflow of the GP-stats simulation and analysis pipeline
-  for estimating false detection rates under null-consistent conditions.} Colored boxes indicate file types: teal = R scripts, blue = data files, orange = plotting scripts, cyan
-= Slurm batch script. Arrows show the flow from parameter estimation (hillclimbing or Shiny app) to Bayesian null model fitting on the cluster, aggregation of simulated results,
-and final visualization. (D) shows the web app interface that takes researcher guesses at
-simple consensus models that generate observations, and the app passes the guess to initialize
-the hillclimbing search algorithm for finding spurious group polarization.}
-  \label{fig:data-model-flow.png}
-\end{figure}
+(METHODS OVERVIEW)
 
-Metascience frames our analysis since metascientific theory provides a framework for
-our goal of calculating the false discovery rate across experimental conditions,
-journal articles, and the ten-paper corpus. The false discovery rate is then
-the Type I error rate relative to the rate of positive discoveries, true or false.
-We calculate via simulation and data fitting
-the Type I error rate for the experimental designs used in experimental conditions
-that, under a preliminary large-$N$ model analysis, generated group
-polarization upon ordinal measurement of simple consensus latent opinion 
-distributions. These can then be aggregated to estimate the false discovery rate
-for each journal article and for the ten-article corpus, which is representative of the
-false discovery rate for group polarization studies.
-
-\subsection{Metascience of the false detection rate in group polarization}
-
-The false detection rate is the scaled probability that a detection, $D$, of
-an effect, $E$, is in fact a false detection and there is \emph{not} an effect,
-denoted $\neg E$. The probability that a detection $D$ has occurred when there
-was no effect, i.e. when $\neg E$, is denoted $\alpha = \Pr(D | \neg E)$.
-This is known as the Type I or family-wise error rate. To calculate the 
-false detection rate ($\fdr$), one must compare $\alpha$ to the statistical power,
-the probability of a detection when there is an effect, denoted $W = \Pr(D | E)$.
-Both these values must be scaled by the rate at which effects occur, denoted $b$.
-Using the notation outlined above, the false detection rate is
-\begin{equation}
-  \fdr = \frac{(1 - b)\alpha}{(1 - b)\alpha + bW}.
-  \label{eq:fdr}
-\end{equation}
-
-In group polarization, $E$ represents the group polarization \emph{effect}, and
-$\neg E$ represents its absence. Simple consensus is one way for group polarization
-to be absent, i.e., simple consensus is one form of $\neg E$. The presence of
-an effect $E$ is independent of whether an effect was \emph{detected}, 
-represented by $D$ if it was detected and $\neg D$ if not.
-
-The group polarization studies are journal articles that present
-the results of multiple experiments to support some hypothesis. To organize
-the presentation of false detection rates across the ten studies in our
-analysis, we first will organize false detection rates by experiment, where
-we created `ExperimentID` tags for each of the sixty experiments across ten
-studies. We also created `StudyID` tags for each of the ten journal articles
-with the key `{Author}{Year}`, e.g., \texttt{Schkade2010} for Schkade, et al.,
-(2010) \cite{Schkade2010}.
-
-We calculate $\alpha = \Pr(D | \neg E)$ through simulation for experimental
-conditions, 
-and use optimistic values of
-$b=0.1$~\cite{Ioannidis2005,McElreath2015} and $W=0.8$~\cite{Smaldino2016NatSel}
-in the main text and in the compare with more pessimistic, yet realistic, estimates
-of these values, as well as even more optimistic estimates.
-
-\subsection{How study design can induce false detections of group polarization}
-\label{subsec:false-detections}
-
-An opinion is never measured directly, only a behavior that indicates an opinion
-in a behavioral experiment or survey.  For example, the location of a mark
-that a participant makes is measured, not the complex neural activity that
-underlies forming opinions (Figure~\ref{fig:experiment_model}A) \mt{Some of this
-needs to be Intro'd}.
-Theoretically, an internal opinion in the form of neural activity is a
-\textit{latent opinion}, from the Latin \emph{latentem} meaning ``lie
-hidden''.  Participants report \text{observed opinions}, typically as one of
-a finite number of ranked, successive integer values that form the
-\emph{selection set}. The minimum possible value in the selection set might
-mean ``strongly disagree'' and the maximum value represents ``strongly
-agree''.
-
-% \begin{wrapfigure}{r}{0.5\textwidth}
-\begin{figure}
-  \caption{\textbf{Opinion measurement model and spurious group polarization.} 
-  Latent model participant opinions are drawn from a normal distribution with the
-  listed parameters (A,C). The pre-deliberation latent distribution has high
-  variance, but appears polarized when measured via integration (A).  The latent
-  distribution's variance is expected to decrease during deliberation (B) via
-  consensus, but in this example the latent mean is static, 
-$\mupre = \mupost$. However, when the post-deliberation distribution is measured,
-the simulated observed mean increases, i.e., spurious
-group polarization occurs (C).}
-  \centering
-  \includegraphics[width=0.65\textwidth]{Figures/Model/latent-ordinal-distros.pdf}
-  \label{fig:distros}
-\end{figure}
-% \end{wrapfigure}
-
-Group polarization experiments share a common experimental design where,
-first, participant opinions are gathered on a topic of interest. Next, groups
-are formed that have some initial opinion extremism. Participants deliberate
-or discuss the topic for some set time. After the set time, participants
-again give their opinion. Group polarization is said to occur if the
-mean group opinion became more extreme, i.e., if the magnitude of the mean
-group opinion is greater \emph{post-deliberation} compared to
-\emph{pre-deliberation}. For example,
-\cite{Moscovici1969} asked Parisian high school students to indicate their
-degree of agreement with the statement, ``American economic aid is always
-used for political pressure'' (Figure~\ref{fig:experiment_model}A). There
-were seven options in the selection set, -3 indicating ``strongly disagree'',
-+3 indicating ``strongly agree'', and 0 was neutral. Moscovici and Zavalloni
-report that the mean pre-deliberation opinion was -0.61 and the mean
-post-deliberation opinion was -1.09 (Moscovici and Zavalloni, 1969; Table 4,
-p. 132). This is a -0.48 shift to more extreme disagreement with the 
-statement, reported by Moscovici and Zavalloni as statistically significant
-in Table 5 on p. 132.
+\subsection{Probability theory of spurious group polarization}
 
@@
 experiments.
+
+\begin{figure}
+  \centering
+    % \includegraphics[width=0.8\textwidth]{Figures/Model/data-model-flow.png}
+    \includegraphics[width=\textwidth]{Figures/Model/data-model-flow-w-webapp.pdf}
+  \caption{\textbf{Workflow of the GP-stats simulation and analysis pipeline
+  for estimating false detection rates under null-consistent conditions.} Colored boxes indicate file types: teal = R scripts, blue = data files, orange = plotting scripts, cyan
+= Slurm batch script. Arrows show the flow from parameter estimation (hillclimbing or Shiny app) to Bayesian null model fitting on the cluster, aggregation of simulated results,
+and final visualization. (D) shows the web app interface that takes researcher guesses at
+simple consensus models that generate observations, and the app passes the guess to initialize
+the hillclimbing search algorithm for finding spurious group polarization.}
+  \label{fig:data-model-flow.png}
+\end{figure}
+
+\subsection{Metascience of the false detection rate in group polarization}
+
+Metascience provides a framework for
+calculating the false discovery rate across experimental conditions,
+aggregated to the level of journal articles and the ten-paper corpus. 
+The false discovery rate is then the Type I error rate relative to the rate of
+positive discoveries, true or false.
+We calculate via simulation and data fitting
+the Type I error rate for the experimental designs used in experimental conditions
+that, under a preliminary large-$N$ model analysis, generated group
+polarization upon ordinal measurement of simple consensus latent opinion 
+distributions. These can then be aggregated to estimate the false discovery rate
+for each journal article and for the ten-article corpus, which is representative of the
+false discovery rate for group polarization studies.
+
+The false detection rate is the scaled probability that a detection, $D$, of
+an effect, $E$, is in fact a false detection and there is \emph{not} an effect,
+denoted $\neg E$. The probability that a detection $D$ has occurred when there
+was no effect, i.e. when $\neg E$, is denoted $\alpha = \Pr(D | \neg E)$.
+This is known as the Type I or family-wise error rate. To calculate the 
+false detection rate ($\fdr$), one must compare $\alpha$ to the statistical power,
+the probability of a detection when there is an effect, denoted $W = \Pr(D | E)$.
+Both these values must be scaled by the rate at which effects occur, denoted $b$.
+Using the notation outlined above, the false detection rate is
+\begin{equation}
+  \fdr = \frac{(1 - b)\alpha}{(1 - b)\alpha + bW}.
+  \label{eq:fdr}
+\end{equation}
+
+In group polarization, $E$ represents the group polarization \emph{effect}, and
+$\neg E$ represents its absence. Simple consensus is one way for group polarization
+to be absent, i.e., simple consensus is one form of $\neg E$. The presence of
+an effect $E$ is independent of whether an effect was \emph{detected}, 
+represented by $D$ if it was detected and $\neg D$ if not.
+
+The group polarization studies are journal articles that present
+the results of multiple experiments to support some hypothesis. To organize
+the presentation of false detection rates across the ten studies in our
+analysis, we first will organize false detection rates by experiment, where
+we created `ExperimentID` tags for each of the sixty experiments across ten
+studies. We also created `StudyID` tags for each of the ten journal articles
+with the key `{Author}{Year}`, e.g., \texttt{Schkade2010} for Schkade, et al.,
+(2010) \cite{Schkade2010}.
+
+We calculate $\alpha = \Pr(D | \neg E)$ through simulation for experimental
+conditions, 
+and use optimistic values of
+$b=0.1$~\cite{Ioannidis2005,McElreath2015} and $W=0.8$~\cite{Smaldino2016NatSel}
+in the main text and in the compare with more pessimistic, yet realistic, estimates
+of these values, as well as even more optimistic estimates.
 

### 2025-08-14 GroupPolarizationStatmod/scripts/plot/analysis.R

@@ -34,19 +34,21 @@ rhg_cols_10 <- rev(c("#771C19", "#AA3929", "#E25033", "#F27314", "#F8A31B",
 
 plot_ordinal_cohens <- function(ordinal_data_dir = "data/probit_fits",
                                 sync_file = "data/probit_fits/all.csv",
-                                output_file = "paper/Figures/Analysis/ordinal_cohens.pdf",
+                                output_file = 
+                                  "paper/Figures/Analysis/ordinal_cohens.pdf",
                                 overwrite = FALSE) {
   
   df <- load_probit_data(probit_fits_dir, sync_file, overwrite)
   
-  df$Cohens_d <- cohens_d(df$LatentMeanPrePosteriorMean, df$LatentMeanPostPosteriorMean,
-                          df$LatentMeanPrePosteriorSD, df$LatentMeanPostPosteriorSD)
+  df$Cohens_d <- cohens_d(df$LatentMeanPrePosteriorMean, 
+                          df$LatentMeanPostPosteriorMean,
+                          df$LatentMeanPrePosteriorSD, 
+                          df$LatentMeanPostPosteriorSD)
   
   ggplot(df, mapping = aes(x = Cohens_d, y = ExperimentID)) + 
     geom_vline(xintercept = 0, color = "red") + 
     geom_vline(xintercept = c(-1, 1), color="red", linetype="dotted") + 
-    geom_boxplot() #+
-    # xlim(-2.0, 2.0)
+    geom_boxplot() 
   
   ggsave(output_file)
 }
@@ -221,8 +223,8 @@ plot_fdr()
 
 plot_fdr(aggregate = TRUE, save_path = "paper/Figures/Analysis/fdr_vs_study.pdf")
 
-plot_fdr(aggregate = TRUE, 
-         save_path = "paper/Figures/Analysis/fdr_vs_study_w_nonid.pdf", 
+plot_fdr(aggregate = TRUE,
+         save_path = "paper/Figures/Analysis/fdr_vs_study_w_nonid.pdf",
          use_non_identified = TRUE)
 
 plot_sigval_for_low_fwer()

### 2025-08-14 GroupPolarizationStatmod/paper/ms.tex
minor typo fix
wrote Methods intro

@@
 
-(METHODS OVERVIEW)
+To find examples of plausibly spurious group polarization we guess at what
+latent mean, combined with a latent pre- and post-deliberation 
+latent standard deviations, could generate it. We calculate the observed group
+polarization for these latent parameters by transforming the latent pre- and
+post-deliberation distributions from continuous normal distributions to ordinal
+"observations" by integrating over each ordinal bin, then calculating the observed
+mean for each experiment *phase*, i.e., both the pre- and post-deliberation phases. 
+If the null model can generate group polarization in this case, then the detection
+is labelled *plausibly false*, i.e., *plausibly spurious*, and we call the 
+generating latent parameters the *plausible null parameters*.
+
+To evaluate whether the plausible null parameters we identify could realistically
+fool a significance test, we implemented a generative simulation of each plausibly
+false detection of group polarization and calculated the *family-wise error rate
+(FWER)* and the *false discovery rate (FDR)*. We use the published number of
+participants $N$ for each of the 54 conditions that yielded plausibly false
+detections of group polarization. The FWER is also known as the Type I
+error, or the false positive rate: it is the probability that a detection of an
+effect occurs given that in reality there was a null effect.  The FDR is the
+probability that a published positive detection of an effect is a false positive,
+i.e., a detection of a null effect. High FWER and FDR for an experimental condition
+indicates that its design is weak, since we identify only one of perhaps infinitely
+many latent parameter combinations that can generate spurious group polarization.
+
+We use this setup to demonstrate how *adversarial* style testing of experimental
+designs with counterfactual null models can be used to estimate important
+metascience parameters. In this study we use our FWER calculations to estimate
+the minimum Cohen's $d$ that should be used to determine whether an effect is
+``significant'' or not that will limit the FWER to 5\%. In the Discussion we
+sketch how our model can then be extended to rigorously calculate statistical power,
+i.e., the probability that a significant effect is detected given that there 
+is indeed a significant effect.
 
@@
 \emph{Spurious} group polarization is when mean observed opinions seem to increase in 
-extremity when the latent mean held constant. We simulate spurious
+extremity when the latent mean is held constant. We simulate spurious
 group polarization by setting $\mupre = \mupost = \mu$ and setting the variances

### 2025-08-14 GroupPolarizationStatmod/paper/intro-latest.tex
further cleanup, expansion, and connection of intro and methods

@@
 
-Paragraph tres: How this procedure has been used in the great pointless theoretical milieu
-that raged for decades with apparently no strong determination on the outcome
-or even what the fight was all about.
+\subsection{How study design can induce false detections of group polarization}
+\label{ssubsec:false-detections}
 
-Paragraph quatro e final: our corpus.
+An opinion is never measured directly, only a behavior that indicates an opinion
+in a behavioral experiment or survey.  For example, the location of a mark
+that a participant makes is measured, not the complex neural activity that
+underlies forming opinions.
+Theoretically, an internal opinion in the form of neural activity is a
+\textit{latent opinion}, from the Latin \emph{latentem} meaning ``lie
+hidden''.  Participants report \text{observed opinions}, typically as one of
+a finite number of ranked, successive integer values that form the
+\emph{selection set}. The minimum possible value in the selection set might
+mean ``strongly disagree'' and the maximum value represents ``strongly
+agree''.
+
+Group polarization experiments share a common experimental design where,
+first, participant opinions are gathered on a topic of interest. Next, groups
+are formed that have some initial opinion extremism. Participants deliberate
+or discuss the topic for some set time. After the set time, participants
+again give their opinion. Group polarization is said to occur if the
+mean group opinion became more extreme, i.e., if the magnitude of the mean
+group opinion is greater \emph{post-deliberation} compared to
+\emph{pre-deliberation}. For example,
+\cite{Moscovici1969} asked Parisian high school students to indicate their
+degree of agreement with the statement, ``American economic aid is always
+used for political pressure''. There
+were seven options in the selection set, -3 indicating ``strongly disagree'',
++3 indicating ``strongly agree'', and 0 was neutral. Moscovici and Zavalloni
+report that the mean pre-deliberation opinion was -0.61 and the mean
+post-deliberation opinion was -1.09 (Moscovici and Zavalloni, 1969; Table 4,
+p. 132). This is a -0.48 shift to more extreme disagreement with the 
+statement, reported by Moscovici and Zavalloni as statistically significant
+in Table 5 on p. 132.
+
+(PREVIOUS TWO PARAGRAPHS SET UP THIS FINAL PARAGRAPH: CONTINUE TO USE MOSCOVICI 1969 TO
+EXPLAIN IN THE SIMPLEST TERMS HOW FAILING TO ACCOUNT FOR MEASUREMENT PROCEDURE
+CAN LEAD TO FALSE POSITIVES; see Figure \ref{fig:distros})
 

### 2025-08-14 GroupPolarizationStatmod/paper/ms.tex
further cleanup, expansion, and connection of intro and methods

@@
 
-To find examples of plausibly spurious group polarization we guess at what
-latent mean, combined with a latent pre- and post-deliberation 
-latent standard deviations, could generate it. We calculate the observed group
-polarization for these latent parameters by transforming the latent pre- and
-post-deliberation distributions from continuous normal distributions to ordinal
-"observations" by integrating over each ordinal bin, then calculating the observed
-mean for each experiment *phase*, i.e., both the pre- and post-deliberation phases. 
-If the null model can generate group polarization in this case, then the detection
-is labelled *plausibly false*, i.e., *plausibly spurious*, and we call the 
-generating latent parameters the *plausible null parameters*.
+To find examples of plausibly spurious group polarization we guess at what latent
+mean, combined with a latent pre- and post-deliberation latent standard deviations,
+could generate it. We calculate the observed group polarization for these latent
+parameters by transforming the latent pre- and post-deliberation distributions from
+continuous normal distributions to ordinal ``observations'' by integrating over each
+ordinal bin, then calculating the observed mean for each experiment \emph{phase},
+i.e., both the pre- and post-deliberation phases.  If the null model can generate
+group polarization in this case, then the detection is labelled \emph{plausibly
+false}, i.e., \emph{plausibly spurious}, and we call the generating latent
+parameters the \emph{plausible null parameters}.
 
@@
 fool a significance test, we implemented a generative simulation of each plausibly
-false detection of group polarization and calculated the *family-wise error rate
-(FWER)* and the *false discovery rate (FDR)*. We use the published number of
+false detection of group polarization and calculated the \emph{family-wise error rate
+(FWER)} and the \emph{false discovery rate (FDR)}. We use the published number of
 participants $N$ for each of the 54 conditions that yielded plausibly false
@@
 
-We use this setup to demonstrate how *adversarial* style testing of experimental
+We use this setup to demonstrate how \emph{adversarial} style testing of experimental
 designs with counterfactual null models can be used to estimate important
@@
 \begin{equation}
-  s = \meanobs_\mathrm{post} - \meanobs_\mathrm{pre}~,
+  g = \meanobs_\mathrm{post} - \meanobs_\mathrm{pre}~,
   \label{eq:group-polarization}
@@
 \noindent
+
 \emph{Spurious} group polarization is when mean observed opinions seem to increase in 
-extremity when the latent mean is held constant. We simulate spurious
-group polarization by setting $\mupre = \mupost = \mu$ and setting the variances
-to values known to induce spurious group polarization. 
-We explain how to find a constant latent mean, $\mu$, and pre- and post-deliberation
-variances, $\sigmapre$ and $\sigmapost$, respectively, that generate spurious 
-detections of group polarization in the next section. 
-% We describe how we find $\mu$,
-% $\sigmapre$ and $\sigmapost$ next in Section~\ref{sec:finding-variances}.
+extremity when the latent mean is held constant. Call the collection of
+latent null model distribution parameters $\lambda = (\mu, \sigma_{pre}, \sigma_{post})$.
+The group polarization opinion shift can then be written as a function of $\lambda$,
+$g(\lambda)$. Formally, then, spurious group polarization is plausible if 
+there exists at least one $\lambda$ for which $g(\lambda) = g_{pub} > 0$, where
+$g_{pub}$ is the published group polarization shift value.
+
+Next, we explain how to find a set of latent null parameters, $\lambda$, that
+generate spurious group polarization when naively measured on an ordinal scale.
+
+\begin{figure}
+  \caption{\textbf{Opinion measurement model and spurious group polarization.} 
+  Latent model participant opinions are drawn from a normal distribution with the
+  listed parameters (A,C). The pre-deliberation latent distribution has high
+  variance, but appears polarized when measured via integration (A).  The latent
+  distribution's variance is expected to decrease during deliberation (B) via
+  consensus, but in this example the latent mean is static, 
+$\mupre = \mupost$. However, when the post-deliberation distribution is measured,
+the simulated observed mean increases, i.e., spurious
+group polarization occurs (C).}
+  \centering
+  \includegraphics[width=0.95\textwidth]{Figures/Model/latent-ordinal-distros.pdf}
+  \label{fig:distros}
+\end{figure}
 

### 2025-08-14 GroupPolarizationStatmod/paper/ms.tex
changed all betas to lambdas

@@
 using Equation~\ref{eq:group-polarization}). Define the collection of these
-parameters $\beta = \{\mu, \sigmapre, \sigmapost\}$. We identify $\beta$ through a
-hillclimbing optimization algorithm that finds $\beta$ that minimizes the difference
+parameters $\lambda = \{\mu, \sigmapre, \sigmapost\}$. We identify $\lambda$ through a
+hillclimbing optimization algorithm that finds $\lambda$ that minimizes the difference
 between reported pre- and post-deliberation mean opinions and simulated ones
@@
 
-The algorithm to find $\beta$ holds $\mu$ constant while searching for $\sigmapre$
+The algorithm to find $\lambda$ holds $\mu$ constant while searching for $\sigmapre$
 and $\sigmapost$ that generate observed $\meanobspre$ and $\meanobspost$, within a
@@
 The algorithm requires a researcher to specify an initial
-guess for $\beta$ and auxiliary parameter settings, which generally vary between
+guess for $\lambda$ and auxiliary parameter settings, which generally vary between
 experimental conditions.  Furthermore, the algorithm does not always produce
-reasonable distributions with a given $\beta$, and sometimes it does not produce
+reasonable distributions with a given $\lambda$, and sometimes it does not produce
 spurious group polarization.  Therefore, it is necessary to sanity check the latent
-and ordinal opinion distributions for the $\beta$ found by the algorithm. If all is
+and ordinal opinion distributions for the $\lambda$ found by the algorithm. If all is
 well, the experimental condition can be marked as plausibly spurious.  To tame the
@@
 some suspicion. If there exists set of latent simple consensus 
-parameters $\beta$ that create binned (Equation~\ref{eq:ordinal-frequency})
+parameters $\lambda$ that create binned (Equation~\ref{eq:ordinal-frequency})
 opinion distributions that naively seem to be group polarized, then the means
@@
 distributions that generated spurious group polarization, i.e., with
-parameters labelled $\beta_e$ found in the hillclimbing step. Binning these
+parameters labelled $\lambda_e$ found in the hillclimbing step. Binning these
 random values simulates observations. $N_e$ is the number
@@
 conditions that did not claim to be positive findings of group polarization.
-Of the remaining 57, we identified simple consensus $\beta$ that generate spurious
+Of the remaining 57, we identified simple consensus $\lambda$ that generate spurious
 group polarization in 54 experimental conditions. In the other three, we did
-identify $\beta$ that generate spurious group polarization, but the initial
-simulated distributions were highly polarized. We rejected $\beta$ generated for
+identify $\lambda$ that generate spurious group polarization, but the initial
+simulated distributions were highly polarized. We rejected $\lambda$ generated for
 two spurious group polarization from Schkade, et al., (2010) because
@@
 polarized~\cite[Figure 1, p. 234]{Schkade2010}. The other condition where we
-found but rejected spurious group polarization $\beta$ was in a question 
+found but rejected spurious group polarization $\lambda$ was in a question 
 about a ``bad teacher'' in Myers (1975)~\cite{Myers1975a}, where lower values

### 2025-08-14 GroupPolarizationStatmod/paper/ms.tex
added date
fixed error in algorithm logic

--- 2025-08-14 GroupPolarizationStatmod/paper/ms.tex
@@
 
-To find examples of plausibly spurious group polarization we guess at what
-latent mean, combined with a latent pre- and post-deliberation 
-latent standard deviations, could generate it. We calculate the observed group
-polarization for these latent parameters by transforming the latent pre- and
-post-deliberation distributions from continuous normal distributions to ordinal
-"observations" by integrating over each ordinal bin, then calculating the observed
-mean for each experiment *phase*, i.e., both the pre- and post-deliberation phases. 
-If the null model can generate group polarization in this case, then the detection
-is labelled *plausibly false*, i.e., *plausibly spurious*, and we call the 
-generating latent parameters the *plausible null parameters*.
-
-To evaluate whether the plausible null parameters we identify could realistically
-fool a significance test, we implemented a generative simulation of each plausibly
-false detection of group polarization and calculated the *family-wise error rate
-(FWER)* and the *false discovery rate (FDR)*. We use the published number of
-participants $N$ for each of the 54 conditions that yielded plausibly false
-detections of group polarization. The FWER is also known as the Type I
-error, or the false positive rate: it is the probability that a detection of an
-effect occurs given that in reality there was a null effect.  The FDR is the
-probability that a published positive detection of an effect is a false positive,
-i.e., a detection of a null effect. High FWER and FDR for an experimental condition
-indicates that its design is weak, since we identify only one of perhaps infinitely
-many latent parameter combinations that can generate spurious group polarization.
-
-We use this setup to demonstrate how *adversarial* style testing of experimental
-designs with counterfactual null models can be used to estimate important
-metascience parameters. In this study we use our FWER calculations to estimate
-the minimum Cohen's $d$ that should be used to determine whether an effect is
-``significant'' or not that will limit the FWER to 5\%. In the Discussion we
-sketch how our model can then be extended to rigorously calculate statistical power,
-i.e., the probability that a significant effect is detected given that there 
-is indeed a significant effect.
+(METHODS OVERVIEW)
 
@@
           \Comment{Update $\sigma$ if the error is less than before.}
-      \EndIf
         \State $\Delta \gets \bar{o}(\mu, \sigma) - \bar{\mathbf{o}}$
           \Comment{Update simulation error if $\sigma$ is updated.}
+      \EndIf
     \EndWhile \\

### 2025-08-14 GroupPolarizationStatmod/paper/ms.tex
edited intro analysis paragraph

--- 2025-08-14 GroupPolarizationStatmod/paper/ms.tex
@@
 
-\date{}
+\date{\today}
 \begin{document}
@@
 
-Our analysis estimates how the probability that a published detection of group
-polarization is actually a false detection. We first identified latent opinion
-parameters that yield spurious group polarization in an asymptotic $N\to\infty$
-model. This gives us our first suggestion that experimental conditions may be
-plausibly false. Briefly, we identified latent opinion distribution parameters
-that generated spurious group polarization in 54 out of 57 experimental
-conditions. In other words, 95\% of all group polarization detections are
-plausibly spurious due to ceiling artifacts from the continuous latent
-opinions-ordinal measurement experimental design. We then analyze the family-wise
-error ($\alpha$) and false detection rate (FDR) 
-for each experimental condition and aggregated 
-up to the study level and across all studies and experimental conditions. 
-We estimate the lower twentieth percentile of experimental conditions 
-have \emph{at least} $\alpha \leq 0.16$, 
-with median $\alpha \geq 0.27$ and upper eightieth percentile with $\alpha \geq
-0.40$ (Table~\ref{tab:quantiles}\mt{A.  –but don't have A and B; TODO: separate
-FWER and FDR tables, only use 0, 20, 50, 80, 100 for probs. Rename 0 -> min, 100 ->
-max.}). For the fiftieth and eightieth percentiles this
-corresponds to false discovery rates of 0.75 and 0.82, respectively. 
-We emphasize that these estimates and others detailed below are lower bounds on
-$\alpha$ and FDR since we are using what is likely an overgeneralized,
-underspecified statistical model that does not account for all the potential
-variance in experimental outcomes, which would increase error rates.
-We close our Analysis by calculating which $d^*$ limits family-wise error rates to
-$\alpha \leq 0.05$. This is greater than $d^* = 0.8$, classically taken to 
-represent ``high'' significance, in all but six experimental conditions.
+We now present our determinations of whether published detections of group
+polarization are plausibly spurious and calculate the best-case false detection
+rate for each experimental design in the corpus. We found that 95\% of all
+detections of group polarization are plausibly false, 54 out of 57. We estimate the
+best-case median false detection rate to be 72\% for those 54 experimental
+conditions. All but seven of the 54 designs were found to require a $d^* > 0.8$
+for ``significant'' Cohen's $d$ to limit the family-wise error rate to 5\%.
+
+% We first identified latent opinion
+% parameters that yield spurious group polarization in an asymptotic $N\to\infty$
+% model. This gives us our first suggestion that experimental conditions may be
+% plausibly false. Briefly, we identified latent opinion distribution parameters
+% that generated spurious group polarization in 54 out of 57 experimental
+% conditions. In other words, 95\% of all group polarization detections are
+% plausibly spurious due to ceiling artifacts from the continuous latent
+% opinions-ordinal measurement experimental design. We then analyze the family-wise
+% error ($\alpha$) and false detection rate (FDR) 
+% for each experimental condition and aggregated 
+% up to the study level and across all studies and experimental conditions. 
+% We estimate the lower twentieth percentile of experimental conditions 
+% have \emph{at least} $\alpha \leq 0.16$, 
+% with median $\alpha \geq 0.27$ and upper eightieth percentile with $\alpha \geq
+% 0.40$ (Table~\ref{tab:quantiles}\mt{A.  –but don't have A and B; TODO: separate
+% FWER and FDR tables, only use 0, 20, 50, 80, 100 for probs. Rename 0 -> min, 100 ->
+% max.}). For the fiftieth and eightieth percentiles this
+% corresponds to false discovery rates of 0.75 and 0.82, respectively. 
+% We emphasize that these estimates and others detailed below are lower bounds on
+% $\alpha$ and FDR since we are using what is likely an overgeneralized,
+% underspecified statistical model that does not account for all the potential
+% variance in experimental outcomes, which would increase error rates.
+% We close our Analysis by calculating which $d^*$ limits family-wise error rates to
+% $\alpha \leq 0.05$. This is greater than $d^* = 0.8$, classically taken to 
+% represent ``high'' significance, in all but six experimental conditions.
 

### 2025-08-15 GroupPolarizationStatmod/paper/intro-latest.tex
unexpectedly tweaking Intro

@@
 % \item
+
 We developed a generative model of opinion change to identify, if possible,
-null opinion change models that could spuriously appear like group polarization
-when naively measured using a Likert scale without accounting for the measurement procedure. 
-We further applied our method to estimate the family-wise error rate of
-detections of group polarization in a corpus of ten influential papers from 
-across several decades and from various theoretical traditions. We turn this criticism of past
-blunders constructive by providing a procedure to calculate what Cohen's $d$ should be
+\emph{simple consensus} latent distributions that could spuriously appear like group polarization.
+This type of problem is known to happen when ordinal measurements are taken of continuous data
+without statistically accounting for the measurement procedure itself~\cite{Liddell2018}.
+We further applied our method to estimate a best-case probability that 
+a published detection of group polarization is false, across 57 experimental
+conditions in a corpus of ten journal articles curated for their prominence and
+representativeness of group polarization research. 
+We turn this criticism of past blunders constructive by providing a procedure to calculate what Cohen's $d$ should be
 used to guarantee a family-wise error rate (i.e., Type I error rate) of 5\% or less.
@@
 
-\subsection{Group polarization research}
+\subsection{Group polarization}
 
 (DO THE MINIMUM AFTER EVERYTHING ELSE)
+
+- Group polarization is a type of \emph{consensus}. \emph\emph{Simple consesus} is consensus where
+\emph{only} the opinion variance of a group changes, not the mean opinion of the group.
+
+\subsubsection{Group polarization research}
 
@@
 
-\subsection{How study design can induce false detections of group polarization}
+\subsubsection{Failure to account for the measurement process can induce false detections}
 \label{ssubsec:false-detections}

### 2025-08-15 GroupPolarizationStatmod/paper/ms.tex
unexpectedly tweaking Intro

@@
 $g_{pub}$ is the published group polarization shift value.
+This could be thought of as sampling the observed ordinal distributions 
+$N \to \infty$, collecting ordinal measurements of samples from 
+each latent distribution corresponding to each phase, pre- and post-deliberation.
 
@@
 
-\subsection{Web app controls algorithm to find parameters inducing spurious group polarization}
+\subsection{Interactive algorithm to find parameters inducing spurious group polarization}
 
-To calculate the false discovery rate for each experimental condition, we first must
-identify a constant latent mean, $\mu$, and latent variances $\sigmapre$ and
-$\sigmapost$ that generate observation distributions\mt{define earlier?} with
-reported mean pre- and post-deliberation opinions and resulting shift $s$ (calculated
-using Equation~\ref{eq:group-polarization}). Define the collection of these
-parameters $\lambda = \{\mu, \sigmapre, \sigmapost\}$. We identify $\lambda$ through a
-hillclimbing optimization algorithm that finds $\lambda$ that minimizes the difference
-between reported pre- and post-deliberation mean opinions and simulated ones
+To identify plausibly false detections of group polarization, we searched numerically
+for latent parameters $\lambda = (\mu, \sigmapre, and \sigmapost)$ that generate
+spurious group polarization, as explained immediately above. We developed an *ad hoc*
+hillclimbing-style algorithm that finds $\lambda$ that finds $\lambda$ by minimizing
+the difference between published group polarization shifts $g_pub$ and calculated shifts,
+$g_\lambda$.
 calculated with Equation~\ref{eq:meanobs}.
@@
 
-% We first identified latent opinion
-% parameters that yield spurious group polarization in an asymptotic $N\to\infty$
-% model. This gives us our first suggestion that experimental conditions may be
-% plausibly false. Briefly, we identified latent opinion distribution parameters
-% that generated spurious group polarization in 54 out of 57 experimental
-% conditions. In other words, 95\% of all group polarization detections are
-% plausibly spurious due to ceiling artifacts from the continuous latent
-% opinions-ordinal measurement experimental design. We then analyze the family-wise
-% error ($\alpha$) and false detection rate (FDR) 
-% for each experimental condition and aggregated 
-% up to the study level and across all studies and experimental conditions. 
-% We estimate the lower twentieth percentile of experimental conditions 
-% have \emph{at least} $\alpha \leq 0.16$, 
-% with median $\alpha \geq 0.27$ and upper eightieth percentile with $\alpha \geq
-% 0.40$ (Table~\ref{tab:quantiles}\mt{A.  –but don't have A and B; TODO: separate
-% FWER and FDR tables, only use 0, 20, 50, 80, 100 for probs. Rename 0 -> min, 100 ->
-% max.}). For the fiftieth and eightieth percentiles this
-% corresponds to false discovery rates of 0.75 and 0.82, respectively. 
-% We emphasize that these estimates and others detailed below are lower bounds on
-% $\alpha$ and FDR since we are using what is likely an overgeneralized,
-% underspecified statistical model that does not account for all the potential
-% variance in experimental outcomes, which would increase error rates.
-% We close our Analysis by calculating which $d^*$ limits family-wise error rates to
-% $\alpha \leq 0.05$. This is greater than $d^* = 0.8$, classically taken to 
-% represent ``high'' significance, in all but six experimental conditions.
-
 \subsection{Nearly all group polarization detections are plausibly spurious}
 
-We used our formal distribution model of group polarization experiments
-to identify plausible latent simple consensus distributions that generate spurious
-group polarization in observed ordinal distributions 
-in the asymptotic case of taking $N \to \infty$ ordinal measurements.
-Across the ten studies we analyzed, we excluded 3 out of 60 experimental
-conditions that did not claim to be positive findings of group polarization.
+% We used our formal distribution model of group polarization experiments
+% to identify plausible latent simple consensus distributions that generate spurious
+% group polarization in 
 Of the remaining 57, we identified simple consensus $\lambda$ that generate spurious
@@
 identify $\lambda$ that generate spurious group polarization, but the initial
-simulated distributions were highly polarized. We rejected $\lambda$ generated for
+simulated distributions were highly polarized. 
+We rejected $\lambda$ generated for
 two spurious group polarization from Schkade, et al., (2010) because
@@
 did not provide details beyond mean values and opinion shifts, and the article was
-published before open data practices.
+published before open data practices. Our corpus had a total of 60 experimental conditions, but we excluded 3 out of 60 
+that did not claim to be positive findings of group polarization.
+
+- Address sigmas, note differences
+
+\begin{figure}
+  \centering
+    \includegraphics[width=\textwidth]{Figures/3.1/latent-sd-cdf.png}
+    \caption{\textbf{}}
+  \label{fig:}
+\end{figure}
 

### 2025-08-17 GroupPolarizationStatmod/paper/intro-latest.tex
expanded discussion and intro

@@
 
-If an extremist's opinion falls and it is measured with a Likert scale,
-will it make a noise? The answer had better be yes if we want to
-properly measure opinion change. Accurate measurements are critical for
-developing rigorous and reliable methods to promote more stable,
-sustainable, responsive, and efficient governments and institutions
-(Mason 2018; Klein 2020). Unfortunately, social and behavioral
-scientists have often used statistical methods that distort opinion
-change when it's measured on an ordinal, Likert-style scale (e.g.~-3
-indicates ``Strongly disagree'', +3 indicates ``Strongly agree'', and 0
-is neutral) (Liddell and Kruschke 2018). This includes a phenomenon
-called \emph{group polarization} that former Obama White House official
-and law professor Cass Sunstein relied on to explain why ``people become
-extremists'' and why ``political and cultural polarization'' is ``so
-pervasive in America'', as the publisher summarizes (Sunstein 2009).
-Sunstein suggests group polarization research provides a ``clue'' to
-explaining how ``facism\ldots student radicalism\ldots Islamic
-terrorism\ldots{[}t{]}he Rwandan genocide\ldots{[}e{]}thnic
-Conflict\ldots acts of torture and humiliation by American soldiers at
-Abu Ghraib'', and more (p.~1).
+\begin{quote}
+  The good thing about science is that it's true, whether or not you believe
+  it (Neil de Grasse Tyson t-shirt, 2010's).
+\end{quote}
 
-% \begin{itemize}
-% \tightlist
-% \item
+If an extremist's opinion falls and it is measured with a Likert scale, will
+it make a noise? The answer had better be yes if we want to properly measure
+opinion change. Accurate measurements of opinion change when tracking
+opinions and beliefs over time in surveys and experiments is critical to
+design more, sustainable, responsive, and efficient governments,
+institutions, and policies (Mason 2018; Klein 2020). One important social
+psychology paradigm for explaining opinion change is called \emph{group
+polarization}.  Former Obama White House administrator of Information and
+Regulatory affairs, and current Harvard Law School professor, Cass Sunstein
+writes that the scientific ``law of group polarization''~\cite{Sunstein1999}
+provides a ``clue'' to explaining how ``facism\ldots student
+radicalism\ldots Islamic terrorism\ldots{[}t{]}he Rwandan
+genocide\ldots{[}e{]}thnic Conflict\ldots acts of torture and humiliation by
+American soldiers at Abu Ghraib'', and more \cite[p. 1]{Sunstein2009}.
+
+We report here that, unfortunately, many studies of group polarization 
+used an ordinal, Likert-style, measurement scale but failed to account 
+for this fact.  
+Naive measurement procedures like this are known to induce
+both false positive and negative detections of changes in opinion, 
+belief, or similar \emph{latent} psychological entities, i.e., unobservable
+theoretical variables~\cite{Liddell2018,Turner2020}.
+Ordinal scales are ordered, integer-valued responses that represent both
+degree and polarity of opinion, e.g., ~-3
+indicates ``Strongly disagree'', +3 indicates ``Strongly agree'', 0
+is neutral, with intermediate numbers indicating different degrees of
+disagreement or agreement, depending on the sign.
+
+To answer the question posed above in the paper's lead sentence, it is
+entirely possible that ordinal scales will not detect all changes in extreme
+opinions. For example, consider the 7-point Likert scale described above. 
+If an extremist's latent opinion changes
+from -5 to -4, but they map their latent opinion onto the closest integer
+option, they will report -3 each time: such opinion shifts towards
+moderation would not be detected. On the other hand, a latent opinion shift
+from -1 to -2 would be detected. 
+Considering a group composed of two individuals whose opinions changed like 
+that, their mean latent opinion did not change: it is -3 in both the pre-
+and post-deliberation phases. We call this \emph{simple consensus}, where
+individuals' opinions agreed more following deliberation, but the mean
+opinion stayed constant. In statistical terms, the \emph{mean} opinion
+stayed the same and the \emph{variance} decreased. 
+The 7-point Likert scale, however, spuriously measures a change in 
+mean opinion from $\frac{1}{2}(-3 + -1) = -2$ to $\frac{1}{2}(-3 + -2) =
+-2.5$. In other words, we have a false detection of group polarization where
+the mean opinion seems to shift from -2 to -2.5.  The same logic applies to
+groups with more than two people.
 
 We developed a generative model of opinion change to identify, if possible,
-\emph{simple consensus} latent distributions that could spuriously appear like group polarization.
-This type of problem is known to happen when ordinal measurements are taken of continuous data
-without statistically accounting for the measurement procedure itself~\cite{Liddell2018}.
-We further applied our method to estimate a best-case probability that 
-a published detection of group polarization is false, across 57 experimental
-conditions in a corpus of ten journal articles curated for their prominence and
-representativeness of group polarization research. 
-We turn this criticism of past blunders constructive by providing a procedure to calculate what Cohen's $d$ should be
-used to guarantee a family-wise error rate (i.e., Type I error rate) of 5\% or less.
+two opinion distributions representing \emph{simple consensus} that seem to
+be \emph{group polarization} when transformed to an ordinal measurement
+scale.  latent distributions that could spuriously appear like group
+polarization.  This type of problem is known to happen when ordinal
+measurements are taken of continuous data without statistically accounting
+for the measurement procedure itself~\cite{Liddell2018}.  We further applied
+our method to estimate a best-case probability that a published detection of
+group polarization is false, across 57 experimental conditions in a corpus
+of ten journal articles curated for their prominence and representativeness
+of group polarization research.  We turn this criticism of past blunders
+constructive by providing a procedure to calculate what Cohen's $d$ should
+be used to guarantee a family-wise error rate (i.e., Type I error rate) of
+5\% or less.
 
 We calculate the severity of the problem in two steps. In the first step we
-provide proof by example that there exists a null opinion change model in terms of \emph{latent, continuous}
-psychological opinions that looks like group polarization when measured via participant behavior, i.e.,
-by their response on an \emph{ordinal survey scale}. If such a null model is found, this means that we cannot
-rule out a null effect, given the published data. This means, therefore, that the published detection is
-just as likely false as it is true, so we say it is a \emph{plausibly false detection of group polarization},
-or a \emph{plausibly spurious detection} for short.
+provide proof by example that there exists a null opinion change model in
+terms of \emph{latent, continuous} psychological opinions that looks like
+group polarization when measured via participant behavior, i.e., by their
+response on an \emph{ordinal survey scale}. If such a null model is found,
+this means that we cannot rule out a null effect, given the published data.
+This means, therefore, that the published detection is just as likely false
+as it is true, so we say it is a \emph{plausibly false detection of group
+polarization}, or a \emph{plausibly spurious detection} for short.
 
-In the second step of the analysis of whether, we simulate observations
-of group polarization experiments where participant opinions are drawn from the 
-null model distributions identified in the first step,
-then these are measured on a Likert scale for simulated group polarization survey
-responses on an ordinal scale. We then calculate the
-family-wise error rate (i.e., Type I error rate) and false detection rate using
-various effect sizes to determine significance of the effect in terms of Cohen's $d$.
-The family-wise error rate is defined as the probability that a known null 
-effect is detected as significant. We use this to then estimate the *false detection rate*, which is the
-probability that a detection of an effect is a false one, which depends on the relative prevalence of
-true and false positive detections in the literature.
-Finally, we put our model to constructive use to show what Cohen's $d$ must be set to
-to limit the family-wise error rate to 5\%, denoted $d*$. This will not help rescue any of
-the published results here found to be plausibly spurious. But, it provides a measure of the weakness of
-the published experimental designs independent of their shared fatal failure to 
-account for the measurement procedure: greater $d*$ indicates weaker designs.
+In the second step of the analysis of whether, we simulate observations of
+group polarization experiments where participant opinions are drawn from the
+null model distributions identified in the first step, then these are
+measured on a Likert scale for simulated group polarization survey responses
+on an ordinal scale. We then calculate the family-wise error rate (i.e.,
+Type I error rate) and false detection rate using various effect sizes to
+determine significance of the effect in terms of Cohen's $d$.  The
+family-wise error rate is defined as the probability that a known null
+effect is detected as significant. We use this to then estimate the *false
+detection rate*, which is the probability that a detection of an effect is a
+false one, which depends on the relative prevalence of true and false
+positive detections in the literature.  Finally, we put our model to
+constructive use to show what Cohen's $d$ must be set to to limit the
+family-wise error rate to 5\%, denoted $d*$. This will not help rescue any
+of the published results here found to be plausibly spurious. But, it
+provides a measure of the weakness of the published experimental designs
+independent of their shared fatal failure to account for the measurement
+procedure: greater $d*$ indicates weaker designs.
 
-We find that 95\% (54 of 57) of group polarization detections are plausibly spurious.
-In other words, empirical support for group polarization has been seriously undermined
-just by removing these results from the count of group polarization replications.
-There is no reason to expect other papers to be more reliable. (Summarize other
-two findings, 1. FWER/FDR and 2. $d*$ for FWER < 5\%)
+We find that 95\% (54 of 57) of group polarization detections are plausibly
+spurious.  In other words, empirical support for group polarization has been
+seriously undermined just by removing these results from the count of group
+polarization replications.  There is no reason to expect other papers to be
+more reliable. (Summarize other two findings, 1. FWER/FDR and 2. $d*$ for
+FWER < 5\%)
 
-Our methods can be turned on any data-driven social science analysis of
-data where participants report a latent, subjective psychological
-variable (mood, physiological state, etc.), and where various theories
-supposedly battle it out to best represent the phenomenon of interest
-via null-hypothesis significance testing using ordinal data as continuous. 
-This may just be the first
-of many bricks to fall since \emph{all 68} of the articles that
-mentioned the word ``Likert'' and analyzed ordinal data used metric
-models to detect latent psychological variable change across three
-influential psychology journals, \emph{Psychological Science}, the
-\emph{Journal of Experimental Psychology: General}, and the
-\emph{Journal of Personality and Social Psychology} (Liddell and
-Kruschke 2018). This further confirms that replication is far from the only
-foundational crisis that behavioral science faces\cite{Yarkoni2022,Turner2022}.
+Group polarization is not the only phenomenon in the social and behavioral
+sciences to naively ignore measurement procedures.  Liddell and Kruschke
+found that the measurment procedures of experimental designs are rarely, if
+ever, accounted for by social psychology researchers in their systematic
+review of select social psychology journals. Our work is important, then,
+both to develop rigorous group polarization studies and provide a template
+for re-analyzing pre-crisis social and behavioral research.  Our methods can
+be turned on any data-driven social science analysis of data where
+participants report a latent, subjective psychological variable (mood,
+physiological state, etc.), and where various theories supposedly battle it
+out to best represent the phenomenon of interest via null-hypothesis
+significance testing using ordinal data as continuous.  This may just be the
+first of many bricks to fall since \emph{all 68} of the articles that
+mentioned the word ``Likert'' and analyzed ordinal data used metric models
+to detect latent psychological variable change across three influential
+psychology journals, \emph{Psychological Science}, the \emph{Journal of
+Experimental Psychology: General}, and the \emph{Journal of Personality and
+Social Psychology} (Liddell and Kruschke 2018). This further confirms that
+replication is far from the only foundational crisis that behavioral science
+faces\cite{Yarkoni2022,Turner2022}.
 
-Replication is not enough:
-strong, mechanistic models must be developed to be able to predict Type I and II
-error rates \emph{ahead of time}, as we demonstrate here. We only demonstrate how to
-calculate expected Type I error rates for a measurement design, i.e., the rate of false 
-positives. Low statistical power exacerbates false detection rates. So, to be confident in replication,
-a pre-registered measurement design should use this style of simulation to 
-estimate the statistical power. Tangled, weak, informal
-verbal models of complex psychological and social phenomena like those that motivated
-empirical studies of group polarization tend to be vulnerable
-to stastical weaknesses. It is difficult to incorporate prior knowledge such
-as experiment structure without a sufficiently rigorous and appropriate 
-formal system \cite{Smaldino2017,Smaldino2020,Flake2020,Deffner2024}.
+\subsection{Group polarization research and corpus}
 
-\subsection{Group polarization}
+To situate this current paper it will help to introduce some relevant
+research traditions and common experimental designs in group polarization,
+using the ten studies included in our corpus as touchstones. To begin,
+*group polarization* is defined as...(continue basic definition with some
+examples from our corpus).
 
-(DO THE MINIMUM AFTER EVERYTHING ELSE)
+Paragraph dos: Common experimental, measurement, and statistical procedures
+to induce and detect group polarization
 
-- Group polarization is a type of \emph{consensus}. \emph\emph{Simple consesus} is consensus where
-\emph{only} the opinion variance of a group changes, not the mean opinion of the group.
+\subsection{Failure to account for the measurement process can induce
+false detections} \label{ssubsec:false-detections}
 
-\subsubsection{Group polarization research}
-
-To situate this current paper it will help to introduce some relevant research
-traditions and common experimental designs in group polarization, using the ten
-studies included in our corpus as touchstones. To begin, *group polarization*
-is defined as...(continue basic definition with some examples from our corpus).
-
-Paragraph dos: 
-Common experimental, measurement, and statistical procedures to induce and
-detect group polarization
-
-\subsubsection{Failure to account for the measurement process can induce false detections}
-\label{ssubsec:false-detections}
-
-An opinion is never measured directly, only a behavior that indicates an opinion
-in a behavioral experiment or survey.  For example, the location of a mark
-that a participant makes is measured, not the complex neural activity that
-underlies forming opinions.
-Theoretically, an internal opinion in the form of neural activity is a
-\textit{latent opinion}, from the Latin \emph{latentem} meaning ``lie
-hidden''.  Participants report \text{observed opinions}, typically as one of
-a finite number of ranked, successive integer values that form the
-\emph{selection set}. The minimum possible value in the selection set might
-mean ``strongly disagree'' and the maximum value represents ``strongly
-agree''.
+To see how failing to account for the measurement procedure could be a
+problem, let's step back to sketch a \emph{model} of group polarization
+experimental design.  This just means we are going to pick some key
+components and processes that we assume are important together. This does
+not mean other factors are not important.  We do this to focus on one
+complex system by \emph{screening out} interactions with several other
+complex systems. This piecemeal process may seem slow, but it has proven
+useful in more mature sciences like physics and biology.
 
 Group polarization experiments share a common experimental design where,
-first, participant opinions are gathered on a topic of interest. Next, groups
-are formed that have some initial opinion extremism. Participants deliberate
-or discuss the topic for some set time. After the set time, participants
-again give their opinion. Group polarization is said to occur if the
-mean group opinion became more extreme, i.e., if the magnitude of the mean
-group opinion is greater \emph{post-deliberation} compared to
-\emph{pre-deliberation}. For example,
-\cite{Moscovici1969} asked Parisian high school students to indicate their
-degree of agreement with the statement, ``American economic aid is always
-used for political pressure''. There
-were seven options in the selection set, -3 indicating ``strongly disagree'',
-+3 indicating ``strongly agree'', and 0 was neutral. Moscovici and Zavalloni
-report that the mean pre-deliberation opinion was -0.61 and the mean
-post-deliberation opinion was -1.09 (Moscovici and Zavalloni, 1969; Table 4,
-p. 132). This is a -0.48 shift to more extreme disagreement with the 
-statement, reported by Moscovici and Zavalloni as statistically significant
-in Table 5 on p. 132.
+first, participant opinions are gathered on a topic of interest. Next,
+groups are formed that have some initial opinion extremism. Participants
+deliberate or discuss the topic for some set time. After the set time,
+participants again give their opinion. Group polarization is said to occur
+if the mean group opinion became more extreme, i.e., if the magnitude of the
+mean group opinion is greater \emph{post-deliberation} compared to
+\emph{pre-deliberation}. For example, \cite{Moscovici1969} asked Parisian
+high school students to indicate their degree of agreement with the
+statement, ``American economic aid is always used for political pressure''.
+There were seven options in the selection set, -3 indicating ``strongly
+disagree'', +3 indicating ``strongly agree'', and 0 was neutral. Moscovici
+and Zavalloni report that the mean pre-deliberation opinion was -0.61 and
+the mean post-deliberation opinion was -1.09 (Moscovici and Zavalloni, 1969;
+Table 4, p. 132). This is a -0.48 shift to more extreme disagreement with
+the statement, reported by Moscovici and Zavalloni as statistically
+significant in Table 5 on p. 132. Sometimes group polarization research uses
+significance testing for determining group polarization, but Moscovici and
+Zavalloni, and others, just report a bare difference in mean opinion. This
+alone should make modern researchers think twice before including it and
+others as canon, but apparently this has gone unnoticed.
 
-(PREVIOUS TWO PARAGRAPHS SET UP THIS FINAL PARAGRAPH: CONTINUE TO USE MOSCOVICI 1969 TO
-EXPLAIN IN THE SIMPLEST TERMS HOW FAILING TO ACCOUNT FOR MEASUREMENT PROCEDURE
-CAN LEAD TO FALSE POSITIVES; see Figure \ref{fig:distros})
+Note that only \emph{behaviors} are measured in this experiment, and not
+\emph{opinions}.  Instead, the participant (paid by the job, not the hour)
+navigates their mouse to hover over a radio button in an online survey, or
+voices their opinion by phone to a survey caller. The \emph{latent}, complex
+neural activity that underlies opinion formation and expression via
+different behaviors is unseen and, so far, unknown. We call internal
+opinions \textit{latent opinions}, from the Latin \emph{latentem} meaning
+``lie hidden''.  In other words, we cannot measure neurobiological activity
+and infer a participant's opinion.  We only ever measure participant
+behavior that signals the participant's opinion. Note this presupposes that
+the participant's opinion is that behavior best matches their opinion on the
+deliberation topic.
+
+The problem arises when we try to compare average \emph{measured} opinion
+between phases, when instead the \emph{latent} opinion is the one that would
+prove group polarization occurred.  In fact, the problem seems to be that
+they are assumed to be identical. However, group polarization of latent
+opinions, or other psycho-social dynamics, are known to get distorted when
+measured with an ordinal scale~\cite{Liddell2018,Turner2020}. Recall that
+theoretically group polarization is one type of consensus, where extremism
+and agreement both increase, meaning the magnitude of the mean opinion in
+the group increases while the variance decreases. We distinguish group
+polarization from what we call \emph{simple consensus}
+
+of If we represent the latent distribution
+
+(PREVIOUS TWO PARAGRAPHS SET UP THIS FINAL PARAGRAPH: CONTINUE TO USE
+MOSCOVICI 1969 TO EXPLAIN IN THE SIMPLEST TERMS HOW FAILING TO ACCOUNT FOR
+MEASUREMENT PROCEDURE CAN LEAD TO FALSE POSITIVES; see Figure
+\ref{fig:distros})
 
@@
 
-In the next section, we introduce the various formal and computational Methods 
-used to obtain our findings outlined above. We then explain our findings in detail in the
-Analysis section. We close with a Discussion of how this is not a defeat, but
-a new beginning for group polariation, given that there are good reasons to expect
-group polarization to occur. We also explain how to expanding this analysis to
-examine other published studies on group polarization and across social and behavioral
-science disciplines.
+In the next section, we introduce the various formal and computational
+Methods used to obtain our findings outlined above. We then explain our
+findings in detail in the Analysis section. We close with a Discussion of
+how this is not a defeat, but a new beginning for group polariation, given
+that there are good reasons to expect group polarization to occur. We also
+explain how to expanding this analysis to examine other published studies on
+group polarization and across social and behavioral science disciplines.

### 2025-08-17 GroupPolarizationStatmod/paper/ms.tex
expanded discussion and intro

@@
 Since there seems to be no selection pressure on false discovery rates,
-experiment design for studying group polarization was free to evolve randomly.
-Various choices for the number of bins and other instrumental details may have
-then become entrenched in different communities defined by 
+experiment design for studying group polarization was free to evolve
+randomly. Various choices for the number of bins and other instrumental
+details may have then become entrenched in different communities defined by
 common theoretical perspectives taken by researchers in those communities.
-Therefore, when 
-Isenberg (1986)~\cite{Isenberg1986} reports that his meta-analysis found more
-support for the XX theory than the YY theory in the form of greater effect sizes,
-it may just be random cultural drift that the XX community instruments had higher
-intrinsic false detection rates. In some cases it may be debatable whether to
-control for the Type I error rate, or may depend on context~\cite{Rubin2021}. 
-In this context, however, it is clear that we must apply such selection pressure 
-to group polarization experiments or it may be nearly impossible to trust 
-statistical analyses of these studies.
+Therefore, when Isenberg (1986)~\cite{Isenberg1986} reports that his
+meta-analysis found more support for the persuasive arguments theory than the 
+social comparisons theory in the form of greater effect sizes. If this were
+indeed the case, perhaps it is only because some researchers hypothesizing
+persuasive arguments happened to use a measurement procedure produced more
+frequent and extreme distortions of opinion measurements. Incentive
+structures known to promote the ``evolution of bad science'' would select
+for research that ignored sticky, time-consuming statistical problems in
+favor of juicy titles and headlines for which practitioners have won several
+professional awards. Specifically, the traditional bias
+towards positive results, and against negative ones, could easily have
+provided a cultural niche in which quick but faulty group polarization
+statistical methods dominate slower, more rigorous ones.
+Now that we are aware of the problem, there must be
+intense selective pressure applied as soon and as widely as possible to 
+avoid further ``replications'' that report spurious group polarization, or
+any similar latent psychological dynamics.
 
-This leads to one of many follow-up research questions generated by this work.
-First, are there general principles of experimental design that could reduce
-the false detection rate for group polarization or similar experimental protocols?
-Second, what is the best way to calculate false detection rate? 
-Likely, even if we cannot identify parameters consistent with simple consensus
-that generate spurious group polarization shifts in the large $N$ case, one
-may still guess at some simple consensus parameters and caluclate the false
-detection rate as we did. On the flip side, we likely found only one of infinitely
-many simple consensus parameters that could result in a false detection. Does this
-have an effect on the calculated false positive rate? More generally, our
-work motivates similar analyses of sub-fields of social 
-psychology and that do not account for Likert-based experimental designs in
-their statistical analyses. We provided a template for this future work.
+We provided tools that experimenters can use. While these tools are
+potentially useful, they are still in the prototype phase that was
+sufficient for a single researcher who was also the software developer. So,
+a first step towards expanding the usefulness of the findings here is to 
+improve the software usability and make the app widely available for
+researchers in group polarization and beyond. Another step would be to go
+beyond reactionary testing of the reliability of experimental designs, and
+identify design principles that reduce the false detection rate without
+needing to adjust the Cohen's $d$ (or whatever measure) 
+used as a significance threshold.
 
-Our analysis brings some additional questions into focus. Most significantly, recall that even
-the more appropriate ordered probit Bayesian model often failed to accurately detect negative
-results with the number of participants used in the original papers. Many of the studies
-analyzed here did not use any statistical test to establish significance; those that did used
-some form of a $t$-test, and it is not clear if the $t$-test that was used 
-accounted for a change in pre- and post-deliberation opinion variance, which we expect due to
-consensus.
+Our analysis brings even more additional questions into focus. Most
+significantly, recall that even the more appropriate ordered probit Bayesian
+model often failed to accurately detect negative results with the number of
+participants used in the original papers. Many of the studies analyzed here
+did not use any statistical test to establish significance; those that did
+used some form of a $t$-test, and it is not clear if the $t$-test that was
+used accounted for a change in pre- and post-deliberation opinion variance,
+which we expect due to consensus.
 
-The ordered probit model we used still likely underestimates the variance in outcomes, which
-could require even larger sample sizes. A more rigorous analysis
-would include additional sources of variance that will likely further inflate the Type I
-error rate, $\alpha$, and the false discovery rate~\cite{Yarkoni2022}. First and foremost,
-there is variance within each experimental deliberation group; no group polarization study
-accounts for this, instead treating participants in each experimental condition as if they were
-one large group. Furthermore, survey data are known to be noisy~\cite{Zaller1992}. People
-report opinions differently over time for no apparent reason (i.e., opinions are
-\emph{unstable}).  Context matters: the order in which a survey question is asked, and which
-question framing is chosen among logically equivalent alternatives, can both significantly
-influence participant responses. The effect of accounting for these sources of variance must be
-understood to estimate $\alpha$ (and power, $W$) for the design of the next generation of group
-polarization experiments.
+The ordered probit model we used still likely underestimates the variance in
+outcomes, which could require even larger sample sizes. A more rigorous
+analysis would include additional sources of variance that will likely
+further inflate the Type I error rate, $\alpha$, and the false discovery
+rate~\cite{Yarkoni2022}. First and foremost, there is variance within each
+experimental deliberation group; no group polarization study accounts for
+this, instead treating participants in each experimental condition as if
+they were one large group. Furthermore, survey data are known to be
+noisy~\cite{Zaller1992}. People report opinions differently over time for no
+apparent reason (i.e., opinions are \emph{unstable}).  Context matters: the
+order in which a survey question is asked, and which question framing is
+chosen among logically equivalent alternatives, can both significantly
+influence participant responses. Finally, little is known about the
+psychological process that converts latent opinions to reporting behaviors
+(e.g., clicking a radio button corresponding to an opinion). 
+The effect of accounting for these sources of variance must be understood to
+estimate $\alpha$ (and power, $W$) for the design of the next generation of
+group polarization experiments.
 
-\mt{concluding paragraph: first sentences summarize the Discussion, then
-in 1-2 sentences summarize the paper.}
+To close, note that replication of weakly- and incommensurately-specified 
+social phenomena is not enough: strong, mechanistic models must be developed to
+be able to predict Type I and II error rates \emph{ahead of time}, as we
+demonstrate here. We only demonstrate how to calculate expected Type I error
+rates for a measurement design, i.e., the rate of false positives. Low
+statistical power exacerbates false detection rates. So, to be confident in
+replication, a pre-registered measurement design should use this style of
+simulation to estimate the statistical power. It is difficult to incorporate
+prior knowledge such as experiment structure without a sufficiently rigorous
+and appropriate formal system
+\cite{Smaldino2017,Smaldino2020,Flake2020,Deffner2024}.
 

### 2025-10-23 GroupPolarizationStatmod/paper/intro-later.tex
tightening intro

--- 2025-08-17 GroupPolarizationStatmod/paper/intro-latest.tex
@@
 
-\emph{Group polarization} is the name for a form of consensus where
-members of a like-minded group become, on average, more extreme in their
-opinions after discussing a topic. Group polarization is important
-because it may increase social and political tensions if moderates
-become more extreme. Decades of studies have replicated detections of
-group polarization, typically using an ordinal scale to measure
-opinions. However, these studies did not account for ordinal
-measurements, which can result in spurious detections of opinion change.
-Lacking original data we can still calculate the probability that a
-published detection of group polarization was spurious, i.e., to
-calculate the \emph{false detection rate}. Across 54 group polarization
-experimental conditions from ten representative journal articles, we
-calculated false detection rates to be between 0.52 and 0.88, with a
-median of 0.75, using a generative model of group polarization
-experiments seeded with empirical data. We also use our model to develop
-experimental designs that achieve an acceptable false discovery rate.
-Much of group polarization research may be unreliable. This analysis can
-help change that by enabling others to avoid this hidden pitfall. More
-broadly this work demonstrates one important way that replication
-success alone does not imply epistemic reliability.
+Echo chambers seem to cause radicalization—and radicalization is risky.
+Radicalization undermines institutions by reducing their diversity, in turn impairing their \emph{adaptability}.
+Social psychologists and others study echo chambers experimentally using \emph{group polarization} experiments.
+Researchers induce \emph{group polarization} by first asking experiment participants their opinions on some topic, then grouping them together if they agree in spirit, but not degree.
+The participants then discuss the topic in their groups.
+Their opinions are then re-measured.
+Group polarization is when the average group opinion increases.
+But there’s a fatal problem: participants report opinions on ordinal scales, but this has never been accounted for statistically.
+This can cause apparent group polarization in observations even if the average latent opinion—the one in people’s heads—doesn’t change, due to ceiling effects.
+Here we prove that across ten journal articles spanning five decades that 57 of 60 group polarization detections are plausibly false due to this effect—therefore they are not true “replications”.
+Our method can be used to evaluate any experimental design that measures latent change with ordinal scales.
+Given how destructive radicalization can be, we must get this right.
 
@@
 
-\begin{quote}
-In our introductory social psychology course, we have for many years
-used the {[}group polarization experimental paradigm{]} as a laboratory
-exercise. The exercise works beautifully, but one must be careful to
-forewarn a class that {[}group polarization{]} does not occur with every
-group\ldots and that the effect is not large (Brown 1986).
-\end{quote}
+If an extremist's opinion falls and it is measured with a Likert scale, will
+it make a noise? The answer is often "no"~\cite{Liddell2018,Turner2020}.
+This means that a group's opinions can appear to radicalize over time even
+when no real shift occurs. The problem matters because Likert and other
+ordinal scales are ubiquitous in research on opinion change and
+radicalization across the social and cognitive sciences~\cite{Liddell2018}.
+One influential theory in this area is \emph{group polarization}—the idea
+that discussion among like-minded people pushes opinions toward extremes,
+producing ``echo chambers.'' Cass Sunstein, Harvard Law professor
+and former White House administrator, has written extensively
+about group polarization~\cite{Sunstein2019}. He said group polarization
+research provides a ``clue'' to prevent ``radicalism'' and its bad effects
+including ``facism\ldots terrorism\ldots genocide\ldots'' and
+``torture''~\cite[p. 1]{Sunstein2009}.  Sunstein claims group polarization is
+a scientific \emph{law}, a moniker meaning a concrete, quantitative
+relationship between variables that will \emph{always be observed as long as
+a set of assumptions are met}~\cite{Cartwright1999}. There is reason to be
+suspicious first on the fact that Roger Brown admitted in his book
+\emph{Social Psychology}, group polarization ``does not occur with every
+group'' and ``the effect is not large.'' Given the scope and influence of
+group polarization, it is imperative to know whether the hundreds of
+``replications'' of group polarization might not really be so because of the
+potential they are just \emph{measurement artifacts} due to a mismatch
+between the
 
-\begin{quote}
-One of the most robust findings in social psychology is that of attitude
-polarization following discussion with like-minded others (Cooper,
-Kelly, and Weaver 2001).
-\end{quote}
+We are sad to report that
 
-\begin{quote}
-  The good thing about science is that it's true, whether or not you believe
-  it (Neil de Grasse Tyson t-shirt, 2010's).
-\end{quote}
+We analyzed sixty representative experimental conditions 
+from the group polarization literature to show that the answer is "no"
+for nearly all of them  \emph{group polarization}
+experiments—\emph{group polarization} is a scientific model of the process by
+which \emph{echo chambers} radicalize their members. By \emph{echo chamber} I
+mean an online or real-life community where likeminded people discuss 
+political or other topics of interest. Group polarization research, then,
+could be of great importance both to develop strategies to counteract 
+radicalization that undermines our institutions, and to increase support for
+sustainable practices. 
+Clearly group polarization research has found 
+Former Obama White House administrator of Information and
+Regulatory affairs, and current Harvard Law School professor, Cass Sunstein
 
-If an extremist's opinion falls and it is measured with a Likert scale, will
-it make a noise? The answer had better be yes if we want to properly measure
-opinion change. Accurate measurements of opinion change when tracking
-opinions and beliefs over time in surveys and experiments is critical to
-design more, sustainable, responsive, and efficient governments,
-institutions, and policies (Mason 2018; Klein 2020). One important social
-psychology paradigm for explaining opinion change is called \emph{group
-polarization}.  Former Obama White House administrator of Information and
-Regulatory affairs, and current Harvard Law School professor, Cass Sunstein
-writes that the scientific ``law of group polarization''~\cite{Sunstein1999}
-provides a ``clue'' to explaining how ``facism\ldots student
-radicalism\ldots Islamic terrorism\ldots{[}t{]}he Rwandan
-genocide\ldots{[}e{]}thnic Conflict\ldots acts of torture and humiliation by
-American soldiers at Abu Ghraib'', and more \cite[p. 1]{Sunstein2009}.
+We show the answer is no, if one uses the predominant experimental design
+adopted almost Liddell and Kruschke (2018) showed that
+this setup can lead to false positives because ordinal scales cannot
+track changes in extreme opinions beyond the ceiling of the measurement
+scale— In another example
+of this experimental design, Burnstein and Vinokur (1975) used a ten-point
+scale from 1 to 10 representing “the lowest odds of success acceptable” to
+try a certain behavior, where 1 means 10\%—Moscovici and Zavalloni (1969) gave participants a seven-point scale, from “strongly disagree” (-3) to “strongly agree” (+3) to report their opinions. 
+reported opinions on the ordinal scale—with discrete numbered bins—were
+treated as if they were continuous-valued, or in other words as if they were
+the internal mental opinion value itself, whatever that might be
+neurobiologically.  In this paper we examine whether group polarization
 
@@
 
-To situate this current paper it will help to introduce some relevant
-research traditions and common experimental designs in group polarization,
-using the ten studies included in our corpus as touchstones. To begin,
-*group polarization* is defined as...(continue basic definition with some
-examples from our corpus).
+Following Cartwright’s (1973) critique, social psychologists and others continued work on group polarization, focused on developing and comparing distinct theories of group polarization, with two becoming particularly influential: social comparisons and persuasive arguments.
+The social comparisons hypothesis says group polarization ultimately occurs due to the human desire to fit in, to hold opinions closer to those around them.
+Group polarization occurs, then, when a group is composed mostly of extremists, with a sufficient number of people with a weak opinion (Myers and Bishop, 1970; Myers, 1975).
+Its main “competitor” from the 1970s was the persuasive arguments theory, which basically claimed that rhetoric was the strongest driver of group polarization, not social comparisons (Burnstein and Vinokur, 1973; 1975).
 
-Paragraph dos: Common experimental, measurement, and statistical procedures
-to induce and detect group polarization
+To compare theories, researchers mainly relied on comparing effect sizes or p-values within the context of single idiosyncratic experiments—even after Cartwright implored researchers to develop a coherent theoretical foundation.
 
-\subsection{Failure to account for the measurement process can induce
+For example, Burnstein and Vinokur (1975, p.
+420) performed the following experiment and found the results supported their persuasive arguments theory.
+In one condition participants were only exposed “to the choices of others” but could not discuss—in another they discussed the survey topic with others, but did not know what numerical value they chose.
+In the final condition, participants could discuss and know each other’s opinions—and this is the only one with a significant p-value indicating group polarization reliably occurred.
+This alone could not confirm that opinion shifts were due to “persuasive arguments”, so Burnstein and Vinokur also counted how many arguments people made during the discussion sessions.
+They found that more arguments correlated with greater group polarization shifts to more extreme opinions, —confirmation that persuasive arguments beats social comparisons, according to them.
+About a decade later, Isenberg (1986) performed an aggregated “significance test” based on reported p-values, comparing the aggregate across papers claiming to prove the persuasive arguments hypothesis versus the social comparisons hypothesis.
+
+Additional hypotheses about what causes group polarization proliferated at the end of the last century, despite zero progress towards commensurability or even awareness of the problem.
+Other notable hypotheses include the self-categorization and social decisions scheme hypotheses.
+Self-categorization augmented the social comparisons hypothesis with a consideration of idiosyncratic beliefs people hold that are the result of their own personal lived experience (Abrams, et al., 1990; Hogg, et al., 1990).
+The social decisions hypothesis posits that contextual factors like social network structure could be at least as important as one’s opinions and the opinions of others .
+Noah Friedkin, a sociologist, notably published perhaps the most rigorous study on group polarization, adopting the social decisions scheme hypothesis, predicting that social connectivity within a group could amplify group polarization (Friedkin, 1999).
+
+\subsection{\mt{CUT??} Failure to account for the measurement process can induce
 false detections} \label{ssubsec:false-detections}
@@
 the statement, reported by Moscovici and Zavalloni as statistically
-significant in Table 5 on p. 132. Sometimes group polarization research uses
-significance testing for determining group polarization, but Moscovici and
-Zavalloni, and others, just report a bare difference in mean opinion. This
-alone should make modern researchers think twice before including it and
-others as canon, but apparently this has gone unnoticed.
+significant in Table 5 on p. 132.
 
@@
 
-of If we represent the latent distribution
-
-(PREVIOUS TWO PARAGRAPHS SET UP THIS FINAL PARAGRAPH: CONTINUE TO USE
-MOSCOVICI 1969 TO EXPLAIN IN THE SIMPLEST TERMS HOW FAILING TO ACCOUNT FOR
-MEASUREMENT PROCEDURE CAN LEAD TO FALSE POSITIVES; see Figure
-\ref{fig:distros})
-
 \subsection{Overview}

### 2025-10-23 GroupPolarizationStatmod/paper/intro-latest.tex
tightening intro

--- 2025-08-14 GroupPolarizationStatmod/paper/intro-latest.tex
@@
 
-\emph{Group polarization} is the name for a form of consensus where
-members of a like-minded group become, on average, more extreme in their
-opinions after discussing a topic. Group polarization is important
-because it may increase social and political tensions if moderates
-become more extreme. Decades of studies have replicated detections of
-group polarization, typically using an ordinal scale to measure
-opinions. However, these studies did not account for ordinal
-measurements, which can result in spurious detections of opinion change.
-Lacking original data we can still calculate the probability that a
-published detection of group polarization was spurious, i.e., to
-calculate the \emph{false detection rate}. Across 54 group polarization
-experimental conditions from ten representative journal articles, we
-calculated false detection rates to be between 0.52 and 0.88, with a
-median of 0.75, using a generative model of group polarization
-experiments seeded with empirical data. We also use our model to develop
-experimental designs that achieve an acceptable false discovery rate.
-Much of group polarization research may be unreliable. This analysis can
-help change that by enabling others to avoid this hidden pitfall. More
-broadly this work demonstrates one important way that replication
-success alone does not imply epistemic reliability.
+Echo chambers seem to cause radicalization—and radicalization is risky.
+Radicalization undermines institutions by reducing their diversity, 
+which in turn impairs their adaptability.  Social psychologists and others study
+echo chambers experimentally using \emph{group polarization} experiments.
+Researchers induce \emph{group polarization} by first asking experiment
+participants their opinions on some topic, then grouping them together if
+they agree in spirit, but not degree. The participants then discuss the topic
+in their groups. Their opinions are then re-measured. Group polarization is
+when the average group opinion increases. But there’s a fatal problem:
+participants report opinions on ordinal scales, but this has never been
+accounted for statistically. This can cause apparent group polarization in
+observations even if the average latent opinion—the one in people’s
+heads—doesn’t change, due to ceiling effects. Here we prove that across ten
+journal articles spanning five decades that 54 of 57 reported polarization effects are plausibly false due to this artifact, and thus cannot be counted as replications.
+Our method can be used to evaluate any experimental design that measures latent change with ordinal scales.
+Given how destructive radicalization can be, we must get this right.
 
@@
 
-\begin{quote}
-In our introductory social psychology course, we have for many years
-used the {[}group polarization experimental paradigm{]} as a laboratory
-exercise. The exercise works beautifully, but one must be careful to
-forewarn a class that {[}group polarization{]} does not occur with every
-group\ldots and that the effect is not large (Brown 1986).
-\end{quote}
+If an extremist’s opinion falls and it is measured with a Likert scale, will
+it make a noise? The answer is often no~\cite{Liddell2018,Turner2020}. This
+means a group’s opinions can appear to radicalize over time even when no real
+shift occurs. The problem matters because Likert and other ordinal scales are
+ubiquitous in research on opinion change and radicalization across the social
+and cognitive sciences~\cite{Liddell2018}. One influential framework in this
+area is group polarization—the idea that discussion among like-minded people
+pushes opinions toward extremes, producing ``echo chambers.'' Cass Sunstein, a
+Harvard Law professor and former White House administrator, has written
+extensively on the topic~\cite{Sunstein2019}, claiming that group
+polarization is a scientific law—a regular, quantitative relationship that
+will always be observed when certain assumptions are
+met~\cite{Cartwright1999}. Yet Roger Brown himself noted in Social Psychology
+that group polarization ``does not occur with every group'' and that ``the
+effect is not large.'' Given the influence of this theory and the hundreds of
+claimed ``replications,'' it is essential to ask whether many of these findings
+may instead be measurement artifacts—spurious effects created when the
+``neutral'' point of the measurement scale fails to align with the group’s own
+center of opinion.
 
-\begin{quote}
-One of the most robust findings in social psychology is that of attitude
-polarization following discussion with like-minded others (Cooper,
-Kelly, and Weaver 2001).
-\end{quote}
+The group polarization hypothesis is that discussion within a group of
+likeminded people causes the group's average opinion to increase, or
+\emph{polarize}~\cite{Brown1986,Brown2000}.
+Group polarization experiments therefore attempt to
+polarize groups as follows: first, participants report their opinion on
+some topic or prompt chosen by the experimenters in advance. Second, 
+participants are put into likeminded groups to
+discuss the topic. Following discussion, participants again report their
+opinions on the topic. Group polarization is defined as an increase in
+the average opinion following discussion. If the group polarized, and we 
+assume that people in the group found consensus—meaning their opinions agreed
+more after the discussion than before—then initially moderate opinions must have
+increased more than more extreme ones.
 
-If an extremist's opinion falls and it is measured with a Likert scale,
-will it make a noise? The answer had better be yes if we want to
-properly measure opinion change. Accurate measurements are critical for
-developing rigorous and reliable methods to promote more stable,
-sustainable, responsive, and efficient governments and institutions
-(Mason 2018; Klein 2020). Unfortunately, social and behavioral
-scientists have often used statistical methods that distort opinion
-change when it's measured on an ordinal, Likert-style scale (e.g.~-3
-indicates ``Strongly disagree'', +3 indicates ``Strongly agree'', and 0
-is neutral) (Liddell and Kruschke 2018). This includes a phenomenon
-called \emph{group polarization} that former Obama White House official
-and law professor Cass Sunstein relied on to explain why ``people become
-extremists'' and why ``political and cultural polarization'' is ``so
-pervasive in America'', as the publisher summarizes (Sunstein 2009).
-Sunstein suggests group polarization research provides a ``clue'' to
-explaining how ``facism\ldots student radicalism\ldots Islamic
-terrorism\ldots{[}t{]}he Rwandan genocide\ldots{[}e{]}thnic
-Conflict\ldots acts of torture and humiliation by American soldiers at
-Abu Ghraib'', and more (p.~1).
+We are sad to report that 95\% of experimental reports of group polarization
+across sixty published experimental conditions cannot be counted as
+replications. We explain here how evidence of group polarization is undermined
+by a basic measurement flaw that obscures opinion change among extremists.
+This opens the possibility that what seems to be group polarization could just
+be \emph{mere agreement}, with no change in mean opinion, if only increases in
+opinion were measured but not decreases.  To evaluate whether the empirical
+record reflects genuine opinion change or merely measurement artifacts, I
+simulated published experiments under conditions of mere agreement. Each
+simulation reproduced the core elements of published designs—pre-discussion
+survey, group interaction, post-discussion survey—while holding the latent
+mean opinion constant. Because the original datasets are lost and published
+results rarely provide enough information to reconstruct them, we cannot know
+whether any reported effect reflects a real shift or a measurement artifact.
+When distinct latent processes can generate indistinguishable results, the
+effect is non-identifiable—and no claim of replication can stand.
 
-% \begin{itemize}
-% \tightlist
-% \item
-We developed a generative model of opinion change to identify, if possible,
-null opinion change models that could spuriously appear like group polarization
-when naively measured using a Likert scale without accounting for the measurement procedure. 
-We further applied our method to estimate the family-wise error rate of
-detections of group polarization in a corpus of ten influential papers from 
-across several decades and from various theoretical traditions. We turn this criticism of past
-blunders constructive by providing a procedure to calculate what Cohen's $d$ should be
-used to guarantee a family-wise error rate (i.e., Type I error rate) of 5\% or less.
+A relatively simple step could fix the problem: re-center the post-discussion
+scale on each group’s own pre-discussion mean, aligning measurement with the
+group’s frame rather than an arbitrary external ``neutral.'' This adjustment
+restores the basic condition for inference: measuring change within a
+consistent frame of reference.
 
-We calculate the severity of the problem in two steps. In the first step we 
-provide proof by example that there exists a null opinion change model in terms of \emph{latent, continuous}
-psychological opinions that looks like group polarization when measured via participant behavior, i.e.,
-by their response on an \emph{ordinal survey scale}. If such a null model is found, this means that we cannot
-rule out a null effect, given the published data. This means, therefore, that the published detection is
-just as likely false as it is true, so we say it is a \emph{plausibly false detection of group polarization},
-or a \emph{plausibly spurious detection} for short.
-
-In the second step of the analysis of whether, we simulate observations
-of group polarization experiments where participant opinions are drawn from the 
-null model distributions identified in the first step,
-then these are measured on a Likert scale for simulated group polarization survey
-responses on an ordinal scale. We then calculate the
-family-wise error rate (i.e., Type I error rate) and false detection rate using
-various effect sizes to determine significance of the effect in terms of Cohen's $d$.
-The family-wise error rate is defined as the probability that a known null 
-effect is detected as significant. We use this to then estimate the *false detection rate*, which is the
-probability that a detection of an effect is a false one, which depends on the relative prevalence of
-true and false positive detections in the literature.
-Finally, we put our model to constructive use to show what Cohen's $d$ must be set to
-to limit the family-wise error rate to 5\%, denoted $d*$. This will not help rescue any of
-the published results here found to be plausibly spurious. But, it provides a measure of the weakness of
-the published experimental designs independent of their shared fatal failure to 
-account for the measurement procedure: greater $d*$ indicates weaker designs.
-
-We find that 95\% (54 of 57) of group polarization detections are plausibly spurious.
-In other words, empirical support for group polarization has been seriously undermined
-just by removing these results from the count of group polarization replications.
-There is no reason to expect other papers to be more reliable. (Summarize other
-two findings, 1. FWER/FDR and 2. $d*$ for FWER < 5\%)
-
-Our methods can be turned on any data-driven social science analysis of
-data where participants report a latent, subjective psychological
-variable (mood, physiological state, etc.), and where various theories
-supposedly battle it out to best represent the phenomenon of interest
-via null-hypothesis significance testing using ordinal data as continuous. 
-This may just be the first
-of many bricks to fall since \emph{all 68} of the articles that
-mentioned the word ``Likert'' and analyzed ordinal data used metric
-models to detect latent psychological variable change across three
-influential psychology journals, \emph{Psychological Science}, the
-\emph{Journal of Experimental Psychology: General}, and the
-\emph{Journal of Personality and Social Psychology} (Liddell and
-Kruschke 2018). This further confirms that replication is far from the only
-foundational crisis that behavioral science faces\cite{Yarkoni2022,Turner2022}.
-
-Replication is not enough:
-strong, mechanistic models must be developed to be able to predict Type I and II
-error rates \emph{ahead of time}, as we demonstrate here. We only demonstrate how to
-calculate expected Type I error rates for a measurement design, i.e., the rate of false 
-positives. Low statistical power exacerbates false detection rates. So, to be confident in replication,
-a pre-registered measurement design should use this style of simulation to 
-estimate the statistical power. Tangled, weak, informal
-verbal models of complex psychological and social phenomena like those that motivated
-empirical studies of group polarization tend to be vulnerable
-to stastical weaknesses. It is difficult to incorporate prior knowledge such
-as experiment structure without a sufficiently rigorous and appropriate 
-formal system \cite{Smaldino2017,Smaldino2020,Flake2020,Deffner2024}.
+But even if this measurement problem is fixed,
+it is one of several severe problems in this kind of social science: Unstable
+opinions, unmodeled input variance, and incommensurate experimental designs
+all uniquely undermine any cumulative insight group polarization might
+claim—problems known about and ignored for decades~\cite{Cartwright1973}. No
+degree of scientific lawfulness can be established without evidence gathered
+from rigorously modeled, commensurate experiments~\cite{Cartwright1999}.
 
@@
 
-(DO THE MINIMUM AFTER EVERYTHING ELSE)
-
-To situate this current paper it will help to introduce some relevant research
-traditions and common experimental designs in group polarization, using the ten
-studies included in our corpus as touchstones. To begin, *group polarization*
-is defined as...(continue basic definition with some examples from our corpus).
-
-Paragraph dos: 
-Common experimental, measurement, and statistical procedures to induce and
-detect group polarization
-
-\subsection{How study design can induce false detections of group polarization}
-\label{ssubsec:false-detections}
-
-An opinion is never measured directly, only a behavior that indicates an opinion
-in a behavioral experiment or survey.  For example, the location of a mark
-that a participant makes is measured, not the complex neural activity that
-underlies forming opinions.
-Theoretically, an internal opinion in the form of neural activity is a
-\textit{latent opinion}, from the Latin \emph{latentem} meaning ``lie
-hidden''.  Participants report \text{observed opinions}, typically as one of
-a finite number of ranked, successive integer values that form the
-\emph{selection set}. The minimum possible value in the selection set might
-mean ``strongly disagree'' and the maximum value represents ``strongly
-agree''.
-
-Group polarization experiments share a common experimental design where,
-first, participant opinions are gathered on a topic of interest. Next, groups
-are formed that have some initial opinion extremism. Participants deliberate
-or discuss the topic for some set time. After the set time, participants
-again give their opinion. Group polarization is said to occur if the
-mean group opinion became more extreme, i.e., if the magnitude of the mean
-group opinion is greater \emph{post-deliberation} compared to
-\emph{pre-deliberation}. For example,
-\cite{Moscovici1969} asked Parisian high school students to indicate their
-degree of agreement with the statement, ``American economic aid is always
-used for political pressure''. There
-were seven options in the selection set, -3 indicating ``strongly disagree'',
-+3 indicating ``strongly agree'', and 0 was neutral. Moscovici and Zavalloni
-report that the mean pre-deliberation opinion was -0.61 and the mean
-post-deliberation opinion was -1.09 (Moscovici and Zavalloni, 1969; Table 4,
-p. 132). This is a -0.48 shift to more extreme disagreement with the 
-statement, reported by Moscovici and Zavalloni as statistically significant
-in Table 5 on p. 132.
-
-(PREVIOUS TWO PARAGRAPHS SET UP THIS FINAL PARAGRAPH: CONTINUE TO USE MOSCOVICI 1969 TO
-EXPLAIN IN THE SIMPLEST TERMS HOW FAILING TO ACCOUNT FOR MEASUREMENT PROCEDURE
-CAN LEAD TO FALSE POSITIVES; see Figure \ref{fig:distros})
-
-\subsection{Overview}
-
-In the next section, we introduce the various formal and computational Methods 
-used to obtain our findings outlined above. We then explain our findings in detail in the
-Analysis section. We close with a Discussion of how this is not a defeat, but
-a new beginning for group polariation, given that there are good reasons to expect
-group polarization to occur. We also explain how to expanding this analysis to
-examine other published studies on group polarization and across social and behavioral
-science disciplines.
+Group polarization research began in the 1960s, which grew rapidly with
+funding from the National Institutes of Mental Health (NIMH), and widespread interest
+in what was then a surprising new science of radicalization.
+Dorian Cartwright (1973) reviewed the first NIMH grants and at the time said,
+`` ''.

### 2025-10-23 GroupPolarizationStatmod/paper/ms.tex
tightening intro

@@
 \makeatletter
-\title{If the Null Fits, You Must Omit: Ubiquitous False Detections of Group Polarization}
+% \title{If the Null Fits, You Must Omit: Ubiquitous False Detections of Group Polarization}
+\title{If the Null Fits, You Must Omit: Measurement Artifacts Undermine Group Polarization Replications}
 \makeatother
@@
 \maketitle
-% \begin{abstract}
-% \noindent 
-% \emph{Group polarization} is the name for a form of consensus where
-% members of a like-minded group become, on average, more extreme in their
-% opinions after discussing a topic. Group polarization is important because
-% it may increase social and political tensions if moderates become more
-% extreme. Decades of studies have replicated detections of group
-% polarization, typically using an ordinal scale to measure opinions.
-% However, these studies did not account for ordinal measurements, which can
-% result in spurious detections of opinion change.  Lacking original data we
-% can still calculate the probability that a published detection of group
-% polarization was spurious, i.e., to calculate the \emph{false detection
-% rate}.  Across 54 group polarization experimental conditions from ten
-% representative journal articles, we calculated false detection rates to be
-% between 0.52 and 0.88, with a median of 0.75, using a generative model of
-% group polarization experiments seeded with empirical data.  
-% We also use our model to develop experimental designs that
-% achieve an acceptable false discovery rate.  Much of group polarization
-% research may be unreliable. This analysis can help change that by enabling others to avoid this
-% hidden pitfall. More
-% broadly this work demonstrates one important way that replication success alone
-% does not imply epistemic reliability.
-% \end{abstract}
 
@@
 \begin{quote}
-One of the most robust findings in social psychology is that of attitude polarization 
-following discussion with like-minded others.
-\par\raggedleft(Cooper et al, 2001, p. 267 \cite[p. 267]{Cooper2001})
-\end{quote}
-
-\begin{quote}
-A medium effect size is conceived as one large enough to be visible to the naked eye.
-\par\raggedleft(Cohen, 1988 \cite[p. 26]{CohenBook1988})
+  The good thing about science is that it's true, whether or not you believe
+  it. (Neil de Grasse Tyson slogan circa 2018)
 \end{quote}
@@
 
-\mt{One more brick to fall in the ongoing replication and generalizability crisis \autocite{Yarkoni2022}.}
-
 Since there seems to be no selection pressure on false discovery rates,
@@
 
-Our analysis brings even more additional questions into focus. Most
-significantly, recall that even the more appropriate ordered probit Bayesian
-model often failed to accurately detect negative results with the number of
-participants used in the original papers. Many of the studies analyzed here
-did not use any statistical test to establish significance; those that did
-used some form of a $t$-test, and it is not clear if the $t$-test that was
-used accounted for a change in pre- and post-deliberation opinion variance,
-which we expect due to consensus.
-
-The ordered probit model we used still likely underestimates the variance in
-outcomes, which could require even larger sample sizes. A more rigorous
-analysis would include additional sources of variance that will likely
-further inflate the Type I error rate, $\alpha$, and the false discovery
-rate~\cite{Yarkoni2022}. First and foremost, there is variance within each
-experimental deliberation group; no group polarization study accounts for
-this, instead treating participants in each experimental condition as if
-they were one large group. Furthermore, survey data are known to be
-noisy~\cite{Zaller1992}. People report opinions differently over time for no
-apparent reason (i.e., opinions are \emph{unstable}).  Context matters: the
-order in which a survey question is asked, and which question framing is
-chosen among logically equivalent alternatives, can both significantly
-influence participant responses. Finally, little is known about the
-psychological process that converts latent opinions to reporting behaviors
-(e.g., clicking a radio button corresponding to an opinion). 
-The effect of accounting for these sources of variance must be understood to
+When we used the appropriate ordered probit Bayesian statistical model, it
+still often failed to accurately identify mere agreement, instead detecting
+group polarization.  The ordered probit model we used still likely
+underestimates the variance in outcomes, which could require even larger
+sample sizes. A more rigorous analysis would include additional sources of
+variance that will likely further inflate the Type I error rate, $\alpha$,
+and the false discovery rate~\cite{Yarkoni2022}. First and foremost, there is
+variance within each experimental deliberation group; no group polarization
+study accounts for this, instead treating participants in each experimental
+condition as if they were one large group. Furthermore, survey data are known
+to be noisy~\cite{Zaller1992}. People report opinions differently over time
+for no apparent reason (i.e., opinions are \emph{unstable}).  Context
+matters: the order in which a survey question is asked, and which question
+framing is chosen among logically equivalent alternatives, can both
+significantly influence participant responses. Finally, little is known about
+the psychological process that converts latent opinions to reporting
+behaviors (e.g., clicking a radio button corresponding to an opinion).  The
+effect of accounting for these sources of variance must be understood to
 estimate $\alpha$ (and power, $W$) for the design of the next generation of
@@
 
-To close, note that replication of weakly- and incommensurately-specified 
-social phenomena is not enough: strong, mechanistic models must be developed to
-be able to predict Type I and II error rates \emph{ahead of time}, as we
-demonstrate here. We only demonstrate how to calculate expected Type I error
-rates for a measurement design, i.e., the rate of false positives. Low
-statistical power exacerbates false detection rates. So, to be confident in
-replication, a pre-registered measurement design should use this style of
-simulation to estimate the statistical power. It is difficult to incorporate
-prior knowledge such as experiment structure without a sufficiently rigorous
-and appropriate formal system
-\cite{Smaldino2017,Smaldino2020,Flake2020,Deffner2024}.
+A simple change in measurement procedure could fix the problem identified
+here: center the post-discussion scale on each group’s own pre-discussion
+mean. This removes the assumption that all groups share a common “neutral”
+midpoint. The situation is like tossing a ball on a moving train: to the
+passengers, the ball’s speed is the same in both directions, but to an
+observer on the ground one throw appears faster and the other slower.
+Likewise, when measurement is anchored to an external scale, simple agreement
+within a group can appear as movement toward one pole. Re-centering the
+measurement frame corrects this distortion. It acknowledges that opinion
+change is relative to the observer’s reference frame—there is no fixed
+background of neutrality against which all groups can be compared. This is,
+in essence, the insight of Newtonian relativity: apparent motion depends on
+the frame from which it is measured.
+
+First we must note that there are even more, perhaps even graver problems
+with group polarization research than I discussed so far: one of note is
+opinion instability. Opinions are not generally stable—people seem to
+construct their opinions based on several factors not directly related to
+whatever underlying topic. Factors affecting reported opinions include the
+language used to frame a prompt or question; the order of questions on a
+multi-item survey; or simply when the question was asked (Kalton and Schuman,
+1982; Zaller and Feldman, 1992; Zaller 1992). Failure to account for sources
+of input variance in any experiment can lead to inflated effect sizes since
+the input variance is missing from the posterior predictions during
+statistical model fitting (Yarkoni, 2022). This further compounds the
+theoretical confusions that fill the void when theories are never formalized
+and grounded in a mathematical or computational model (Smaldino 2020; Turner
+and Smaldino, 2022).
+
+A second additional problem with group polarization research is that we literally cannot learn anything by comparing summary statistics obtained with incommensurate experimental designs—but this is exactly how group polarization research has tried to make progress.
+Comparing summary statistics only has explanatory power if datasets were generated under the same experimental designs—it seems this was never done or even considered in the group polarization literature.
+In order to compare studies, their datasets must be generated from commensurate experiments, meaning the studies must share the same predictive and statistical models and experimental design.
+Otherwise, such comparisons are meaningless, they literally tell us nothing, as shown by philosopher of science Nancy Cartwright in Chapter 5 of her book The Dappled World: A Study of the Boundaries of Science.
+
+\subsection{Humble Hearts Promote Rigorous Science}
+
+Cass Sunstein kicked off the 2000s by elevating group polarization to be a scientific law and advocating the application of group polarization research to serious real-world problems in two books (Sunstein, 2009; Sunstein, 2019) and two recent high-profile articles in Nature Human Behavior: one on the social foundations of pandemic preparedness (Van Bavel, et al., 2020) and one to “promote truth, autonomy, and democratic discourse online” (Lorenz-Spreen, Lewandowski, Sunstein, and Hertwig, 2020).
+Let us try to stop the spread of this bad science—and let us explore what must be done to achieve Dorian Cartwright’s vision of a rigorous science of group polarization he articulated fifty years ago.
+
+To be practically useful, science must be rigorous.
+Rigorous science comes most naturally when scientists remain humble, especially when they serve at the pleasure of the public, the government, or other donors.
+Here’s another example from Burnstein and Vinokur (1973) that we can learn from—let’s try to avoid the sort of arrogance and haughtiness that seep from these acknowledgements they wrote on p.
+123:
+
+\begin{quote}
+This research was supported by a Grant from the National Institute of Mental Health (MH-16950-03) and by a Guggenheim Foundation Fellowship awarded to the first author.
+We gratefully acknowledge the incisive, witty, and vinous commentary provided by colleagues at the University of Provence, in particular Robert Abelson, Claude Flament, and Jean-Pierre Poitou, as well as the spirited and intelligent efforts of Livia Mezrick and Irene Graczyk, who helped us carry out the study.
+\end{quote}
+
+Science is not glamorous—it takes guts and sacrifice to achieve the technical skills and discipline to sharpen one’s prose, methods, mathematics, and computer code to the point they can really cut reality at its joints, investigate different parts, then put it all back together with a message for others about how the world really works.
+These days, to do rigorous social science work one must wisely curate the best of the available social science by harmonizing cross-disciplinary jargon.
+But disincentives still exist against careful scholarship.
+Prizes and professorships go to those with the most numerous publications in the right journals—period.
+
+Unpaid, anonymous reviewers are supposed to act as gatekeepers to catch problems like the ones I reviewed.
+But, journal article reviewers have little to gain from enforcing rigor—actually there is a lazy strategy that can pay off big for freeloaders.
+It’s easy to imagine a cultural norm evolving where reviewers save themselves time and energy by asking little of an author, in hopes that little will be demanded of them when the reviewer submits their next paper.
+This could result in a publication process that rewards the laziest and most willing to flood the zone and self-aggrandize.
+
+\subsection{Conclusion}
+
+Social science can use physics as a model discipline to transition from idiosyncratic and Balkanized to informative and united.
+Physics only changed the world because it is useful—it is useful only because it reliably predicts things.
+Physics became more useful as it became more united, finding shared representations for concepts developed independently—for example the science of optics was developed independently from the science of dynamics, but the physics of light is explained partly in analogy to a pendulum or spring—the mathematics of harmonic motion can be used to predict the motion of springs or pendulums—or why we see different colors of light through a prism.
+With careful, coordinated tweaks to group polarization experimental design and theoretical modeling, we will build a more rigorous theory of group polarization and its large-scale expression, echo chamber radicalization.
+
+I expect that group polarization experiments, when properly constructed and analyzed, will help sustainability campaigns optimize their social influence campaigns.
+We can start by developing simpler experimental designs based on formal or computational models capable of predicting outcomes in an experiment.
+I have developed a model that could motivate an experimental design: I showed that if extremists are more stubborn, this is sufficient to cause group polarization when likeminded groups interact over time.
+There is even a free stubborn extremism parameter that could be fit to data to measure the relationship between stubbornness and extremism on different issues.
+
+Group polarization research so far has mostly provided us with a checklist of what not to do.
+The alternative approach I outline builds on rock—the rock of rigor—and will supersede past group polarization research built on a thousand grains of inexact verbal speculation.
 

### 2025-10-24 GroupPolarizationStatmod/paper/intro-latest.tex
nearing complete Intro

@@
 
-But even if this measurement problem is fixed,
-it is one of several severe problems in this kind of social science: Unstable
+But even if this measurement problem were fixed,
+there are several problems in this kind of social science: Unstable
 opinions, unmodeled input variance, and incommensurate experimental designs
@@
 Group polarization research began in the 1960s, which grew rapidly with
-funding from the National Institutes of Mental Health (NIMH), and widespread interest
-in what was then a surprising new science of radicalization.
-Dorian Cartwright (1973) reviewed the first NIMH grants and at the time said,
-`` ''.
+focused government funding coupled with widespread
+interest in what was then a surprising new science of
+radicalization~\cite{Cartwright1971}. Dorian Cartwright (1973) reviewed the
+first decade of group polarization research and found problems: he lamented
+the ``disturbing fact'' that group polarization research has produced no
+cumulative theoretical undersatnding of the effects of discussion on
+group opinions~\cite{Cartwright1973}. Cartwright was somewhat dismayed, but
+optimistically hoped that group polarizaiton research would eventually
+produce ``an enduring\ldots more comprehensive paradigm'' \cite[p.
+230]{Cartwright1973}.
+
+Cartwright (1973) suggested that a ``vigorous search for alternative
+formulations'' could help group polarization become more useful. More
+researchers developed specific theories of group
+polarization, with two becoming particularly influential: social comparisons
+and persuasive arguments. The social comparisons hypothesis says group
+polarization occurs because humans want to fit in, and so generally like to . Group polarization occurs, then, when a
+group is composed mostly of extremists, with a sufficient number of people
+with a weak opinion \cite{Myers1970; Myers1975}. Its main
+“competitor” from the 1970s was the persuasive arguments theory, which
+basically claimed that rhetoric was the strongest driver of group
+polarization, not social comparisons \cite{Burnstein1973,Burnstein1975}.
+
+Other notable hypotheses include the
+self-categorization and social decisions scheme hypotheses.
+Self-categorization augmented the social comparisons hypothesis with a
+consideration of idiosyncratic beliefs people hold that are the result of
+their own personal lived experience \cite{Abrams1990,Hogg1990}.
+The social decisions hypothesis posits that contextual factors like
+social network structure could be at least as important as the distribution
+of opinions and . Noah Friedkin, a sociologist, notably published
+the study with the least potential measurement artifacts for a rigorous study on group polarization, adopting the social
+decisions scheme hypothesis, predicting that social connectivity within a
+group could amplify group polarization \cite{Friedkin1999}.
+
+\subsection{Measurement procedure}
+
+In most group polarization experiments, opinions are measured on an
+ordinal scale: Burnstein and Vinokur (1975) for example used a ten-point
+scale from 1 to 10 representing “the lowest odds of success acceptable” to
+try a certain behavior, where 1 means 10\%—Moscovici and Zavalloni (1969) gave participants a seven-point scale, from “strongly disagree” (-3) to “strongly agree” (+3) to report their opinions.
+In these and every other survey-based group polarization study I know, the reported opinions on the ordinal scale—with discrete numbered bins—were treated as if they were continuous-valued, or in other words as if they were the internal mental opinion value itself, whatever that might be neurobiologically.
+Liddell and Kruschke (2018) showed that in the general case this setup can
+lead to false positives because ordinal scales cannot track changes in
+extreme opinions beyond the ceiling of the measurement scale—this approach is
+ubiquitous in social psychology and survey research, not at all limited to
+group polarization.
+
+\subsection{Study Corpus}
+
+We selected ten journal articles with sixty total experimental conditions
+across the ten.
+
+\begin{table}[ht]
+\caption{\textbf{Journal articles in our corpus.}}
+\centering
+\begin{tabular}{clcc}
+\toprule
+Authors (Year) & Title & Journal & Category \\
+\midrule
+  Moscovici \& Zavalloni (1969) & The Group as a Polarizer of Attitudes &
+    Journal of Personality and Social Psychology & Politics \\
+    Schkade, Sunstein, \& Hastie & When Deliberation Produces Extremism & Critical Review
+                                 & Politics \\
+Burnstein \& Vinokur (1973) & Testing Two Classes of Theories About Group Induced Shifts
+In Individual Choice & Journal of Experimental Social Psychology & Persuasive Arguments \\
+
+Burnstein \& Vinokur (1975) & What a person thinks upon learning he has chosen differently from others: Nice evidence for the persuasive-arguments explanation of choice shifts & Journal of Experimental Social Psychology & Persuasive Arguments \\
+\bottomrule
+\end{tabular}
+\label{table:mr}
+\end{table}

### 2025-10-25 GroupPolarizationStatmod/paper/intro-latest.tex
finished corpus table

@@
 
-A relatively simple step could fix the problem: re-center the post-discussion
-scale on each group’s own pre-discussion mean, aligning measurement with the
-group’s frame rather than an arbitrary external ``neutral.'' This adjustment
-restores the basic condition for inference: measuring change within a
-consistent frame of reference.
+A relatively simple step could fix the problem, we believe: measure twice
+before discussion. Before forming groups, experimenters could poll opinions
+using a scale centered arbitrarily at 0, with varying strengths of disagree
+or agree with negative and positive integers, respectively. 
+Then after likeminded groups are
+selected, measure group opinions again before they discuss, but now with the
+scale centered on the average group opinion. To keep the same number of bins,
+add more bins to express more extreme views and take away bins from
+the opposite side as necessary.
 
-But even if this measurement problem were fixed,
-there are several problems in this kind of social science: Unstable
-opinions, unmodeled input variance, and incommensurate experimental designs
-all uniquely undermine any cumulative insight group polarization might
-claim—problems known about and ignored for decades~\cite{Cartwright1973}. No
-degree of scientific lawfulness can be established without evidence gathered
-from rigorously modeled, commensurate experiments~\cite{Cartwright1999}.
+Even if this measurement problem were fixed, equally severe problems remain
+for group polarization research. First there seems to be no stable 
+``opinions'' per
+se—it is more accurate to say people construct opinions when they are
+prompted or need them for some other reason, as revealed by respondents
+answering questions differently if the framing or ordering of them changes,
+and by the fact that people respond differently to the same question over
+time. These are just a few of several potential sources of variance that were never accounted for, which is also known to generate false detections of change. 
+Finally, because group polarization experimental design was never
+standardized its research corpus cannot cumulatively or systematically tell 
+us anything general about the world—Dorian
+Cartrwright (1973) tried to alert group polarization researchers to this
+problem half a century ago~\cite{Cartwright1973}. General scientific claims
+of the existence and functioning of some scientifically induced phenomenon
+only work if the evidentiary observations were collected from
+\emph{commensurate} experiments, or in other words, only from experiments
+that share the exact same causal structure as expressed in a formal or 
+mechanistic causal model~\cite{Cartwright1999}.
 
@@
 
+\subsection{Study Corpus}
+
+We selected ten representative journal articles with sixty total experimental
+conditions among them. For a representative sample of group polarization
+research across decades, we selected two articles for each of five article
+categories: \emph{Social Comparisons}, \emph{Persuasive Arguments},
+\emph{Self-Categorization}, \emph{Social Decisions}, and \emph{Politics}. The
+first four of these categories reference the theoretical tradition that the
+authors were trying to prove right—the fifth, \emph{Politics}, was the
+subject matter of the studies, but the two studies took no clear side in the
+group polarization theory wars. This resulted in a representative sample of
+group polarization research over time, as well: one article is from 1969,
+four are from the 1970s, . Information about our corpus is summarized below
+(Table~\ref{tab:articles}).  Please see the Supplemental Information
+
+% \begin{table}[ht]
+
+%   \caption{\textbf{Journal articles in our corpus.}}
+%   \label{tab:articles}
+
+%   \centering 
+%   \rowcolors{2}{white}{gray!7}
+%   \begin{tabular}{m{1.35in} m{2.45in}
+%     m{1.3in} m{0.85in}}
+
+%     \toprule
+
+%     \textbf{Authors (Year)}& \textbf{Title} & \textbf{Journal} &
+%     \textbf{Category} \\
+
+%     \midrule 
+%     Moscovici \& Zavalloni (1969) & The Group as a Polarizer of
+%     Attitudes & \emph{JPSP} & Politics \\ 
+%     Schkade, Sunstein, \& Hastie (2010) & When
+%     Deliberation Produces Extremism & \emph{Critical~Review} & Politics \\
+
+%     Bishop \& Myers (1970) & Discussion Effects on Racial Attitudes &
+%     \emph{Science} & Social Comp's \\ 
+%     Myers (1975) & Discussion-Induced Attitude Polarization & 
+%     \emph{Human Relations} & Social Comp's \\
+
+%     Burnstein \& Vinokur (1973) & Testing Two Classes of Theories About Group
+%     Induced Shifts In Individual Choice & \emph{JESP} & Persuasive Arg's \\
+%     Burnstein \& Vinokur (1975) & What a person thinks upon learning he has
+%     chosen differently from others: Nice evidence for the
+%     persuasive-arguments\ldots & \emph{JESP} & Persuasive Arg's \\
+
+%     Abrams, et al., (1990) & Knowing What to Think by Knowing Who You Are:
+%     Self‐Categorization and the Nature of Norm Formation\ldots & 
+%     \emph{Brit. J. of Soc. Psych.} & Self-Categ'n \\
+%     Hogg, Turner, \& Davidson (1990) & Polarized Norms and Social Frames of 
+%     Reference: A Test of the Self-Categorization Theory of Group 
+%     Polarization & \emph{B. \& Appl. Soc. Psych.} & Self-Categ'n \\
+%     Krizan \& Baron (2007) & Group Polarization and Choice Dilemmas: How
+%     Important is Self-Categorization? & \emph{Eur. J. of Soc. Psych.} &
+%     Self-Categ'n \\
+
+%     Friedkin (1999) & Choice Shifts \& Group Polarization & 
+%     \emph{Amer. Soc.  Rev.} & Soc'l Decisions \\
+
+% \bottomrule
+% \end{tabular}
+% \end{table}
+\begin{table}[ht]
+\centering
+\caption{\textbf{Journal articles in our corpus.}}
+\label{tab:articles}
+
+\begin{tabular}{m{0.9in} m{1.35in} m{2.45in} m{1.3in}}
+\toprule
+\textbf{Category} & \textbf{Authors (Year)} & \textbf{Title} & \textbf{Journal} \\
+\midrule
+
+\rowcolor{gray!10}
+\multirowcell{2}{Politics} 
+  & Moscovici \& Zavalloni (1969) & The Group as a Polarizer of Attitudes & \emph{JPSP} \\
+\rowcolor{gray!10}
+  & Schkade, Sunstein, \& Hastie (2010) & When Deliberation Produces Extremism & \emph{Critical~Review} \\[0.2em]
+
+\rowcolor{gray!5}
+\multirowcell{2}{Social Comp's \vspace{0.7em}}
+% \multirowcell{2}{Social Comp's}
+  & Bishop \& Myers (1970) & Discussion Effects on Racial Attitudes & \emph{Science} \\[0.2em]
+\rowcolor{gray!5}
+  & Myers (1975) & Discussion-Induced Attitude Polarization & \emph{Human Relations} \\[0.3em]
+
+\rowcolor{gray!10}
+\multirowcell{2}{Persuasive Arg's}
+  & Burnstein \& Vinokur (1973) & Testing Two Classes of Theories About Group Induced Shifts In Individual Choice & \emph{JESP} \\
+\rowcolor{gray!10}
+  & Burnstein \& Vinokur (1975) & What a person thinks upon learning he has chosen differently from others: Nice evidence for the persuasive-arguments\ldots & \emph{JESP} \\
+
+\rowcolor{gray!5}
+\multirowcell{3}{Self-Categ'n}
+  & Abrams, et al., (1990) & Knowing What to Think by Knowing Who You Are: Self‐Categorization and the Nature of Norm Formation\ldots & \emph{Brit. J. of Soc. Psy.} \\
+\rowcolor{gray!5}
+  & Hogg, Turner, \& Davidson (1990) & Polarized Norms and Social Frames of Reference: A Test of the Self-Categorization Theory of Group Polarization & \emph{B. \& Appl. Soc. Psy.} \\
+\rowcolor{gray!5}
+  & Krizan \& Baron (2007) & Group Polarization and Choice Dilemmas: How Important is Self-Categorization? & \emph{Eur. J. of Soc. Psy.} \\
+
+\rowcolor{gray!10}
+\multirowcell{1}{Soc'l Decisions}
+  & Friedkin (1999) & Choice Shifts \& Group Polarization & \emph{Amer. Soc. Rev.} \\
+
+\bottomrule
+\end{tabular}
+\end{table}
+
 \subsection{Measurement procedure}
 
-In most group polarization experiments, opinions are measured on an
-ordinal scale: Burnstein and Vinokur (1975) for example used a ten-point
+In studies of group polarization, opinion measurements are typically ordinal, reflecting
+ordered positions rather than measurable distances between them—
+Moscovici and Zavalloni (1969) gave participants a seven-point scale, from “strongly disagree” (-3) to “strongly agree” (+3) to report their opinions.
+Burnstein and Vinokur (1975) used a ten-point
 scale from 1 to 10 representing “the lowest odds of success acceptable” to
-try a certain behavior, where 1 means 10\%—Moscovici and Zavalloni (1969) gave participants a seven-point scale, from “strongly disagree” (-3) to “strongly agree” (+3) to report their opinions.
+try a certain behavior, where 1 means 10\%.
 In these and every other survey-based group polarization study I know, the reported opinions on the ordinal scale—with discrete numbered bins—were treated as if they were continuous-valued, or in other words as if they were the internal mental opinion value itself, whatever that might be neurobiologically.
-Liddell and Kruschke (2018) showed that in the general case this setup can
+
+Note how ordinal scales clip all latent opinions with a magnitude greater than the
+magnitude of its extreme values. 
 lead to false positives because ordinal scales cannot track changes in
@@
 group polarization.
-
-\subsection{Study Corpus}
-
-We selected ten journal articles with sixty total experimental conditions
-across the ten.
-
-\begin{table}[ht]
-\caption{\textbf{Journal articles in our corpus.}}
-\centering
-\begin{tabular}{clcc}
-\toprule
-Authors (Year) & Title & Journal & Category \\
-\midrule
-  Moscovici \& Zavalloni (1969) & The Group as a Polarizer of Attitudes &
-    Journal of Personality and Social Psychology & Politics \\
-    Schkade, Sunstein, \& Hastie & When Deliberation Produces Extremism & Critical Review
-                                 & Politics \\
-Burnstein \& Vinokur (1973) & Testing Two Classes of Theories About Group Induced Shifts
-In Individual Choice & Journal of Experimental Social Psychology & Persuasive Arguments \\
-
-Burnstein \& Vinokur (1975) & What a person thinks upon learning he has chosen differently from others: Nice evidence for the persuasive-arguments explanation of choice shifts & Journal of Experimental Social Psychology & Persuasive Arguments \\
-\bottomrule
-\end{tabular}
-\label{table:mr}
-\end{table}

### 2025-10-26 GroupPolarizationStatmod/paper/intro-latest.tex
finished intro draft

@@
 that group polarization ``does not occur with every group'' and that ``the
-effect is not large.'' Given the influence of this theory and the hundreds of
-claimed ``replications,'' it is essential to ask whether many of these findings
-may instead be measurement artifacts—spurious effects created when the
+effect is not large''~\cite{Brown1986}. 
+Given the influence of this theory and the hundreds of claimed ``replications'' of
+group polarization, it is essential to ask whether any of these replications may
+instead be \emph{measurement artifacts}—spurious effects created when the
 ``neutral'' point of the measurement scale fails to align with the group’s own
-center of opinion.
+center of opinion. The science of opinion change has a central role in designing new
+methods for social organization and social change to face pressing existential
+challenges~\cite{Galesic2021,Galesic2023}.
 
@@
 conditions among them. For a representative sample of group polarization
-research across decades, we selected two articles for each of five article
+research across decades, we two articles for three of the five article
 categories: \emph{Social Comparisons}, \emph{Persuasive Arguments},
-\emph{Self-Categorization}, \emph{Social Decisions}, and \emph{Politics}. The
-first four of these categories reference the theoretical tradition that the
-authors were trying to prove right—the fifth, \emph{Politics}, was the
-subject matter of the studies, but the two studies took no clear side in the
-group polarization theory wars. This resulted in a representative sample of
-group polarization research over time, as well: one article is from 1969,
-four are from the 1970s, . Information about our corpus is summarized below
-(Table~\ref{tab:articles}).  Please see the Supplemental Information
+ and \emph{Politics}. We included three \emph{Self-Categorization} studies and one \emph{Social Decisions}.
+All but the \emph{Politics} category refers to the theoretical tradition that the
+authors were trying to prove right—the two \emph{Politics} articles were agnostic
+to group psychology theories. This resulted in a representative sample research
+over time, as well: we have one study from 1969,
+four from the 1970s, three from the 1990s, one from 2007 and one from 2010. 
+Information about our corpus is summarized below with references 
+(Table~\ref{tab:articles}).
 
-% \begin{table}[ht]
+\include{article-table.tex}
 
-%   \caption{\textbf{Journal articles in our corpus.}}
-%   \label{tab:articles}
+Each study used subtly different experimental methods that make it difficult
+to build cumulative general knowledge. But all of the studies share one thing
+in common: they all measure participant opinions on an ordinal scale. We start
+with a non-technical review of how ordinal opinion measurements work in group
+polarization, and then how the distributions of measurements are analyzed to
+determine if group opinions became polarized. Then we can explain our
+close-reading analysis to extract the necessary experimental design parameters
+we need to simulate all the experiments in our corpus.
 
-%   \centering 
-%   \rowcolors{2}{white}{gray!7}
-%   \begin{tabular}{m{1.35in} m{2.45in}
-%     m{1.3in} m{0.85in}}
+The corpus contains
+sixty different experimental conditions in the ten corpus journal articles. We 
+simulate experiments for all 57 experimental conditions reported to have 
+replicated group polarization, out of 60 total conditions (three were
+null control conditions).
 
-%     \toprule
+Using simulations we will estimate the number of these 57 replications that
+must be retracted because null-polarization simulations appear polarized due
+to measurement artifacts. We collect experiment design parameters that fill out 
+details of the simulation model of group polarization experiments.
 
-%     \textbf{Authors (Year)}& \textbf{Title} & \textbf{Journal} &
-%     \textbf{Category} \\
+\subsection{Ordinal Opinion Measurements: Problems and Simulations}
 
-%     \midrule 
-%     Moscovici \& Zavalloni (1969) & The Group as a Polarizer of
-%     Attitudes & \emph{JPSP} & Politics \\ 
-%     Schkade, Sunstein, \& Hastie (2010) & When
-%     Deliberation Produces Extremism & \emph{Critical~Review} & Politics \\
+In studies of group polarization, opinion measurements are typically ordinal,
+reflecting ordered positions rather than measurable distances between them—
+Moscovici and Zavalloni (1969) gave participants a seven-point scale, from
+“strongly disagree” (-3) to “strongly agree” (+3) to report their opinions.
+Burnstein and Vinokur (1975) used a ten-point scale from 1 to 10 representing
+“the lowest odds of success acceptable” to try a certain behavior, where 1
+means 10\%.  In these and every other survey-based group polarization study I
+know, the reported opinions on the ordinal scale—with discrete numbered
+bins—were treated as if they were continuous-valued, or in other words as if
+they were the internal mental opinion value itself, whatever that might be
+neurobiologically.
 
-%     Bishop \& Myers (1970) & Discussion Effects on Racial Attitudes &
-%     \emph{Science} & Social Comp's \\ 
-%     Myers (1975) & Discussion-Induced Attitude Polarization & 
-%     \emph{Human Relations} & Social Comp's \\
+Note that ordinal scales clip all latent opinions with a magnitude greater
+than the magnitude of its extreme values. This can make null-polarization
+outcomes look like group polarization. Consider the following example 
+with a group of just two people to see why: say they start with latent
+opinions of 2.1 and 4.7, which they report as +2 and +3, let's say, on Moscovici
+and Zavalloni's scale that tops out at +3. If they then discuss the topic and their
+opinions move toward one another by 1 each, then their latent opinions would be 3.1
+and 3.7. Before and after discussion the latent mean was a constant 3.4. 
+When reported on an ordinal scale, both opinions are now +3—the average measured
+opinion \emph{spuriously increased} from 2.75 to 3.0!
 
-%     Burnstein \& Vinokur (1973) & Testing Two Classes of Theories About Group
-%     Induced Shifts In Individual Choice & \emph{JESP} & Persuasive Arg's \\
-%     Burnstein \& Vinokur (1975) & What a person thinks upon learning he has
-%     chosen differently from others: Nice evidence for the
-%     persuasive-arguments\ldots & \emph{JESP} & Persuasive Arg's \\
+To determine whether a replication must be retracted, we developed a model to
+simulate measurements of null-polarizations, labelled \emph{mere agreement}, where
+the mean group opinion stays constant, but variance decreases. In other words,
+simulated participants report more agreement following discussion, 
+but they do not become radicalized.
 
-%     Abrams, et al., (1990) & Knowing What to Think by Knowing Who You Are:
-%     Self‐Categorization and the Nature of Norm Formation\ldots & 
-%     \emph{Brit. J. of Soc. Psych.} & Self-Categ'n \\
-%     Hogg, Turner, \& Davidson (1990) & Polarized Norms and Social Frames of 
-%     Reference: A Test of the Self-Categorization Theory of Group 
-%     Polarization & \emph{B. \& Appl. Soc. Psych.} & Self-Categ'n \\
-%     Krizan \& Baron (2007) & Group Polarization and Choice Dilemmas: How
-%     Important is Self-Categorization? & \emph{Eur. J. of Soc. Psych.} &
-%     Self-Categ'n \\
+The simulations are seeded by actual group polarization experiment parameters.
+We copied the pre- and post-discussion opinion means reported in each article.
+There is no original data available for any of the experiments. Sometimes more
+information was available about the pre- and post-discussion
+distributions—we used this information to decide whether replications plausibly 
+have null-polarization. We recorded all available data with a guide to finding it in
+its original article in the file `StudiesAnalysis.csv`, available with the
+Supplemental Information or to view/download on
+GitHub\footnote{\url{https://github.com/mt-digital/gp-statmod/GroupPolarizationStatmod/data/StudiesAnalysis.csv}}.
 
-%     Friedkin (1999) & Choice Shifts \& Group Polarization & 
-%     \emph{Amer. Soc.  Rev.} & Soc'l Decisions \\
+\subsection{Research Overview}
 
-% \bottomrule
-% \end{tabular}
-% \end{table}
-\begin{table}[ht]
-\centering
-\caption{\textbf{Journal articles in our corpus.}}
-\label{tab:articles}
+We collected original data across 10 journal articles to show that 95\% of 
+the replications reported must be retracted due to non-identifiability with a
+null-polarization model. These replications cannot be salvaged because the
+original data is no longer available, and thus cannot be re-analyzed with a 
+better statistical model that properly accounts for the measurement procedure.
 
-\begin{tabular}{m{0.9in} m{1.35in} m{2.45in} m{1.3in}}
-\toprule
-\textbf{Category} & \textbf{Authors (Year)} & \textbf{Title} & \textbf{Journal} \\
-\midrule
+In the Analysis we go further: we use simulations to show that 
+the experimental designs in our corpus are beset by high false positive rates 
+\emph{even with a better statistical model}. The only way to reduce false
+positive rates with the experimental designs used in the papers is to increase
+the significance threshold (in terms of Cohen's $d$ as explained in more
+detail in the Methods).
 
-\rowcolor{gray!10}
-\multirowcell{2}{Politics} 
-  & Moscovici \& Zavalloni (1969) & The Group as a Polarizer of Attitudes & \emph{JPSP} \\
-\rowcolor{gray!10}
-  & Schkade, Sunstein, \& Hastie (2010) & When Deliberation Produces Extremism & \emph{Critical~Review} \\[0.2em]
+Group polarization was taken as one of the most reliable group effects in social
+science—but we showed it is impossible to distinguish 95\% of group
+polarization detections from a null-polarization model. This means 54
+replications must be retracted—and recall that this was a selection of the
+highest quality, most representative group polarization research. 
+To understand group polarization, new
+methods will be required.
 
-\rowcolor{gray!5}
-\multirowcell{2}{Social Comp's \vspace{0.7em}}
-% \multirowcell{2}{Social Comp's}
-  & Bishop \& Myers (1970) & Discussion Effects on Racial Attitudes & \emph{Science} \\[0.2em]
-\rowcolor{gray!5}
-  & Myers (1975) & Discussion-Induced Attitude Polarization & \emph{Human Relations} \\[0.3em]
-
-\rowcolor{gray!10}
-\multirowcell{2}{Persuasive Arg's}
-  & Burnstein \& Vinokur (1973) & Testing Two Classes of Theories About Group Induced Shifts In Individual Choice & \emph{JESP} \\
-\rowcolor{gray!10}
-  & Burnstein \& Vinokur (1975) & What a person thinks upon learning he has chosen differently from others: Nice evidence for the persuasive-arguments\ldots & \emph{JESP} \\
-
-\rowcolor{gray!5}
-\multirowcell{3}{Self-Categ'n}
-  & Abrams, et al., (1990) & Knowing What to Think by Knowing Who You Are: Self‐Categorization and the Nature of Norm Formation\ldots & \emph{Brit. J. of Soc. Psy.} \\
-\rowcolor{gray!5}
-  & Hogg, Turner, \& Davidson (1990) & Polarized Norms and Social Frames of Reference: A Test of the Self-Categorization Theory of Group Polarization & \emph{B. \& Appl. Soc. Psy.} \\
-\rowcolor{gray!5}
-  & Krizan \& Baron (2007) & Group Polarization and Choice Dilemmas: How Important is Self-Categorization? & \emph{Eur. J. of Soc. Psy.} \\
-
-\rowcolor{gray!10}
-\multirowcell{1}{Soc'l Decisions}
-  & Friedkin (1999) & Choice Shifts \& Group Polarization & \emph{Amer. Soc. Rev.} \\
-
-\bottomrule
-\end{tabular}
-\end{table}
-
-\subsection{Measurement procedure}
-
-In studies of group polarization, opinion measurements are typically ordinal, reflecting
-ordered positions rather than measurable distances between them—
-Moscovici and Zavalloni (1969) gave participants a seven-point scale, from “strongly disagree” (-3) to “strongly agree” (+3) to report their opinions.
-Burnstein and Vinokur (1975) used a ten-point
-scale from 1 to 10 representing “the lowest odds of success acceptable” to
-try a certain behavior, where 1 means 10\%.
-In these and every other survey-based group polarization study I know, the reported opinions on the ordinal scale—with discrete numbered bins—were treated as if they were continuous-valued, or in other words as if they were the internal mental opinion value itself, whatever that might be neurobiologically.
-
-Note how ordinal scales clip all latent opinions with a magnitude greater than the
-magnitude of its extreme values. 
-lead to false positives because ordinal scales cannot track changes in
-extreme opinions beyond the ceiling of the measurement scale—this approach is
-ubiquitous in social psychology and survey research, not at all limited to
-group polarization.
+This sort of error threatens our ability to construct a
+cumulative science of opinion change, social influence, and social change
+generally. We need reliable social science as the foundation for new
+strategies for social organization capable of reversing decline and powering
+sustainable futures. We have a long way to go.

### 2025-10-29 GroupPolarizationStatmod/paper/ms.tex
finished corpus table
tables and mostly put in new symbology

@@
   abstract]{article}
+
+\usepackage{floatrow}
+\DeclareFloatFont{small}{\small}% "scriptsize" is defined by floatrow, "tiny" not
+\floatsetup[table]{font=small}
+
+\usepackage[table]{xcolor}
+\usepackage{multirow}
+\usepackage{makecell}
+\usepackage{array}
 
@@
 
+\begin{table}[h]
+  \caption{\textbf{Latent psychological variables and distributions.}}
+  \begin{tabular}{cll} 
+  \toprule
+    \textbf{Variable} & \textbf{Description} & \textbf{Values} \\
+  \midrule  
+    $\omega_{i,t}$ & Latent opinion of individual $i$ at time $t$ & $\mathbb{R}$ \\
+    $\mu_t$ & Mean latent opinion at time $t$ (constant in null-polarization) & $\mathbb{R}$ \\
+    $\sigma_t$ & Variance of observed opnions at time $t$ & $\mathbb{R}$ \\
+  \bottomrule
+  \end{tabular} 
+\end{table}
+
+\begin{table}[h]
+  \caption{\textbf{Experimental design and numerical model parameters.}}
+  \begin{tabular}{cll} 
+  \toprule
+    \textbf{Variable} & \textbf{Description} & \textbf{Values} \\
+  \midrule  
+    $B$ & Number of bins & $5,6,7,\ldots$ \\
+    $\theta_b$ & The $B+1$ threshold values that separate bins & 
+      e.g., $-\infty, -,\ldots$ \\
+    $b$ & Bin index & $0,1,2,\ldots$ \\
+  \bottomrule
+  \end{tabular} 
+\end{table}
+
+\begin{table}[h]
+  \caption{\textbf{Simulated observed variables.}}
+  \begin{tabular}{cll} 
+  \toprule
+    \textbf{Variable} & \textbf{Description} & \textbf{Values} \\
+  \midrule  
+    $\meanobst$ & Mean of all reported opinions (Eq. \ref{eq:meanobs}) & $\mathbb{R}$ \\
+    $g$ & Observed group opinion polarization (Eq. \ref{eq:group-polarization})
+        & $\mathbb{R}$ \\
+    $d$ & Cohen's $d$ used as significance threshold 
+        & $\mathbb{R}$ \\
+    $\alpha_e,\alpha_s,\alpha_{\text{all}}$ & Family-wise/Type I error rate
+    for an experiment, study, or across all experiments
+             & $[0, 1]$ \\
+    FDR & False discovery rate
+        & $[0, 1]$ \\
+  \bottomrule
+  \end{tabular} 
+\end{table}
+
 \subsection{Probability theory of spurious group polarization}
@@
 \begin{equation} 
-  o_{i,t} \sim p(o; \mu_t, \sigma_t) = \mathcal{N}(\mu_t, \sigma_t).
+  \omega_{i,t} \sim p(\omega; \mu_t, \sigma_t) = \mathcal{N}(\mu_t, \sigma_t).
   \label{eq:opinionDistribution} 
@@
 of which there are $B+1$, where $B$ is the number of opinion bins, with sequential
-integer values $b$. In the Moscovici and Zavalloni (1969) example, $b \in
-\{-3,-2,\ldots,2,3\}$ and $B = 7$.
+integer values $b \in \{0, B\}$. In the Moscovici and Zavalloni (1969)
+example, the thresholds are \in
+\{-\infty,-2.5,-1.5,\ldots,2.5,\infty\}$, with $B = 7$.
 Bin $b$ is defined by its lower and upper continuous-valued thresholds, 
@@
 $\theta_{b}$, respectively. Thresholds are given by the following equations
-(and visualized as dotted lines in Figure~\ref{fig:distros}).
+(visualized as dotted lines in Figure~\ref{fig:distros}).
 \begin{equation}
@@
   \infty          & \text{ if } b=B \\
-  b + \frac{1}{2} & \text{ otherwise.}
+  \theta_{\text{Base}} + \frac{b}{2} & \text{ otherwise.}
 \end{cases}
@@
 \noindent
+In the Moscovici and Zavalloni example, the base bin value is
+$\theta_{\text{Base}} = -3$.
 With thresholds defined, the frequency of ordinal measurment $b$ at time $t$ 
@@
 \begin{equation}
-  p(b;~\mu_t, \sigma_t, B) = \int_{\theta_{b-1}}^{\theta_b} p(o; \mu_t, \sigma_t)
-  do.
+  p(b;~\mu_t, \sigma_t, B) = \int_{\theta_{b-1}}^{\theta_b} p(\omega; \mu_t, \sigma_t)
+  d\omega.
   \label{eq:ordinal-frequency}
@@
 
-We can now define spurious group polarization and give examples of when it occurs.
-First, the observed group polarization opinion shift is
-\begin{equation}
-  g = \meanobs_\mathrm{post} - \meanobs_\mathrm{pre}~,
-  \label{eq:group-polarization}
-\end{equation}
-\noindent
-where the expected value of observed opinions at time $t$ is
+We can now define spurious group polarization and give examples of when it
+occurs. First, the expected value of observed opinions at time $t$ is
 \begin{equation}
@@
 \noindent
+The observed group polarization opinion shift is then
+\begin{equation}
+  g = \meanobs_\mathrm{post} - \meanobs_\mathrm{pre}.
+  \label{eq:group-polarization}
+\end{equation}
+\noindent
 
-\emph{Spurious} group polarization is when mean observed opinions seem to increase in 
-extremity when the latent mean is held constant. Call the collection of
-latent null model distribution parameters $\lambda = (\mu, \sigma_{pre}, \sigma_{post})$.
-The group polarization opinion shift can then be written as a function of $\lambda$,
-$g(\lambda)$. Formally, then, spurious group polarization is plausible if 
-there exists at least one $\lambda$ for which $g(\lambda) = g_{pub} > 0$, where
+Measurement artifacts can make null-polarization spuriously appear like group
+polarization. To see how, we will build up a formal model of the
+process. Consider a collection of
+null-polarization latent opinion distribution parameters $\nu = (\mu, \sigma_{pre}, \sigma_{post})$.
+The group polarization opinion shift can then be written as a function of
+$\nu$,
+$g(\nu)$.
+
+If we can find ..... $\nu$ for which $g(\nu) = g_{pub} > 0$, where
 $g_{pub}$ is the published group polarization shift value.
@@
 
-Next, we explain how to find a set of latent null parameters, $\lambda$, that
+Next, we explain how to find a set of latent null parameters, $\nu$, that
 generate spurious group polarization when naively measured on an ordinal scale.
@@
 To identify plausibly false detections of group polarization, we searched numerically
-for latent parameters $\lambda = (\mu, \sigmapre, and \sigmapost)$ that generate
+for latent parameters $\nu = (\mu, \sigmapre, and \sigmapost)$ that generate
 spurious group polarization, as explained immediately above. We developed an *ad hoc*
-hillclimbing-style algorithm that finds $\lambda$ that finds $\lambda$ by minimizing
+hillclimbing-style algorithm that finds $\nu$ that finds $\nu$ by minimizing
 the difference between published group polarization shifts $g_pub$ and calculated shifts,
-$g_\lambda$.
+$g_\nu$.
 calculated with Equation~\ref{eq:meanobs}.
 
-The algorithm to find $\lambda$ holds $\mu$ constant while searching for $\sigmapre$
+The algorithm to find $\nu$ holds $\mu$ constant while searching for $\sigmapre$
 and $\sigmapost$ that generate observed $\meanobspre$ and $\meanobspost$, within a
@@
 The algorithm requires a researcher to specify an initial
-guess for $\lambda$ and auxiliary parameter settings, which generally vary between
+guess for $\nu$ and auxiliary parameter settings, which generally vary between
 experimental conditions.  Furthermore, the algorithm does not always produce
-reasonable distributions with a given $\lambda$, and sometimes it does not produce
+reasonable distributions with a given $\nu$, and sometimes it does not produce
 spurious group polarization.  Therefore, it is necessary to sanity check the latent
-and ordinal opinion distributions for the $\lambda$ found by the algorithm. If all is
+and ordinal opinion distributions for the $\nu$ found by the algorithm. If all is
 well, the experimental condition can be marked as plausibly spurious.  To tame the
@@
 some suspicion. If there exists set of latent simple consensus 
-parameters $\lambda$ that create binned (Equation~\ref{eq:ordinal-frequency})
+parameters $\nu$ that create binned (Equation~\ref{eq:ordinal-frequency})
 opinion distributions that naively seem to be group polarized, then the means
@@
 distributions that generated spurious group polarization, i.e., with
-parameters labelled $\lambda_e$ found in the hillclimbing step. Binning these
+parameters labelled $\nu_e$ found in the hillclimbing step. Binning these
 random values simulates observations. $N_e$ is the number
@@
 % group polarization in 
-Of the remaining 57, we identified simple consensus $\lambda$ that generate spurious
+Of the remaining 57, we identified simple consensus $\nu$ that generate spurious
 group polarization in 54 experimental conditions. In the other three, we did
-identify $\lambda$ that generate spurious group polarization, but the initial
+identify $\nu$ that generate spurious group polarization, but the initial
 simulated distributions were highly polarized. 
-We rejected $\lambda$ generated for
+We rejected $\nu$ generated for
 two spurious group polarization from Schkade, et al., (2010) because
@@
 polarized~\cite[Figure 1, p. 234]{Schkade2010}. The other condition where we
-found but rejected spurious group polarization $\lambda$ was in a question 
+found but rejected spurious group polarization $\nu$ was in a question 
 about a ``bad teacher'' in Myers (1975)~\cite{Myers1975a}, where lower values

### 2025-10-30 GroupPolarizationStatmod/paper/intro-latest.tex
intro updates

--- 2025-10-24 GroupPolarizationStatmod/paper/intro-latest.tex
@@
 
-If an extremist’s opinion falls and it is measured with a Likert scale, will
-it make a noise? The answer is often no~\cite{Liddell2018,Turner2020}. This
-means a group’s opinions can appear to radicalize over time even when no real
-shift occurs. The problem matters because Likert and other ordinal scales are
-ubiquitous in research on opinion change and radicalization across the social
-and cognitive sciences~\cite{Liddell2018}. One influential framework in this
-area is group polarization—the idea that discussion among like-minded people
-pushes opinions toward extremes, producing ``echo chambers.'' Cass Sunstein, a
+If an extremist moderates their opinion and it is measured with a Likert scale, will
+this move to moderation be detected? The answer is often no in social psychology 
+experiments~\cite{Liddell2018}. This
+means that a group can appear to become radicalized over time even if the 
+average opinion does not actually change~\cite{Turner2020}. 
+This has huge consequences for the study of \emph{group polarization}, the hypothesis
+that discussion among like-minded people pushes opinions toward extremes—effectively 
+making it the science of ``echo chambers'', online or in-person communities 
+composed of likeminded folks.  Cass Sunstein, a
 Harvard Law professor and former White House administrator, has written
 extensively on the topic~\cite{Sunstein2019}, claiming that group
-polarization is a scientific law—a regular, quantitative relationship that
-will always be observed when certain assumptions are
-met~\cite{Cartwright1999}. Yet Roger Brown himself noted in Social Psychology
-that group polarization ``does not occur with every group'' and that ``the
-effect is not large.'' Given the influence of this theory and the hundreds of
-claimed ``replications,'' it is essential to ask whether many of these findings
-may instead be measurement artifacts—spurious effects created when the
-``neutral'' point of the measurement scale fails to align with the group’s own
-center of opinion.
+polarization is a scientific law—but Roger Brown (1986, Ch. 5) wrote that group polarization ``does not occur with every group'' and that ``the
+effect is not large,'' conflicting with the standard definition that lawlike
+behavior means certain observations will obtain given that specific assumptions
+are satisfied~\cite{Cartwright1999}. Given the prominence of group polarization 
+theory, we decided to analyze whether its hundreds of claimed replications
+may instead be measurement artifacts using a formal model and numerical simulations of
+group polarization experiments.
 
@@
 
-We are sad to report that 95\% of experimental reports of group polarization
+We found that 95\% of experimental reports of group polarization
 across sixty published experimental conditions cannot be counted as
@@
 \centering
-\begin{tabular}{clcc}
+\begin{tabular}{p{1in}p{3.5in}p{2in}c}
 \toprule

### 2025-10-30 GroupPolarizationStatmod/paper/ms.tex
intro updates

--- 2025-10-23 GroupPolarizationStatmod/paper/ms.tex
@@
 
-\begin{quote}
-In our introductory social psychology course, 
-we have for many years used the [group polarization experimental paradigm] as
-a laboratory exercise. The exercise works beautifully, but one must be
-careful to forewarn a class that [group polarization] does not occur with every 
-group\ldots and that the effect is not large. 
-\par\raggedleft(Brown, 1986, p.\cite[p. 205]{Brown1986})
-\end{quote}
+% \begin{quote}
+% In our introductory social psychology course, 
+% we have for many years used the [group polarization experimental paradigm] as
+% a laboratory exercise. The exercise works beautifully, but one must be
+% careful to forewarn a class that [group polarization] does not occur with every 
+% group\ldots and that the effect is not large. 
+% \par\raggedleft(Brown, 1986, p.\cite[p. 205]{Brown1986})
+% \end{quote}
 
-\begin{quote}
-  The good thing about science is that it's true, whether or not you believe
-  it. (Neil de Grasse Tyson slogan circa 2018)
-\end{quote}
+% \begin{quote}
+%   The good thing about science is that it's true, whether or not you believe
+  % it. (Neil de Grasse Tyson slogan circa 2018)
+% \end{quote}
 

### 2025-10-30 GroupPolarizationStatmod/paper/article-table.tex
added article table

\begin{table}[ht]
\centering
\caption{\textbf{Journal articles in our corpus.}}
\label{tab:articles}

\begin{tabular}{m{0.9in} m{1.35in} m{2.45in} m{1.3in}}
\toprule
\textbf{Category} & \textbf{Authors (Year)} & \textbf{Title} & \textbf{Journal} \\
\midrule

\rowcolor{gray!10}
\multirowcell{2}{Politics} 
                  & Moscovici \& Zavalloni (1969)~\cite{Moscovici1969} & The Group as a Polarizer of Attitudes & \emph{JPSP} \\
\rowcolor{gray!10}
                  & Schkade, Sunstein, \& Hastie (2010)~\cite{Schkade2010} & When Deliberation Produces Extremism & \emph{Critical~Review} \\[0.2em]

\rowcolor{gray!5}
\multirowcell{2}{Social Comp's \vspace{0.7em}}
% \multirowcell{2}{Social Comp's}
                  & Myers \& Bishop (1970)~\cite{Myers1970} & Discussion Effects on Racial Attitudes & \emph{Science} \\[0.2em]
\rowcolor{gray!5}
                  & Myers (1975)~\cite{Myers1975} & Discussion-Induced Attitude Polarization & \emph{Human Relations} \\[0.3em]

\rowcolor{gray!10}
\multirowcell{2}{Persuasive Arg's}
                  & Burnstein \& Vinokur (1973)~\cite{Burnstein1973} & Testing Two Classes of Theories About Group Induced Shifts In Individual Choice & \emph{JESP} \\
\rowcolor{gray!10}
                  & Burnstein \& Vinokur (1975)~\cite{Burnstein1975} & What a person thinks upon learning he has chosen differently from others: Nice evidence for the persuasive-arguments\ldots & \emph{JESP} \\

\rowcolor{gray!5}
\multirowcell{3}{Self-Categ'n}
                  & Abrams, et al., (1990)~\cite{Abrams1990} & Knowing What to Think by Knowing Who You Are: Self‐Categorization and the Nature of Norm Formation\ldots & \emph{Brit. J. of Soc. Psy.} \\
\rowcolor{gray!5}
                  & Hogg, Turner, \& Davidson (1990)~\cite{Hogg1990} & Polarized Norms and Social Frames of Reference: A Test of the Self-Categorization Theory of Group Polarization & \emph{B. \& Appl. Soc. Psy.} \\
\rowcolor{gray!5}
                  & Krizan \& Baron (2007)~\cite{Krizan2007} & Group Polarization and Choice Dilemmas: How Important is Self-Categorization? & \emph{Eur. J. of Soc. Psy.} \\

\rowcolor{gray!10}
\multirowcell{1}{Soc'l Decisions}
                  & Friedkin (1999)~\cite{Friedkin1999} & Choice Shifts \& Group Polarization & \emph{Amer. Soc. Rev.} \\

\bottomrule
\end{tabular}
\end{table}

### 2025-10-30 GroupPolarizationStatmod/paper/ms.tex
fixed math problem

--- 2025-10-29 GroupPolarizationStatmod/paper/ms.tex
@@
 of which there are $B+1$, where $B$ is the number of opinion bins, with sequential
-integer values $b \in \{0, B\}$. In the Moscovici and Zavalloni (1969)
-example, the thresholds are \in
-\{-\infty,-2.5,-1.5,\ldots,2.5,\infty\}$, with $B = 7$.
+integer values $b\{0, B\}$. In the Moscovici and Zavalloni (1969)
+example, the thresholds are $\theta_b \in \{-\infty,-2.5,-1.5,\ldots,2.5,\infty\}$, with $B = 7$.
 Bin $b$ is defined by its lower and upper continuous-valued thresholds, 
-$\theta_{b-1}$ and
+% $\theta_{b-1}$ and
 $\theta_{b}$, respectively. Thresholds are given by the following equations

### 2025-10-30 GroupPolarizationStatmod/paper/analysis.tex
on to method

--- 2025-10-30 GroupPolarizationStatmod/paper/ms.tex
@@
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
-\PassOptionsToPackage{hyphens}{url}
-\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
-%
-\documentclass[
-  abstract]{article}
-
-\usepackage{floatrow}
-\DeclareFloatFont{small}{\small}% "scriptsize" is defined by floatrow, "tiny" not
-\floatsetup[table]{font=small}
-
-\usepackage[table]{xcolor}
-\usepackage{multirow}
-\usepackage{makecell}
-\usepackage{array}
-
-\usepackage{amsmath,amssymb}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-\usepackage{iftex}
-\input{preamble.tex}
-\hypersetup{
-  pdftitle={Group polarization replications may be marred by high false discovery rates},
-  colorlinks=true,
-  linkcolor={blue},
-  filecolor={Maroon},
-  citecolor={Blue},
-  urlcolor={Blue}}
-
-\newcommand{\mupre}{\mu_\mathrm{pre}}
-\newcommand{\mupost}{\mu_\mathrm{post}}
-
-\newcommand{\sigmapre}{\sigma_\mathrm{pre}}
-\newcommand{\sigmapost}{\sigma_\mathrm{post}}
-
-\newcommand{\thetafit}{\theta^\mathrm{fit}}
-
-\newcommand{\fdr}{\mathrm{FDR}}
-
-\newcommand{\meanobs}{\bar{o}}
-\newcommand{\meanobsemp}{\bar{\mathbf{o}}}
-\newcommand{\meanobst}{\meanobs_t}
-\newcommand{\meanobsjt}{\meanobs_{j,t}}
-
-\newcommand{\meanobspre}{\meanobs_\mathrm{pre}}
-\newcommand{\meanobspost}{\meanobs_\mathrm{post}}
-
-\definecolor{myorange}{RGB}{240, 96, 0}
-\newcommand{\mt}[1]{{\textcolor{myorange} {({\tiny MT:} #1)}}}
-
-\usepackage{authblk}
-\usepackage{etoolbox}
-\usepackage{wrapfig}
-\makeatletter
-% \title{If the Null Fits, You Must Omit: Ubiquitous False Detections of Group Polarization}
-\title{If the Null Fits, You Must Omit: Measurement Artifacts Undermine Group Polarization Replications}
-\makeatother
-\author[1,*]{{Matthew A.~Turner}}
-\affil[1]{\small Environmental Social Sciences, Stanford Doerr School of Sustainability, Stanford University}
-
-\author[2,3]{{Paul E.~Smaldino}}
-\affil[2]{\small Cognitive and Information Sciences, University of California, Merced}
-\affil[3]{\small Santa Fe Institute} 
-% \vspace{2em}
-\affil[*]{\small Correspondence: \href{mailto:maturner@stanford.edu}{maturner@stanford.edu}}
-
-\date{\today}
-\begin{document}
-\maketitle
-
-\renewcommand*\contentsname{Table of contents}
-{
-\hypersetup{linkcolor=gray}
-\setcounter{tocdepth}{2}
-\tableofcontents
-}
-
-\begin{center} \noindent\rule{4cm}{0.4pt} \end{center}
-% \clearpage
-
-\begin{quote}
-In our introductory social psychology course, 
-we have for many years used the [group polarization experimental paradigm] as
-a laboratory exercise. The exercise works beautifully, but one must be
-careful to forewarn a class that [group polarization] does not occur with every 
-group\ldots and that the effect is not large. 
-\par\raggedleft(Brown, 1986, p.\cite[p. 205]{Brown1986})
-\end{quote}
-
-\begin{quote}
-  The good thing about science is that it's true, whether or not you believe
-  it. (Neil de Grasse Tyson slogan circa 2018)
-\end{quote}
-
-\input{intro-latest.tex}
-
-% \input{old-intro.tex}
-
-\section{Methods}\label{methods}
-
-To find examples of plausibly spurious group polarization we guess at what latent
-mean, combined with a latent pre- and post-deliberation latent standard deviations,
-could generate it. We calculate the observed group polarization for these latent
-parameters by transforming the latent pre- and post-deliberation distributions from
-continuous normal distributions to ordinal ``observations'' by integrating over each
-ordinal bin, then calculating the observed mean for each experiment \emph{phase},
-i.e., both the pre- and post-deliberation phases.  If the null model can generate
-group polarization in this case, then the detection is labelled \emph{plausibly
-false}, i.e., \emph{plausibly spurious}, and we call the generating latent
-parameters the \emph{plausible null parameters}.
-
-To evaluate whether the plausible null parameters we identify could realistically
-fool a significance test, we implemented a generative simulation of each plausibly
-false detection of group polarization and calculated the \emph{family-wise error rate
-(FWER)} and the \emph{false discovery rate (FDR)}. We use the published number of
-participants $N$ for each of the 54 conditions that yielded plausibly false
-detections of group polarization. The FWER is also known as the Type I
-error, or the false positive rate: it is the probability that a detection of an
-effect occurs given that in reality there was a null effect.  The FDR is the
-probability that a published positive detection of an effect is a false positive,
-i.e., a detection of a null effect. High FWER and FDR for an experimental condition
-indicates that its design is weak, since we identify only one of perhaps infinitely
-many latent parameter combinations that can generate spurious group polarization.
-
-We use this setup to demonstrate how \emph{adversarial} style testing of experimental
-designs with counterfactual null models can be used to estimate important
-metascience parameters. In this study we use our FWER calculations to estimate
-the minimum Cohen's $d$ that should be used to determine whether an effect is
-``significant'' or not that will limit the FWER to 5\%. In the Discussion we
-sketch how our model can then be extended to rigorously calculate statistical power,
-i.e., the probability that a significant effect is detected given that there 
-is indeed a significant effect.
-
-\begin{table}[h]
-  \caption{\textbf{Latent psychological variables and distributions.}}
-  \begin{tabular}{cll} 
-  \toprule
-    \textbf{Variable} & \textbf{Description} & \textbf{Values} \\
-  \midrule  
-    $\omega_{i,t}$ & Latent opinion of individual $i$ at time $t$ & $\mathbb{R}$ \\
-    $\mu_t$ & Mean latent opinion at time $t$ (constant in null-polarization) & $\mathbb{R}$ \\
-    $\sigma_t$ & Variance of observed opnions at time $t$ & $\mathbb{R}$ \\
-  \bottomrule
-  \end{tabular} 
-\end{table}
-
-\begin{table}[h]
-  \caption{\textbf{Experimental design and numerical model parameters.}}
-  \begin{tabular}{cll} 
-  \toprule
-    \textbf{Variable} & \textbf{Description} & \textbf{Values} \\
-  \midrule  
-    $B$ & Number of bins & $5,6,7,\ldots$ \\
-    $\theta_b$ & The $B+1$ threshold values that separate bins & 
-      e.g., $-\infty, -,\ldots$ \\
-    $b$ & Bin index & $0,1,2,\ldots$ \\
-  \bottomrule
-  \end{tabular} 
-\end{table}
-
-\begin{table}[h]
-  \caption{\textbf{Simulated observed variables.}}
-  \begin{tabular}{cll} 
-  \toprule
-    \textbf{Variable} & \textbf{Description} & \textbf{Values} \\
-  \midrule  
-    $\meanobst$ & Mean of all reported opinions (Eq. \ref{eq:meanobs}) & $\mathbb{R}$ \\
-    $g$ & Observed group opinion polarization (Eq. \ref{eq:group-polarization})
-        & $\mathbb{R}$ \\
-    $d$ & Cohen's $d$ used as significance threshold 
-        & $\mathbb{R}$ \\
-    $\alpha_e,\alpha_s,\alpha_{\text{all}}$ & Family-wise/Type I error rate
-    for an experiment, study, or across all experiments
-             & $[0, 1]$ \\
-    FDR & False discovery rate
-        & $[0, 1]$ \\
-  \bottomrule
-  \end{tabular} 
-\end{table}
-
-\subsection{Probability theory of spurious group polarization}
-
-We formalize this experimental design in the $N\to\infty$ case where the
-distribution of ordinal measurements is created by integrating continuous latent
-opinion distributions.  Latent opinions in the model are normally distributed with
-mean $\mu_t$ and variance $\sigma_t$ for time $t$. We are concerned with only two
-times, \emph{pre-} and \emph{post-} deliberation ($t=0$ and $t=T$, respectively).
-For short we can write $\mu_\mathrm{pre}$ and $\sigma_\mathrm{pre}$ for
-pre-deliberation mean and variance (and similar for \emph{post}-deliberation).  
-The latent opinion of participant $i$ at time $t$ is drawn from a normal 
-distribution in the model, 
-\begin{equation} 
-  \omega_{i,t} \sim p(\omega; \mu_t, \sigma_t) = \mathcal{N}(\mu_t, \sigma_t).
-  \label{eq:opinionDistribution} 
-\end{equation} 
-\noindent 
-Latent opinions are assumed to be normally distributed since they are unbounded 
-aggregations of neurobiological activity.
-
-The measurement process with $N\to\infty$ participants is simulated by 
-integrating the latent opinion distribution between ordinal bin thresholds, 
-of which there are $B+1$, where $B$ is the number of opinion bins, with sequential
-integer values $b\{0, B\}$. In the Moscovici and Zavalloni (1969)
-example, the thresholds are $\theta_b \in \{-\infty,-2.5,-1.5,\ldots,2.5,\infty\}$, with $B = 7$.
-Bin $b$ is defined by its lower and upper continuous-valued thresholds, 
-% $\theta_{b-1}$ and
-$\theta_{b}$, respectively. Thresholds are given by the following equations
-(visualized as dotted lines in Figure~\ref{fig:distros}).
-\begin{equation}
-\theta_b = \begin{cases}
-  -\infty         & \text{ if } b=0 \\
-  \infty          & \text{ if } b=B \\
-  \theta_{\text{Base}} + \frac{b}{2} & \text{ otherwise.}
-\end{cases}
-\end{equation}
-\noindent
-In the Moscovici and Zavalloni example, the base bin value is
-$\theta_{\text{Base}} = -3$.
-With thresholds defined, the frequency of ordinal measurment $b$ at time $t$ 
-is (Figure~\ref{fig:distros})
-\begin{equation}
-  p(b;~\mu_t, \sigma_t, B) = \int_{\theta_{b-1}}^{\theta_b} p(\omega; \mu_t, \sigma_t)
-  d\omega.
-  \label{eq:ordinal-frequency}
-\end{equation}
-
-We can now define spurious group polarization and give examples of when it
-occurs. First, the expected value of observed opinions at time $t$ is
-\begin{equation}
-  \meanobst = \frac{1}{B} \sum_{b=1}^B b \cdot p(b; \mu_t, \sigma_t, B).
-  \label{eq:meanobs}
-\end{equation}
-\noindent
-The observed group polarization opinion shift is then
-\begin{equation}
-  g = \meanobs_\mathrm{post} - \meanobs_\mathrm{pre}.
-  \label{eq:group-polarization}
-\end{equation}
-\noindent
-
-Measurement artifacts can make null-polarization spuriously appear like group
-polarization. To see how, we will build up a formal model of the
-process. Consider a collection of
-null-polarization latent opinion distribution parameters $\nu = (\mu, \sigma_{pre}, \sigma_{post})$.
-The group polarization opinion shift can then be written as a function of
-$\nu$,
-$g(\nu)$.
-
-If we can find ..... $\nu$ for which $g(\nu) = g_{pub} > 0$, where
-$g_{pub}$ is the published group polarization shift value.
-This could be thought of as sampling the observed ordinal distributions 
-$N \to \infty$, collecting ordinal measurements of samples from 
-each latent distribution corresponding to each phase, pre- and post-deliberation.
-
-Next, we explain how to find a set of latent null parameters, $\nu$, that
-generate spurious group polarization when naively measured on an ordinal scale.
-
-\begin{figure}
-  \caption{\textbf{Opinion measurement model and spurious group polarization.} 
-  Latent model participant opinions are drawn from a normal distribution with the
-  listed parameters (A,C). The pre-deliberation latent distribution has high
-  variance, but appears polarized when measured via integration (A).  The latent
-  distribution's variance is expected to decrease during deliberation (B) via
-  consensus, but in this example the latent mean is static, 
-$\mupre = \mupost$. However, when the post-deliberation distribution is measured,
-the simulated observed mean increases, i.e., spurious
-group polarization occurs (C).}
-  \centering
-  \includegraphics[width=0.95\textwidth]{Figures/Model/latent-ordinal-distros.pdf}
-  \label{fig:distros}
-\end{figure}
-
-\subsection{Interactive algorithm to find parameters inducing spurious group polarization}
-
-To identify plausibly false detections of group polarization, we searched numerically
-for latent parameters $\nu = (\mu, \sigmapre, and \sigmapost)$ that generate
-spurious group polarization, as explained immediately above. We developed an *ad hoc*
-hillclimbing-style algorithm that finds $\nu$ that finds $\nu$ by minimizing
-the difference between published group polarization shifts $g_pub$ and calculated shifts,
-$g_\nu$.
-calculated with Equation~\ref{eq:meanobs}.
-
-The algorithm to find $\nu$ holds $\mu$ constant while searching for $\sigmapre$
-and $\sigmapost$ that generate observed $\meanobspre$ and $\meanobspost$, within a
-tolerance of $\pm10^{-3}$ (Algorithm~\ref{alg:hillclimbing}). 
-The algorithm requires a researcher to specify an initial
-guess for $\nu$ and auxiliary parameter settings, which generally vary between
-experimental conditions.  Furthermore, the algorithm does not always produce
-reasonable distributions with a given $\nu$, and sometimes it does not produce
-spurious group polarization.  Therefore, it is necessary to sanity check the latent
-and ordinal opinion distributions for the $\nu$ found by the algorithm. If all is
-well, the experimental condition can be marked as plausibly spurious.  To tame the
-complexity involved in data entry and model fitting for each of these sixty
-experimental conditions, we developed a web application to input data from control
-the algorithm and inspect its results (Figure~\ref{fig:webapp}). Web application code is
-available online in our GitHub
-repository\footnote{\url{https://github.com/mt-digital/gp-statmod/blob/main/GroupPolarizationStatmod/app.R}}.
-
-\begin{algorithm}
-  \caption{Routine to find latent mean and variance that generate spurious group polarization.}
-  \label{alg:hillclimbing}
-  \begin{algorithmic}
-
-    \Require $\mu$, $\sigma$, $\bar{\mathbf{o}}$ 
-      \Comment{Set latent mean, guess variance that yields, input empirical mean.}
-
-    \Require $\epsilon$, $\delta$, $i_\mathrm{max}$ 
-      \Comment{Hillclimbing tolerance, step size, and maximum iterations.}
-
-    \State $\Delta \gets \meanobs(\mu, \sigma) - \meanobsemp$ 
-      \Comment{Initialize simulation error.}
-    \While {$\Delta^2 > \epsilon$ and $i < i_\mathrm{max}$} \Comment{Search until desired
-    accuracy or max iterations reached.}
-      \State $\sigma' \sim \mathcal{N}(\sigma, \delta)$ \Comment{Draw new variance to test.}
-      \State $\Delta' \gets \bar{o}(\mu, \sigma') - \bar{\mathbf{o}}$ 
-        \Comment{Calculate error using new variance.}
-      \If {$\Delta' - \Delta < 0$}
-        \State $\sigma \gets \sigma'$
-          \Comment{Update $\sigma$ if the error is less than before.}
-        \State $\Delta \gets \bar{o}(\mu, \sigma) - \bar{\mathbf{o}}$
-          \Comment{Update simulation error if $\sigma$ is updated.}
-      \EndIf
-    \EndWhile \\
-
-    \Return $\sigma$
-  \end{algorithmic}
-\end{algorithm}
-
-% \begin{figure}
-%   \caption{\textbf{Shiny web application for user-guided hillclimbing to find
-%   parameters that generate spurious group polarization.} This web application enables users to
-% create new analyses of experimental conditions from published studies. Users 
-% provide initial guesses for the algorithm to use as starting points in 
-% finding variances that generate spurious group polarization. Users inspect 
-% model fits provided by the algorithm, adjusting parameters if necessary, and
-% mark an experimental condition as ``Plausibly spurious'' if a reasonable fit
-% is indeed found.\mt{Maybe add some A, B, C guides to the figure and include
-% in parentheses here to go with the text?}}
-%   \centering
-%     \includegraphics[width=0.75\textwidth]{Figures/Model/WebApp.png}
-%   \label{fig:webapp}
-% \end{figure}
-
-Note that so far, we have only identified cases where spurious group polarization plausibly
-occurs in the asymptotic case where $N\to\infty$ ($N$ being the number of participants).
-Experimental conditions yielding plausibly spurious group polarization warrant 
-some suspicion. If there exists set of latent simple consensus 
-parameters $\nu$ that create binned (Equation~\ref{eq:ordinal-frequency})
-opinion distributions that naively seem to be group polarized, then the means
-of at least some samples from these distibutions with empirical numbers of 
-participants would also appear to be shifted consistent with group
-polarization. This process itself does not tell us the false detection rate.
-We use the identified simple consensus parameters to define pre- and post-deliberation
-latent opinion distributions in simulated trials of group polarization 
-experiments.
-
-\begin{figure}
-  \centering
-    % \includegraphics[width=0.8\textwidth]{Figures/Model/data-model-flow.png}
-    \includegraphics[width=\textwidth]{Figures/Model/data-model-flow-w-webapp.pdf}
-  \caption{\textbf{Workflow of the GP-stats simulation and analysis pipeline
-  for estimating false detection rates under null-consistent conditions.} Colored boxes indicate file types: teal = R scripts, blue = data files, orange = plotting scripts, cyan
-= Slurm batch script. Arrows show the flow from parameter estimation (hillclimbing or Shiny app) to Bayesian null model fitting on the cluster, aggregation of simulated results,
-and final visualization. (D) shows the web app interface that takes researcher guesses at
-simple consensus models that generate observations, and the app passes the guess to initialize
-the hillclimbing search algorithm for finding spurious group polarization.}
-  \label{fig:data-model-flow.png}
-\end{figure}
-
-\subsection{Metascience of the false detection rate in group polarization}
-
-Metascience provides a framework for
-calculating the false discovery rate across experimental conditions,
-aggregated to the level of journal articles and the ten-paper corpus. 
-The false discovery rate is then the Type I error rate relative to the rate of
-positive discoveries, true or false.
-We calculate via simulation and data fitting
-the Type I error rate for the experimental designs used in experimental conditions
-that, under a preliminary large-$N$ model analysis, generated group
-polarization upon ordinal measurement of simple consensus latent opinion 
-distributions. These can then be aggregated to estimate the false discovery rate
-for each journal article and for the ten-article corpus, which is representative of the
-false discovery rate for group polarization studies.
-
-The false detection rate is the scaled probability that a detection, $D$, of
-an effect, $E$, is in fact a false detection and there is \emph{not} an effect,
-denoted $\neg E$. The probability that a detection $D$ has occurred when there
-was no effect, i.e. when $\neg E$, is denoted $\alpha = \Pr(D | \neg E)$.
-This is known as the Type I or family-wise error rate. To calculate the 
-false detection rate ($\fdr$), one must compare $\alpha$ to the statistical power,
-the probability of a detection when there is an effect, denoted $W = \Pr(D | E)$.
-Both these values must be scaled by the rate at which effects occur, denoted $b$.
-Using the notation outlined above, the false detection rate is
-\begin{equation}
-  \fdr = \frac{(1 - b)\alpha}{(1 - b)\alpha + bW}.
-  \label{eq:fdr}
-\end{equation}
-
-In group polarization, $E$ represents the group polarization \emph{effect}, and
-$\neg E$ represents its absence. Simple consensus is one way for group polarization
-to be absent, i.e., simple consensus is one form of $\neg E$. The presence of
-an effect $E$ is independent of whether an effect was \emph{detected}, 
-represented by $D$ if it was detected and $\neg D$ if not.
-
-The group polarization studies are journal articles that present
-the results of multiple experiments to support some hypothesis. To organize
-the presentation of false detection rates across the ten studies in our
-analysis, we first will organize false detection rates by experiment, where
-we created `ExperimentID` tags for each of the sixty experiments across ten
-studies. We also created `StudyID` tags for each of the ten journal articles
-with the key `{Author}{Year}`, e.g., \texttt{Schkade2010} for Schkade, et al.,
-(2010) \cite{Schkade2010}.
-
-We calculate $\alpha = \Pr(D | \neg E)$ through simulation for experimental
-conditions, 
-and use optimistic values of
-$b=0.1$~\cite{Ioannidis2005,McElreath2015} and $W=0.8$~\cite{Smaldino2016NatSel}
-in the main text and in the compare with more pessimistic, yet realistic, estimates
-of these values, as well as even more optimistic estimates.
-
-\subsection{Estimating and limiting false detection rates}
-
-To calculate the false detection rate we begin by simulating 
-pre- and post-deliberation 
-opinions by drawing $N_e$ samples from normal latent opinion
-distributions that generated spurious group polarization, i.e., with
-parameters labelled $\nu_e$ found in the hillclimbing step. Binning these
-random values simulates observations. $N_e$ is the number
-of participants in experimental condition $e$. Each trial $i$ of $N_e$
-simulated pre- and post-deliberation observations is then inspected to see
-if spurious group polarization occurred in the simulation. Spurious group
-polarization is determined by first inferring the latent parameters of each
-simulated experimental trial by fitting an ordered probit model~\cite{Liddell2018}. 
-The inferred parameters are then used to calculate the effect size for that
-trial, Cohen's $d$, defined as~\cite[p. 331]{Liddell2018}
-\begin{equation}
-  d = \frac{\mupost - \mupre}{\sqrt{\frac{1}{2}(\sigmapost^2 + \sigmapre^2)}}.
-  \label{eq:cohens}
-\end{equation}
-\noindent
-The family-wise error rate is, $\alpha_{e} = \Pr(D|\neg E)$, is the 
-fraction of trials, 
-with $d_{ei} > d^*$ for the $i^\text{th}$ trial for experimental
-condition $e$, written
-\begin{equation}
-  \alpha_e = \frac{1}{N_T} \sum_i^{N_T} \Theta(d_{ei}; d^*),
-  \label{eq:study_aggregate}
-\end{equation}
-\noindent
-where
-\[
-  \Theta(d_{ei}; d^*) = 
-    \begin{cases}
-      1 & \text{ if } d_{ei} > d^* \\
-      0 & \text{ otherwise.}
-    \end{cases}
-\]
-and $N_T = 1000$ is the number of trials. For simplicity we assume that the sign of
-$d_{ei}$ is positive, meaning a shift to greater extremisim. It is possible that
-$d_{ei}$ could be negative, indicating a shift to lesser extremity. Indeed we
-observe many simulations with a shift to lesser extremity. For now we do not
-include these as indicating a ``significant'' effect, which we return to in the
-Analysis below.
-
-To calculate the family-wise error rate
-for a study, $s$ which contains several experimental conditions, $e$,
-\begin{equation}
-  \alpha_s = \frac{1}{|s|} \sum_{e \in s} \alpha_e
-\end{equation}
-\noindent
-The mean across all results identified as plausibly spurious is 
-\begin{equation}
-  \alpha_\mathrm{all} = \frac{1}{|E|} \sum_{e \in E} \alpha_e.
-\end{equation}
-We finally may calculate the false discovery rate by plugging this $\alpha$
-into Equation~\ref{eq:fdr} and assuming empirically-motivated values for 
-base rate, $b$, and statistical power, $W$, as well as more pessimistic and
-optimistic settings for base rate and power.
-
-We close our Analysis by inverting this procedure to find what significance value
-$d^*$ is necessary to achieve a low family-wise error rate $\alpha_e = 0.05$ for
-experimental condition, $e$. To do this we first calculate $\alpha_e(d^*)$ for a
-sufficiently wide range of $d^*$. We then find the lowest $d^*$ for which
-$\alpha_e(d^*) \leq 0.05$.
-
 \section{Analysis}\label{analysis}
@@
 \end{figure}
-
-\section{Discussion}\label{discussion}
-
-\mt{Restate results and importance and outline this section.}
-
-Since there seems to be no selection pressure on false discovery rates,
-experiment design for studying group polarization was free to evolve
-randomly. Various choices for the number of bins and other instrumental
-details may have then become entrenched in different communities defined by
-common theoretical perspectives taken by researchers in those communities.
-Therefore, when Isenberg (1986)~\cite{Isenberg1986} reports that his
-meta-analysis found more support for the persuasive arguments theory than the 
-social comparisons theory in the form of greater effect sizes. If this were
-indeed the case, perhaps it is only because some researchers hypothesizing
-persuasive arguments happened to use a measurement procedure produced more
-frequent and extreme distortions of opinion measurements. Incentive
-structures known to promote the ``evolution of bad science'' would select
-for research that ignored sticky, time-consuming statistical problems in
-favor of juicy titles and headlines for which practitioners have won several
-professional awards. Specifically, the traditional bias
-towards positive results, and against negative ones, could easily have
-provided a cultural niche in which quick but faulty group polarization
-statistical methods dominate slower, more rigorous ones.
-Now that we are aware of the problem, there must be
-intense selective pressure applied as soon and as widely as possible to 
-avoid further ``replications'' that report spurious group polarization, or
-any similar latent psychological dynamics.
-
-We provided tools that experimenters can use. While these tools are
-potentially useful, they are still in the prototype phase that was
-sufficient for a single researcher who was also the software developer. So,
-a first step towards expanding the usefulness of the findings here is to 
-improve the software usability and make the app widely available for
-researchers in group polarization and beyond. Another step would be to go
-beyond reactionary testing of the reliability of experimental designs, and
-identify design principles that reduce the false detection rate without
-needing to adjust the Cohen's $d$ (or whatever measure) 
-used as a significance threshold.
-
-When we used the appropriate ordered probit Bayesian statistical model, it
-still often failed to accurately identify mere agreement, instead detecting
-group polarization.  The ordered probit model we used still likely
-underestimates the variance in outcomes, which could require even larger
-sample sizes. A more rigorous analysis would include additional sources of
-variance that will likely further inflate the Type I error rate, $\alpha$,
-and the false discovery rate~\cite{Yarkoni2022}. First and foremost, there is
-variance within each experimental deliberation group; no group polarization
-study accounts for this, instead treating participants in each experimental
-condition as if they were one large group. Furthermore, survey data are known
-to be noisy~\cite{Zaller1992}. People report opinions differently over time
-for no apparent reason (i.e., opinions are \emph{unstable}).  Context
-matters: the order in which a survey question is asked, and which question
-framing is chosen among logically equivalent alternatives, can both
-significantly influence participant responses. Finally, little is known about
-the psychological process that converts latent opinions to reporting
-behaviors (e.g., clicking a radio button corresponding to an opinion).  The
-effect of accounting for these sources of variance must be understood to
-estimate $\alpha$ (and power, $W$) for the design of the next generation of
-group polarization experiments.
-
-A simple change in measurement procedure could fix the problem identified
-here: center the post-discussion scale on each group’s own pre-discussion
-mean. This removes the assumption that all groups share a common “neutral”
-midpoint. The situation is like tossing a ball on a moving train: to the
-passengers, the ball’s speed is the same in both directions, but to an
-observer on the ground one throw appears faster and the other slower.
-Likewise, when measurement is anchored to an external scale, simple agreement
-within a group can appear as movement toward one pole. Re-centering the
-measurement frame corrects this distortion. It acknowledges that opinion
-change is relative to the observer’s reference frame—there is no fixed
-background of neutrality against which all groups can be compared. This is,
-in essence, the insight of Newtonian relativity: apparent motion depends on
-the frame from which it is measured.
-
-First we must note that there are even more, perhaps even graver problems
-with group polarization research than I discussed so far: one of note is
-opinion instability. Opinions are not generally stable—people seem to
-construct their opinions based on several factors not directly related to
-whatever underlying topic. Factors affecting reported opinions include the
-language used to frame a prompt or question; the order of questions on a
-multi-item survey; or simply when the question was asked (Kalton and Schuman,
-1982; Zaller and Feldman, 1992; Zaller 1992). Failure to account for sources
-of input variance in any experiment can lead to inflated effect sizes since
-the input variance is missing from the posterior predictions during
-statistical model fitting (Yarkoni, 2022). This further compounds the
-theoretical confusions that fill the void when theories are never formalized
-and grounded in a mathematical or computational model (Smaldino 2020; Turner
-and Smaldino, 2022).
-
-A second additional problem with group polarization research is that we literally cannot learn anything by comparing summary statistics obtained with incommensurate experimental designs—but this is exactly how group polarization research has tried to make progress.
-Comparing summary statistics only has explanatory power if datasets were generated under the same experimental designs—it seems this was never done or even considered in the group polarization literature.
-In order to compare studies, their datasets must be generated from commensurate experiments, meaning the studies must share the same predictive and statistical models and experimental design.
-Otherwise, such comparisons are meaningless, they literally tell us nothing, as shown by philosopher of science Nancy Cartwright in Chapter 5 of her book The Dappled World: A Study of the Boundaries of Science.
-
-\subsection{Humble Hearts Promote Rigorous Science}
-
-Cass Sunstein kicked off the 2000s by elevating group polarization to be a scientific law and advocating the application of group polarization research to serious real-world problems in two books (Sunstein, 2009; Sunstein, 2019) and two recent high-profile articles in Nature Human Behavior: one on the social foundations of pandemic preparedness (Van Bavel, et al., 2020) and one to “promote truth, autonomy, and democratic discourse online” (Lorenz-Spreen, Lewandowski, Sunstein, and Hertwig, 2020).
-Let us try to stop the spread of this bad science—and let us explore what must be done to achieve Dorian Cartwright’s vision of a rigorous science of group polarization he articulated fifty years ago.
-
-To be practically useful, science must be rigorous.
-Rigorous science comes most naturally when scientists remain humble, especially when they serve at the pleasure of the public, the government, or other donors.
-Here’s another example from Burnstein and Vinokur (1973) that we can learn from—let’s try to avoid the sort of arrogance and haughtiness that seep from these acknowledgements they wrote on p.
-123:
-
-\begin{quote}
-This research was supported by a Grant from the National Institute of Mental Health (MH-16950-03) and by a Guggenheim Foundation Fellowship awarded to the first author.
-We gratefully acknowledge the incisive, witty, and vinous commentary provided by colleagues at the University of Provence, in particular Robert Abelson, Claude Flament, and Jean-Pierre Poitou, as well as the spirited and intelligent efforts of Livia Mezrick and Irene Graczyk, who helped us carry out the study.
-\end{quote}
-
-Science is not glamorous—it takes guts and sacrifice to achieve the technical skills and discipline to sharpen one’s prose, methods, mathematics, and computer code to the point they can really cut reality at its joints, investigate different parts, then put it all back together with a message for others about how the world really works.
-These days, to do rigorous social science work one must wisely curate the best of the available social science by harmonizing cross-disciplinary jargon.
-But disincentives still exist against careful scholarship.
-Prizes and professorships go to those with the most numerous publications in the right journals—period.
-
-Unpaid, anonymous reviewers are supposed to act as gatekeepers to catch problems like the ones I reviewed.
-But, journal article reviewers have little to gain from enforcing rigor—actually there is a lazy strategy that can pay off big for freeloaders.
-It’s easy to imagine a cultural norm evolving where reviewers save themselves time and energy by asking little of an author, in hopes that little will be demanded of them when the reviewer submits their next paper.
-This could result in a publication process that rewards the laziest and most willing to flood the zone and self-aggrandize.
-
-\subsection{Conclusion}
-
-Social science can use physics as a model discipline to transition from idiosyncratic and Balkanized to informative and united.
-Physics only changed the world because it is useful—it is useful only because it reliably predicts things.
-Physics became more useful as it became more united, finding shared representations for concepts developed independently—for example the science of optics was developed independently from the science of dynamics, but the physics of light is explained partly in analogy to a pendulum or spring—the mathematics of harmonic motion can be used to predict the motion of springs or pendulums—or why we see different colors of light through a prism.
-With careful, coordinated tweaks to group polarization experimental design and theoretical modeling, we will build a more rigorous theory of group polarization and its large-scale expression, echo chamber radicalization.
-
-I expect that group polarization experiments, when properly constructed and analyzed, will help sustainability campaigns optimize their social influence campaigns.
-We can start by developing simpler experimental designs based on formal or computational models capable of predicting outcomes in an experiment.
-I have developed a model that could motivate an experimental design: I showed that if extremists are more stubborn, this is sufficient to cause group polarization when likeminded groups interact over time.
-There is even a free stubborn extremism parameter that could be fit to data to measure the relationship between stubbornness and extremism on different issues.
-
-Group polarization research so far has mostly provided us with a checklist of what not to do.
-The alternative approach I outline builds on rock—the rock of rigor—and will supersede past group polarization research built on a thousand grains of inexact verbal speculation.
-
-\printbibliography[title=References]
-
-\appendix
-
-\renewcommand{\thefigure}{A\arabic{figure}}
-
-\setcounter{figure}{0}
-
-\clearpage
-
-%%%%%%%%%% Supplement %%%%%%%%%%
-\pagebreak
-\begin{center}
-  \textbf{\Large \textsf{Group polarization replications may be marred by high false discovery rates (Supplementary Material)}}
-\end{center}
-
-%%%%%%%%%% Prefix a "S" to all equations, figures, tables and reset the counter %%%%%%%%%%
-\setcounter{equation}{1}
-\setcounter{figure}{0}
-\setcounter{section}{0}
-\setcounter{table}{0}
-\setcounter{page}{1}
-\makeatletter
-\renewcommand{\theequation}{S\arabic{equation}}
-\renewcommand{\thefigure}{S\arabic{figure}}
-\renewcommand{\thetable}{S\arabic{table}}
-\renewcommand{\thesection}{S\arabic{section}}
-\renewcommand{\thepage}{S\arabic{page}}
-
-\section{Analysis}
-
-\begin{figure}
-  \caption{Boxplot of Cohen's $d$ for the ordered probit model over 1000
-    simulation trias for each experimental condition (y-axis in A and B). Each
-    trial represents a possible outcome of a group polarization experiment where
-    the true opinion shift is zero. The mean across trials is closer to 0 than
-    in the metric case, but many simulated zero-shift experiments still result in
-    an observed shift. This illustrates the weak statistical power of these group
-    polarization experiments.}
-  \label{fig:OrdinalBoxplot}
-  \centering
-    \includegraphics[width=1.0\textwidth]{Figures/Analysis/ordinal_cohens.pdf}
-\end{figure}
-
-\end{document}

### 2025-10-30 GroupPolarizationStatmod/paper/discussion.tex
on to method

--- 2025-10-30 GroupPolarizationStatmod/paper/ms.tex
@@
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
-\PassOptionsToPackage{hyphens}{url}
-\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
-%
-\documentclass[
-  abstract]{article}
-
-\usepackage{amsmath,amssymb}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-\usepackage{iftex}
-\input{preamble.tex}
-\hypersetup{
-  pdftitle={Group polarization replications may be marred by high false discovery rates},
-  colorlinks=true,
-  linkcolor={blue},
-  filecolor={Maroon},
-  citecolor={Blue},
-  urlcolor={Blue}}
-
-\newcommand{\mupre}{\mu_\mathrm{pre}}
-\newcommand{\mupost}{\mu_\mathrm{post}}
-
-\newcommand{\sigmapre}{\sigma_\mathrm{pre}}
-\newcommand{\sigmapost}{\sigma_\mathrm{post}}
-
-\newcommand{\thetafit}{\theta^\mathrm{fit}}
-
-\newcommand{\fdr}{\mathrm{FDR}}
-
-\newcommand{\meanobs}{\bar{o}}
-\newcommand{\meanobsemp}{\bar{\mathbf{o}}}
-\newcommand{\meanobst}{\meanobs_t}
-\newcommand{\meanobsjt}{\meanobs_{j,t}}
-
-\newcommand{\meanobspre}{\meanobs_\mathrm{pre}}
-\newcommand{\meanobspost}{\meanobs_\mathrm{post}}
-
-\definecolor{myorange}{RGB}{240, 96, 0}
-\newcommand{\mt}[1]{{\textcolor{myorange} {({\tiny MT:} #1)}}}
-
-\usepackage{authblk}
-\usepackage{etoolbox}
-\usepackage{wrapfig}
-\makeatletter
-% \title{If the Null Fits, You Must Omit: Ubiquitous False Detections of Group Polarization}
-\title{If the Null Fits, You Must Omit: Measurement Artifacts Undermine Group Polarization Replications}
-\makeatother
-\author[1,*]{{Matthew A.~Turner}}
-\affil[1]{\small Environmental Social Sciences, Stanford Doerr School of Sustainability, Stanford University}
-
-\author[2,3]{{Paul E.~Smaldino}}
-\affil[2]{\small Cognitive and Information Sciences, University of California, Merced}
-\affil[3]{\small Santa Fe Institute} 
-% \vspace{2em}
-\affil[*]{\small Correspondence: \href{mailto:maturner@stanford.edu}{maturner@stanford.edu}}
-
-\date{\today}
-\begin{document}
-\maketitle
-
-\renewcommand*\contentsname{Table of contents}
-{
-\hypersetup{linkcolor=gray}
-\setcounter{tocdepth}{2}
-\tableofcontents
-}
-
-\begin{center} \noindent\rule{4cm}{0.4pt} \end{center}
-% \clearpage
-
-% \begin{quote}
-% In our introductory social psychology course, 
-% we have for many years used the [group polarization experimental paradigm] as
-% a laboratory exercise. The exercise works beautifully, but one must be
-% careful to forewarn a class that [group polarization] does not occur with every 
-% group\ldots and that the effect is not large. 
-% \par\raggedleft(Brown, 1986, p.\cite[p. 205]{Brown1986})
-% \end{quote}
-
-% \begin{quote}
-%   The good thing about science is that it's true, whether or not you believe
-  % it. (Neil de Grasse Tyson slogan circa 2018)
-% \end{quote}
-
-\input{intro-latest.tex}
-
-% \input{old-intro.tex}
-
-\section{Methods}\label{methods}
-
-To find examples of plausibly spurious group polarization we guess at what latent
-mean, combined with a latent pre- and post-deliberation latent standard deviations,
-could generate it. We calculate the observed group polarization for these latent
-parameters by transforming the latent pre- and post-deliberation distributions from
-continuous normal distributions to ordinal ``observations'' by integrating over each
-ordinal bin, then calculating the observed mean for each experiment \emph{phase},
-i.e., both the pre- and post-deliberation phases.  If the null model can generate
-group polarization in this case, then the detection is labelled \emph{plausibly
-false}, i.e., \emph{plausibly spurious}, and we call the generating latent
-parameters the \emph{plausible null parameters}.
-
-To evaluate whether the plausible null parameters we identify could realistically
-fool a significance test, we implemented a generative simulation of each plausibly
-false detection of group polarization and calculated the \emph{family-wise error rate
-(FWER)} and the \emph{false discovery rate (FDR)}. We use the published number of
-participants $N$ for each of the 54 conditions that yielded plausibly false
-detections of group polarization. The FWER is also known as the Type I
-error, or the false positive rate: it is the probability that a detection of an
-effect occurs given that in reality there was a null effect.  The FDR is the
-probability that a published positive detection of an effect is a false positive,
-i.e., a detection of a null effect. High FWER and FDR for an experimental condition
-indicates that its design is weak, since we identify only one of perhaps infinitely
-many latent parameter combinations that can generate spurious group polarization.
-
-We use this setup to demonstrate how \emph{adversarial} style testing of experimental
-designs with counterfactual null models can be used to estimate important
-metascience parameters. In this study we use our FWER calculations to estimate
-the minimum Cohen's $d$ that should be used to determine whether an effect is
-``significant'' or not that will limit the FWER to 5\%. In the Discussion we
-sketch how our model can then be extended to rigorously calculate statistical power,
-i.e., the probability that a significant effect is detected given that there 
-is indeed a significant effect.
-
-\subsection{Probability theory of spurious group polarization}
-
-We formalize this experimental design in the $N\to\infty$ case where the
-distribution of ordinal measurements is created by integrating continuous latent
-opinion distributions.  Latent opinions in the model are normally distributed with
-mean $\mu_t$ and variance $\sigma_t$ for time $t$. We are concerned with only two
-times, \emph{pre-} and \emph{post-} deliberation ($t=0$ and $t=T$, respectively).
-For short we can write $\mu_\mathrm{pre}$ and $\sigma_\mathrm{pre}$ for
-pre-deliberation mean and variance (and similar for \emph{post}-deliberation).  
-The latent opinion of participant $i$ at time $t$ is drawn from a normal 
-distribution in the model, 
-\begin{equation} 
-  o_{i,t} \sim p(o; \mu_t, \sigma_t) = \mathcal{N}(\mu_t, \sigma_t).
-  \label{eq:opinionDistribution} 
-\end{equation} 
-\noindent 
-Latent opinions are assumed to be normally distributed since they are unbounded 
-aggregations of neurobiological activity.
-
-The measurement process with $N\to\infty$ participants is simulated by 
-integrating the latent opinion distribution between ordinal bin thresholds, 
-of which there are $B+1$, where $B$ is the number of opinion bins, with sequential
-integer values $b$. In the Moscovici and Zavalloni (1969) example, $b \in
-\{-3,-2,\ldots,2,3\}$ and $B = 7$.
-Bin $b$ is defined by its lower and upper continuous-valued thresholds, 
-$\theta_{b-1}$ and
-$\theta_{b}$, respectively. Thresholds are given by the following equations
-(and visualized as dotted lines in Figure~\ref{fig:distros}).
-\begin{equation}
-\theta_b = \begin{cases}
-  -\infty         & \text{ if } b=0 \\
-  \infty          & \text{ if } b=B \\
-  b + \frac{1}{2} & \text{ otherwise.}
-\end{cases}
-\end{equation}
-\noindent
-With thresholds defined, the frequency of ordinal measurment $b$ at time $t$ 
-is (Figure~\ref{fig:distros})
-\begin{equation}
-  p(b;~\mu_t, \sigma_t, B) = \int_{\theta_{b-1}}^{\theta_b} p(o; \mu_t, \sigma_t)
-  do.
-  \label{eq:ordinal-frequency}
-\end{equation}
-
-We can now define spurious group polarization and give examples of when it occurs.
-First, the observed group polarization opinion shift is
-\begin{equation}
-  g = \meanobs_\mathrm{post} - \meanobs_\mathrm{pre}~,
-  \label{eq:group-polarization}
-\end{equation}
-\noindent
-where the expected value of observed opinions at time $t$ is
-\begin{equation}
-  \meanobst = \frac{1}{B} \sum_{b=1}^B b \cdot p(b; \mu_t, \sigma_t, B).
-  \label{eq:meanobs}
-\end{equation}
-\noindent
-
-\emph{Spurious} group polarization is when mean observed opinions seem to increase in 
-extremity when the latent mean is held constant. Call the collection of
-latent null model distribution parameters $\lambda = (\mu, \sigma_{pre}, \sigma_{post})$.
-The group polarization opinion shift can then be written as a function of $\lambda$,
-$g(\lambda)$. Formally, then, spurious group polarization is plausible if 
-there exists at least one $\lambda$ for which $g(\lambda) = g_{pub} > 0$, where
-$g_{pub}$ is the published group polarization shift value.
-This could be thought of as sampling the observed ordinal distributions 
-$N \to \infty$, collecting ordinal measurements of samples from 
-each latent distribution corresponding to each phase, pre- and post-deliberation.
-
-Next, we explain how to find a set of latent null parameters, $\lambda$, that
-generate spurious group polarization when naively measured on an ordinal scale.
-
-\begin{figure}
-  \caption{\textbf{Opinion measurement model and spurious group polarization.} 
-  Latent model participant opinions are drawn from a normal distribution with the
-  listed parameters (A,C). The pre-deliberation latent distribution has high
-  variance, but appears polarized when measured via integration (A).  The latent
-  distribution's variance is expected to decrease during deliberation (B) via
-  consensus, but in this example the latent mean is static, 
-$\mupre = \mupost$. However, when the post-deliberation distribution is measured,
-the simulated observed mean increases, i.e., spurious
-group polarization occurs (C).}
-  \centering
-  \includegraphics[width=0.95\textwidth]{Figures/Model/latent-ordinal-distros.pdf}
-  \label{fig:distros}
-\end{figure}
-
-\subsection{Interactive algorithm to find parameters inducing spurious group polarization}
-
-To identify plausibly false detections of group polarization, we searched numerically
-for latent parameters $\lambda = (\mu, \sigmapre, and \sigmapost)$ that generate
-spurious group polarization, as explained immediately above. We developed an *ad hoc*
-hillclimbing-style algorithm that finds $\lambda$ that finds $\lambda$ by minimizing
-the difference between published group polarization shifts $g_pub$ and calculated shifts,
-$g_\lambda$.
-calculated with Equation~\ref{eq:meanobs}.
-
-The algorithm to find $\lambda$ holds $\mu$ constant while searching for $\sigmapre$
-and $\sigmapost$ that generate observed $\meanobspre$ and $\meanobspost$, within a
-tolerance of $\pm10^{-3}$ (Algorithm~\ref{alg:hillclimbing}). 
-The algorithm requires a researcher to specify an initial
-guess for $\lambda$ and auxiliary parameter settings, which generally vary between
-experimental conditions.  Furthermore, the algorithm does not always produce
-reasonable distributions with a given $\lambda$, and sometimes it does not produce
-spurious group polarization.  Therefore, it is necessary to sanity check the latent
-and ordinal opinion distributions for the $\lambda$ found by the algorithm. If all is
-well, the experimental condition can be marked as plausibly spurious.  To tame the
-complexity involved in data entry and model fitting for each of these sixty
-experimental conditions, we developed a web application to input data from control
-the algorithm and inspect its results (Figure~\ref{fig:webapp}). Web application code is
-available online in our GitHub
-repository\footnote{\url{https://github.com/mt-digital/gp-statmod/blob/main/GroupPolarizationStatmod/app.R}}.
-
-\begin{algorithm}
-  \caption{Routine to find latent mean and variance that generate spurious group polarization.}
-  \label{alg:hillclimbing}
-  \begin{algorithmic}
-
-    \Require $\mu$, $\sigma$, $\bar{\mathbf{o}}$ 
-      \Comment{Set latent mean, guess variance that yields, input empirical mean.}
-
-    \Require $\epsilon$, $\delta$, $i_\mathrm{max}$ 
-      \Comment{Hillclimbing tolerance, step size, and maximum iterations.}
-
-    \State $\Delta \gets \meanobs(\mu, \sigma) - \meanobsemp$ 
-      \Comment{Initialize simulation error.}
-    \While {$\Delta^2 > \epsilon$ and $i < i_\mathrm{max}$} \Comment{Search until desired
-    accuracy or max iterations reached.}
-      \State $\sigma' \sim \mathcal{N}(\sigma, \delta)$ \Comment{Draw new variance to test.}
-      \State $\Delta' \gets \bar{o}(\mu, \sigma') - \bar{\mathbf{o}}$ 
-        \Comment{Calculate error using new variance.}
-      \If {$\Delta' - \Delta < 0$}
-        \State $\sigma \gets \sigma'$
-          \Comment{Update $\sigma$ if the error is less than before.}
-        \State $\Delta \gets \bar{o}(\mu, \sigma) - \bar{\mathbf{o}}$
-          \Comment{Update simulation error if $\sigma$ is updated.}
-      \EndIf
-    \EndWhile \\
-
-    \Return $\sigma$
-  \end{algorithmic}
-\end{algorithm}
-
-% \begin{figure}
-%   \caption{\textbf{Shiny web application for user-guided hillclimbing to find
-%   parameters that generate spurious group polarization.} This web application enables users to
-% create new analyses of experimental conditions from published studies. Users 
-% provide initial guesses for the algorithm to use as starting points in 
-% finding variances that generate spurious group polarization. Users inspect 
-% model fits provided by the algorithm, adjusting parameters if necessary, and
-% mark an experimental condition as ``Plausibly spurious'' if a reasonable fit
-% is indeed found.\mt{Maybe add some A, B, C guides to the figure and include
-% in parentheses here to go with the text?}}
-%   \centering
-%     \includegraphics[width=0.75\textwidth]{Figures/Model/WebApp.png}
-%   \label{fig:webapp}
-% \end{figure}
-
-Note that so far, we have only identified cases where spurious group polarization plausibly
-occurs in the asymptotic case where $N\to\infty$ ($N$ being the number of participants).
-Experimental conditions yielding plausibly spurious group polarization warrant 
-some suspicion. If there exists set of latent simple consensus 
-parameters $\lambda$ that create binned (Equation~\ref{eq:ordinal-frequency})
-opinion distributions that naively seem to be group polarized, then the means
-of at least some samples from these distibutions with empirical numbers of 
-participants would also appear to be shifted consistent with group
-polarization. This process itself does not tell us the false detection rate.
-We use the identified simple consensus parameters to define pre- and post-deliberation
-latent opinion distributions in simulated trials of group polarization 
-experiments.
-
-\begin{figure}
-  \centering
-    % \includegraphics[width=0.8\textwidth]{Figures/Model/data-model-flow.png}
-    \includegraphics[width=\textwidth]{Figures/Model/data-model-flow-w-webapp.pdf}
-  \caption{\textbf{Workflow of the GP-stats simulation and analysis pipeline
-  for estimating false detection rates under null-consistent conditions.} Colored boxes indicate file types: teal = R scripts, blue = data files, orange = plotting scripts, cyan
-= Slurm batch script. Arrows show the flow from parameter estimation (hillclimbing or Shiny app) to Bayesian null model fitting on the cluster, aggregation of simulated results,
-and final visualization. (D) shows the web app interface that takes researcher guesses at
-simple consensus models that generate observations, and the app passes the guess to initialize
-the hillclimbing search algorithm for finding spurious group polarization.}
-  \label{fig:data-model-flow.png}
-\end{figure}
-
-\subsection{Metascience of the false detection rate in group polarization}
-
-Metascience provides a framework for
-calculating the false discovery rate across experimental conditions,
-aggregated to the level of journal articles and the ten-paper corpus. 
-The false discovery rate is then the Type I error rate relative to the rate of
-positive discoveries, true or false.
-We calculate via simulation and data fitting
-the Type I error rate for the experimental designs used in experimental conditions
-that, under a preliminary large-$N$ model analysis, generated group
-polarization upon ordinal measurement of simple consensus latent opinion 
-distributions. These can then be aggregated to estimate the false discovery rate
-for each journal article and for the ten-article corpus, which is representative of the
-false discovery rate for group polarization studies.
-
-The false detection rate is the scaled probability that a detection, $D$, of
-an effect, $E$, is in fact a false detection and there is \emph{not} an effect,
-denoted $\neg E$. The probability that a detection $D$ has occurred when there
-was no effect, i.e. when $\neg E$, is denoted $\alpha = \Pr(D | \neg E)$.
-This is known as the Type I or family-wise error rate. To calculate the 
-false detection rate ($\fdr$), one must compare $\alpha$ to the statistical power,
-the probability of a detection when there is an effect, denoted $W = \Pr(D | E)$.
-Both these values must be scaled by the rate at which effects occur, denoted $b$.
-Using the notation outlined above, the false detection rate is
-\begin{equation}
-  \fdr = \frac{(1 - b)\alpha}{(1 - b)\alpha + bW}.
-  \label{eq:fdr}
-\end{equation}
-
-In group polarization, $E$ represents the group polarization \emph{effect}, and
-$\neg E$ represents its absence. Simple consensus is one way for group polarization
-to be absent, i.e., simple consensus is one form of $\neg E$. The presence of
-an effect $E$ is independent of whether an effect was \emph{detected}, 
-represented by $D$ if it was detected and $\neg D$ if not.
-
-The group polarization studies are journal articles that present
-the results of multiple experiments to support some hypothesis. To organize
-the presentation of false detection rates across the ten studies in our
-analysis, we first will organize false detection rates by experiment, where
-we created `ExperimentID` tags for each of the sixty experiments across ten
-studies. We also created `StudyID` tags for each of the ten journal articles
-with the key `{Author}{Year}`, e.g., \texttt{Schkade2010} for Schkade, et al.,
-(2010) \cite{Schkade2010}.
-
-We calculate $\alpha = \Pr(D | \neg E)$ through simulation for experimental
-conditions, 
-and use optimistic values of
-$b=0.1$~\cite{Ioannidis2005,McElreath2015} and $W=0.8$~\cite{Smaldino2016NatSel}
-in the main text and in the compare with more pessimistic, yet realistic, estimates
-of these values, as well as even more optimistic estimates.
-
-\subsection{Estimating and limiting false detection rates}
-
-To calculate the false detection rate we begin by simulating 
-pre- and post-deliberation 
-opinions by drawing $N_e$ samples from normal latent opinion
-distributions that generated spurious group polarization, i.e., with
-parameters labelled $\lambda_e$ found in the hillclimbing step. Binning these
-random values simulates observations. $N_e$ is the number
-of participants in experimental condition $e$. Each trial $i$ of $N_e$
-simulated pre- and post-deliberation observations is then inspected to see
-if spurious group polarization occurred in the simulation. Spurious group
-polarization is determined by first inferring the latent parameters of each
-simulated experimental trial by fitting an ordered probit model~\cite{Liddell2018}. 
-The inferred parameters are then used to calculate the effect size for that
-trial, Cohen's $d$, defined as~\cite[p. 331]{Liddell2018}
-\begin{equation}
-  d = \frac{\mupost - \mupre}{\sqrt{\frac{1}{2}(\sigmapost^2 + \sigmapre^2)}}.
-  \label{eq:cohens}
-\end{equation}
-\noindent
-The family-wise error rate is, $\alpha_{e} = \Pr(D|\neg E)$, is the 
-fraction of trials, 
-with $d_{ei} > d^*$ for the $i^\text{th}$ trial for experimental
-condition $e$, written
-\begin{equation}
-  \alpha_e = \frac{1}{N_T} \sum_i^{N_T} \Theta(d_{ei}; d^*),
-  \label{eq:study_aggregate}
-\end{equation}
-\noindent
-where
-\[
-  \Theta(d_{ei}; d^*) = 
-    \begin{cases}
-      1 & \text{ if } d_{ei} > d^* \\
-      0 & \text{ otherwise.}
-    \end{cases}
-\]
-and $N_T = 1000$ is the number of trials. For simplicity we assume that the sign of
-$d_{ei}$ is positive, meaning a shift to greater extremisim. It is possible that
-$d_{ei}$ could be negative, indicating a shift to lesser extremity. Indeed we
-observe many simulations with a shift to lesser extremity. For now we do not
-include these as indicating a ``significant'' effect, which we return to in the
-Analysis below.
-
-To calculate the family-wise error rate
-for a study, $s$ which contains several experimental conditions, $e$,
-\begin{equation}
-  \alpha_s = \frac{1}{|s|} \sum_{e \in s} \alpha_e
-\end{equation}
-\noindent
-The mean across all results identified as plausibly spurious is 
-\begin{equation}
-  \alpha_\mathrm{all} = \frac{1}{|E|} \sum_{e \in E} \alpha_e.
-\end{equation}
-We finally may calculate the false discovery rate by plugging this $\alpha$
-into Equation~\ref{eq:fdr} and assuming empirically-motivated values for 
-base rate, $b$, and statistical power, $W$, as well as more pessimistic and
-optimistic settings for base rate and power.
-
-We close our Analysis by inverting this procedure to find what significance value
-$d^*$ is necessary to achieve a low family-wise error rate $\alpha_e = 0.05$ for
-experimental condition, $e$. To do this we first calculate $\alpha_e(d^*)$ for a
-sufficiently wide range of $d^*$. We then find the lowest $d^*$ for which
-$\alpha_e(d^*) \leq 0.05$.
-
-\section{Analysis}\label{analysis}
-
-We now present our determinations of whether published detections of group
-polarization are plausibly spurious and calculate the best-case false detection
-rate for each experimental design in the corpus. We found that 95\% of all
-detections of group polarization are plausibly false, 54 out of 57. We estimate the
-best-case median false detection rate to be 72\% for those 54 experimental
-conditions. All but seven of the 54 designs were found to require a $d^* > 0.8$
-for ``significant'' Cohen's $d$ to limit the family-wise error rate to 5\%.
-
-\subsection{Nearly all group polarization detections are plausibly spurious}
-
-% We used our formal distribution model of group polarization experiments
-% to identify plausible latent simple consensus distributions that generate spurious
-% group polarization in 
-Of the remaining 57, we identified simple consensus $\lambda$ that generate spurious
-group polarization in 54 experimental conditions. In the other three, we did
-identify $\lambda$ that generate spurious group polarization, but the initial
-simulated distributions were highly polarized. 
-We rejected $\lambda$ generated for
-two spurious group polarization from Schkade, et al., (2010) because
-the histogram plots of the observed opinion distributions are not
-polarized~\cite[Figure 1, p. 234]{Schkade2010}. The other condition where we
-found but rejected spurious group polarization $\lambda$ was in a question 
-about a ``bad teacher'' in Myers (1975)~\cite{Myers1975a}, where lower values
-signify increasingly worse impressions of the teacher and higher values signify
-increasingly good impressions. It is unlikely that a
-``bad teacher'' would be rated as good in this case, so we excluded this 
-from our analysis as well. This exclusion may be overly generous since Myers (1975)
-did not provide details beyond mean values and opinion shifts, and the article was
-published before open data practices. Our corpus had a total of 60 experimental conditions, but we excluded 3 out of 60 
-that did not claim to be positive findings of group polarization.
-
-- Address sigmas, note differences
-
-\begin{figure}
-  \centering
-    \includegraphics[width=\textwidth]{Figures/3.1/latent-sd-cdf.png}
-    \caption{\textbf{}}
-  \label{fig:}
-\end{figure}
-
-\subsection{Half of experiments exceed a 70\% false discovery rate at ``medium'' significance}
-
-We simulated trials from each experimental condition, $e$, by sampling and binning
-pre- and post-deliberation latent opinions from distributions with parameters
-identified with the $N\to\infty$ model as causing plausibly spurious group
-polarization. For each $e$, we drew a number of samples equal to the number of
-participants in the original experiments, $N_\mathrm{emp}$. To estimate family-wise
-error and false detection rates, we fit an ordered probit model to the simulated
-observations and obtained estimates of the group polarization effect size, $d$
-(Equation~\ref{eq:cohens}). If the effect size was greater than a significance
-value, $d > d^*$, then by definition there was a detection $D$ of group
-polarization, even though there was no effect by design, written $\neg E$.  The
-family-wise error rate is $\Pr(D|\neg E)$, calculated as the frequency with which $d
-> d^*$ out of 1000 simulation trials. The false detection rate is the family-wise
-error rate scaled using the base rate of effects and the statistical power
-(Equation~\ref{eq:fdr}). We test three difference significance values, $d =
-0.2,~0.5$ and $0.8$, corresponding to what Cohen prescribed to use for ``low'',
-``medium'', and ``high'' significance. If not otherwise noted, we set the base
-rate to be $b=0.1$ and the power $W = 0.8$ as metascience parameters to calculate 
-the false discovery rate (Equation~\ref{eq:fdr}). Note that if the family-wise error
-rate is low at $\alpha = 0.05$, the false discovery rate is $\fdr = 0.36$.
-
-Different conditions yield different effect size distributions
-(Figure~\ref{fig:effect_size_distros}), which 
-determine the magnitude of error rates and difference between the error rates for low,
-medium, and high significance values $d^*=0.2,0.5,0.8$ magnitude of error rates 
-(Figure~\ref{fig:fwer_fdr_synthesis}B). 
-Family-wise error rates varied from a minimum of one in ten for the
-\texttt{Feminists-Experimental} condition from Myers (1975)~\cite{Myers1975a}
-to a maximum of two in three for the \texttt{COSprings-CivilUnions} condition from
-Schkade, Sunstein, and Hastie (2010)~\cite{Schkade2010} under medium
-significance effect size $d^* = 0.5$ (Figure~\ref{fig:fwer_fdr_synthesis} and
-Table~\ref{tab:quantiles}). The median family-wise error
-rate is 0.28 for Krizan and Baron's (2007) experimental condition
-\texttt{NoOutgroupScenario1}~\cite{Krizan2007}. This translated to a minimum
-false detection rate of 0.52 for the Myers (1975) condition, a maximum false
-detection rate of 0.88 for the Schkade, et al., (2010) condition, and a median false
-detection rate of 0.75 for the Krizan and Baron (2007) condition.
-
-\begin{table}[ht]
-  \caption{{\textbf{Quantiles for the two error-rate measures under different
-  significance values.}}}
-  \label{tab:quantiles}
-  \centering
-  \input{tex/quantile_table_min_20_50_80_max.tex}
-\end{table}
-
-In the conditions
-with lower family-wise error rates, there were several conditions where using the low 
-significance value $\alpha$ resulted in a much greater error rate than the medium or 
-high significance values. In the ten conditions with the lowest $\alpha$ for 
-$d^* = 0.5$, with $\alpha$ near 0.1, five of these conditions have $\alpha > 0.25$
-when $d^* = 0.2$ corresponding to a small effect size. For low $\alpha$, the
-difference in false detection rates for each $d^*$ value are more widely
-distributed, following the distribution of $\alpha$, meaning that 
-a modest increase in $d^*$ can have greater effects on the false discovery rate. In
-the conditions above the 50th percentile, $\alpha$ approaches and exceeds one in
-two, even when requiring ``high'' significance with $d^*=0.8$. With $d^* = 0.8$,
-$\alpha$ approaches and exceeds one in four, with four near one in two. When 
-requiring ``low'' signifance,
-
-Experimental designs within each study or article tend to share certain
-characteristics, such as the number of bins in the ordinal measurement design
-or the number of participants per condition. Therefore, the 
-average false detection rate across experimental trials within a study
-(Equation~\ref{eq:study_aggregate}) show that
-
-\begin{figure}
-
-  \caption{\textbf{Family-wise error rates and false discovery rates across studies and experiments for
-    low, medium, and high Cohen's $d$ significance.}}
-  \label{fig:fwer_fdr_synthesis}
-
-  \centering
-  \includegraphics[width=1.1\textwidth]{Figures/Analysis/fwer_fdr_synthesis.pdf}
-\end{figure}
-
-\begin{itemize}
-  \item 
-    Other $b$ and $W$ values in quantiles.
-  \item
-    Note that there were many ``significant'' results---in the wrong direction
-    (Figure~\ref{fig:OrdinalBoxplot}).
-\end{itemize}
-
-\subsection{High significance values ($d^*$) often necessary for FWER of 0.05}
-
-Our simulations can be used a different way, in this case to calculate what
-significance value $d^*$ limits the family-wise error rate to the low value of
-0.05. 
-Solving the inverse problem of which $d^*$ achieves a low family-wise error rate,
-which drives the false detection rate. Here we present our calculations to find
-$d^*$ that achieve $\alpha(d^*) = 0.05$.
-
-\begin{figure}
-  \caption{
-    \textbf{Significance value, $d^*$, to limit family-wise error 
-    rate of 0.05 for each identified study.} The Moscovici and Zavalloni (1969)
-    ``Americans'' question would require the greatest $d^*$ to achieve $\alpha \leq
-    0.05$, even though it does not have the highest error rates. This is because of
-    larger outliers in simulated $d$ for this condition compared to others with
-    greater error rates (Figure~\ref{fig:OrdinalBoxplot}).
-  }
-  \centering
-    \includegraphics[width=0.75\textwidth]{Figures/Analysis/sigval_for_low_fwer.pdf}
-  \label{fig:sigval_for_low_fwer}
-\end{figure}
-
 \section{Discussion}\label{discussion}
@@
 The alternative approach I outline builds on rock—the rock of rigor—and will supersede past group polarization research built on a thousand grains of inexact verbal speculation.
-
-\printbibliography[title=References]
-
-\appendix
-
-\renewcommand{\thefigure}{A\arabic{figure}}
-
-\setcounter{figure}{0}
-
-\clearpage
-
-%%%%%%%%%% Supplement %%%%%%%%%%
-\pagebreak
-\begin{center}
-  \textbf{\Large \textsf{Group polarization replications may be marred by high false discovery rates (Supplementary Material)}}
-\end{center}
-
-%%%%%%%%%% Prefix a "S" to all equations, figures, tables and reset the counter %%%%%%%%%%
-\setcounter{equation}{1}
-\setcounter{figure}{0}
-\setcounter{section}{0}
-\setcounter{table}{0}
-\setcounter{page}{1}
-\makeatletter
-\renewcommand{\theequation}{S\arabic{equation}}
-\renewcommand{\thefigure}{S\arabic{figure}}
-\renewcommand{\thetable}{S\arabic{table}}
-\renewcommand{\thesection}{S\arabic{section}}
-\renewcommand{\thepage}{S\arabic{page}}
-
-\section{Analysis}
-
-\begin{figure}
-  \caption{Boxplot of Cohen's $d$ for the ordered probit model over 1000
-    simulation trias for each experimental condition (y-axis in A and B). Each
-    trial represents a possible outcome of a group polarization experiment where
-    the true opinion shift is zero. The mean across trials is closer to 0 than
-    in the metric case, but many simulated zero-shift experiments still result in
-    an observed shift. This illustrates the weak statistical power of these group
-    polarization experiments.}
-  \label{fig:OrdinalBoxplot}
-  \centering
-    \includegraphics[width=1.0\textwidth]{Figures/Analysis/ordinal_cohens.pdf}
-\end{figure}
-
-\end{document}

### 2025-10-30 GroupPolarizationStatmod/paper/method.tex
on to method

--- 2025-10-30 GroupPolarizationStatmod/paper/ms.tex
@@
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
-\PassOptionsToPackage{hyphens}{url}
-\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
-%
-\documentclass[
-  abstract]{article}
+\section{Method}\label{method}
 
-\usepackage{floatrow}
-\DeclareFloatFont{small}{\small}% "scriptsize" is defined by floatrow, "tiny" not
-\floatsetup[table]{font=small}
+Our method combines several formal and computational tools to identify which
+replications should be retracted and evaluate experimental designs across our
+corpus. The first step in the analysis is to find a reasonable pair of latent
+null-polarization distributions that appear polarized due to clipping.  To do
+this we designed a web app that allows the user to try different latent
+distribution parameters (Figure~\ref{fig:data-model-flow}) to seed a
+hillclimbing algorithm that explores latent distribution parameter space
+until it finds null parameters that look like group polarization under the
+experiment's ordinal measurement procedure, or the algorithm stops after a
+certain number of steps (Algorithm~\ref{alg:hillclimbing}).
 
-\usepackage[table]{xcolor}
-\usepackage{multirow}
-\usepackage{makecell}
-\usepackage{array}
+—latent mean and pre- and post-discussion 
+variances that  loads experimental design and observation data
+from a user-provided CSV file
 
-\usepackage{amsmath,amssymb}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-\usepackage{iftex}
-\input{preamble.tex}
-\hypersetup{
-  pdftitle={Group polarization replications may be marred by high false discovery rates},
-  colorlinks=true,
-  linkcolor={blue},
-  filecolor={Maroon},
-  citecolor={Blue},
-  urlcolor={Blue}}
-
-\newcommand{\mupre}{\mu_\mathrm{pre}}
-\newcommand{\mupost}{\mu_\mathrm{post}}
-
-\newcommand{\sigmapre}{\sigma_\mathrm{pre}}
-\newcommand{\sigmapost}{\sigma_\mathrm{post}}
-
-\newcommand{\thetafit}{\theta^\mathrm{fit}}
-
-\newcommand{\fdr}{\mathrm{FDR}}
-
-\newcommand{\meanobs}{\bar{o}}
-\newcommand{\meanobsemp}{\bar{\mathbf{o}}}
-\newcommand{\meanobst}{\meanobs_t}
-\newcommand{\meanobsjt}{\meanobs_{j,t}}
-
-\newcommand{\meanobspre}{\meanobs_\mathrm{pre}}
-\newcommand{\meanobspost}{\meanobs_\mathrm{post}}
-
-\definecolor{myorange}{RGB}{240, 96, 0}
-\newcommand{\mt}[1]{{\textcolor{myorange} {({\tiny MT:} #1)}}}
-
-\usepackage{authblk}
-\usepackage{etoolbox}
-\usepackage{wrapfig}
-\makeatletter
-% \title{If the Null Fits, You Must Omit: Ubiquitous False Detections of Group Polarization}
-\title{If the Null Fits, You Must Omit: Measurement Artifacts Undermine Group Polarization Replications}
-\makeatother
-\author[1,*]{{Matthew A.~Turner}}
-\affil[1]{\small Environmental Social Sciences, Stanford Doerr School of Sustainability, Stanford University}
-
-\author[2,3]{{Paul E.~Smaldino}}
-\affil[2]{\small Cognitive and Information Sciences, University of California, Merced}
-\affil[3]{\small Santa Fe Institute} 
-% \vspace{2em}
-\affil[*]{\small Correspondence: \href{mailto:maturner@stanford.edu}{maturner@stanford.edu}}
-
-\date{\today}
-\begin{document}
-\maketitle
-
-\renewcommand*\contentsname{Table of contents}
-{
-\hypersetup{linkcolor=gray}
-\setcounter{tocdepth}{2}
-\tableofcontents
-}
-
-\begin{center} \noindent\rule{4cm}{0.4pt} \end{center}
-% \clearpage
-
-\begin{quote}
-In our introductory social psychology course, 
-we have for many years used the [group polarization experimental paradigm] as
-a laboratory exercise. The exercise works beautifully, but one must be
-careful to forewarn a class that [group polarization] does not occur with every 
-group\ldots and that the effect is not large. 
-\par\raggedleft(Brown, 1986, p.\cite[p. 205]{Brown1986})
-\end{quote}
-
-\begin{quote}
-  The good thing about science is that it's true, whether or not you believe
-  it. (Neil de Grasse Tyson slogan circa 2018)
-\end{quote}
-
-\input{intro-latest.tex}
-
-% \input{old-intro.tex}
-
-\section{Methods}\label{methods}
+ordinal measurements distort to look 
+polarized.
 
@@
     $\theta_b$ & The $B+1$ threshold values that separate bins & 
-      e.g., $-\infty, -,\ldots$ \\
-    $b$ & Bin index & $0,1,2,\ldots$ \\
+      e.g., $-\infty, -2.5, -1.5,\ldots,2.5, \infty$ \\
+    $b$ & Bin index—there are $B+1$ to mark $B$ bins & $0,1,2,\ldots,B$ \\
   \bottomrule
@@
     % \includegraphics[width=0.8\textwidth]{Figures/Model/data-model-flow.png}
-    \includegraphics[width=\textwidth]{Figures/Model/data-model-flow-w-webapp.pdf}
-  \caption{\textbf{Workflow of the GP-stats simulation and analysis pipeline
-  for estimating false detection rates under null-consistent conditions.} Colored boxes indicate file types: teal = R scripts, blue = data files, orange = plotting scripts, cyan
-= Slurm batch script. Arrows show the flow from parameter estimation (hillclimbing or Shiny app) to Bayesian null model fitting on the cluster, aggregation of simulated results,
-and final visualization. (D) shows the web app interface that takes researcher guesses at
-simple consensus models that generate observations, and the app passes the guess to initialize
-the hillclimbing search algorithm for finding spurious group polarization.}
-  \label{fig:data-model-flow.png}
+  \includegraphics[width=\textwidth]{Figures/Model/SoftwareMap_Mashup.png}
+  \caption{\textbf{Simulation and analysis pipeline}
+  for finding when measurement artifacts.  
+  Arrows show the flow of input and output data from parameter estimation.
+  First, we used the web app (A) to find null-polarization parameters for 
+  each experimental condition, $\nu_e$,
+  that ordinal measurements distort to look like polarization
+  (B—described in Algorithm~\ref{alg:hillclimbing}).
+  We fit ordered probit models to simulated null-polarization outcomes,
+  seeded by the $\nu$ found in the previous step (C). Analysis and 
+  visualization routines finish the pipeline (D).}
+  \label{fig:data-model-flow}
 \end{figure}
@@
 $\alpha_e(d^*) \leq 0.05$.
-
-\section{Analysis}\label{analysis}
-
-We now present our determinations of whether published detections of group
-polarization are plausibly spurious and calculate the best-case false detection
-rate for each experimental design in the corpus. We found that 95\% of all
-detections of group polarization are plausibly false, 54 out of 57. We estimate the
-best-case median false detection rate to be 72\% for those 54 experimental
-conditions. All but seven of the 54 designs were found to require a $d^* > 0.8$
-for ``significant'' Cohen's $d$ to limit the family-wise error rate to 5\%.
-
-\subsection{Nearly all group polarization detections are plausibly spurious}
-
-% We used our formal distribution model of group polarization experiments
-% to identify plausible latent simple consensus distributions that generate spurious
-% group polarization in 
-Of the remaining 57, we identified simple consensus $\nu$ that generate spurious
-group polarization in 54 experimental conditions. In the other three, we did
-identify $\nu$ that generate spurious group polarization, but the initial
-simulated distributions were highly polarized. 
-We rejected $\nu$ generated for
-two spurious group polarization from Schkade, et al., (2010) because
-the histogram plots of the observed opinion distributions are not
-polarized~\cite[Figure 1, p. 234]{Schkade2010}. The other condition where we
-found but rejected spurious group polarization $\nu$ was in a question 
-about a ``bad teacher'' in Myers (1975)~\cite{Myers1975a}, where lower values
-signify increasingly worse impressions of the teacher and higher values signify
-increasingly good impressions. It is unlikely that a
-``bad teacher'' would be rated as good in this case, so we excluded this 
-from our analysis as well. This exclusion may be overly generous since Myers (1975)
-did not provide details beyond mean values and opinion shifts, and the article was
-published before open data practices. Our corpus had a total of 60 experimental conditions, but we excluded 3 out of 60 
-that did not claim to be positive findings of group polarization.
-
-- Address sigmas, note differences
-
-\begin{figure}
-  \centering
-    \includegraphics[width=\textwidth]{Figures/3.1/latent-sd-cdf.png}
-    \caption{\textbf{}}
-  \label{fig:}
-\end{figure}
-
-\subsection{Half of experiments exceed a 70\% false discovery rate at ``medium'' significance}
-
-We simulated trials from each experimental condition, $e$, by sampling and binning
-pre- and post-deliberation latent opinions from distributions with parameters
-identified with the $N\to\infty$ model as causing plausibly spurious group
-polarization. For each $e$, we drew a number of samples equal to the number of
-participants in the original experiments, $N_\mathrm{emp}$. To estimate family-wise
-error and false detection rates, we fit an ordered probit model to the simulated
-observations and obtained estimates of the group polarization effect size, $d$
-(Equation~\ref{eq:cohens}). If the effect size was greater than a significance
-value, $d > d^*$, then by definition there was a detection $D$ of group
-polarization, even though there was no effect by design, written $\neg E$.  The
-family-wise error rate is $\Pr(D|\neg E)$, calculated as the frequency with which $d
-> d^*$ out of 1000 simulation trials. The false detection rate is the family-wise
-error rate scaled using the base rate of effects and the statistical power
-(Equation~\ref{eq:fdr}). We test three difference significance values, $d =
-0.2,~0.5$ and $0.8$, corresponding to what Cohen prescribed to use for ``low'',
-``medium'', and ``high'' significance. If not otherwise noted, we set the base
-rate to be $b=0.1$ and the power $W = 0.8$ as metascience parameters to calculate 
-the false discovery rate (Equation~\ref{eq:fdr}). Note that if the family-wise error
-rate is low at $\alpha = 0.05$, the false discovery rate is $\fdr = 0.36$.
-
-Different conditions yield different effect size distributions
-(Figure~\ref{fig:effect_size_distros}), which 
-determine the magnitude of error rates and difference between the error rates for low,
-medium, and high significance values $d^*=0.2,0.5,0.8$ magnitude of error rates 
-(Figure~\ref{fig:fwer_fdr_synthesis}B). 
-Family-wise error rates varied from a minimum of one in ten for the
-\texttt{Feminists-Experimental} condition from Myers (1975)~\cite{Myers1975a}
-to a maximum of two in three for the \texttt{COSprings-CivilUnions} condition from
-Schkade, Sunstein, and Hastie (2010)~\cite{Schkade2010} under medium
-significance effect size $d^* = 0.5$ (Figure~\ref{fig:fwer_fdr_synthesis} and
-Table~\ref{tab:quantiles}). The median family-wise error
-rate is 0.28 for Krizan and Baron's (2007) experimental condition
-\texttt{NoOutgroupScenario1}~\cite{Krizan2007}. This translated to a minimum
-false detection rate of 0.52 for the Myers (1975) condition, a maximum false
-detection rate of 0.88 for the Schkade, et al., (2010) condition, and a median false
-detection rate of 0.75 for the Krizan and Baron (2007) condition.
-
-\begin{table}[ht]
-  \caption{{\textbf{Quantiles for the two error-rate measures under different
-  significance values.}}}
-  \label{tab:quantiles}
-  \centering
-  \input{tex/quantile_table_min_20_50_80_max.tex}
-\end{table}
-
-In the conditions
-with lower family-wise error rates, there were several conditions where using the low 
-significance value $\alpha$ resulted in a much greater error rate than the medium or 
-high significance values. In the ten conditions with the lowest $\alpha$ for 
-$d^* = 0.5$, with $\alpha$ near 0.1, five of these conditions have $\alpha > 0.25$
-when $d^* = 0.2$ corresponding to a small effect size. For low $\alpha$, the
-difference in false detection rates for each $d^*$ value are more widely
-distributed, following the distribution of $\alpha$, meaning that 
-a modest increase in $d^*$ can have greater effects on the false discovery rate. In
-the conditions above the 50th percentile, $\alpha$ approaches and exceeds one in
-two, even when requiring ``high'' significance with $d^*=0.8$. With $d^* = 0.8$,
-$\alpha$ approaches and exceeds one in four, with four near one in two. When 
-requiring ``low'' signifance,
-
-Experimental designs within each study or article tend to share certain
-characteristics, such as the number of bins in the ordinal measurement design
-or the number of participants per condition. Therefore, the 
-average false detection rate across experimental trials within a study
-(Equation~\ref{eq:study_aggregate}) show that
-
-\begin{figure}
-
-  \caption{\textbf{Family-wise error rates and false discovery rates across studies and experiments for
-    low, medium, and high Cohen's $d$ significance.}}
-  \label{fig:fwer_fdr_synthesis}
-
-  \centering
-  \includegraphics[width=1.1\textwidth]{Figures/Analysis/fwer_fdr_synthesis.pdf}
-\end{figure}
-
-\begin{itemize}
-  \item 
-    Other $b$ and $W$ values in quantiles.
-  \item
-    Note that there were many ``significant'' results---in the wrong direction
-    (Figure~\ref{fig:OrdinalBoxplot}).
-\end{itemize}
-
-\subsection{High significance values ($d^*$) often necessary for FWER of 0.05}
-
-Our simulations can be used a different way, in this case to calculate what
-significance value $d^*$ limits the family-wise error rate to the low value of
-0.05. 
-Solving the inverse problem of which $d^*$ achieves a low family-wise error rate,
-which drives the false detection rate. Here we present our calculations to find
-$d^*$ that achieve $\alpha(d^*) = 0.05$.
-
-\begin{figure}
-  \caption{
-    \textbf{Significance value, $d^*$, to limit family-wise error 
-    rate of 0.05 for each identified study.} The Moscovici and Zavalloni (1969)
-    ``Americans'' question would require the greatest $d^*$ to achieve $\alpha \leq
-    0.05$, even though it does not have the highest error rates. This is because of
-    larger outliers in simulated $d$ for this condition compared to others with
-    greater error rates (Figure~\ref{fig:OrdinalBoxplot}).
-  }
-  \centering
-    \includegraphics[width=0.75\textwidth]{Figures/Analysis/sigval_for_low_fwer.pdf}
-  \label{fig:sigval_for_low_fwer}
-\end{figure}
-
-\section{Discussion}\label{discussion}
-
-\mt{Restate results and importance and outline this section.}
-
-Since there seems to be no selection pressure on false discovery rates,
-experiment design for studying group polarization was free to evolve
-randomly. Various choices for the number of bins and other instrumental
-details may have then become entrenched in different communities defined by
-common theoretical perspectives taken by researchers in those communities.
-Therefore, when Isenberg (1986)~\cite{Isenberg1986} reports that his
-meta-analysis found more support for the persuasive arguments theory than the 
-social comparisons theory in the form of greater effect sizes. If this were
-indeed the case, perhaps it is only because some researchers hypothesizing
-persuasive arguments happened to use a measurement procedure produced more
-frequent and extreme distortions of opinion measurements. Incentive
-structures known to promote the ``evolution of bad science'' would select
-for research that ignored sticky, time-consuming statistical problems in
-favor of juicy titles and headlines for which practitioners have won several
-professional awards. Specifically, the traditional bias
-towards positive results, and against negative ones, could easily have
-provided a cultural niche in which quick but faulty group polarization
-statistical methods dominate slower, more rigorous ones.
-Now that we are aware of the problem, there must be
-intense selective pressure applied as soon and as widely as possible to 
-avoid further ``replications'' that report spurious group polarization, or
-any similar latent psychological dynamics.
-
-We provided tools that experimenters can use. While these tools are
-potentially useful, they are still in the prototype phase that was
-sufficient for a single researcher who was also the software developer. So,
-a first step towards expanding the usefulness of the findings here is to 
-improve the software usability and make the app widely available for
-researchers in group polarization and beyond. Another step would be to go
-beyond reactionary testing of the reliability of experimental designs, and
-identify design principles that reduce the false detection rate without
-needing to adjust the Cohen's $d$ (or whatever measure) 
-used as a significance threshold.
-
-When we used the appropriate ordered probit Bayesian statistical model, it
-still often failed to accurately identify mere agreement, instead detecting
-group polarization.  The ordered probit model we used still likely
-underestimates the variance in outcomes, which could require even larger
-sample sizes. A more rigorous analysis would include additional sources of
-variance that will likely further inflate the Type I error rate, $\alpha$,
-and the false discovery rate~\cite{Yarkoni2022}. First and foremost, there is
-variance within each experimental deliberation group; no group polarization
-study accounts for this, instead treating participants in each experimental
-condition as if they were one large group. Furthermore, survey data are known
-to be noisy~\cite{Zaller1992}. People report opinions differently over time
-for no apparent reason (i.e., opinions are \emph{unstable}).  Context
-matters: the order in which a survey question is asked, and which question
-framing is chosen among logically equivalent alternatives, can both
-significantly influence participant responses. Finally, little is known about
-the psychological process that converts latent opinions to reporting
-behaviors (e.g., clicking a radio button corresponding to an opinion).  The
-effect of accounting for these sources of variance must be understood to
-estimate $\alpha$ (and power, $W$) for the design of the next generation of
-group polarization experiments.
-
-A simple change in measurement procedure could fix the problem identified
-here: center the post-discussion scale on each group’s own pre-discussion
-mean. This removes the assumption that all groups share a common “neutral”
-midpoint. The situation is like tossing a ball on a moving train: to the
-passengers, the ball’s speed is the same in both directions, but to an
-observer on the ground one throw appears faster and the other slower.
-Likewise, when measurement is anchored to an external scale, simple agreement
-within a group can appear as movement toward one pole. Re-centering the
-measurement frame corrects this distortion. It acknowledges that opinion
-change is relative to the observer’s reference frame—there is no fixed
-background of neutrality against which all groups can be compared. This is,
-in essence, the insight of Newtonian relativity: apparent motion depends on
-the frame from which it is measured.
-
-First we must note that there are even more, perhaps even graver problems
-with group polarization research than I discussed so far: one of note is
-opinion instability. Opinions are not generally stable—people seem to
-construct their opinions based on several factors not directly related to
-whatever underlying topic. Factors affecting reported opinions include the
-language used to frame a prompt or question; the order of questions on a
-multi-item survey; or simply when the question was asked (Kalton and Schuman,
-1982; Zaller and Feldman, 1992; Zaller 1992). Failure to account for sources
-of input variance in any experiment can lead to inflated effect sizes since
-the input variance is missing from the posterior predictions during
-statistical model fitting (Yarkoni, 2022). This further compounds the
-theoretical confusions that fill the void when theories are never formalized
-and grounded in a mathematical or computational model (Smaldino 2020; Turner
-and Smaldino, 2022).
-
-A second additional problem with group polarization research is that we literally cannot learn anything by comparing summary statistics obtained with incommensurate experimental designs—but this is exactly how group polarization research has tried to make progress.
-Comparing summary statistics only has explanatory power if datasets were generated under the same experimental designs—it seems this was never done or even considered in the group polarization literature.
-In order to compare studies, their datasets must be generated from commensurate experiments, meaning the studies must share the same predictive and statistical models and experimental design.
-Otherwise, such comparisons are meaningless, they literally tell us nothing, as shown by philosopher of science Nancy Cartwright in Chapter 5 of her book The Dappled World: A Study of the Boundaries of Science.
-
-\subsection{Humble Hearts Promote Rigorous Science}
-
-Cass Sunstein kicked off the 2000s by elevating group polarization to be a scientific law and advocating the application of group polarization research to serious real-world problems in two books (Sunstein, 2009; Sunstein, 2019) and two recent high-profile articles in Nature Human Behavior: one on the social foundations of pandemic preparedness (Van Bavel, et al., 2020) and one to “promote truth, autonomy, and democratic discourse online” (Lorenz-Spreen, Lewandowski, Sunstein, and Hertwig, 2020).
-Let us try to stop the spread of this bad science—and let us explore what must be done to achieve Dorian Cartwright’s vision of a rigorous science of group polarization he articulated fifty years ago.
-
-To be practically useful, science must be rigorous.
-Rigorous science comes most naturally when scientists remain humble, especially when they serve at the pleasure of the public, the government, or other donors.
-Here’s another example from Burnstein and Vinokur (1973) that we can learn from—let’s try to avoid the sort of arrogance and haughtiness that seep from these acknowledgements they wrote on p.
-123:
-
-\begin{quote}
-This research was supported by a Grant from the National Institute of Mental Health (MH-16950-03) and by a Guggenheim Foundation Fellowship awarded to the first author.
-We gratefully acknowledge the incisive, witty, and vinous commentary provided by colleagues at the University of Provence, in particular Robert Abelson, Claude Flament, and Jean-Pierre Poitou, as well as the spirited and intelligent efforts of Livia Mezrick and Irene Graczyk, who helped us carry out the study.
-\end{quote}
-
-Science is not glamorous—it takes guts and sacrifice to achieve the technical skills and discipline to sharpen one’s prose, methods, mathematics, and computer code to the point they can really cut reality at its joints, investigate different parts, then put it all back together with a message for others about how the world really works.
-These days, to do rigorous social science work one must wisely curate the best of the available social science by harmonizing cross-disciplinary jargon.
-But disincentives still exist against careful scholarship.
-Prizes and professorships go to those with the most numerous publications in the right journals—period.
-
-Unpaid, anonymous reviewers are supposed to act as gatekeepers to catch problems like the ones I reviewed.
-But, journal article reviewers have little to gain from enforcing rigor—actually there is a lazy strategy that can pay off big for freeloaders.
-It’s easy to imagine a cultural norm evolving where reviewers save themselves time and energy by asking little of an author, in hopes that little will be demanded of them when the reviewer submits their next paper.
-This could result in a publication process that rewards the laziest and most willing to flood the zone and self-aggrandize.
-
-\subsection{Conclusion}
-
-Social science can use physics as a model discipline to transition from idiosyncratic and Balkanized to informative and united.
-Physics only changed the world because it is useful—it is useful only because it reliably predicts things.
-Physics became more useful as it became more united, finding shared representations for concepts developed independently—for example the science of optics was developed independently from the science of dynamics, but the physics of light is explained partly in analogy to a pendulum or spring—the mathematics of harmonic motion can be used to predict the motion of springs or pendulums—or why we see different colors of light through a prism.
-With careful, coordinated tweaks to group polarization experimental design and theoretical modeling, we will build a more rigorous theory of group polarization and its large-scale expression, echo chamber radicalization.
-
-I expect that group polarization experiments, when properly constructed and analyzed, will help sustainability campaigns optimize their social influence campaigns.
-We can start by developing simpler experimental designs based on formal or computational models capable of predicting outcomes in an experiment.
-I have developed a model that could motivate an experimental design: I showed that if extremists are more stubborn, this is sufficient to cause group polarization when likeminded groups interact over time.
-There is even a free stubborn extremism parameter that could be fit to data to measure the relationship between stubbornness and extremism on different issues.
-
-Group polarization research so far has mostly provided us with a checklist of what not to do.
-The alternative approach I outline builds on rock—the rock of rigor—and will supersede past group polarization research built on a thousand grains of inexact verbal speculation.
-
-\printbibliography[title=References]
-
-\appendix
-
-\renewcommand{\thefigure}{A\arabic{figure}}
-
-\setcounter{figure}{0}
-
-\clearpage
-
-%%%%%%%%%% Supplement %%%%%%%%%%
-\pagebreak
-\begin{center}
-  \textbf{\Large \textsf{Group polarization replications may be marred by high false discovery rates (Supplementary Material)}}
-\end{center}
-
-%%%%%%%%%% Prefix a "S" to all equations, figures, tables and reset the counter %%%%%%%%%%
-\setcounter{equation}{1}
-\setcounter{figure}{0}
-\setcounter{section}{0}
-\setcounter{table}{0}
-\setcounter{page}{1}
-\makeatletter
-\renewcommand{\theequation}{S\arabic{equation}}
-\renewcommand{\thefigure}{S\arabic{figure}}
-\renewcommand{\thetable}{S\arabic{table}}
-\renewcommand{\thesection}{S\arabic{section}}
-\renewcommand{\thepage}{S\arabic{page}}
-
-\section{Analysis}
-
-\begin{figure}
-  \caption{Boxplot of Cohen's $d$ for the ordered probit model over 1000
-    simulation trias for each experimental condition (y-axis in A and B). Each
-    trial represents a possible outcome of a group polarization experiment where
-    the true opinion shift is zero. The mean across trials is closer to 0 than
-    in the metric case, but many simulated zero-shift experiments still result in
-    an observed shift. This illustrates the weak statistical power of these group
-    polarization experiments.}
-  \label{fig:OrdinalBoxplot}
-  \centering
-    \includegraphics[width=1.0\textwidth]{Figures/Analysis/ordinal_cohens.pdf}
-\end{figure}
-
-\end{document}

### 2025-10-30 GroupPolarizationStatmod/paper/ms.tex
on to method

--- 2025-10-30 GroupPolarizationStatmod/paper/ms.tex
@@
 %
-\documentclass[
-  abstract]{article}
+\documentclass[preview]{article}
+
+\usepackage{floatrow}
+\DeclareFloatFont{small}{\small}% "scriptsize" is defined by floatrow, "tiny" not
+\floatsetup[table]{font=small}
+
+\usepackage[table]{xcolor}
+\usepackage{multirow}
+\usepackage{makecell}
+\usepackage{array}
 
@@
 
-\section{Methods}\label{methods}
+\input{method.tex}
 
-To find examples of plausibly spurious group polarization we guess at what latent
-mean, combined with a latent pre- and post-deliberation latent standard deviations,
-could generate it. We calculate the observed group polarization for these latent
-parameters by transforming the latent pre- and post-deliberation distributions from
-continuous normal distributions to ordinal ``observations'' by integrating over each
-ordinal bin, then calculating the observed mean for each experiment \emph{phase},
-i.e., both the pre- and post-deliberation phases.  If the null model can generate
-group polarization in this case, then the detection is labelled \emph{plausibly
-false}, i.e., \emph{plausibly spurious}, and we call the generating latent
-parameters the \emph{plausible null parameters}.
+\input{analysis.tex}
 
-To evaluate whether the plausible null parameters we identify could realistically
-fool a significance test, we implemented a generative simulation of each plausibly
-false detection of group polarization and calculated the \emph{family-wise error rate
-(FWER)} and the \emph{false discovery rate (FDR)}. We use the published number of
-participants $N$ for each of the 54 conditions that yielded plausibly false
-detections of group polarization. The FWER is also known as the Type I
-error, or the false positive rate: it is the probability that a detection of an
-effect occurs given that in reality there was a null effect.  The FDR is the
-probability that a published positive detection of an effect is a false positive,
-i.e., a detection of a null effect. High FWER and FDR for an experimental condition
-indicates that its design is weak, since we identify only one of perhaps infinitely
-many latent parameter combinations that can generate spurious group polarization.
-
-We use this setup to demonstrate how \emph{adversarial} style testing of experimental
-designs with counterfactual null models can be used to estimate important
-metascience parameters. In this study we use our FWER calculations to estimate
-the minimum Cohen's $d$ that should be used to determine whether an effect is
-``significant'' or not that will limit the FWER to 5\%. In the Discussion we
-sketch how our model can then be extended to rigorously calculate statistical power,
-i.e., the probability that a significant effect is detected given that there 
-is indeed a significant effect.
-
-\subsection{Probability theory of spurious group polarization}
-
-We formalize this experimental design in the $N\to\infty$ case where the
-distribution of ordinal measurements is created by integrating continuous latent
-opinion distributions.  Latent opinions in the model are normally distributed with
-mean $\mu_t$ and variance $\sigma_t$ for time $t$. We are concerned with only two
-times, \emph{pre-} and \emph{post-} deliberation ($t=0$ and $t=T$, respectively).
-For short we can write $\mu_\mathrm{pre}$ and $\sigma_\mathrm{pre}$ for
-pre-deliberation mean and variance (and similar for \emph{post}-deliberation).  
-The latent opinion of participant $i$ at time $t$ is drawn from a normal 
-distribution in the model, 
-\begin{equation} 
-  o_{i,t} \sim p(o; \mu_t, \sigma_t) = \mathcal{N}(\mu_t, \sigma_t).
-  \label{eq:opinionDistribution} 
-\end{equation} 
-\noindent 
-Latent opinions are assumed to be normally distributed since they are unbounded 
-aggregations of neurobiological activity.
-
-The measurement process with $N\to\infty$ participants is simulated by 
-integrating the latent opinion distribution between ordinal bin thresholds, 
-of which there are $B+1$, where $B$ is the number of opinion bins, with sequential
-integer values $b$. In the Moscovici and Zavalloni (1969) example, $b \in
-\{-3,-2,\ldots,2,3\}$ and $B = 7$.
-Bin $b$ is defined by its lower and upper continuous-valued thresholds, 
-$\theta_{b-1}$ and
-$\theta_{b}$, respectively. Thresholds are given by the following equations
-(and visualized as dotted lines in Figure~\ref{fig:distros}).
-\begin{equation}
-\theta_b = \begin{cases}
-  -\infty         & \text{ if } b=0 \\
-  \infty          & \text{ if } b=B \\
-  b + \frac{1}{2} & \text{ otherwise.}
-\end{cases}
-\end{equation}
-\noindent
-With thresholds defined, the frequency of ordinal measurment $b$ at time $t$ 
-is (Figure~\ref{fig:distros})
-\begin{equation}
-  p(b;~\mu_t, \sigma_t, B) = \int_{\theta_{b-1}}^{\theta_b} p(o; \mu_t, \sigma_t)
-  do.
-  \label{eq:ordinal-frequency}
-\end{equation}
-
-We can now define spurious group polarization and give examples of when it occurs.
-First, the observed group polarization opinion shift is
-\begin{equation}
-  g = \meanobs_\mathrm{post} - \meanobs_\mathrm{pre}~,
-  \label{eq:group-polarization}
-\end{equation}
-\noindent
-where the expected value of observed opinions at time $t$ is
-\begin{equation}
-  \meanobst = \frac{1}{B} \sum_{b=1}^B b \cdot p(b; \mu_t, \sigma_t, B).
-  \label{eq:meanobs}
-\end{equation}
-\noindent
-
-\emph{Spurious} group polarization is when mean observed opinions seem to increase in 
-extremity when the latent mean is held constant. Call the collection of
-latent null model distribution parameters $\lambda = (\mu, \sigma_{pre}, \sigma_{post})$.
-The group polarization opinion shift can then be written as a function of $\lambda$,
-$g(\lambda)$. Formally, then, spurious group polarization is plausible if 
-there exists at least one $\lambda$ for which $g(\lambda) = g_{pub} > 0$, where
-$g_{pub}$ is the published group polarization shift value.
-This could be thought of as sampling the observed ordinal distributions 
-$N \to \infty$, collecting ordinal measurements of samples from 
-each latent distribution corresponding to each phase, pre- and post-deliberation.
-
-Next, we explain how to find a set of latent null parameters, $\lambda$, that
-generate spurious group polarization when naively measured on an ordinal scale.
-
-\begin{figure}
-  \caption{\textbf{Opinion measurement model and spurious group polarization.} 
-  Latent model participant opinions are drawn from a normal distribution with the
-  listed parameters (A,C). The pre-deliberation latent distribution has high
-  variance, but appears polarized when measured via integration (A).  The latent
-  distribution's variance is expected to decrease during deliberation (B) via
-  consensus, but in this example the latent mean is static, 
-$\mupre = \mupost$. However, when the post-deliberation distribution is measured,
-the simulated observed mean increases, i.e., spurious
-group polarization occurs (C).}
-  \centering
-  \includegraphics[width=0.95\textwidth]{Figures/Model/latent-ordinal-distros.pdf}
-  \label{fig:distros}
-\end{figure}
-
-\subsection{Interactive algorithm to find parameters inducing spurious group polarization}
-
-To identify plausibly false detections of group polarization, we searched numerically
-for latent parameters $\lambda = (\mu, \sigmapre, and \sigmapost)$ that generate
-spurious group polarization, as explained immediately above. We developed an *ad hoc*
-hillclimbing-style algorithm that finds $\lambda$ that finds $\lambda$ by minimizing
-the difference between published group polarization shifts $g_pub$ and calculated shifts,
-$g_\lambda$.
-calculated with Equation~\ref{eq:meanobs}.
-
-The algorithm to find $\lambda$ holds $\mu$ constant while searching for $\sigmapre$
-and $\sigmapost$ that generate observed $\meanobspre$ and $\meanobspost$, within a
-tolerance of $\pm10^{-3}$ (Algorithm~\ref{alg:hillclimbing}). 
-The algorithm requires a researcher to specify an initial
-guess for $\lambda$ and auxiliary parameter settings, which generally vary between
-experimental conditions.  Furthermore, the algorithm does not always produce
-reasonable distributions with a given $\lambda$, and sometimes it does not produce
-spurious group polarization.  Therefore, it is necessary to sanity check the latent
-and ordinal opinion distributions for the $\lambda$ found by the algorithm. If all is
-well, the experimental condition can be marked as plausibly spurious.  To tame the
-complexity involved in data entry and model fitting for each of these sixty
-experimental conditions, we developed a web application to input data from control
-the algorithm and inspect its results (Figure~\ref{fig:webapp}). Web application code is
-available online in our GitHub
-repository\footnote{\url{https://github.com/mt-digital/gp-statmod/blob/main/GroupPolarizationStatmod/app.R}}.
-
-\begin{algorithm}
-  \caption{Routine to find latent mean and variance that generate spurious group polarization.}
-  \label{alg:hillclimbing}
-  \begin{algorithmic}
-
-    \Require $\mu$, $\sigma$, $\bar{\mathbf{o}}$ 
-      \Comment{Set latent mean, guess variance that yields, input empirical mean.}
-
-    \Require $\epsilon$, $\delta$, $i_\mathrm{max}$ 
-      \Comment{Hillclimbing tolerance, step size, and maximum iterations.}
-
-    \State $\Delta \gets \meanobs(\mu, \sigma) - \meanobsemp$ 
-      \Comment{Initialize simulation error.}
-    \While {$\Delta^2 > \epsilon$ and $i < i_\mathrm{max}$} \Comment{Search until desired
-    accuracy or max iterations reached.}
-      \State $\sigma' \sim \mathcal{N}(\sigma, \delta)$ \Comment{Draw new variance to test.}
-      \State $\Delta' \gets \bar{o}(\mu, \sigma') - \bar{\mathbf{o}}$ 
-        \Comment{Calculate error using new variance.}
-      \If {$\Delta' - \Delta < 0$}
-        \State $\sigma \gets \sigma'$
-          \Comment{Update $\sigma$ if the error is less than before.}
-        \State $\Delta \gets \bar{o}(\mu, \sigma) - \bar{\mathbf{o}}$
-          \Comment{Update simulation error if $\sigma$ is updated.}
-      \EndIf
-    \EndWhile \\
-
-    \Return $\sigma$
-  \end{algorithmic}
-\end{algorithm}
-
-% \begin{figure}
-%   \caption{\textbf{Shiny web application for user-guided hillclimbing to find
-%   parameters that generate spurious group polarization.} This web application enables users to
-% create new analyses of experimental conditions from published studies. Users 
-% provide initial guesses for the algorithm to use as starting points in 
-% finding variances that generate spurious group polarization. Users inspect 
-% model fits provided by the algorithm, adjusting parameters if necessary, and
-% mark an experimental condition as ``Plausibly spurious'' if a reasonable fit
-% is indeed found.\mt{Maybe add some A, B, C guides to the figure and include
-% in parentheses here to go with the text?}}
-%   \centering
-%     \includegraphics[width=0.75\textwidth]{Figures/Model/WebApp.png}
-%   \label{fig:webapp}
-% \end{figure}
-
-Note that so far, we have only identified cases where spurious group polarization plausibly
-occurs in the asymptotic case where $N\to\infty$ ($N$ being the number of participants).
-Experimental conditions yielding plausibly spurious group polarization warrant 
-some suspicion. If there exists set of latent simple consensus 
-parameters $\lambda$ that create binned (Equation~\ref{eq:ordinal-frequency})
-opinion distributions that naively seem to be group polarized, then the means
-of at least some samples from these distibutions with empirical numbers of 
-participants would also appear to be shifted consistent with group
-polarization. This process itself does not tell us the false detection rate.
-We use the identified simple consensus parameters to define pre- and post-deliberation
-latent opinion distributions in simulated trials of group polarization 
-experiments.
-
-\begin{figure}
-  \centering
-    % \includegraphics[width=0.8\textwidth]{Figures/Model/data-model-flow.png}
-    \includegraphics[width=\textwidth]{Figures/Model/data-model-flow-w-webapp.pdf}
-  \caption{\textbf{Workflow of the GP-stats simulation and analysis pipeline
-  for estimating false detection rates under null-consistent conditions.} Colored boxes indicate file types: teal = R scripts, blue = data files, orange = plotting scripts, cyan
-= Slurm batch script. Arrows show the flow from parameter estimation (hillclimbing or Shiny app) to Bayesian null model fitting on the cluster, aggregation of simulated results,
-and final visualization. (D) shows the web app interface that takes researcher guesses at
-simple consensus models that generate observations, and the app passes the guess to initialize
-the hillclimbing search algorithm for finding spurious group polarization.}
-  \label{fig:data-model-flow.png}
-\end{figure}
-
-\subsection{Metascience of the false detection rate in group polarization}
-
-Metascience provides a framework for
-calculating the false discovery rate across experimental conditions,
-aggregated to the level of journal articles and the ten-paper corpus. 
-The false discovery rate is then the Type I error rate relative to the rate of
-positive discoveries, true or false.
-We calculate via simulation and data fitting
-the Type I error rate for the experimental designs used in experimental conditions
-that, under a preliminary large-$N$ model analysis, generated group
-polarization upon ordinal measurement of simple consensus latent opinion 
-distributions. These can then be aggregated to estimate the false discovery rate
-for each journal article and for the ten-article corpus, which is representative of the
-false discovery rate for group polarization studies.
-
-The false detection rate is the scaled probability that a detection, $D$, of
-an effect, $E$, is in fact a false detection and there is \emph{not} an effect,
-denoted $\neg E$. The probability that a detection $D$ has occurred when there
-was no effect, i.e. when $\neg E$, is denoted $\alpha = \Pr(D | \neg E)$.
-This is known as the Type I or family-wise error rate. To calculate the 
-false detection rate ($\fdr$), one must compare $\alpha$ to the statistical power,
-the probability of a detection when there is an effect, denoted $W = \Pr(D | E)$.
-Both these values must be scaled by the rate at which effects occur, denoted $b$.
-Using the notation outlined above, the false detection rate is
-\begin{equation}
-  \fdr = \frac{(1 - b)\alpha}{(1 - b)\alpha + bW}.
-  \label{eq:fdr}
-\end{equation}
-
-In group polarization, $E$ represents the group polarization \emph{effect}, and
-$\neg E$ represents its absence. Simple consensus is one way for group polarization
-to be absent, i.e., simple consensus is one form of $\neg E$. The presence of
-an effect $E$ is independent of whether an effect was \emph{detected}, 
-represented by $D$ if it was detected and $\neg D$ if not.
-
-The group polarization studies are journal articles that present
-the results of multiple experiments to support some hypothesis. To organize
-the presentation of false detection rates across the ten studies in our
-analysis, we first will organize false detection rates by experiment, where
-we created `ExperimentID` tags for each of the sixty experiments across ten
-studies. We also created `StudyID` tags for each of the ten journal articles
-with the key `{Author}{Year}`, e.g., \texttt{Schkade2010} for Schkade, et al.,
-(2010) \cite{Schkade2010}.
-
-We calculate $\alpha = \Pr(D | \neg E)$ through simulation for experimental
-conditions, 
-and use optimistic values of
-$b=0.1$~\cite{Ioannidis2005,McElreath2015} and $W=0.8$~\cite{Smaldino2016NatSel}
-in the main text and in the compare with more pessimistic, yet realistic, estimates
-of these values, as well as even more optimistic estimates.
-
-\subsection{Estimating and limiting false detection rates}
-
-To calculate the false detection rate we begin by simulating 
-pre- and post-deliberation 
-opinions by drawing $N_e$ samples from normal latent opinion
-distributions that generated spurious group polarization, i.e., with
-parameters labelled $\lambda_e$ found in the hillclimbing step. Binning these
-random values simulates observations. $N_e$ is the number
-of participants in experimental condition $e$. Each trial $i$ of $N_e$
-simulated pre- and post-deliberation observations is then inspected to see
-if spurious group polarization occurred in the simulation. Spurious group
-polarization is determined by first inferring the latent parameters of each
-simulated experimental trial by fitting an ordered probit model~\cite{Liddell2018}. 
-The inferred parameters are then used to calculate the effect size for that
-trial, Cohen's $d$, defined as~\cite[p. 331]{Liddell2018}
-\begin{equation}
-  d = \frac{\mupost - \mupre}{\sqrt{\frac{1}{2}(\sigmapost^2 + \sigmapre^2)}}.
-  \label{eq:cohens}
-\end{equation}
-\noindent
-The family-wise error rate is, $\alpha_{e} = \Pr(D|\neg E)$, is the 
-fraction of trials, 
-with $d_{ei} > d^*$ for the $i^\text{th}$ trial for experimental
-condition $e$, written
-\begin{equation}
-  \alpha_e = \frac{1}{N_T} \sum_i^{N_T} \Theta(d_{ei}; d^*),
-  \label{eq:study_aggregate}
-\end{equation}
-\noindent
-where
-\[
-  \Theta(d_{ei}; d^*) = 
-    \begin{cases}
-      1 & \text{ if } d_{ei} > d^* \\
-      0 & \text{ otherwise.}
-    \end{cases}
-\]
-and $N_T = 1000$ is the number of trials. For simplicity we assume that the sign of
-$d_{ei}$ is positive, meaning a shift to greater extremisim. It is possible that
-$d_{ei}$ could be negative, indicating a shift to lesser extremity. Indeed we
-observe many simulations with a shift to lesser extremity. For now we do not
-include these as indicating a ``significant'' effect, which we return to in the
-Analysis below.
-
-To calculate the family-wise error rate
-for a study, $s$ which contains several experimental conditions, $e$,
-\begin{equation}
-  \alpha_s = \frac{1}{|s|} \sum_{e \in s} \alpha_e
-\end{equation}
-\noindent
-The mean across all results identified as plausibly spurious is 
-\begin{equation}
-  \alpha_\mathrm{all} = \frac{1}{|E|} \sum_{e \in E} \alpha_e.
-\end{equation}
-We finally may calculate the false discovery rate by plugging this $\alpha$
-into Equation~\ref{eq:fdr} and assuming empirically-motivated values for 
-base rate, $b$, and statistical power, $W$, as well as more pessimistic and
-optimistic settings for base rate and power.
-
-We close our Analysis by inverting this procedure to find what significance value
-$d^*$ is necessary to achieve a low family-wise error rate $\alpha_e = 0.05$ for
-experimental condition, $e$. To do this we first calculate $\alpha_e(d^*)$ for a
-sufficiently wide range of $d^*$. We then find the lowest $d^*$ for which
-$\alpha_e(d^*) \leq 0.05$.
-
-\section{Analysis}\label{analysis}
-
-We now present our determinations of whether published detections of group
-polarization are plausibly spurious and calculate the best-case false detection
-rate for each experimental design in the corpus. We found that 95\% of all
-detections of group polarization are plausibly false, 54 out of 57. We estimate the
-best-case median false detection rate to be 72\% for those 54 experimental
-conditions. All but seven of the 54 designs were found to require a $d^* > 0.8$
-for ``significant'' Cohen's $d$ to limit the family-wise error rate to 5\%.
-
-\subsection{Nearly all group polarization detections are plausibly spurious}
-
-% We used our formal distribution model of group polarization experiments
-% to identify plausible latent simple consensus distributions that generate spurious
-% group polarization in 
-Of the remaining 57, we identified simple consensus $\lambda$ that generate spurious
-group polarization in 54 experimental conditions. In the other three, we did
-identify $\lambda$ that generate spurious group polarization, but the initial
-simulated distributions were highly polarized. 
-We rejected $\lambda$ generated for
-two spurious group polarization from Schkade, et al., (2010) because
-the histogram plots of the observed opinion distributions are not
-polarized~\cite[Figure 1, p. 234]{Schkade2010}. The other condition where we
-found but rejected spurious group polarization $\lambda$ was in a question 
-about a ``bad teacher'' in Myers (1975)~\cite{Myers1975a}, where lower values
-signify increasingly worse impressions of the teacher and higher values signify
-increasingly good impressions. It is unlikely that a
-``bad teacher'' would be rated as good in this case, so we excluded this 
-from our analysis as well. This exclusion may be overly generous since Myers (1975)
-did not provide details beyond mean values and opinion shifts, and the article was
-published before open data practices. Our corpus had a total of 60 experimental conditions, but we excluded 3 out of 60 
-that did not claim to be positive findings of group polarization.
-
-- Address sigmas, note differences
-
-\begin{figure}
-  \centering
-    \includegraphics[width=\textwidth]{Figures/3.1/latent-sd-cdf.png}
-    \caption{\textbf{}}
-  \label{fig:}
-\end{figure}
-
-\subsection{Half of experiments exceed a 70\% false discovery rate at ``medium'' significance}
-
-We simulated trials from each experimental condition, $e$, by sampling and binning
-pre- and post-deliberation latent opinions from distributions with parameters
-identified with the $N\to\infty$ model as causing plausibly spurious group
-polarization. For each $e$, we drew a number of samples equal to the number of
-participants in the original experiments, $N_\mathrm{emp}$. To estimate family-wise
-error and false detection rates, we fit an ordered probit model to the simulated
-observations and obtained estimates of the group polarization effect size, $d$
-(Equation~\ref{eq:cohens}). If the effect size was greater than a significance
-value, $d > d^*$, then by definition there was a detection $D$ of group
-polarization, even though there was no effect by design, written $\neg E$.  The
-family-wise error rate is $\Pr(D|\neg E)$, calculated as the frequency with which $d
-> d^*$ out of 1000 simulation trials. The false detection rate is the family-wise
-error rate scaled using the base rate of effects and the statistical power
-(Equation~\ref{eq:fdr}). We test three difference significance values, $d =
-0.2,~0.5$ and $0.8$, corresponding to what Cohen prescribed to use for ``low'',
-``medium'', and ``high'' significance. If not otherwise noted, we set the base
-rate to be $b=0.1$ and the power $W = 0.8$ as metascience parameters to calculate 
-the false discovery rate (Equation~\ref{eq:fdr}). Note that if the family-wise error
-rate is low at $\alpha = 0.05$, the false discovery rate is $\fdr = 0.36$.
-
-Different conditions yield different effect size distributions
-(Figure~\ref{fig:effect_size_distros}), which 
-determine the magnitude of error rates and difference between the error rates for low,
-medium, and high significance values $d^*=0.2,0.5,0.8$ magnitude of error rates 
-(Figure~\ref{fig:fwer_fdr_synthesis}B). 
-Family-wise error rates varied from a minimum of one in ten for the
-\texttt{Feminists-Experimental} condition from Myers (1975)~\cite{Myers1975a}
-to a maximum of two in three for the \texttt{COSprings-CivilUnions} condition from
-Schkade, Sunstein, and Hastie (2010)~\cite{Schkade2010} under medium
-significance effect size $d^* = 0.5$ (Figure~\ref{fig:fwer_fdr_synthesis} and
-Table~\ref{tab:quantiles}). The median family-wise error
-rate is 0.28 for Krizan and Baron's (2007) experimental condition
-\texttt{NoOutgroupScenario1}~\cite{Krizan2007}. This translated to a minimum
-false detection rate of 0.52 for the Myers (1975) condition, a maximum false
-detection rate of 0.88 for the Schkade, et al., (2010) condition, and a median false
-detection rate of 0.75 for the Krizan and Baron (2007) condition.
-
-\begin{table}[ht]
-  \caption{{\textbf{Quantiles for the two error-rate measures under different
-  significance values.}}}
-  \label{tab:quantiles}
-  \centering
-  \input{tex/quantile_table_min_20_50_80_max.tex}
-\end{table}
-
-In the conditions
-with lower family-wise error rates, there were several conditions where using the low 
-significance value $\alpha$ resulted in a much greater error rate than the medium or 
-high significance values. In the ten conditions with the lowest $\alpha$ for 
-$d^* = 0.5$, with $\alpha$ near 0.1, five of these conditions have $\alpha > 0.25$
-when $d^* = 0.2$ corresponding to a small effect size. For low $\alpha$, the
-difference in false detection rates for each $d^*$ value are more widely
-distributed, following the distribution of $\alpha$, meaning that 
-a modest increase in $d^*$ can have greater effects on the false discovery rate. In
-the conditions above the 50th percentile, $\alpha$ approaches and exceeds one in
-two, even when requiring ``high'' significance with $d^*=0.8$. With $d^* = 0.8$,
-$\alpha$ approaches and exceeds one in four, with four near one in two. When 
-requiring ``low'' signifance,
-
-Experimental designs within each study or article tend to share certain
-characteristics, such as the number of bins in the ordinal measurement design
-or the number of participants per condition. Therefore, the 
-average false detection rate across experimental trials within a study
-(Equation~\ref{eq:study_aggregate}) show that
-
-\begin{figure}
-
-  \caption{\textbf{Family-wise error rates and false discovery rates across studies and experiments for
-    low, medium, and high Cohen's $d$ significance.}}
-  \label{fig:fwer_fdr_synthesis}
-
-  \centering
-  \includegraphics[width=1.1\textwidth]{Figures/Analysis/fwer_fdr_synthesis.pdf}
-\end{figure}
-
-\begin{itemize}
-  \item 
-    Other $b$ and $W$ values in quantiles.
-  \item
-    Note that there were many ``significant'' results---in the wrong direction
-    (Figure~\ref{fig:OrdinalBoxplot}).
-\end{itemize}
-
-\subsection{High significance values ($d^*$) often necessary for FWER of 0.05}
-
-Our simulations can be used a different way, in this case to calculate what
-significance value $d^*$ limits the family-wise error rate to the low value of
-0.05. 
-Solving the inverse problem of which $d^*$ achieves a low family-wise error rate,
-which drives the false detection rate. Here we present our calculations to find
-$d^*$ that achieve $\alpha(d^*) = 0.05$.
-
-\begin{figure}
-  \caption{
-    \textbf{Significance value, $d^*$, to limit family-wise error 
-    rate of 0.05 for each identified study.} The Moscovici and Zavalloni (1969)
-    ``Americans'' question would require the greatest $d^*$ to achieve $\alpha \leq
-    0.05$, even though it does not have the highest error rates. This is because of
-    larger outliers in simulated $d$ for this condition compared to others with
-    greater error rates (Figure~\ref{fig:OrdinalBoxplot}).
-  }
-  \centering
-    \includegraphics[width=0.75\textwidth]{Figures/Analysis/sigval_for_low_fwer.pdf}
-  \label{fig:sigval_for_low_fwer}
-\end{figure}
-
-\section{Discussion}\label{discussion}
-
-\mt{Restate results and importance and outline this section.}
-
-Since there seems to be no selection pressure on false discovery rates,
-experiment design for studying group polarization was free to evolve
-randomly. Various choices for the number of bins and other instrumental
-details may have then become entrenched in different communities defined by
-common theoretical perspectives taken by researchers in those communities.
-Therefore, when Isenberg (1986)~\cite{Isenberg1986} reports that his
-meta-analysis found more support for the persuasive arguments theory than the 
-social comparisons theory in the form of greater effect sizes. If this were
-indeed the case, perhaps it is only because some researchers hypothesizing
-persuasive arguments happened to use a measurement procedure produced more
-frequent and extreme distortions of opinion measurements. Incentive
-structures known to promote the ``evolution of bad science'' would select
-for research that ignored sticky, time-consuming statistical problems in
-favor of juicy titles and headlines for which practitioners have won several
-professional awards. Specifically, the traditional bias
-towards positive results, and against negative ones, could easily have
-provided a cultural niche in which quick but faulty group polarization
-statistical methods dominate slower, more rigorous ones.
-Now that we are aware of the problem, there must be
-intense selective pressure applied as soon and as widely as possible to 
-avoid further ``replications'' that report spurious group polarization, or
-any similar latent psychological dynamics.
-
-We provided tools that experimenters can use. While these tools are
-potentially useful, they are still in the prototype phase that was
-sufficient for a single researcher who was also the software developer. So,
-a first step towards expanding the usefulness of the findings here is to 
-improve the software usability and make the app widely available for
-researchers in group polarization and beyond. Another step would be to go
-beyond reactionary testing of the reliability of experimental designs, and
-identify design principles that reduce the false detection rate without
-needing to adjust the Cohen's $d$ (or whatever measure) 
-used as a significance threshold.
-
-When we used the appropriate ordered probit Bayesian statistical model, it
-still often failed to accurately identify mere agreement, instead detecting
-group polarization.  The ordered probit model we used still likely
-underestimates the variance in outcomes, which could require even larger
-sample sizes. A more rigorous analysis would include additional sources of
-variance that will likely further inflate the Type I error rate, $\alpha$,
-and the false discovery rate~\cite{Yarkoni2022}. First and foremost, there is
-variance within each experimental deliberation group; no group polarization
-study accounts for this, instead treating participants in each experimental
-condition as if they were one large group. Furthermore, survey data are known
-to be noisy~\cite{Zaller1992}. People report opinions differently over time
-for no apparent reason (i.e., opinions are \emph{unstable}).  Context
-matters: the order in which a survey question is asked, and which question
-framing is chosen among logically equivalent alternatives, can both
-significantly influence participant responses. Finally, little is known about
-the psychological process that converts latent opinions to reporting
-behaviors (e.g., clicking a radio button corresponding to an opinion).  The
-effect of accounting for these sources of variance must be understood to
-estimate $\alpha$ (and power, $W$) for the design of the next generation of
-group polarization experiments.
-
-A simple change in measurement procedure could fix the problem identified
-here: center the post-discussion scale on each group’s own pre-discussion
-mean. This removes the assumption that all groups share a common “neutral”
-midpoint. The situation is like tossing a ball on a moving train: to the
-passengers, the ball’s speed is the same in both directions, but to an
-observer on the ground one throw appears faster and the other slower.
-Likewise, when measurement is anchored to an external scale, simple agreement
-within a group can appear as movement toward one pole. Re-centering the
-measurement frame corrects this distortion. It acknowledges that opinion
-change is relative to the observer’s reference frame—there is no fixed
-background of neutrality against which all groups can be compared. This is,
-in essence, the insight of Newtonian relativity: apparent motion depends on
-the frame from which it is measured.
-
-First we must note that there are even more, perhaps even graver problems
-with group polarization research than I discussed so far: one of note is
-opinion instability. Opinions are not generally stable—people seem to
-construct their opinions based on several factors not directly related to
-whatever underlying topic. Factors affecting reported opinions include the
-language used to frame a prompt or question; the order of questions on a
-multi-item survey; or simply when the question was asked (Kalton and Schuman,
-1982; Zaller and Feldman, 1992; Zaller 1992). Failure to account for sources
-of input variance in any experiment can lead to inflated effect sizes since
-the input variance is missing from the posterior predictions during
-statistical model fitting (Yarkoni, 2022). This further compounds the
-theoretical confusions that fill the void when theories are never formalized
-and grounded in a mathematical or computational model (Smaldino 2020; Turner
-and Smaldino, 2022).
-
-A second additional problem with group polarization research is that we literally cannot learn anything by comparing summary statistics obtained with incommensurate experimental designs—but this is exactly how group polarization research has tried to make progress.
-Comparing summary statistics only has explanatory power if datasets were generated under the same experimental designs—it seems this was never done or even considered in the group polarization literature.
-In order to compare studies, their datasets must be generated from commensurate experiments, meaning the studies must share the same predictive and statistical models and experimental design.
-Otherwise, such comparisons are meaningless, they literally tell us nothing, as shown by philosopher of science Nancy Cartwright in Chapter 5 of her book The Dappled World: A Study of the Boundaries of Science.
-
-\subsection{Humble Hearts Promote Rigorous Science}
-
-Cass Sunstein kicked off the 2000s by elevating group polarization to be a scientific law and advocating the application of group polarization research to serious real-world problems in two books (Sunstein, 2009; Sunstein, 2019) and two recent high-profile articles in Nature Human Behavior: one on the social foundations of pandemic preparedness (Van Bavel, et al., 2020) and one to “promote truth, autonomy, and democratic discourse online” (Lorenz-Spreen, Lewandowski, Sunstein, and Hertwig, 2020).
-Let us try to stop the spread of this bad science—and let us explore what must be done to achieve Dorian Cartwright’s vision of a rigorous science of group polarization he articulated fifty years ago.
-
-To be practically useful, science must be rigorous.
-Rigorous science comes most naturally when scientists remain humble, especially when they serve at the pleasure of the public, the government, or other donors.
-Here’s another example from Burnstein and Vinokur (1973) that we can learn from—let’s try to avoid the sort of arrogance and haughtiness that seep from these acknowledgements they wrote on p.
-123:
-
-\begin{quote}
-This research was supported by a Grant from the National Institute of Mental Health (MH-16950-03) and by a Guggenheim Foundation Fellowship awarded to the first author.
-We gratefully acknowledge the incisive, witty, and vinous commentary provided by colleagues at the University of Provence, in particular Robert Abelson, Claude Flament, and Jean-Pierre Poitou, as well as the spirited and intelligent efforts of Livia Mezrick and Irene Graczyk, who helped us carry out the study.
-\end{quote}
-
-Science is not glamorous—it takes guts and sacrifice to achieve the technical skills and discipline to sharpen one’s prose, methods, mathematics, and computer code to the point they can really cut reality at its joints, investigate different parts, then put it all back together with a message for others about how the world really works.
-These days, to do rigorous social science work one must wisely curate the best of the available social science by harmonizing cross-disciplinary jargon.
-But disincentives still exist against careful scholarship.
-Prizes and professorships go to those with the most numerous publications in the right journals—period.
-
-Unpaid, anonymous reviewers are supposed to act as gatekeepers to catch problems like the ones I reviewed.
-But, journal article reviewers have little to gain from enforcing rigor—actually there is a lazy strategy that can pay off big for freeloaders.
-It’s easy to imagine a cultural norm evolving where reviewers save themselves time and energy by asking little of an author, in hopes that little will be demanded of them when the reviewer submits their next paper.
-This could result in a publication process that rewards the laziest and most willing to flood the zone and self-aggrandize.
-
-\subsection{Conclusion}
-
-Social science can use physics as a model discipline to transition from idiosyncratic and Balkanized to informative and united.
-Physics only changed the world because it is useful—it is useful only because it reliably predicts things.
-Physics became more useful as it became more united, finding shared representations for concepts developed independently—for example the science of optics was developed independently from the science of dynamics, but the physics of light is explained partly in analogy to a pendulum or spring—the mathematics of harmonic motion can be used to predict the motion of springs or pendulums—or why we see different colors of light through a prism.
-With careful, coordinated tweaks to group polarization experimental design and theoretical modeling, we will build a more rigorous theory of group polarization and its large-scale expression, echo chamber radicalization.
-
-I expect that group polarization experiments, when properly constructed and analyzed, will help sustainability campaigns optimize their social influence campaigns.
-We can start by developing simpler experimental designs based on formal or computational models capable of predicting outcomes in an experiment.
-I have developed a model that could motivate an experimental design: I showed that if extremists are more stubborn, this is sufficient to cause group polarization when likeminded groups interact over time.
-There is even a free stubborn extremism parameter that could be fit to data to measure the relationship between stubbornness and extremism on different issues.
-
-Group polarization research so far has mostly provided us with a checklist of what not to do.
-The alternative approach I outline builds on rock—the rock of rigor—and will supersede past group polarization research built on a thousand grains of inexact verbal speculation.
+\input{discussion.tex}
 

### 2025-10-31 GroupPolarizationStatmod/paper/method.tex
update method
updated method

@@
 corpus. The first step in the analysis is to find a reasonable pair of latent
-null-polarization distributions that appear polarized due to clipping.  To do
+null-polarization distributions that appear polarized due to clipping.
+
+\subsection{Identify replications for removal}
+
+The first step is to find which replications of the 57 must be retracted 
+due to non-identifiability. We establish non-identifiability by finding one
+set of latent null-polarization opinion distributions where the mean doesn't change,
+but appear polarized when measured on the ordinal scale used in the experiment.
+
+To do
 this we designed a web app that allows the user to try different latent
@@
 experiment's ordinal measurement procedure, or the algorithm stops after a
-certain number of steps (Algorithm~\ref{alg:hillclimbing}).
+certain number of steps (Algorithm~\ref{alg:hillclimbing}). If the web app
+finds
 
@@
 
-\begin{figure}
+\begin{figure}[ht]
   \caption{\textbf{Opinion measurement model and spurious group polarization.} 

### 2025-10-31 GroupPolarizationStatmod/paper/method.tex
updates to method

@@
 corpus. The first step in the analysis is to find a reasonable pair of latent
-null-polarization distributions that appear polarized due to clipping.
+null-polarization distributions that appear polarized due to clipping
+(Section~\ref{subsec:identify}). To do this, we develop a formal model and a numeric
+rootfinding algorithm to power a Shiny web app we used to interactively identify and
+record null-polarization parameters generating such measurement artifacts.
+
+The second step in the analysis is to evaluate the quality of the experimental designs
+when we properly account for the ordinal measurement procedure
+(Section~\ref{subsec:evaluate}). We use the null-polarization parameters we found in
+the first step to generate simulated experimental data, then estimate the difference
+between the pre- and post-discussion means using ordered probit models to fit each
+distribution—ordered probit models account for the ordinal measurement procedure.  We
+estimate the false positive rate and false detection rate for low, medium, and high
+significance thresholds in terms of Cohen's $d$. Finally, we use the effect sizes thus
+calculated to determine the significance threshold one must use to achieve a 5\% false
+positive rate.
 
 \subsection{Identify replications for removal}
+\label{subsec:identify}
 
-The first step is to find which replications of the 57 must be retracted 
-due to non-identifiability. We establish non-identifiability by finding one
-set of latent null-polarization opinion distributions where the mean doesn't change,
-but appear polarized when measured on the ordinal scale used in the experiment.
+The first step is to find which replications of the 57 must be retracted due to
+non-identifiability. We establish non-identifiability by finding one set of latent
+null-polarization opinion distributions where the mean doesn't change, but appear
+polarized when measured on the ordinal scale used in the experiment.  In other words,
+if we find a null-polarization model that explains the outcome as well as a group
+polarization one, the group polarization hypothesis is \emph{non-identifiable} and the
+replication must be rejected—the data cannot confirm or disconfirm the target
+hypothesis.
 
-To do
-this we designed a web app that allows the user to try different latent
+To do this we designed a web app that allows the user to try different latent
 distribution parameters (Figure~\ref{fig:data-model-flow}) to seed a
@@
 experiment's ordinal measurement procedure, or the algorithm stops after a
-certain number of steps (Algorithm~\ref{alg:hillclimbing}). If the web app
-finds
+certain number of steps (Algorithm~\ref{alg:hillclimbing}).
 
-—latent mean and pre- and post-discussion 
-variances that  loads experimental design and observation data
-from a user-provided CSV file
+% To find examples of plausibly spurious group polarization we guess at what latent
+% mean, combined with a latent pre- and post-deliberation latent standard deviations,
+% could generate it. We calculate the observed group polarization for these latent
+% parameters by transforming the latent pre- and post-deliberation distributions from
+% continuous normal distributions to ordinal ``observations'' by integrating over each
+% ordinal bin, then calculating the observed mean for each experiment \emph{phase},
+% i.e., both the pre- and post-deliberation phases.  If the null model can generate
+% group polarization in this case, then the detection is labelled \emph{plausibly
+% false}, i.e., \emph{plausibly spurious}, and we call the generating latent
+% parameters the \emph{plausible null parameters}.
 
-ordinal measurements distort to look 
-polarized.
-
-To find examples of plausibly spurious group polarization we guess at what latent
-mean, combined with a latent pre- and post-deliberation latent standard deviations,
-could generate it. We calculate the observed group polarization for these latent
-parameters by transforming the latent pre- and post-deliberation distributions from
-continuous normal distributions to ordinal ``observations'' by integrating over each
-ordinal bin, then calculating the observed mean for each experiment \emph{phase},
-i.e., both the pre- and post-deliberation phases.  If the null model can generate
-group polarization in this case, then the detection is labelled \emph{plausibly
-false}, i.e., \emph{plausibly spurious}, and we call the generating latent
-parameters the \emph{plausible null parameters}.
-
-To evaluate whether the plausible null parameters we identify could realistically
-fool a significance test, we implemented a generative simulation of each plausibly
-false detection of group polarization and calculated the \emph{family-wise error rate
-(FWER)} and the \emph{false discovery rate (FDR)}. We use the published number of
-participants $N$ for each of the 54 conditions that yielded plausibly false
-detections of group polarization. The FWER is also known as the Type I
-error, or the false positive rate: it is the probability that a detection of an
-effect occurs given that in reality there was a null effect.  The FDR is the
-probability that a published positive detection of an effect is a false positive,
-i.e., a detection of a null effect. High FWER and FDR for an experimental condition
-indicates that its design is weak, since we identify only one of perhaps infinitely
-many latent parameter combinations that can generate spurious group polarization.
+% To evaluate whether the plausible null parameters we identify could realistically
+% fool a significance test, we implemented a generative simulation of each plausibly
+% false detection of group polarization and calculated the \emph{family-wise error rate
+% (FWER)} and the \emph{false discovery rate (FDR)}. We use the published number of
+% participants $N$ for each of the 54 conditions that yielded plausibly false
+% detections of group polarization. The FWER is also known as the Type I
+% error, or the false positive rate: it is the probability that a detection of an
+% effect occurs given that in reality there was a null effect.  The FDR is the
+% probability that a published positive detection of an effect is a false positive,
+% i.e., a detection of a null effect. High FWER and FDR for an experimental condition
+% indicates that its design is weak, since we identify only one of perhaps infinitely
+% many latent parameter combinations that can generate spurious group polarization.
 
@@
 
-\subsection{Probability theory of spurious group polarization}
+\subsubsection{Probability theory of spurious group polarization}
 
-We formalize this experimental design in the $N\to\infty$ case where the
-distribution of ordinal measurements is created by integrating continuous latent
-opinion distributions.  Latent opinions in the model are normally distributed with
+% We formalize this experimental design in the $N\to\infty$ case where the
+% distribution of ordinal measurements is created by integrating continuous latent
+% opinion distributions.
+
+Latent opinions are assumed to be normally distributed in general since they are
+aggregations of many neurobiological potentials of
+different polarities and magnitudes. We have two opinion distributions for 
+two different times in the group polarization experiment process: pre-discussion,
+$\tpre$, and post-discussion, $\tpost$.
+The 
 mean $\mu_t$ and variance $\sigma_t$ for time $t$. We are concerned with only two
@@
 \noindent
-Latent opinions are assumed to be normally distributed since they are unbounded 
-aggregations of neurobiological activity.
 
@@
 
-\subsection{Interactive algorithm to find parameters inducing spurious group polarization}
+\subsubsection{Interactive algorithm to find parameters inducing spurious group polarization}
 
@@
 
-\subsection{Metascience of the false detection rate in group polarization}
+\subsection{Evaluate experimental designs}
 
-Metascience provides a framework for
-calculating the false discovery rate across experimental conditions,
-aggregated to the level of journal articles and the ten-paper corpus. 
-The false discovery rate is then the Type I error rate relative to the rate of
-positive discoveries, true or false.
 We calculate via simulation and data fitting
@@
 
+The false discovery rate is then the Type I error rate relative to the rate of
+positive discoveries, true or false.
 The group polarization studies are journal articles that present
@@
 
-\subsection{Estimating and limiting false detection rates}
+\subsubsection{Estimating and limiting false detection rates}
 

### 2025-11-03 GroupPolarizationStatmod/paper/method.tex
method updates

--- 2025-10-31 GroupPolarizationStatmod/paper/method.tex
@@
 
-To do
-this we designed a web app that allows the user to try different latent
+To do this we designed a web app that allows the user to try different latent
 distribution parameters (Figure~\ref{fig:data-model-flow}) to seed a
@@
 certain number of steps (Algorithm~\ref{alg:hillclimbing}). If the web app
-finds
-
-—latent mean and pre- and post-discussion 
-variances that  loads experimental design and observation data
-from a user-provided CSV file
-
-ordinal measurements distort to look 
-polarized.
+finds a constant latent mean and pre- and post-discussion 
+variances to specify a pre- and post-discussion opinion distributions that
+generate spurious group polarization, this is communicated to the user who
+can then mark the experimental condition as such.
 
 To find examples of plausibly spurious group polarization we guess at what latent
-mean, combined with a latent pre- and post-deliberation latent standard deviations,
-could generate it. We calculate the observed group polarization for these latent
-parameters by transforming the latent pre- and post-deliberation distributions from
-continuous normal distributions to ordinal ``observations'' by integrating over each
-ordinal bin, then calculating the observed mean for each experiment \emph{phase},
-i.e., both the pre- and post-deliberation phases.  If the null model can generate
-group polarization in this case, then the detection is labelled \emph{plausibly
+mean, combined with a latent pre- and post-deliberation latent standard
+deviations, could generate it. We calculate the observed group polarization
+for these latent parameters by transforming the latent pre- and
+post-deliberation distributions from continuous normal distributions to
+ordinal ``observations'' by integrating over each ordinal bin, then
+calculating the observed mean for each experiment \emph{phase}, i.e., both
+the pre- and post-deliberation phases.  If the null model can generate group
+polarization in this case, then the detection is labelled \emph{plausibly
 false}, i.e., \emph{plausibly spurious}, and we call the generating latent
@@
 
-To evaluate whether the plausible null parameters we identify could realistically
-fool a significance test, we implemented a generative simulation of each plausibly
-false detection of group polarization and calculated the \emph{family-wise error rate
-(FWER)} and the \emph{false discovery rate (FDR)}. We use the published number of
-participants $N$ for each of the 54 conditions that yielded plausibly false
-detections of group polarization. The FWER is also known as the Type I
-error, or the false positive rate: it is the probability that a detection of an
-effect occurs given that in reality there was a null effect.  The FDR is the
-probability that a published positive detection of an effect is a false positive,
-i.e., a detection of a null effect. High FWER and FDR for an experimental condition
-indicates that its design is weak, since we identify only one of perhaps infinitely
-many latent parameter combinations that can generate spurious group polarization.
+To evaluate whether the plausible null parameters we identify could
+realistically fool a significance test, we implemented a generative
+simulation of each plausibly false detection of group polarization and
+calculated the \emph{family-wise error rate (FWER)} and the \emph{false
+discovery rate (FDR)}. We use the published number of participants $N$ for
+each of the 54 conditions that yielded plausibly false detections of group
+polarization. The FWER is also known as the Type I error, or the false
+positive rate: it is the probability that a detection of an effect occurs
+given that in reality there was a null effect.  The FDR is the probability
+that a published positive detection of an effect is a false positive, i.e., a
+detection of a null effect. High FWER and FDR for an experimental condition
+indicates that its design is weak, since we identify only one of perhaps
+infinitely many latent parameter combinations that can generate spurious
+group polarization.
 
-We use this setup to demonstrate how \emph{adversarial} style testing of experimental
-designs with counterfactual null models can be used to estimate important
-metascience parameters. In this study we use our FWER calculations to estimate
-the minimum Cohen's $d$ that should be used to determine whether an effect is
-``significant'' or not that will limit the FWER to 5\%. In the Discussion we
-sketch how our model can then be extended to rigorously calculate statistical power,
-i.e., the probability that a significant effect is detected given that there 
-is indeed a significant effect.
+We use this setup to demonstrate how \emph{adversarial} style testing of
+experimental designs with counterfactual null models can be used to estimate
+important metascience parameters. In this study we use our FWER calculations
+to estimate the minimum Cohen's $d$ that should be used to determine whether
+an effect is ``significant'' or not that will limit the FWER to 5\%. In the
+Discussion we sketch how our model can then be extended to rigorously
+calculate statistical power, i.e., the probability that a significant effect
+is detected given that there is indeed a significant effect.
 
@@
 
-% \begin{figure}
-%   \caption{\textbf{Shiny web application for user-guided hillclimbing to find
-%   parameters that generate spurious group polarization.} This web application enables users to
-% create new analyses of experimental conditions from published studies. Users 
-% provide initial guesses for the algorithm to use as starting points in 
-% finding variances that generate spurious group polarization. Users inspect 
-% model fits provided by the algorithm, adjusting parameters if necessary, and
-% mark an experimental condition as ``Plausibly spurious'' if a reasonable fit
-% is indeed found.\mt{Maybe add some A, B, C guides to the figure and include
-% in parentheses here to go with the text?}}
-%   \centering
-%     \includegraphics[width=0.75\textwidth]{Figures/Model/WebApp.png}
-%   \label{fig:webapp}
-% \end{figure}
-
-Note that so far, we have only identified cases where spurious group polarization plausibly
-occurs in the asymptotic case where $N\to\infty$ ($N$ being the number of participants).
-Experimental conditions yielding plausibly spurious group polarization warrant 
-some suspicion. If there exists set of latent simple consensus 
-parameters $\nu$ that create binned (Equation~\ref{eq:ordinal-frequency})
-opinion distributions that naively seem to be group polarized, then the means
-of at least some samples from these distibutions with empirical numbers of 
-participants would also appear to be shifted consistent with group
-polarization. This process itself does not tell us the false detection rate.
-We use the identified simple consensus parameters to define pre- and post-deliberation
-latent opinion distributions in simulated trials of group polarization 
-experiments.
+Note that so far, we have only identified cases where spurious group
+polarization plausibly occurs in the asymptotic case where $N\to\infty$ ($N$
+being the number of participants).  Experimental conditions yielding
+plausibly spurious group polarization warrant some suspicion. If there exists
+set of latent simple consensus parameters $\nu$ that create binned
+(Equation~\ref{eq:ordinal-frequency}) opinion distributions that naively seem
+to be group polarized, then the means of at least some samples from these
+distibutions with empirical numbers of participants would also appear to be
+shifted consistent with group polarization. This process itself does not tell
+us the false detection rate.  We use the identified simple consensus
+parameters to define pre- and post-deliberation latent opinion distributions
+in simulated trials of group polarization experiments.
 

### 2025-11-03 GroupPolarizationStatmod/paper/method.tex

@@
 corpus. The first step in the analysis is to find a reasonable pair of latent
-null-polarization distributions that appear polarized due to clipping.
+null-polarization distributions that appear polarized due to clipping
+(Section~\ref{subsec:identify}). To do this, we develop a formal model and a numeric
+rootfinding algorithm to power a Shiny web app we used to interactively identify and
+record null-polarization parameters generating such measurement artifacts.
+
+The second step in the analysis is to evaluate the quality of the experimental designs
+when we properly account for the ordinal measurement procedure
+(Section~\ref{subsec:evaluate}). We use the null-polarization parameters we found in
+the first step to generate simulated experimental data, then estimate the difference
+between the pre- and post-discussion means using ordered probit models to fit each
+distribution—ordered probit models account for the ordinal measurement procedure.  We
+estimate the false positive rate and false detection rate for low, medium, and high
+significance thresholds in terms of Cohen's $d$. Finally, we use the effect sizes thus
+calculated to determine the significance threshold one must use to achieve a 5\% false
+positive rate.
 
 \subsection{Identify replications for removal}
+\label{subsec:identify}
 
-The first step is to find which replications of the 57 must be retracted 
-due to non-identifiability. We establish non-identifiability by finding one
-set of latent null-polarization opinion distributions where the mean doesn't change,
-but appear polarized when measured on the ordinal scale used in the experiment.
+The first step is to find which replications of the 57 must be retracted due to
+non-identifiability. We establish non-identifiability by finding one set of latent
+null-polarization opinion distributions where the mean doesn't change, but appear
+polarized when measured on the ordinal scale used in the experiment.  In other words,
+if we find a null-polarization model that explains the outcome as well as a group
+polarization one, the group polarization hypothesis is \emph{non-identifiable} and the
+replication must be rejected—the data cannot confirm or disconfirm the target
+hypothesis.
 
@@
 
-We formalize this experimental design in the $N\to\infty$ case where the
-distribution of ordinal measurements is created by integrating continuous latent
-opinion distributions.  Latent opinions in the model are normally distributed with
+% We formalize this experimental design in the $N\to\infty$ case where the
+% distribution of ordinal measurements is created by integrating continuous latent
+% opinion distributions.
+
+Latent opinions are assumed to be normally distributed in general since they are
+aggregations of many neurobiological potentials of
+different polarities and magnitudes. We have two opinion distributions for 
+two different times in the group polarization experiment process: pre-discussion,
+$\tpre$, and post-discussion, $\tpost$.
+The 
 mean $\mu_t$ and variance $\sigma_t$ for time $t$. We are concerned with only two
@@
 
-\subsection{Metascience of the false detection rate in group polarization}
-
-Metascience provides a framework for
-calculating the false discovery rate across experimental conditions,
-aggregated to the level of journal articles and the ten-paper corpus. 
-The false discovery rate is then the Type I error rate relative to the rate of
-positive discoveries, true or false.
 We calculate via simulation and data fitting

### 2025-11-06 GroupPolarizationStatmod/paper/article-table.tex
updates

@@
 
-\rowcolor{gray!10}
+\rowcolor{gray!5}
 \multirowcell{2}{Politics} 
                   & Moscovici \& Zavalloni (1969)~\cite{Moscovici1969} & The Group as a Polarizer of Attitudes & \emph{JPSP} \\
-\rowcolor{gray!10}
+\rowcolor{gray!5}
                   & Schkade, Sunstein, \& Hastie (2010)~\cite{Schkade2010} & When Deliberation Produces Extremism & \emph{Critical~Review} \\[0.2em]
 
-\rowcolor{gray!5}
 \multirowcell{2}{Social Comp's \vspace{0.7em}}
@@
                   & Myers \& Bishop (1970)~\cite{Myers1970} & Discussion Effects on Racial Attitudes & \emph{Science} \\[0.2em]
-\rowcolor{gray!5}
                   & Myers (1975)~\cite{Myers1975} & Discussion-Induced Attitude Polarization & \emph{Human Relations} \\[0.3em]
 
-\rowcolor{gray!10}
+\rowcolor{gray!5}
 \multirowcell{2}{Persuasive Arg's}
                   & Burnstein \& Vinokur (1973)~\cite{Burnstein1973} & Testing Two Classes of Theories About Group Induced Shifts In Individual Choice & \emph{JESP} \\
-\rowcolor{gray!10}
+\rowcolor{gray!5}
                   & Burnstein \& Vinokur (1975)~\cite{Burnstein1975} & What a person thinks upon learning he has chosen differently from others: Nice evidence for the persuasive-arguments\ldots & \emph{JESP} \\
 
-\rowcolor{gray!5}
 \multirowcell{3}{Self-Categ'n}
@@
 Psy.} \\
-\rowcolor{gray!5}
                   & Hogg, Turner, \& Davidson (1990)~\cite{Hogg1990} & Polarized Norms and Social Frames of Reference: A Test of the Self-Categorization Theory of Group Polarization & \emph{B.
@@
 Psy.} \\
-\rowcolor{gray!5}
                   & Krizan \& Baron (2007)~\cite{Krizan2007} & Group Polarization and Choice Dilemmas: How Important is Self-Categorization? & \emph{Eur. J. of Soc. Psy.} \\
 
-\rowcolor{gray!10}
+\rowcolor{gray!5}
 \multirowcell{1}{Soc'l Decisions}

### 2025-11-06 GroupPolarizationStatmod/paper/intro.tex
updates

--- 2025-10-26 GroupPolarizationStatmod/paper/intro-latest.tex
@@
 
-If an extremist’s opinion falls and it is measured with a Likert scale, will
-it make a noise? The answer is often no~\cite{Liddell2018,Turner2020}. This
-means a group’s opinions can appear to radicalize over time even when no real
-shift occurs. The problem matters because Likert and other ordinal scales are
-ubiquitous in research on opinion change and radicalization across the social
-and cognitive sciences~\cite{Liddell2018}. One influential framework in this
-area is group polarization—the idea that discussion among like-minded people
-pushes opinions toward extremes, producing ``echo chambers.'' Cass Sunstein, a
+If an extremist moderates their opinion and it is measured with a Likert scale, will
+this move to moderation be detected? The answer is often no in social psychology 
+experiments~\cite{Liddell2018}. This
+means that a group can appear to become radicalized over time even if the 
+average opinion does not actually change~\cite{Turner2020}. 
+This has huge consequences for the study of \emph{group polarization}, the hypothesis
+that discussion among like-minded people pushes opinions toward extremes—effectively 
+making it the science of ``echo chambers'', online or in-person communities 
+composed of likeminded folks.  Cass Sunstein, a
 Harvard Law professor and former White House administrator, has written

### 2025-11-06 GroupPolarizationStatmod/paper/method.tex
updates

@@
 
-\begin{table}[h]
-  \caption{\textbf{Latent psychological variables and distributions.}}
-  \begin{tabular}{cll} 
-  \toprule
-    \textbf{Variable} & \textbf{Description} & \textbf{Values} \\
-  \midrule  
-    $\omega_{i,t}$ & Latent opinion of individual $i$ at time $t$ & $\mathbb{R}$ \\
-    $\mu_t$ & Mean latent opinion at time $t$ (constant in null-polarization) & $\mathbb{R}$ \\
-    $\sigma_t$ & Variance of observed opnions at time $t$ & $\mathbb{R}$ \\
-  \bottomrule
-  \end{tabular} 
-\end{table}
-
-\begin{table}[h]
-  \caption{\textbf{Experimental design and numerical model parameters.}}
-  \begin{tabular}{cll} 
-  \toprule
-    \textbf{Variable} & \textbf{Description} & \textbf{Values} \\
-  \midrule  
-    $B$ & Number of bins & $5,6,7,\ldots$ \\
-    $\theta_b$ & The $B+1$ threshold values that separate bins & 
-      e.g., $-\infty, -2.5, -1.5,\ldots,2.5, \infty$ \\
-    $b$ & Bin index—there are $B+1$ to mark $B$ bins & $0,1,2,\ldots,B$ \\
-  \bottomrule
-  \end{tabular} 
-\end{table}
-
-\begin{table}[h]
-  \caption{\textbf{Simulated observed variables.}}
-  \begin{tabular}{cll} 
-  \toprule
-    \textbf{Variable} & \textbf{Description} & \textbf{Values} \\
-  \midrule  
-    $\meanobst$ & Mean of all reported opinions (Eq. \ref{eq:meanobs}) & $\mathbb{R}$ \\
-    $g$ & Observed group opinion polarization (Eq. \ref{eq:group-polarization})
-        & $\mathbb{R}$ \\
-    $d$ & Cohen's $d$ used as significance threshold 
-        & $\mathbb{R}$ \\
-    $\alpha_e,\alpha_s,\alpha_{\text{all}}$ & Family-wise/Type I error rate
-    for an experiment, study, or across all experiments
-             & $[0, 1]$ \\
-    FDR & False discovery rate
-        & $[0, 1]$ \\
-  \bottomrule
-  \end{tabular} 
-\end{table}
+\input{method-tables.tex}
 

### 2025-11-06 GroupPolarizationStatmod/paper/method-tables.tex
added method tables

\begin{table}[h]
  \caption{\textbf{Latent psychological variables and distributions.}}
  \begin{tabular}{cll} 
  \toprule
    \textbf{Variable} & \textbf{Description} & \textbf{Values} \\
  \midrule  
    $\omega_{i,t}$ & Latent opinion of individual $i$ at time $t$ & $\mathbb{R}$ \\
    $\mu_t$ & Mean latent opinion at time $t$ (constant in null-polarization) & $\mathbb{R}$ \\
    $\sigma_t$ & Variance of observed opnions at time $t$ & $\mathbb{R}$ \\
  \bottomrule
  \end{tabular} 
\end{table}

\begin{table}[h]
  \caption{\textbf{Experimental design and numerical model parameters.}}
  \begin{tabular}{cll} 
  \toprule
    \textbf{Variable} & \textbf{Description} & \textbf{Values} \\
  \midrule  
    $B$ & Number of bins & $5,6,7,\ldots$ \\
    $\theta_b$ & The $B+1$ threshold values that separate bins & 
      e.g., $-\infty, -2.5, -1.5,\ldots,2.5, \infty$ \\
    $b$ & Bin index—there are $B+1$ to mark $B$ bins & $0,1,2,\ldots,B$ \\
  \bottomrule
  \end{tabular} 
\end{table}

\begin{table}[h]
  \caption{\textbf{Simulated observed variables.}}
  \begin{tabular}{cll} 
  \toprule
    \textbf{Variable} & \textbf{Description} & \textbf{Values} \\
  \midrule  
    $\meanobst$ & Mean of all reported opinions (Eq. \ref{eq:meanobs}) & $\mathbb{R}$ \\
    $g$ & Observed group opinion polarization (Eq. \ref{eq:group-polarization})
        & $\mathbb{R}$ \\
    $d$ & Cohen's $d$ used as significance threshold 
        & $\mathbb{R}$ \\
    $\alpha_e,\alpha_s,\alpha_{\text{all}}$ & Family-wise/Type I error rate
    for an experiment, study, or across all experiments
             & $[0, 1]$ \\
    FDR & False discovery rate
        & $[0, 1]$ \\
  \bottomrule
  \end{tabular} 
\end{table}

### 2025-11-14 GroupPolarizationStatmod/paper/method-tables.tex
method rewrite commence

@@
   \midrule  
-    $\meanobst$ & Mean of all reported opinions (Eq. \ref{eq:meanobs}) & $\mathbb{R}$ \\
-    $g$ & Observed group opinion polarization (Eq. \ref{eq:group-polarization})
+    $\meanobst$ & Mean of all reported opinions (Equation \ref{eq:meanobs}) & $\mathbb{R}$ \\
+    $g(\nu)$ & Spurious group opinion polarization for null-polarization params $\nu$
+    (Equation \ref{eq:group-polarization})
         & $\mathbb{R}$ \\
-    $d$ & Cohen's $d$ used as significance threshold 
-        & $\mathbb{R}$ \\
-    $\alpha_e,\alpha_s,\alpha_{\text{all}}$ & Family-wise/Type I error rate
-    for an experiment, study, or across all experiments
+    $d_{ei}$ & Cohen's $d$ for the $i^{\mathrm{th}}$ simulation of experiment $e$ & $\mathbb{R}$ \\
+    $d*$ & Significance value in terms of Cohen's $d$, e.g., $d^* \geq 0.5$ as a
+    "medium" effect\footnote{According to Cohen (1986)\cite{CohenBook1988}.} & $\mathbb{R}$ \\
+    $\alpha_e,\alpha_s,\alpha_{\text{all}}$ & False positive rate
+    for an experiment, $e$; study, $s$; or across \emph{all} experiments.
              & $[0, 1]$ \\
-    FDR & False discovery rate
+    FDR$_e$ & False detection rate: calculated using $\alpha_e$ to get chance experimental
+    detection is false\footnote{See Equation~\ref{eq:fdr}}
         & $[0, 1]$ \\

### 2025-11-14 GroupPolarizationStatmod/paper/method.tex
method rewrite commence

@@
 
-Our method combines several formal and computational tools to identify which
-replications should be retracted and evaluate experimental designs across our
-corpus. The first step in the analysis is to find a reasonable pair of latent
-null-polarization distributions that appear polarized due to clipping
-(Section~\ref{subsec:identify}). To do this, we develop a formal model and a numeric
-rootfinding algorithm to power a Shiny web app we used to interactively identify and
-record null-polarization parameters generating such measurement artifacts.
+A group polarization replication is proven undecidable, and therefore not
+truly a replication, if we can find a model of \emph{mere agreement} that
+appears to be polarized under that replication's experimental design. If we
+cannot find such a model, we cannot determine a replication to be
+undecidable. Mere agreement in this model is a dynamic where agreement
+increases within the group, but extremism does not—in other words, A
+\emph{null-polarization distributions} is a set of two Gaussian
+distributions: one representing the pre-discussion opinions of the group, and
+the other representing the post-discussion opinions. Our
+\emph{null-polarization distributions} represent the process of the group
+finding \emph{mere agreement}, which is when the average group opinion is
+constant over time, but opinion variance decreases to represent agreement
+(or, equivalently, \emph{consensus}~\cite{DeGroot1974}).
 
-The second step in the analysis is to evaluate the quality of the experimental designs
-when we properly account for the ordinal measurement procedure
-(Section~\ref{subsec:evaluate}). We use the null-polarization parameters we found in
-the first step to generate simulated experimental data, then estimate the difference
-between the pre- and post-discussion means using ordered probit models to fit each
-distribution—ordered probit models account for the ordinal measurement procedure.  We
-estimate the false positive rate and false detection rate for low, medium, and high
-significance thresholds in terms of Cohen's $d$. Finally, we use the effect sizes thus
-calculated to determine the significance threshold one must use to achieve a 5\% false
-positive rate.
+Null-polarization  We thus model opinions at each time, $t=\mathrm{pre}$ and
+$t=\mathrm{post}$, as random variables drawn from a normal distribution with
+mean $\mu_t$ and standard deviation $\sigma_t$.  we drop the time
 
-\subsection{Identify replications for removal}
-\label{subsec:identify}
+The null-polarization distributions are represented by
 
-The first step is to find which replications of the 57 must be retracted due to
-non-identifiability. We establish non-identifiability by finding one set of latent
-null-polarization opinion distributions where the mean doesn't change, but appear
-polarized when measured on the ordinal scale used in the experiment.  In other words,
-if we find a null-polarization model that explains the outcome as well as a group
-polarization one, the group polarization hypothesis is \emph{non-identifiable} and the
-replication must be rejected—the data cannot confirm or disconfirm the target
-hypothesis.
+in a simple, general way—i.e., a decrease in opinion variance but a static
+opinion mean—we represent . In the general group polarization hypothesis, the
+group is assumed to start out mostly agreeing in principle (i.e., mostly all
+agree or disagree) but to different extremes. After discussion
 
-To do this we designed a web app that allows the user to try different latent
-distribution parameters (Figure~\ref{fig:data-model-flow}) to seed a
-hillclimbing algorithm that explores latent distribution parameter space
-until it finds null parameters that look like group polarization under the
-experiment's ordinal measurement procedure, or the algorithm stops after a
-certain number of steps (Algorithm~\ref{alg:hillclimbing}). If the web app
-finds a constant latent mean and pre- and post-discussion 
-variances to specify a pre- and post-discussion opinion distributions that
-generate spurious group polarization, this is communicated to the user who
-can then mark the experimental condition as such.
+Showing a replication is undecidable is sufficient to undermine it. However,
+it does not tell us what is the probability that a particular experimental
+design produces false positives, due to clipping or any other problem.
 
-To find examples of plausibly spurious group polarization we guess at what latent
-mean, combined with a latent pre- and post-deliberation latent standard
-deviations, could generate it. We calculate the observed group polarization
-for these latent parameters by transforming the latent pre- and
-post-deliberation distributions from continuous normal distributions to
-ordinal ``observations'' by integrating over each ordinal bin, then
-calculating the observed mean for each experiment \emph{phase}, i.e., both
-the pre- and post-deliberation phases.  If the null model can generate group
-polarization in this case, then the detection is labelled \emph{plausibly
-false}, i.e., \emph{plausibly spurious}, and we call the generating latent
-parameters the \emph{plausible null parameters}.
+first develop a mathematical model of estimate the best-case scenario false
+positive rate for each experimental design,
 
-To evaluate whether the plausible null parameters we identify could
-realistically fool a significance test, we implemented a generative
-simulation of each plausibly false detection of group polarization and
-calculated the \emph{family-wise error rate (FWER)} and the \emph{false
-discovery rate (FDR)}. We use the published number of participants $N$ for
-each of the 54 conditions that yielded plausibly false detections of group
-polarization. The FWER is also known as the Type I error, or the false
-positive rate: it is the probability that a detection of an effect occurs
-given that in reality there was a null effect.  The FDR is the probability
-that a published positive detection of an effect is a false positive, i.e., a
-detection of a null effect. High FWER and FDR for an experimental condition
-indicates that its design is weak, since we identify only one of perhaps
-infinitely many latent parameter combinations that can generate spurious
-group polarization.
+\subsection{Mathematical Model of Null-Polarization Measurement Artifacts}
 
-We use this setup to demonstrate how \emph{adversarial} style testing of
-experimental designs with counterfactual null models can be used to estimate
-important metascience parameters. In this study we use our FWER calculations
-to estimate the minimum Cohen's $d$ that should be used to determine whether
-an effect is ``significant'' or not that will limit the FWER to 5\%. In the
-Discussion we sketch how our model can then be extended to rigorously
-calculate statistical power, i.e., the probability that a significant effect
-is detected given that there is indeed a significant effect.
-
-\input{method-tables.tex}
-
-\subsection{Probability theory of spurious group polarization}
-
-% We formalize this experimental design in the $N\to\infty$ case where the
-% distribution of ordinal measurements is created by integrating continuous latent
-% opinion distributions.
-
-Latent opinions are assumed to be normally distributed in general since they are
-aggregations of many neurobiological potentials of
-different polarities and magnitudes. We have two opinion distributions for 
-two different times in the group polarization experiment process: pre-discussion,
-$\tpre$, and post-discussion, $\tpost$.
-The 
-mean $\mu_t$ and variance $\sigma_t$ for time $t$. We are concerned with only two
-times, \emph{pre-} and \emph{post-} deliberation ($t=0$ and $t=T$, respectively).
-For short we can write $\mu_\mathrm{pre}$ and $\sigma_\mathrm{pre}$ for
-pre-deliberation mean and variance (and similar for \emph{post}-deliberation).  
-The latent opinion of participant $i$ at time $t$ is drawn from a normal 
-distribution in the model, 
+Latent opinions are modeled as normally-distributed random variables.  
+This is a simple, sufficiently general way to represent group-level changes
+in opinions as encoded in the general hypothesis of group polarization, that
+likeminded groups tend to become more extreme on average as they discuss.
+For time, $t$, then, we represent each participant $i$'s opinion as 
 \begin{equation} 
-  \omega_{i,t} \sim p(\omega; \mu_t, \sigma_t) = \mathcal{N}(\mu_t, \sigma_t).
+  \omega_{i,t} \sim \mathcal{N}(\mu_t, \sigma_t).
   \label{eq:opinionDistribution} 
@@
 
-The measurement process with $N\to\infty$ participants is simulated by 
-integrating the latent opinion distribution between ordinal bin thresholds, 
-of which there are $B+1$, where $B$ is the number of opinion bins, with sequential
-integer values $b\{0, B\}$. In the Moscovici and Zavalloni (1969)
-example, the thresholds are $\theta_b \in \{-\infty,-2.5,-1.5,\ldots,2.5,\infty\}$, with $B = 7$.
-Bin $b$ is defined by its lower and upper continuous-valued thresholds, 
-% $\theta_{b-1}$ and
-$\theta_{b}$, respectively. Thresholds are given by the following equations
-(visualized as dotted lines in Figure~\ref{fig:distros}).
+Group-level opinion change, then, can be represented as a change in opinion
+mean, opinion variance, or both. Group polarization occurs 
+when a group's average opinion increases over time. When likeminded people
+interact, they tend to come to agree more over time—they \emph{find consensus}. 
+By definition this means opinion variance decreases, 
+whether the average changes or not.
+
+We can define a null-polarization model, then, as one where the group's
+average opinion stays constant, but the variance decreases: $\sigmapost
+\leq \sigmapre$. Since the mean opinion is constant we just call it 
+$\mu = \mupre = \mupost$. For conciseness, we define a null-polarization
+model as a 3-tuple, $\nu = (\mu, \sigmapre, \sigmapost)$.
+
+\input{method-tables.tex}
+
+\subsubsection{Simulated opinion measurements}
+
+We simulate the collection of an \emph{ordinal distribution} of \emph{ordinal
+observations}, as was performed in a prototypical group polarization experiment 
+by integrating the latent opinion probability density function, $f(\omega)$,
+over each bin's corresponding range of opinion values. For example, consider an
+ordinal scale that runs from -3 (strongly disagree) to +3 (strongly agree)
+(as used by \textcite{Moscovici1969}, for example).
+the frequency with which the ``2'' bin is chosen is just the total
+area underneath the probability density curve between 1.5 and 2.5.
+
+Formally, the probability density function is
 \begin{equation}
-\theta_b = \begin{cases}
-  -\infty         & \text{ if } b=0 \\
-  \infty          & \text{ if } b=B \\
-  \theta_{\text{Base}} + \frac{b}{2} & \text{ otherwise.}
-\end{cases}
+  f(\omega) = 
+    \frac{1}{\sigma \sqrt{2 \pi}} e^{-\frac{(\omega - \mu)^2}{2 \sigma^2}}.
+  \label{eq:normal_pdf}
 \end{equation}
 \noindent
-In the Moscovici and Zavalloni example, the base bin value is
-$\theta_{\text{Base}} = -3$.
-With thresholds defined, the frequency of ordinal measurment $b$ at time $t$ 
-is (Figure~\ref{fig:distros})
-\begin{equation}
-  p(b;~\mu_t, \sigma_t, B) = \int_{\theta_{b-1}}^{\theta_b} p(\omega; \mu_t, \sigma_t)
-  d\omega.
-  \label{eq:ordinal-frequency}
-\end{equation}
+Assume there are $B$ bins, meaning there are $B+1$ \emph{thresholds} that
+define which latent opinions get mapped to which bin values. The bin value,
+$b_i$, is the numerical value of the $i$th bin.
 
-We can now define spurious group polarization and give examples of when it
-occurs. First, the expected value of observed opinions at time $t$ is
-\begin{equation}
-  \meanobst = \frac{1}{B} \sum_{b=1}^B b \cdot p(b; \mu_t, \sigma_t, B).
-  \label{eq:meanobs}
-\end{equation}
-\noindent
-The observed group polarization opinion shift is then
-\begin{equation}
-  g = \meanobs_\mathrm{post} - \meanobs_\mathrm{pre}.
-  \label{eq:group-polarization}
-\end{equation}
-\noindent
+\subsection{Root-finding Algorithm to Find Null-Polarization Models with Artifacts}
 
-Measurement artifacts can make null-polarization spuriously appear like group
-polarization. To see how, we will build up a formal model of the
-process. Consider a collection of
-null-polarization latent opinion distribution parameters $\nu = (\mu, \sigma_{pre}, \sigma_{post})$.
-The group polarization opinion shift can then be written as a function of
-$\nu$,
-$g(\nu)$.
+\subsection{Corpus Analysis to Determine Undecidable Replications}
 
-If we can find ..... $\nu$ for which $g(\nu) = g_{pub} > 0$, where
-$g_{pub}$ is the published group polarization shift value.
-This could be thought of as sampling the observed ordinal distributions 
-$N \to \infty$, collecting ordinal measurements of samples from 
-each latent distribution corresponding to each phase, pre- and post-deliberation.
-
-Next, we explain how to find a set of latent null parameters, $\nu$, that
-generate spurious group polarization when naively measured on an ordinal scale.
-
-\begin{figure}[ht]
-  \caption{\textbf{Opinion measurement model and spurious group polarization.} 
-  Latent model participant opinions are drawn from a normal distribution with the
-  listed parameters (A,C). The pre-deliberation latent distribution has high
-  variance, but appears polarized when measured via integration (A).  The latent
-  distribution's variance is expected to decrease during deliberation (B) via
-  consensus, but in this example the latent mean is static, 
-$\mupre = \mupost$. However, when the post-deliberation distribution is measured,
-the simulated observed mean increases, i.e., spurious
-group polarization occurs (C).}
-  \centering
-  \includegraphics[width=0.95\textwidth]{Figures/Model/latent-ordinal-distros.pdf}
-  \label{fig:distros}
-\end{figure}
-
-\subsubsection{Interactive algorithm to find parameters inducing spurious group polarization}
-
-To identify plausibly false detections of group polarization, we searched numerically
-for latent parameters $\nu = (\mu, \sigmapre, and \sigmapost)$ that generate
-spurious group polarization, as explained immediately above. We developed an *ad hoc*
-hillclimbing-style algorithm that finds $\nu$ that finds $\nu$ by minimizing
-the difference between published group polarization shifts $g_pub$ and calculated shifts,
-$g_\nu$.
-calculated with Equation~\ref{eq:meanobs}.
-
-The algorithm to find $\nu$ holds $\mu$ constant while searching for $\sigmapre$
-and $\sigmapost$ that generate observed $\meanobspre$ and $\meanobspost$, within a
-tolerance of $\pm10^{-3}$ (Algorithm~\ref{alg:hillclimbing}). 
-The algorithm requires a researcher to specify an initial
-guess for $\nu$ and auxiliary parameter settings, which generally vary between
-experimental conditions.  Furthermore, the algorithm does not always produce
-reasonable distributions with a given $\nu$, and sometimes it does not produce
-spurious group polarization.  Therefore, it is necessary to sanity check the latent
-and ordinal opinion distributions for the $\nu$ found by the algorithm. If all is
-well, the experimental condition can be marked as plausibly spurious.  To tame the
-complexity involved in data entry and model fitting for each of these sixty
-experimental conditions, we developed a web application to input data from control
-the algorithm and inspect its results (Figure~\ref{fig:webapp}). Web application code is
-available online in our GitHub
-repository\footnote{\url{https://github.com/mt-digital/gp-statmod/blob/main/GroupPolarizationStatmod/app.R}}.
-
-\begin{algorithm}
-  \caption{Routine to find latent mean and variance that generate spurious group polarization.}
-  \label{alg:hillclimbing}
-  \begin{algorithmic}
-
-    \Require $\mu$, $\sigma$, $\bar{\mathbf{o}}$ 
-      \Comment{Set latent mean, guess variance that yields, input empirical mean.}
-
-    \Require $\epsilon$, $\delta$, $i_\mathrm{max}$ 
-      \Comment{Hillclimbing tolerance, step size, and maximum iterations.}
-
-    \State $\Delta \gets \meanobs(\mu, \sigma) - \meanobsemp$ 
-      \Comment{Initialize simulation error.}
-    \While {$\Delta^2 > \epsilon$ and $i < i_\mathrm{max}$} \Comment{Search until desired
-    accuracy or max iterations reached.}
-      \State $\sigma' \sim \mathcal{N}(\sigma, \delta)$ \Comment{Draw new variance to test.}
-      \State $\Delta' \gets \bar{o}(\mu, \sigma') - \bar{\mathbf{o}}$ 
-        \Comment{Calculate error using new variance.}
-      \If {$\Delta' - \Delta < 0$}
-        \State $\sigma \gets \sigma'$
-          \Comment{Update $\sigma$ if the error is less than before.}
-        \State $\Delta \gets \bar{o}(\mu, \sigma) - \bar{\mathbf{o}}$
-          \Comment{Update simulation error if $\sigma$ is updated.}
-      \EndIf
-    \EndWhile \\
-
-    \Return $\sigma$
-  \end{algorithmic}
-\end{algorithm}
-
-Note that so far, we have only identified cases where spurious group
-polarization plausibly occurs in the asymptotic case where $N\to\infty$ ($N$
-being the number of participants).  Experimental conditions yielding
-plausibly spurious group polarization warrant some suspicion. If there exists
-set of latent simple consensus parameters $\nu$ that create binned
-(Equation~\ref{eq:ordinal-frequency}) opinion distributions that naively seem
-to be group polarized, then the means of at least some samples from these
-distibutions with empirical numbers of participants would also appear to be
-shifted consistent with group polarization. This process itself does not tell
-us the false detection rate.  We use the identified simple consensus
-parameters to define pre- and post-deliberation latent opinion distributions
-in simulated trials of group polarization experiments.
-
-\begin{figure}
-  \centering
-    % \includegraphics[width=0.8\textwidth]{Figures/Model/data-model-flow.png}
-  \includegraphics[width=\textwidth]{Figures/Model/SoftwareMap_Mashup.png}
-  \caption{\textbf{Simulation and analysis pipeline}
-  for finding when measurement artifacts.  
-  Arrows show the flow of input and output data from parameter estimation.
-  First, we used the web app (A) to find null-polarization parameters for 
-  each experimental condition, $\nu_e$,
-  that ordinal measurements distort to look like polarization
-  (B—described in Algorithm~\ref{alg:hillclimbing}).
-  We fit ordered probit models to simulated null-polarization outcomes,
-  seeded by the $\nu$ found in the previous step (C). Analysis and 
-  visualization routines finish the pipeline (D).}
-  \label{fig:data-model-flow}
-\end{figure}
-
-We calculate via simulation and data fitting
-the Type I error rate for the experimental designs used in experimental conditions
-that, under a preliminary large-$N$ model analysis, generated group
-polarization upon ordinal measurement of simple consensus latent opinion 
-distributions. These can then be aggregated to estimate the false discovery rate
-for each journal article and for the ten-article corpus, which is representative of the
-false discovery rate for group polarization studies.
-
-The false detection rate is the scaled probability that a detection, $D$, of
-an effect, $E$, is in fact a false detection and there is \emph{not} an effect,
-denoted $\neg E$. The probability that a detection $D$ has occurred when there
-was no effect, i.e. when $\neg E$, is denoted $\alpha = \Pr(D | \neg E)$.
-This is known as the Type I or family-wise error rate. To calculate the 
-false detection rate ($\fdr$), one must compare $\alpha$ to the statistical power,
-the probability of a detection when there is an effect, denoted $W = \Pr(D | E)$.
-Both these values must be scaled by the rate at which effects occur, denoted $b$.
-Using the notation outlined above, the false detection rate is
-\begin{equation}
-  \fdr = \frac{(1 - b)\alpha}{(1 - b)\alpha + bW}.
-  \label{eq:fdr}
-\end{equation}
-
-In group polarization, $E$ represents the group polarization \emph{effect}, and
-$\neg E$ represents its absence. Simple consensus is one way for group polarization
-to be absent, i.e., simple consensus is one form of $\neg E$. The presence of
-an effect $E$ is independent of whether an effect was \emph{detected}, 
-represented by $D$ if it was detected and $\neg D$ if not.
-
-The false discovery rate is then the Type I error rate relative to the rate of
-positive discoveries, true or false.
-The group polarization studies are journal articles that present
-the results of multiple experiments to support some hypothesis. To organize
-the presentation of false detection rates across the ten studies in our
-analysis, we first will organize false detection rates by experiment, where
-we created `ExperimentID` tags for each of the sixty experiments across ten
-studies. We also created `StudyID` tags for each of the ten journal articles
-with the key `{Author}{Year}`, e.g., \texttt{Schkade2010} for Schkade, et al.,
-(2010) \cite{Schkade2010}.
-
-We calculate $\alpha = \Pr(D | \neg E)$ through simulation for experimental
-conditions, 
-and use optimistic values of
-$b=0.1$~\cite{Ioannidis2005,McElreath2015} and $W=0.8$~\cite{Smaldino2016NatSel}
-in the main text and in the compare with more pessimistic, yet realistic, estimates
-of these values, as well as even more optimistic estimates.
-
-\subsubsection{Estimating and limiting false detection rates}
-
-To calculate the false detection rate we begin by simulating 
-pre- and post-deliberation 
-opinions by drawing $N_e$ samples from normal latent opinion
-distributions that generated spurious group polarization, i.e., with
-parameters labelled $\nu_e$ found in the hillclimbing step. Binning these
-random values simulates observations. $N_e$ is the number
-of participants in experimental condition $e$. Each trial $i$ of $N_e$
-simulated pre- and post-deliberation observations is then inspected to see
-if spurious group polarization occurred in the simulation. Spurious group
-polarization is determined by first inferring the latent parameters of each
-simulated experimental trial by fitting an ordered probit model~\cite{Liddell2018}. 
-The inferred parameters are then used to calculate the effect size for that
-trial, Cohen's $d$, defined as~\cite[p. 331]{Liddell2018}
-\begin{equation}
-  d = \frac{\mupost - \mupre}{\sqrt{\frac{1}{2}(\sigmapost^2 + \sigmapre^2)}}.
-  \label{eq:cohens}
-\end{equation}
-\noindent
-The family-wise error rate is, $\alpha_{e} = \Pr(D|\neg E)$, is the 
-fraction of trials, 
-with $d_{ei} > d^*$ for the $i^\text{th}$ trial for experimental
-condition $e$, written
-\begin{equation}
-  \alpha_e = \frac{1}{N_T} \sum_i^{N_T} \Theta(d_{ei}; d^*),
-  \label{eq:study_aggregate}
-\end{equation}
-\noindent
-where
-\[
-  \Theta(d_{ei}; d^*) = 
-    \begin{cases}
-      1 & \text{ if } d_{ei} > d^* \\
-      0 & \text{ otherwise.}
-    \end{cases}
-\]
-and $N_T = 1000$ is the number of trials. For simplicity we assume that the sign of
-$d_{ei}$ is positive, meaning a shift to greater extremisim. It is possible that
-$d_{ei}$ could be negative, indicating a shift to lesser extremity. Indeed we
-observe many simulations with a shift to lesser extremity. For now we do not
-include these as indicating a ``significant'' effect, which we return to in the
-Analysis below.
-
-To calculate the family-wise error rate
-for a study, $s$ which contains several experimental conditions, $e$,
-\begin{equation}
-  \alpha_s = \frac{1}{|s|} \sum_{e \in s} \alpha_e
-\end{equation}
-\noindent
-The mean across all results identified as plausibly spurious is 
-\begin{equation}
-  \alpha_\mathrm{all} = \frac{1}{|E|} \sum_{e \in E} \alpha_e.
-\end{equation}
-We finally may calculate the false discovery rate by plugging this $\alpha$
-into Equation~\ref{eq:fdr} and assuming empirically-motivated values for 
-base rate, $b$, and statistical power, $W$, as well as more pessimistic and
-optimistic settings for base rate and power.
-
-We close our Analysis by inverting this procedure to find what significance value
-$d^*$ is necessary to achieve a low family-wise error rate $\alpha_e = 0.05$ for
-experimental condition, $e$. To do this we first calculate $\alpha_e(d^*)$ for a
-sufficiently wide range of $d^*$. We then find the lowest $d^*$ for which
-$\alpha_e(d^*) \leq 0.05$.
+\subsection{Evaluating Grouop Polarization Experimental Designs}

### 2025-11-18 GroupPolarizationStatmod/paper/method.tex
drafting method cont'd

@@
 
-\input{method-tables.tex}
+\begin{center}
+\vspace{1.5em}
+[ TABLE \ref{tab:latent-variables} ABOUT HERE ]
+\vspace{1.5em}
+\end{center}
 
@@
 \noindent
-Assume there are $B$ bins, meaning there are $B+1$ \emph{thresholds} that
-define which latent opinions get mapped to which bin values. The bin value,
-$b_i$, is the numerical value of the $i$th bin.
+To set up our measurement simulation, we will represent
+oridnal bin values as $b_i$ with $i=1,\ldots,B$—i.e., there are $B$ bins. 
+To create $B$ bins there must be $B+1$ \emph{thresholds} that define what
+range of latent opinions correspond to which ordinal bin value.
+
+\begin{center}
+  \vspace{1.5em}
+  [ TABLE \ref{tab:experiment-parameters} ABOUT HERE ]
+  \vspace{1.5em}
+\end{center}
+
+All bins have a width of 1 in latent opinion space, except for the two most
+extreme bins at opposite ends of the sentiment spectrum—this works because
+bin values are always separated by 1. All continuous latent opinion values
+$\omega < b_1 + \frac{1}{2} = \theta_1$ get mapped to the ordinal measurement
+value $b_1$—so, $b_1$'s lower threshold is $\theta_0 = -\infty$.  Similarly,
+all opinions $\omega > b_B - 0.5 = \theta_B$ are reported as the maximum bin
+value, $b_B$, with an effective upper bound of $\theta_{B+1} = +\infty$.  We
+can define all thresholds compactly as \cite{Liddell2018}
+\begin{equation}
+\theta_j = \begin{cases}
+  -\infty         & \text{ if } j=0 \\
+  \infty          & \text{ if } j=B \\
+  b_1 + \frac{j + 1}{2} & \text{ otherwise.}
+\end{cases}
+\label{eq:thresholds}
+\end{equation}
+\noindent
+
+We model the measurement of ordinal opinions by calculating the probability
+density function over bins, written
+\begin{equation}
+  f[b_i] = \int_{\theta_{i - 1}}^{\theta_i} f(\omega) d\omega~,
+  \label{eq:measurement-math}
+\end{equation}
+\noindent
+switching to square brackets for the argument to $f$ now to emphasize that 
+bins are discrete. With this ordinal measurement model we can caluclate the 
+mean opinion of the simulated measurements like so
+\begin{equation}
+  \bar{o}_t = \frac{1}{B} \sum_{b=b_1}^{b_B} b \cdot f_t[b].
+\end{equation}
+\noindent
+
+\begin{center}
+\vspace{1.5em}
+[ FIGURE \ref{fig:distros} ABOUT HERE ]
+\vspace{1.5em}
+\end{center}
+
+Group polarization is the difference between the post-deliberation mean
+observed opinion and the pre-deliberation one, 
+which we can write as 
+\begin{equation}
+  g = \bar{o}_\mathrm{post} - \bar{o}_\mathrm{pre}.
+  \label{eq:group-polarization}
+\end{equation}
+\noindent
+
+\begin{center}
+\vspace{1.5em}
+[ TABLE \ref{tab:simulated-observations} ABOUT HERE ]
+\vspace{1.5em}
+\end{center}
 
@@
 
+To analyze published replications we need to expand our notation to index
+individual \emph{experiments} with replications, and from which 
+different journal articles, track and analyze it,
+
 \subsection{Evaluating Grouop Polarization Experimental Designs}

### 2025-11-18 GroupPolarizationStatmod/paper/figures.tex
added fig and tables module

\section*{Figures}

\begin{figure}[ht]
  \caption{\textbf{Opinion measurement model and spurious group polarization.} 
  Latent model participant opinions are drawn from a normal distribution with the
  listed parameters (A,C). The pre-deliberation latent distribution has high
  variance, but appears polarized when measured via integration (A).  The latent
  distribution's variance is expected to decrease during deliberation (B) via
  consensus, but in this example the latent mean is static, 
$\mupre = \mupost$. However, when the post-deliberation distribution is measured,
the simulated observed mean increases, i.e., spurious
group polarization occurs (C).}
  \centering
  \includegraphics[width=0.95\textwidth]{Figures/Model/latent-ordinal-distros.pdf}
  \label{fig:distros}
\end{figure}

### 2025-11-18 GroupPolarizationStatmod/paper/tables.tex
added fig and tables module

--- 2025-11-06 GroupPolarizationStatmod/paper/article-table.tex
@@
+\section*{Tables}
+
 \begin{table}[ht]
@@
 \end{table}
+
+%%%%%%%%%%%%%%%%%%%
+% TABLES OF VARIABLES IN METHOD
+\begin{table}[h]
+  \caption{\textbf{Latent psychological variables and distributions.}}
+  \label{tab:latent-variables}
+  \begin{tabular}{cll} 
+  \toprule
+    \textbf{Variable} & \textbf{Description} & \textbf{Values} \\
+  \midrule  
+    $\omega_{i,t}$ & Latent opinion of individual $i$ at time $t$ & $\mathbb{R}$ \\
+    $\mu_t$ & Mean latent opinion at time $t$ (constant in null-polarization) & $\mathbb{R}$ \\
+    $\sigma_t$ & Variance of observed opnions at time $t$ & $\mathbb{R}$ \\
+  \bottomrule
+  \end{tabular} 
+\end{table}
+
+\begin{table}[h]
+  \caption{\textbf{Experimental design and numerical model parameters.}}
+  \label{tab:experiment-parameters}
+  \begin{tabular}{cll} 
+  \toprule
+    \textbf{Variable} & \textbf{Description} & \textbf{Values} \\
+  \midrule  
+    $B$ & Number of bins & $5,6,7,\ldots$ \\
+    $\theta_b$ & The $B+1$ threshold values that separate bins & 
+      e.g., $-\infty, -2.5, -1.5,\ldots,2.5, \infty$ \\
+    $b$ & Bin index—there are $B+1$ to mark $B$ bins & $0,1,2,\ldots,B$ \\
+  \bottomrule
+  \end{tabular} 
+\end{table}
+
+\begin{table}[h]
+  \caption{\textbf{Simulated observed variables.}}
+  \label{tab:simulated-observations}
+  \begin{tabular}{cll} 
+  \toprule
+    \textbf{Variable} & \textbf{Description} & \textbf{Values} \\
+  \midrule  
+    $\meanobst$ & Mean of all reported opinions (Equation \ref{eq:meanobs}) & $\mathbb{R}$ \\
+    $g(\nu)$ & Spurious group opinion polarization for null-polarization params $\nu$
+    (Equation \ref{eq:group-polarization})
+        & $\mathbb{R}$ \\
+    $d_{ei}$ & Cohen's $d$ for the $i^{\mathrm{th}}$ simulation of experiment $e$ & $\mathbb{R}$ \\
+    $d*$ & Significance value in terms of Cohen's $d$, e.g., $d^* \geq 0.5$ as a
+    "medium" effect\footnote{According to Cohen (1986)\cite{CohenBook1988}.} & $\mathbb{R}$ \\
+    $\alpha_e,\alpha_s,\alpha_{\text{all}}$ & False positive rate
+    for an experiment, $e$; study, $s$; or across \emph{all} experiments.
+             & $[0, 1]$ \\
+    FDR$_e$ & False detection rate: calculated using $\alpha_e$ to get chance experimental
+    detection is false\footnote{See Equation~\ref{eq:fdr}}
+        & $[0, 1]$ \\
+  \bottomrule
+  \end{tabular} 
+\end{table}

### 2025-11-19 GroupPolarizationStatmod/paper/code_examples.qmd
Created paper/code_examples.qmd and continued on method

---
title: Code examples to accompany "If the Null Fits, You Must Omit"
author: Matt Turner · [matt@mat.phd](mailto:matt@mat.phd)
editor: 
  markdown: 
    wrap: 90
bibliography: this.bib
---

# Purpose: demonstrate gp-stats codebase

Here are some quick notes on the group polarization codebase in R supporting our article,
*If the Null Fits, You Must Omit*. The code is permanently [hosted on
GitHub](https://github.com/mt-digital/gp-statmod/tree/final-draft/GroupPolarizationStatmod).

To follow along, please clone the `gp-statmod` directory.

```         
$ git clone https://github.com/mt-digital/gp-statmod.git
```

Change directories to go to the `paper` directory where this document lives:

```         
cd gp-statmod/GroupPolarizationStatmod/paper
```

Then you can run the examples in this document, edit them, etc., on your own computer.

## Calculate probability density vector for `Moscovici1969_Americans` condition

See Method section of the paper where we calculate the probabilty density vector for
different measurements based on the latent opinion probability density function.

Here we show how to calculate the mean observed opinion at pre- and post-discussion to
show how spurious group polarization occurs under a null polarization hypothesis.

@Moscovici1969 report observing a mean opinion of -0.6 in pre-discussion opinions and mean
opinion -1.04 post-discussion, measured using an ordinal scale from -3 (strongly disagree)
to +3 (strongly agree). Let's calculate the probability density vector for the pre- and
post-discussion distributions then calculate the mean opinion for each one. The latent
distributions have the same mean, -1.2, but different variances, which results in a
shifted observed mean due to opinion clipping.

![](images/clipboard-1443885641.png){width="628"}

We call this a *null polarization model* (defined in terms of latent opinions) that
generates *spurious group polarization.* *Spurious group polarization* is when the
observed opinion distributions appear to polarize (i.e., *radicalize*, or *become more
extreme*) that does not represent the true change in latent opinions. The measurement
procedure *clips* extreme opinions less than the minimum bin value, or greater than the
maximum, causing the average observed opinion to *spuriously* appear polarized.

Let's use some modeling code from the repository to demonstrate. For this part we need to
source (*use code from*) the `model.R` file that has the the
[`binProb`](https://github.com/mt-digital/gp-statmod/blob/final-draft/GroupPolarizationStatmod/model.R#L7)
function I use below. Also make sure you have set the working directory to the
`GroupPolarizationStatmod` directory within the `gp-stats` project/repository directory.

For a preliminary step, adapt my code below for your computer. Change the argument to
`root.dir` to be the location of the `GroupPolarizationStatmod` directory on your own
computer.

```{r setup}

```

Now in the following chunk, the `source` function will know where to look for the file
with the `makeProbVec` function we need, `model.R`.

```{r warning=FALSE}
# This is how I set the working directory
# Adapt for your own computer
setwd("~/workspace/gp-statmod/GroupPolarizationStatmod/")
# Load makeProbVec from file model.R
source("model.R")
```

Now read in the pre- and post-discussion means from `Moscovici1969_Americans`

```{r}
# Experimental means
mosc1969_ave_pre <- -0.6
mosc1969_ave_post <- -1.04
```

Create a hypothesized null model that will spuriously look like polarization, with the
same pre- and post-discussion means above. These correspond to $\mu$,
$\sigma_\mathrm{pre}$, and $\sigma_\mathrm{post}$ from the paper that define a null
polarization model, $\nu = (\mu, \sigma_\mathrm{pre}, \sigma_\mathrm{post})$.

```{r}
# Hypothesized null model that looks polarized
mu <- -1.2
sd_pre <- 4.3
sd_post <- 1.9
```

These

```{r}
# Vector w/ probability density for each bin
f_pre <- makeProbVec(-3:3, -1.2, 4.3)
# Print to two significant digits
sprintf("%.2f", f_pre)
```

We can plot this simply using just

```{r}
barplot(f_pre, names.arg = -3:3)
```

### 2025-11-19 GroupPolarizationStatmod/paper/figures.tex
Created paper/code_examples.qmd and continued on method

@@
 \end{figure}
+
+\begin{figure}
+  \centering
+    % \includegraphics[width=0.8\textwidth]{Figures/Model/data-model-flow.png}
+  \includegraphics[width=\textwidth]{Figures/Model/SoftwareMap_Mashup.png}
+  \caption{\textbf{Simulation and analysis pipeline}
+  for finding when measurement artifacts.  
+  Arrows show the flow of input and output data from parameter estimation.
+  First, we used the web app (A) to find null-polarization parameters for 
+  each experimental condition, $\nu_e$,
+  that ordinal measurements distort to look like polarization
+  (B—described in Algorithm~\ref{alg:hillclimbing}).
+  We fit ordered probit models to simulated null-polarization outcomes,
+  seeded by the $\nu$ found in the previous step (C). Analysis and 
+  visualization routines finish the pipeline (D).}
+  \label{fig:data-model-flow}
+\end{figure}

### 2025-11-19 GroupPolarizationStatmod/paper/method.tex
Created paper/code_examples.qmd and continued on method

@@
 
-The null-polarization distributions are represented by
+The null-polarization distributions are represented by $\nu$.
 
@@
 
-\subsubsection{Simulated opinion measurements}
+\subsubsection{Simulated group polarization measurements}
 
@@
   \infty          & \text{ if } j=B \\
-  b_1 + \frac{j + 1}{2} & \text{ otherwise.}
+  b_j + \frac{1}{2} & \text{ otherwise.}
 \end{cases}
@@
 
-\begin{center}
-\vspace{1.5em}
-[ FIGURE \ref{fig:distros} ABOUT HERE ]
-\vspace{1.5em}
-\end{center}
-
 Group polarization is the difference between the post-deliberation mean
@@
 
+\subsubsection{Simulated spurious group polarization}
+
+Group polarization measures radicalization in experimental echo chambers in terms of
+the difference between mean group opinions before and after the group discusses some
+survey topic. The following model shows the problem of how the measurement scheme 
+above can make \emph{mere agreement} spuriously look like polarization.
+
+To make things more exact, we define a \emph{null model}, $\nu$, of mere agreement
+defined as a 3-tuple: a constant mean for the pre- and post-discussion latent opinion
+distributions, and a pre-deliberation latent opinion variance that is greater than the
+post-deliberation variance to represent finding agreement,
+\begin{equation}
+  \nu = (\mu,~\sigmapre,~\sigmapost).
+  \label{eq:nu-definition}
+\end{equation}
+\noindent
+Group polarization is \emph{spurious} in this case if measurement clipping causes 
+the observed difference in pre- and post-discussion opinions to be greater than zero,
+i.e., $g > 0$ (Equation \ref{eq:group-polarization}, 
+even if the true latent change is zero.
+
+We will show how this arrangement generates spurious polarization in an
+experimental condition in our corpus, 
+the \emph{Americans} experimental condition in
+\textcite{Moscovici1969}, to which we gave the Experiment ID
+\texttt{Moscovici1969\_Americans}. In that condition, the authors used a seven-bin
+scale ($B=7$) with numerical values from -3 to +3 indicating opinions from ``strongly
+disagree'' to ``strongly agree''. Pre-consensus opinions were reported to have
+mean -0.6 and the post-consensus mean was -1.04 \cite[p. 132, Table 4]{Moscovici1969}.
+
+And now we will show that a null model with
+$\nu = (\mu = -1.2, \sigmapre = 4.3, and \sigmapost = 1.9)$ generates spurious
+polarization with exactly the pre- and post-discussion means Moscovici and Zavalloni
+report (Figure~\ref{fig:distros}). First, calculate the vector of probability density
+values for each bin, $f[b]$, for the pre- and post-discussion latent opinion
+distributions using Equation \ref{eq:} (calculated with
+[`makeProbVec`](https://github.com/mt-digital/gp-statmod/blob/final-draft/GroupPolarizationStatmod/model.R\#L43) from 
+the [project code repository](https://github.com/mt-digital/gp-statmod/tree/final-draft/GroupPolarizationStatmod):
+\begin{subequations}
+  \begin{equation}
+    f_\mathrm{pre}[b] = 
+  \end{equation}
+  \begin{equation}
+    f_\mathrm{post}[b] = 
+  \end{equation}
+\end{subequations}
+
+\begin{center}
+\vspace{1.5em}
+[ FIGURE \ref{fig:distros} ABOUT HERE ]
+\vspace{1.5em}
+\end{center}
 \subsection{Root-finding Algorithm to Find Null-Polarization Models with Artifacts}

### 2025-11-20 GroupPolarizationStatmod/paper/code_examples.qmd
worked out code example for calculating observed means from latent opinion distributions

@@
 ---
-title: Code examples to accompany "If the Null Fits, You Must Omit"
-author: Matt Turner · [matt@mat.phd](mailto:matt@mat.phd)
+title: Notes and code examples to accompany "If the Null Fits, You Must Omit"
+author: Matt Turner · [https://mat.phd](https://mat.phd)
+date: today
+toc: true
 editor: 
@@
 
-# Purpose: demonstrate gp-stats codebase
+### Purpose: demonstrate gp-stats codebase
 
@@
 
-## Calculate probability density vector for `Moscovici1969_Americans` condition
+# Simulations
 
-See Method section of the paper where we calculate the probabilty density vector for
-different measurements based on the latent opinion probability density function.
+I developed two models of group polarization to serve complementary purposes.
+
+## Expectation model of group polarization experiments
+
+See Method section of the paper where we explain how to calculate the probability density
+vector representing the probability that a participant gives the opinion represented by
+each measurement bin, $b$.
 
@@
 computer.
-
-```{r setup}
-
-```
 
@@
 
-These
+The function `makeProbVec` creates the vector $f[b]$ of probabilities that bin $b_i \in b$
+will be reported by a randomly-selected participant, given the measurement bin values and
+the distribution of latent opinions, assumed normal, defined by the mean and standard
+deviation, $\mu$ and $\sigma$. and bin values. Here bin values are $b=-3,-2,\ldots,3$;
+$\mu=-1.2$; and $\sigma=4.3$.
 
@@
 
-We can plot this simply using just
+Let's `barplot` to have a look:
+
+```{r out.width=500}
+barplot(f_pre, 
+        names.arg = -3:3,
+        ylim = c(0, 0.4),
+        ylab = "Pct. Responding",
+        xlab = "Bin Value")
+```
+
+Let's check the simulated average observed opinion, $\bar{o}$, using the expected value of
+the distribution, i.e., we assume $\bar{o} = \langle o \rangle$, so we use the standard
+formula for expected value, which assumes that $f[b]$ is normalized.
+
+$$
+\bar{o} = \sum_{b=b_1}^{b_B} b \cdot f[b].
+$$
+
+In code,
 
 ```{r}
-barplot(f_pre, names.arg = -3:3)
+# Calculate simulated and compare with actual
+sim_obs_pre <- sum((-3:3) * f_pre)
+print(abs(sim_obs_pre - mosc1969_ave_pre))
 ```
+
+This is within one significant digit greater than the reporting scale, meaning it is an
+acceptable fit.
+
+This shows that even just a little spread in the distribution of latent opinions results
+in a seriously distorted mean [@Liddell2018]—one that reproduces the mean reported by
+@Moscovici1969, but actually confirming the null polarization hypothesis!
+
+Let's check the simulated average observed opinion for post-deliberation, repeating the
+previous steps for $t=\mathrm{post}$.
+
+```{r}
+f_post <- makeProbVec(-3:3, mu, sd_post)
+sprintf("%.2f", f_post)
+# Calculate simulated and compare with actual
+sim_obs_post <- sum((-3:3) * f_post)
+print(abs(sim_obs_post - mosc1969_ave_post))
+```
+
+Here we get even closer agreement between the observed post-discussion average opinion in
+@Moscovici1969 and the one we simulated.

### 2025-11-20 GroupPolarizationStatmod/paper/method.qmd
worked out code example for calculating observed means from latent opinion distributions

---
editor: 
  markdown: 
    wrap: 100
---

# Method

A group polarization replication is proven undecidable, and therefore not truly a replication, if we
can find a model of *mere agreement* that appears to be polarized under that replication's
experimental design. If we cannot find such a model, we cannot determine a replication to be
undecidable. Mere agreement in this model is a dynamic where agreement increases within the group,
but extremism does not---in other words, A *null-polarization distributions* is a set of two
Gaussian distributions: one representing the pre-discussion opinions of the group, and the other
representing the post-discussion opinions. Our *null-polarization distributions* represent the
process of the group finding *mere agreement*, which is when the average group opinion is constant
over time, but opinion variance decreases to represent agreement (or, equivalently,
*consensus* [@DeGroot1974]).

Null-polarization We thus model opinions at each time, $t=\mathrm{pre}$ and $t=\mathrm{post}$, as
random variables drawn from a normal distribution with mean $\mu_t$ and standard deviation
$\sigma_t$. we drop the time

The null-polarization distributions are represented by $\nu$.

in a simple, general way---i.e., a decrease in opinion variance but a static opinion mean---we
represent . In the general group polarization hypothesis, the group is assumed to start out mostly
agreeing in principle (i.e., mostly all agree or disagree) but to different extremes. After
discussion

Showing a replication is undecidable is sufficient to undermine it. However, it does not tell us
what is the probability that a particular experimental design produces false positives, due to
clipping or any other problem.

## Simulation of Mere Consensus and Spurious Polarization

Our analysis models \emph{mere consensus} to examine whether published detections of group polarization may plausibly be \emph{spurious}, caused instead by artificats introduced by opinion measurement design.

Latent opinions are modeled as normally-distributed random variables. This is a simple, sufficiently
general way to represent group-level changes in opinions as encoded in the general hypothesis of
group polarization, that likeminded groups tend to become more extreme on average as they discuss.
For time, $t$, then, we represent each participant $i$'s opinion as 
$$
\omega_{i,t} \sim \mathcal{N}(\mu_t, \sigma_t).
$$

Group-level opinion change, then, can be represented as a change in opinion mean, opinion variance,
or both. Group polarization occurs when a group's average opinion increases over time. When
likeminded people interact, they tend to come to agree more over time---they *find consensus*. By
definition this means opinion variance decreases, whether the average changes or not.

We can define a null-polarization model, then, as one where the group's average opinion stays
constant, but the variance decreases: $\sigmapost
\leq \sigmapre$. Since the mean opinion is constant we just call it $\mu = \mupre = \mupost$. For
conciseness, we define a null-polarization model as a 3-tuple, $\nu = (\mu, \sigmapre, \sigmapost)$.

::: {style="text-align: center;"}
\[ TABLE \[tab:latent-variables\](#tab:latent-variables){reference-type="ref"
reference="tab:latent-variables"} ABOUT HERE \]
:::

### Simulated group polarization measurements

We simulate the collection of an *ordinal distribution* of *ordinal observations*, as was performed
in a prototypical group polarization experiment by integrating the latent opinion probability
density function, $f(\omega)$, over each bin's corresponding range of opinion values. For example,
consider an ordinal scale that runs from -3 (strongly disagree) to +3 (strongly agree) (as used by
@Moscovici1969, for example). the frequency with which the "2" bin is chosen is just the total area
underneath the probability density curve between 1.5 and 2.5.

Formally, the probability density function is $$
f(\omega) = 
    \frac{1}{\sigma \sqrt{2 \pi}} e^{-\frac{(\omega - \mu)^2}{2 \sigma^2}}.
$$ To set up our measurement simulation, we will represent oridnal bin values as $b_i$ with
$i=1,\ldots,B$---i.e., there are $B$ bins. To create $B$ bins there must be $B+1$ *thresholds* that
define what range of latent opinions correspond to which ordinal bin value.

::: {text-align="center"}
\[ TABLE [\[tab:experiment-parameters\]](#tab:experiment-parameters){reference-type="ref"
reference="tab:experiment-parameters"} ABOUT HERE \]
:::

All bins have a width of 1 in latent opinion space, except for the two most extreme bins at opposite
ends of the sentiment spectrum---this works because bin values are always separated by 1. All
continuous latent opinion values $\omega < b_1 + \frac{1}{2} = \theta_1$ get mapped to the ordinal
measurement value $b_1$---so, $b_1$'s lower threshold is $\theta_0 = -\infty$. Similarly, all
opinions $\omega > b_B - 0.5 = \theta_B$ are reported as the maximum bin value, $b_B$, with an
effective upper bound of $\theta_{B+1} = +\infty$. We can define all thresholds compactly as
[@Liddell2018] $$
\theta_j = \begin{cases}
  -\infty         & \text{ if } j=0 \\
  \infty          & \text{ if } j=B \\
  b_j + \frac{1}{2} & \text{ otherwise.}
\end{cases}
$$

We model the measurement of ordinal opinions by calculating the probability density function over
bins, written $$
f[b_i] = \int_{\theta_{i - 1}}^{\theta_i} f(\omega) d\omega,
$$ switching to square brackets for the argument to $f$ now to emphasize that bins are discrete.
With this ordinal measurement model we can caluclate the mean opinion of the simulated measurements
like so $$\bar{o}_t = \frac{1}{B} \sum_{b=b_1}^{b_B} b \cdot f_t[b].$$

Group polarization is the difference between the post-deliberation mean observed opinion and the
pre-deliberation one, which we can write as $$g = \bar{o}_\mathrm{post} - \bar{o}_\mathrm{pre}.
  \label{eq:group-polarization}$$

::: center
\[ TABLE [\[tab:simulated-observations\]](#tab:simulated-observations){reference-type="ref"
reference="tab:simulated-observations"} ABOUT HERE \]
:::

### Simulated spurious group polarization

Group polarization measures radicalization in experimental echo chambers in terms of the difference
between mean group opinions before and after the group discusses some survey topic. The following
model shows the problem of how the measurement scheme above can make *mere agreement* spuriously
look like polarization.

To make things more exact, we define a *null model*, $\nu$, of mere agreement defined as a 3-tuple:
a constant mean for the pre- and post-discussion latent opinion distributions, and a
pre-deliberation latent opinion variance that is greater than the post-deliberation variance to
represent finding agreement, $$
\nu = (\mu,~\sigmapre,~\sigmapost).
$$ Group polarization is *spurious* in this case if measurement clipping causes the observed
difference in pre- and post-discussion opinions to be greater than zero, i.e., $g > 0$ (Equation
[\[eq:group-polarization\]](#eq:group-polarization){reference-type="ref"
reference="eq:group-polarization"}, even if the true latent change is zero.

We will show how this arrangement generates spurious polarization in an experimental condition in
our corpus, the *Americans* experimental condition in @Moscovici1969, to which we gave the
Experiment ID `Moscovici1969_Americans`. In that condition, the authors used a seven-bin scale
($B=7$) with numerical values from -3 to +3 indicating opinions from "strongly disagree" to
"strongly agree". Pre-consensus opinions were reported to have mean -0.6 and the post-consensus mean
was -1.04 [@Moscovici1969 p. 132, Table 4].

And now we will show that a null model with
$\nu = (\mu = -1.2, \sigmapre = 4.3, and \sigmapost = 1.9)$ generates spurious polarization with
exactly the pre- and post-discussion means Moscovici and Zavalloni report
(Figure [\[fig:distros\]](#fig:distros){reference-type="ref" reference="fig:distros"}). First,
calculate the vector of probability density values for each bin, $f[b]$, for the pre- and
post-discussion latent opinion distributions using Equation [\[eq:\]](#eq:){reference-type="ref"
reference="eq:"} (calculated with
\['makeProbVec'\](https://github.com/mt-digital/gp-statmod/blob/final-draft/GroupPolarizationStatmod/model.R#L43)
from the \[project code
repository\](https://github.com/mt-digital/gp-statmod/tree/final-draft/GroupPolarizationStatmod)

$$
    f_\mathrm{pre}[b] = 
$$ $$
    f_\mathrm{post}[b] = 
$$

::: {style="text-align: center;"}
FIGURE [\[fig:distros\]](#fig:distros){reference-type="ref" reference="fig:distros"} ABOUT HERE
:::

## Root-finding Algorithm to Find Null-Polarization Models with Artifacts

## Corpus Analysis to Determine Undecidable Replications

To analyze published replications we need to expand our notation to index individual *experiments*
with replications, and from which different journal articles, track and analyze it,

## Evaluating Grouop Polarization Experimental Designs

### 2025-11-20 GroupPolarizationStatmod/analysis.R

@@ -179,14 +179,10 @@ latexifyTTestExperiment <- function(tTestTablePath = "data/output/TtestSummaryTa
                                         Est.FPRate = ExpectedPower)
     
     ttestTable_latex <- 
-        kable(tTestTable, "latex", align = "lccccc", longtable = T, booktabs = T, digits=2) %>%
+        kable(tTestTable, "latex", align = "lccccc", 
+              longtable = T, booktabs = T, digits=2) %>%
         kable_styling(latex_options = c("repeat_header"))
 
     writeLines(ttestTable_latex, latexifiedPath)
 }
 
-
-doAnalyses <- function(studiesAnalysisDfPath = "data/StudiesAnalysis.csv")
-{
-  
-}

### 2025-11-20 GroupPolarizationStatmod/numerical.R

@@ -14,7 +14,7 @@ solveForLatentSD <- function(kVec, latentMean, observedMean, guess,
                              step = 0.01, maxIts = 10000, tol = 1e-6)
 {
     # Keep the error function to be minimized through hillclimbing
-    # separate from hillclimbing itself. ### XXXX BUT WHYYYY seems confusing...
+    # separate from hillclimbing itself.
     errfunc <- function(latentSD)
     {
         simulatedObservedMean <- meanObs(
@@ -22,7 +22,6 @@ solveForLatentSD <- function(kVec, latentMean, observedMean, guess,
             makeProbVec(kVec, latentMean, latentSD)
         )
         
-        # return ((observedMean - simulatedObservedMean)^2);
         return (simulatedObservedMean - observedMean);
     }
 

### 2025-11-20 GroupPolarizationStatmod/scripts/plot/model.R

@@ -15,8 +15,8 @@ plot_latent_pdf_integration = function(mu = 0, sd = 2.5, min_bin, max_bin,
                                        bins, bin_colors, # xlim = c(-6.0, 6.0), 
                                        save_path = "test_latent_pdf.pdf", 
                                        max_prob = 1.0) {
+  
   print(save_path)
-
 
   xmin = min_bin - 0.5
   xmax = max_bin + 0.5

### 2025-11-24 GroupPolarizationStatmod/paper/method.qmd
finished expectation model subsection

@@
   markdown: 
-    wrap: 100
+    wrap: 80
+bibliography: this.bib
 ---
+
+\newcommand{\mupre}{\mu_\mathrm{pre}}
+\newcommand{\mupost}{\mu_\mathrm{post}}
+\newcommand{\sigmapre}{\sigma_\mathrm{pre}}
+\newcommand{\sigmapost}{\sigma_\mathrm{post}}
+
+\newcommand{\thetafit}{\theta^\mathrm{fit}}
+
+\newcommand{\fdr}{\mathrm{FDR}}
+
+\newcommand{\meanobs}{\bar{o}}
+\newcommand{\meanobsemp}{\bar{\mathbf{o}}}
+\newcommand{\meanobst}{\meanobs_t}
+\newcommand{\meanobsjt}{\meanobs_{j,t}}
+\newcommand{\meanobspre}{\meanobs_\mathrm{pre}}
+\newcommand{\meanobspost}{\meanobs_\mathrm{post}}
+\newcommand{\tpre}{t_\mathrm{pre}}
+\newcommand{\tpost}{t_\mathrm{post}}
 
@@
 
-A group polarization replication is proven undecidable, and therefore not truly a replication, if we
-can find a model of *mere agreement* that appears to be polarized under that replication's
-experimental design. If we cannot find such a model, we cannot determine a replication to be
-undecidable. Mere agreement in this model is a dynamic where agreement increases within the group,
-but extremism does not---in other words, A *null-polarization distributions* is a set of two
-Gaussian distributions: one representing the pre-discussion opinions of the group, and the other
-representing the post-discussion opinions. Our *null-polarization distributions* represent the
-process of the group finding *mere agreement*, which is when the average group opinion is constant
-over time, but opinion variance decreases to represent agreement (or, equivalently,
+A group polarization replication is proven undecidable, and therefore not truly
+a replication, if we can find a model of *mere agreement* that appears to be
+polarized under that replication's experimental design. If we cannot find such a
+model, we cannot determine a replication to be undecidable. Mere agreement in
+this model is a dynamic where agreement increases within the group, but
+extremism does not—in other words, A *null-polarization distributions* is a set
+of two Gaussian distributions: one representing the pre-discussion opinions of
+the group, and the other representing the post-discussion opinions. Our
+*null-polarization distributions* represent the process of the group finding
+*mere agreement*, which is when the average group opinion is constant over time,
+but opinion variance decreases to represent agreement (or, equivalently,
 *consensus* [@DeGroot1974]).
 
-Null-polarization We thus model opinions at each time, $t=\mathrm{pre}$ and $t=\mathrm{post}$, as
-random variables drawn from a normal distribution with mean $\mu_t$ and standard deviation
-$\sigma_t$. we drop the time
+Null-polarization We thus model opinions at each time, $t=\mathrm{pre}$ and
+$t=\mathrm{post}$, as random variables drawn from a normal distribution with
+mean $\mu_t$ and standard deviation $\sigma_t$. we drop the time
 
@@
 
-in a simple, general way---i.e., a decrease in opinion variance but a static opinion mean---we
-represent . In the general group polarization hypothesis, the group is assumed to start out mostly
-agreeing in principle (i.e., mostly all agree or disagree) but to different extremes. After
-discussion
+in a simple, general way---i.e., a decrease in opinion variance but a static
+opinion mean---we represent . In the general group polarization hypothesis, the
+group is assumed to start out mostly agreeing in principle (i.e., mostly all
+agree or disagree) but to different extremes. After discussion
 
-Showing a replication is undecidable is sufficient to undermine it. However, it does not tell us
-what is the probability that a particular experimental design produces false positives, due to
-clipping or any other problem.
+Showing a replication is undecidable is sufficient to undermine it. However, it
+does not tell us what is the probability that a particular experimental design
+produces false positives, due to clipping or any other problem.
 
-## Simulation of Mere Consensus and Spurious Polarization
+## Simulation of Mere Consensus and Opinion Change Measurement
 
-Our analysis models \emph{mere consensus} to examine whether published detections of group polarization may plausibly be \emph{spurious}, caused instead by artificats introduced by opinion measurement design.
+We model \emph{mere consensus} and opinion change measurement in group
+polarization experiments to examine whether published replications can be
+explained by this particular null polarization model. We simulate consensus and
+measurement two ways for complementary purposes. First is this *expectation*
+model, and the second is the *stochastic* model. Both models assume group
+opinions are normally distributed, which is practical due to its simplicity.
 
-Latent opinions are modeled as normally-distributed random variables. This is a simple, sufficiently
-general way to represent group-level changes in opinions as encoded in the general hypothesis of
-group polarization, that likeminded groups tend to become more extreme on average as they discuss.
-For time, $t$, then, we represent each participant $i$'s opinion as 
+We assume that participants' true latent opinions at some time, $t$, are random
+variables drawn from a normal distribution with mean $\mu_t$ and standard
+deviation $\sigma_t$. A normal distribution is scientifically justifiable, since
+opinions are aggregations of many constituent electrical and chemical signals in
+the brain—the distribution of aggregations like this tend to be *normal* (hence
+the name). A normal distribution also makes sense because several of the
+analysis procedures from the corpus use a $t$-test or other basic linear
+regression-style test, which implicitly assume data are distributed normally.
+
+The two models differ in how they simulate opinions, opinion change, and opinion
+change measurement. We use the *expectation model* as a computationally
+efficient way to simulate the observed mean opinion as the mathematical
+*expectation value* of the distribution of measured opinions. In the second
+model, the *stochastic model*, we simulate participant survey reporting as
+drawing a latent opinion from the latent opinion distribution, then binning that
+opinion in the associated opinion bin. We use the expectation model for its
+efficiency in the first step of our analysis: to find a *null-polarization
+model* that also fits the data for each experimental condition studied—if we
+find one that means we must retract this replication since this proves that the
+group polarization is undecidable with the data we have. Once we find any such
+undecidable results, we use the null-polarization model to help us further
+evaluate the experimental design itself.
+
+In the *stochastic model*, we simulate group polarization data for the
+*Evaluation* step of the analysis. We simulate experimental data, with the same
+number of participants as the original experiment, using the null-polarization
+model found in the *Identification* step, if one was found.
+
+We *evaluate* the experimental design, $e$, in terms of the false positive rate,
+also known as the Type I error rate or family-wise error rate, denoted
+$\alpha_e$. We close our analysis by demonstrating how our method can be
+inverted to *evaluate* experimental designs in the pre-registration stage to
+avoid preventable replication failures like we identify here.
+
+### Latent Opinion Model
+
+Throughout this paper we model a participant $i$'s latent opinion at time $t$ as
+a normally-distributed random variable,
+
 $$ 
-\omega_{i,t} \sim \mathcal{N}(\mu_t, \sigma_t).
+\omega_{i,t} \sim
+\mathcal{N}(\mu_t, \sigma_t).  
 $$
+Group polarization in latent opinions is defined as any pair of pre- and
+post-discussion opinion distributions where the mean increases in magnitude over
+time—formally this is $| \mupre | < | \mupost |$. We need to take magnitudes since radicalization may be towards increasingly negative or positive opinion values.
 
-Group-level opinion change, then, can be represented as a change in opinion mean, opinion variance,
-or both. Group polarization occurs when a group's average opinion increases over time. When
-likeminded people interact, they tend to come to agree more over time---they *find consensus*. By
-definition this means opinion variance decreases, whether the average changes or not.
+For group polarization experiments, only two times are relevant:
+*pre-discussion*, $t=\mathrm{pre}$, and *post-discussion*, $t=\mathrm{post}$.
+Since participant groups are assumed to be likeminded, we could also assume they come to agree more over time as well as become more extreme in our group polarization model, or in terms of the latent opinion variance, $\sigmapre > \sigmapost$—but only one study in our ten actually measured opinion variance to check this [@Schkade2010].
 
-We can define a null-polarization model, then, as one where the group's average opinion stays
-constant, but the variance decreases: $\sigmapost
-\leq \sigmapre$. Since the mean opinion is constant we just call it $\mu = \mupre = \mupost$. For
-conciseness, we define a null-polarization model as a 3-tuple, $\nu = (\mu, \sigmapre, \sigmapost)$.
+A *null-polarization hypothesis*, or *model*, is one where the latent mean stays
+constant, i.e. $\mu = \mu_\mathrm{pre} = \mu_\mathrm{post}$. *Mere agreement* is
+one type of null-polarization model, 
+the variance decreases over time to model agreement as described above.
+The mere agreement model is the cornerstone of our analysis, since it is empirically plausible and generates *spurious* observations of group polarization under mere agreement models, due to measurement artifacts induced by clipping all latent opinion values outside the measurement range—we call this *spurious polarization* for short.
+We simulate opinion measurements in two different ways, but both use the mere agreement model of latent opinions to test whether and by how much the measurement procedure distorts null-polarization to look like group polarization.
 
@@
 
-### Simulated group polarization measurements
+### The Expectation Model of Opinion Measurement
 
-We simulate the collection of an *ordinal distribution* of *ordinal observations*, as was performed
-in a prototypical group polarization experiment by integrating the latent opinion probability
-density function, $f(\omega)$, over each bin's corresponding range of opinion values. For example,
-consider an ordinal scale that runs from -3 (strongly disagree) to +3 (strongly agree) (as used by
-@Moscovici1969, for example). the frequency with which the "2" bin is chosen is just the total area
-underneath the probability density curve between 1.5 and 2.5.
+We use the \emph{expectation model} of opinion measurement to \emph{Identify} which, if any, mere agreement null-polarization models generate spurious group polarization in a particular experiment, $e$.
+The trick to this model, and the source of its name, is that we equate the theoretical \emph{expected value} of opinion measurements with the simulated \emph{observed average} of participant opinions.
 
-Formally, the probability density function is $$
-f(\omega) = 
-    \frac{1}{\sigma \sqrt{2 \pi}} e^{-\frac{(\omega - \mu)^2}{2 \sigma^2}}.
-$$ To set up our measurement simulation, we will represent oridnal bin values as $b_i$ with
-$i=1,\ldots,B$---i.e., there are $B$ bins. To create $B$ bins there must be $B+1$ *thresholds* that
-define what range of latent opinions correspond to which ordinal bin value.
+We simulate the measurement of the mean observed opinion at time $t$ 
+by first calculating integrating the normal probability density function over measurement bin boundaries $f(\omega)$, over each bin's corresponding range of opinion values.
+For example, consider an ordinal scale  that runs from -3 (strongly disagree) to +3 (strongly agree) (as
+used by @Moscovici1969, for example). the frequency with which the "2" bin is
+chosen is just the total area underneath the probability density curve between
+1.5 and 2.5.
 
-::: {text-align="center"}
-\[ TABLE [\[tab:experiment-parameters\]](#tab:experiment-parameters){reference-type="ref"
-reference="tab:experiment-parameters"} ABOUT HERE \]
-:::
-
-All bins have a width of 1 in latent opinion space, except for the two most extreme bins at opposite
-ends of the sentiment spectrum---this works because bin values are always separated by 1. All
-continuous latent opinion values $\omega < b_1 + \frac{1}{2} = \theta_1$ get mapped to the ordinal
-measurement value $b_1$---so, $b_1$'s lower threshold is $\theta_0 = -\infty$. Similarly, all
-opinions $\omega > b_B - 0.5 = \theta_B$ are reported as the maximum bin value, $b_B$, with an
-effective upper bound of $\theta_{B+1} = +\infty$. We can define all thresholds compactly as
-[@Liddell2018] $$
+All bins have a width of 1 in latent opinion space, except for the two most
+extreme bins at opposite ends of the sentiment spectrum---this works because bin
+values are always separated by 1. All continuous latent opinion values
+$\omega < b_1 + \frac{1}{2} = \theta_1$ get mapped to the ordinal measurement
+value $b_1$---so, $b_1$'s lower threshold is $\theta_0 = -\infty$. Similarly,
+all opinions $\omega > b_B - 0.5 = \theta_B$ are reported as the maximum bin
+value, $b_B$, with an effective upper bound of $\theta_{B+1} = +\infty$. We can
+define all thresholds compactly as [@Liddell2018] 
+$$
 \theta_j = \begin{cases}
@@
 
-We model the measurement of ordinal opinions by calculating the probability density function over
-bins, written $$
-f[b_i] = \int_{\theta_{i - 1}}^{\theta_i} f(\omega) d\omega,
-$$ switching to square brackets for the argument to $f$ now to emphasize that bins are discrete.
-With this ordinal measurement model we can caluclate the mean opinion of the simulated measurements
-like so $$\bar{o}_t = \frac{1}{B} \sum_{b=b_1}^{b_B} b \cdot f_t[b].$$
+The probability density that a participant reports the opinion in bin index $i$ at time $t$ is
+$$
+f_{it}(\mu, \sigma_t) = \int_{\theta_{i - 1}}^{\theta_{i}} f(\omega; \mu, \sigma_t)~d\omega,
+$$
+which we calculated in R to evaluate the integral at the two threshold values $\theta_{i-1}$ and $\theta_i$.
+Note that the normal probability density function we used is parameterized by mean $\mu$ and standard deviation $\sigma_t$.
+$$
+f(\omega; \mu, \sigma_t) = 
+    \frac{1}{\sigma_t \sqrt{2 \pi}} e^{-\frac{(\omega - \mu)^2}{2 \sigma_t^2}}.
+$$
 
-Group polarization is the difference between the post-deliberation mean observed opinion and the
-pre-deliberation one, which we can write as $$g = \bar{o}_\mathrm{post} - \bar{o}_\mathrm{pre}.
-  \label{eq:group-polarization}$$
+To simulate the measurement of the *average* observed opinion at time $t$, $\bar{o}_t$, as the expected value of the probability density function of observed ordinal opinions.
+In other words, we simulate $\bar{o}_t$ by calculating the weighted sum of products between the probability of observing each bin value and the bin value itself:
+$$
+\bar{o}_t(\nu, e) = \sum_{i=1}^B b_i f_{it}(\nu, e).
+$${#eq-sim-ave-observed}
+
+Group polarization is the difference between the post-deliberation mean observed
+opinion and the pre-deliberation one.
+We will calculate the magnitude of spurious group polarization under a mere agreement null model represented by its collected distribution parameters, $\nu = (\mu, \sigmapre, \sigmapost)$, under experimental design, $e$:
+$$
+g(\nu, e) = |\bar{o}_\mathrm{post}(\nu, e)| - |\bar{o}_\mathrm{pre}(\nu, e)|
+$${#eq-spurious-gp}
+
+In the *Inspect* stage of our analysis we will determine which group polarization replications are invalidated due to equifinality between the *mere agreement null-polarization hypothesis* on the one hand, and the *group polarization hypothesis*
+on the other.
+This requires more than just showing there is some $\nu$ that causes spurious polarization for some experiment where $g(\nu, e) > 0$.
+We have to find a $\nu$ that generates the exact $g(\nu, e)$ reported in the original journal article *and* check that simulated $\bar{o}_\mathrm{pre}$ and $\bar{o}_\mathrm{pre}$ match published values.
+
+::: {text-align="center"}
+\[ TABLE
+[\[tab:experiment-parameters\]](#tab:experiment-parameters){reference-type="ref"
+reference="tab:experiment-parameters"} ABOUT HERE \]
+:::
 
 ::: center
-\[ TABLE [\[tab:simulated-observations\]](#tab:simulated-observations){reference-type="ref"
+\[ TABLE
+[\[tab:simulated-observations\]](#tab:simulated-observations){reference-type="ref"
 reference="tab:simulated-observations"} ABOUT HERE \]
@@
 
-### Simulated spurious group polarization
+#### Example: the "Americans" condition in Moscovici and Zavalloni (1969)
 
-Group polarization measures radicalization in experimental echo chambers in terms of the difference
-between mean group opinions before and after the group discusses some survey topic. The following
-model shows the problem of how the measurement scheme above can make *mere agreement* spuriously
-look like polarization.
+One of the replications we invalidate here is from Moscovici and Zavalloni (1969),
+the "Americans" condition as they called it—we coded it `Moscovici1969_Americans`. 
+They report that participant opinions polarized after discussing the survey topic, "American economic aid is always used for political pressure." They report survey 
+responses increased in extremity from an average of -0.9 before discussion to
+-1.6 after, on a seven-bin scale
+from -3 indicating "strongly disagree" to +3 indicating "strongly agree".
 
-To make things more exact, we define a *null model*, $\nu$, of mere agreement defined as a 3-tuple:
-a constant mean for the pre- and post-discussion latent opinion distributions, and a
-pre-deliberation latent opinion variance that is greater than the post-deliberation variance to
-represent finding agreement, $$
-\nu = (\mu,~\sigmapre,~\sigmapost).
-$$ Group polarization is *spurious* in this case if measurement clipping causes the observed
-difference in pre- and post-discussion opinions to be greater than zero, i.e., $g > 0$ (Equation
-[\[eq:group-polarization\]](#eq:group-polarization){reference-type="ref"
-reference="eq:group-polarization"}, even if the true latent change is zero.
+Moscovici and Zavalloni say their results confirm the basic hypothesis, that \emph{group discussion causes polarization}.
+But we found a mere agreement null-polarization model that generates spurious group polarization using our model—specifically, $\nu = (\mu = -1.2, \sigma_\mathrm{pre}= 4.3, and \sigma_\mathrm{post}= 1.9)$.
+This means that the hypothesis is actually undecidable given the data, invalidating Moscovici and Zavalloni's replication.
+When a hypothesis and its null explain data equally well, the data tell us nothing.
 
-We will show how this arrangement generates spurious polarization in an experimental condition in
-our corpus, the *Americans* experimental condition in @Moscovici1969, to which we gave the
-Experiment ID `Moscovici1969_Americans`. In that condition, the authors used a seven-bin scale
-($B=7$) with numerical values from -3 to +3 indicating opinions from "strongly disagree" to
-"strongly agree". Pre-consensus opinions were reported to have mean -0.6 and the post-consensus mean
-was -1.04 [@Moscovici1969 p. 132, Table 4].
+Now let's use the expectation model to prove that this $\nu$ generates spurious group polarization in this experiment, i.e., $g(\nu, e= \mathrm{Moscovici1969\_Americans}) > 0$. 
+In the process we demonstrate how the expectation model works to simulate the measurement of spurious group polarization (@eq-spurious-gp).
 
-And now we will show that a null model with
-$\nu = (\mu = -1.2, \sigmapre = 4.3, and \sigmapost = 1.9)$ generates spurious polarization with
-exactly the pre- and post-discussion means Moscovici and Zavalloni report
-(Figure [\[fig:distros\]](#fig:distros){reference-type="ref" reference="fig:distros"}). First,
-calculate the vector of probability density values for each bin, $f[b]$, for the pre- and
-post-discussion latent opinion distributions using Equation [\[eq:\]](#eq:){reference-type="ref"
-reference="eq:"} (calculated with
-\['makeProbVec'\](https://github.com/mt-digital/gp-statmod/blob/final-draft/GroupPolarizationStatmod/model.R#L43)
-from the \[project code
-repository\](https://github.com/mt-digital/gp-statmod/tree/final-draft/GroupPolarizationStatmod)
-
+First, we calculate the probability density vector using the [`makeProbVec`](https://github.com/mt-digital/gp-statmod/blob/main/GroupPolarizationStatmod/model.R#L44) function in the file [`GroupPolarizationStatmod/model.R`](https://github.com/mt-digital/gp-statmod/blob/main/GroupPolarizationStatmod/model.R) in the [project code repository on GitHub](https://github.com/mt-digital/gp-statmod) for both experiment times: pre- and post-discussion:
 $$
-    f_\mathrm{pre}[b] = 
-$$ $$
-    f_\mathrm{post}[b] = 
+f_\mathrm{pre} = 
+\begin{pmatrix}
+  0.38 & 0.09 & 0.09 & 0.09 & 0.08 & 0.07 & 0.19
+\end{pmatrix}
+$$ 
+and
+$$
+f_\mathrm{post} = \
+\begin{pmatrix} 
+0.25 & 0.19 & 0.21 & 0.17 & 0.11 & 0.05 & 0.03 
+\end{pmatrix}.
 $$
 
+When we plug in the bin values and copmuted $f_\mathrm{pre}$ values into the simulated average pre-discussion opinion definition (@eq-sim-ave-observed),
+$$
+\bar{o}_\mathrm{pre} = -3 \cdot 0.38~~+-2\cdot 0.09~~+\ldots+3\cdot0.19 = -0.6
+$$
+$$
+\bar{o}_\mathrm{post} = -3 \cdot 0.25~~+-2\cdot 0.19~~+\ldots+3\cdot0.03 = -1.04
+$$
+So, we find that for the `Moscovici1969_Americans`, our null-polarization model appears polarized, with the exact same average pre- and post-discussion opinions reported by Moscovici and Zavalloni (1969), and the same magnitude of radicalization (CHECK THE PRESENTATION OF ALL THREE ABOVE)
+$$
+g(\nu, e) = \bar{o}_\mathrm{post} - \bar{o}_\mathrm{pre} = -1.04 - (-0.6) = - 0.4.
+$$
 ::: {style="text-align: center;"}
-FIGURE [\[fig:distros\]](#fig:distros){reference-type="ref" reference="fig:distros"} ABOUT HERE
+FIGURE [\[fig:distros\]](#fig:distros){reference-type="ref"
+reference="fig:distros"} ABOUT HERE 
 :::
 
+The replication of the `Moscovici1969_Americans` experimental condition is therefore invalidated since we proved equifinality between the group polarization hypothesis 
+(discussion induces polarization) and its null (discussion does not induce polarization).
+\emph{If the null fits, we must omit} the result from the set of results supporting any particular hypothesis.
+Therefore, despite being cited over 1,000 times as a replication, these citations are in fact erroneous—even if powerful authors hold them up as such.
+
+But how did we actually \emph{find} the null-polarization model that generated spurious group polarization?
+Below we demonstrate how we use a stochastic descent-style root finding that calculates spurious group polarization for a given $\nu$, then takes a step in a random direction, biased towards the direction of less error.
+It is therefore essential to use this efficient calculation simulation method, since tens, hundreds, or more evaluations are required for this root-finding algorithm.
+We show exactly how we put this to work in the following subsection explaining the algorithm to finish the \emph{Identification} step of our analysis.
+
+<!-- OLD STUFF BELOW BEFORE QMD REWRITE -->
+<!-- Moscovici and Zavalloni report -->
+<!-- (Figure [\[fig:distros\]](#fig:distros){reference-type="ref" -->
+<!-- reference="fig:distros"}). First, calculate the vector of probability density -->
+<!-- values for each bin, $f[b]$, for the pre- and post-discussion latent opinion -->
+<!-- distributions using Equation [\[eq:\]](#eq:){reference-type="ref" -->
+<!-- reference="eq:"} (calculated with -->
+<!-- \['makeProbVec'\](https://github.com/mt-digital/gp-statmod/blob/final-draft/GroupPolarizationStatmod/model.R#L43) -->
+<!-- from the \[project code -->
+<!-- repository\](https://github.com/mt-digital/gp-statmod/tree/final-draft/GroupPolarizationStatmod) -->
 ## Root-finding Algorithm to Find Null-Polarization Models with Artifacts
@@
 
-To analyze published replications we need to expand our notation to index individual *experiments*
-with replications, and from which different journal articles, track and analyze it,
+To analyze published replications we need to expand our notation to index
+individual *experiments* with replications, and from which different journal
+articles, track and analyze it,
 

### 2025-11-24 GroupPolarizationStatmod/paper/method.tex
worked out code example for calculating observed means from latent opinion distributions
finished expectation model subsection

@@
 
-Group polarization measures radicalization in experimental echo chambers in terms of
-the difference between mean group opinions before and after the group discusses some
-survey topic. The following model shows the problem of how the measurement scheme 
-above can make \emph{mere agreement} spuriously look like polarization.
+Group polarization measures radicalization in experimental echo chambers in
+terms of the difference between mean group opinions before and after the
+group discusses some survey topic. The following model shows the problem of
+how the measurement scheme above can make \emph{mere agreement} spuriously
+look like polarization.
 
-To make things more exact, we define a \emph{null model}, $\nu$, of mere agreement
-defined as a 3-tuple: a constant mean for the pre- and post-discussion latent opinion
-distributions, and a pre-deliberation latent opinion variance that is greater than the
-post-deliberation variance to represent finding agreement,
-\begin{equation}
-  \nu = (\mu,~\sigmapre,~\sigmapost).
-  \label{eq:nu-definition}
-\end{equation}
-\noindent
-Group polarization is \emph{spurious} in this case if measurement clipping causes 
-the observed difference in pre- and post-discussion opinions to be greater than zero,
-i.e., $g > 0$ (Equation \ref{eq:group-polarization}, 
+To make things more exact, we define a \emph{null model}, $\nu$, of mere
+agreement defined as a 3-tuple: a constant mean for the pre- and
+post-discussion latent opinion distributions, and a pre-deliberation latent
+opinion variance that is greater than the post-deliberation variance to
+represent finding agreement, \begin{equation} \nu =
+(\mu,~\sigmapre,~\sigmapost).  \label{eq:nu-definition} \end{equation}
+\noindent Group polarization is \emph{spurious} in this case if measurement
+clipping causes the observed difference in pre- and post-discussion opinions
+to be greater than zero, i.e., $g > 0$ (Equation \ref{eq:group-polarization},
 even if the true latent change is zero.
@@
 We will show how this arrangement generates spurious polarization in an
-experimental condition in our corpus, 
-the \emph{Americans} experimental condition in
-\textcite{Moscovici1969}, to which we gave the Experiment ID
-\texttt{Moscovici1969\_Americans}. In that condition, the authors used a seven-bin
-scale ($B=7$) with numerical values from -3 to +3 indicating opinions from ``strongly
-disagree'' to ``strongly agree''. Pre-consensus opinions were reported to have
-mean -0.6 and the post-consensus mean was -1.04 \cite[p. 132, Table 4]{Moscovici1969}.
+experimental condition in our corpus, the \emph{Americans} experimental
+condition in \textcite{Moscovici1969}, to which we gave the Experiment ID
+\texttt{Moscovici1969\_Americans}. In that condition, the authors used a
+seven-bin scale ($B=7$) with numerical values from -3 to +3 indicating
+opinions from ``strongly disagree'' to ``strongly agree''. Pre-consensus
+opinions were reported to have mean -0.6 and the post-consensus mean was
+-1.04 \cite[p. 132, Table 4]{Moscovici1969}.
 

### 2025-11-25 GroupPolarizationStatmod/paper/method.qmd
started on specify sig thresh

@@
 ---
+title: Method
 editor: 
@@
     wrap: 80
+author: Matt Turner · [https://mat.phd](https://mat.phd)
+date: today
+toc: true
 bibliography: this.bib
@@
 
-# Method
-
 A group polarization replication is proven undecidable, and therefore not truly
@@
 
-Null-polarization We thus model opinions at each time, $t=\mathrm{pre}$ and
-$t=\mathrm{post}$, as random variables drawn from a normal distribution with
-mean $\mu_t$ and standard deviation $\sigma_t$. we drop the time
+# Latent Opinion Model
 
-The null-polarization distributions are represented by $\nu$.
+Throughout this paper we model a participant $i$'s latent opinion at time $t$ as
+a normally-distributed random variable,
+$$ 
+\omega_{i,t} \sim
+\mathcal{N}(\mu_t, \sigma_t),
+$$ {#eq-random-variable-def}
+with mean $\mu_t$ and standard deviation $\sigma_t$. 
+Group polarization in terms of latent opinions is then defined as any pair of pre- and
+post-discussion opinion distributions where the mean increases in magnitude over
+time—formally, $| \mupost | > | \mupre |$. We need to take magnitudes since radicalization may be towards increasingly negative or positive opinion values.
 
-in a simple, general way---i.e., a decrease in opinion variance but a static
-opinion mean---we represent . In the general group polarization hypothesis, the
-group is assumed to start out mostly agreeing in principle (i.e., mostly all
-agree or disagree) but to different extremes. After discussion
+For group polarization experiments, only two times are relevant:
+*pre-discussion*, $t=\mathrm{pre}$, and *post-discussion*, $t=\mathrm{post}$.
+Since participant groups are assumed to be likeminded, we could also assume they come to agree more over time as well as become more extreme in our group polarization model, or in terms of the latent opinion variance, $\sigmapre > \sigmapost$—but only one study in our ten actually measured opinion variance to check this [@Schkade2010].
 
-Showing a replication is undecidable is sufficient to undermine it. However, it
-does not tell us what is the probability that a particular experimental design
-produces false positives, due to clipping or any other problem.
+A *null-polarization model* is a model that is consistent with the *negation*, or *null*, of the  *group polarization hypothesis*.
+The group polarization hypothesis is that group discussion causes the average group opinion to increase, assuming the group is aligned on a particular issue, though opinion strength varies.
+Therefore a null model is one where the latent mean stays
+constant, i.e., $\mu = \mu_\mathrm{pre} = \mu_\mathrm{post}$.
 
-## Simulation of Mere Consensus and Opinion Change Measurement
+*Mere agreement* is the type of null-polarization model we use here.
+Mere agreement is a null-model because the mean latent opinion is set to be constant before and after discussion.
+We represent *agreement*—or equivalently, *consensus*—as a decrease in the standard deviation that defines the pre- and post-discussion latent opinion distributions.
+The mere agreement model is the cornerstone of our analysis, since it is empirically plausible and generates *spurious* observations of group polarization under mere agreement models, due to measurement artifacts induced by clipping all latent opinion values outside the measurement range—we call this *spurious polarization* for short.
+We simulate opinion measurements in two different ways, but both use the mere agreement model of latent opinions to test whether and by how much the measurement procedure distorts null-polarization to look like group polarization.
 
-We model \emph{mere consensus} and opinion change measurement in group
-polarization experiments to examine whether published replications can be
-explained by this particular null polarization model. We simulate consensus and
-measurement two ways for complementary purposes. First is this *expectation*
-model, and the second is the *stochastic* model. Both models assume group
-opinions are normally distributed, which is practical due to its simplicity.
+::: {style="text-align: center;"}
+\[ TABLE \[tab:latent-variables\](#tab:latent-variables){reference-type="ref"
+reference="tab:latent-variables"} ABOUT HERE \]
+:::
 
-We assume that participants' true latent opinions at some time, $t$, are random
-variables drawn from a normal distribution with mean $\mu_t$ and standard
-deviation $\sigma_t$. A normal distribution is scientifically justifiable, since
-opinions are aggregations of many constituent electrical and chemical signals in
-the brain—the distribution of aggregations like this tend to be *normal* (hence
-the name). A normal distribution also makes sense because several of the
-analysis procedures from the corpus use a $t$-test or other basic linear
-regression-style test, which implicitly assume data are distributed normally.
+## Two Ways to Model Measurement that Extend the Latent Opinion Model
 
-The two models differ in how they simulate opinions, opinion change, and opinion
-change measurement. We use the *expectation model* as a computationally
-efficient way to simulate the observed mean opinion as the mathematical
-*expectation value* of the distribution of measured opinions. In the second
-model, the *stochastic model*, we simulate participant survey reporting as
-drawing a latent opinion from the latent opinion distribution, then binning that
-opinion in the associated opinion bin. We use the expectation model for its
+We simulate measurement in two different ways—i.e., with two different models—for the two parts of our analysis. The first
+is the *expectation model*: an
+efficient way to *simulate measurement of the observed average* opinion of a group, using the *mathematical expectation value* of the distribution of measured opinions as a proxy. In the second
+model, the *stochastic model*, we simulate individual participant responses by first drawing an opinion for each participant from the latent opinion distribution (@eq-.
+The opinions are binned to the appropriate ordinal value.
+
+We use the expectation model for its
 efficiency in the first step of our analysis: to find a *null-polarization
@@
 
-### Latent Opinion Model
+The measurement models are introduced in detail along with the corresponding stage of the analysis that uses them.
+First, the *Identification* stage that uses the *expectation* measurement model; second, the *Evaluation* stage that uses the *stochastic* measurement model.
 
-Throughout this paper we model a participant $i$'s latent opinion at time $t$ as
-a normally-distributed random variable,
+# *Identification* of Invalid Replications
 
-$$ 
-\omega_{i,t} \sim
-\mathcal{N}(\mu_t, \sigma_t).  
-$$
-Group polarization in latent opinions is defined as any pair of pre- and
-post-discussion opinion distributions where the mean increases in magnitude over
-time—formally this is $| \mupre | < | \mupost |$. We need to take magnitudes since radicalization may be towards increasingly negative or positive opinion values.
+In the first stage of our analysis, the *Identification* stage, we check if we can *invalidate* a replication of group polarization achieved in some experiment/experimental condition, which we label $e$.
+Our invalidation strategy is to find a mere agreement null-polarization model, call it $\nu$, that predicts the exact observations reported in each experimental replication in our corpus.
+If one is found, then we have found a model consistent with the *negation* of the group polarization hypothesis that explains the data as well as the group polarization hypothesis itself.
+In other words, the hypothesis is not proved or disproved given the data, but *undecidable*: the data cannot *decide* the hypothesis one way or another because of potential measurement artifacts.
+Because the experiment failed to produce a valid replication, the results are effectively stripped of evidentiary value and warrant exclusion from the aggregate evidence base.
 
-For group polarization experiments, only two times are relevant:
-*pre-discussion*, $t=\mathrm{pre}$, and *post-discussion*, $t=\mathrm{post}$.
-Since participant groups are assumed to be likeminded, we could also assume they come to agree more over time as well as become more extreme in our group polarization model, or in terms of the latent opinion variance, $\sigmapre > \sigmapost$—but only one study in our ten actually measured opinion variance to check this [@Schkade2010].
+But how do we find a $\nu$ that generates spurious group polarization?
+And how do we even simulate the group polarization experiments and measurements that make mere agreement in latent opinions look like group polarization?
+We simulate
+experimental measurements of group polarization by using the mathematical *expectation value* of latent opinions defined by $\nu$ as a proxy for the observed average value of ordinal-valued reported opinions (described in the first subsection below).
+We then use this measurement model in a numerical root-finding algorithm that identifies a $\nu$ that matches published observations, if possible.
+If we find such a $\nu$, this means we have a model
+consistent with the null hypothesis that explains published observations.
+In other words, the replication is invalidated since the group polarization hypothesis is *undecidable* given the data—i.e., it cannot decide whether the hypothesis or its null is correct.
 
-A *null-polarization hypothesis*, or *model*, is one where the latent mean stays
-constant, i.e. $\mu = \mu_\mathrm{pre} = \mu_\mathrm{post}$. *Mere agreement* is
-one type of null-polarization model, 
-the variance decreases over time to model agreement as described above.
-The mere agreement model is the cornerstone of our analysis, since it is empirically plausible and generates *spurious* observations of group polarization under mere agreement models, due to measurement artifacts induced by clipping all latent opinion values outside the measurement range—we call this *spurious polarization* for short.
-We simulate opinion measurements in two different ways, but both use the mere agreement model of latent opinions to test whether and by how much the measurement procedure distorts null-polarization to look like group polarization.
+So, to review, this subsection introduces how the *Identification* stage of the analysis works.
+In this stage we identify for which experiments, $e$, does a mere agreement null model look like group polarization under $e$'s measurement procedure.
 
-::: {style="text-align: center;"}
-\[ TABLE \[tab:latent-variables\](#tab:latent-variables){reference-type="ref"
-reference="tab:latent-variables"} ABOUT HERE \]
-:::
+## The Expectation Model of Opinion Measurement
 
-### The Expectation Model of Opinion Measurement
-
-We use the \emph{expectation model} of opinion measurement to \emph{Identify} which, if any, mere agreement null-polarization models generate spurious group polarization in a particular experiment, $e$.
-The trick to this model, and the source of its name, is that we equate the theoretical \emph{expected value} of opinion measurements with the simulated \emph{observed average} of participant opinions.
+We use the *expectation model* of opinion measurement to *Identify* which, if any, mere agreement null-polarization models generate spurious group polarization in a particular experiment, $e$.
+The trick to this model, and the source of its name, is that we use the theoretical *expected value* of opinion measurements as a proxy for the *observed average* of participant opinions under an experimental design, $e$.
 
 We simulate the measurement of the mean observed opinion at time $t$ 
-by first calculating integrating the normal probability density function over measurement bin boundaries $f(\omega)$, over each bin's corresponding range of opinion values.
+by first calculating integrating the normal probability density function,  $f(\omega)$, over measurement bin boundaries, over each bin's corresponding range of opinion values.
 For example, consider an ordinal scale  that runs from -3 (strongly disagree) to +3 (strongly agree) (as
@@
 
-#### Example: the "Americans" condition in Moscovici and Zavalloni (1969)
+## Example: the "Americans" condition in Moscovici and Zavalloni (1969)
 
@@
 $$
+
 ::: {style="text-align: center;"}
@@
 
-The replication of the `Moscovici1969_Americans` experimental condition is therefore invalidated since we proved equifinality between the group polarization hypothesis 
-(discussion induces polarization) and its null (discussion does not induce polarization).
-\emph{If the null fits, we must omit} the result from the set of results supporting any particular hypothesis.
-Therefore, despite being cited over 1,000 times as a replication, these citations are in fact erroneous—even if powerful authors hold them up as such.
+Since the simulated average opinions match the observed ones—note that as a consequence, the spurious polarization magnitude also matches the observed one.
+In other words, we *identified* a null model that explains the data. 
+If both a hypothesis and its null explain data equally well, then the hypothesis is *undecidable*
 
-But how did we actually \emph{find} the null-polarization model that generated spurious group polarization?
-Below we demonstrate how we use a stochastic descent-style root finding that calculates spurious group polarization for a given $\nu$, then takes a step in a random direction, biased towards the direction of less error.
-It is therefore essential to use this efficient calculation simulation method, since tens, hundreds, or more evaluations are required for this root-finding algorithm.
-We show exactly how we put this to work in the following subsection explaining the algorithm to finish the \emph{Identification} step of our analysis.
+In this example we started with
 
-<!-- OLD STUFF BELOW BEFORE QMD REWRITE -->
-<!-- Moscovici and Zavalloni report -->
-<!-- (Figure [\[fig:distros\]](#fig:distros){reference-type="ref" -->
-<!-- reference="fig:distros"}). First, calculate the vector of probability density -->
-<!-- values for each bin, $f[b]$, for the pre- and post-discussion latent opinion -->
-<!-- distributions using Equation [\[eq:\]](#eq:){reference-type="ref" -->
-<!-- reference="eq:"} (calculated with -->
-<!-- \['makeProbVec'\](https://github.com/mt-digital/gp-statmod/blob/final-draft/GroupPolarizationStatmod/model.R#L43) -->
-<!-- from the \[project code -->
-<!-- repository\](https://github.com/mt-digital/gp-statmod/tree/final-draft/GroupPolarizationStatmod) -->
 ## Root-finding Algorithm to Find Null-Polarization Models with Artifacts
 
-## Corpus Analysis to Determine Undecidable Replications
+# Evaluation of Experimental Designs
 
-To analyze published replications we need to expand our notation to index
-individual *experiments* with replications, and from which different journal
-articles, track and analyze it,
+- The evaluation stage complements our *Identify* stage by filling in more details about the experimental design.
 
-## Evaluating Grouop Polarization Experimental Designs
+## False positive rate
+
+First, how frequently do the $\nu$ that we found generate simulated data
+  that fool a rigorous statistical design that properly accounts for the measurement process? This serves two roles: 
+  1.
+Confirm that the $\nu$ we found could create simulated datasets that appear polarized under statistical analysis, which is different from using the expected value of an exact latent opinion distribution, integrated with the bin values to simulate the average opinion.
+  2. Evaluate the underlying experimental design independent of the failure to account for how ordinal measurements inevitably clip extreme latent opinions.
+  3. We do this by simulating experiment data using $\nu$ we identified in the previous stage. We simulate 1000 experiments indexed by $i$ for each experiment $e$.
+
+To calculate the false detection rate we begin by simulating 
+pre- and post-deliberation 
+opinions by drawing $N_e$ samples from normal latent opinion
+distributions that generated spurious group polarization, i.e., one sample at each experiment stage for each simulated participant.
+The latent opinion distributions are specified with the mere agreement null model labelled $\nu_e$.
+
+Each trial $i$ of $N_e$ simulated pre- and post-deliberation observations is then inspected to see
+if spurious group polarization occurred in the simulation. Spurious group
+polarization is determined by first inferring the latent parameters of each
+simulated experimental trial by fitting an ordered probit model~\cite{Liddell2018}.
+The procedure infers the latent distribution parameters that we in fact specified with $\nu$: $\mu_\mathrm{pre}$, $\sigma_\mathrm{pre}$, $\mu_\mathrm{post}$, and $\sigma_\mathrm{post}$.
+
+The inferred parameters are then used to calculate the effect size for that
+trial, Cohen's $d$, defined as~\cite[p. 331]{Liddell2018}
+$$
+  d = \frac{\mupost - \mupre}{\sqrt{\frac{1}{2}(\sigmapost^2 + \sigmapre^2)}}.
+$$ {#eq-cohens}
+It is implied that these values are the inferred values, not the values from $\nu$, when used in the context of calculating Cohen's $d$. Note $d=0$ always for the true $\nu$.
+
+The family-wise error rate is the fraction of trials 
+with $d_{ei} > d^*$, written
+$$
+  \alpha_e = \frac{1}{N_T} \sum_i^{N_T} \Theta(d_{ei}; d^*),
+$$ {#eq-study-aggregate}
+where
+$$
+  \Theta(d_{ei}; d^*) = 
+    \begin{cases}
+      1 & \text{ if } d_{ei} > d^* \\
+      0 & \text{ otherwise.}
+    \end{cases}
+$$
+$d^*$ is the *significance threshold*, which must be specified by the experimenter.
+There are three common significance levels which we adopt here: $d^* = 0.2$ for "low", $d^*=0.5$ for "medium", and $d^* = 0.8$ for "high".
+
+To calculate the family-wise error rate
+across all experimental conditions of a study, $s$, 
+\begin{equation}
+  \alpha_s = \frac{1}{|s|} \sum_{e \in s} \alpha_e,
+\end{equation}
+\noindent
+where $e \in s$ indicates sum over all experimental conditions $e$ presented in the article associated with study $s$.
+The mean across all results identified as plausibly spurious is 
+\begin{equation}
+  \alpha_\mathrm{all} = \frac{1}{|E|} \sum_{e \in E} \alpha_e.
+\end{equation}
+
+In the final part of our *Evaluation* analysis we adapt this analysis to re-align the question of statistical rigor: we can be specific now, and instead of wondering, "what significance level is truly rigorous", we can use our techniques to make the choice of $d^*$ to conform to a specific design requirement.
+Specifically we will calculate what $d^*$ must be in order to limit $\alpha \leq 0.05$ as an example.
+
+## False discovery rate
+
+We finally may calculate the false discovery rate by plugging this $\alpha$
+into Equation~\ref{eq:fdr} and assuming empirically-motivated values for 
+base rate, $b$, and statistical power, $W$, as well as more pessimistic and
+optimistic settings for base rate and power.
+
+Using the notation outlined above, the false detection rate is
+\begin{equation}
+  \fdr = \frac{(1 - b)\alpha}{(1 - b)\alpha + bW}.
+  \label{eq:fdr}
+\end{equation}
+
+## *Specify* Significance Threshold for Low False Positive Rates
+
+In the final analysis step we calculate what significance threshold, $d^*$,
+would limit the false positive rate to 0.05 or less.
+
+There are many possible
+null models we could choose.
+
+what significance threshold would have limited inverting this procedure to find what significance value
+$d^*$ is necessary to achieve a low family-wise error rate $\alpha_e = 0.05$ for
+experimental condition, $e$.
+
+To do this we first calculate $\alpha_e(d^*)$ for a
+sufficiently wide range of $d^*$. We then find the lowest $d^*$ for which
+$\alpha_e(d^*) \leq 0.05$.

### 2025-11-25 GroupPolarizationStatmod/scripts/analysis.R

@@ -111,8 +111,6 @@ NON_IDENTIFIED =
 #
 sigval_for_low_fwer = function(fdr_vs_sig_tbl, target_fwer = 0.05) {
 
-  # return (fdr_vs_sig_tbl)
-
   return(
 
     fdr_vs_sig_tbl %>%
@@ -120,12 +118,9 @@ sigval_for_low_fwer = function(fdr_vs_sig_tbl, target_fwer = 0.05) {
       mutate(StudyID = fct_reorder(StudyID, FWER, mean), 
              ExperimentID = fct_reorder(ExperimentID, FWER, mean)) %>%
       filter(FWER <= target_fwer) %>% 
-      # group_by(StudyID, ExperimentID) %>% 
       group_by(ExperimentID) %>% 
       filter(SigVal == min(SigVal))  %>% 
-      mutate(SigVal = as_factor(SigVal), ExperimentID = fct_reorder(ExperimentID, FDR)) # %>%
-      # arrange(desc(SigVal)) 
-
+      mutate(SigVal = as_factor(SigVal), ExperimentID = fct_reorder(ExperimentID, FDR)) 
   )
 }
 

### 2025-11-25 GroupPolarizationStatmod/paper/method.qmd
magic expanding methods section

@@
 bibliography: this.bib
+format:
+  html:
+    grid:
+      body-width: 880px  
 ---
@@
 Throughout this paper we model a participant $i$'s latent opinion at time $t$ as
-a normally-distributed random variable,
-$$ 
+a normally-distributed random variable, $$ 
 \omega_{i,t} \sim
 \mathcal{N}(\mu_t, \sigma_t),
-$$ {#eq-random-variable-def}
-with mean $\mu_t$ and standard deviation $\sigma_t$. 
-Group polarization in terms of latent opinions is then defined as any pair of pre- and
-post-discussion opinion distributions where the mean increases in magnitude over
-time—formally, $| \mupost | > | \mupre |$. We need to take magnitudes since radicalization may be towards increasingly negative or positive opinion values.
+$$ {#eq-random-variable-def} with mean $\mu_t$ and standard deviation
+$\sigma_t$. Group polarization in terms of latent opinions is then defined as
+any pair of pre- and post-discussion opinion distributions where the mean
+increases in magnitude over time—formally,
+$| \mu_\mathrm{post}| > | \mu_\mathrm{pre}|$. We need to take magnitudes since
+radicalization may be towards increasingly negative or positive opinion values.
 
@@
 *pre-discussion*, $t=\mathrm{pre}$, and *post-discussion*, $t=\mathrm{post}$.
-Since participant groups are assumed to be likeminded, we could also assume they come to agree more over time as well as become more extreme in our group polarization model, or in terms of the latent opinion variance, $\sigmapre > \sigmapost$—but only one study in our ten actually measured opinion variance to check this [@Schkade2010].
+Since participant groups are assumed to be likeminded, we could also assume they
+come to agree more over time as well as become more extreme in our group
+polarization model, or in terms of the latent opinion variance,
+$\sigma_\mathrm{pre}> \sigma_\mathrm{post}$—but only one study in our ten
+actually measured opinion variance to check this [@Schkade2010].
 
-A *null-polarization model* is a model that is consistent with the *negation*, or *null*, of the  *group polarization hypothesis*.
-The group polarization hypothesis is that group discussion causes the average group opinion to increase, assuming the group is aligned on a particular issue, though opinion strength varies.
-Therefore a null model is one where the latent mean stays
+A *null-polarization model* is a model that is consistent with the *negation*,
+or *null*, of the *group polarization hypothesis*. The group polarization
+hypothesis is that group discussion causes the average group opinion to
+increase, assuming the group is aligned on a particular issue, though opinion
+strength varies. Therefore a null model is one where the latent mean stays
 constant, i.e., $\mu = \mu_\mathrm{pre} = \mu_\mathrm{post}$.
 
-*Mere agreement* is the type of null-polarization model we use here.
-Mere agreement is a null-model because the mean latent opinion is set to be constant before and after discussion.
-We represent *agreement*—or equivalently, *consensus*—as a decrease in the standard deviation that defines the pre- and post-discussion latent opinion distributions.
-The mere agreement model is the cornerstone of our analysis, since it is empirically plausible and generates *spurious* observations of group polarization under mere agreement models, due to measurement artifacts induced by clipping all latent opinion values outside the measurement range—we call this *spurious polarization* for short.
-We simulate opinion measurements in two different ways, but both use the mere agreement model of latent opinions to test whether and by how much the measurement procedure distorts null-polarization to look like group polarization.
+*Mere agreement* is the type of null-polarization model we use here. Mere
+agreement is a null-model because the mean latent opinion is set to be constant
+before and after discussion. We represent *agreement*—or equivalently,
+*consensus*—as a decrease in the standard deviation that defines the pre- and
+post-discussion latent opinion distributions. The mere agreement model is the
+cornerstone of our analysis, since it is empirically plausible and generates
+*spurious* observations of group polarization under mere agreement models, due
+to measurement artifacts induced by clipping all latent opinion values outside
+the measurement range—we call this *spurious polarization* for short. We
+simulate opinion measurements in two different ways, but both use the mere
+agreement model of latent opinions to test whether and by how much the
+measurement procedure distorts null-polarization to look like group
+polarization.
 
@@
 
-## Two Ways to Model Measurement that Extend the Latent Opinion Model
+# Two Models of Group Polarization Measurement
 
-We simulate measurement in two different ways—i.e., with two different models—for the two parts of our analysis. The first
-is the *expectation model*: an
-efficient way to *simulate measurement of the observed average* opinion of a group, using the *mathematical expectation value* of the distribution of measured opinions as a proxy. In the second
-model, the *stochastic model*, we simulate individual participant responses by first drawing an opinion for each participant from the latent opinion distribution (@eq-.
-The opinions are binned to the appropriate ordinal value.
+We simulate measurement in two different ways—i.e., with two different
+models—for different parts of our analysis. The first is the *expectation
+model*: an efficient way to *simulate measurement of the observed average*
+opinion of a group, using the *mathematical expectation value* of the
+distribution of measured opinions as a proxy. In the second model, the
+*stochastic model*, we simulate individual participant responses by first
+drawing an opinion for each participant from the latent opinion distribution
+(@eq-random-variable-def). The opinions are binned to the appropriate ordinal
+value.
 
-We use the expectation model for its
-efficiency in the first step of our analysis: to find a *null-polarization
-model* that also fits the data for each experimental condition studied—if we
-find one that means we must retract this replication since this proves that the
-group polarization is undecidable with the data we have. Once we find any such
-undecidable results, we use the null-polarization model to help us further
-evaluate the experimental design itself.
+We use the expectation model for its efficiency in the first step of our
+analysis: to find a *null-polarization model* that also fits the data for each
+experimental condition studied—if we find one that means we must retract this
+replication since this proves that the group polarization is undecidable with
+the data we have. Once we find any such undecidable results, we use the
+null-polarization model to help us further evaluate the experimental design
+itself.
 
@@
 
-The measurement models are introduced in detail along with the corresponding stage of the analysis that uses them.
-First, the *Identification* stage that uses the *expectation* measurement model; second, the *Evaluation* stage that uses the *stochastic* measurement model.
+The measurement models are introduced in detail along with the corresponding
+stage of the analysis that uses them. First, the *Identification* stage that
+uses the *expectation* measurement model; second, the *Evaluation* stage that
+uses the *stochastic* measurement model.
 
-# *Identification* of Invalid Replications
+# Part 1: *Identify* Invalid Replications
 
-In the first stage of our analysis, the *Identification* stage, we check if we can *invalidate* a replication of group polarization achieved in some experiment/experimental condition, which we label $e$.
-Our invalidation strategy is to find a mere agreement null-polarization model, call it $\nu$, that predicts the exact observations reported in each experimental replication in our corpus.
-If one is found, then we have found a model consistent with the *negation* of the group polarization hypothesis that explains the data as well as the group polarization hypothesis itself.
-In other words, the hypothesis is not proved or disproved given the data, but *undecidable*: the data cannot *decide* the hypothesis one way or another because of potential measurement artifacts.
-Because the experiment failed to produce a valid replication, the results are effectively stripped of evidentiary value and warrant exclusion from the aggregate evidence base.
+In the first stage of our analysis, we try to *identify* a null polarization
+models, $\nu$, that invalidate published detections of group polarization if the
+measurement procedure induces *spurious group polarization* for that null
+polarization model. Null polarization can take many forms—all that is required
+is that the latent mean opinion stays constant before and after discussion. Here
+we use a *mere agreement* null polarization model, where we simulate
+*consensus*, the tendency for sufficiently similar opinions to become more
+similar—but we leave out polarization. In mere agreement, moderate and extreme
+opinions approach change by equal magnitude towards each other. We therefore
+represent consensus as a decrease in the standard deviation of the latent
+opinion distribution from pre- to post-discussion. We simulate the measurement
+of the group's average opinion before and after discussion and measure the
+spurious group polarization, if there is any. We use this procedure in a search
+algorithm we developed to identify which combination of constant latent mean and
+pre- and post-discussion opinion variances generate spurious group polarization,
+if there are any.
 
-But how do we find a $\nu$ that generates spurious group polarization?
-And how do we even simulate the group polarization experiments and measurements that make mere agreement in latent opinions look like group polarization?
-We simulate
-experimental measurements of group polarization by using the mathematical *expectation value* of latent opinions defined by $\nu$ as a proxy for the observed average value of ordinal-valued reported opinions (described in the first subsection below).
-We then use this measurement model in a numerical root-finding algorithm that identifies a $\nu$ that matches published observations, if possible.
-If we find such a $\nu$, this means we have a model
-consistent with the null hypothesis that explains published observations.
-In other words, the replication is invalidated since the group polarization hypothesis is *undecidable* given the data—i.e., it cannot decide whether the hypothesis or its null is correct.
+Our invalidation strategy, then, is to find a mere agreement null polarization
+model, $\nu$, that predicts the exact observations reported in each experimental
+replication in our corpus. If one is found, then we have found a model
+consistent with the *negation* of the group polarization hypothesis that
+explains the data as well as the group polarization hypothesis itself. In other
+words, the hypothesis is not proved or disproved given the data, but
+*undecidable*: the data cannot *decide* the hypothesis one way or another
+because of potential measurement artifacts. Because the experiment failed to
+produce a valid replication, the results are effectively stripped of evidentiary
+value and warrant exclusion from the aggregate evidence base. We apply our
+analysis to every experimental condition across all ten journal articles in our
+corpus, which we label $e$.
 
-So, to review, this subsection introduces how the *Identification* stage of the analysis works.
-In this stage we identify for which experiments, $e$, does a mere agreement null model look like group polarization under $e$'s measurement procedure.
+But how do we find a $\nu$ that generates spurious group polarization? And how
+do we even simulate the group polarization experiments and measurements that
+make mere agreement in latent opinions look like group polarization? We simulate
+experimental measurements of group polarization by using the mathematical
+*expectation value* of latent opinions defined by $\nu$ as a proxy for the
+observed average value of ordinal-valued reported opinions (described in the
+first subsection below). We then use this measurement model in a numerical
+root-finding algorithm that identifies a $\nu$ that matches published
+observations, if possible. If we find such a $\nu$, this means we have a model
+consistent with the null hypothesis that explains published observations. In
+other words, the replication is invalidated since the group polarization
+hypothesis is *undecidable* given the data—i.e., it cannot decide whether the
+hypothesis or its null is correct.
 
-## The Expectation Model of Opinion Measurement
+So, to review, this subsection introduces how we identify which experimental
+data can be explained by a mere agreement null model, where the ordinal survey
+measurement procedure induces a clipping artifact that hides changes in extreme
+opinions that fall outside the measurement scale.
 
-We use the *expectation model* of opinion measurement to *Identify* which, if any, mere agreement null-polarization models generate spurious group polarization in a particular experiment, $e$.
-The trick to this model, and the source of its name, is that we use the theoretical *expected value* of opinion measurements as a proxy for the *observed average* of participant opinions under an experimental design, $e$.
+## The Expectation Model of Spurious Group Polarization
 
-We simulate the measurement of the mean observed opinion at time $t$ 
-by first calculating integrating the normal probability density function,  $f(\omega)$, over measurement bin boundaries, over each bin's corresponding range of opinion values.
-For example, consider an ordinal scale  that runs from -3 (strongly disagree) to +3 (strongly agree) (as
-used by @Moscovici1969, for example). the frequency with which the "2" bin is
-chosen is just the total area underneath the probability density curve between
-1.5 and 2.5.
+We use the *expectation model* of opinion measurement to *Identify* which, if
+any, mere agreement null-polarization models generate spurious group
+polarization in a particular experiment, $e$. The trick to this model, and the
+source of its name, is that we use the theoretical *expected value* of opinion
+measurements as a proxy for the *observed average* of participant opinions under
+an experimental design, $e$.
+
+We simulate the measurement of the mean observed opinion at time $t$ by first
+calculating integrating the normal probability density function, $f(\omega)$,
+over measurement bin boundaries, over each bin's corresponding range of opinion
+values. For example, consider an ordinal scale that runs from -3 (strongly
+disagree) to +3 (strongly agree) (as used by @Moscovici1969, for example). the
+frequency with which the "2" bin is chosen is just the total area underneath the
+probability density curve between 1.5 and 2.5.
 
@@
 value, $b_B$, with an effective upper bound of $\theta_{B+1} = +\infty$. We can
-define all thresholds compactly as [@Liddell2018] 
-$$
+define all thresholds compactly as [@Liddell2018] $$
 \theta_j = \begin{cases}
@@
 
-The probability density that a participant reports the opinion in bin index $i$ at time $t$ is
-$$
+The probability density that a participant reports the opinion in bin index $i$
+at time $t$ is $$
 f_{it}(\mu, \sigma_t) = \int_{\theta_{i - 1}}^{\theta_{i}} f(\omega; \mu, \sigma_t)~d\omega,
-$$
-which we calculated in R to evaluate the integral at the two threshold values $\theta_{i-1}$ and $\theta_i$.
-Note that the normal probability density function we used is parameterized by mean $\mu$ and standard deviation $\sigma_t$.
-$$
+$$ which we calculated in R to evaluate the integral at the two threshold values
+$\theta_{i-1}$ and $\theta_i$. Note that the normal probability density function
+we used is parameterized by mean $\mu$ and standard deviation $\sigma_t$. $$
 f(\omega; \mu, \sigma_t) = 
@@
 
-To simulate the measurement of the *average* observed opinion at time $t$, $\bar{o}_t$, as the expected value of the probability density function of observed ordinal opinions.
-In other words, we simulate $\bar{o}_t$ by calculating the weighted sum of products between the probability of observing each bin value and the bin value itself:
-$$
+To simulate the measurement of the *average* observed opinion at time $t$,
+$\bar{o}_t$, as the expected value of the probability density function of
+observed ordinal opinions. In other words, we simulate $\bar{o}_t$ by
+calculating the weighted sum of products between the probability of observing
+each bin value and the bin value itself: $$
 \bar{o}_t(\nu, e) = \sum_{i=1}^B b_i f_{it}(\nu, e).
-$${#eq-sim-ave-observed}
+$$ {#eq-sim-ave-observed}
 
 Group polarization is the difference between the post-deliberation mean observed
-opinion and the pre-deliberation one.
-We will calculate the magnitude of spurious group polarization under a mere agreement null model represented by its collected distribution parameters, $\nu = (\mu, \sigmapre, \sigmapost)$, under experimental design, $e$:
-$$
+opinion and the pre-deliberation one. We will calculate the magnitude of
+spurious group polarization under a mere agreement null model represented by its
+collected distribution parameters,
+$\nu = (\mu, \sigma_\mathrm{pre}, \sigma_\mathrm{post})$, under experimental
+design, $e$: $$
 g(\nu, e) = |\bar{o}_\mathrm{post}(\nu, e)| - |\bar{o}_\mathrm{pre}(\nu, e)|
-$${#eq-spurious-gp}
+$$ {#eq-spurious-gp}
 
-In the *Inspect* stage of our analysis we will determine which group polarization replications are invalidated due to equifinality between the *mere agreement null-polarization hypothesis* on the one hand, and the *group polarization hypothesis*
-on the other.
-This requires more than just showing there is some $\nu$ that causes spurious polarization for some experiment where $g(\nu, e) > 0$.
-We have to find a $\nu$ that generates the exact $g(\nu, e)$ reported in the original journal article *and* check that simulated $\bar{o}_\mathrm{pre}$ and $\bar{o}_\mathrm{pre}$ match published values.
+In the *Inspect* stage of our analysis we will determine which group
+polarization replications are invalidated due to equifinality between the *mere
+agreement null-polarization hypothesis* on the one hand, and the *group
+polarization hypothesis* on the other. This requires more than just showing
+there is some $\nu$ that causes spurious polarization for some experiment where
+$g(\nu, e) > 0$. We have to find a $\nu$ that generates the exact $g(\nu, e)$
+reported in the original journal article *and* check that simulated
+$\bar{o}_\mathrm{pre}$ and $\bar{o}_\mathrm{pre}$ match published values.
 
@@
 
-## Example: the "Americans" condition in Moscovici and Zavalloni (1969)
+## Example: Invalidating the "Americans" condition in Moscovici and Zavalloni (1969)
 
-One of the replications we invalidate here is from Moscovici and Zavalloni (1969),
-the "Americans" condition as they called it—we coded it `Moscovici1969_Americans`. 
-They report that participant opinions polarized after discussing the survey topic, "American economic aid is always used for political pressure." They report survey 
-responses increased in extremity from an average of -0.9 before discussion to
--1.6 after, on a seven-bin scale
-from -3 indicating "strongly disagree" to +3 indicating "strongly agree".
+One of the replications we invalidate here is from Moscovici and Zavalloni
+(1969), the "Americans" condition as they called it—we coded it
+`Moscovici1969_Americans`. They report that participant opinions polarized after
+discussing the survey topic, "American economic aid is always used for political
+pressure." They report survey responses increased in extremity from an average
+of -0.9 before discussion to -1.6 after, on a seven-bin scale from -3 indicating
+"strongly disagree" to +3 indicating "strongly agree".
 
-Moscovici and Zavalloni say their results confirm the basic hypothesis, that \emph{group discussion causes polarization}.
-But we found a mere agreement null-polarization model that generates spurious group polarization using our model—specifically, $\nu = (\mu = -1.2, \sigma_\mathrm{pre}= 4.3, and \sigma_\mathrm{post}= 1.9)$.
-This means that the hypothesis is actually undecidable given the data, invalidating Moscovici and Zavalloni's replication.
-When a hypothesis and its null explain data equally well, the data tell us nothing.
+Moscovici and Zavalloni say their results confirm the basic hypothesis, that
+\emph{group discussion causes polarization}. But we found a mere agreement
+null-polarization model that generates spurious group polarization using our
+model—specifically,
+$\nu = (\mu = -1.2, \sigma_\mathrm{pre}= 4.3, and \sigma_\mathrm{post}= 1.9)$.
+This means that the hypothesis is actually undecidable given the data,
+invalidating Moscovici and Zavalloni's replication. When a hypothesis and its
+null explain data equally well, the data tell us nothing.
 
-Now let's use the expectation model to prove that this $\nu$ generates spurious group polarization in this experiment, i.e., $g(\nu, e= \mathrm{Moscovici1969\_Americans}) > 0$. 
-In the process we demonstrate how the expectation model works to simulate the measurement of spurious group polarization (@eq-spurious-gp).
+Now let's use the expectation model to prove that this $\nu$ generates spurious
+group polarization in this experiment, i.e.,
+$g(\nu, e= \mathrm{Moscovici1969\_Americans}) > 0$. In the process we
+demonstrate how the expectation model works to simulate the measurement of
+spurious group polarization (@eq-spurious-gp).
 
-First, we calculate the probability density vector using the [`makeProbVec`](https://github.com/mt-digital/gp-statmod/blob/main/GroupPolarizationStatmod/model.R#L44) function in the file [`GroupPolarizationStatmod/model.R`](https://github.com/mt-digital/gp-statmod/blob/main/GroupPolarizationStatmod/model.R) in the [project code repository on GitHub](https://github.com/mt-digital/gp-statmod) for both experiment times: pre- and post-discussion:
-$$
+First, we calculate the probability density vector using the
+[`makeProbVec`](https://github.com/mt-digital/gp-statmod/blob/main/GroupPolarizationStatmod/model.R#L44)
+function in the file
+[`GroupPolarizationStatmod/model.R`](https://github.com/mt-digital/gp-statmod/blob/main/GroupPolarizationStatmod/model.R)
+in the [project code repository on
+GitHub](https://github.com/mt-digital/gp-statmod) for both experiment times:
+pre- and post-discussion: $$
 f_\mathrm{pre} = 
@@
 \end{pmatrix}
-$$ 
-and
-$$
+$$ and $$
 f_\mathrm{post} = \
@@
 
-When we plug in the bin values and copmuted $f_\mathrm{pre}$ values into the simulated average pre-discussion opinion definition (@eq-sim-ave-observed),
-$$
+When we plug in the bin values and copmuted $f_\mathrm{pre}$ values into the
+simulated average pre-discussion opinion definition (@eq-sim-ave-observed), $$
 \bar{o}_\mathrm{pre} = -3 \cdot 0.38~~+-2\cdot 0.09~~+\ldots+3\cdot0.19 = -0.6
-$$
-$$
+$$ $$
 \bar{o}_\mathrm{post} = -3 \cdot 0.25~~+-2\cdot 0.19~~+\ldots+3\cdot0.03 = -1.04
-$$
-So, we find that for the `Moscovici1969_Americans`, our null-polarization model appears polarized, with the exact same average pre- and post-discussion opinions reported by Moscovici and Zavalloni (1969), and the same magnitude of radicalization (CHECK THE PRESENTATION OF ALL THREE ABOVE)
-$$
+$$ So, we find that for the `Moscovici1969_Americans`, our null-polarization
+model appears polarized, with the exact same average pre- and post-discussion
+opinions reported by Moscovici and Zavalloni (1969), and the same magnitude of
+radicalization (CHECK THE PRESENTATION OF ALL THREE ABOVE) $$
 g(\nu, e) = \bar{o}_\mathrm{post} - \bar{o}_\mathrm{pre} = -1.04 - (-0.6) = - 0.4.
@@
 
-Since the simulated average opinions match the observed ones—note that as a consequence, the spurious polarization magnitude also matches the observed one.
-In other words, we *identified* a null model that explains the data. 
-If both a hypothesis and its null explain data equally well, then the hypothesis is *undecidable*
+Since the simulated average opinions match the observed ones—note that as a
+consequence, the spurious polarization magnitude also matches the observed one.
+In other words, we *identified* a null model that explains the data. If both a
+hypothesis and its null explain data equally well, then the hypothesis is
+*undecidable* under the available data. This invalidates this experiment's claim
+to be a replication of support for the group polarization hypothesis. We
+replicate this analysis for all 57 experimental conditions that purport to
+support the group polarization hypothesis. To facilitate this analysis, we
+created a web app to manage the data entry, the process of identifying invalid
+replications, and managing the outcome data of which replications were
+invalidated, and what null model parameters generated the spurious polarization.
 
-In this example we started with
+## System for identifying invalid replications
 
-## Root-finding Algorithm to Find Null-Polarization Models with Artifacts
+The next step is to scale up the analysis we did manually in the example above,
+including develop a systematic way to find, and not guess, the set of null model
+parameters that
 
-# Evaluation of Experimental Designs
+![**Computational system to interactively simulate opinion change, identify
+invalid replications, and evaluate experimental
+designs.**](Figures/Model/SoftwareMap_Mashup.png){#fig-computational-system
+fig-align="center" width="1011"}
 
-- The evaluation stage complements our *Identify* stage by filling in more details about the experimental design.
+# Part 2: *Evaluate* Experimental Designs
+
+In the previous step we identified which, if any, null polarization models
+generate spurious group polarization with the expectation model of opinion
+measurement. This is enough to invalidate a replication, but it leaves
+unanswered questions. First, we would like to know how what the chance was that
+a published detection of group polarization was actually a false positive,
+assuming that participant opinions were distributed according to the null model
+identified in the first part of the analysis. This is equivalent to the false
+positive rate of the experimental design, which we denote $\alpha_e$ for
+experiment $e$. Second, we would like to estimate the false discovery rate for
+our corpus, which is a good proxy for the field as a whole, since we selected
+top quality, highly influential work that has been cited thousands of times. The
+false discovery rate in this case tells us the probability that a published
+detection of group polarization is actually a false detection. (@eq-fdr).
+
+## Generative Model of Spurious Group Polarization
+
+To evaluate an experimental design's ability to properly identify null effects,
+we need to simulate participant responses assuming the mere agreement null model
+that invalidated the replication, if one was found. We simulate data from a
+single group polarization experiment by sampling participants' ordinal opinions
+at pre-/post-discussion from the pre- and post-discussion opinion distributions
+specified by the null model that invalidated the replication in that experiment,
+$\nu_e$. If there are $N_e$ participants in experiment $e$ then our simulated
+data sets contained $2N_e$ rows. We recorded $N_e$ in the web app as metadata
+along with data about average opinions and opinion shifts for each experimental
+condition across the corpus.
+
+It can be helpful to
 
@@
 
-First, how frequently do the $\nu$ that we found generate simulated data
-  that fool a rigorous statistical design that properly accounts for the measurement process? This serves two roles: 
-  1.
-Confirm that the $\nu$ we found could create simulated datasets that appear polarized under statistical analysis, which is different from using the expected value of an exact latent opinion distribution, integrated with the bin values to simulate the average opinion.
-  2. Evaluate the underlying experimental design independent of the failure to account for how ordinal measurements inevitably clip extreme latent opinions.
-  3. We do this by simulating experiment data using $\nu$ we identified in the previous stage. We simulate 1000 experiments indexed by $i$ for each experiment $e$.
+To calculate the false detection rate we begin by simulating pre- and
+post-deliberation opinions by drawing $N_e$ samples from normal latent opinion
+distributions that generated spurious group polarization, i.e., one sample at
+each experiment stage for each simulated participant. The latent opinion
+distributions are specified with the mere agreement null model labelled $\nu_e$.
 
-To calculate the false detection rate we begin by simulating 
-pre- and post-deliberation 
-opinions by drawing $N_e$ samples from normal latent opinion
-distributions that generated spurious group polarization, i.e., one sample at each experiment stage for each simulated participant.
-The latent opinion distributions are specified with the mere agreement null model labelled $\nu_e$.
-
-Each trial $i$ of $N_e$ simulated pre- and post-deliberation observations is then inspected to see
-if spurious group polarization occurred in the simulation. Spurious group
-polarization is determined by first inferring the latent parameters of each
-simulated experimental trial by fitting an ordered probit model~\cite{Liddell2018}.
-The procedure infers the latent distribution parameters that we in fact specified with $\nu$: $\mu_\mathrm{pre}$, $\sigma_\mathrm{pre}$, $\mu_\mathrm{post}$, and $\sigma_\mathrm{post}$.
+Each trial $i$ of $N_e$ simulated pre- and post-deliberation observations is
+then inspected to see if spurious group polarization occurred in the simulation.
+Spurious group polarization is determined by first inferring the latent
+parameters of each simulated experimental trial by fitting an ordered probit
+model [@Liddell2018]. The procedure infers the latent distribution parameters
+that we in fact specified with $\nu$: $\mu_\mathrm{pre}$, $\sigma_\mathrm{pre}$,
+$\mu_\mathrm{post}$, and $\sigma_\mathrm{post}$.
 
 The inferred parameters are then used to calculate the effect size for that
-trial, Cohen's $d$, defined as~\cite[p. 331]{Liddell2018}
-$$
-  d = \frac{\mupost - \mupre}{\sqrt{\frac{1}{2}(\sigmapost^2 + \sigmapre^2)}}.
-$$ {#eq-cohens}
-It is implied that these values are the inferred values, not the values from $\nu$, when used in the context of calculating Cohen's $d$. Note $d=0$ always for the true $\nu$.
+trial, Cohen's $d$, defined as\~\cite[p. 331]{Liddell2018} $$
+  d = \frac{\mu_\mathrm{post}- \mu_\mathrm{pre}}{\sqrt{\frac{1}{2}(\sigma_\mathrm{post}^2 + \sigma_\mathrm{pre}^2)}}.
+$$ {#eq-cohens} It is implied that these values are the inferred values, not the
+values from $\nu$, when used in the context of calculating Cohen's $d$. Note
+$d=0$ always for the true $\nu$.
 
-The family-wise error rate is the fraction of trials 
-with $d_{ei} > d^*$, written
-$$
+The family-wise error rate is the fraction of trials with $d_{ei} > d^*$,
+written $$
   \alpha_e = \frac{1}{N_T} \sum_i^{N_T} \Theta(d_{ei}; d^*),
-$$ {#eq-study-aggregate}
-where
-$$
+$$ {#eq-study-aggregate} where $$
   \Theta(d_{ei}; d^*) = 
@@
     \end{cases}
+$$ $d^*$ is the *significance threshold*, which must be specified by the
+experimenter. There are three common significance levels which we adopt here:
+$d^* = 0.2$ for "low", $d^*=0.5$ for "medium", and $d^* = 0.8$ for "high".
+
+To calculate the family-wise error rate across all experimental conditions of a
+study, $s$, $$
+\alpha_s = \frac{1}{|s|} \sum_{e \in s} \alpha_e,
+$$ where $e \in s$ indicates sum over all experimental conditions $e$ presented
+in the article associated with study $s$. The mean across all results identified
+as plausibly spurious is $$
+ \alpha_\mathrm{all} = \frac{1}{|E|} \sum_{e \in E} \alpha_e.
 $$
-$d^*$ is the *significance threshold*, which must be specified by the experimenter.
-There are three common significance levels which we adopt here: $d^* = 0.2$ for "low", $d^*=0.5$ for "medium", and $d^* = 0.8$ for "high".
 
-To calculate the family-wise error rate
-across all experimental conditions of a study, $s$, 
-\begin{equation}
-  \alpha_s = \frac{1}{|s|} \sum_{e \in s} \alpha_e,
-\end{equation}
-\noindent
-where $e \in s$ indicates sum over all experimental conditions $e$ presented in the article associated with study $s$.
-The mean across all results identified as plausibly spurious is 
-\begin{equation}
-  \alpha_\mathrm{all} = \frac{1}{|E|} \sum_{e \in E} \alpha_e.
-\end{equation}
-
-In the final part of our *Evaluation* analysis we adapt this analysis to re-align the question of statistical rigor: we can be specific now, and instead of wondering, "what significance level is truly rigorous", we can use our techniques to make the choice of $d^*$ to conform to a specific design requirement.
-Specifically we will calculate what $d^*$ must be in order to limit $\alpha \leq 0.05$ as an example.
+In the final part of our *Evaluation* analysis we adapt this analysis to
+re-align the question of statistical rigor: we can be specific now, and instead
+of wondering, "what significance level is truly rigorous", we can use our
+techniques to make the choice of $d^*$ to conform to a specific design
+requirement. Specifically we will calculate what $d^*$ must be in order to limit
+$\alpha \leq 0.05$ as an example.
 
@@
 
-We finally may calculate the false discovery rate by plugging this $\alpha$
-into Equation~\ref{eq:fdr} and assuming empirically-motivated values for 
-base rate, $b$, and statistical power, $W$, as well as more pessimistic and
-optimistic settings for base rate and power.
+The false discovery rate is the false positive rate, scaled by the rate of true
+negative effects and normalized by the frequency of true positive effects\~
+(@eq-fdr) and assuming empirically-motivated values for base rate, $b$, and
+statistical power, $W$, as well as more pessimistic and optimistic settings for
+base rate and power. Using this notation, the false detection rate is
 
-Using the notation outlined above, the false detection rate is
-\begin{equation}
-  \fdr = \frac{(1 - b)\alpha}{(1 - b)\alpha + bW}.
-  \label{eq:fdr}
-\end{equation}
+$$
+\mathrm{FDR}= \frac{(1 - b)\alpha}{(1 - b)\alpha + bW}.
+$$ {#eq-fdr}
 
-## *Specify* Significance Threshold for Low False Positive Rates
+# Part 3: *Specify* Significance Threshold for Experimental Design
 
@@
 
-There are many possible
-null models we could choose.
+There are many possible null models we could choose.
 
-what significance threshold would have limited inverting this procedure to find what significance value
-$d^*$ is necessary to achieve a low family-wise error rate $\alpha_e = 0.05$ for
-experimental condition, $e$.
+what significance threshold would have limited inverting this procedure to find
+what significance value $d^*$ is necessary to achieve a low family-wise error
+rate $\alpha_e = 0.05$ for experimental condition, $e$.
 
-To do this we first calculate $\alpha_e(d^*)$ for a
-sufficiently wide range of $d^*$. We then find the lowest $d^*$ for which
-$\alpha_e(d^*) \leq 0.05$.
+To do this we first calculate $\alpha_e(d^*)$ for a sufficiently wide range of
+$d^*$. We then find the lowest $d^*$ for which $\alpha_e(d^*) \leq 0.05$.

### 2025-11-26 GroupPolarizationStatmod/paper/Discussion.qmd
updated method, started Discussion standalone

---
title: Discussion
editor: 
  markdown: 
    wrap: 80
author: Matt Turner · [https://mat.phd](https://mat.phd)
date: today
toc: true
bibliography: this.bib
format:
  html:
    grid:
      body-width: 880px  
---

\newcommand{\mupre}{\mu_\mathrm{pre}}
\newcommand{\mupost}{\mu_\mathrm{post}}

\newcommand{\sigmapre}{\sigma_\mathrm{pre}}
\newcommand{\sigmapost}{\sigma_\mathrm{post}}

\newcommand{\thetafit}{\theta^\mathrm{fit}}

\newcommand{\fdr}{\mathrm{FDR}}

\newcommand{\meanobs}{\bar{o}}
\newcommand{\meanobsemp}{\bar{\mathbf{o}}}
\newcommand{\meanobst}{\meanobs_t}
\newcommand{\meanobsjt}{\meanobs_{j,t}}
\newcommand{\meanobspre}{\meanobs_\mathrm{pre}}
\newcommand{\meanobspost}{\meanobs_\mathrm{post}}
\newcommand{\tpre}{t_\mathrm{pre}}
\newcommand{\tpost}{t_\mathrm{post}}

# The Full Bayesian Rigor: Deriving Effect Size Posteriors and Justifying the FPR Approximation

In evaluating the False Positive Rate (FPR) of a simulated null model, the primary challenge is moving beyond the raw parameter posterior distributions ($p(\mu_1 | \text{data})$, $p(\sigma_2 | \text{data})$, etc.) to accurately quantify the posterior distribution of the derived effect size, $d$. This section outlines the statistically rigorous method for this derivation and justifies the pragmatic approximation necessary for the evaluation stage of this paper.

## The Correct Derivation of $\boldsymbol{\delta}$

In our model, the effect size $d$ is a function of four estimated parameters ($\mu_1, \mu_2, \sigma_1, \sigma_2$), defined by the authors whose work we replicate as:

$$d = \frac{\mu_1 - \mu_2}{\sqrt{(\sigma_1^2 + \sigma_2^2)/2}}$$

The statistically rigorous method for obtaining the posterior distribution of $d$, denoted $p(d | \text{data})$, is through **marginalization by sampling** across the joint posterior distribution of the primary parameters $p(\mu_1, \mu_2, \sigma_1, \sigma_2 | \text{data})$.

### Necessity of MCMC Samples

The parameters are fundamentally **sequentially dependent** in the MCMC chain, as the estimation of means is conditional on the estimation of variances. This dependency is captured in the joint posterior distribution.

To maintain statistical rigor, we must avoid calculating $d$ by combining the isolated summary statistics (e.g., the posterior median of $\mu_1$ and the posterior median of $\sigma_1$) because this procedure, known as **plug-in estimation**, ignores the critical correlation between parameters, resulting in an unquantified and likely underestimated uncertainty.

The correct procedure requires using the full MCMC samples:

1.  For each usable iteration $k$ of the Markov Chain (where $K$ is the total number of samples), a plausible set of values is drawn: $(\mu_1^{(k)}, \mu_2^{(k)}, \sigma_1^{(k)}, \sigma_2^{(k)})$.
2.  The effect size is calculated for that specific set of values:
    $$d^{(k)} = \frac{\mu_1^{(k)} - \mu_2^{(k)}}{\sqrt{\frac{(\sigma_1^{(k)})^2 + (\sigma_2^{(k)})^2}{2}}}$$
3.  The set $\{d^{(1)}, d^{(2)}, \ldots, d^{(K)}\}$ then forms the complete posterior distribution $p(d | \text{data})$.

The center of this distribution, the **posterior median ($\tilde{d}$)**, is the most robust point estimate for the effect size, as it minimizes expected absolute error and is less sensitive to potential skewness than the posterior mean.

## Justifying the FPR Approximation

Due to the computational complexity and the unavailability of the original MCMC chains for all $57,000$ fits, a pragmatic simplification was necessary for the subsequent evaluation of the False Positive Rate (FPR).

### The FPR Criterion

A standard method for estimating the FPR (or Type I error rate) in simulation studies is to count the proportion of null trials where the observed effect size meets or exceeds a predefined threshold, $d^*$.

$$\text{FPR} = \frac{\text{Number of Null Fits where } |d| > d^*_{\text{threshold}}}{\text{Total Number of Null Simulations}}$$

In this analysis, $d^*_{\text{threshold}}$ is set to the commonly adopted small effect size threshold of $d^*=0.2$ (and varied for sensitivity analysis), avoiding the reliance on approximated $95\%$ credible intervals, which would require further, less reliable approximations.

### The Compromise: Plug-in and Heuristics

We employed a simplified plug-in value for $d$ using the stored posterior means:

$$d_{\text{plug}} = \frac{\hat{\mu}_1 - \hat{\mu}_2}{\sqrt{(\hat{\sigma}_1^2 + \hat{\sigma}_2^2)/2}}$$

Where $\hat{\theta}$ represents the posterior mean of the parameter $\theta$.

While this approach is a statistical shortcut, the error is mitigated by two factors: the use of the $d^*_{\text{threshold}}$ heuristic, and the large simulated sample size ($N=500$ per group), which ensures that the posteriors for $\mu$ and $\sigma$ are sufficiently narrow. This minimizes the difference between $d_{\text{plug}}$ and the true posterior median $\tilde{d}$, making the shortcut defensible as a pragmatic necessity given the study's scale.

### 2025-11-26 GroupPolarizationStatmod/paper/method.qmd
updated method, started Discussion standalone

@@
 \end{cases}
-$$
+$$ {#eq-bin-thresholds}
 
@@
 When we plug in the bin values and copmuted $f_\mathrm{pre}$ values into the
-simulated average pre-discussion opinion definition (@eq-sim-ave-observed), $$
+simulated average pre-discussion opinion definition (@eq-sim-ave-observed), we
+get $$
 \bar{o}_\mathrm{pre} = -3 \cdot 0.38~~+-2\cdot 0.09~~+\ldots+3\cdot0.19 = -0.6
-$$ $$
-\bar{o}_\mathrm{post} = -3 \cdot 0.25~~+-2\cdot 0.19~~+\ldots+3\cdot0.03 = -1.04
+$$ and $$
+\bar{o}_\mathrm{post} = -3 \cdot 0.25~~+-2\cdot 0.19~~+\ldots+3\cdot0.03 = -1.04.
 $$ So, we find that for the `Moscovici1969_Americans`, our null-polarization
@@
 opinions reported by Moscovici and Zavalloni (1969), and the same magnitude of
-radicalization (CHECK THE PRESENTATION OF ALL THREE ABOVE) $$
+radicalization: $$
 g(\nu, e) = \bar{o}_\mathrm{post} - \bar{o}_\mathrm{pre} = -1.04 - (-0.6) = - 0.4.
@@
 we need to simulate participant responses assuming the mere agreement null model
-that invalidated the replication, if one was found. We simulate data from a
-single group polarization experiment by sampling participants' ordinal opinions
-at pre-/post-discussion from the pre- and post-discussion opinion distributions
-specified by the null model that invalidated the replication in that experiment,
-$\nu_e$. If there are $N_e$ participants in experiment $e$ then our simulated
-data sets contained $2N_e$ rows. We recorded $N_e$ in the web app as metadata
-along with data about average opinions and opinion shifts for each experimental
-condition across the corpus.
+that invalidated the replication ($\nu_e$). We use the stochastic model for this
+simulation stage, as it correctly models the noise inherent in individual
+participant measurement and response. The ordinal, measured opinions are
+generated by executing the following steps for $N_e$ simulated participants
+across $N_T$ independent trials:
 
-It can be helpful to
+1.  **Latent Opinion Sampling:** For each participant $i$ and time
+    $t \in \{\text{pre, post}\}$, we draw a continuous latent opinion
+    $\omega_{i,t}$ from the corresponding normal distribution defined by the
+    null model $\nu_e$: $$\omega_{i,t} \sim \mathcal{N}(\mu, \sigma_t)$$ where
+    $\mu$ is the constant latent mean and $\sigma_t$ is the time-specific
+    standard deviation derived from the mere agreement null model $\nu_e$.
+2.  **Measurement Simulation (Ordinal Binning):** Each continuous latent opinion
+    $\omega_{i,t}$ is mapped onto the experiment's ordinal scale (e.g., $-3$ to
+    $+3$). This mapping is determined by the fixed measurement thresholds,
+    $\theta_j$, established in the *Expectation Model* subsection
+    (@eq-bin-thresholds). This process simulates the clipping artifact that
+    generates spurious polarization in the observable data.
+
+This results in a simulated dataset of $2N_e$ rows (two time steps per
+participant) for a single trial, consisting of ordinal opinion values that are
+then passed to the statistical inference pipeline.
 
@@
 
-In the final analysis step we calculate what significance threshold, $d^*$,
-would limit the false positive rate to 0.05 or less.
+In the final stage of our analysis, we pivot from evaluating the outcome of
+published designs to *specifying the statistical requirements for future
+designs*. The primary goal of this specification is to invert the standard
+question of statistical rigor: instead of asking if an effect is 'significant'
+given a fixed effect-size threshold (e.g., $d^{*} = 0.5$), we ask what
+significance threshold is necessary to limit the false positive rate to a
+desired maximum, such as $5\%$ (or $\alpha = 0.05$).
 
-There are many possible null models we could choose.
+We leverage the evaluation framework developed in Part 2, based on the mere
+agreement null model $\nu_e$ identified for each experimental condition $e$. Our
+objective is to calculate the function $\alpha_e(d^*)$ across a sufficient range
+of $d^*$ values. This allows us to determine the lowest $d^*$ that satisfies the
+design requirement $\alpha_e(d^*) \leq 0.05$.
 
-what significance threshold would have limited inverting this procedure to find
-what significance value $d^*$ is necessary to achieve a low family-wise error
-rate $\alpha_e = 0.05$ for experimental condition, $e$.
+The function $\alpha_e(d^*)$ is the simulated false positive rate as a function
+of the significance threshold, $d^*$, defined as: $$
+\alpha_e(d^*) = \frac{1}{N_T} \sum_i^{N_T} \Theta(d_{ei}; d^*),
+$$ where $N_T$ is the number of null trials, $d_{ei}$ is the calculated effect
+size for trial $i$, and $\Theta$ is the indicator function that flags trials
+where the effect size $d_{ei}$ exceeds the significance threshold $d^*$. In the
+interest of giving the most generous best-case scenario possible, we only count
+positive effect sizes because only these indicate increased extremism over time.
+Negative effect sizes indicate group opinions became more moderate—such
+observations do not support the group polarization hypothesis that opinions
+become more extreme following discussion.
 
-To do this we first calculate $\alpha_e(d^*)$ for a sufficiently wide range of
-$d^*$. We then find the lowest $d^*$ for which $\alpha_e(d^*) \leq 0.05$.
+Because our system is automated, and the code publicly available, it can be
+adapted and used to design future group polarization experiments with
+empirically-grounded statistical rigor. Experimenters can specify a desired
+maximum false positive rate, such as $\alpha = 0.05$, and our system will
+calculate the minimum effect size threshold $d^*$ that must be exceeded to
+achieve that level of rigor under the mere agreement null model identified for
+their experimental design. This prescriptive step transforms the simulation from
+a post-hoc critique into a forward-looking design tool, giving experimenters an
+empirically-derived minimum effect size they must exceed to maintain statistical
+confidence in their results.
+
+# Analysis Outline and Summary

### 2025-11-26 GroupPolarizationStatmod/paper/method.qmd
old updates to method

--- 2025-11-25 GroupPolarizationStatmod/paper/method.qmd
@@
 $\nu_e$. If there are $N_e$ participants in experiment $e$ then our simulated
-data sets contained $2N_e$ rows. We recorded $N_e$ in the web app as metadata
-along with data about average opinions and opinion shifts for each experimental
-condition across the corpus.
+data sets contained $2N_e$ rows: two time step rows per participant. We recorded
+$N_e$ used by each experiment in the web app as metadata along with data about
+average opinions and opinion shifts for each experimental condition across the
+corpus.
 
-It can be helpful to
+Group polarization is "detected" if the group polarization effect size meets or
+exceeds a threshold value. We use Cohen's $d$ for the effect size, defined as
+[@Liddell2018, 331] $$
+  d = \frac{\mu_\mathrm{post}- \mu_\mathrm{pre}}{\sqrt{\frac{1}{2}(\sigma_\mathrm{post}^2 + \sigma_\mathrm{pre}^2)}}.
+$$ It is implied that these values are the inferred values, not the values from
+$\nu$, when used in the context of calculating Cohen's $d$. Note $d=0$ always
+for the true $\nu$.
 
@@
 distributions are specified with the mere agreement null model labelled $\nu_e$.
-
-Each trial $i$ of $N_e$ simulated pre- and post-deliberation observations is
-then inspected to see if spurious group polarization occurred in the simulation.
-Spurious group polarization is determined by first inferring the latent
-parameters of each simulated experimental trial by fitting an ordered probit
-model [@Liddell2018]. The procedure infers the latent distribution parameters
-that we in fact specified with $\nu$: $\mu_\mathrm{pre}$, $\sigma_\mathrm{pre}$,
-$\mu_\mathrm{post}$, and $\sigma_\mathrm{post}$.
-
-The inferred parameters are then used to calculate the effect size for that
-trial, Cohen's $d$, defined as\~\cite[p. 331]{Liddell2018} $$
-  d = \frac{\mu_\mathrm{post}- \mu_\mathrm{pre}}{\sqrt{\frac{1}{2}(\sigma_\mathrm{post}^2 + \sigma_\mathrm{pre}^2)}}.
-$$ {#eq-cohens} It is implied that these values are the inferred values, not the
-values from $\nu$, when used in the context of calculating Cohen's $d$. Note
-$d=0$ always for the true $\nu$.
 

### 2025-11-26 GroupPolarizationStatmod/paper/method.qmd
expanded method with error rates table

--- 2025-11-26 GroupPolarizationStatmod/paper/method.qmd
@@
 
+The core of our evaluation is the quantification of the probability that a spurious result—an effect generated by the null model—is falsely interpreted as genuine group polarization.
+We calculate this probability across three distinct units of analysis, building from the specific to the general:
+
+1.
+The Type I Error Rate ($\alpha_e$): The rate of falsely detecting group polarization in a single experimental condition ($e$).
+This directly quantifies the risk posed by the mere agreement model identified for that specific experiment.
+2.
+The study-level Family-Wise Error Rate ($\alpha_s$): The mean Type I Error Rate aggregated across all conditions within a single study ($s$).
+This estimates the false claim rate across the set of tests performed in that study.
+3.
+The Family-Wise Error Rate ($\alpha_{\text{all}}$): The mean Type I Error Rate aggregated across all 57 experimental conditions in the corpus ($E$).
+This estimates the expected false claim rate for a researcher adopting the general experimental design found throughout the literature.
+
+These calculated Type I Error Rates ($\alpha$) are then combined with fixed external parameters to estimate the False Discovery Rate.
+Following convention, we adopt a fixed Statistical Power of $W=0.80$ (and test sensitivity with alternative values) and a Base Rate of $b=0.10$.
+The relationship between the units of analysis and the estimated error rates is summarized in Table @tbl-error-rates.
+
+{{< include Error-Rates-Table.html >}}
+
 ## Generative Model of Spurious Group Polarization
@@
 then passed to the statistical inference pipeline.
+
+### Statistical Inference Pipeline
+
+To complete our our simulated experimental design we need to be able to
+statistically infer the effect sizes that measure the strength of group
+polarization in an experiment. To keep the presentation simple and transparent,
+we infer effect sizes by inferring the parameters $\mu_t$ and $\sigma_t$ of the
+hypothesized normal distribution that generates the data—*literally* generates
+the data in our simulations, but is just a model of the distribution of group
+opinions in real-world experiments.
 

### 2025-12-02 GroupPolarizationStatmod/paper/method.qmd

--- 2025-11-26 GroupPolarizationStatmod/paper/method.qmd
@@
 
-Each trial $i$ of $N_e$ simulated pre- and post-deliberation observations is
-then inspected to see if spurious group polarization occurred in the simulation.
-Spurious group polarization is determined by first inferring the latent
-parameters of each simulated experimental trial by fitting an ordered probit
-model [@Liddell2018]. The procedure infers the latent distribution parameters
-that we in fact specified with $\nu$: $\mu_\mathrm{pre}$, $\sigma_\mathrm{pre}$,
-$\mu_\mathrm{post}$, and $\sigma_\mathrm{post}$.
-
-The inferred parameters are then used to calculate the effect size for that
-trial, Cohen's $d$, defined as\~\cite[p. 331]{Liddell2018} $$
-  d = \frac{\mu_\mathrm{post}- \mu_\mathrm{pre}}{\sqrt{\frac{1}{2}(\sigma_\mathrm{post}^2 + \sigma_\mathrm{pre}^2)}}.
-$$ {#eq-cohens} It is implied that these values are the inferred values, not the
-values from $\nu$, when used in the context of calculating Cohen's $d$. Note
-$d=0$ always for the true $\nu$.
-
 The family-wise error rate is the fraction of trials with $d_{ei} > d^*$,

### 2025-12-02 GroupPolarizationStatmod/paper/Discussion.qmd

@@
 \newcommand{\mupost}{\mu_\mathrm{post}}
-
 \newcommand{\sigmapre}{\sigma_\mathrm{pre}}
@@
 
-In evaluating the False Positive Rate (FPR) of a simulated null model, the primary challenge is moving beyond the raw parameter posterior distributions ($p(\mu_1 | \text{data})$, $p(\sigma_2 | \text{data})$, etc.) to accurately quantify the posterior distribution of the derived effect size, $d$.
-This section outlines the statistically rigorous method for this derivation and justifies the pragmatic approximation necessary for the evaluation stage of this paper.
+In evaluating the False Positive Rate (FPR) of a simulated null model, the
+primary challenge is moving beyond the raw parameter posterior distributions
+($p(\mu_1 | \text{data})$, $p(\sigma_2 | \text{data})$, etc.) to accurately
+quantify the posterior distribution of the derived effect size, $d$. This
+section outlines the statistically rigorous method for this derivation and
+justifies the pragmatic approximation necessary for the evaluation stage of this
+paper.
 
@@
 
-In our model, the effect size $d$ is a function of four estimated parameters ($\mu_1, \mu_2, \sigma_1, \sigma_2$), defined by the authors whose work we replicate as:
+In our model, the effect size $d$ is a function of four estimated parameters
+($\mu_1, \mu_2, \sigma_1, \sigma_2$), defined by the authors whose work we
+replicate as:
 
@@
 
-The statistically rigorous method for obtaining the posterior distribution of $d$, denoted $p(d | \text{data})$, is through **marginalization by sampling** across the joint posterior distribution of the primary parameters $p(\mu_1, \mu_2, \sigma_1, \sigma_2 | \text{data})$.
+The statistically rigorous method for obtaining the posterior distribution of
+$d$, denoted $p(d | \text{data})$, is through **marginalization by sampling**
+across the joint posterior distribution of the primary parameters
+$p(\mu_1, \mu_2, \sigma_1, \sigma_2 | \text{data})$.
 
@@
 
-The parameters are fundamentally **sequentially dependent** in the MCMC chain, as the estimation of means is conditional on the estimation of variances.
-This dependency is captured in the joint posterior distribution.
+The parameters are fundamentally **sequentially dependent** in the MCMC chain,
+as the estimation of means is conditional on the estimation of variances. This
+dependency is captured in the joint posterior distribution.
 
-To maintain statistical rigor, we must avoid calculating $d$ by combining the isolated summary statistics (e.g., the posterior median of $\mu_1$ and the posterior median of $\sigma_1$) because this procedure, known as **plug-in estimation**, ignores the critical correlation between parameters, resulting in an unquantified and likely underestimated uncertainty.
+To maintain statistical rigor, we must avoid calculating $d$ by combining the
+isolated summary statistics (e.g., the posterior median of $\mu_1$ and the
+posterior median of $\sigma_1$) because this procedure, known as **plug-in
+estimation**, ignores the critical correlation between parameters, resulting in
+an unquantified and likely underestimated uncertainty.
 
@@
 
-1.  For each usable iteration $k$ of the Markov Chain (where $K$ is the total number of samples), a plausible set of values is drawn: $(\mu_1^{(k)}, \mu_2^{(k)}, \sigma_1^{(k)}, \sigma_2^{(k)})$.
+1.  For each usable iteration $k$ of the Markov Chain (where $K$ is the total
+    number of samples), a plausible set of values is drawn:
+    $(\mu_1^{(k)}, \mu_2^{(k)}, \sigma_1^{(k)}, \sigma_2^{(k)})$.
 2.  The effect size is calculated for that specific set of values:
     $$d^{(k)} = \frac{\mu_1^{(k)} - \mu_2^{(k)}}{\sqrt{\frac{(\sigma_1^{(k)})^2 + (\sigma_2^{(k)})^2}{2}}}$$
-3.  The set $\{d^{(1)}, d^{(2)}, \ldots, d^{(K)}\}$ then forms the complete posterior distribution $p(d | \text{data})$.
+3.  The set $\{d^{(1)}, d^{(2)}, \ldots, d^{(K)}\}$ then forms the complete
+    posterior distribution $p(d | \text{data})$.
 
-The center of this distribution, the **posterior median ($\tilde{d}$)**, is the most robust point estimate for the effect size, as it minimizes expected absolute error and is less sensitive to potential skewness than the posterior mean.
+The center of this distribution, the **posterior median (**$\tilde{d}$), is the
+most robust point estimate for the effect size, as it minimizes expected
+absolute error and is less sensitive to potential skewness than the posterior
+mean.
 
@@
 
-Due to the computational complexity and the unavailability of the original MCMC chains for all $57,000$ fits, a pragmatic simplification was necessary for the subsequent evaluation of the False Positive Rate (FPR).
+Due to the computational complexity and the unavailability of the original MCMC
+chains for all $57,000$ fits, a pragmatic simplification was necessary for the
+subsequent evaluation of the False Positive Rate (FPR).
 
@@
 
-A standard method for estimating the FPR (or Type I error rate) in simulation studies is to count the proportion of null trials where the observed effect size meets or exceeds a predefined threshold, $d^*$.
+A standard method for estimating the FPR (or Type I error rate) in simulation
+studies is to count the proportion of null trials where the observed effect size
+meets or exceeds a predefined threshold, $d^*$.
 
@@
 
-In this analysis, $d^*_{\text{threshold}}$ is set to the commonly adopted small effect size threshold of $d^*=0.2$ (and varied for sensitivity analysis), avoiding the reliance on approximated $95\%$ credible intervals, which would require further, less reliable approximations.
+In this analysis, $d^*_{\text{threshold}}$ is set to the commonly adopted small
+effect size threshold of $d^*=0.2$ (and varied for sensitivity analysis),
+avoiding the reliance on approximated $95\%$ credible intervals, which would
+require further, less reliable approximations.
 
@@
 
-While this approach is a statistical shortcut, the error is mitigated by two factors: the use of the $d^*_{\text{threshold}}$ heuristic, and the large simulated sample size ($N=500$ per group), which ensures that the posteriors for $\mu$ and $\sigma$ are sufficiently narrow.
-This minimizes the difference between $d_{\text{plug}}$ and the true posterior median $\tilde{d}$, making the shortcut defensible as a pragmatic necessity given the study's scale.
+While this approach is a statistical shortcut, the error is mitigated by two
+factors: the use of the $d^*_{\text{threshold}}$ heuristic, and the large
+simulated sample size ($N=500$ per group), which ensures that the posteriors for
+$\mu$ and $\sigma$ are sufficiently narrow. This minimizes the difference
+between $d_{\text{plug}}$ and the true posterior median $\tilde{d}$, making the
+shortcut defensible as a pragmatic necessity given the study's scale.

### 2025-12-04 GroupPolarizationStatmod/paper/method.qmd
cleaned up and simplified for faster build in texshop

@@
     grid:
-      body-width: 880px  
+      body-width: 725px  
 ---
@@
 The next step is to scale up the analysis we did manually in the example above,
-including develop a systematic way to find, and not guess, the set of null model
-parameters that
+including develop a systematic way to find, and not guess, the set of mere
+agreement null model parameters that generate spurious group polarization due to
+measurement artifacts.
 

### 2025-12-04 GroupPolarizationStatmod/paper/ms.tex
updates
drafting method cont'd
cleaned up and simplified for faster build in texshop

@@
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
-\PassOptionsToPackage{hyphens}{url}
-\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
-%
+% ==============================================================================
+% CLEANED ms.tex (Optimized for XeLaTeX/Fira Fonts)
+% - Consolidated preamble and ms.tex.
+% - Removed all conflicting packages (lmodern, redundant font settings).
+% - Uses Fira Sans/Mono/Math consistently.
+% ==============================================================================
+
 \documentclass[preview]{article}
 
-\usepackage{floatrow}
-\DeclareFloatFont{small}{\small}% "scriptsize" is defined by floatrow, "tiny" not
+% --- CORE FONT SETUP (Sans-Serif Headings, Serif Body, Professional Math) ---
+\usepackage{fontspec}        
+\usepackage{unicode-math}    
+%
+%% 1. Main Text: Set to Latin Modern Roman (Stable, Professional Serif)
+%\setmainfont{Latin Modern Roman} 
+%
+% 2. Monospaced/Code Text: Fira Mono
+%\setmonofont{Fira Mono} 
+%
+%% 3. Mathematical Notation: Latin Modern Math (Stable, Conventional Serif Math)
+%\setmathfont{Latin Modern Math}
+
+% 4. Heading Font Setup (Fira Sans)
+\newfontfamily{\sansheadings}{Fira Sans} 
+\usepackage{sectsty}
+\sectionfont{\sansheadings \raggedright \Large}
+\subsectionfont{\sansheadings \raggedright \large}
+\subsubsectionfont{\sansheadings \raggedright \normalsize}
+% Ensure your title command uses \sansheadings as defined in the previous step.
+% ...
+
+% --- ESSENTIAL PACKAGES (Cleaned and Prioritized) ---
+
+% 2. Core Text Utilities and Geometry
+\usepackage[margin=1.25in,letterpaper]{geometry} % Defines page layout
+\usepackage[]{microtype}    % High-quality font protrusion and spacing
+\usepackage{upquote}        % Straight quotes in verbatim/code
+\usepackage{calc}           % Utility for calculations
+
+% 3. Graphics, Color, and Table Fundamentals
+\usepackage[table,dvipsnames,svgnames,x11names]{xcolor} % Color definitions
+\usepackage{graphicx}       % For including graphics
+\usepackage{booktabs}       % Professional quality tables
+\usepackage{longtable}      % Multi-page tables
+\usepackage{array}          % Enhanced array and column definitions
+\usepackage{multirow}       % Merging rows in tables
+\usepackage{makecell}       % Formatting cells in tables
+
+% 4. Caption Handling (Load basic caption definitions BEFORE floatrow)
+\usepackage{caption}        
+\usepackage{subcaption}     
+\usepackage{floatrow}       % Complex float/caption placement and formatting
+
+% 5. Math Environments and Algorithm Structure
+\usepackage{amsmath}        % Essential math environments (align, gather, etc.)
+\usepackage{algorithm}
+\usepackage{algpseudocode}
+
+% 6. Hyperlinks and References
+\usepackage{hyperref}       % Hyperlinks and PDF metadata (must be loaded relatively late)
+\usepackage{bookmark}       % PDF bookmark control
+\usepackage{xurl}           % Better URL line breaks
+\usepackage{footnotehyper}  % Footnotes integration with hyperref
+
+% 7. Custom Utilities (Needed for your document structure)
+\usepackage{iftex}          
+\usepackage{authblk}        % Author affiliations
+\usepackage{etoolbox}       % Utility commands for patching/macros
+\usepackage{wrapfig}        % Text wrapping around figures
+
+% --- End of ESSENTIAL PACKAGES ---
+
+% --- BIBLIOGRAPHY CONFIGURATION ---
+% Use the biblatex settings you had in preamble.tex
+\usepackage[sorting=none, style=numeric-comp]{biblatex}
+\addbibresource{/Users/mt/workspace/Writing/library.bib}
+%\addbibresource{this.bib}
+
+% --- FLOATROW / TABLE SETUP (From ms.tex) ---
+\DeclareFloatFont{small}{\small}
 \floatsetup[table]{font=small}
 
-\usepackage[table]{xcolor}
-\usepackage{multirow}
-\usepackage{makecell}
-\usepackage{array}
-
-\usepackage{amsmath,amssymb}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-\usepackage{iftex}
-\input{preamble.tex}
+% --- HYPERREF SETUP (From ms.tex) ---
 \hypersetup{
@@
 
+% --- CUSTOM COMMANDS (From ms.tex) ---
 \newcommand{\mupre}{\mu_\mathrm{pre}}
 \newcommand{\mupost}{\mu_\mathrm{post}}
-
 \newcommand{\sigmapre}{\sigma_\mathrm{pre}}
 \newcommand{\sigmapost}{\sigma_\mathrm{post}}
-
 \newcommand{\thetafit}{\theta^\mathrm{fit}}
-
 \newcommand{\fdr}{\mathrm{FDR}}
-
 \newcommand{\meanobs}{\bar{o}}
@@
 \newcommand{\meanobsjt}{\meanobs_{j,t}}
-
 \newcommand{\meanobspre}{\meanobs_\mathrm{pre}}
 \newcommand{\meanobspost}{\meanobs_\mathrm{post}}
-
+\newcommand{\tpre}{t_\mathrm{pre}}
+\newcommand{\tpost}{t_\mathrm{yopost}}
 \definecolor{myorange}{RGB}{240, 96, 0}
@@
 
-\usepackage{authblk}
-\usepackage{etoolbox}
-\usepackage{wrapfig}
+% --- TITLE AND AUTHOR INFO (From ms.tex) ---
 \makeatletter
-% \title{If the Null Fits, You Must Omit: Ubiquitous False Detections of Group Polarization}
-\title{If the Null Fits, You Must Omit: Measurement Artifacts Undermine Group Polarization Replications}
+\title{\Large If the Null Fits, You Must Omit: Measurement Artifacts Undermine Group Polarization Replications}
 \makeatother
@@
 \affil[1]{\small Environmental Social Sciences, Stanford Doerr School of Sustainability, Stanford University}
-
 \author[2,3]{{Paul E.~Smaldino}}
@@
 \affil[3]{\small Santa Fe Institute} 
-% \vspace{2em}
 \affil[*]{\small Correspondence: \href{mailto:maturner@stanford.edu}{maturner@stanford.edu}}
+\date{\today}
 
-\date{\today}
+% ==============================================================================
+% BEGIN DOCUMENT
+% ==============================================================================
 \begin{document}
@@
 
-\renewcommand*\contentsname{Table of contents}
+\renewcommand*\contentsname{\small Table of contents}
 {
 \hypersetup{linkcolor=gray}
-\setcounter{tocdepth}{2}
+\setcounter{tocdepth}{3}
+\footnotesize
 \tableofcontents
@@
 
+% Center rule to separate TOC from rest of paper
+\vspace{-2em}
 \begin{center} \noindent\rule{4cm}{0.4pt} \end{center}
-% \clearpage
+\vspace{-2em}
 
-% \begin{quote}
-% In our introductory social psychology course, 
-% we have for many years used the [group polarization experimental paradigm] as
-% a laboratory exercise. The exercise works beautifully, but one must be
-% careful to forewarn a class that [group polarization] does not occur with every 
-% group\ldots and that the effect is not large. 
-% \par\raggedleft(Brown, 1986, p.\cite[p. 205]{Brown1986})
-% \end{quote}
+% --- INPUT CHAPTERS ---
+\input{intro.tex}
+\input{method.tex}
+\input{analysis.tex}
+\input{discussion.tex}
+\input{tables.tex}
+\input{figures.tex}
 
-% \begin{quote}
-%   The good thing about science is that it's true, whether or not you believe
-%   it. (Neil de Grasse Tyson slogan circa 2018)
-% \end{quote}
-
-\input{intro-latest.tex}
-
-% \input{old-intro.tex}
-
-\input{method.tex}
-
-\input{analysis.tex}
-
-\input{discussion.tex}
+\clearpage
 
@@
 \renewcommand{\thefigure}{A\arabic{figure}}
-
 \setcounter{figure}{0}
-
 \clearpage
@@
 \renewcommand{\thepage}{S\arabic{page}}
+\makeatother
 
@@
   \caption{Boxplot of Cohen's $d$ for the ordered probit model over 1000
-    simulation trias for each experimental condition (y-axis in A and B). Each
-    trial represents a possible outcome of a group polarization experiment where
+    simulation trials for each experimental condition (y-axis in A and B).
+    Each trial represents a possible outcome of a group polarization experiment where
     the true opinion shift is zero. The mean across trials is closer to 0 than

### 2025-12-05 GroupPolarizationStatmod/paper/discussion.tex
starting penultimate revision

@@
 
+\subsection{The Full Bayesian Rigor: Deriving Effect Size Posteriors and Justifying the FPR Approximation}
+
+In evaluating the \textbf{False Positive Rate (FPR)} of a simulated null model, the primary challenge is moving beyond the raw parameter posterior distributions ($p(\mu_1 | \text{data})$, $p(\sigma_2 | \text{data})$, etc.) to accurately quantify the posterior distribution of the derived \textbf{effect size}, $d$.
+This section outlines the statistically rigorous method for this derivation and justifies the pragmatic approximation necessary for the evaluation stage of this paper.
+
+\subsubsection{The most rigorous formulation of $d$}
+
+In our model, the effect size $d$ is a function of four estimated parameters ($\mu_1, \mu_2, \sigma_1, \sigma_2$), defined by the authors whose work we replicate as:
+
+\begin{equation}
+d = \frac{\mu_1 - \mu_2}{\sqrt{(\sigma_1^2 + \sigma_2^2)/2}}
+\end{equation}
+
+The statistically rigorous method for obtaining the posterior distribution of $d$, denoted $p(d | \text{data})$, is through \textbf{marginalization by sampling} across the joint posterior distribution of the primary parameters $p(\mu_1, \mu_2, \sigma_1, \sigma_2 | \text{data})$.
+
+\paragraph{Necessity of MCMC Samples}
+
+The parameters are fundamentally \textbf{sequentially dependent} in the MCMC chain, as the estimation of means is conditional on the estimation of variances.
+This dependency is captured in the \textbf{joint posterior distribution}.
+
+To maintain statistical rigor, we must avoid calculating $d$ by combining the isolated summary statistics (e.g., the posterior median of $\mu_1$ and the posterior median of $\sigma_1$) because this procedure, known as \textbf{plug-in estimation}, ignores the critical correlation between parameters, resulting in an unquantified and likely underestimated uncertainty.
+
+The correct procedure requires using the full MCMC samples:
+
+\begin{enumerate}
+    \item For each usable iteration $k$ of the Markov Chain (where $K$ is the total number of samples), a plausible set of values is drawn: $(\mu_1^{(k)}, \mu_2^{(k)}, \sigma_1^{(k)}, \sigma_2^{(k)})$.
+    \item The effect size is calculated for that specific set of values:
+    $$d^{(k)} = \frac{\mu_1^{(k)} - \mu_2^{(k)}}{\sqrt{\frac{(\sigma_1^{(k)})^2 + (\sigma_2^{(k)})^2}{2}}}$$
+    \item The set $\{d^{(1)}, d^{(2)}, \ldots, d^{(K)}\}$ then forms the complete posterior distribution $p(d | \text{data})$.
+\end{enumerate}
+
+The center of this distribution, the \textbf{posterior median} ($\tilde{d}$), is the most robust point estimate for the effect size, as it minimizes expected absolute error and is less sensitive to potential skewness than the posterior mean.
+
+\subsubsection{Justifying the FPR Approximation}
+
+Due to the computational complexity and the unavailability of the original MCMC chains for all $57,000$ fits, a pragmatic simplification was necessary for the subsequent evaluation of the False Positive Rate (FPR).
+
+\paragraph{The FPR Criterion}
+
+A standard method for estimating the FPR (or Type I error rate) in simulation studies is to count the proportion of null trials where the observed effect size meets or exceeds a predefined threshold, $d^*$.
+
+$$\text{FPR} = \frac{\text{Number of Null Fits where } |d| > d^*_{\text{threshold}}}{\text{Total Number of Null Simulations}}$$
+
+In this analysis, $d^*_{\text{threshold}}$ is set to the commonly adopted small effect size threshold of $d^*=0.2$ (and varied for sensitivity analysis), avoiding the reliance on approximated $95\%$ credible intervals, which would require further, less reliable approximations.
+
+\paragraph{The Compromise: Plug-in and Heuristics}
+
+We employed a simplified plug-in value for $d$ using the stored posterior means:
+
+$$d_{\text{plug}} = \frac{\hat{\mu}_1 - \hat{\mu}_2}{\sqrt{(\hat{\sigma}_1^2 + \hat{\sigma}_2^2)/2}}$$
+
+Where $\hat{\theta}$ represents the \textbf{posterior mean} of the parameter $\theta$.
+
+While this approach is a statistical shortcut, the error is mitigated by two factors: the use of the $d^*_{\text{threshold}}$ heuristic, and the large simulated sample size ($N=500$ per group), which ensures that the posteriors for $\mu$ and $\sigma$ are sufficiently narrow.
+This minimizes the difference between $d_{\text{plug}}$ and the true posterior median $\tilde{d}$, making the shortcut defensible as a pragmatic necessity given the study's scale.
+
 \subsection{Humble Hearts Promote Rigorous Science}

### 2025-12-05 GroupPolarizationStatmod/paper/method.tex
starting penultimate revision

--- 2025-11-18 GroupPolarizationStatmod/paper/method.tex
@@
-\section{Method}\label{method}
+\section{Method}
 
-A group polarization replication is proven undecidable, and therefore not
-truly a replication, if we can find a model of \emph{mere agreement} that
-appears to be polarized under that replication's experimental design. If we
-cannot find such a model, we cannot determine a replication to be
-undecidable. Mere agreement in this model is a dynamic where agreement
-increases within the group, but extremism does not—in other words, A
-\emph{null-polarization distributions} is a set of two Gaussian
-distributions: one representing the pre-discussion opinions of the group, and
-the other representing the post-discussion opinions. Our
-\emph{null-polarization distributions} represent the process of the group
-finding \emph{mere agreement}, which is when the average group opinion is
-constant over time, but opinion variance decreases to represent agreement
-(or, equivalently, \emph{consensus}~\cite{DeGroot1974}).
+A group polarization replication is proven undecidable, and therefore
+not truly a replication, if we can find a model of \emph{mere agreement}
+that appears to be polarized under that replication's experimental
+design. If we cannot find such a model, we cannot determine a
+replication to be undecidable. Mere agreement in this model is a dynamic
+where agreement increases within the group, but extremism does not---in
+other words, A \emph{null-polarization distributions} is a set of two
+Gaussian distributions: one representing the pre-discussion opinions of
+the group, and the other representing the post-discussion opinions. Our
+\emph{null-polarization distributions} represent the process of the
+group finding \emph{mere agreement}, which is when the average group
+opinion is constant over time, but opinion variance decreases to
+represent agreement (or, equivalently, \emph{consensus}~(DeGroot 1974)).
 
-Null-polarization  We thus model opinions at each time, $t=\mathrm{pre}$ and
-$t=\mathrm{post}$, as random variables drawn from a normal distribution with
-mean $\mu_t$ and standard deviation $\sigma_t$.  we drop the time
+\subsection{Latent Opinion Model}\label{latent-opinion-model}
 
-The null-polarization distributions are represented by
+Throughout this paper we model a participant \(i\)'s latent opinion at
+time \(t\) as a normally-distributed random variable,
+\begin{equation}\phantomsection\label{eq-random-variable-def}{ 
+\omega_{i,t} \sim
+\mathcal{N}(\mu_t, \sigma_t),
+}\end{equation} with mean \(\mu_t\) and standard deviation \(\sigma_t\).
+Group polarization in terms of latent opinions is then defined as any
+pair of pre- and post-discussion opinion distributions where the mean
+increases in magnitude over time---formally,
+\(| \mu_\mathrm{post}| > | \mu_\mathrm{pre}|\). We need to take
+magnitudes since radicalization may be towards increasingly negative or
+positive opinion values.
 
-in a simple, general way—i.e., a decrease in opinion variance but a static
-opinion mean—we represent . In the general group polarization hypothesis, the
-group is assumed to start out mostly agreeing in principle (i.e., mostly all
-agree or disagree) but to different extremes. After discussion
+For group polarization experiments, only two times are relevant:
+\emph{pre-discussion}, \(t=\mathrm{pre}\), and \emph{post-discussion},
+\(t=\mathrm{post}\). Since participant groups are assumed to be
+likeminded, we could also assume they come to agree more over time as
+well as become more extreme in our group polarization model, or in terms
+of the latent opinion variance,
+\(\sigma_\mathrm{pre}> \sigma_\mathrm{post}\)---but only one study in
+our ten actually measured opinion variance to check this (Schkade,
+Sunstein, and Hastie 2010).
 
-Showing a replication is undecidable is sufficient to undermine it. However,
-it does not tell us what is the probability that a particular experimental
-design produces false positives, due to clipping or any other problem.
+A \emph{null-polarization model} is a model that is consistent with the
+\emph{negation}, or \emph{null}, of the \emph{group polarization
+hypothesis}. The group polarization hypothesis is that group discussion
+causes the average group opinion to increase, assuming the group is
+aligned on a particular issue, though opinion strength varies. Therefore
+a null model is one where the latent mean stays constant, i.e.,
+\(\mu = \mu_\mathrm{pre} = \mu_\mathrm{post}\).
 
-first develop a mathematical model of estimate the best-case scenario false
-positive rate for each experimental design,
+\emph{Mere agreement} is the type of null-polarization model we use
+here. Mere agreement is a null-model because the mean latent opinion is
+set to be constant before and after discussion. We represent
+\emph{agreement}---or equivalently, \emph{consensus}---as a decrease in
+the standard deviation that defines the pre- and post-discussion latent
+opinion distributions. The mere agreement model is the cornerstone of
+our analysis, since it is empirically plausible and generates
+\emph{spurious} observations of group polarization under mere agreement
+models, due to measurement artifacts induced by clipping all latent
+opinion values outside the measurement range---we call this
+\emph{spurious polarization} for short. We simulate opinion measurements
+in two different ways, but both use the mere agreement model of latent
+opinions to test whether and by how much the measurement procedure
+distorts null-polarization to look like group polarization.
 
-\subsection{Mathematical Model of Null-Polarization Measurement Artifacts}
+{[} TABLE
+{[}tab:latent-variables{]}(\#tab:latent-variables)\{reference-type=``ref''
+reference=``tab:latent-variables''\} ABOUT HERE {]}
 
-Latent opinions are modeled as normally-distributed random variables.  
-This is a simple, sufficiently general way to represent group-level changes
-in opinions as encoded in the general hypothesis of group polarization, that
-likeminded groups tend to become more extreme on average as they discuss.
-For time, $t$, then, we represent each participant $i$'s opinion as 
-\begin{equation} 
-  \omega_{i,t} \sim \mathcal{N}(\mu_t, \sigma_t).
-  \label{eq:opinionDistribution} 
-\end{equation} 
-\noindent
+\subsection{Two Models of Group Polarization
+Measurement}\label{two-models-of-group-polarization-measurement}
 
-Group-level opinion change, then, can be represented as a change in opinion
-mean, opinion variance, or both. Group polarization occurs 
-when a group's average opinion increases over time. When likeminded people
-interact, they tend to come to agree more over time—they \emph{find consensus}. 
-By definition this means opinion variance decreases, 
-whether the average changes or not.
+We simulate measurement in two different ways---i.e., with two different
+models---for different parts of our analysis. The first is the
+\emph{expectation model}: an efficient way to \emph{simulate measurement
+of the observed average} opinion of a group, using the
+\emph{mathematical expectation value} of the distribution of measured
+opinions as a proxy. In the second model, the \emph{stochastic model},
+we simulate individual participant responses by first drawing an opinion
+for each participant from the latent opinion distribution
+(Equation~\ref{eq-random-variable-def}). The opinions are binned to the
+appropriate ordinal value.
 
-We can define a null-polarization model, then, as one where the group's
-average opinion stays constant, but the variance decreases: $\sigmapost
-\leq \sigmapre$. Since the mean opinion is constant we just call it 
-$\mu = \mupre = \mupost$. For conciseness, we define a null-polarization
-model as a 3-tuple, $\nu = (\mu, \sigmapre, \sigmapost)$.
+We use the expectation model for its efficiency in the first step of our
+analysis: to find a \emph{null-polarization model} that also fits the
+data for each experimental condition studied---if we find one that means
+we must retract this replication since this proves that the group
+polarization is undecidable with the data we have. Once we find any such
+undecidable results, we use the null-polarization model to help us
+further evaluate the experimental design itself.
 
-\begin{center}
-\vspace{1.5em}
-[ TABLE \ref{tab:latent-variables} ABOUT HERE ]
-\vspace{1.5em}
-\end{center}
+In the \emph{stochastic model}, we simulate group polarization data for
+the \emph{Evaluation} step of the analysis. We simulate experimental
+data, with the same number of participants as the original experiment,
+using the null-polarization model found in the \emph{Identification}
+step, if one was found.
 
-\subsubsection{Simulated opinion measurements}
+We \emph{evaluate} the experimental design, \(e\), in terms of the false
+positive rate, also known as the Type I error rate or family-wise error
+rate, denoted \(\alpha_e\). We close our analysis by demonstrating how
+our method can be inverted to \emph{evaluate} experimental designs in
+the pre-registration stage to avoid preventable replication failures
+like we identify here.
 
-We simulate the collection of an \emph{ordinal distribution} of \emph{ordinal
-observations}, as was performed in a prototypical group polarization experiment 
-by integrating the latent opinion probability density function, $f(\omega)$,
-over each bin's corresponding range of opinion values. For example, consider an
-ordinal scale that runs from -3 (strongly disagree) to +3 (strongly agree)
-(as used by \textcite{Moscovici1969}, for example).
-the frequency with which the ``2'' bin is chosen is just the total
-area underneath the probability density curve between 1.5 and 2.5.
+The measurement models are introduced in detail along with the
+corresponding stage of the analysis that uses them. First, the
+\emph{Identification} stage that uses the \emph{expectation} measurement
+model; second, the \emph{Evaluation} stage that uses the
+\emph{stochastic} measurement model.
 
-Formally, the probability density function is
-\begin{equation}
-  f(\omega) = 
-    \frac{1}{\sigma \sqrt{2 \pi}} e^{-\frac{(\omega - \mu)^2}{2 \sigma^2}}.
-  \label{eq:normal_pdf}
-\end{equation}
-\noindent
-To set up our measurement simulation, we will represent
-oridnal bin values as $b_i$ with $i=1,\ldots,B$—i.e., there are $B$ bins. 
-To create $B$ bins there must be $B+1$ \emph{thresholds} that define what
-range of latent opinions correspond to which ordinal bin value.
+\subsection{\texorpdfstring{Part 1: \emph{Identify} Invalid
+Replications}{Part 1: Identify Invalid Replications}}\label{part-1-identify-invalid-replications}
 
-\begin{center}
-  \vspace{1.5em}
-  [ TABLE \ref{tab:experiment-parameters} ABOUT HERE ]
-  \vspace{1.5em}
-\end{center}
+In the first stage of our analysis, we try to \emph{identify} a null
+polarization models, \(\nu\), that invalidate published detections of
+group polarization if the measurement procedure induces \emph{spurious
+group polarization} for that null polarization model. Null polarization
+can take many forms---all that is required is that the latent mean
+opinion stays constant before and after discussion. Here we use a
+\emph{mere agreement} null polarization model, where we simulate
+\emph{consensus}, the tendency for sufficiently similar opinions to
+become more similar---but we leave out polarization. In mere agreement,
+moderate and extreme opinions approach change by equal magnitude towards
+each other. We therefore represent consensus as a decrease in the
+standard deviation of the latent opinion distribution from pre- to
+post-discussion. We simulate the measurement of the group's average
+opinion before and after discussion and measure the spurious group
+polarization, if there is any. We use this procedure in a search
+algorithm we developed to identify which combination of constant latent
+mean and pre- and post-discussion opinion variances generate spurious
+group polarization, if there are any.
 
-All bins have a width of 1 in latent opinion space, except for the two most
-extreme bins at opposite ends of the sentiment spectrum—this works because
-bin values are always separated by 1. All continuous latent opinion values
-$\omega < b_1 + \frac{1}{2} = \theta_1$ get mapped to the ordinal measurement
-value $b_1$—so, $b_1$'s lower threshold is $\theta_0 = -\infty$.  Similarly,
-all opinions $\omega > b_B - 0.5 = \theta_B$ are reported as the maximum bin
-value, $b_B$, with an effective upper bound of $\theta_{B+1} = +\infty$.  We
-can define all thresholds compactly as \cite{Liddell2018}
-\begin{equation}
+Our invalidation strategy, then, is to find a mere agreement null
+polarization model, \(\nu\), that predicts the exact observations
+reported in each experimental replication in our corpus. If one is
+found, then we have found a model consistent with the \emph{negation} of
+the group polarization hypothesis that explains the data as well as the
+group polarization hypothesis itself. In other words, the hypothesis is
+not proved or disproved given the data, but \emph{undecidable}: the data
+cannot \emph{decide} the hypothesis one way or another because of
+potential measurement artifacts. Because the experiment failed to
+produce a valid replication, the results are effectively stripped of
+evidentiary value and warrant exclusion from the aggregate evidence
+base. We apply our analysis to every experimental condition across all
+ten journal articles in our corpus, which we label \(e\).
+
+But how do we find a \(\nu\) that generates spurious group polarization?
+And how do we even simulate the group polarization experiments and
+measurements that make mere agreement in latent opinions look like group
+polarization? We simulate experimental measurements of group
+polarization by using the mathematical \emph{expectation value} of
+latent opinions defined by \(\nu\) as a proxy for the observed average
+value of ordinal-valued reported opinions (described in the first
+subsection below). We then use this measurement model in a numerical
+root-finding algorithm that identifies a \(\nu\) that matches published
+observations, if possible. If we find such a \(\nu\), this means we have
+a model consistent with the null hypothesis that explains published
+observations. In other words, the replication is invalidated since the
+group polarization hypothesis is \emph{undecidable} given the
+data---i.e., it cannot decide whether the hypothesis or its null is
+correct.
+
+So, to review, this subsection introduces how we identify which
+experimental data can be explained by a mere agreement null model, where
+the ordinal survey measurement procedure induces a clipping artifact
+that hides changes in extreme opinions that fall outside the measurement
+scale.
+
+\subsubsection{The Expectation Model of Spurious Group
+Polarization}\label{the-expectation-model-of-spurious-group-polarization}
+
+We use the \emph{expectation model} of opinion measurement to
+\emph{Identify} which, if any, mere agreement null-polarization models
+generate spurious group polarization in a particular experiment, \(e\).
+The trick to this model, and the source of its name, is that we use the
+theoretical \emph{expected value} of opinion measurements as a proxy for
+the \emph{observed average} of participant opinions under an
+experimental design, \(e\).
+
+We simulate the measurement of the mean observed opinion at time \(t\)
+by first calculating integrating the normal probability density
+function, \(f(\omega)\), over measurement bin boundaries, over each
+bin's corresponding range of opinion values. For example, consider an
+ordinal scale that runs from -3 (strongly disagree) to +3 (strongly
+agree) (as used by Moscovici and Zavalloni (1969), for example). the
+frequency with which the ``2'' bin is chosen is just the total area
+underneath the probability density curve between 1.5 and 2.5.
+
+All bins have a width of 1 in latent opinion space, except for the two
+most extreme bins at opposite ends of the sentiment spectrum---this
+works because bin values are always separated by 1. All continuous
+latent opinion values \(\omega < b_1 + \frac{1}{2} = \theta_1\) get
+mapped to the ordinal measurement value \(b_1\)---so, \(b_1\)'s lower
+threshold is \(\theta_0 = -\infty\). Similarly, all opinions
+\(\omega > b_B - 0.5 = \theta_B\) are reported as the maximum bin value,
+\(b_B\), with an effective upper bound of \(\theta_{B+1} = +\infty\). We
+can define all thresholds compactly as (Liddell and Kruschke 2018)
+\begin{equation}\phantomsection\label{eq-bin-thresholds}{
 \theta_j = \begin{cases}
@@
   \infty          & \text{ if } j=B \\
-  b_1 + \frac{j + 1}{2} & \text{ otherwise.}
+  b_j + \frac{1}{2} & \text{ otherwise.}
 \end{cases}
-\label{eq:thresholds}
-\end{equation}
-\noindent
+}\end{equation}
 
-We model the measurement of ordinal opinions by calculating the probability
-density function over bins, written
-\begin{equation}
-  f[b_i] = \int_{\theta_{i - 1}}^{\theta_i} f(\omega) d\omega~,
-  \label{eq:measurement-math}
-\end{equation}
-\noindent
-switching to square brackets for the argument to $f$ now to emphasize that 
-bins are discrete. With this ordinal measurement model we can caluclate the 
-mean opinion of the simulated measurements like so
-\begin{equation}
-  \bar{o}_t = \frac{1}{B} \sum_{b=b_1}^{b_B} b \cdot f_t[b].
-\end{equation}
-\noindent
+The probability density that a participant reports the opinion in bin
+index \(i\) at time \(t\) is \[
+f_{it}(\mu, \sigma_t) = \int_{\theta_{i - 1}}^{\theta_{i}} f(\omega; \mu, \sigma_t)~d\omega,
+\] which we calculated in R to evaluate the integral at the two
+threshold values \(\theta_{i-1}\) and \(\theta_i\). Note that the normal
+probability density function we used is parameterized by mean \(\mu\)
+and standard deviation \(\sigma_t\). \[
+f(\omega; \mu, \sigma_t) = 
+    \frac{1}{\sigma_t \sqrt{2 \pi}} e^{-\frac{(\omega - \mu)^2}{2 \sigma_t^2}}.
+\]
 
-\begin{center}
-\vspace{1.5em}
-[ FIGURE \ref{fig:distros} ABOUT HERE ]
-\vspace{1.5em}
-\end{center}
+To simulate the measurement of the \emph{average} observed opinion at
+time \(t\), \(\bar{o}_t\), as the expected value of the probability
+density function of observed ordinal opinions. In other words, we
+simulate \(\bar{o}_t\) by calculating the weighted sum of products
+between the probability of observing each bin value and the bin value
+itself: \begin{equation}\phantomsection\label{eq-sim-ave-observed}{
+\bar{o}_t(\nu, e) = \sum_{i=1}^B b_i f_{it}(\nu, e).
+}\end{equation}
 
 Group polarization is the difference between the post-deliberation mean
-observed opinion and the pre-deliberation one, 
-which we can write as 
-\begin{equation}
-  g = \bar{o}_\mathrm{post} - \bar{o}_\mathrm{pre}.
-  \label{eq:group-polarization}
-\end{equation}
-\noindent
+observed opinion and the pre-deliberation one. We will calculate the
+magnitude of spurious group polarization under a mere agreement null
+model represented by its collected distribution parameters,
+\(\nu = (\mu, \sigma_\mathrm{pre}, \sigma_\mathrm{post})\), under
+experimental design, \(e\):
+\begin{equation}\phantomsection\label{eq-spurious-gp}{
+g(\nu, e) = |\bar{o}_\mathrm{post}(\nu, e)| - |\bar{o}_\mathrm{pre}(\nu, e)|
+}\end{equation}
 
-\begin{center}
-\vspace{1.5em}
-[ TABLE \ref{tab:simulated-observations} ABOUT HERE ]
-\vspace{1.5em}
-\end{center}
+In the \emph{Inspect} stage of our analysis we will determine which
+group polarization replications are invalidated due to equifinality
+between the \emph{mere agreement null-polarization hypothesis} on the
+one hand, and the \emph{group polarization hypothesis} on the other.
+This requires more than just showing there is some \(\nu\) that causes
+spurious polarization for some experiment where \(g(\nu, e) > 0\). We
+have to find a \(\nu\) that generates the exact \(g(\nu, e)\) reported
+in the original journal article \emph{and} check that simulated
+\(\bar{o}_\mathrm{pre}\) and \(\bar{o}_\mathrm{pre}\) match published
+values.
 
-\subsection{Root-finding Algorithm to Find Null-Polarization Models with Artifacts}
+{[} TABLE
+\hyperref[tab:experiment-parameters]{{[}tab:experiment-parameters{]}}
+ABOUT HERE {]}
 
-\subsection{Corpus Analysis to Determine Undecidable Replications}
+{[} TABLE
+\hyperref[tab:simulated-observations]{{[}tab:simulated-observations{]}}
+ABOUT HERE {]}
 
-To analyze published replications we need to expand our notation to index
-individual \emph{experiments} with replications, and from which 
-different journal articles, track and analyze it,
+\subsubsection{Example: Invalidating the ``Americans'' condition in
+Moscovici and Zavalloni
+(1969)}\label{example-invalidating-the-americans-condition-in-moscovici-and-zavalloni-1969}
 
-\subsection{Evaluating Grouop Polarization Experimental Designs}
+One of the replications we invalidate here is from Moscovici and
+Zavalloni (1969), the ``Americans'' condition as they called it---we
+coded it \texttt{Moscovici1969\_Americans}. They report that participant
+opinions polarized after discussing the survey topic, ``American
+economic aid is always used for political pressure.'' They report survey
+responses increased in extremity from an average of -0.9 before
+discussion to -1.6 after, on a seven-bin scale from -3 indicating
+``strongly disagree'' to +3 indicating ``strongly agree''.
+
+Moscovici and Zavalloni say their results confirm the basic hypothesis,
+that \emph{group discussion causes polarization}. But we found a mere
+agreement null-polarization model that generates spurious group
+polarization using our model---specifically,
+\(\nu = (\mu = -1.2, \sigma_\mathrm{pre}= 4.3, and \sigma_\mathrm{post}= 1.9)\).
+This means that the hypothesis is actually undecidable given the data,
+invalidating Moscovici and Zavalloni's replication. When a hypothesis
+and its null explain data equally well, the data tell us nothing.
+
+Now let's use the expectation model to prove that this \(\nu\) generates
+spurious group polarization in this experiment, i.e.,
+\(g(\nu, e= \mathrm{Moscovici1969\_Americans}) > 0\). In the process we
+demonstrate how the expectation model works to simulate the measurement
+of spurious group polarization (Equation~\ref{eq-spurious-gp}).
+
+First, we calculate the probability density vector using the
+\href{https://github.com/mt-digital/gp-statmod/blob/main/GroupPolarizationStatmod/model.R\#L44}{\texttt{makeProbVec}}
+function in the file
+\href{https://github.com/mt-digital/gp-statmod/blob/main/GroupPolarizationStatmod/model.R}{\texttt{GroupPolarizationStatmod/model.R}}
+in the \href{https://github.com/mt-digital/gp-statmod}{project code
+repository on GitHub} for both experiment times: pre- and
+post-discussion: \[
+f_\mathrm{pre} = 
+\begin{pmatrix}
+  0.38 & 0.09 & 0.09 & 0.09 & 0.08 & 0.07 & 0.19
+\end{pmatrix}
+\] and \[
+f_\mathrm{post} = \
+\begin{pmatrix} 
+0.25 & 0.19 & 0.21 & 0.17 & 0.11 & 0.05 & 0.03 
+\end{pmatrix}.
+\]
+
+When we plug in the bin values and copmuted \(f_\mathrm{pre}\) values
+into the simulated average pre-discussion opinion definition
+(Equation~\ref{eq-sim-ave-observed}), we get \[
+\bar{o}_\mathrm{pre} = -3 \cdot 0.38~~+-2\cdot 0.09~~+\ldots+3\cdot0.19 = -0.6
+\] and \[
+\bar{o}_\mathrm{post} = -3 \cdot 0.25~~+-2\cdot 0.19~~+\ldots+3\cdot0.03 = -1.04.
+\] So, we find that for the \texttt{Moscovici1969\_Americans}, our
+null-polarization model appears polarized, with the exact same average
+pre- and post-discussion opinions reported by Moscovici and Zavalloni
+(1969), and the same magnitude of radicalization: \[
+g(\nu, e) = \bar{o}_\mathrm{post} - \bar{o}_\mathrm{pre} = -1.04 - (-0.6) = - 0.4.
+\]
+
+FIGURE \hyperref[fig:distros]{{[}fig:distros{]}} ABOUT HERE
+
+Since the simulated average opinions match the observed ones---note that
+as a consequence, the spurious polarization magnitude also matches the
+observed one. In other words, we \emph{identified} a null model that
+explains the data. If both a hypothesis and its null explain data
+equally well, then the hypothesis is \emph{undecidable} under the
+available data. This invalidates this experiment's claim to be a
+replication of support for the group polarization hypothesis. We
+replicate this analysis for all 57 experimental conditions that purport
+to support the group polarization hypothesis. To facilitate this
+analysis, we created a web app to manage the data entry, the process of
+identifying invalid replications, and managing the outcome data of which
+replications were invalidated, and what null model parameters generated
+the spurious polarization.
+
+\subsubsection{System for identifying invalid
+replications}\label{system-for-identifying-invalid-replications}
+
+The next step is to scale up the analysis we did manually in the example
+above, including develop a systematic way to find, and not guess, the
+set of null model parameters that
+
+\begin{figure}
+
+\centering{
+
+\includegraphics[width=10.53125in,height=\textheight,keepaspectratio]{Figures/Model/SoftwareMap_Mashup.png}
+
+}
+
+\caption{\label{fig-computational-system}\textbf{Computational system to
+interactively simulate opinion change, identify invalid replications,
+and evaluate experimental designs.}}
+
+\end{figure}%
+
+\subsection{\texorpdfstring{Part 2: \emph{Evaluate} Experimental
+Designs}{Part 2: Evaluate Experimental Designs}}\label{part-2-evaluate-experimental-designs}
+
+In the previous step we identified which, if any, null polarization
+models generate spurious group polarization with the expectation model
+of opinion measurement. This is enough to invalidate a replication, but
+it leaves unanswered questions. First, we would like to know how what
+the chance was that a published detection of group polarization was
+actually a false positive, assuming that participant opinions were
+distributed according to the null model identified in the first part of
+the analysis. This is equivalent to the false positive rate of the
+experimental design, which we denote \(\alpha_e\) for experiment \(e\).
+Second, we would like to estimate the false discovery rate for our
+corpus, which is a good proxy for the field as a whole, since we
+selected top quality, highly influential work that has been cited
+thousands of times. The false discovery rate in this case tells us the
+probability that a published detection of group polarization is actually
+a false detection. (Equation~\ref{eq-fdr}).
+
+The core of our evaluation is the quantification of the probability that
+a spurious result---an effect generated by the null model---is falsely
+interpreted as genuine group polarization. We calculate this probability
+across three distinct units of analysis, building from the specific to
+the general:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  The Type I Error Rate (\(\alpha_e\)): The rate of falsely detecting
+  group polarization in a single experimental condition (\(e\)). This
+  directly quantifies the risk posed by the mere agreement model
+  identified for that specific experiment.
+\item
+  The study-level Family-Wise Error Rate (\(\alpha_s\)): The mean Type I
+  Error Rate aggregated across all conditions within a single study
+  (\(s\)). This estimates the false claim rate across the set of tests
+  performed in that study.
+\item
+  The Family-Wise Error Rate (\(\alpha_{\text{all}}\)): The mean Type I
+  Error Rate aggregated across all 57 experimental conditions in the
+  corpus (\(E\)). This estimates the expected false claim rate for a
+  researcher adopting the general experimental design found throughout
+  the literature.
+\end{enumerate}
+
+These calculated Type I Error Rates (\(\alpha\)) are then combined with
+fixed external parameters to estimate the False Discovery Rate.
+Following convention, we adopt a fixed Statistical Power of \(W=0.80\)
+(and test sensitivity with alternative values) and a Base Rate of
+\(b=0.10\). The relationship between the units of analysis and the
+estimated error rates is summarized in Table \textbf{?@tbl-error-rates}.
+
+Statistical Error Rates and Units of Analysis
+
+Unit of Analysis
+
+Type
+
+Notation
+
+Definition
+
+Experiment, \(e\)
+
+Type I
+
+\(\alpha_e\)
+
+The rate of falsely rejecting the null hypothesis (\(d=0\)) for that
+single test (used in aggregation).
+
+All experiments in a study, \(s\)
+
+Family-Wise
+
+\(\alpha_s\)
+
+The mean Type I Error Rate across all conditions in that specific study.
+
+All Experiments
+
+Family-Wise
+
+\(\alpha_{\text{E}}\)
+
+The mean Type I Error Rate across all conditions identified as plausibly
+spurious.
+
+\subsubsection{Generative Model of Spurious Group
+Polarization}\label{generative-model-of-spurious-group-polarization}
+
+To evaluate an experimental design's ability to properly identify null
+effects, we need to simulate participant responses assuming the mere
+agreement null model that invalidated the replication (\(\nu_e\)). We
+use the stochastic model for this simulation stage, as it correctly
+models the noise inherent in individual participant measurement and
+response. The ordinal, measured opinions are generated by executing the
+following steps for \(N_e\) simulated participants across \(N_T\)
+independent trials:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  \textbf{Latent Opinion Sampling:} For each participant \(i\) and time
+  \(t \in \{\text{pre, post}\}\), we draw a continuous latent opinion
+  \(\omega_{i,t}\) from the corresponding normal distribution defined by
+  the null model \(\nu_e\):
+  \[\omega_{i,t} \sim \mathcal{N}(\mu, \sigma_t)\] where \(\mu\) is the
+  constant latent mean and \(\sigma_t\) is the time-specific standard
+  deviation derived from the mere agreement null model \(\nu_e\).
+\item
+  \textbf{Measurement Simulation (Ordinal Binning):} Each continuous
+  latent opinion \(\omega_{i,t}\) is mapped onto the experiment's
+  ordinal scale (e.g., \(-3\) to \(+3\)). This mapping is determined by
+  the fixed measurement thresholds, \(\theta_j\), established in the
+  \emph{Expectation Model} subsection
+  (Equation~\ref{eq-bin-thresholds}). This process simulates the
+  clipping artifact that generates spurious polarization in the
+  observable data.
+\end{enumerate}
+
+This results in a simulated dataset of \(2N_e\) rows (two time steps per
+participant) for a single trial, consisting of ordinal opinion values
+that are then passed to the statistical inference pipeline.
+
+\subsubsection{Statistical Inference
+Pipeline}\label{statistical-inference-pipeline}
+
+To complete our our simulated experimental design we need to be able to
+statistically infer the effect sizes that measure the strength of group
+polarization in an experiment. To keep the presentation simple and
+transparent, we infer effect sizes by inferring the parameters \(\mu_t\)
+and \(\sigma_t\) of the hypothesized normal distribution that generates
+the data---\emph{literally} generates the data in our simulations, but
+is just a model of the distribution of group opinions in real-world
+experiments.
+
+\subsubsection{False positive rate}\label{false-positive-rate}
+
+To calculate the false detection rate we begin by simulating pre- and
+post-deliberation opinions by drawing \(N_e\) samples from normal latent
+opinion distributions that generated spurious group polarization, i.e.,
+one sample at each experiment stage for each simulated participant. The
+latent opinion distributions are specified with the mere agreement null
+model labelled \(\nu_e\).
+
+The family-wise error rate is the fraction of trials with
+\(d_{ei} > d^*\), written
+\begin{equation}\phantomsection\label{eq-study-aggregate}{
+  \alpha_e = \frac{1}{N_T} \sum_i^{N_T} \Theta(d_{ei}; d^*),
+}\end{equation} where \[
+  \Theta(d_{ei}; d^*) = 
+    \begin{cases}
+      1 & \text{ if } d_{ei} > d^* \\
+      0 & \text{ otherwise.}
+    \end{cases}
+\] \(d^*\) is the \emph{significance threshold}, which must be specified
+by the experimenter. There are three common significance levels which we
+adopt here: \(d^* = 0.2\) for ``low'', \(d^*=0.5\) for ``medium'', and
+\(d^* = 0.8\) for ``high''.
+
+To calculate the family-wise error rate across all experimental
+conditions of a study, \(s\), \[
+\alpha_s = \frac{1}{|s|} \sum_{e \in s} \alpha_e,
+\] where \(e \in s\) indicates sum over all experimental conditions
+\(e\) presented in the article associated with study \(s\). The mean
+across all results identified as plausibly spurious is \[
+ \alpha_\mathrm{all} = \frac{1}{|E|} \sum_{e \in E} \alpha_e.
+\]
+
+In the final part of our \emph{Evaluation} analysis we adapt this
+analysis to re-align the question of statistical rigor: we can be
+specific now, and instead of wondering, ``what significance level is
+truly rigorous'', we can use our techniques to make the choice of
+\(d^*\) to conform to a specific design requirement. Specifically we
+will calculate what \(d^*\) must be in order to limit
+\(\alpha \leq 0.05\) as an example.
+
+\subsubsection{False discovery rate}\label{false-discovery-rate}
+
+The false discovery rate is the false positive rate, scaled by the rate
+of true negative effects and normalized by the frequency of true
+positive effects\textasciitilde{} (Equation~\ref{eq-fdr}) and assuming
+empirically-motivated values for base rate, \(b\), and statistical
+power, \(W\), as well as more pessimistic and optimistic settings for
+base rate and power. Using this notation, the false detection rate is
+
+\begin{equation}\phantomsection\label{eq-fdr}{
+\mathrm{FDR}= \frac{(1 - b)\alpha}{(1 - b)\alpha + bW}.
+}\end{equation}
+
+\subsection{\texorpdfstring{Part 3: \emph{Specify} Significance Threshold
+for Experimental
+Design}{Part 3: Specify Significance Threshold for Experimental Design}}\label{part-3-specify-significance-threshold-for-experimental-design}
+
+In the final stage of our analysis, we pivot from evaluating the outcome
+of published designs to \emph{specifying the statistical requirements
+for future designs}. The primary goal of this specification is to invert
+the standard question of statistical rigor: instead of asking if an
+effect is `significant' given a fixed effect-size threshold (e.g.,
+\(d^{*} = 0.5\)), we ask what significance threshold is necessary to
+limit the false positive rate to a desired maximum, such as \(5\%\) (or
+\(\alpha = 0.05\)).
+
+We leverage the evaluation framework developed in Part 2, based on the
+mere agreement null model \(\nu_e\) identified for each experimental
+condition \(e\). Our objective is to calculate the function
+\(\alpha_e(d^*)\) across a sufficient range of \(d^*\) values. This
+allows us to determine the lowest \(d^*\) that satisfies the design
+requirement \(\alpha_e(d^*) \leq 0.05\).
+
+The function \(\alpha_e(d^*)\) is the simulated false positive rate as a
+function of the significance threshold, \(d^*\), defined as: \[
+\alpha_e(d^*) = \frac{1}{N_T} \sum_i^{N_T} \Theta(d_{ei}; d^*),
+\] where \(N_T\) is the number of null trials, \(d_{ei}\) is the
+calculated effect size for trial \(i\), and \(\Theta\) is the indicator
+function that flags trials where the effect size \(d_{ei}\) exceeds the
+significance threshold \(d^*\). In the interest of giving the most
+generous best-case scenario possible, we only count positive effect
+sizes because only these indicate increased extremism over time.
+Negative effect sizes indicate group opinions became more
+moderate---such observations do not support the group polarization
+hypothesis that opinions become more extreme following discussion.
+
+Because our system is automated, and the code publicly available, it can
+be adapted and used to design future group polarization experiments with
+empirically-grounded statistical rigor. Experimenters can specify a
+desired maximum false positive rate, such as \(\alpha = 0.05\), and our
+system will calculate the minimum effect size threshold \(d^*\) that
+must be exceeded to achieve that level of rigor under the mere agreement
+null model identified for their experimental design. This prescriptive
+step transforms the simulation from a post-hoc critique into a
+forward-looking design tool, giving experimenters an empirically-derived
+minimum effect size they must exceed to maintain statistical confidence
+in their results.
+
+\subsection{Analysis Outline and
+Summary}\label{analysis-outline-and-summary}
+\addcontentsline{toc}{section}{Analysis Outline and Summary}
+
+%\phantomsection\label{refs}
+%\begin{CSLReferences}{1}{0}
+%\bibitem[\citeproctext]{ref-DeGroot1974}
+%DeGroot, Morris H. 1974. {``{Reaching a Consensus}.''} \emph{Journal of
+%the American Statistical Association} 69 (345): 118--21.
+%
+%\bibitem[\citeproctext]{ref-Liddell2018}
+%Liddell, Torrin M., and John K. Kruschke. 2018. {``{Analyzing ordinal
+%data with metric models: What could possibly go wrong?}''} \emph{Journal
+%of Experimental Social Psychology} 79: 328--48.
+%\url{https://doi.org/10.1016/j.jesp.2018.08.009}.
+%
+%\bibitem[\citeproctext]{ref-Moscovici1969}
+%Moscovici, Serge, and Marisa Zavalloni. 1969. {``{The group as a
+%polarizer of attitudes}.''} \emph{Journal of Personality and Social
+%Psychology} 12 (2): 125--35.
+%
+%\bibitem[\citeproctext]{ref-Schkade2010}
+%Schkade, David, Cass R. Sunstein, and Reid Hastie. 2010. {``{When
+%deliberation produces extremism}.''} \emph{Critical Review} 22 (2):
+%227--52. \url{https://doi.org/10.1080/08913811.2010.508634}.
+%
+%\end{CSLReferences}
+%
+%
+%
+%
+%\end{document}

### 2025-12-09 GroupPolarizationStatmod/code_examples.qmd
working on stats inference

--- 2025-11-20 GroupPolarizationStatmod/paper/code_examples.qmd
@@
     wrap: 90
-bibliography: this.bib
+bibliography: "paper/this.bib"
+
+execute:
+  # echo: false     # hide code
+  warning: false  # hide warnings
+  message: false  # hide messages
+  # error: false
 ---
@@
 
-![](images/clipboard-1443885641.png){width="628"}
+![**Mean group opinions pre- and post-discussion from Moscovici & Zavalloni
+(1969).**](images/clipboard-1443885641.png){width="575"}
 
@@
 observed opinion distributions appear to polarize (i.e., *radicalize*, or *become more
-extreme*) that does not represent the true change in latent opinions. The measurement
+extreme*) that does notrepresent the true change in latent opinions. The measurement
 procedure *clips* extreme opinions less than the minimum bin value, or greater than the
@@
 # Adapt for your own computer
-setwd("~/workspace/gp-statmod/GroupPolarizationStatmod/")
+# setwd("~/workspace/gp-statmod/GroupPolarizationStatmod/")
 # Load makeProbVec from file model.R
@@
 @Moscovici1969 and the one we simulated.
+
+# The Generative Model and Bayesian Inference for False Detections
+
+The analysis of the generative model has two major components:
+
+1.  **Calculate false positive/detection rates for the simulated null polarization
+    model**, *mere agreement*—where polarization experiment participant opinions agree
+    more following discussion than before. Recall this is modeled as a decrease in the
+    standard deviation of latent opinions, but a constant mean latent opinion.
+2.  **Find an effect–detection threshold**, $d^*(\alpha)$, such that, under the null
+    model, the probability of observing an estimated effect size exceeding
+    \$d\^\*(\\alpha)\$ is \$\alpha\$: $$
+     \Pr\!\left( \hat{d} \ge
+     d^{*}(\alpha) \mid \mathcal{M}_{0} \right) = \alpha.
+     $$ 
+     This yields a null-model–calibrated false–positive rate for detecting a group
+    polarization effect.
+
+## Simulate each experimental design's false positive rate
+
+Here's how to calculate the Type I or family-wise error rate for all the experiments
+(though I misnamed the function back in the day, so it calls all of the error rates
+"family-wise" because I had understood them to be synonyms).
+
+```{r}
+source("src/fwer.R")
+
+probit_fits_dir <- "data/probit_fits"
+# Set a "moderate" significance, according to Cohen's definitions 
+significance_val <- 0.5
+fwer_tbl <- calculate_fwer(probit_fits_dir, significance_val);
+
+print(head(fwer_tbl))
+```
+
+This is used inside the `calc_fdr_vs_significance` function that adds another column to
+this dataframe, the false discovery rate (FDR) for a set of significance values. Here it
+is being called with the default metascience parameters with a base rate of $b=0.1$ and a
+statistical power of $W=0.8$. We consider three effect sizes in this paper according to
+the : "small", "medium", and "large" effect sizes traditionally assumed practically
+meaningful across social sciences [@CohenBook1988].
+I put this function in the `scripts` directory because to run the full analyses presented in the paper it's only necessary to run `calc_fdr_vs_significance` since it creates a column for the false positive rate and the estimated false detection rate calculated for each experimental design we studied.
+
+This is reported for a variety of effect-detection thresholds, $\delta^*$, which the function name implies are "significance" values. When I first wrote this function
+which are really effect-detection thresholds—so the code slightly misuses the term "significance", which typically refers only to a p-value based effect-detection threshold in the *null-hypothesis significance testing* approach.
+I updated the prose in the article to reflect this distinction, but left the code as-is.
+I again call the function with the defaults for illustration.
+
+```{r}
+source("scripts/analysis.R")
+
+fdr_over_thresholds <- calc_fdr_vs_significance(
+  probit_fits_dir, 
+)
+```
+
+## Find effect-detection thresholds to achieve $\alpha = 0.05$
+
+In the second step of our generative model analysis I calculate

### 2025-12-09 GroupPolarizationStatmod/scripts/analysis.R

@@ -28,13 +28,11 @@ fdr = function(fwer, b, W) {
 }
 
 
-
 add_default_rows = function(fdr_vs_sig_tbl, default_fwer = 0.05) {
 
   # Figure out the number of sig values used.
   sigvals = unique(fdr_vs_sig_tbl$SigVal)
   n_sigvals = length(sigvals)
-  # print(sigvals)
 
   # Extract power and base rate.
   W = pull(fdr_vs_sig_tbl[1, "Power"])
@@ -46,26 +44,7 @@ add_default_rows = function(fdr_vs_sig_tbl, default_fwer = 0.05) {
   n_non_id = n_distinct(non_id)
   n_default_rows = n_non_id * n_sigvals
 
-  # # Set default_fwer to be half of the least fwer calculated for that article
-  # # by first finding the minimum fwer for the study.
-  # study_ids = non_id$StudyID
-
-  # default_fwers_tbl = 
-  #   fdr_vs_sig_tbl %>% 
-  #   filter(StudyID %in% study_ids) %>%
-  #   group_by(StudyID) %>%
-  #   summarize(default_fwer = min(FWER))
-
-  # # Then create a vector that fits with default_rows.
-  # default_fwers_vec = map(default_fwers_tbl$default_fwer, 
-  #                         \(fwer) rep(fwer / 2.0, n_sigvals))
-  # print(default_fwers_vec)
-
-  # default_fdr_vec = fdr(default_fwers_vec, b, W)
-
   default_fdr = fdr(default_fwer, b, W)
-
-  print(default_fdr)
 
   default_rows = tibble(StudyID = rep(non_id$StudyID, n_sigvals),
                         ExperimentID = rep(non_id$ExperimentID, n_sigvals),
@@ -130,7 +109,7 @@ calc_fdr_vs_significance = function(
   significance_vals = c(0.2, 0.5, 0.8)
 ) {
 
-  probit_fits_data = load_probit_data()
+  probit_fits_data = load_probit_data(probit_fits_dir)
   n_experiments = length(unique(probit_fits_data$ExperimentID))
 
   # Initialize return tibble.

### 2025-12-09 GroupPolarizationStatmod/code_examples.qmd
finally documenting well to avoid backtracking through codebase every time

@@
 
-### Purpose: demonstrate gp-stats codebase
+## Purpose: demonstrate gp-stats codebase
 
@@
 
-Change directories to go to the `paper` directory where this document lives:
+Change directories to go to the `gp-statmod/GroupPolarizationStatmod` directory where this document lives:
 
 ```         
-cd gp-statmod/GroupPolarizationStatmod/paper
+cd gp-statmod/GroupPolarizationStatmod
 ```
@@
 ![**Mean group opinions pre- and post-discussion from Moscovici & Zavalloni
-(1969).**](images/clipboard-1443885641.png){width="575"}
+(1969).**](code_examples_images/clipboard-1443885641.png){width="575"}
 
@@
 
+The first step in our _generative_ model of false detections is to _generate_ simulated 
+experimental data from a mere agreement null-polarization model.
+Then we can calculate how frequently an effect size was inferred that exceeded the effect-detection threshold.
+simulated null-model data our simulated null-models.
+
+## Simulate Experimental Data
+
 ## Simulate each experimental design's false positive rate
@@
 
-print(head(fwer_tbl))
+print(head(fwer_tbl, n = 10))
 ```
@@
 
-This is reported for a variety of effect-detection thresholds, $\delta^*$, which the function name implies are "significance" values. When I first wrote this function
-which are really effect-detection thresholds—so the code slightly misuses the term "significance", which typically refers only to a p-value based effect-detection threshold in the *null-hypothesis significance testing* approach.
+The `calc_fdr_vs_significance` function calculates the false-positive rate ($\alpha$) and false detection rate ($\mathrm{FDR}$) reported for a variety of effect-detection thresholds, $d^*$, as we call them in the paper.
+When I first wrote the code, though, I was thinking about detection thresholds as synonymous with "significance values".
+So that's why the function name, `calc_fdr_vs_significance`, includes the word `significance`, which is a stand-in for the maximally descriptive "effect-detection thresholds" terminology.
+This seemed fine to keep things simple and not change the name everywhere in the code.
+But it was right to change it for the paper, since *significance* is only understood in a p-value based  *null-hypothesis significance testing* approach, where a hypothesis is confirmed if the p-value is less than some particular threshold.
 I updated the prose in the article to reflect this distinction, but left the code as-is.
-I again call the function with the defaults for illustration.
+
+I again call the function with the defaults specified, for illustration. Remember the `significance_vals` map to _effect-detection thresholds_ ($d^*$) in the paper.
 
@@
 fdr_over_thresholds <- calc_fdr_vs_significance(
-  probit_fits_dir, 
+  probit_fits_dir = probit_fits_dir,  # defined in above code block
+  base_rate = 0.1, power = 0.8, significance_vals = c(0.2, 0.5, 0.8)
 )
+
+print(head(fdr_over_thresholds, n=10))
+```
+To see `SigVal = 0.8` (ie $d^* = 0.8$), check the `tail` of the table:
+```{r}
+print(tail(fdr_over_thresholds, n = 10))
 ```
@@
 In the second step of our generative model analysis I calculate
+
+# Analyses
+
+## _Identify_ Invalid Replications: All But 5% Were Invalidated
+
+## _Evaluate_ Experimental Designs: ~70% Chance a Polarization Detection was Erroneous
+
+## _Specify_ Effect-Detection Thresholds: Spec Experiments—Don't Just Design Them—and Avoid Informed Guesswork
+
+# References

### 2025-12-09 GroupPolarizationStatmod/scripts/plot/analysis.R

@@ -56,7 +56,7 @@ plot_ordinal_cohens <- function(ordinal_data_dir = "data/probit_fits",
 
 
 
-plot_fdr = function(save_path = "paper/Figures/Analysis/fdr_vs_experiment.pdf", 
+plot_fdr <- function(save_path = "paper/Figures/Analysis/fdr_vs_experiment.pdf", 
                     aggregate = FALSE, use_non_identified = FALSE,
                     default_fwer = 0.05,
                     base_rate = 0.1, power = 0.8, 
@@ -143,7 +143,7 @@ plot_fdr_aggregated = function(tbl, focal_sig_val,
 plot_fdr_nonaggregated = function(tbl, focal_sig_val, save_path) {
   
   # Filter to keep just "medium" effect sizes and sort factors.
-  tbl_foc = 
+  tbl_foc <- 
     filter(tbl, SigVal == focal_sig_val) %>% 
     mutate(StudyID = fct_reorder(StudyID, FWER, mean), 
            ExperimentID = fct_reorder(ExperimentID, FWER, max))
@@ -181,20 +181,20 @@ plot_sigval_for_low_fwer = function(target_fwer = 0.05,
   
   # Going to create ggplot, p.
   # First load and pre-process calculation of FWER ≤ target_fwer by experiment.
-  pre_tbl = calc_fdr_vs_significance(significance_vals = significance_vals)
-  pre_tbl_focal = filter(pre_tbl, SigVal == focal_sig_val) %>%
+  pre_tbl <- calc_fdr_vs_significance(significance_vals = significance_vals)
+  pre_tbl_focal <- filter(pre_tbl, SigVal == focal_sig_val) %>%
     mutate(StudyID = fct_reorder(StudyID, FWER, mean),
            ExperimentID = fct_reorder(ExperimentID, FWER, max))
 
   print(pre_tbl_focal)
-  study_id_fct = levels(pre_tbl_focal$StudyID)
+  study_id_fct <- levels(pre_tbl_focal$StudyID)
 
-  tbl = sigval_for_low_fwer(pre_tbl) %>%
+  tbl <- sigval_for_low_fwer(pre_tbl) %>%
     mutate(ExperimentID = as.factor(paste(StudyID, ExperimentID, sep = "_")),
            SigVal = as.numeric(as.character(SigVal))) 
   
   # Then plot the result.
[…]

### 2025-12-09 LK/R Code/Averaged Ordinal Values/DBDA2E-utilities.R
removed LK code, preparing for restructuring

@@ -1,559 +1,0 @@
-# Utility programs for use with the book,
-#   Kruschke, J. K. (2015). Doing Bayesian Data Analysis, Second Edition: 
-#   A Tutorial with R, JAGS, and Stan. Academic Press / Elsevier.
-# This file contains several functions that are called by other programs
-# or can be called directly by the user. To load all the functions into
-# R's working memory, at R's command line type:
-# source("DBDA2E-utilities.R")
-
-#------------------------------------------------------------------------------
-
-bookInfo = "Kruschke, J. K. (2015). Doing Bayesian Data Analysis, Second Edition:\nA Tutorial with R, JAGS, and Stan. Academic Press / Elsevier."
-bannerBreak = "\n*********************************************************************\n"
-cat(paste0(bannerBreak,bookInfo,bannerBreak,"\n"))
-
-#------------------------------------------------------------------------------
-# Check that required packages are installed:
-want = c("parallel","rjags","runjags","compute.es")
-have = want %in% rownames(installed.packages())
-if ( any(!have) ) { install.packages( want[!have] ) }
-
-# Load rjags. Assumes JAGS is already installed.
-try( library(rjags) )
-# Load runjags. Assumes JAGS is already installed.
-try( library(runjags) )
-try( runjags.options( inits.warning=FALSE , rng.warning=FALSE ) )
-
-# set default number of chains and parallelness for MCMC:
-library(parallel) # for detectCores().
-nCores = detectCores() 
-if ( !is.finite(nCores) ) { nCores = 1 } 
-if ( nCores > 4 ) { 
-  nChainsDefault = 4  # because JAGS has only 4 rng's.
-  runjagsMethodDefault = "rjparallel"
-}
-if ( nCores == 4 ) { 
-  nChainsDefault = 3  # save 1 core for other processes.
-  runjagsMethodDefault = "rjparallel"
-}
-if ( nCores < 4 ) { 
-  nChainsDefault = 3 
-  runjagsMethodDefault = "rjags" # NOT parallel
-}
-
-#------------------------------------------------------------------------------
-# Functions for opening and saving graphics that operate the same for 
[…]

### 2025-12-09 LK/R Code/Averaged Ordinal Values/OrdinalScaleGroupJags.R

@@ -1,1126 +1,0 @@
-
-# OrdinalScaleGroupJags.R
-# John Kruschke, April 2017 and February-March 2018.
-#
-# This is a lengthy script, but the vast majority of it is data manipulation (for 
-# input) and graphics manipulation (for output). The Bayesian and frequentist 
-# modeling constitutes only a small portion of the script.
-
-#------------------------------------------------------------------------------
-# Optional generic preliminaries:
-graphics.off() # This closes all of R's graphics windows.
-rm(list=ls())  # Careful! This clears all of R's memory!
-
-source("DBDA2E-utilities.R")
-# Define new function for saving multiple formats with one command:
-saveGraph2 = function( fileName , typeVec=c("png","pdf","eps") , ... ) {
-  for ( typeVal in typeVec ) {
-    saveGraph( file=fileName , type=typeVal , ... )
-  }
-}
-
-#------------------------------------------------------------------------------
-# Load/create the data file.
-# The ordered-probit JAGS model expects data to be in long format, with one
-# rating per row. There are columns for ordinal response, question ID, subject
-# ID, group ID. The responses must be integers, from 1 to K. The ID's can be
-# non-numeric, and will be converted to integer indices internally.
-# N.B.: This version of the program assumes that all items have the same number 
-# of ordinal levels.
-
-# Function for generating simulated data:
-genGroupData = function( 
-  # specify means on underlying scales:
-  mQ = c( 1.0 , 1.5 , 2.5 , 3.0 ) , # length is number of questions
-  # specify standard deviations of responses on underlying scales:
-  sdQ =  c( 1.5 , 2.0 , 2.5 , 3.0 ) , # length is number of questions
-  # specify correlations of responses on underlying scales
-  rQval = 0.9 ,
-  # specify thresholds:
-  thresh = matrix( c( 1.5 , 2.5 , 3.5 , 4.5 , 5.5 , 6.5 ,
[…]

### 2025-12-09 LK/R Code/Movie Ratings/OrderedProbitModel-Example.R

@@ -1,97 +1,0 @@
-# OrderedProbitModel-Example.R
-# John K. Kruschke, January-August 2018.
-# Sept 28, 2018: Revised by John K. Kruschke; uses new argument caseIDcolName.
-#
-# This is an example of using the function ordinalAndMetricAnalysis, defined in
-# the R script OrderedProbitModel.R. It analyzes ordinal data with ordered probit
-# and (inappropriately) metric models.
-#
-# These R scripts accompany the article, "Analyzing ordinal data with metric
-# models: What could possibly go wrong?" by Torrin M. Liddell and John K.
-# Kruschke. Files can be found at https://osf.io/53ce9/
-# 
-# To use this script, you must have in R's working directory the following three
-# files:
-# (1) OrderedProbitModel.R
-# (2) DBDA2E-utilities.R
-# (3) the data file, which in this example is MoviesData.csv. The data file
-# must be in comma-separated-value (CSV) format, with each row having the
-# frequency counts of the levels of a single group or case. The columns must be
-# named, but the names can be whatever you want. For example, the top of the
-# MovieData.csv file looks like this:
-#
-#   "ID","Descrip","n1","n2","n3","n4","n5"
-#   1,"The Whole Truth",49,70,119,217,245
-#   2,"Priceless",67,22,22,60,574
-#
-# The first row must be the column names. Each subsequent row is for a
-# particular movie. This file happens to have movie ID and Description columns,
-# but these are not used by the script. The columns named "n1" through "n5"
-# contain the frequenices observed at each ordinal level.
-#
-# See important comments at the top of OrderedProbitModel.R for more information
-# about the function.
-#
-#-----------------------------------------------------------------------------------
-# Optional housekeeping:
-graphics.off() # This closes all of R's graphics windows.
-rm(list=ls())  # Careful! This clears all of R's memory!
-#-----------------------------------------------------------------------------------
-
-source("OrderedProbitModel.R")
-
-# Example with the movies data.
-OrdModelResults = ordinalAndMetricAnalysis( 
-  dataFileName = "MoviesData.csv" ,
-  yColNames = c("n1","n2","n3","n4","n5") , # column names in data file
-  caseIDcolName = "Descrip" ,
-  compareCases = list( c(34,4) , # a case of inverted means
-                       c(5,6) , # another case of inverted means
[…]

### 2025-12-09 LK/R Code/Movie Ratings/OrderedProbitModel.R

@@ -1,923 +1,0 @@
-####################################################################################
-#
-# OrderedProbitModel.R
-# John K. Kruschke, January-August 2018. 
-# Sept 18, 2018: Revised by John K. Kruschke; graphical output revised.
-# Sept 28, 2018: Revised by John K. Kruschke; graphical output revised more, and
-#                argument caseIDcolName added.
-#
-### Description:
-#
-# This R script defines a function, ordinalAndMetricAnalysis(), that analyzes
-# groups of ordinal data with an ordered probit model and an inappropriate
-# metric model.
-#
-# The function is used for an example in the article, "Analyzing ordinal data
-# with metric models: What could possibly go wrong?" by Torrin M. Liddell and
-# John K. Kruschke. Files can be found at https://osf.io/53ce9/
-#
-### Usage:
-#
-# source("OrderedProbitModel.R") # source this file to load the function
-#
-# ordinalResults = ordinalAndMetricAnalysis( dataFileName , yColNames ,
-#                                            caseIDcolName=NULL ,
-#                                            compareCases=NULL , 
-#                                            graphFileType="pdf" ,
-#                                            hierarchSD=FALSE )
-#
-# For a detailed example of using this function, see the script
-# OrderedProbitModel-Example.R
-#
-### Arguments:
-#
-# dataFileName: The name of the data file. Assumed to be comma separated values.
-#               See below for more details.
-#
-# yColNames: A vector of column names of the frequencies of ordinal levels.
-#
-# caseIDcolName: Column name of row identifiers or labels. Optional, defaults to
-#                NULL.
-#
-# compareCases: A list of case pairs to compare. Optional, defaults to NULL.
-#
-# graphFileType: The graphics format of saved graphs. Defaults to "pdf".
-#
-# hierarchSD: Whether or not to put hierarchical structure on the group standard
-#             deviations. Logical TRUE/FALSE, defaults to FALSE.
-#
-### Output (Value):
-#
-# A list of two components:
-#
-# OrdSummaryMat: A matrix, with rows being the parameters in the ordinal model
[…]

### 2025-12-09 LK/R Code/polrToOrdScale.R

@@ -1,23 +1,0 @@
-# John K. Kruschke, 2014-Nov-21.
-# This function is described at the blog post
-# http://doingbayesiandataanalysis.blogspot.com/2014/11/ordinal-probit-regression-transforming.html
-# A PDF copy of the blog post is available at
-# https://osf.io/fc6zd/
-# This function comes with no warranties, guarantees, etc. Use at your own risk.
-polrToOrdScale = function( polrObject ) {
-  polrThresh = polrObject$zeta
-  polrSlopes = polrObject$coefficients
-  polrInter = 0.0
-  polrSigma = 1.0
-  K = length(polrThresh) + 1  # K is number of ordinal levels
-  sigmaMult = unname( (K-2)/(polrThresh[K-1]-polrThresh[1]) )
-  inter = unname( 1.5 - ( sigmaMult*polrThresh[1] ) )
-  respThresh = sigmaMult*polrThresh + inter
-  respSigma = sigmaMult*polrSigma
-  respB0 = sigmaMult*polrInter + inter
-  respSlopes = sigmaMult*polrSlopes
-  return( list( sigma=respSigma ,
-                b0=respB0 ,
-                coefficients=respSlopes ,
-                zeta=respThresh ) )
-}

### 2025-12-10 index.qmd
restructured

--- 2025-12-09 GroupPolarizationStatmod/code_examples.qmd
@@
 
-Change directories to go to the `gp-statmod/GroupPolarizationStatmod` directory where this document lives:
+Change directories to go to the `gp-statmod/docs` directory where this document lives:
 
 ```         
-cd gp-statmod/GroupPolarizationStatmod
+cd gp-statmod/docs
 ```
@@
 Then you can run the examples in this document, edit them, etc., on your own computer.
+
+This tutorial is still valuable if you're just reading it on the [code documentation associated with the journal article](https://mt-digital.github.io/gp-statmod).
 
@@
 
+I simulate experimental group polarization data for a given time, $t$, with latent group opinion mean and standard deviation, $\mu_t$ and $\sigma_t$, and with a survey response scale defined by $B$ bin values, $b_i$, with $i=1,\ldots,B$.
+
 ## Simulate each experimental design's false positive rate
@@
 is being called with the default metascience parameters with a base rate of $b=0.1$ and a
-statistical power of $W=0.8$. We consider three effect sizes in this paper according to
-the : "small", "medium", and "large" effect sizes traditionally assumed practically
+statistical power of $W=0.8$. We consider three effect-detection thresholds in the paper: "small", "medium", and "large" effect sizes traditionally assumed practically
 meaningful across social sciences [@CohenBook1988].
-I put this function in the `scripts` directory because to run the full analyses presented in the paper it's only necessary to run `calc_fdr_vs_significance` since it creates a column for the false positive rate and the estimated false detection rate calculated for each experimental design we studied.
 
@@
 
-I again call the function with the defaults specified, for illustration. Remember the `significance_vals` map to _effect-detection thresholds_ ($d^*$) in the paper.
+I put this function in the `scripts` directory because to run the full analyses presented in the paper it's only necessary to run `calc_fdr_vs_significance` since it creates a column for the false positive rate and the estimated false detection rate calculated for each experimental design we studied.
+
+Below I call the function with the defaults specified, for illustration. Remember the `significance_vals` map to _effect-detection thresholds_ ($d^*$) in the paper.
 
@@
 
+# This 
 fdr_over_thresholds <- calc_fdr_vs_significance(

### 2025-12-10 paper/method-old.tex
restructured

\section{Method}\label{method}

Our method combines several formal and computational tools to identify which
replications should be retracted and evaluate experimental designs across our
corpus. The first step in the analysis is to find a reasonable pair of latent
null-polarization distributions that appear polarized due to clipping
(Section~\ref{subsec:identify}). To do this, we develop a formal model and a numeric
rootfinding algorithm to power a Shiny web app we used to interactively identify and
record null-polarization parameters generating such measurement artifacts.

The second step in the analysis is to evaluate the quality of the experimental designs
when we properly account for the ordinal measurement procedure
(Section~\ref{subsec:evaluate}). We use the null-polarization parameters we found in
the first step to generate simulated experimental data, then estimate the difference
between the pre- and post-discussion means using ordered probit models to fit each
distribution—ordered probit models account for the ordinal measurement procedure.  We
estimate the false positive rate and false detection rate for low, medium, and high
significance thresholds in terms of Cohen's $d$. Finally, we use the effect sizes thus
calculated to determine the significance threshold one must use to achieve a 5\% false
positive rate.

\subsection{Identify replications for removal}
\label{subsec:identify}

The first step is to find which replications of the 57 must be retracted due to
non-identifiability. We establish non-identifiability by finding one set of latent
null-polarization opinion distributions where the mean doesn't change, but appear
polarized when measured on the ordinal scale used in the experiment.  In other words,
if we find a null-polarization model that explains the outcome as well as a group
polarization one, the group polarization hypothesis is \emph{non-identifiable} and the
replication must be rejected—the data cannot confirm or disconfirm the target
hypothesis.

To do this we designed a web app that allows the user to try different latent
distribution parameters (Figure~\ref{fig:data-model-flow}) to seed a
hillclimbing algorithm that explores latent distribution parameter space
until it finds null parameters that look like group polarization under the
experiment's ordinal measurement procedure, or the algorithm stops after a
certain number of steps (Algorithm~\ref{alg:hillclimbing}). If the web app
finds a constant latent mean and pre- and post-discussion 
variances to specify a pre- and post-discussion opinion distributions that
generate spurious group polarization, this is communicated to the user who
can then mark the experimental condition as such.

To find examples of plausibly spurious group polarization we guess at what latent
mean, combined with a latent pre- and post-deliberation latent standard
deviations, could generate it. We calculate the observed group polarization
for these latent parameters by transforming the latent pre- and
post-deliberation distributions from continuous normal distributions to
ordinal ``observations'' by integrating over each ordinal bin, then
calculating the observed mean for each experiment \emph{phase}, i.e., both
the pre- and post-deliberation phases.  If the null model can generate group
polarization in this case, then the detection is labelled \emph{plausibly
false}, i.e., \emph{plausibly spurious}, and we call the generating latent
parameters the \emph{plausible null parameters}.

To evaluate whether the plausible null parameters we identify could
realistically fool a significance test, we implemented a generative
simulation of each plausibly false detection of group polarization and
calculated the \emph{family-wise error rate (FWER)} and the \emph{false
discovery rate (FDR)}. We use the published number of participants $N$ for
each of the 54 conditions that yielded plausibly false detections of group
polarization. The FWER is also known as the Type I error, or the false
positive rate: it is the probability that a detection of an effect occurs
given that in reality there was a null effect.  The FDR is the probability
that a published positive detection of an effect is a false positive, i.e., a
detection of a null effect. High FWER and FDR for an experimental condition
indicates that its design is weak, since we identify only one of perhaps
infinitely many latent parameter combinations that can generate spurious
group polarization.

We use this setup to demonstrate how \emph{adversarial} style testing of
experimental designs with counterfactual null models can be used to estimate
important metascience parameters. In this study we use our FWER calculations
to estimate the minimum Cohen's $d$ that should be used to determine whether
an effect is ``significant'' or not that will limit the FWER to 5\%. In the
Discussion we sketch how our model can then be extended to rigorously
calculate statistical power, i.e., the probability that a significant effect
is detected given that there is indeed a significant effect.

\input{method-tables.tex}

\subsection{Probability theory of spurious group polarization}

% We formalize this experimental design in the $N\to\infty$ case where the
% distribution of ordinal measurements is created by integrating continuous latent
% opinion distributions.

Latent opinions are assumed to be normally distributed in general since they are
aggregations of many neurobiological potentials of
different polarities and magnitudes. We have two opinion distributions for 
two different times in the group polarization experiment process: pre-discussion,
$\tpre$, and post-discussion, $\tpost$.
The 
mean $\mu_t$ and variance $\sigma_t$ for time $t$. We are concerned with only two
times, \emph{pre-} and \emph{post-} deliberation ($t=0$ and $t=T$, respectively).
For short we can write $\mu_\mathrm{pre}$ and $\sigma_\mathrm{pre}$ for
pre-deliberation mean and variance (and similar for \emph{post}-deliberation).  
The latent opinion of participant $i$ at time $t$ is drawn from a normal 
distribution in the model, 
\begin{equation} 
  \omega_{i,t} \sim p(\omega; \mu_t, \sigma_t) = \mathcal{N}(\mu_t, \sigma_t).
  \label{eq:opinionDistribution} 
\end{equation} 
\noindent

The measurement process with $N\to\infty$ participants is simulated by 
integrating the latent opinion distribution between ordinal bin thresholds, 
of which there are $B+1$, where $B$ is the number of opinion bins, with sequential
integer values $b\{0, B\}$. In the Moscovici and Zavalloni (1969)
article, the thresholds are $\theta_b \in \{-\infty,-2.5,-1.5,\ldots,2.5,\infty\}$, with $B = 7$.
Bin $b$ is defined by its lower and upper continuous-valued thresholds, 
% $\theta_{b-1}$ and
$\theta_{b}$, respectively. Thresholds are given by the following equations
(visualized as dotted lines in Figure~\ref{fig:distros}).
\begin{equation}
\theta_b = \begin{cases}
  -\infty         & \text{ if } b=0 \\
  \infty          & \text{ if } b=B \\
  \theta_{\text{Base}} + \frac{b}{2} & \text{ otherwise.}
\end{cases}
\end{equation}
\noindent
In the Moscovici and Zavalloni example, the base bin value is
$\theta_{\text{Base}} = -3$.
With thresholds defined, the frequency of ordinal measurment $b$ at time $t$ 
is (Figure~\ref{fig:distros})
\begin{equation}
  p(b;~\mu_t, \sigma_t, B) = \int_{\theta_{b-1}}^{\theta_b} p(\omega; \mu_t, \sigma_t)
  d\omega.
  \label{eq:ordinal-frequency}
\end{equation}

We can now define spurious group polarization and give examples of when it
occurs. First, the expected value of observed opinions at time $t$ is
\begin{equation}
  \meanobst = \frac{1}{B} \sum_{b=1}^B b \cdot p(b; \mu_t, \sigma_t, B).
  \label{eq:meanobs}
\end{equation}
\noindent
The observed group polarization opinion shift is then
\begin{equation}
  g = \meanobs_\mathrm{post} - \meanobs_\mathrm{pre}.
  \label{eq:group-polarization}
\end{equation}
\noindent

Measurement artifacts can make null-polarization spuriously appear like group
polarization. To see how, we will build up a formal model of the
process. Consider a collection of
null-polarization latent opinion distribution parameters $\nu = (\mu, \sigma_{pre}, \sigma_{post})$.
The group polarization opinion shift can then be written as a function of
$\nu$,
$g(\nu)$.

If we can find ..... $\nu$ for which $g(\nu) = g_{pub} > 0$, where
$g_{pub}$ is the published group polarization shift value.
This could be thought of as sampling the observed ordinal distributions 
$N \to \infty$, collecting ordinal measurements of samples from 
each latent distribution corresponding to each phase, pre- and post-deliberation.

Next, we explain how to find a set of latent null parameters, $\nu$, that
generate spurious group polarization when naively measured on an ordinal scale.

\begin{figure}[ht]
  \caption{\textbf{Opinion measurement model and spurious group polarization.} 
  Latent model participant opinions are drawn from a normal distribution with the
  listed parameters (A,C). The pre-deliberation latent distribution has high
  variance, but appears polarized when measured via integration (A).  The latent
  distribution's variance is expected to decrease during deliberation (B) via
  consensus, but in this example the latent mean is static, 
$\mupre = \mupost$. However, when the post-deliberation distribution is measured,
the simulated observed mean increases, i.e., spurious
group polarization occurs (C).}
  \centering
  \includegraphics[width=0.95\textwidth]{Figures/Model/latent-ordinal-distros.pdf}
  \label{fig:distros}
\end{figure}

\subsubsection{Interactive algorithm to find parameters inducing spurious group polarization}

To identify plausibly false detections of group polarization, we searched numerically
for latent parameters $\nu = (\mu, \sigmapre, and \sigmapost)$ that generate
spurious group polarization, as explained immediately above. We developed an *ad hoc*
hillclimbing-style algorithm that finds $\nu$ that finds $\nu$ by minimizing
the difference between published group polarization shifts $g_pub$ and calculated shifts,
$g_\nu$.
calculated with Equation~\ref{eq:meanobs}.

The algorithm to find $\nu$ holds $\mu$ constant while searching for $\sigmapre$
and $\sigmapost$ that generate observed $\meanobspre$ and $\meanobspost$, within a
tolerance of $\pm10^{-3}$ (Algorithm~\ref{alg:hillclimbing}). 
The algorithm requires a researcher to specify an initial
guess for $\nu$ and auxiliary parameter settings, which generally vary between
experimental conditions.  Furthermore, the algorithm does not always produce
reasonable distributions with a given $\nu$, and sometimes it does not produce
spurious group polarization.  Therefore, it is necessary to sanity check the latent
and ordinal opinion distributions for the $\nu$ found by the algorithm. If all is
well, the experimental condition can be marked as plausibly spurious.  To tame the
complexity involved in data entry and model fitting for each of these sixty
experimental conditions, we developed a web application to input data from control
the algorithm and inspect its results (Figure~\ref{fig:webapp}). Web application code is
available online in our GitHub
repository\footnote{\url{https://github.com/mt-digital/gp-statmod/blob/main/GroupPolarizationStatmod/app.R}}.

\begin{algorithm}
  \caption{Routine to find latent mean and variance that generate spurious group polarization.}
  \label{alg:hillclimbing}
  \begin{algorithmic}

    \Require $\mu$, $\sigma$, $\bar{\mathbf{o}}$ 
      \Comment{Set latent mean, guess variance that yields, input empirical mean.}

    \Require $\epsilon$, $\delta$, $i_\mathrm{max}$ 
      \Comment{Hillclimbing tolerance, step size, and maximum iterations.}

    \State $\Delta \gets \meanobs(\mu, \sigma) - \meanobsemp$ 
      \Comment{Initialize simulation error.}
    \While {$\Delta^2 > \epsilon$ and $i < i_\mathrm{max}$} \Comment{Search until desired
    accuracy or max iterations reached.}
      \State $\sigma' \sim \mathcal{N}(\sigma, \delta)$ \Comment{Draw new variance to test.}
      \State $\Delta' \gets \bar{o}(\mu, \sigma') - \bar{\mathbf{o}}$ 
        \Comment{Calculate error using new variance.}
      \If {$\Delta' - \Delta < 0$}
        \State $\sigma \gets \sigma'$
          \Comment{Update $\sigma$ if the error is less than before.}
        \State $\Delta \gets \bar{o}(\mu, \sigma) - \bar{\mathbf{o}}$
          \Comment{Update simulation error if $\sigma$ is updated.}
      \EndIf
    \EndWhile \\

    \Return $\sigma$
  \end{algorithmic}
\end{algorithm}

Note that so far, we have only identified cases where spurious group
polarization plausibly occurs in the asymptotic case where $N\to\infty$ ($N$
being the number of participants).  Experimental conditions yielding
plausibly spurious group polarization warrant some suspicion. If there exists
set of latent simple consensus parameters $\nu$ that create binned
(Equation~\ref{eq:ordinal-frequency}) opinion distributions that naively seem
to be group polarized, then the means of at least some samples from these
distibutions with empirical numbers of participants would also appear to be
shifted consistent with group polarization. This process itself does not tell
us the false detection rate.  We use the identified simple consensus
parameters to define pre- and post-deliberation latent opinion distributions
in simulated trials of group polarization experiments.

\begin{figure}
  \centering
    % \includegraphics[width=0.8\textwidth]{Figures/Model/data-model-flow.png}
  \includegraphics[width=\textwidth]{Figures/Model/SoftwareMap_Mashup.png}
  \caption{\textbf{Simulation and analysis pipeline}
  for finding when measurement artifacts.  
  Arrows show the flow of input and output data from parameter estimation.
  First, we used the web app (A) to find null-polarization parameters for 
  each experimental condition, $\nu_e$,
  that ordinal measurements distort to look like polarization
  (B—described in Algorithm~\ref{alg:hillclimbing}).
  We fit ordered probit models to simulated null-polarization outcomes,
  seeded by the $\nu$ found in the previous step (C). Analysis and 
  visualization routines finish the pipeline (D).}
  \label{fig:data-model-flow}
\end{figure}

We calculate via simulation and data fitting
the Type I error rate for the experimental designs used in experimental conditions
that, under a preliminary large-$N$ model analysis, generated group
polarization upon ordinal measurement of simple consensus latent opinion 
distributions. These can then be aggregated to estimate the false discovery rate
for each journal article and for the ten-article corpus, which is representative of the
false discovery rate for group polarization studies.

The false detection rate is the scaled probability that a detection, $D$, of
an effect, $E$, is in fact a false detection and there is \emph{not} an effect,
denoted $\neg E$. The probability that a detection $D$ has occurred when there
was no effect, i.e. when $\neg E$, is denoted $\alpha = \Pr(D | \neg E)$.
This is known as the Type I or family-wise error rate. To calculate the 
false detection rate ($\fdr$), one must compare $\alpha$ to the statistical power,
the probability of a detection when there is an effect, denoted $W = \Pr(D | E)$.
Both these values must be scaled by the rate at which effects occur, denoted $b$.
Using the notation outlined above, the false detection rate is
\begin{equation}
  \fdr = \frac{(1 - b)\alpha}{(1 - b)\alpha + bW}.
  \label{eq:fdr}
\end{equation}

In group polarization, $E$ represents the group polarization \emph{effect}, and
$\neg E$ represents its absence. Simple consensus is one way for group polarization
to be absent, i.e., simple consensus is one form of $\neg E$. The presence of
an effect $E$ is independent of whether an effect was \emph{detected}, 
represented by $D$ if it was detected and $\neg D$ if not.

The false discovery rate is then the Type I error rate relative to the rate of
positive discoveries, true or false.
The group polarization studies are journal articles that present
the results of multiple experiments to support some hypothesis. To organize
the presentation of false detection rates across the ten studies in our
analysis, we first will organize false detection rates by experiment, where
we created `ExperimentID` tags for each of the sixty experiments across ten
studies. We also created `StudyID` tags for each of the ten journal articles
with the key `{Author}{Year}`, e.g., \texttt{Schkade2010} for Schkade, et al.,
(2010) \cite{Schkade2010}.

We calculate $\alpha = \Pr(D | \neg E)$ through simulation for experimental
conditions, 
and use optimistic values of
$b=0.1$~\cite{Ioannidis2005,McElreath2015} and $W=0.8$~\cite{Smaldino2016NatSel}
in the main text and in the compare with more pessimistic, yet realistic, estimates
of these values, as well as even more optimistic estimates.

\subsubsection{Estimating and limiting false detection rates}

To calculate the false detection rate we begin by simulating 
pre- and post-deliberation 
opinions by drawing $N_e$ samples from normal latent opinion
distributions that generated spurious group polarization, i.e., with
parameters labelled $\nu_e$ found in the hillclimbing step. Binning these
random values simulates observations. $N_e$ is the number
of participants in experimental condition $e$. Each trial $i$ of $N_e$
simulated pre- and post-deliberation observations is then inspected to see
if spurious group polarization occurred in the simulation. Spurious group
polarization is determined by first inferring the latent parameters of each
simulated experimental trial by fitting an ordered probit model~\cite{Liddell2018}. 
The inferred parameters are then used to calculate the effect size for that
trial, Cohen's $d$, defined as~\cite[p. 331]{Liddell2018}
\begin{equation}
  d = \frac{\mupost - \mupre}{\sqrt{\frac{1}{2}(\sigmapost^2 + \sigmapre^2)}}.
  \label{eq:cohens}
\end{equation}
\noindent
The family-wise error rate is, $\alpha_{e} = \Pr(D|\neg E)$, is the 
fraction of trials, 
with $d_{ei} > d^*$ for the $i^\text{th}$ trial for experimental
condition $e$, written
\begin{equation}
  \alpha_e = \frac{1}{N_T} \sum_i^{N_T} \Theta(d_{ei}; d^*),
  \label{eq:study_aggregate}
\end{equation}
\noindent
where
\[
  \Theta(d_{ei}; d^*) = 
    \begin{cases}
      1 & \text{ if } d_{ei} > d^* \\
      0 & \text{ otherwise.}
    \end{cases}
\]
and $N_T = 1000$ is the number of trials. For simplicity we assume that the sign of
$d_{ei}$ is positive, meaning a shift to greater extremisim. It is possible that
$d_{ei}$ could be negative, indicating a shift to lesser extremity. Indeed we
observe many simulations with a shift to lesser extremity. For now we do not
include these as indicating a ``significant'' effect, which we return to in the
Analysis below.

To calculate the family-wise error rate
for a study, $s$ which contains several experimental conditions, $e$,
\begin{equation}
  \alpha_s = \frac{1}{|s|} \sum_{e \in s} \alpha_e
\end{equation}
\noindent
The mean across all results identified as plausibly spurious is 
\begin{equation}
  \alpha_\mathrm{all} = \frac{1}{|E|} \sum_{e \in E} \alpha_e.
\end{equation}
We finally may calculate the false discovery rate by plugging this $\alpha$
into Equation~\ref{eq:fdr} and assuming empirically-motivated values for 
base rate, $b$, and statistical power, $W$, as well as more pessimistic and
optimistic settings for base rate and power.

We close our Analysis by inverting this procedure to find what significance value
$d^*$ is necessary to achieve a low family-wise error rate $\alpha_e = 0.05$ for
experimental condition, $e$. To do this we first calculate $\alpha_e(d^*)$ for a
sufficiently wide range of $d^*$. We then find the lowest $d^*$ for which
$\alpha_e(d^*) \leq 0.05$.

### 2025-12-10 paper/ms.tex
restructured

% ==============================================================================
% CLEANED ms.tex (Optimized for XeLaTeX/Fira Fonts)
% - Consolidated preamble and ms.tex.
% - Removed all conflicting packages (lmodern, redundant font settings).
% - Uses Fira Sans/Mono/Math consistently.
% ==============================================================================

\documentclass[preview]{article}

% --- CORE FONT SETUP (Sans-Serif Headings, Serif Body, Professional Math) ---
\usepackage{fontspec}        
% Use Path=fonts/ to look inside the fonts folder
\setmainfont{FiraSans.ttf}[
    Path = fonts/,
    UprightFont = *-Regular,
    BoldFont = *-Bold,
    ItalicFont = *-Italic,
    BoldItalicFont = *-BoldItalic,
    Extension = .ttf
]
\newfontfamily{\sansheadings}{Fira Sans}
\usepackage{sectsty}
\sectionfont{\sansheadings \raggedright \Large}
\subsectionfont{\sansheadings \raggedright \large}
\subsubsectionfont{\sansheadings \raggedright \normalsize}

\usepackage{unicode-math}    
%
%% 1. Main Text: Set to Latin Modern Roman (Stable, Professional Serif)
%\setmainfont{Latin Modern Roman} 
%
% 2. Monospaced/Code Text: Fira Mono
%\setmonofont{Fira Mono} 
%
%% 3. Mathematical Notation: Latin Modern Math (Stable, Conventional Serif Math)
%\setmathfont{Latin Modern Math}

% 4. Heading Font Setup (Fira Sans)

% \usepackage{sectsty}
% \sectionfont{\sansheadings \raggedright \Large}
% \subsectionfont{\sansheadings \raggedright \large}
% \subsubsectionfont{\sansheadings \raggedright \normalsize}
% Ensure your title command uses \sansheadings as defined in the previous step.
% ...

% --- ESSENTIAL PACKAGES ---

% 2. Core Text Utilities and Geometry
\usepackage[margin=1.25in,letterpaper]{geometry} % Defines page layout
\usepackage[]{microtype}    % High-quality font protrusion and spacing
\usepackage{upquote}        % Straight quotes in verbatim/code
\usepackage{calc}           % Utility for calculations

% 3. Graphics, Color, and Table Fundamentals
\usepackage[table,dvipsnames,svgnames,x11names]{xcolor} % Color definitions
\usepackage{graphicx}       % For including graphics
\usepackage{booktabs}       % Professional quality tables
\usepackage{longtable}      % Multi-page tables
\usepackage{array}          % Enhanced array and column definitions
\usepackage{multirow}       % Merging rows in tables
\usepackage{makecell}       % Formatting cells in tables

% 4. Caption Handling (Load basic caption definitions BEFORE floatrow)
\usepackage{caption}        
\usepackage{subcaption}     
\usepackage{floatrow}       % Complex float/caption placement and formatting

% 5. Math Environments and Algorithm Structure
\usepackage{amsmath}        % Essential math environments (align, gather, etc.)
\usepackage{algorithm}
\usepackage{algpseudocode}

% 6. Hyperlinks and References
\usepackage{hyperref}       % Hyperlinks and PDF metadata (must be loaded relatively late)
\usepackage{bookmark}       % PDF bookmark control
\usepackage{xurl}           % Better URL line breaks
\usepackage{footnotehyper}  % Footnotes integration with hyperref

% 7. Custom Utilities (Needed for your document structure)
\usepackage{iftex}          
\usepackage{authblk}        % Author affiliations
\usepackage{etoolbox}       % Utility commands for patching/macros
\usepackage{wrapfig}        % Text wrapping around figures

% --- End of ESSENTIAL PACKAGES ---

% --- BIBLIOGRAPHY CONFIGURATION ---
% Use the biblatex settings you had in preamble.tex
\usepackage[sorting=none, style=numeric-comp]{biblatex}
\addbibresource{/Users/mt/workspace/Writing/library.bib}
%\addbibresource{this.bib}

% --- FLOATROW / TABLE SETUP (From ms.tex) ---
\DeclareFloatFont{small}{\small}
\floatsetup[table]{font=small}

% --- HYPERREF SETUP (From ms.tex) ---
\hypersetup{
  pdftitle={Group polarization replications may be marred by high false discovery rates},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue}}

% --- CUSTOM COMMANDS (From ms.tex) ---
\newcommand{\mupre}{\mu_\mathrm{pre}}
\newcommand{\mupost}{\mu_\mathrm{post}}
\newcommand{\sigmapre}{\sigma_\mathrm{pre}}
\newcommand{\sigmapost}{\sigma_\mathrm{post}}
\newcommand{\thetafit}{\theta^\mathrm{fit}}
\newcommand{\fdr}{\mathrm{FDR}}
\newcommand{\meanobs}{\bar{o}}
\newcommand{\meanobsemp}{\bar{\mathbf{o}}}
\newcommand{\meanobst}{\meanobs_t}
\newcommand{\meanobsjt}{\meanobs_{j,t}}
\newcommand{\meanobspre}{\meanobs_\mathrm{pre}}
\newcommand{\meanobspost}{\meanobs_\mathrm{post}}
\newcommand{\tpre}{t_\mathrm{pre}}
\newcommand{\tpost}{t_\mathrm{yopost}}
\definecolor{myorange}{RGB}{240, 96, 0}
\newcommand{\mt}[1]{{\textcolor{myorange} {({\tiny MT:} #1)}}}

% --- TITLE AND AUTHOR INFO (From ms.tex) ---
\makeatletter
\title{\Large If the Null Fits, You Must Omit: Measurement Artifacts Undermine Group Polarization Replications}
\makeatother
\author[1,*]{{Matthew A.~Turner}}
\affil[1]{\small Environmental Social Sciences, Stanford Doerr School of Sustainability, Stanford University}
\author[2,3]{{Paul E.~Smaldino}}
\affil[2]{\small Cognitive and Information Sciences, University of California, Merced}
\affil[3]{\small Santa Fe Institute} 
\affil[*]{\small Correspondence: \href{mailto:maturner@stanford.edu}{maturner@stanford.edu}}
\date{\today}

% ==============================================================================
% BEGIN DOCUMENT
% ==============================================================================
\begin{document}
\maketitle

\renewcommand*\contentsname{\small Table of contents}
{
\hypersetup{linkcolor=gray}
\setcounter{tocdepth}{3}
\footnotesize
\tableofcontents
}

% Center rule to separate TOC from rest of paper
\vspace{-2em}
\begin{center} \noindent\rule{4cm}{0.4pt} \end{center}
\vspace{-2em}

% --- INPUT CHAPTERS ---
% \input{intro.tex}

\subsection*{Abstract}\label{abstract}
\addcontentsline{toc}{subsection}{Abstract}

Echo chambers seem to cause radicalization—and radicalization is risky.
Radicalization undermines institutions by reducing their diversity, 
which in turn impairs their adaptability.  Social psychologists and others study
echo chambers experimentally using \emph{group polarization} experiments.
Researchers induce \emph{group polarization} by first asking experiment
participants their opinions on some topic, then grouping them together if
they agree in spirit, but not degree. The participants then discuss the topic
in their groups. Their opinions are then re-measured. Group polarization is
when the average group opinion increases. But there’s a fatal problem:
participants report opinions on ordinal scales, but this has never been
accounted for statistically. This can cause apparent group polarization in
observations even if the average latent opinion—the one in people’s
heads—doesn’t change, due to ceiling effects. Here we prove that across ten
journal articles spanning five decades that 54 of 57 reported polarization effects are plausibly false due to this artifact, and thus cannot be counted as replications. Our method can be used to evaluate any experimental design that measures latent change with ordinal scales. Given how destructive radicalization can be, we must get this right.

\section{Introduction}\label{introduction}

If an extremist moderates their opinion and it is measured with a Likert scale, will
this move to moderation be detected? The answer is often no in social psychology 
experiments~\cite{Liddell2018}. This
means that a group can appear to become radicalized over time even if the 
average opinion does not actually change~\cite{Turner2020}. 
This has huge consequences for the study of \emph{group polarization}, the hypothesis
that discussion among like-minded people pushes opinions toward extremes—effectively 
making it the science of ``echo chambers'', online or in-person communities 
composed of likeminded folks.  Cass Sunstein, a
Harvard Law professor and former White House administrator, has written
extensively on the topic~\cite{Sunstein2019}, claiming that group
polarization is a scientific law—a regular, quantitative relationship that
will always be observed when certain assumptions are
met~\cite{Cartwright1999}. Yet Roger Brown himself noted in Social Psychology
that group polarization ``does not occur with every group'' and that ``the
effect is not large''~\cite{Brown1986}. 
Given the influence of this theory and the hundreds of claimed ``replications'' of
group polarization, it is essential to ask whether any of these replications may
instead be \emph{measurement artifacts}—spurious effects created when the
``neutral'' point of the measurement scale fails to align with the group’s own
center of opinion. The science of opinion change has a central role in designing new
methods for social organization and social change to face pressing existential
challenges~\cite{Galesic2021,Galesic2023}.

The group polarization hypothesis is that discussion within a group of
likeminded people causes the group's average opinion to increase, or
\emph{polarize}~\cite{Brown1986,Brown2000}.
Group polarization experiments therefore attempt to
polarize groups as follows: first, participants report their opinion on
some topic or prompt chosen by the experimenters in advance. Second, 
participants are put into likeminded groups to
discuss the topic. Following discussion, participants again report their
opinions on the topic. Group polarization is defined as an increase in
the average opinion following discussion. If the group polarized, and we 
assume that people in the group found consensus—meaning their opinions agreed
more after the discussion than before—then initially moderate opinions must have
increased more than more extreme ones.

We found that 95\% of experimental reports of group polarization
across sixty published experimental conditions cannot be counted as
replications. We explain here how evidence of group polarization is undermined
by a basic measurement flaw that obscures opinion change among extremists.
This opens the possibility that what seems to be group polarization could just
be \emph{mere agreement}, with no change in mean opinion, if only increases in
opinion were measured but not decreases.  To evaluate whether the empirical
record reflects genuine opinion change or merely measurement artifacts, I
simulated published experiments under conditions of mere agreement. Each
simulation reproduced the core elements of published designs—pre-discussion
survey, group interaction, post-discussion survey—while holding the latent
mean opinion constant. Because the original datasets are lost and published
results rarely provide enough information to reconstruct them, we cannot know
whether any reported effect reflects a real shift or a measurement artifact.
When distinct latent processes can generate indistinguishable results, the
effect is non-identifiable—and no claim of replication can stand.

A relatively simple step could fix the problem, we believe: measure twice
before discussion. Before forming groups, experimenters could poll opinions
using a scale centered arbitrarily at 0, with varying strengths of disagree
or agree with negative and positive integers, respectively. 
Then after likeminded groups are
selected, measure group opinions again before they discuss, but now with the
scale centered on the average group opinion. To keep the same number of bins,
add more bins to express more extreme views and take away bins from
the opposite side as necessary.

Even if this measurement problem were fixed, equally severe problems remain
for group polarization research. First there seems to be no stable 
``opinions'' per
se—it is more accurate to say people construct opinions when they are
prompted or need them for some other reason, as revealed by respondents
answering questions differently if the framing or ordering of them changes,
and by the fact that people respond differently to the same question over
time. These are just a few of several potential sources of variance that were never accounted for, which is also known to generate false detections of change. 
Finally, because group polarization experimental design was never
standardized its research corpus cannot cumulatively or systematically tell 
us anything general about the world—Dorian
Cartrwright (1973) tried to alert group polarization researchers to this
problem half a century ago~\cite{Cartwright1973}. General scientific claims
of the existence and functioning of some scientifically induced phenomenon
only work if the evidentiary observations were collected from
\emph{commensurate} experiments, or in other words, only from experiments
that share the exact same causal structure as expressed in a formal or 
mechanistic causal model~\cite{Cartwright1999}.

\subsection{Group polarization research}

Group polarization research began in the 1960s, which grew rapidly with
focused government funding coupled with widespread
interest in what was then a surprising new science of
radicalization~\cite{Cartwright1971}. Dorian Cartwright (1973) reviewed the
first decade of group polarization research and found problems: he lamented
the ``disturbing fact'' that group polarization research has produced no
cumulative theoretical undersatnding of the effects of discussion on
group opinions~\cite{Cartwright1973}. Cartwright was somewhat dismayed, but
optimistically hoped that group polarizaiton research would eventually
produce ``an enduring\ldots more comprehensive paradigm'' \cite[p.
230]{Cartwright1973}.

Cartwright (1973) suggested that a ``vigorous search for alternative
formulations'' could help group polarization become more useful. More
researchers developed specific theories of group
polarization, with two becoming particularly influential: social comparisons
and persuasive arguments. The social comparisons hypothesis says group
polarization occurs because humans want to fit in, and so generally like to . Group polarization occurs, then, when a
group is composed mostly of extremists, with a sufficient number of people
with a weak opinion \cite{Myers1970,Myers1975}. Its main
“competitor” from the 1970s was the persuasive arguments theory, which
basically claimed that rhetoric was the strongest driver of group
polarization, not social comparisons \cite{Burnstein1973,Burnstein1975}.

Other notable hypotheses include the
self-categorization and social decisions scheme hypotheses.
Self-categorization augmented the social comparisons hypothesis with a
consideration of idiosyncratic beliefs people hold that are the result of
their own personal lived experience \cite{Abrams1990,Hogg1990}.
The social decisions hypothesis posits that contextual factors like
social network structure could be at least as important as the distribution
of opinions and . Noah Friedkin, a sociologist, notably published
the study with the least potential measurement artifacts for a rigorous study on group polarization, adopting the social
decisions scheme hypothesis, predicting that social connectivity within a
group could amplify group polarization \cite{Friedkin1999}.

\subsection{Study Corpus}

We selected ten representative journal articles with sixty total experimental
conditions among them. For a representative sample of group polarization
research across decades, we two articles for three of the five article
categories: \emph{Social Comparisons}, \emph{Persuasive Arguments},
 and \emph{Politics}. We included three \emph{Self-Categorization} studies and one \emph{Social Decisions}.
All but the \emph{Politics} category refers to the theoretical tradition that the
authors were trying to prove right—the two \emph{Politics} articles were agnostic
to group psychology theories. This resulted in a representative sample research
over time, as well: we have one study from 1969,
four from the 1970s, three from the 1990s, one from 2007 and one from 2010. 
Information about our corpus is summarized below with references 
(Table~\ref{tab:articles}).

\begin{center}
\vspace{1.5em}
[ TABLE \ref{tbl:article-table} ABOUT HERE ]
\vspace{1.5em}
\end{center}

Each study used subtly different experimental methods that make it difficult
to build cumulative general knowledge. But all of the studies share one thing
in common: they all measure participant opinions on an ordinal scale. We start
with a non-technical review of how ordinal opinion measurements work in group
polarization, and then how the distributions of measurements are analyzed to
determine if group opinions became polarized. Then we can explain our
close-reading analysis to extract the necessary experimental design parameters
we need to simulate all the experiments in our corpus.

The corpus contains
sixty different experimental conditions in the ten corpus journal articles. We 
simulate experiments for all 57 experimental conditions reported to have 
replicated group polarization, out of 60 total conditions (three were
null control conditions).

Using simulations we will estimate the number of these 57 replications that
must be retracted because null-polarization simulations appear polarized due
to measurement artifacts. We collect experiment design parameters that fill out 
details of the simulation model of group polarization experiments.

\subsection{Ordinal Opinion Measurements: Problems and Simulations}

In studies of group polarization, opinion measurements are typically ordinal,
reflecting ordered positions rather than measurable distances between them—
Moscovici and Zavalloni (1969) gave participants a seven-point scale, from
“strongly disagree” (-3) to “strongly agree” (+3) to report their opinions.
Burnstein and Vinokur (1975) used a ten-point scale from 1 to 10 representing
“the lowest odds of success acceptable” to try a certain behavior, where 1
means 10\%.  In these and every other survey-based group polarization study I
know, the reported opinions on the ordinal scale—with discrete numbered
bins—were treated as if they were continuous-valued, or in other words as if
they were the internal mental opinion value itself, whatever that might be
neurobiologically.

Note that ordinal scales clip all latent opinions with a magnitude greater
than the magnitude of its extreme values. This can make null-polarization
outcomes look like group polarization. Consider the following example 
with a group of just two people to see why: say they start with latent
opinions of 2.1 and 4.7, which they report as +2 and +3, let's say, on Moscovici
and Zavalloni's scale that tops out at +3. If they then discuss the topic and their
opinions move toward one another by 1 each, then their latent opinions would be 3.1
and 3.7. Before and after discussion the latent mean was a constant 3.4. 
When reported on an ordinal scale, both opinions are now +3—the average measured
opinion \emph{spuriously increased} from 2.75 to 3.0!

To determine whether a replication must be retracted, we developed a model to
simulate measurements of null-polarizations, labelled \emph{mere agreement}, where
the mean group opinion stays constant, but variance decreases. In other words,
simulated participants report more agreement following discussion, 
but they do not become radicalized.

The simulations are seeded by actual group polarization experiment parameters.
We copied the pre- and post-discussion opinion means reported in each article.
There is no original data available for any of the experiments. Sometimes more
information was available about the pre- and post-discussion
distributions—we used this information to decide whether replications plausibly 
have null-polarization. We recorded all available data with a guide to finding it in
its original article in the file `StudiesAnalysis.csv`, available with the
Supplemental Information or to view/download on
GitHub\footnote{\url{https://github.com/mt-digital/gp-statmod/GroupPolarizationStatmod/data/StudiesAnalysis.csv}}.

\subsection{Research Overview}

We collected original data across 10 journal articles to show that 95\% of 
the replications reported must be retracted due to non-identifiability with a
null-polarization model. These replications cannot be salvaged because the
original data is no longer available, and thus cannot be re-analyzed with a 
better statistical model that properly accounts for the measurement procedure.

In the Analysis we go further: we use simulations to show that 
the experimental designs in our corpus are beset by high false positive rates 
\emph{even with a better statistical model}. The only way to reduce false
positive rates with the experimental designs used in the papers is to increase
the significance threshold (in terms of Cohen's $d$ as explained in more
detail in the Methods).

Group polarization was taken as one of the most reliable group effects in social
science—but we showed it is impossible to distinguish 95\% of group
polarization detections from a null-polarization model. This means 54
replications must be retracted—and recall that this was a selection of the
highest quality, most representative group polarization research. 
To understand group polarization, new
methods will be required.

This sort of error threatens our ability to construct a
cumulative science of opinion change, social influence, and social change
generally. We need reliable social science as the foundation for new
strategies for social organization capable of reversing decline and powering
sustainable futures. We have a long way to go.
\section{Method}

A group polarization replication is proven undecidable, and therefore
not truly a replication, if we can find a model of \emph{mere agreement}
that appears to be polarized under that replication's experimental
design. If we cannot find such a model, we cannot determine a
replication to be undecidable. Mere agreement in this model is a dynamic
where agreement increases within the group, but extremism does not---in
other words, A \emph{null-polarization distributions} is a set of two
Gaussian distributions: one representing the pre-discussion opinions of
the group, and the other representing the post-discussion opinions. Our
\emph{null-polarization distributions} represent the process of the
group finding \emph{mere agreement}, which is when the average group
opinion is constant over time, but opinion variance decreases to
represent agreement (or, equivalently, \emph{consensus}~(DeGroot 1974)).

\subsection{Latent Opinion Model}\label{latent-opinion-model}

Throughout this paper we model a participant \(i\)'s latent opinion at
time \(t\) as a normally-distributed random variable,

\begin{equation}\phantomsection\label{eq:normal-opinion-distribution}
{ 
    \omega_{i,t} \sim
    \mathcal{N}(\mu_t, \sigma_t),
}
\end{equation} 
with mean \(\mu_t\) and standard deviation \(\sigma_t\).
Group polarization in terms of latent opinions is then defined as any
pair of pre- and post-discussion opinion distributions where the mean
increases in magnitude over time---formally,
\(| \mu_\mathrm{post}| > | \mu_\mathrm{pre}|\). We need to take
magnitudes since radicalization may be towards increasingly negative or
positive opinion values.

For group polarization experiments, only two times are relevant:
\emph{pre-discussion}, \(t=\mathrm{pre}\), and \emph{post-discussion},
\(t=\mathrm{post}\). Since participant groups are assumed to be
likeminded, we could also assume they come to agree more over time as
well as become more extreme in our group polarization model, or in terms
of the latent opinion variance,
\(\sigma_\mathrm{pre}> \sigma_\mathrm{post}\)---but only one study in
our ten actually measured opinion variance to check this (Schkade,
Sunstein, and Hastie 2010).

A \emph{null-polarization model} is a model that is consistent with the
\emph{negation}, or \emph{null}, of the \emph{group polarization
hypothesis}. The group polarization hypothesis is that group discussion
causes the average group opinion to increase, assuming the group is
aligned on a particular issue, though opinion strength varies. Therefore
a null model is one where the latent mean stays constant, i.e.,
\(\mu = \mu_\mathrm{pre} = \mu_\mathrm{post}\).

\emph{Mere agreement} is the type of null-polarization model we use
here. Mere agreement is a null-model because the mean latent opinion is
set to be constant before and after discussion. We represent
\emph{agreement}---or equivalently, \emph{consensus}---as a decrease in
the standard deviation that defines the pre- and post-discussion latent
opinion distributions. The mere agreement model is the cornerstone of
our analysis, since it is empirically plausible and generates
\emph{spurious} observations of group polarization under mere agreement
models, due to measurement artifacts induced by clipping all latent
opinion values outside the measurement range---we call this
\emph{spurious polarization} for short. We simulate opinion measurements
in two different ways, but both use the mere agreement model of latent
opinions to test whether and by how much the measurement procedure
distorts null-polarization to look like group polarization.

%{[} TABLE
%{[}tab:latent-variables{]}(\#tab:latent-variables)\{reference-type=``ref''
%reference=``tab:latent-variables''\} ABOUT HERE {]}

\begin{centering}
[ TABLE~\ref{eq:latent-variables} about here ]
\end{centering}

\subsection{Two Models of Group Polarization
Measurement}\label{two-models-of-group-polarization-measurement}

We simulate measurement in two different ways---i.e., with two different
models---for different parts of our analysis. The first is the
\emph{expectation model}: an efficient way to \emph{simulate measurement
of the observed average} opinion of a group, using the
\emph{mathematical expectation value} of the distribution of measured
opinions as a proxy. In the second model, the \emph{stochastic model},
we simulate individual participant responses by first drawing an opinion
for each participant from the latent opinion distribution
(Equation~\ref{eq-random-variable-def}). The opinions are binned to the
appropriate ordinal value.

We use the expectation model for its efficiency in the first step of our
analysis: to find a \emph{null-polarization model} that also fits the
data for each experimental condition studied---if we find one that means
we must retract this replication since this proves that the group
polarization is undecidable with the data we have. Once we find any such
undecidable results, we use the null-polarization model to help us
further evaluate the experimental design itself.

In the \emph{stochastic model}, we simulate group polarization data for
the \emph{Evaluation} step of the analysis. We simulate experimental
data, with the same number of participants as the original experiment,
using the null-polarization model found in the \emph{Identification}
step, if one was found.

We \emph{evaluate} the experimental design, \(e\), in terms of the false
positive rate, also known as the Type I error rate or family-wise error
rate, denoted \(\alpha_e\). We close our analysis by demonstrating how
our method can be inverted to \emph{specify} experimental designs in
the pre-registration stage using our evaluation framework
to avoid preventable replication failures like the ones we identify here.

The measurement models are introduced in detail along with the
corresponding stage of the analysis that uses them. First, the
\emph{Identification} stage that uses the \emph{expectation} measurement
model; second, the \emph{Evaluation} stage that uses the
\emph{stochastic} measurement model.

\subsection{\texorpdfstring{Part 1: \emph{Identify} Invalid
Replications}{Part 1: Identify Invalid Replications}}\label{part-1-identify-invalid-replications}

In the first stage of our analysis, we try to \emph{identify} a null
polarization models, \(\nu\), that invalidate published detections of
group polarization if the measurement procedure induces \emph{spurious
group polarization} for that null polarization model. Null polarization
can take many forms---all that is required is that the latent mean
opinion stays constant before and after discussion. Here we use a
\emph{mere agreement} null polarization model, where we simulate
\emph{consensus}, the tendency for sufficiently similar opinions to
become more similar---but we leave out polarization. In mere agreement,
moderate and extreme opinions approach change by equal magnitude towards
each other. We therefore represent consensus as a decrease in the
standard deviation of the latent opinion distribution from pre- to
post-discussion. We simulate the measurement of the group's average
opinion before and after discussion and measure the spurious group
polarization, if there is any. We use this procedure in a search
algorithm we developed to identify which combination of constant latent
mean and pre- and post-discussion opinion variances generate spurious
group polarization, if there are any.

Our invalidation strategy, then, is to find a mere agreement null
polarization model, \(\nu\), that predicts the exact observations
reported in each experimental replication in our corpus. If one is
found, then we have found a model consistent with the \emph{negation} of
the group polarization hypothesis that explains the data as well as the
group polarization hypothesis itself. In other words, the hypothesis is
not proved or disproved given the data, but \emph{undecidable}: the data
cannot \emph{decide} the hypothesis one way or another because of
potential measurement artifacts. Because the experiment failed to
produce a valid replication, the results are effectively stripped of
evidentiary value and warrant exclusion from the aggregate evidence
base. We apply our analysis to every experimental condition across all
ten journal articles in our corpus, which we label \(e\). But how do we find a \(\nu\) that generates spurious group polarization?
And how do we even simulate the group polarization experiments and
measurements that make mere agreement in latent opinions look like group
polarization?

We simulate experimental measurements of group
polarization by using the mathematical \emph{expectation value} of
latent opinions defined by \(\nu\) as a proxy for the observed average
value of ordinal-valued reported opinions (described in the first
subsection below). We then use this measurement model in a numerical
root-finding algorithm that identifies a \(\nu\) that matches published
observations, if possible. If we find such a \(\nu\), this means we have
a model consistent with the null hypothesis that explains published
observations. In other words, the replication is invalidated since the
group polarization hypothesis is \emph{undecidable} given the
data---i.e., it cannot decide whether the hypothesis or its null is
correct.

So, to review, this subsection introduces how we identify which
experimental data can be explained by a mere agreement null model, where
the ordinal survey measurement procedure induces a clipping artifact
that hides changes in extreme opinions that fall outside the measurement
scale.

\subsubsection{The Expectation Model of Spurious Group
Polarization}\label{the-expectation-model-of-spurious-group-polarization}

We use the \emph{expectation model} of opinion measurement to
\emph{Identify} which, if any, mere agreement null-polarization models
generate spurious group polarization in a particular experiment, \(e\).
The trick to this model, and the source of its name, is that we use the
theoretical \emph{expected value} of opinion measurements as a proxy for
the \emph{observed average} of participant opinions under an
experimental design, \(e\).

We simulate the measurement of the mean observed opinion at time \(t\)
by first calculating integrating the normal probability density
function, \(f(\omega)\), over measurement bin boundaries, over each
bin's corresponding range of opinion values. For example, consider an
ordinal scale that runs from -3 (strongly disagree) to +3 (strongly
agree) (as used by Moscovici and Zavalloni (1969), for example). the
frequency with which the ``2'' bin is chosen is just the total area
underneath the probability density curve between 1.5 and 2.5.

All bins have a width of 1 in latent opinion space, except for the two
most extreme bins at opposite ends of the sentiment spectrum---this
works because bin values are always separated by 1. All continuous
latent opinion values \(\omega < b_1 + \frac{1}{2} = \theta_1\) get
mapped to the ordinal measurement value \(b_1\)---so, \(b_1\)'s lower
threshold is \(\theta_0 = -\infty\). Similarly, all opinions
\(\omega > b_B - 0.5 = \theta_B\) are reported as the maximum bin value,
\(b_B\), with an effective upper bound of \(\theta_{B+1} = +\infty\). We
can define all thresholds compactly as (Liddell and Kruschke 2018)
\begin{equation}\phantomsection\label{eq-bin-thresholds}{
\theta_j = \begin{cases}
  -\infty         & \text{ if } j=0 \\
  \infty          & \text{ if } j=B \\
  b_j + \frac{1}{2} & \text{ otherwise.}
\end{cases}
}\end{equation}

The probability density that a participant reports the opinion in bin
index \(i\) at time \(t\) is \[
f_{it}(\mu, \sigma_t) = \int_{\theta_{i - 1}}^{\theta_{i}} f(\omega; \mu, \sigma_t)~d\omega,
\] which we calculated in R to evaluate the integral at the two
threshold values \(\theta_{i-1}\) and \(\theta_i\). Note that the normal
probability density function we used is parameterized by mean \(\mu\)
and standard deviation \(\sigma_t\). \[
f(\omega; \mu, \sigma_t) = 
    \frac{1}{\sigma_t \sqrt{2 \pi}} e^{-\frac{(\omega - \mu)^2}{2 \sigma_t^2}}.
\]

To simulate the measurement of the \emph{average} observed opinion at
time \(t\), \(\bar{o}_t\), as the expected value of the probability
density function of observed ordinal opinions. In other words, we
simulate \(\bar{o}_t\) by calculating the weighted sum of products
between the probability of observing each bin value and the bin value
itself: \begin{equation}\phantomsection\label{eq-sim-ave-observed}{
\bar{o}_t(\nu, e) = \sum_{i=1}^B b_i f_{it}(\nu, e).
}\end{equation}

Group polarization is the difference between the post-deliberation mean
observed opinion and the pre-deliberation one. We will calculate the
magnitude of spurious group polarization under a mere agreement null
model represented by its collected distribution parameters,
\(\nu = (\mu, \sigma_\mathrm{pre}, \sigma_\mathrm{post})\), under
experimental design, \(e\):
\begin{equation}\phantomsection\label{eq-spurious-gp}{
g(\nu, e) = |\bar{o}_\mathrm{post}(\nu, e)| - |\bar{o}_\mathrm{pre}(\nu, e)|
}\end{equation}

In the \emph{Inspect} stage of our analysis we will determine which
group polarization replications are invalidated due to equifinality
between the \emph{mere agreement null-polarization hypothesis} on the
one hand, and the \emph{group polarization hypothesis} on the other.
This requires more than just showing there is some \(\nu\) that causes
spurious polarization for some experiment where \(g(\nu, e) > 0\). We
have to find a \(\nu\) that generates the exact \(g(\nu, e)\) reported
in the original journal article \emph{and} check that simulated
\(\bar{o}_\mathrm{pre}\) and \(\bar{o}_\mathrm{pre}\) match published
values.

{[} TABLE
\hyperref[tab:experiment-parameters]{{[}tab:experiment-parameters{]}}
ABOUT HERE {]}

{[} TABLE
\hyperref[tab:simulated-observations]{{[}tab:simulated-observations{]}}
ABOUT HERE {]}

\subsubsection{Example Invalidation of a Replication in Moscovici and Zavalloni
(1969)}\label{subsubsec:example-invalidation}

One of the replications we invalidate here is from Moscovici and
Zavalloni (1969), the ``Americans'' condition as they called it---we
coded it \texttt{Moscovici1969\_Americans}. They report that participant
opinions polarized after discussing the survey topic, ``American
economic aid is always used for political pressure.'' They report survey
responses increased in extremity from an average of -0.9 before
discussion to -1.6 after, on a seven-bin scale from -3 indicating
``strongly disagree'' to +3 indicating ``strongly agree''.

Moscovici and Zavalloni say their results confirm the basic hypothesis,
that \emph{group discussion causes polarization}. But we found a mere
agreement null-polarization model that generates spurious group
polarization using our model---specifically,
\(\nu = (\mu = -1.2, \sigma_\mathrm{pre}= 4.3, and \sigma_\mathrm{post}= 1.9)\).
This means that the hypothesis is actually undecidable given the data,
invalidating Moscovici and Zavalloni's replication. When a hypothesis
and its null explain data equally well, the data tell us nothing.

Now let's use the expectation model to prove that this \(\nu\) generates
spurious group polarization in this experiment, i.e.,
\(g(\nu, e= \mathrm{Moscovici1969\_Americans}) > 0\). In the process we
demonstrate how the expectation model works to simulate the measurement
of spurious group polarization (Equation~\ref{eq-spurious-gp}).

First, we calculate the probability density vector using the
\href{https://github.com/mt-digital/gp-statmod/blob/main/GroupPolarizationStatmod/model.R\#L44}{\texttt{makeProbVec}}
function in the file
\href{https://github.com/mt-digital/gp-statmod/blob/main/GroupPolarizationStatmod/model.R}{\texttt{GroupPolarizationStatmod/model.R}}
in the \href{https://github.com/mt-digital/gp-statmod}{project code
repository on GitHub} for both experiment times: pre- and
post-discussion: \[
f_\mathrm{pre} = 
\begin{pmatrix}
  0.38 & 0.09 & 0.09 & 0.09 & 0.08 & 0.07 & 0.19
\end{pmatrix}
\] and \[
f_\mathrm{post} = \
\begin{pmatrix} 
0.25 & 0.19 & 0.21 & 0.17 & 0.11 & 0.05 & 0.03 
\end{pmatrix}.
\]

When we plug in the bin values and copmuted \(f_\mathrm{pre}\) values
into the simulated average pre-discussion opinion definition
(Equation~\ref{eq-sim-ave-observed}), we get \[
\bar{o}_\mathrm{pre} = -3 \cdot 0.38~~+-2\cdot 0.09~~+\ldots+3\cdot0.19 = -0.6
\] and \[
\bar{o}_\mathrm{post} = -3 \cdot 0.25~~+-2\cdot 0.19~~+\ldots+3\cdot0.03 = -1.04.
\] So, we find that for the \texttt{Moscovici1969\_Americans}, our
null-polarization model appears polarized, with the exact same average
pre- and post-discussion opinions reported by Moscovici and Zavalloni
(1969), and the same magnitude of radicalization: \[
g(\nu, e) = \bar{o}_\mathrm{post} - \bar{o}_\mathrm{pre} = -1.04 - (-0.6) = - 0.4.
\]

FIGURE \hyperref[fig:distros]{{[}fig:distros{]}} ABOUT HERE

Since the simulated average opinions match the observed ones---note that
as a consequence, the spurious polarization magnitude also matches the
observed one. In other words, we \emph{identified} a null model that
explains the data. If both a hypothesis and its null explain data
equally well, then the hypothesis is \emph{undecidable} under the
available data. This invalidates this experiment's claim to be a
replication of support for the group polarization hypothesis. We
replicate this analysis for all 57 experimental conditions that purport
to support the group polarization hypothesis. To facilitate this
analysis, we created a web app to manage the data entry, the process of
identifying invalid replications, and managing the outcome data of which
replications were invalidated, and what null model parameters generated
the spurious polarization.

\begin{figure}

    \centering{
    	\includegraphics[width=0.8\textwidth]{Figures/Model/SoftwareMap_Mashup.png}
    }

    \caption{\label{fig-computational-system}\textbf{Computational system to
    interactively simulate opinion change, identify invalid replications,
    and evaluate experimental designs.}}

\end{figure}

\subsubsection{System for identifying invalid
replications}\label{system-for-identifying-invalid-replications}

We now demonstrate how to scale up and automate the analysis we did manually for just one experimental condition in the example above. The foundation of this step is an algorithm that starts from an initial user-provided guess at a mere agreement model $\nu$ that will generate spurious group polarization that matches the same average pre- and post-discussion opinions as reported in the original article.

\begin{algorithm}
  \caption{Routine to find latent mean and variance that generate spurious group polarization.}
  \label{alg:hillclimbing}
  \begin{algorithmic}

    \Require $\mu$, $\sigma$, $\bar{\mathbf{o}}$ 
      \Comment{Set latent mean, guess variance that yields, input empirical mean.}

    \Require $\epsilon$, $\delta$, $i_\mathrm{max}$ 
      \Comment{Hillclimbing tolerance, step size, and maximum iterations.}

    \State $\Delta \gets \meanobs(\mu, \sigma) - \meanobsemp$ 
      \Comment{Initialize simulation error.}
    \While {$\Delta^2 > \epsilon$ and $i < i_\mathrm{max}$} \Comment{Search until desired
    accuracy or max iterations reached.}
      \State $\sigma' \sim \mathcal{N}(\sigma, \delta)$ \Comment{Draw new variance to test.}
      \State $\Delta' \gets \bar{o}(\mu, \sigma') - \bar{\mathbf{o}}$ 
        \Comment{Calculate error using new variance.}
      \If {$\Delta' - \Delta < 0$}
        \State $\sigma \gets \sigma'$
          \Comment{Update $\sigma$ if the error is less than before.}
        \State $\Delta \gets \bar{o}(\mu, \sigma) - \bar{\mathbf{o}}$
          \Comment{Update simulation error if $\sigma$ is updated.}
      \EndIf
    \EndWhile \\

    \Return $\sigma$
  \end{algorithmic}
\end{algorithm}

\subsection{\texorpdfstring{Part 2: \emph{Evaluate} Experimental
Designs}{Part 2: Evaluate Experimental Designs}}\label{part-2-evaluate-experimental-designs}

In the previous step we identified which, if any, null polarization
models generate spurious group polarization with the expectation model
of opinion measurement. This is enough to invalidate a replication, but
it leaves unanswered questions. First, we would like to know 
the chance that a published detection of group polarization was
actually a false positive, assuming the null model identified in the first part of
the analysis. This is equivalent to the false positive rate of the
experimental design, which we denote \(\alpha_e\) for experiment \(e\).
Second, we would like to estimate the false discovery rate for our
corpus, which is a good proxy for the field as a whole, since we
selected top quality, highly influential work that has been cited
thousands of times. The false discovery rate in this case tells us the
probability that a published detection of group polarization is actually
a false detection. (Equation~\ref{eq-fdr}).

The core of our evaluation is the quantification of the probability that
a spurious result---an effect generated by the null model---is falsely
interpreted as genuine group polarization. We calculate this probability
across three distinct units of analysis, building from the specific to
the general (summarized in Table \ref{tab:error-rate-outcomes}).

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  The Type I Error Rate (\(\alpha_e\)): The rate of falsely detecting
  group polarization in a single experimental condition (\(e\)). This
  directly quantifies the risk posed by the mere agreement model
  identified for that specific experiment.
\item
  The study-level Family-Wise Error Rate (\(\alpha_s\)): The mean Type I
  Error Rate aggregated across all conditions within a single study
  (\(s\)). This estimates the false claim rate across the set of tests
  performed in that study.
\item
  The Family-Wise Error Rate (\(\alpha_{\text{all}}\)): The mean Type I
  Error Rate aggregated across all 57 experimental conditions in the
  corpus (\(E\)). This estimates the expected false claim rate for a
  researcher adopting the general experimental design found throughout
  the literature.
\end{enumerate}

These calculated Type I Error Rates (\(\alpha\)) are then combined with
fixed external parameters to estimate the False Discovery Rate.
Following convention, we adopt a fixed Statistical Power of \(W=0.80\)
(and test sensitivity with alternative values) and a Base Rate of
\(b=0.10\).

\begin{table}[htp]

    \caption{\textbf{Statistical Error Rates and Units of Analysis}}
    \begin{center}

    \begin{tabular}{llcp{2.15in}}
      \toprule
      %-----  
      \textbf{Unit of Analysis} & \textbf{Type} & \textbf{Notation} & \textbf{Definition} \\
      \midrule
      %-----
      Experiment, \(e\) 
        & Type I & \(\alpha_e\) 
        & Rate of false rejections of $e$'s null hypothesis \\

      All experiments in one study, \(s\) 
        & Family-Wise 
        & \(\alpha_s\) 
        & Mean Type I Error Rate across all conditions in study, $s$. \\

      All experiments across studies & Family-Wise & \(\alpha_{\mathrm{All}}\) &
        The mean Type I Error Rate for conditions with plausibly spurious results \\

      %-----
      \bottomrule
    \end{tabular}

    \end{center}
    \label{tab:error-rate-outcomes}
\end{table}%

\subsubsection{Generative Model of Spurious Group
Polarization}\label{generative-model-of-spurious-group-polarization}

To evaluate an experimental design's ability to properly identify null
effects, we need to simulate participant responses assuming the mere
agreement null model that invalidated the replication (\(\nu_e\)). We
use the stochastic model for this simulation stage, as it correctly
models the noise inherent in individual participant measurement and
response. The ordinal, measured opinions are generated by executing the
following steps for \(N_e\) simulated participants across \(N_T\)
independent trials:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Latent Opinion Sampling:} For each participant \(i\) and time
  \(t \in \{\text{pre, post}\}\), we draw a continuous latent opinion
  \(\omega_{i,t}\) from the corresponding normal distribution defined by
  the null model \(\nu_e\) (Equation~\ref{eq:normal-opinion-distribution}).
\item
  \textbf{Measurement Simulation (Ordinal Binning):} Each continuous
  latent opinion \(\omega_{i,t}\) is mapped onto the experiment's
  ordinal scale (e.g., \(-3\) to \(+3\)). This mapping is determined by
  the fixed measurement thresholds, \(\theta_j\), established in the
  \emph{Expectation Model} subsection
  (Equation~\ref{eq-bin-thresholds}). This process simulates the
  clipping artifact that generates spurious polarization in the
  observable data.
\end{enumerate}

This results in a simulated dataset of \(2N_e\) rows (two time steps per
participant) for a single trial, consisting of ordinal opinion values
that are then passed to the statistical inference pipeline.

\subsubsection{Statistical Inference
Pipeline}\label{statistical-inference-pipeline}

To complete our our simulated experimental design we need to be able to
statistically infer the effect sizes that measure the strength of group
polarization in an experiment. To keep the presentation simple and
transparent, we infer effect sizes by inferring the parameters \(\mu_t\)
and \(\sigma_t\) of the hypothesized normal distribution that generates
the data---\emph{literally} generates the data in our simulations, but
is just a model of the distribution of group opinions in real-world
experiments.

\subsubsection{False positive rate}\label{false-positive-rate}

To calculate the false detection rate we begin by simulating pre- and
post-deliberation opinions by drawing \(N_e\) samples from normal latent
opinion distributions that generated spurious group polarization, i.e.,
one sample at each experiment stage for each simulated participant. The
latent opinion distributions are specified with the mere agreement null
model labelled \(\nu_e\).

The family-wise error rate is the fraction of trials with
\(d_{ei} > d^*\), written
\begin{equation}\phantomsection\label{eq-study-aggregate}{
  \alpha_e = \frac{1}{N_T} \sum_i^{N_T} \Theta(d_{ei}; d^*),
}\end{equation} where \[
  \Theta(d_{ei}; d^*) = 
    \begin{cases}
      1 & \text{ if } d_{ei} > d^* \\
      0 & \text{ otherwise.}
    \end{cases}
\] \(d^*\) is the \emph{significance threshold}, which must be specified
by the experimenter. There are three common significance levels which we
adopt here: \(d^* = 0.2\) for ``low'', \(d^*=0.5\) for ``medium'', and
\(d^* = 0.8\) for ``high''.

To calculate the family-wise error rate across all experimental
conditions of a study, \(s\), \[
\alpha_s = \frac{1}{|s|} \sum_{e \in s} \alpha_e,
\] where \(e \in s\) indicates sum over all experimental conditions
\(e\) presented in the article associated with study \(s\). The mean
across all results identified as plausibly spurious is \[
 \alpha_\mathrm{all} = \frac{1}{|E|} \sum_{e \in E} \alpha_e.
\]

In the final part of our \emph{Evaluation} analysis we adapt this
analysis to re-align the question of statistical rigor: we can be
specific now, and instead of wondering, ``what significance level is
truly rigorous'', we can use our techniques to make the choice of
\(d^*\) to conform to a specific design requirement. Specifically we
will calculate what \(d^*\) must be in order to limit
\(\alpha \leq 0.05\) as an example.

\subsubsection{False discovery rate}\label{false-discovery-rate}

The false discovery rate is the false positive rate, scaled by the rate
of true negative effects and normalized by the frequency of true
positive effects\textasciitilde{} (Equation~\ref{eq-fdr}) and assuming
empirically-motivated values for base rate, \(b\), and statistical
power, \(W\), as well as more pessimistic and optimistic settings for
base rate and power. Using this notation, the false detection rate is

\begin{equation}\phantomsection\label{eq-fdr}{
\mathrm{FDR}= \frac{(1 - b)\alpha}{(1 - b)\alpha + bW}.
}\end{equation}

\subsection{\texorpdfstring{Part 3: \emph{Specify} Significance Threshold
for Experimental
Design}{Part 3: Specify Significance Threshold for Experimental Design}}\label{part-3-specify-significance-threshold-for-experimental-design}

In the final stage of our analysis, we pivot from evaluating the outcome
of published designs to \emph{specifying the statistical requirements
for future designs}. The primary goal of this specification is to invert
the standard question of statistical rigor: instead of asking if an
effect is `significant' given a fixed effect-size threshold (e.g.,
\(d^{*} = 0.5\)), we ask what significance threshold is necessary to
limit the false positive rate to a desired maximum, such as \(5\%\) (or
\(\alpha = 0.05\)).

We leverage the evaluation framework developed in Part 2, based on the
mere agreement null model \(\nu_e\) identified for each experimental
condition \(e\). Our objective is to calculate the function
\(\alpha_e(d^*)\) across a sufficient range of \(d^*\) values. This
allows us to determine the lowest \(d^*\) that satisfies the design
requirement \(\alpha_e(d^*) \leq 0.05\).

The function \(\alpha_e(d^*)\) is the simulated false positive rate as a
function of the significance threshold, \(d^*\), defined as: \[
\alpha_e(d^*) = \frac{1}{N_T} \sum_i^{N_T} \Theta(d_{ei}; d^*),
\] where \(N_T\) is the number of null trials, \(d_{ei}\) is the
calculated effect size for trial \(i\), and \(\Theta\) is the indicator
function that flags trials where the effect size \(d_{ei}\) exceeds the
significance threshold \(d^*\). In the interest of giving the most
generous best-case scenario possible, we only count positive effect
sizes because only these indicate increased extremism over time.
Negative effect sizes indicate group opinions became more
moderate---such observations do not support the group polarization
hypothesis that opinions become more extreme following discussion.

Because our system is automated, and the code publicly available, it can
be adapted and used to design future group polarization experiments with
empirically-grounded statistical rigor. Experimenters can specify a
desired maximum false positive rate, such as \(\alpha = 0.05\), and our
system will calculate the minimum effect size threshold \(d^*\) that
must be exceeded to achieve that level of rigor under the mere agreement
null model identified for their experimental design. This prescriptive
step transforms the simulation from a post-hoc critique into a
forward-looking design tool, giving experimenters an empirically-derived
minimum effect size they must exceed to maintain statistical confidence
in their results.

\subsection{Analysis Outline and Summary}

\input{analysis.tex}
% \input{discussion.tex}

\section{Discussion}\label{discussion}

\mt{Restate results and importance and outline this section.}

Since there seems to be no selection pressure on false discovery rates,
experiment design for studying group polarization was free to evolve
randomly. Various choices for the number of bins and other instrumental
details may have then become entrenched in different communities defined by
common theoretical perspectives taken by researchers in those communities.
Therefore, when Isenberg (1986)~\cite{Isenberg1986} reports that his
meta-analysis found more support for the persuasive arguments theory than the 
social comparisons theory in the form of greater effect sizes. If this were
indeed the case, perhaps it is only because some researchers hypothesizing
persuasive arguments happened to use a measurement procedure produced more
frequent and extreme distortions of opinion measurements. Incentive
structures known to promote the ``evolution of bad science'' would select
for research that ignored sticky, time-consuming statistical problems in
favor of juicy titles and headlines for which practitioners have won several
professional awards. Specifically, the traditional bias
towards positive results, and against negative ones, could easily have
provided a cultural niche in which quick but faulty group polarization
statistical methods dominate slower, more rigorous ones.
Now that we are aware of the problem, there must be
intense selective pressure applied as soon and as widely as possible to 
avoid further ``replications'' that report spurious group polarization, or
any similar latent psychological dynamics.

We provided tools that experimenters can use. While these tools are
potentially useful, they are still in the prototype phase that was
sufficient for a single researcher who was also the software developer. So,
a first step towards expanding the usefulness of the findings here is to 
improve the software usability and make the app widely available for
researchers in group polarization and beyond. Another step would be to go
beyond reactionary testing of the reliability of experimental designs, and
identify design principles that reduce the false detection rate without
needing to adjust the Cohen's $d$ (or whatever measure) 
used as a significance threshold.

When we used the appropriate ordered probit Bayesian statistical model, it
still often failed to accurately identify mere agreement, instead detecting
group polarization.  The ordered probit model we used still likely
underestimates the variance in outcomes, which could require even larger
sample sizes. A more rigorous analysis would include additional sources of
variance that will likely further inflate the Type I error rate, $\alpha$,
and the false discovery rate~\cite{Yarkoni2022}. First and foremost, there is
variance within each experimental deliberation group; no group polarization
study accounts for this, instead treating participants in each experimental
condition as if they were one large group. Furthermore, survey data are known
to be noisy~\cite{Zaller1992}. People report opinions differently over time
for no apparent reason (i.e., opinions are \emph{unstable}).  Context
matters: the order in which a survey question is asked, and which question
framing is chosen among logically equivalent alternatives, can both
significantly influence participant responses. Finally, little is known about
the psychological process that converts latent opinions to reporting
behaviors (e.g., clicking a radio button corresponding to an opinion).  The
effect of accounting for these sources of variance must be understood to
estimate $\alpha$ (and power, $W$) for the design of the next generation of
group polarization experiments.

A simple change in measurement procedure could fix the problem identified
here: center the post-discussion scale on each group’s own pre-discussion
mean. This removes the assumption that all groups share a common “neutral”
midpoint. The situation is like tossing a ball on a moving train: to the
passengers, the ball’s speed is the same in both directions, but to an
observer on the ground one throw appears faster and the other slower.
Likewise, when measurement is anchored to an external scale, simple agreement
within a group can appear as movement toward one pole. Re-centering the
measurement frame corrects this distortion. It acknowledges that opinion
change is relative to the observer’s reference frame—there is no fixed
background of neutrality against which all groups can be compared. This is,
in essence, the insight of Newtonian relativity: apparent motion depends on
the frame from which it is measured.

First we must note that there are even more, perhaps even graver problems
with group polarization research than I discussed so far: one of note is
opinion instability. Opinions are not generally stable—people seem to
construct their opinions based on several factors not directly related to
whatever underlying topic. Factors affecting reported opinions include the
language used to frame a prompt or question; the order of questions on a
multi-item survey; or simply when the question was asked (Kalton and Schuman,
1982; Zaller and Feldman, 1992; Zaller 1992). Failure to account for sources
of input variance in any experiment can lead to inflated effect sizes since
the input variance is missing from the posterior predictions during
statistical model fitting (Yarkoni, 2022). This further compounds the
theoretical confusions that fill the void when theories are never formalized
and grounded in a mathematical or computational model (Smaldino 2020; Turner
and Smaldino, 2022).

A second additional problem with group polarization research is that we literally cannot learn anything by comparing summary statistics obtained with incommensurate experimental designs—but this is exactly how group polarization research has tried to make progress. Comparing summary statistics only has explanatory power if datasets were generated under the same experimental designs—it seems this was never done or even considered in the group polarization literature. In order to compare studies, their datasets must be generated from commensurate experiments, meaning the studies must share the same predictive and statistical models and experimental design. Otherwise, such comparisons are meaningless, they literally tell us nothing, as shown by philosopher of science Nancy Cartwright in Chapter 5 of her book The Dappled World: A Study of the Boundaries of Science.

\subsection{The Full Bayesian Rigor: Deriving Effect Size Posteriors and Justifying the FPR Approximation}

In evaluating the \textbf{False Positive Rate (FPR)} of a simulated null model, the primary challenge is moving beyond the raw parameter posterior distributions ($p(\mu_1 | \text{data})$, $p(\sigma_2 | \text{data})$, etc.) to accurately quantify the posterior distribution of the derived \textbf{effect size}, $d$. This section outlines the statistically rigorous method for this derivation and justifies the pragmatic approximation necessary for the evaluation stage of this paper.

\subsubsection{The most rigorous formulation of $d$}

In our model, the effect size $d$ is a function of four estimated parameters ($\mu_1, \mu_2, \sigma_1, \sigma_2$), defined by the authors whose work we replicate as:

\begin{equation}
d = \frac{\mu_1 - \mu_2}{\sqrt{(\sigma_1^2 + \sigma_2^2)/2}}
\end{equation}

The statistically rigorous method for obtaining the posterior distribution of $d$, denoted $p(d | \text{data})$, is through \textbf{marginalization by sampling} across the joint posterior distribution of the primary parameters $p(\mu_1, \mu_2, \sigma_1, \sigma_2 | \text{data})$.

\paragraph{Necessity of MCMC Samples}

The parameters are fundamentally \textbf{sequentially dependent} in the MCMC chain, as the estimation of means is conditional on the estimation of variances. This dependency is captured in the \textbf{joint posterior distribution}.

To maintain statistical rigor, we must avoid calculating $d$ by combining the isolated summary statistics (e.g., the posterior median of $\mu_1$ and the posterior median of $\sigma_1$) because this procedure, known as \textbf{plug-in estimation}, ignores the critical correlation between parameters, resulting in an unquantified and likely underestimated uncertainty.

The correct procedure requires using the full MCMC samples:

\begin{enumerate}
    \item For each usable iteration $k$ of the Markov Chain (where $K$ is the total number of samples), a plausible set of values is drawn: $(\mu_1^{(k)}, \mu_2^{(k)}, \sigma_1^{(k)}, \sigma_2^{(k)})$.
    \item The effect size is calculated for that specific set of values:
    $$d^{(k)} = \frac{\mu_1^{(k)} - \mu_2^{(k)}}{\sqrt{\frac{(\sigma_1^{(k)})^2 + (\sigma_2^{(k)})^2}{2}}}$$
    \item The set $\{d^{(1)}, d^{(2)}, \ldots, d^{(K)}\}$ then forms the complete posterior distribution $p(d | \text{data})$.
\end{enumerate}

The center of this distribution, the \textbf{posterior median} ($\tilde{d}$), is the most robust point estimate for the effect size, as it minimizes expected absolute error and is less sensitive to potential skewness than the posterior mean.

\subsubsection{Justifying the FPR Approximation}

Due to the computational complexity and the unavailability of the original MCMC chains for all $57,000$ fits, a pragmatic simplification was necessary for the subsequent evaluation of the False Positive Rate (FPR).

\paragraph{The FPR Criterion}

A standard method for estimating the FPR (or Type I error rate) in simulation studies is to count the proportion of null trials where the observed effect size meets or exceeds a predefined threshold, $d^*$.

$$\text{FPR} = \frac{\text{Number of Null Fits where } |d| > d^*_{\text{threshold}}}{\text{Total Number of Null Simulations}}$$

In this analysis, $d^*_{\text{threshold}}$ is set to the commonly adopted small effect size threshold of $d^*=0.2$ (and varied for sensitivity analysis), avoiding the reliance on approximated $95\%$ credible intervals, which would require further, less reliable approximations.

\paragraph{The Compromise: Plug-in and Heuristics}

We employed a simplified plug-in value for $d$ using the stored posterior means:

$$d_{\text{plug}} = \frac{\hat{\mu}_1 - \hat{\mu}_2}{\sqrt{(\hat{\sigma}_1^2 + \hat{\sigma}_2^2)/2}}$$

Where $\hat{\theta}$ represents the \textbf{posterior mean} of the parameter $\theta$.

While this approach is a statistical shortcut, the error is mitigated by two factors: the use of the $d^*_{\text{threshold}}$ heuristic, and the large simulated sample size ($N=500$ per group), which ensures that the posteriors for $\mu$ and $\sigma$ are sufficiently narrow. This minimizes the difference between $d_{\text{plug}}$ and the true posterior median $\tilde{d}$, making the shortcut defensible as a pragmatic necessity given the study's scale.

\subsection{Humble Hearts Promote Rigorous Science}

Cass Sunstein kicked off the 2000s by elevating group polarization to be a scientific law and advocating the application of group polarization research to serious real-world problems in two books (Sunstein, 2009; Sunstein, 2019) and two recent high-profile articles in Nature Human Behavior: one on the social foundations of pandemic preparedness (Van Bavel, et al., 2020) and one to “promote truth, autonomy, and democratic discourse online” (Lorenz-Spreen, Lewandowski, Sunstein, and Hertwig, 2020). Let us try to stop the spread of this bad science—and let us explore what must be done to achieve Dorian Cartwright’s vision of a rigorous science of group polarization he articulated fifty years ago.

To be practically useful, science must be rigorous. Rigorous science comes most naturally when scientists remain humble, especially when they serve at the pleasure of the public, the government, or other donors. Here’s another example from Burnstein and Vinokur (1973) that we can learn from—let’s try to avoid the sort of arrogance and haughtiness that seep from these acknowledgements they wrote on p. 123:

\begin{quote}
This research was supported by a Grant from the National Institute of Mental Health (MH-16950-03) and by a Guggenheim Foundation Fellowship awarded to the first author. We gratefully acknowledge the incisive, witty, and vinous commentary provided by colleagues at the University of Provence, in particular Robert Abelson, Claude Flament, and Jean-Pierre Poitou, as well as the spirited and intelligent efforts of Livia Mezrick and Irene Graczyk, who helped us carry out the study.
\end{quote}

Science is not glamorous—it takes guts and sacrifice to achieve the technical skills and discipline to sharpen one’s prose, methods, mathematics, and computer code to the point they can really cut reality at its joints, investigate different parts, then put it all back together with a message for others about how the world really works. These days, to do rigorous social science work one must wisely curate the best of the available social science by harmonizing cross-disciplinary jargon. But disincentives still exist against careful scholarship. Prizes and professorships go to those with the most numerous publications in the right journals—period.

Unpaid, anonymous reviewers are supposed to act as gatekeepers to catch problems like the ones I reviewed. But, journal article reviewers have little to gain from enforcing rigor—actually there is a lazy strategy that can pay off big for freeloaders. It’s easy to imagine a cultural norm evolving where reviewers save themselves time and energy by asking little of an author, in hopes that little will be demanded of them when the reviewer submits their next paper. This could result in a publication process that rewards the laziest and most willing to flood the zone and self-aggrandize.

\subsection{Conclusion}

Social science can use physics as a model discipline to transition from idiosyncratic and Balkanized to informative and united. Physics only changed the world because it is useful—it is useful only because it reliably predicts things. Physics became more useful as it became more united, finding shared representations for concepts developed independently—for example the science of optics was developed independently from the science of dynamics, but the physics of light is explained partly in analogy to a pendulum or spring—the mathematics of harmonic motion can be used to predict the motion of springs or pendulums—or why we see different colors of light through a prism. With careful, coordinated tweaks to group polarization experimental design and theoretical modeling, we will build a more rigorous theory of group polarization and its large-scale expression, echo chamber radicalization.

I expect that group polarization experiments, when properly constructed and analyzed, will help sustainability campaigns optimize their social influence campaigns. We can start by developing simpler experimental designs based on formal or computational models capable of predicting outcomes in an experiment. I have developed a model that could motivate an experimental design: I showed that if extremists are more stubborn, this is sufficient to cause group polarization when likeminded groups interact over time. There is even a free stubborn extremism parameter that could be fit to data to measure the relationship between stubbornness and extremism on different issues.

Group polarization research so far has mostly provided us with a checklist of what not to do. The alternative approach I outline builds on rock—the rock of rigor—and will supersede past group polarization research built on a thousand grains of inexact verbal speculation.

\input{tables.tex}
\input{figures.tex}

\clearpage

\printbibliography[title=References]

\appendix

\renewcommand{\thefigure}{A\arabic{figure}}
\setcounter{figure}{0}
\clearpage

%%%%%%%%%% Supplement %%%%%%%%%%
\pagebreak
\begin{center}
  \textbf{\Large \textsf{Group polarization replications may be marred by high false discovery rates (Supplementary Material)}}
\end{center}

%%%%%%%%%% Prefix a "S" to all equations, figures, tables and reset the counter %%%%%%%%%%
\setcounter{equation}{1}
\setcounter{figure}{0}
\setcounter{section}{0}
\setcounter{table}{0}
\setcounter{page}{1}
\makeatletter
\renewcommand{\theequation}{S\arabic{equation}}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thesection}{S\arabic{section}}
\renewcommand{\thepage}{S\arabic{page}}
\makeatother

\section{Supplemental Analyses}

\begin{figure}
  \caption{Boxplot of Cohen's $d$ for the ordered probit model over 1000
    simulation trials for each experimental condition (y-axis in A and B).
    Each trial represents a possible outcome of a group polarization experiment where
    the true opinion shift is zero. The mean across trials is closer to 0 than
    in the metric case, but many simulated zero-shift experiments still result in
    an observed shift. This illustrates the weak statistical power of these group
    polarization experiments.}
  \label{fig:OrdinalBoxplot}
  \centering
    \includegraphics[width=1.0\textwidth]{Figures/Analysis/ordinal_cohens.pdf}
\end{figure}

\end{document}

### 2025-12-10 paper/old-intro.tex
restructured

\section{Introduction}\label{introduction}

If an extremist's opinion falls and it is measured with a Likert scale, will it
make a noise? The answer had better be yes if we want to properly measure opinion
change. Accurate measurements are critical
for developing rigorous and reliable methods to promote more centrist,
sustainable, and responsive governments and institutions
\autocite{Mason2018UncivilAgreementBook,Klein2020}. Unfortunately, social and
behavioral scientists have often used statistical methods 
that distort
opinion change when it's measured on an ordinal, Likert-style scale (e.g. -3
indicates ``Strongly disagree'', +3 indicates ``Strongly agree'', and 0 is
neutral)~\autocite{Liddell2018}. 
This includes research that Obama official and law professor Cass Sunstein
marshals to explain why ``people become extremists'' and why ``political and
cultural polarization'' is ``so pervasive in America''\footnote{From the
  Publisher's Summary. Sunstein does not make such general claims in the book,
  but does suggest that the research he reviews collectively provides
  a ``clue'' to explaining how
  ``facism...student radicalism...Islamic terrorism...[t]he Rwandan
genocide...[e]thnic Conflict.}
administration official and law professor Cass Sunstein writing a book called
*researchers construct

This work is motivated by a core conviction that better thinking leads to better
science. Theoretical clarity, careful measurement, and principled statistical
modeling are not academic luxuries, but essential tools for empirical science.
Theoretical and statistical confusions in the group polarization literature have
lain dormant, undermining measurements of opinion change in group polarization
unbeknownst to generations of researchers~\autocite{Brown1986}.

literature has long suffered from ambiguities in theory, ordinal-to-continuous mismatches in measurement, and default inferential techniques that mask model non-identifiability. By reconstructing these issues from the ground up, this paper aims not only to assess the credibility of existing evidence, but to show that rigorous thinking — about what we measure, how we model, and what we can legitimately conclude — has direct, practical consequences for the claims we make about the social world.

\emph{Group polarization} is that essentially
measures how much being in an opinion echo-chamber the name given the social phenomenon of 
when a group becomes more extreme in mean opinion over
time, assuming the mean opinion is biased to one extreme to begin with.
\mt{now hit them with the real shit to intro the problem, our analysis, and
how this can be used constructively going forward.}

\mt{Seems like a good second paragraph structurally. We should give more
details about the scope of the study, group polarization and common
elements of experimental designs.} 
Group polarization experiments tend to employ a common set of design
elements incorporated in our computational model analysis. Group
polarization experiments mostly use ordinal (see exceptions
\cite{Blascovich1974,Blascovich1975,Blascovich1976,Keating2016}). The opinions (e.g., ``Do you approve,
disapprove, or neither approve nor disapprove of the current
administration's policies?''). Extremity only increases, though, assuming
that the mean opinion of the group is initially biased towards one extreme
or the other \autocite{Brown1986,Brown2020}. Group polarization is important
to understand rigorously, therefore, since if groups become increasingly
extreme over time, while group opinions are consolidated, this would
exacerbate political polarization between political opposition groups, which
is already undermines the ability of partisans in democratic governments to
work together. Dozens of studies and meta-analyses over decades have
apparently established the empirical reality of group polarization
\autocite{Moscovici1969,Myers1970,Isenberg1986,Sunstein2009,Sunstein2019}
\autocite[Ch. 5]{Brown2000}. We believe group polarization should occur in
certain contexts since more extreme group members are likely also to be more
stubborn \autocite{Guazzini2015,Lewandowsky2019}, which should pull
consensus to extremes \autocite{Turner2020}. However, many or most of the
evidence for group polarization relies on experimental designs that use
ordinal measurements that are then analyzed as if they were continuous.
Unfortunately, this procedure is unfortunately known to induce false
positives if it is not properly accounted for \autocite{Liddell2018}.

The null-consistency procedure functions as a test of model indistinguishability: it demonstrates that a reported group polarization effect could plausibly arise from a stable-latent process combined with changing opinion precision. This logic echoes foundational concerns in causal inference, where two competing models—one causal, one not—can imply the same observed data distribution and are therefore empirically indistinguishable without further assumptions \cite{pearl2000causality,spirtes2000causation}. In such cases, inference from observation to mechanism is invalid. Similarly, in psychometrics, it is well known that different configurations of latent traits and response thresholds can yield identical observed ordinal patterns, making the true structure unidentifiable without additional constraints \cite{borsboom2005measuring}. The present approach does not attempt to estimate the probability of false detection, but rather shows that under ordinal measurement, a null model can reproduce the observed pattern—rendering the result non-diagnostic of a true latent shift. This establishes a stringent standard: an effect must not only be detected, but shown to be incompatible with plausible null-generating mechanisms.

Experimental designs in group polarization studies share a certain
experimental boilerplate (Figure 1 --- NEED TO INCLUDE) with the goal of
testing the group polarization hypothesis that a group of like-minded
people will develop more extreme opinions on some salient topic over the
course of discussing some topic (Figure~\ref{fig:ModelSketch}). 
Group polarization experiments are thus
initialized by forming like-minded groups of people and surveying their
initial opinions on some topic. In some cases these initial opinions are
used for assigning participants to like-minded groups, but other methods
have been used, such as geographic location for producing, on average,
like-minded groups \autocite{Schkade2010}. Participants then discuss the
topic in their groups. After discussion, participant opinions are again
measured. Several studies we analyze here determined group polarization
occurred (i.e., defined a positive detection) only as a non-zero
difference in mean opinion, with no other statistical testing. Some
others performed null-hypothesis \(t\)-tests to test whether there was
enough evidence to reject the null-hypothesis that \emph{simple
consensus} occurredand not group polarization. In our meta-analysis
simulations, we measure simulated group polarization effect size in
terms of Cohen's \(d\), which is an abstraction beyond these two
practices from the group polarization literature that can also be used
to compare model fits between metric models that treat group
polarization measurements as continuous although they are ordinal, and
more appropriate statistical models that account for data ordinality.

% \begin{figure}
%   \caption{\textbf{Opinions are implicitly behavioral choices that are most consistent with some
%   complex aggregation of brain states and dynamics, and so are theoretically
%   continuous (A). Common components of a group polarization experiment.}
%   Participants first give their opinion in response to some prompt (B). Afterwards,
%   participants deliberate the topic (C). Participants then give their opinion again
%   (D). }
%   \centering
%     \includegraphics[width=0.65\textwidth]{Figures/Model/ModelSketchesA.pdf}
%     \includegraphics[width=0.7\textwidth]{Figures/Model/ModelSketchesB-D.pdf}
%   \label{fig:ModelSketch}
% \end{figure}

In this paper, we perform a meta-analysis of sixty detections of group
polarization from ten published research articles and show that
fifty-five of the sixty cannot be reliably identified as group
polarization. We show that these 55 are plausibly false detections
because they are equally well-explained by a \emph{simple consensus}
model, where the mean group opinion is unchanged by group discussion but
opinion variance decreases. Theoretically, consensus occurs when
sufficiently like-minded individuals influence, which decreases group
opinion variance \autocite{DeGroot1974,Turner2018}. Group polarization,
then, is consensus plus an overall increase in group opinion extremity,
i.e., an increase in mean group opinion after discussion. We use a novel
approach, described herein, that enabled us to identify continuous pre-
and post-discussion simple consensus opinion distributions that result
in data that appear to be group-polarized when opinions are drawn from
these distributions and measured with an ordinal scale. This approach
could be adapted for similar social psychological studies that used
ordinal measures to determine whether change occurred in distributions
of continuous psychological variables such as opinions and beliefs. For
social psychology and group polarization, this means that an accepted,
heretofore unquestioned, empirical fact must be re-evaluated.
Experimental designs in the field must be updated if we are to
rigorously understand whether and when group polarization occurs.

\subsection{Studies included in this
meta-analysis}\label{studies-included-in-this-meta-analysis}

We chose the ten studies guided by a desire to identify foundational
studies that contribute to the general consensus in social psychology
that group polarization reliably occurs across experimental
settings---one author of one of the ten journal articles is a law
professor who has written two trade books on group polarization to argue
that group polarization occurs in real-world settings in addition to the
laboratory, with political and other practical importance
\autocite{Sunstein2009,Sunstein2019}. We chose studies that either had a
significant influence on the study of group polarization and social
psychology, in terms of a high citation count or adaptation in
popularized trade books, or were designed to support one of four
prominent theoretical explanations of group polarization. The studies
must also have used an ordinal scale for measuring participant opinions
(all but a handful of exceptions use ordinal measurements to measure
group polarization; e.g.~blackjack bets
\autocite{Blascovich1973,Blascovich1976} and mock jury monetary damages
\autocite{Schkade2000}). We included at least one study that supported
or was motivated by one of four established theoretical explanations
(introduced below) of group polarization to show all have at least some
supporting evidence that must be discarded. We also included two studies
that lack a clear theoretical motivation. For the sake of orienting the
reader to the group polarization literature and as a useful organizing
feature for our analysis, we now introduce the theories that motivate
eight of the ten studies. To further orient the reader, we then provide
details of the experimental designs used to produce the 60 published
positive detections of group polarization across 10 research articles
that we scrutinize here. Group polarization experimental designs are
diverse and inconsistent with one another in many ways, which
complicates this or other potential meta-analyses and may imperil our
ability to rigorously understand how regular the group polarization
effect really is \autocite{Yarkoni2022,Almaatouq2022}.

The first theoretical explanation whose supporting findings we include
here is (1) social comparisons theory, which posits that an individual's
opinions are updated so they stand out in a group enough to be
``desirably distinctive'' (Myers, 1978, p.~562) but not too distinct to
be viewed as anti-conformist \autocite{Brown1974,Sanders1977,Myers1978}.
Second, \emph{persuasive arguments theory} posits that more extreme
individuals have a greater number of more convincing arguments for their
view, which draw more moderate group members to extreme opinions
\autocite{Burnstein1973,Burnstein1975,Burnstein1977}. Third, the
\emph{self-categorization theory} predicts that group polarization will
occur as more moderate individuals judge the mean group opinion to be
more extreme \autocite{Turner1989,Abrams1990,Krizan2007}. Fourth and
finally, the social decision schemes explanation posits that the
structure of \autocite{Davis1973,Zuber1992,Friedkin1999a}. We do not
evaluate or compare these theories here, though a simpler explanation
that extremists are more stubborn may be a more parsimonious explanation
consistent with these four alternatives
\autocite{Acemoglu2013,Guazzini2015,Turner2020}.

Each of the ten group polarization studies we analyze here varied
significantly in how groups were formed, how opinions were gathered, and
other details. There was significant variation across experimental
designs, participant population sizes, group sizes, and how like-minded
groups were formed in the first place. We review some of these details
below to provide the reader with some insight into the experimental
designs that produced what we show are plausibly false positive
detections of group polarization. We focus here on the questions that
were asked, how like-minded novel experimental groups were formed,
organizing our presentation by theoretical tradition.

\subsubsection{Details of experimental conditions included in this
meta-analysis}\label{details-of-experimental-conditions-included-in-this-meta-analysis}

\paragraph*{Atheoretical}

The first of ten studies we included in our meta-analysis was by
Moscovici and Zavalloni (1969) \autocite{Moscovici1969} who sought to
measure group polarization among groups of French high school students
who discussed domestic and foreign politics. Moscovici and Zavalloni
tested four conditions whether group polarization occurred for two
different questions and two different rating systems: participants were
asked their opinions on then-president Charles De Gaulle and their
opinion about foreign aid from the USA either in the ``Opinion'' setting
or the ``Judgement'' condition (see Table 4 on p.~132).

The second, relatively more recent article is Schkade, Sunstein, and
Hastie (2010) \autocite{Schkade2010} who used geographic location to
construct like-minded groups by running a group polarization experiment
in two cities in the state of Colorado in the United States: Colorado
Springs, whose residents historically voted for republican candidates,
and Boulder, whose residents historically voted for democratic
candidates. Discussion groups were formed in which they measured group
polarization when group members discussed global warming, affirmative
action, and same-sex civil unions.

\paragraph{Social comparisons}

In developing the social comparisons theory of group polarization
(described in Myers (1978) \autocite{Myers1978}), Myers and Bishop
(1970), writing in \emph{Science,} \autocite{Myers1970} sought to
measure group polarization among high school students in the United
States who discussed racial attitudes and Myers (1975)
\autocite{Myers1975} measured group polarization in groups of college
students discussing the quality of a faculty member in one condition,
and women's rights and issues in another condition.

\paragraph{Persuasive arguments}

\paragraph{Self-categorization}

\paragraph{Social decisions scheme}

The ten studies we included each tested several experimental conditions
meant to elicit group polarization in novel experimental groups, 60 in
total, which we briefly introduce here.\\

\subsection{Research overview}\label{research-overview}

To evaluate whether published detections of group polarization are
plausibly false we developed and combined formal and generative models
of group polarization to identify plausible simple consensus pre- and
post-discussion opinion distributions that seem like group polarization
when binned or measured on ordinal scales. Opinion distributions are
assumed to be normal. We define a \emph{simple consensus pair} of
\emph{latent opinion} distributions to be a pair of pre- and
post-discussion distributions where the mean opinion is constant, but
the pre-discussion variance is greater than the post-discussion
variance. A \emph{latent distribution} of opinions represents the
internal, unobservable distribution of opinions for a group or
collection of groups. The \emph{observed distribution} is the
distribution of reported ordinal opinion measurements. We then define a
\emph{plausibly false detection} of group polarization to be one where
the mean post-discussion observed distribution is greater in magnitude
(i.e., more extreme) than the post-discussion observed distribution's,
but both were measurements of a pair of latent simple consensus
distributions.

In the formal model, a simulated observed distribution is generated by
binning a latent distribution into a finite number of ordinal scale
bins, then integrating over the probability density in each bin. In the
generative model, an observed distribution is simulated by first drawing
participant opinions from a latent distribution, then counting the
number of opinions falling within each bin. The formal model is used to
identify which, if any, simple consensus pairs exist for a given
reported detection of group polarization through an optimization routine
that finds pre- and post-discussion variances that generate simulated
observations that appear to be group polarization (or reports a failure
to such variances). If the optimization routine finds a constant latent
mean that generates simulated observed group polarization, then we say
the published group polarization detection is plausibly false. If the
routine fails to find a constant latent mean that generates simulated
group polarization, then we cannot and do not claim the detection is
plausibly false. We use the generative model to evaluate how frequently
we would expect plausibly false detections when there are finite-size
groups, since the formal model implicitly applies infinite-size groups.
Therefore in our analyses, we also generate one thousand simulated group
polarization datasets for each of the 55 plausibly false detections
identified with the formal model and examine the distribution of group
polarization effect sizes in terms of Cohen's \(d\). In the first round
of simulations we set the number of observations to match those in the
ten published studies we analyzed. Many or most simulated effect sizes
are non-zero with these numbers of observations. To see if perhaps these
non-zero effects were due to low statistical power from too few
observations, we also tested a hypothetical study with ten times more
participants. Finally, we evaluate whether an appropriate statistical
model that accounts for the ordinal measurement of latent opinions can
reliably identify simulated simple consensus in latent opinions when
simulated observations appear as group polarization based on mean group
opinion alone.

% \begin{wrapfigure}{l}{0.5\textwidth}
\begin{figure}
  \caption{\textbf{Group polarization experimental model.} Latent opinions are 
    represented as numbers corresponding to the valence and extremism of one's opinion
(A). Before deliberating a topic of interest, all participants report their 
opinion (B). Participants then deliberate the topic with others (C). Following
deliberation participants report their opinion once more (D).}
  \centering
  \includegraphics[width=0.65\textwidth]{Figures/Model/ModelSketches.pdf}
  \label{fig:experiment_model}
% \end{wrapfigure}
\end{figure}

### 2025-12-10 GroupPolarizationStatmod/analysis.R

@@ -1,188 +1,0 @@
-library(dplyr)
-library(tidystats)
-library(knitr)
-library(kableExtra)
-library(foreach)
-library(parallel)
-library(doParallel)
-
-source("experiments.R")
-
-
-##
-# Arguments:
-#  studiesAnalysisDfPath (string): path where exact analysis was done using web app
-#    to determine whether a result was a false positive or not.
-#  
-makePlausibleFPTable <- function(studiesAnalysisDfPath = "data/StudiesAnalysis.csv", 
-                                 useSynced = FALSE) 
-{
-  plausibleFPTable <-
-    read.csv(studiesAnalysisDfPath) %>%
-    filter(IncludeInt == 1) %>% 
-    group_by(ArticleTag) %>%
-    summarize(PlausibleFPs = sum(PlausibleInt), ConditionCount = n(), PlausibleFPRate = mean(PlausibleInt), )
-  
-  write.csv(plausibleFPTable, file = "data/output/PlausibleFPTable.csv", row.names = FALSE)
-  
-  return (plausibleFPTable)
-}
-
-
-##
-# Generate synthetic data based using parameters from experimental conditions
-# that yielded plausibly false positives to test how frequently simulated simple
-# consensus data yields false positives.
-#
-# Arguments:
-#  studiesAnalysisDfPath (string): path where exact analysis was done using web app
-#    to determine whether a result was a false positive or not.
-#  ntrials (int): Number of synthetic datasets to generate and t-test for each of the
-#    treatments yielding plausibly false positive results.
-#  limit (int): Limit the number of treatments to test for initial development.
-#
-make_metric_cohens_d_table <- function(studiesAnalysisDfPath = "data/StudiesAnalysis.csv",
-                                      outputPath = "data/output/cohens_sample_test.csv",
-                                      diagnosticSavePath = "data/diagnostic/tTestFits.RDS",
-                                      ntrials = 2, limit = 2, N_multiplier = 1)
-{
-  # Load analysis of studies from web interface stored as CSV.
-  studiesDf <- read.csv(studiesAnalysisDfPath)
-
-  # Filter out implausible false positives.
-  studiesDf <- filter(studiesDf, (IncludeInt == 1) & (PlausibleInt == 1))
-  
-  # Limit number of plausible false positives for preliminary development.
-  if (limit > 0)
-  {
-    studiesDf <- head(studiesDf, limit)
-  }
-  
[…]

### 2025-12-10 GroupPolarizationStatmod/app.R

@@ -1,298 +1,0 @@
-library(ggplot2)
-library(ggplotify)
-library(reshape2)
-library(readxl)
-library(shiny)
-
-source("model.R")
-source("numerical.R")
-
-# STUDIES_DB = "data/CaseStudies_ProtoDB-TEST.xlsx"
-# STUDIES <- read_excel(STUDIES_DB, "ArticleStudiesMinimal")
-STUDIES_DB = "data/StudiesAnalysis.csv"
-STUDIES <- read.csv(STUDIES_DB)
-# STUDIES <- read_excel("data/CaseStudies_ProtoDB.xlsx", "ArticleStudiesMinimal")
-FULL_TAGS <- sort(paste(STUDIES$ArticleTag, STUDIES$TreatmentTag, sep=" - "))
-
-# Define UI for application that draws a histogram
-ui <- function(request) { fluidPage(
-
-    tags$head(tags$style(HTML("
-        #plausible-div {
-            text-align: center;
-        }
-        #save-btn-div {
-            text-align: center;
-        }
-        #Notes {
-            height: 125px;
-        }
-        #saveBtn {
-            font-size: 25px;   
-            width: 50%;
-            color: white;
-            background-color: dodgerblue;
-        }
-    
-    "))),
-  
-    titlePanel("Spurious group polarization generator."),
-
-    h3(textOutput("treatmentTag")),
-
-    # Sidebar with a slider input for number of bins 
-    sidebarLayout(
-        sidebarPanel(
-          selectInput("treatmentTag", "Choose a study within an article:",
-                      choices = FULL_TAGS, size=25, selectize = FALSE),
-            width = 3
-        ),
-
-        # Show a plot of the generated distribution
-        mainPanel(
-        # Initialize parameter controls defaulted to a Schkade result.
-            fluidRow(
-                column(6,
-                    numericInput("ObservedMeanPre",
-                                "Reported, target pre-deliberation mean",
-                                step = 0.01,
-                                value = 9.2),
-                    numericInput("ObservedMeanPost",
-                                "Reported, target post-deliberation mean",
-                                step = 0.01,
-                                value = 9.4),
-                    numericInput("LatentMean",
-                               "Hypothesized latent mean:",
-                               step = 0.01,
-                               value = 5.5),
-                    numericInput("N",
-                               "Sample size: ",
[…]

### 2025-12-10 GroupPolarizationStatmod/bayesian_fit_trial.sh

@@ -1,13 +1,0 @@
-#!/bin/bash
-#SBATCH --nodes=1
-#SBATCH --ntasks=1
-#SBATCH --cpus-per-task=4
-#SBATCH --mem-per-cpu=8G
-#SBATCH --time=12:00:00
-#SBATCH --output=bayes_fits.out
-#SBATCH --partition=serc,normal
-
-module load jags
-module load R/4.2.0
-
-R -e "source(\"experiments.R\"); singleBayesianFitTrial()"

### 2025-12-10 GroupPolarizationStatmod/experiments.R

@@ -1,451 +1,0 @@
-##
-# Experiments to understand the interpretation of existing evidence for 
-# group polarization and to design new experiments.
-#
-
-# require(glue)
-require(MASS)
-
-
-
-## INITIALIZE JAGS PARAMETERS (Following Kruschke (2015) DBDA2E-utilities.R)'
-# 
-# Author: Matthew A. Turner <maturner01@gmail.com>
-# Date: 8/23/2022
-#
-
-
-# Check that required packages are installed:
-want = c("parallel","rjags","runjags","compute.es")
-have = want %in% rownames(installed.packages())
-if ( any(!have) ) { install.packages( want[!have] ) }
-
-# Load rjags. Assumes JAGS is already installed.
-try( library(rjags) )
-
-# Load runjags. Assumes JAGS is already installed.
-try( library(runjags) )
-try( runjags.options( inits.warning=FALSE , rng.warning=FALSE ) )
-
-# library(tidyverse)
-library(dplyr)
-library(tibble)
-
-library(uuid)
-
-# set default number of chains and parallelness for MCMC:
-library(parallel) # for detectCores().
-nCores = detectCores() 
-
-if ( !is.finite(nCores) ) { nCores = 1 } 
-
-if ( nCores > 4 ) 
-{ 
-  nChainsDefault = 4  # because JAGS has only 4 rng's.
-  runjagsMethodDefault = "rjparallel"
-} else if ( nCores == 4 ) { 
-  nChainsDefault = 3  # save 1 core for other processes.
-  runjagsMethodDefault = "rjparallel"
-} else { 
-  nChainsDefault = 3 
-  runjagsMethodDefault = "rjags" # NOT parallel
-}  
-
-fileNameRoot = "JAGSOutput-"
-
-
-singleBayesianFitTrial <- function(output.dir = "data/probit_fits", test = FALSE) 
-{
-    trial_file_name <- paste(UUIDgenerate(), ".csv", sep="")
-    full_path <- paste(output.dir, trial_file_name, sep="/")
-    makeBayesianFitTable(output.csv = full_path, test = test)
-}
-
-
-makeBayesianFitTable <- function(studies.data.csv = "data/StudiesAnalysis.csv", 
-                                 output.csv = "data/BayesianAnalysis.csv",
-                                 test = FALSE)
-{
-    studiesDf <- read.csv(studies.data.csv)
-    
-    outputDf <- tibble("TreatmentTag" = character(0), "ArticleTag" = character(0), 
[…]

### 2025-12-10 GroupPolarizationStatmod/model.R

@@ -1,73 +1,0 @@
-##
-# Large-N exact calculations.
-##
-
-# Probability that a participant reports opinion in bin k, i.e., between
-# theta_{k-1} and theta_k. 
-binProb <- function(thisBin, bin1, K, latentMean, latentSD)
-{
-    # Thresholds in theta vector start at bin1 + 0.5 (see Eq. 6 in L&K supplement).
-    interiorThresholds <- c(bin1:(bin1 + K - 1) + 0.5)
-
-    # Get this by solving thisbin = (bin1 - 1) + binIdx, where binIdx starts from 1.
-    binIdx <- thisBin - bin1 + 1
-    
-    # But it only works if k is not the first...
-    if (thisBin == bin1)
-    {
-        theta_km1 = -Inf
-    }
-    else
-    {
-        theta_km1 <- interiorThresholds[binIdx - 1]
-    }
-    # ... or the last values.
-    # print(thisBin)
-    # print(bin1)
-    # print(K)
-    if (thisBin == (bin1 + K - 1))
-    {
-        theta_k = Inf
-    }
-    else
-    {
-        theta_k <- interiorThresholds[binIdx]
-    }
-    
-    # From Equation 2 in Liddell & Kruschke (2018), p. 340.
-    return (pnorm((theta_k - latentMean) / latentSD) - 
-            pnorm((theta_km1 - latentMean) / latentSD))
-}
-
-
-makeProbVec <- function(kVec, latentMean, latentSD)
-{
-    L <- kVec[1]
-    K <- length(kVec)
-
-    probVec <- numeric(K)
-
-    for (ii in 1:K)
-    {
-        res <- binProb(kVec[ii], L, K, latentMean, latentSD)
-        probVec[ii] <- binProb(kVec[ii], L, K, latentMean, latentSD)
-    }
-
-    return (probVec);
-}
-
-
-# Simulated observed mean denoted <o_t> in paper.
-meanObs <- function(kVec, probVec)
-{
-    return (sum(kVec * probVec));
-}
-
-
-# Simulated observed standard deviation denoted S_t in paper.
-sdObs <- function(kVec, probVec) {
-    expectation <- meanObs(kVec, probVec)
-    variVec <- (kVec - expectation)^2
-
-    return ( sum(variVec * probVec)^(0.5) );
-}

### 2025-12-10 GroupPolarizationStatmod/numerical.R

@@ -1,75 +1,0 @@
-##
-# This module implements the numerical procedure to find what latent standard
-# deviation results in the observed mean in the large N limit, given latent
-# mean µ, the bin vectors, and thresholds.
-#
-# This estimation from the large N limit can then be compared with finite N
-# simulations with N specified by referring to studies we review in our
-# paper. 
-#
-source('model.R')
-
-
-solveForLatentSD <- function(kVec, latentMean, observedMean, guess, 
-                             step = 0.01, maxIts = 10000, tol = 1e-6)
-{
-    # Keep the error function to be minimized through hillclimbing
-    # separate from hillclimbing itself.
-    errfunc <- function(latentSD)
-    {
-        simulatedObservedMean <- meanObs(
-            kVec,
-            makeProbVec(kVec, latentMean, latentSD)
-        )
-        
-        return (simulatedObservedMean - observedMean);
-    }
-
-    # This structure is kind of weird with this implicitly defined "errfunc" that
-    # only takes one variable because the other one has been set "privately"
-    # and implicitly within this function definition.
-    return (hillclimbing(errfunc, guess, step, maxIts, tol))
-}
-
-
-##
-# Called hill climbing, but it's really hill tumbling to find the minimum.
-#
-hillclimbing <- function(errfunc, sd_guess, stepSize = 0.1, maxIts = 1e5, tol = 1e-4)
-{
-    sd_curr <- sd_guess
-    # "Error" between the published shift and the one simulated here via simple consensus.
-    err_curr <- errfunc(sd_curr)
-    sqerr_curr <- err_curr^2
-    
-    its <- 0
-    change <- 0.0
-
-    while(its < maxIts)
-    {
-        its <- its + 1
-
-        # Undirected, random search for a better standard deviation to 
-        # generate pre or post observed mean.
-        sd_next <- rnorm(1, sd_curr, stepSize)
-        err_next <- errfunc(sd_next)
-        sqerr_next <- err_next^2
-
-        change <- sqerr_next - sqerr_curr
-
-        if (abs(change) < tol)
-        {
-            return (c(sd_curr, err_curr, its));
-        }
-        
-        # A negative change means err_next is less that sqerr_curr, and closer/better
-        # to minimum solution.
-        if (change < 0)
-        {
-            sd_curr <- sd_next
-            err_curr <- err_next
-            sqerr_curr <- sqerr_next
-        }
-    }
-    return (c(sd_curr, err_curr, its));
-}

### 2025-12-10 GroupPolarizationStatmod/plot.R

@@ -1,104 +1,0 @@
-##
-# Some methods for plotting results from statmods.
-#
-
-require(fs)
-require(tidyverse)
-
-source("experiments.R")
-
-
-LIKERT_SCALE_DENS_MAX <- 0.7
-
-##
-# Plot the frequentist analysis of difference between two groups.
-#
-# Arguments:
-#     inputDf (data.frame): the simulated data to model
-#     fittedFreq (list): list of fitted frequency model parameters and summary
-#     nBins (int): number of Likert-type bins used for opinion binning
-#
-plotFreq <- function(inputDf, fittedFreq, nBins)
-{
-
-    # Open window for plot:
-    openGraph(width=3.5*2, height=2.5*1+0.75)
-    layout(matrix(1:2, nrow=1))
-    par(mar=c(3.5,3.5,2.5,0.5), mgp=c(2.0,0.7,0), oma=c(0,0,3.5,0)) # , xpd=NA )
-
-    # histBreakInc <- 1/nQ
-    # Some that seem to be defaults.
-    histBreakInc <- 1
-    xlabText="Ordinal Resp." 
-
-    # make histograms of data:
-    mainTitle <- c("Pre-discussion", "Post-discussion")
-    for ( gIdx in 1:2 ) 
-    {
-        ydata <- as.numeric(inputDf$resp[inputDf$cond == gIdx])
-        histInfo <- hist( 
-            ydata, main=mainTitle[gIdx], 
-            xlab=xlabText, ylab="Probability",
-            freq=FALSE, col="pink", border="white", cex.lab=1.5,
-            breaks=seq(-histBreakInc/2, 
-                       nBins+1+histBreakInc/2, 
-                       by=histBreakInc), 
-            ylim=c(0, LIKERT_SCALE_DENS_MAX) # NB: must set manually at this time
-        )
-        #text( max(histInfo$breaks) , likertScaleDensMax, adj=c(1,1),
-        #      labels=bquote(N==.(length(ydata))) )
-        # xcomb <- seq(0,max(nYlevels)+1,length=201)
-        xcomb <- seq(0, nBins + 1, length=201)
-        lines(xcomb, dnorm(xcomb-mean(ydata), 0, sd(ydata)), 
-              col="skyblue", lwd=2)
-    }
-    
-    tInfo <- fittedFreq$tInfo
-    effSz <- fittedFreq$effSz
[…]

### 2025-12-10 GroupPolarizationStatmod/python/get_num_citations.py

@@ -1,40 +1,0 @@
-from datetime import date
-from pandas import read_csv, DataFrame
-
-from scholarly import scholarly
-
-
-ARTICLESDF = read_csv('data/titles.csv')
-
-
-def make_num_citations_csv(output_file_root = 'data/output/citation_count'):
-
-    outdf = ARTICLESDF.copy()
-    outdf['num_citations'] = outdf.Title.map(lambda title: get_num_citations(title))
-
-    outdf.to_csv(output_file_root + str(date.today()) + '.csv', index=False)
-
-    return outdf
-
-
-def get_num_citations(title):
-
-    with open('data/diagnostic/num_citations_log.txt', 'a') as log_f:
-        result = next(scholarly.search_pubs(title))
-
-        # Uncomment to see full result to double check citation count.
-        # print(result)
-
-        found_title = result['bib']['title']
-        num_citations = result['num_citations']
-
-        # Tell the user what's been found to use as entry for
-        diagnostic_msg = \
-            f'* Found "{found_title}" result\nfor searched title "{title}"\nwith ' \
-            f'{num_citations} citations\n'
-
-        print(diagnostic_msg)
-
-        log_f.write(diagnostic_msg + '\n')
-
-        return num_citations

### 2025-12-10 GroupPolarizationStatmod/runBayesTrialCluster.R

@@ -1,1 +1,0 @@
-source("experiments.R")

### 2025-12-10 GroupPolarizationStatmod/scripts/analysis.R

@@ -1,156 +1,0 @@
-##
-# Two functions. First, calculate a tibble that calculates the family-wise
-# error rate (FWER) and false detection rate (FDR) for three significance values
-# d^*=0.2,0.5,0.8 at three different levels of aggregation (experimental condition, 
-# journal article, and all). The significance values correspond to Cohen's
-# (1988; pp. ~25-30) recommendation. 
-#
-# Second, visualize the tibble at the different aggregation levels for building
-# into a composite figure.
-# 
-# Author: Matthew A. Turner <maturner01@gmail.com>
-# Date: 2024-11-05
-#
-
-library(purrr)
-
-source("src/fwer.R")
-source("src/util.R")
-
-
-# b for base rate, W for power. See 
-fdr = function(fwer, b, W) {
-
-  num = (1 - b) * fwer
-  denom = (num + (b * W))
-
-  return (num / denom)
-}
-
-
-add_default_rows = function(fdr_vs_sig_tbl, default_fwer = 0.05) {
-
-  # Figure out the number of sig values used.
-  sigvals = unique(fdr_vs_sig_tbl$SigVal)
-  n_sigvals = length(sigvals)
-
-  # Extract power and base rate.
-  W = pull(fdr_vs_sig_tbl[1, "Power"])
-  b = pull(fdr_vs_sig_tbl[1, "BaseRate"])
-  
-  # Load and use non-identified plausible.
-  non_id = NON_IDENTIFIED
-
-  n_non_id = n_distinct(non_id)
-  n_default_rows = n_non_id * n_sigvals
-
-  default_fdr = fdr(default_fwer, b, W)
-
-  default_rows = tibble(StudyID = rep(non_id$StudyID, n_sigvals),
-                        ExperimentID = rep(non_id$ExperimentID, n_sigvals),
-                        FWER = default_fwer,
-                        FDR = default_fdr,
-                        SigVal = rep(sigvals, n_non_id),
-                        Power = W,
-                        BaseRate = b)
-                        # Power = rep(W, n_default_rows),
-                        # BaseRate = rep(b, n_default_rows))
-  
-  # Use default FWER, FDR for all SigVals for non-plausibly-identified studies.
-  return (bind_rows(fdr_vs_sig_tbl, default_rows))
-}
-
-
-aggregate_fdr_vs_sig = function(fdr_vs_sig_tbl, use_non_identified = FALSE, 
[…]

### 2025-12-10 GroupPolarizationStatmod/scripts/calculations.R

@@ -1,111 +1,0 @@
-##
-# Calculations that are included in paper prose.
-#
-# Author: Matthew A. Turner
-# Date: 2024-12-27
-#
-
-
-source("scripts/analysis.R")
-
-
-# Create a CSV list of quantile calculations that may be included in publications.
-quantiles_tbl = function(tex_output_file = "paper/tex/quantile_table.tex", 
-                         q_resolution_or_probs = 0.05, 
-                         significance_vals = c(0.2, 0.5, 0.8),
-                         base_rate = 0.1, power = 0.8
-                         ) {
-  
-  # Hacky but just need this to proceed for now.
-  if (length(q_resolution_or_probs) > 1) {
-    probs <- q_resolution_or_probs
-  } else {
-    probs <- seq(0.0, 1.0, q_resolution_or_probs)
-  }
-
-  initialized = FALSE
-  for (sigval in significance_vals) {
-
-    tbl <- calc_fdr_vs_significance(significance_vals = sigval, 
-                                    base_rate = base_rate, power = power)
-
-    for (measure in c("FWER", "FDR")) {
-
-      # Get quantile for this measure, base rate, and power.
-      q <- quantile(as_vector(tbl[, measure]), probs = probs)
-
-      if (!initialized) {
-        # Load quantile named list into tbl with quantiles as colnames sans "%".
-        q_tbl <- tibble(!!!q, .name_repair = ~ str_replace(., "%", ""))
-        q_tbl$BaseRate <- base_rate
-        q_tbl$Power <- power
-        q_tbl$SigVal <- sigval
-        q_tbl$Measure <- measure
-
-        initialized = TRUE
-      }
-      else {
-        # Create new row.
-        new_row <- tibble(!!!q, .name_repair = ~ str_replace(., "%", ""))
-        new_row$BaseRate <- base_rate
-        new_row$Power <- power
-        new_row$SigVal <- sigval
-        new_row$Measure <- measure
-
-        # Bind new row to existing.
-        q_tbl <- rbind(q_tbl, new_row)
-      }
-    }
-  }
-
-  # Reorder columns to be easier to read.
-  q_tbl <- 
-    q_tbl %>% 
-      relocate(Measure, .before = `0`) %>% 
-      relocate(SigVal, .after = Measure) %>% 
[…]

### 2025-12-10 GroupPolarizationStatmod/scripts/plot/analysis.R

@@ -1,231 +1,0 @@
-##
-#
-# Plot analysis of family-wise error rate simulations.
-# 
-# Currently model illustrations are in the root project directory and will
-# soon be ported here.
-# 
-# Author: Matthew A. Turner <maturner01@gmail.com>
-# Date: 2024-12-07
-#
-
-library(forcats)
-library(ggplot2)
-library(latex2exp)
-library(emojifont)
-
-source("src/fwer.R")
-source("scripts/analysis.R")
-
-mytheme = theme(
-  axis.line = element_line(), legend.key=element_rect(fill = NA),
-  text = element_text(size=18),# family = 'PT Sans'),
-  # legend.key.width = unit(2, 'cm'),
-  # legend.key.size = unit(1.5, 'lines'),
-  panel.background = element_rect(fill = "white"),
-  panel.grid.major.x = element_line(color = "lightgrey", linewidth = 0.1, linetype = 2)
-)
-
-rhg_cols <- rev(c("#771C19", "#AA3929", "#E25033", "#F27314", "#F8A31B", 
-              "#E2C59F", "#B6C5CC", "#8E9CA3", "#556670", "#556AAA", "#000000"))
-
-rhg_cols_10 <- rev(c("#771C19", "#AA3929", "#E25033", "#F27314", "#F8A31B", 
-              "#E2C59F", "#B6C5CC", "#8E9CA3", "#556670", "#556AAA"))
-
-plot_ordinal_cohens <- function(ordinal_data_dir = "data/probit_fits",
-                                sync_file = "data/probit_fits/all.csv",
-                                output_file = 
-                                  "paper/Figures/Analysis/ordinal_cohens.pdf",
-                                overwrite = FALSE) {
-  
-  df <- load_probit_data(probit_fits_dir, sync_file, overwrite)
-  
-  df$Cohens_d <- cohens_d(df$LatentMeanPrePosteriorMean, 
-                          df$LatentMeanPostPosteriorMean,
-                          df$LatentMeanPrePosteriorSD, 
-                          df$LatentMeanPostPosteriorSD)
-  
-  ggplot(df, mapping = aes(x = Cohens_d, y = ExperimentID)) + 
-    geom_vline(xintercept = 0, color = "red") + 
-    geom_vline(xintercept = c(-1, 1), color="red", linetype="dotted") + 
-    geom_boxplot() 
-  
-  ggsave(output_file)
-}
[…]

### 2025-12-10 GroupPolarizationStatmod/scripts/plot/consensus-params-cdf.R

@@ -1,68 +1,0 @@
-library(readr)
-library(dplyr)
-library(tidyr)
-library(ggplot2)
-
-# Read data
-df <- read_csv("data/StudiesAnalysis.csv")
-
-# Filter to plausibly spurious & included
-df_plot <- df %>%
-  filter(PlausibleInt == 1, IncludeInt == 1) %>%
-  select(ArticleTag, LatentSDPre, LatentSDPost) %>%
-  pivot_longer(cols = c(LatentSDPre, LatentSDPost),
-               names_to = "phase", values_to = "latent_sd") %>%
-  mutate(phase = ifelse(phase == "LatentSDPre", "Pre", "Post"))
-
-# Base plot object
-p <- ggplot() +
-  # Per-article ECDFs (thin + transparent)
-  stat_ecdf(data = df_plot, aes(x = latent_sd, group = interaction(ArticleTag, phase),
-                                color = phase, linetype = phase),
-            geom = "step", alpha = 0.2, size = 0.4) +
-  # Overall ECDFs (bold)
-  stat_ecdf(data = df_plot, aes(x = latent_sd, color = phase, linetype = phase),
-            geom = "step", size = 1.2) +
-  scale_color_manual(values = c("Pre" = "#1b9e77", "Post" = "#d95f02")) +
-  labs(x = "Latent standard deviation",
-       y = "Cumulative fraction of conditions",
-       title = "Pre vs. Post deliberation latent SDs (plausibly spurious cases)") +
-  theme_minimal(base_size = 14) +
-  theme(panel.grid.minor = element_blank())
-
-p
-
-
-library(readr)
-library(dplyr)
-library(tidyr)
-library(ggplot2)
-
-# Read data
-df <- read_csv("data/StudiesAnalysis.csv")
-
-# Filter to plausibly spurious and included
-df_plot <- df %>%
-  filter(PlausibleInt == 1, IncludeInt == 1) %>%
-  select(TreatmentTag, ArticleTag, LatentSDPre, LatentSDPost) %>%
-  pivot_longer(cols = c(LatentSDPre, LatentSDPost),
-               names_to = "Phase", values_to = "latent_sd") %>%
-  mutate(Phase = ifelse(Phase == "LatentSDPre", "Pre", "Post"))
-
-# Plot CDF
-p <- ggplot(df_plot, aes(x = latent_sd, color = Phase, linetype = Phase)) +
-  stat_ecdf(size = 1) +
[…]

### 2025-12-10 GroupPolarizationStatmod/scripts/plot/model.R

@@ -1,180 +1,0 @@
-library(ggplot2)
-library(readr)
-library(dplyr)
-
-source("model.R")  # for binProb function used in ordinal plot function below.
-
-
-mytheme <- theme(
-  panel.border = element_blank(), axis.line = element_line(), 
-  text = element_text(size=16, family = 'PT Sans'), 
-  panel.background = element_rect(fill = "white")
-)
-
-plot_latent_pdf_integration = function(mu = 0, sd = 2.5, min_bin, max_bin, 
-                                       bins, bin_colors, # xlim = c(-6.0, 6.0), 
-                                       save_path = "test_latent_pdf.pdf", 
-                                       max_prob = 1.0) {
-  
-  print(save_path)
-
-  xmin = min_bin - 0.5
-  xmax = max_bin + 0.5
-
-  xlim = c(xmin, xmax)
-
-  print(xlim)
-
-  p = ggplot(data.frame(x = xlim), aes(x)) 
-
-  p = p + stat_function(fun = dnorm,
-                        args = c(mu, sd),
-                        geom = "line",
-                        xlim = c(xmin, xmax)) 
-
-  # First bin from -infinity to min_bin + 0.5
-  p = p + stat_function(fun = dnorm,
-                        geom = "area",
-                        args = c(mu, sd),
-                        fill = bin_colors[1],
-                        xlim = c(xmin, min_bin + 0.5))
-
-  n_bins = length(bins)
-
-  # Middle bins.
-  for (idx in 2:(n_bins - 1)) {
-
-    p = p + stat_function(fun = dnorm,
-                          geom = "area",
-                          args = c(mu, sd),
-                          fill = bin_colors[idx],
-                          xlim = c(bins[idx - 1] + 0.5, bins[idx] + 0.5))
-  }
-
-  # Final bin from last_bin - 0.5 to infinity.
-  p = p + stat_function(fun = dnorm,
-                        geom = "area",
-                        args = c(mu, sd),
-                        fill = bin_colors[n_bins],
-                        xlim = c(xmax, max_bin - 0.5))
-
-  p = p + xlab("Latent opinion") + ylab("Probability density\n\n") + 
-  geom_vline(xintercept=c((bins[1:(length(bins) - 1)] + 0.5)), linetype="dotted") + 
[…]

### 2025-12-10 GroupPolarizationStatmod/src/fwer.R

@@ -1,62 +1,0 @@
-##
-# Calculate the false detection rate from bayesian fits data.
-#
-# Author: Matthew A. Turner
-# Date: 2024-10-30
-#
-
-library(dplyr)
-library(readr)
-library(tidyr)
-
-source("src/util.R")
-
-
-cohens_d <- function(pre_mean_estimate, post_mean_estimate, pre_sd, post_sd) {
-  
-  numerator <- post_mean_estimate - pre_mean_estimate
-  
-  denominator = sqrt((pre_sd**2 + post_sd**2) / 2.0)
-  
-  return (numerator / denominator)
-}
-
-
-##
-# Family-wise error rate calculation using whatever ExperimentIDs are present.
-# Note probit_fits may be a tibble for avoiding reading data every calculation.
-#
-calculate_fwer <- function (probit_fits = "data/probit_fits", sigval = 0.8) {
-
-  # Load probit data if necessary.
-  if (typeof(probit_fits) == "character") {
-    probit_data <- load_probit_data(probit_fits)
-  } else {
-    probit_data <- probit_fits
-  }
-     
-  probit_data$Cohens_d <- cohens_d(probit_data$LatentMeanPrePosteriorMean, 
-                                  probit_data$LatentMeanPostPosteriorMean,
-                                  probit_data$LatentMeanPrePosteriorSD, 
-                                  probit_data$LatentMeanPostPosteriorSD)
-
-  # Get the sign of the hypothesized shift, where these columns come from 
-  # the original data.
-  shift_sign <- sign(probit_data$ObservedMeanPost - probit_data$ObservedMeanPre)
-
-  # Only significant if it's positive in direction of shift. TODO: alternative
-  # formulation could check absolute value.
-  probit_data$Significant <- 
-    ifelse((shift_sign * probit_data$Cohens_d) >= sigval, 1, 0)
-
-  ret = probit_data %>%
-    group_by(ExperimentID) %>% 
-    summarise(FWER = mean(Significant)) %>%
-    separate_wider_delim(ExperimentID, delim = "_", 
-                         names = c("StudyID", "ExperimentID")) %>%
-    arrange(desc(FWER))
-
-  return (ret)
-}
-
-

### 2025-12-10 GroupPolarizationStatmod/src/util.R

@@ -1,32 +1,0 @@
-##
-# Utility functions used in other src and scripts files.
-#
-# Author: Matthew A. Turner
-# Date: 2024-12-27
-#
-
-library(dplyr)
-library(readr)
-library(tidyr)
-library(fs)
-
-
-load_probit_data = function(probit_fits_dir = "data/probit_fits", 
-                            sync_file = "data/probit_fits/all.csv",
-                            overwrite = FALSE) {
-     
-  if (!file.exists(sync_file) || overwrite) {
-
-    ret <- dir_ls(probit_fits_dir, glob = "*.csv") %>%
-      read_csv() %>% unite(ExperimentID, ArticleTag, TreatmentTag)
-
-    write_csv(ret, sync_file)
-
-  } else {
-
-    ret <- read_csv(sync_file)
-  }
-
-  return (ret)
-}
-

### 2025-12-10 GroupPolarizationStatmod/util.R

@@ -1,136 +1,0 @@
-##
-# The main display element is one row of controls and two rows of plots. 
-#
-# The top row will be the large N exact calculations of the 
-# latent standard deviations given observed pre- and post-deliberation
-# means, a hypothesized latent mean, and a reporting model and measurement 
-# scheme. The latent mean and observed pre- and post-deliberation means
-# can be set using the top row controls, or loaded from preset values
-# from published result case studies that contain plausibly
-# false demonstrations of the group polarization effect.
-#
-# The second row displays simulated empirical observations for an N given in 
-# the controls at the top of the main display. The numbers and plots 
-# are generated using the same latent and observed 
-# pre- and post-deliberation means as for the large N exact calculations.
-#
-# The "reporting model" is operationalized as thresholds that determine
-# which opinion scale (e.g. Likert) bin a particpant's opinion is reported as.
-# See the code for how this works. The rating scale would be, e.g., a
-# 7-point Likert, or the CDQ responses, which rate riskiness from 1 to 10.
-#  
-# getCaseStudyData <- function() {
-#     read.csv('caseStudies.csv') 
-#     data.names <- data$CaseStudyName
-#     return (data)
-# }
-# CaseStudyData <- getCaseStudyData();
-
-library(shiny)
-library(grid)
-library(gridExtra)
-library(lattice)
-library(ggplot2)
-
-source("model.R")
-source("numerical.R")
-
-# CaseStudyData <- read.csv('caseStudies.csv', row.names='CaseStudyTag')
-
-
-##
-# Arguments:
-#     caseStudy (string): See CSV for values to use, e.g., 
-#         Moscovici1969-DeGaulle.
-setCaseStudyValues <- function(caseStudyName, session)
-{
-    data <- CaseStudyData[caseStudyName, ]
-    updateSliderInput(session, "latentMean", 
-                      value = data$LatentMean, min = data$MinBinValue - 2,
-                      max = data$MaxBinValue + 2)
-                      
-    updateSliderInput(session, "observedPreDelibMean", 
-                      value = data$ObservedPreDelibMean, 
-                      min = data$MinBinValue,
-                      max = data$MaxBinValue)
-
-    updateSliderInput(session, "observedPostDelibMean", 
-                      value = data$ObservedPostDelibMean, 
[…]

### 2025-12-10 GroupPolarizationStatmod/www/message-handler.js

@@ -1,5 +1,0 @@
-Shiny.addCustomMessageHandler("testmessageHandlerForJS",
-  function(message) {
-    alert(JSON.stringify(message));
-  }
-);

### 2025-12-10 analysis.R

@@ -1,0 +1,188 @@
+library(dplyr)
+library(tidystats)
+library(knitr)
+library(kableExtra)
+library(foreach)
+library(parallel)
+library(doParallel)
+
+source("experiments.R")
+
+
+##
+# Arguments:
+#  studiesAnalysisDfPath (string): path where exact analysis was done using web app
+#    to determine whether a result was a false positive or not.
+#  
+makePlausibleFPTable <- function(studiesAnalysisDfPath = "data/StudiesAnalysis.csv", 
+                                 useSynced = FALSE) 
+{
+  plausibleFPTable <-
+    read.csv(studiesAnalysisDfPath) %>%
+    filter(IncludeInt == 1) %>% 
+    group_by(ArticleTag) %>%
+    summarize(PlausibleFPs = sum(PlausibleInt), ConditionCount = n(), PlausibleFPRate = mean(PlausibleInt), )
+  
+  write.csv(plausibleFPTable, file = "data/output/PlausibleFPTable.csv", row.names = FALSE)
+  
+  return (plausibleFPTable)
+}
+
+
+##
+# Generate synthetic data based using parameters from experimental conditions
+# that yielded plausibly false positives to test how frequently simulated simple
+# consensus data yields false positives.
+#
+# Arguments:
+#  studiesAnalysisDfPath (string): path where exact analysis was done using web app
+#    to determine whether a result was a false positive or not.
+#  ntrials (int): Number of synthetic datasets to generate and t-test for each of the
+#    treatments yielding plausibly false positive results.
+#  limit (int): Limit the number of treatments to test for initial development.
+#
+make_metric_cohens_d_table <- function(studiesAnalysisDfPath = "data/StudiesAnalysis.csv",
+                                      outputPath = "data/output/cohens_sample_test.csv",
+                                      diagnosticSavePath = "data/diagnostic/tTestFits.RDS",
+                                      ntrials = 2, limit = 2, N_multiplier = 1)
+{
+  # Load analysis of studies from web interface stored as CSV.
+  studiesDf <- read.csv(studiesAnalysisDfPath)
+
+  # Filter out implausible false positives.
+  studiesDf <- filter(studiesDf, (IncludeInt == 1) & (PlausibleInt == 1))
+  
+  # Limit number of plausible false positives for preliminary development.
+  if (limit > 0)
+  {
+    studiesDf <- head(studiesDf, limit)
+  }
+  
[…]

### 2025-12-10 app.R

@@ -1,0 +1,298 @@
+library(ggplot2)
+library(ggplotify)
+library(reshape2)
+library(readxl)
+library(shiny)
+
+source("model.R")
+source("numerical.R")
+
+# STUDIES_DB = "data/CaseStudies_ProtoDB-TEST.xlsx"
+# STUDIES <- read_excel(STUDIES_DB, "ArticleStudiesMinimal")
+STUDIES_DB = "data/StudiesAnalysis.csv"
+STUDIES <- read.csv(STUDIES_DB)
+# STUDIES <- read_excel("data/CaseStudies_ProtoDB.xlsx", "ArticleStudiesMinimal")
+FULL_TAGS <- sort(paste(STUDIES$ArticleTag, STUDIES$TreatmentTag, sep=" - "))
+
+# Define UI for application that draws a histogram
+ui <- function(request) { fluidPage(
+
+    tags$head(tags$style(HTML("
+        #plausible-div {
+            text-align: center;
+        }
+        #save-btn-div {
+            text-align: center;
+        }
+        #Notes {
+            height: 125px;
+        }
+        #saveBtn {
+            font-size: 25px;   
+            width: 50%;
+            color: white;
+            background-color: dodgerblue;
+        }
+    
+    "))),
+  
+    titlePanel("Spurious group polarization generator."),
+
+    h3(textOutput("treatmentTag")),
+
+    # Sidebar with a slider input for number of bins 
+    sidebarLayout(
+        sidebarPanel(
+          selectInput("treatmentTag", "Choose a study within an article:",
+                      choices = FULL_TAGS, size=25, selectize = FALSE),
+            width = 3
+        ),
+
+        # Show a plot of the generated distribution
+        mainPanel(
+        # Initialize parameter controls defaulted to a Schkade result.
+            fluidRow(
+                column(6,
+                    numericInput("ObservedMeanPre",
+                                "Reported, target pre-deliberation mean",
+                                step = 0.01,
+                                value = 9.2),
+                    numericInput("ObservedMeanPost",
+                                "Reported, target post-deliberation mean",
+                                step = 0.01,
+                                value = 9.4),
+                    numericInput("LatentMean",
+                               "Hypothesized latent mean:",
+                               step = 0.01,
+                               value = 5.5),
+                    numericInput("N",
+                               "Sample size: ",
[…]

### 2025-12-10 experiments.R

@@ -1,0 +1,451 @@
+##
+# Experiments to understand the interpretation of existing evidence for 
+# group polarization and to design new experiments.
+#
+
+# require(glue)
+require(MASS)
+
+
+
+## INITIALIZE JAGS PARAMETERS (Following Kruschke (2015) DBDA2E-utilities.R)'
+# 
+# Author: Matthew A. Turner <maturner01@gmail.com>
+# Date: 8/23/2022
+#
+
+
+# Check that required packages are installed:
+want = c("parallel","rjags","runjags","compute.es")
+have = want %in% rownames(installed.packages())
+if ( any(!have) ) { install.packages( want[!have] ) }
+
+# Load rjags. Assumes JAGS is already installed.
+try( library(rjags) )
+
+# Load runjags. Assumes JAGS is already installed.
+try( library(runjags) )
+try( runjags.options( inits.warning=FALSE , rng.warning=FALSE ) )
+
+# library(tidyverse)
+library(dplyr)
+library(tibble)
+
+library(uuid)
+
+# set default number of chains and parallelness for MCMC:
+library(parallel) # for detectCores().
+nCores = detectCores() 
+
+if ( !is.finite(nCores) ) { nCores = 1 } 
+
+if ( nCores > 4 ) 
+{ 
+  nChainsDefault = 4  # because JAGS has only 4 rng's.
+  runjagsMethodDefault = "rjparallel"
+} else if ( nCores == 4 ) { 
+  nChainsDefault = 3  # save 1 core for other processes.
+  runjagsMethodDefault = "rjparallel"
+} else { 
+  nChainsDefault = 3 
+  runjagsMethodDefault = "rjags" # NOT parallel
+}  
+
+fileNameRoot = "JAGSOutput-"
+
+
+singleBayesianFitTrial <- function(output.dir = "data/probit_fits", test = FALSE) 
+{
+    trial_file_name <- paste(UUIDgenerate(), ".csv", sep="")
+    full_path <- paste(output.dir, trial_file_name, sep="/")
+    makeBayesianFitTable(output.csv = full_path, test = test)
+}
+
+
+makeBayesianFitTable <- function(studies.data.csv = "data/StudiesAnalysis.csv", 
+                                 output.csv = "data/BayesianAnalysis.csv",
+                                 test = FALSE)
+{
+    studiesDf <- read.csv(studies.data.csv)
+    
+    outputDf <- tibble("TreatmentTag" = character(0), "ArticleTag" = character(0), 
[…]

### 2025-12-10 model.R

@@ -1,0 +1,73 @@
+##
+# Large-N exact calculations.
+##
+
+# Probability that a participant reports opinion in bin k, i.e., between
+# theta_{k-1} and theta_k. 
+binProb <- function(thisBin, bin1, K, latentMean, latentSD)
+{
+    # Thresholds in theta vector start at bin1 + 0.5 (see Eq. 6 in L&K supplement).
+    interiorThresholds <- c(bin1:(bin1 + K - 1) + 0.5)
+
+    # Get this by solving thisbin = (bin1 - 1) + binIdx, where binIdx starts from 1.
+    binIdx <- thisBin - bin1 + 1
+    
+    # But it only works if k is not the first...
+    if (thisBin == bin1)
+    {
+        theta_km1 = -Inf
+    }
+    else
+    {
+        theta_km1 <- interiorThresholds[binIdx - 1]
+    }
+    # ... or the last values.
+    # print(thisBin)
+    # print(bin1)
+    # print(K)
+    if (thisBin == (bin1 + K - 1))
+    {
+        theta_k = Inf
+    }
+    else
+    {
+        theta_k <- interiorThresholds[binIdx]
+    }
+    
+    # From Equation 2 in Liddell & Kruschke (2018), p. 340.
+    return (pnorm((theta_k - latentMean) / latentSD) - 
+            pnorm((theta_km1 - latentMean) / latentSD))
+}
+
+
+makeProbVec <- function(kVec, latentMean, latentSD)
+{
+    L <- kVec[1]
+    K <- length(kVec)
+
+    probVec <- numeric(K)
+
+    for (ii in 1:K)
+    {
+        res <- binProb(kVec[ii], L, K, latentMean, latentSD)
+        probVec[ii] <- binProb(kVec[ii], L, K, latentMean, latentSD)
+    }
+
+    return (probVec);
+}
+
+
+# Simulated observed mean denoted <o_t> in paper.
+meanObs <- function(kVec, probVec)
+{
+    return (sum(kVec * probVec));
+}
+
+
+# Simulated observed standard deviation denoted S_t in paper.
+sdObs <- function(kVec, probVec) {
+    expectation <- meanObs(kVec, probVec)
+    variVec <- (kVec - expectation)^2
+
+    return ( sum(variVec * probVec)^(0.5) );
+}

### 2025-12-10 numerical.R

@@ -1,0 +1,75 @@
+##
+# This module implements the numerical procedure to find what latent standard
+# deviation results in the observed mean in the large N limit, given latent
+# mean µ, the bin vectors, and thresholds.
+#
+# This estimation from the large N limit can then be compared with finite N
+# simulations with N specified by referring to studies we review in our
+# paper. 
+#
+source('model.R')
+
+
+solveForLatentSD <- function(kVec, latentMean, observedMean, guess, 
+                             step = 0.01, maxIts = 10000, tol = 1e-6)
+{
+    # Keep the error function to be minimized through hillclimbing
+    # separate from hillclimbing itself.
+    errfunc <- function(latentSD)
+    {
+        simulatedObservedMean <- meanObs(
+            kVec,
+            makeProbVec(kVec, latentMean, latentSD)
+        )
+        
+        return (simulatedObservedMean - observedMean);
+    }
+
+    # This structure is kind of weird with this implicitly defined "errfunc" that
+    # only takes one variable because the other one has been set "privately"
+    # and implicitly within this function definition.
+    return (hillclimbing(errfunc, guess, step, maxIts, tol))
+}
+
+
+##
+# Called hill climbing, but it's really hill tumbling to find the minimum.
+#
+hillclimbing <- function(errfunc, sd_guess, stepSize = 0.1, maxIts = 1e5, tol = 1e-4)
+{
+    sd_curr <- sd_guess
+    # "Error" between the published shift and the one simulated here via simple consensus.
+    err_curr <- errfunc(sd_curr)
+    sqerr_curr <- err_curr^2
+    
+    its <- 0
+    change <- 0.0
+
+    while(its < maxIts)
+    {
+        its <- its + 1
+
+        # Undirected, random search for a better standard deviation to 
+        # generate pre or post observed mean.
+        sd_next <- rnorm(1, sd_curr, stepSize)
+        err_next <- errfunc(sd_next)
+        sqerr_next <- err_next^2
+
+        change <- sqerr_next - sqerr_curr
+
+        if (abs(change) < tol)
+        {
+            return (c(sd_curr, err_curr, its));
+        }
+        
+        # A negative change means err_next is less that sqerr_curr, and closer/better
+        # to minimum solution.
+        if (change < 0)
+        {
+            sd_curr <- sd_next
+            err_curr <- err_next
+            sqerr_curr <- sqerr_next
+        }
+    }
+    return (c(sd_curr, err_curr, its));
+}

### 2025-12-10 plot.R

@@ -1,0 +1,104 @@
+##
+# Some methods for plotting results from statmods.
+#
+
+require(fs)
+require(tidyverse)
+
+source("experiments.R")
+
+
+LIKERT_SCALE_DENS_MAX <- 0.7
+
+##
+# Plot the frequentist analysis of difference between two groups.
+#
+# Arguments:
+#     inputDf (data.frame): the simulated data to model
+#     fittedFreq (list): list of fitted frequency model parameters and summary
+#     nBins (int): number of Likert-type bins used for opinion binning
+#
+plotFreq <- function(inputDf, fittedFreq, nBins)
+{
+
+    # Open window for plot:
+    openGraph(width=3.5*2, height=2.5*1+0.75)
+    layout(matrix(1:2, nrow=1))
+    par(mar=c(3.5,3.5,2.5,0.5), mgp=c(2.0,0.7,0), oma=c(0,0,3.5,0)) # , xpd=NA )
+
+    # histBreakInc <- 1/nQ
+    # Some that seem to be defaults.
+    histBreakInc <- 1
+    xlabText="Ordinal Resp." 
+
+    # make histograms of data:
+    mainTitle <- c("Pre-discussion", "Post-discussion")
+    for ( gIdx in 1:2 ) 
+    {
+        ydata <- as.numeric(inputDf$resp[inputDf$cond == gIdx])
+        histInfo <- hist( 
+            ydata, main=mainTitle[gIdx], 
+            xlab=xlabText, ylab="Probability",
+            freq=FALSE, col="pink", border="white", cex.lab=1.5,
+            breaks=seq(-histBreakInc/2, 
+                       nBins+1+histBreakInc/2, 
+                       by=histBreakInc), 
+            ylim=c(0, LIKERT_SCALE_DENS_MAX) # NB: must set manually at this time
+        )
+        #text( max(histInfo$breaks) , likertScaleDensMax, adj=c(1,1),
+        #      labels=bquote(N==.(length(ydata))) )
+        # xcomb <- seq(0,max(nYlevels)+1,length=201)
+        xcomb <- seq(0, nBins + 1, length=201)
+        lines(xcomb, dnorm(xcomb-mean(ydata), 0, sd(ydata)), 
+              col="skyblue", lwd=2)
+    }
+    
+    tInfo <- fittedFreq$tInfo
+    effSz <- fittedFreq$effSz
[…]

### 2025-12-10 python/get_num_citations.py

@@ -1,0 +1,40 @@
+from datetime import date
+from pandas import read_csv, DataFrame
+
+from scholarly import scholarly
+
+
+ARTICLESDF = read_csv('data/titles.csv')
+
+
+def make_num_citations_csv(output_file_root = 'data/output/citation_count'):
+
+    outdf = ARTICLESDF.copy()
+    outdf['num_citations'] = outdf.Title.map(lambda title: get_num_citations(title))
+
+    outdf.to_csv(output_file_root + str(date.today()) + '.csv', index=False)
+
+    return outdf
+
+
+def get_num_citations(title):
+
+    with open('data/diagnostic/num_citations_log.txt', 'a') as log_f:
+        result = next(scholarly.search_pubs(title))
+
+        # Uncomment to see full result to double check citation count.
+        # print(result)
+
+        found_title = result['bib']['title']
+        num_citations = result['num_citations']
+
+        # Tell the user what's been found to use as entry for
+        diagnostic_msg = \
+            f'* Found "{found_title}" result\nfor searched title "{title}"\nwith ' \
+            f'{num_citations} citations\n'
+
+        print(diagnostic_msg)
+
+        log_f.write(diagnostic_msg + '\n')
+
+        return num_citations

### 2025-12-10 runBayesTrialCluster.R

@@ -1,0 +1,1 @@
+source("experiments.R")

### 2025-12-10 scripts/analysis.R

@@ -1,0 +1,156 @@
+##
+# Two functions. First, calculate a tibble that calculates the family-wise
+# error rate (FWER) and false detection rate (FDR) for three significance values
+# d^*=0.2,0.5,0.8 at three different levels of aggregation (experimental condition, 
+# journal article, and all). The significance values correspond to Cohen's
+# (1988; pp. ~25-30) recommendation. 
+#
+# Second, visualize the tibble at the different aggregation levels for building
+# into a composite figure.
+# 
+# Author: Matthew A. Turner <maturner01@gmail.com>
+# Date: 2024-11-05
+#
+
+library(purrr)
+
+source("src/fwer.R")
+source("src/util.R")
+
+
+# b for base rate, W for power. See 
+fdr = function(fwer, b, W) {
+
+  num = (1 - b) * fwer
+  denom = (num + (b * W))
+
+  return (num / denom)
+}
+
+
+add_default_rows = function(fdr_vs_sig_tbl, default_fwer = 0.05) {
+
+  # Figure out the number of sig values used.
+  sigvals = unique(fdr_vs_sig_tbl$SigVal)
+  n_sigvals = length(sigvals)
+
+  # Extract power and base rate.
+  W = pull(fdr_vs_sig_tbl[1, "Power"])
+  b = pull(fdr_vs_sig_tbl[1, "BaseRate"])
+  
+  # Load and use non-identified plausible.
+  non_id = NON_IDENTIFIED
+
+  n_non_id = n_distinct(non_id)
+  n_default_rows = n_non_id * n_sigvals
+
+  default_fdr = fdr(default_fwer, b, W)
+
+  default_rows = tibble(StudyID = rep(non_id$StudyID, n_sigvals),
+                        ExperimentID = rep(non_id$ExperimentID, n_sigvals),
+                        FWER = default_fwer,
+                        FDR = default_fdr,
+                        SigVal = rep(sigvals, n_non_id),
+                        Power = W,
+                        BaseRate = b)
+                        # Power = rep(W, n_default_rows),
+                        # BaseRate = rep(b, n_default_rows))
+  
+  # Use default FWER, FDR for all SigVals for non-plausibly-identified studies.
+  return (bind_rows(fdr_vs_sig_tbl, default_rows))
+}
+
+
+aggregate_fdr_vs_sig = function(fdr_vs_sig_tbl, use_non_identified = FALSE, 
[…]

### 2025-12-10 scripts/calculations.R

@@ -1,0 +1,111 @@
+##
+# Calculations that are included in paper prose.
+#
+# Author: Matthew A. Turner
+# Date: 2024-12-27
+#
+
+
+source("scripts/analysis.R")
+
+
+# Create a CSV list of quantile calculations that may be included in publications.
+quantiles_tbl = function(tex_output_file = "paper/tex/quantile_table.tex", 
+                         q_resolution_or_probs = 0.05, 
+                         significance_vals = c(0.2, 0.5, 0.8),
+                         base_rate = 0.1, power = 0.8
+                         ) {
+  
+  # Hacky but just need this to proceed for now.
+  if (length(q_resolution_or_probs) > 1) {
+    probs <- q_resolution_or_probs
+  } else {
+    probs <- seq(0.0, 1.0, q_resolution_or_probs)
+  }
+
+  initialized = FALSE
+  for (sigval in significance_vals) {
+
+    tbl <- calc_fdr_vs_significance(significance_vals = sigval, 
+                                    base_rate = base_rate, power = power)
+
+    for (measure in c("FWER", "FDR")) {
+
+      # Get quantile for this measure, base rate, and power.
+      q <- quantile(as_vector(tbl[, measure]), probs = probs)
+
+      if (!initialized) {
+        # Load quantile named list into tbl with quantiles as colnames sans "%".
+        q_tbl <- tibble(!!!q, .name_repair = ~ str_replace(., "%", ""))
+        q_tbl$BaseRate <- base_rate
+        q_tbl$Power <- power
+        q_tbl$SigVal <- sigval
+        q_tbl$Measure <- measure
+
+        initialized = TRUE
+      }
+      else {
+        # Create new row.
+        new_row <- tibble(!!!q, .name_repair = ~ str_replace(., "%", ""))
+        new_row$BaseRate <- base_rate
+        new_row$Power <- power
+        new_row$SigVal <- sigval
+        new_row$Measure <- measure
+
+        # Bind new row to existing.
+        q_tbl <- rbind(q_tbl, new_row)
+      }
+    }
+  }
+
+  # Reorder columns to be easier to read.
+  q_tbl <- 
+    q_tbl %>% 
+      relocate(Measure, .before = `0`) %>% 
+      relocate(SigVal, .after = Measure) %>% 
[…]

### 2025-12-10 scripts/plot/analysis.R

@@ -1,0 +1,231 @@
+##
+#
+# Plot analysis of family-wise error rate simulations.
+# 
+# Currently model illustrations are in the root project directory and will
+# soon be ported here.
+# 
+# Author: Matthew A. Turner <maturner01@gmail.com>
+# Date: 2024-12-07
+#
+
+library(forcats)
+library(ggplot2)
+library(latex2exp)
+library(emojifont)
+
+source("src/fwer.R")
+source("scripts/analysis.R")
+
+mytheme = theme(
+  axis.line = element_line(), legend.key=element_rect(fill = NA),
+  text = element_text(size=18),# family = 'PT Sans'),
+  # legend.key.width = unit(2, 'cm'),
+  # legend.key.size = unit(1.5, 'lines'),
+  panel.background = element_rect(fill = "white"),
+  panel.grid.major.x = element_line(color = "lightgrey", linewidth = 0.1, linetype = 2)
+)
+
+rhg_cols <- rev(c("#771C19", "#AA3929", "#E25033", "#F27314", "#F8A31B", 
+              "#E2C59F", "#B6C5CC", "#8E9CA3", "#556670", "#556AAA", "#000000"))
+
+rhg_cols_10 <- rev(c("#771C19", "#AA3929", "#E25033", "#F27314", "#F8A31B", 
+              "#E2C59F", "#B6C5CC", "#8E9CA3", "#556670", "#556AAA"))
+
+plot_ordinal_cohens <- function(ordinal_data_dir = "data/probit_fits",
+                                sync_file = "data/probit_fits/all.csv",
+                                output_file = 
+                                  "paper/Figures/Analysis/ordinal_cohens.pdf",
+                                overwrite = FALSE) {
+  
+  df <- load_probit_data(probit_fits_dir, sync_file, overwrite)
+  
+  df$Cohens_d <- cohens_d(df$LatentMeanPrePosteriorMean, 
+                          df$LatentMeanPostPosteriorMean,
+                          df$LatentMeanPrePosteriorSD, 
+                          df$LatentMeanPostPosteriorSD)
+  
+  ggplot(df, mapping = aes(x = Cohens_d, y = ExperimentID)) + 
+    geom_vline(xintercept = 0, color = "red") + 
+    geom_vline(xintercept = c(-1, 1), color="red", linetype="dotted") + 
+    geom_boxplot() 
+  
+  ggsave(output_file)
+}
[…]

### 2025-12-10 scripts/plot/model.R

@@ -1,0 +1,180 @@
+library(ggplot2)
+library(readr)
+library(dplyr)
+
+source("model.R")  # for binProb function used in ordinal plot function below.
+
+
+mytheme <- theme(
+  panel.border = element_blank(), axis.line = element_line(), 
+  text = element_text(size=16, family = 'PT Sans'), 
+  panel.background = element_rect(fill = "white")
+)
+
+plot_latent_pdf_integration = function(mu = 0, sd = 2.5, min_bin, max_bin, 
+                                       bins, bin_colors, # xlim = c(-6.0, 6.0), 
+                                       save_path = "test_latent_pdf.pdf", 
+                                       max_prob = 1.0) {
+  
+  print(save_path)
+
+  xmin = min_bin - 0.5
+  xmax = max_bin + 0.5
+
+  xlim = c(xmin, xmax)
+
+  print(xlim)
+
+  p = ggplot(data.frame(x = xlim), aes(x)) 
+
+  p = p + stat_function(fun = dnorm,
+                        args = c(mu, sd),
+                        geom = "line",
+                        xlim = c(xmin, xmax)) 
+
+  # First bin from -infinity to min_bin + 0.5
+  p = p + stat_function(fun = dnorm,
+                        geom = "area",
+                        args = c(mu, sd),
+                        fill = bin_colors[1],
+                        xlim = c(xmin, min_bin + 0.5))
+
+  n_bins = length(bins)
+
+  # Middle bins.
+  for (idx in 2:(n_bins - 1)) {
+
+    p = p + stat_function(fun = dnorm,
+                          geom = "area",
+                          args = c(mu, sd),
+                          fill = bin_colors[idx],
+                          xlim = c(bins[idx - 1] + 0.5, bins[idx] + 0.5))
+  }
+
+  # Final bin from last_bin - 0.5 to infinity.
+  p = p + stat_function(fun = dnorm,
+                        geom = "area",
+                        args = c(mu, sd),
+                        fill = bin_colors[n_bins],
+                        xlim = c(xmax, max_bin - 0.5))
+
+  p = p + xlab("Latent opinion") + ylab("Probability density\n\n") + 
+  geom_vline(xintercept=c((bins[1:(length(bins) - 1)] + 0.5)), linetype="dotted") + 
[…]

### 2025-12-10 src/fwer.R

@@ -1,0 +1,62 @@
+##
+# Calculate the false detection rate from bayesian fits data.
+#
+# Author: Matthew A. Turner
+# Date: 2024-10-30
+#
+
+library(dplyr)
+library(readr)
+library(tidyr)
+
+source("src/util.R")
+
+
+cohens_d <- function(pre_mean_estimate, post_mean_estimate, pre_sd, post_sd) {
+  
+  numerator <- post_mean_estimate - pre_mean_estimate
+  
+  denominator = sqrt((pre_sd**2 + post_sd**2) / 2.0)
+  
+  return (numerator / denominator)
+}
+
+
+##
+# Family-wise error rate calculation using whatever ExperimentIDs are present.
+# Note probit_fits may be a tibble for avoiding reading data every calculation.
+#
+calculate_fwer <- function (probit_fits = "data/probit_fits", sigval = 0.8) {
+
+  # Load probit data if necessary.
+  if (typeof(probit_fits) == "character") {
+    probit_data <- load_probit_data(probit_fits)
+  } else {
+    probit_data <- probit_fits
+  }
+     
+  probit_data$Cohens_d <- cohens_d(probit_data$LatentMeanPrePosteriorMean, 
+                                  probit_data$LatentMeanPostPosteriorMean,
+                                  probit_data$LatentMeanPrePosteriorSD, 
+                                  probit_data$LatentMeanPostPosteriorSD)
+
+  # Get the sign of the hypothesized shift, where these columns come from 
+  # the original data.
+  shift_sign <- sign(probit_data$ObservedMeanPost - probit_data$ObservedMeanPre)
+
+  # Only significant if it's positive in direction of shift. TODO: alternative
+  # formulation could check absolute value.
+  probit_data$Significant <- 
+    ifelse((shift_sign * probit_data$Cohens_d) >= sigval, 1, 0)
+
+  ret = probit_data %>%
+    group_by(ExperimentID) %>% 
+    summarise(FWER = mean(Significant)) %>%
+    separate_wider_delim(ExperimentID, delim = "_", 
+                         names = c("StudyID", "ExperimentID")) %>%
+    arrange(desc(FWER))
+
+  return (ret)
+}
+
+

### 2025-12-10 src/util.R

@@ -1,0 +1,32 @@
+##
+# Utility functions used in other src and scripts files.
+#
+# Author: Matthew A. Turner
+# Date: 2024-12-27
+#
+
+library(dplyr)
+library(readr)
+library(tidyr)
+library(fs)
+
+
+load_probit_data = function(probit_fits_dir = "data/probit_fits", 
+                            sync_file = "data/probit_fits/all.csv",
+                            overwrite = FALSE) {
+     
+  if (!file.exists(sync_file) || overwrite) {
+
+    ret <- dir_ls(probit_fits_dir, glob = "*.csv") %>%
+      read_csv() %>% unite(ExperimentID, ArticleTag, TreatmentTag)
+
+    write_csv(ret, sync_file)
+
+  } else {
+
+    ret <- read_csv(sync_file)
+  }
+
+  return (ret)
+}
+

### 2025-12-10 util.R

@@ -1,0 +1,136 @@
+##
+# The main display element is one row of controls and two rows of plots. 
+#
+# The top row will be the large N exact calculations of the 
+# latent standard deviations given observed pre- and post-deliberation
+# means, a hypothesized latent mean, and a reporting model and measurement 
+# scheme. The latent mean and observed pre- and post-deliberation means
+# can be set using the top row controls, or loaded from preset values
+# from published result case studies that contain plausibly
+# false demonstrations of the group polarization effect.
+#
+# The second row displays simulated empirical observations for an N given in 
+# the controls at the top of the main display. The numbers and plots 
+# are generated using the same latent and observed 
+# pre- and post-deliberation means as for the large N exact calculations.
+#
+# The "reporting model" is operationalized as thresholds that determine
+# which opinion scale (e.g. Likert) bin a particpant's opinion is reported as.
+# See the code for how this works. The rating scale would be, e.g., a
+# 7-point Likert, or the CDQ responses, which rate riskiness from 1 to 10.
+#  
+# getCaseStudyData <- function() {
+#     read.csv('caseStudies.csv') 
+#     data.names <- data$CaseStudyName
+#     return (data)
+# }
+# CaseStudyData <- getCaseStudyData();
+
+library(shiny)
+library(grid)
+library(gridExtra)
+library(lattice)
+library(ggplot2)
+
+source("model.R")
+source("numerical.R")
+
+# CaseStudyData <- read.csv('caseStudies.csv', row.names='CaseStudyTag')
+
+
+##
+# Arguments:
+#     caseStudy (string): See CSV for values to use, e.g., 
+#         Moscovici1969-DeGaulle.
+setCaseStudyValues <- function(caseStudyName, session)
+{
+    data <- CaseStudyData[caseStudyName, ]
+    updateSliderInput(session, "latentMean", 
+                      value = data$LatentMean, min = data$MinBinValue - 2,
+                      max = data$MaxBinValue + 2)
+                      
+    updateSliderInput(session, "observedPreDelibMean", 
+                      value = data$ObservedPreDelibMean, 
+                      min = data$MinBinValue,
+                      max = data$MaxBinValue)
+
+    updateSliderInput(session, "observedPostDelibMean", 
+                      value = data$ObservedPostDelibMean, 
[…]

### 2025-12-10 www/message-handler.js

@@ -1,0 +1,5 @@
+Shiny.addCustomMessageHandler("testmessageHandlerForJS",
+  function(message) {
+    alert(JSON.stringify(message));
+  }
+);

### 2025-12-10 index.qmd
continued restructuring
ready to ship online gh-pages supplement

@@
 title: Notes and code examples to accompany "If the Null Fits, You Must Omit"
-author: Matt Turner · [https://mat.phd](https://mat.phd)
+author: Matt Turner · [Email Me!](mailto:matt@mat.phd) · [https://mat.phd](https://mat.phd) 
 date: today
 toc: true
+toc-expand: 2
 editor: 
@@
     wrap: 90
-bibliography: "paper/this.bib"
+bibliography: "this.bib"
 
@@
 
-## Purpose: demonstrate gp-stats codebase
+## Overview
 
 Here are some quick notes on the group polarization codebase in R supporting our article,
-*If the Null Fits, You Must Omit*. The code is permanently [hosted on
+*If the Null Fits, You Must Omit*. The code is [hosted on
 GitHub](https://github.com/mt-digital/gp-statmod/tree/final-draft/GroupPolarizationStatmod).
@@
 
-I developed two models of group polarization to serve complementary purposes.
+I developed two models of experimental observations in group polarization experiments
+to serve complementary purposes. First is the _expectation_ model where we model the observation of the mean
 
@@
 ![**Mean group opinions pre- and post-discussion from Moscovici & Zavalloni
-(1969).**](code_examples_images/clipboard-1443885641.png){width="575"}
+(1969).**](images/clipboard-1443885641.png){width="575"}
 

### 2025-12-11 index.qmd
updated docs

@@
 
+<!-- Define the tex command for the null model -->
+\newcommand{\nullmodel}{\mathcal{M}_0}
+
+<!-- Begin main document -->
 ## Overview
@@
 
+You will learn how the code works that simulates measurements of opinion change
+in group polarization experiments that test if and by how much group discussion polarizes opinions.
+I used this modeling system to simulate 57 different group polarization experimental conditions across a corpus of ten journal articles published over several decades, and across social science subdisciplines for my paper with Paul Smaldino, "Null Fits?
+You Must Omit!
+How Measurement Artifacts Undermine Many Group Polarization Replications".
+
 # Simulations
@@
 I developed two models of experimental observations in group polarization experiments
-to serve complementary purposes. First is the _expectation_ model where we model the observation of the mean
+to serve complementary purposes. These models provide two ways to simulate observations of _mere agreement_, which serves as our null-polarization model, $\nullmodel$,
+
+First is the _expectation_ model where we model the observation of the average group opinion as the _expectation value_ of the distribution of ordinal opinions.
+We calculate the ordinal distribution of opinions by writing the endpoints of each measurement bins as limits of integration in a series of $B$ integrals, where $B$ is the number of measurement bins.
+For example, an example of an ordinal measurement scale might be a Likert-style series of disagree-agree levels, associated with numerical values:
+-3 indicates
 

### 2025-12-11 index.qmd
improved explanation of generative model analyses; next: how simulating experimental data works

@@
 \newcommand{\nullmodel}{\mathcal{M}_0}
+
+\DeclareMathOperator*{\argmin}{argmin}
+\DeclareMathOperator*{\argmax}{argmax} % Optional: also define argmax
 
@@
     standard deviation of latent opinions, but a constant mean latent opinion.
+Here, then, we calculate the false positive rate, $\alpha_e$ as a function of the effect-detection threshold, $d^*$, i.e.,
+    $$
+      \alpha_e(d^*) = \Pr\!\left( \hat{d} \ge d^{*} \mid \mathcal{M}_{0} \right).
+    $$
+    In the study, we calculate the right-hand side as the frequency of the 1000 trials where $\hat{d} \ge d^*$ for simulations of experiment, $e$. 
 2.  **Find an effect–detection threshold**, $d^*(\alpha)$, such that, under the null
     model, the probability of observing an estimated effect size exceeding
-    \$d\^\*(\\alpha)\$ is \$\alpha\$: $$
-     \Pr\!\left( \hat{d} \ge
-     d^{*}(\alpha) \mid \mathcal{M}_{0} \right) = \alpha.
+    \$d\^\*(\\alpha)\$ is some target false positive rate, $\alpha^* = 0.05$, for example: 
+    $$
+     d^* = \min~d,~~\text{subject to}~\Pr\!\left( \hat{d} \ge d \mid \mathcal{M}_{0} \right)  \le\alpha^*.
      $$ 
@@
 The first step in our _generative_ model of false detections is to _generate_ simulated 
-experimental data from a mere agreement null-polarization model.
-Then we can calculate how frequently an effect size was inferred that exceeded the effect-detection threshold.
-simulated null-model data our simulated null-models.
+experimental data from a mere agreement null-polarization model. Then we can calculate how frequently an effect size was inferred that exceeded the effect-detection threshold.
 
@@
 I simulate experimental group polarization data for a given time, $t$, with latent group opinion mean and standard deviation, $\mu_t$ and $\sigma_t$, and with a survey response scale defined by $B$ bin values, $b_i$, with $i=1,\ldots,B$.
+
+<!-------- CONTINUE HERE 12/11! ------>
 

### 2025-12-11 index.qmd
minor updates and cleanup of analysis; about to post docs

@@
       \alpha_e(d^*) = \Pr\!\left( \hat{d} \ge d^{*} \mid \mathcal{M}_{0} \right).
-    $$
+    $$ {#eq-false-positive-rate}
     In the study, we calculate the right-hand side as the frequency of the 1000 trials where $\hat{d} \ge d^*$ for simulations of experiment, $e$. 
@@
      d^* = \min~d,~~\text{subject to}~\Pr\!\left( \hat{d} \ge d \mid \mathcal{M}_{0} \right)  \le\alpha^*.
-     $$ 
+     $$ {#eq-threshold-spec}
      This yields a null-model–calibrated false–positive rate for detecting a group
@@
 
-In the second step of our generative model analysis I calculate
+In the second step of our generative model analysis I calculate the solution to
+the problem in @eq-threshold-spec, finding the effect-detection threshold, $d^*$, that achieves a target false positive rate, $\alpha^* = 0.05$.
 

### 2025-12-11 scripts/analysis.R

@@ -12,6 +12,7 @@
 # Date: 2024-11-05
 #
 
+library(forcats)
 library(purrr)
 
 source("src/fwer.R")

### 2025-12-11 index.qmd
updated docs

@@
 
-\DeclareMathOperator*{\argmin}{argmin}
-\DeclareMathOperator*{\argmax}{argmax} % Optional: also define argmax
+<!-- Begin main document -->
+# Introduction
 
-<!-- Begin main document -->
-## Overview
+This document supports my article with Paul Smaldino, *Null Fits?
+Must Omit!* It follows the structure of the article's _Method_ and _Analysis_ sections, providing a tutorial explanation, first, of the code used to model group polarization experiments and opinion measurement; and, second, how the code works to summarize, analyze, and visualize simulation results.
 
-Here are some quick notes on the group polarization codebase in R supporting our article,
-*If the Null Fits, You Must Omit*. The code is [hosted on
-GitHub](https://github.com/mt-digital/gp-statmod/tree/final-draft/GroupPolarizationStatmod).
+Our article describes how we analyzed simulations of group polarization experiments to show that ~95% of experimental replications of the group polarization effect are invalid due to non-identifiability from measurement artifacts across ten journal articles that comprised our curated corpus.
+There were a total of 57 independent experimental conditions across these studies.
 
-To follow along, please clone the `gp-statmod` directory.
+We furter estimated at least a 70% chance that any published replication of group polarization is, in fact, a false detection.
 
-```
-$ git clone https://github.com/mt-digital/gp-statmod.git
+We also show how this simulation system can be used constructively to specify what effect size should be used as a detection threshold to limit FPR to 0.05.
+For more details of the motivation, theory, and results of our analysis, please see our paper.
+
+**You can read this documentation online, hosted by GitHub:** https://mt-digital.github.io/gp-statmod
+
+**Or, you can read this document and run the code examples yourself.** To follow along and run code on your own, please clone the `gp-statmod` directory:
+
+```{bash}
+#| eval: false
+git clone https://github.com/mt-digital/gp-statmod.git
 ```
 
-Change directories to go to the `gp-statmod/docs` directory where this document lives:
+Change directories to go to the `gp-statmod` directory where this document lives:
 
-```
-cd gp-statmod/docs
+```{sh}
+#| eval: false
+cd gp-statmod
 ```
@@
 
-This tutorial is still valuable if you're just reading it on the [code documentation associated with the journal article](https://mt-digital.github.io/gp-statmod).
-
-You will learn how the code works that simulates measurements of opinion change
+Either way, you will learn how the code works that simulates measurements of opinion change
 in group polarization experiments that test if and by how much group discussion polarizes opinions.
@@
 
-# Simulations
+# Method
 
@@
 $\sigma_\mathrm{pre}$, and $\sigma_\mathrm{post}$ from the paper that define a null
-polarization model, $\nu = (\mu, \sigma_\mathrm{pre}, \sigma_\mathrm{post})$.
+polarization model, $\nullmodel = (\mu, \sigma_\mathrm{pre}, \sigma_\mathrm{post})$.
 
@@
 the problem in @eq-threshold-spec, finding the effect-detection threshold, $d^*$, that achieves a target false positive rate, $\alpha^* = 0.05$.
+We need an updated dataset of Type I error rates calculated with a finer-grained set of candidate effect-detection thresholds, which we just call $d$ if they're candidates, than the simple three, 0.2, 0.5, and 0.8, used in the previous section.
 
-# Analyses
+```{r}
+# The fwer is equivalently Type I error—written before I understood the difference,
+# not worth changing the name for this long-running project before it's done.
+det_thresh_tbl <- sigval_for_low_fwer(fdr_over_thresholds, target_fwer = 0.05)
+
+print(det_thresh_tbl)
+```
+
+# Analysis
 
